### 2017-03-24 cross-validation.Rnw
share files of the first draft

\documentclass{article}
\usepackage{amssymb,amsmath,natbib,graphicx,enumerate,subcaption,tikz,url,booktabs}

\usepackage[margin=1in]{geometry}

\newcommand{\logit}{\mbox{logit}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\ones}{\mathbbm{1}}
\newcommand{\indic}[1]{\boldsymbol{1}_{\{ #1 \}}}
\newcommand{\mvec}{\mbox{vec}}
\newcommand{\cov}{\mbox{Cov}}
\newcommand{\eqdist}{\overset{{\cal D}}{=}}
\newcommand{\const}{\mbox{const}}
\newcommand{\iid}{\overset{iid}{\sim}}
\newcommand{\diag}{\mbox{{\bf diag}}}
\newcommand{\Diag}{\mbox{Diag}}

\newcommand{\bmu}{\boldsymbol\mu}
\newcommand{\balpha}{\boldsymbol\alpha}
\newcommand{\br}{\boldsymbol{r}}
\newcommand{\bs}{\boldsymbol{s}}
\newcommand{\by}{\boldsymbol{y}}
\newcommand{\bbeta}{\boldsymbol{\beta}}
\newcommand{\btheta}{\boldsymbol{\theta}}
\newcommand{\btau}{\boldsymbol{\tau}}
\newcommand{\bx}{\boldsymbol{x}}
\newcommand{\beps}{\boldsymbol\epsilon}

\newcommand{\cX}{{\cal X}}
\newcommand{\cA}{{\cal A}}
\newcommand{\cAmin}{{\cal A}^{\mathsmaller{-}}}

\newcommand{\vm}[1]{\mbox{vec}^\mathsmaller{-}\hspace{-0.25pc}\left(#1\right)}
\newcommand{\Imin}{I^\mathsmaller{-}}
\newcommand{\Itri}{{I^\mathsmaller{\triangle}}}

\newcommand{\deriv}[1]{\noindent {\it Derivation:} #1 $\hfill\square$}
\newcommand{\pbe}[1]{{\cal P}_{\epsilon,#1}}
\newcommand{\be}{{\cal B}_{\epsilon}}

\newcommand{\adjSet}{\{A_{k_\ell}\}_{\ell=1}^K}
\newcommand{\abSet}{\{\alpha_{k_\ell},\bbeta_{k_\ell}\}_{\ell=1}^K}
\newcommand{\bZ}{{\bf Z}}

\title{Comparing Cross Validation Methods for Penalized Cox Regression}
\author{
}
\date{March 24, 2017}

\begin{document}
\maketitle

\section*{Abstract}
Due to the semi-parametric nature of Cox Regression, conducting cross validation for Cox models has always been a challenge. While cross validation is a commonly used approach for selecting tuning parameters in penalized regression, little research has been done to study cross validation methods for penalized Cox regression. We propose two new cross-validation methods for Cox Regression, and compare them to traditional information criteria as well as a cross-validated partial likelihood approach originally proposed by Verweij et al. Our simulation studies show that, in general, cross-validation tends to be conservative (i.e., select smaller models than the ideal choice of tuning parameters) for penalized Cox regression models. However, our proposed approach of cross-validating the linear predictors generally offers the best balance of stability and performance. We also illustrate these approaches on data from studies of gene expression and progression-free survival in cancer patients.

\section{Introduction}
Cox propotional hazard regression is one of the most commonly used statistical model for analyzing data with survival outcomes. Despite its popularity, the cox model has its own limitations due to the semiparametric nature of the model. Since the model only estimates relative risks, it is challenging to evaluate the model's predictive accuracy.

In the Cox model, the hazard function for individual observation is given by \begin{equation}h_{j}(t) = h_{0}(t) exp( X_{j}^{T} \beta),\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{j}^{T} \beta}$ is the relative risk. The estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood \begin{equation}L(\beta) = \prod_{j=1}^{n} \frac{exp ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}exp ( X_{k}^{T} \beta)}.\end{equation} As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard. Hence the interpretation of the Cox model is only valid in a relative sense.

Penalized Cox Regression is an extension of the regular cox regression model. Instead of dealing with the partial likelihood, estimates of the $\beta$ coefficients are obtained by minizing the objective function \begin{equation}  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{n} log \left \{\frac{exp ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}exp ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta). \end{equation} The first part of the objective function is based on the partial likelihood. The second part is a penalty term that depends on tuning parameter $\lambda$. With LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$, the penalized cox regression extends cox regression to deal with variable selections in high dimensional settings, where the number of variable p $\gg$ n, the number of observation. An example of its application would be to study how the overall survival time of a particular cancer is associated with gene expressions.

To get LASSO estimations for the $\beta$ coefficients, the choice of the tuning parameter $\lambda$ is crutial. As $\lambda$ increases, more $\beta$ coefficients would be set to 0 and fewer variables will be selected by the model. When $\lambda$ is too small, the model tends to be redundant. And $\lambda$ also should not be set too large. In this case, the model would be too conservative to select any variables.

One common approach for selecting the tuning parameter $\lambda$ for linear regression and logitic regression is via cross validation. The idea of K - fold cross validation is illustrated in Figure 1. The data set would be split into K folds. One fold would be treated as the test set and the other K - 1 folds as the training set. The model would first be built on the training set, then fitted to the test set to obtain cross validation error (CVE). To select the most suitable lambda,the CVE would be calculated for each candidate lambda. And the one that has the smallest cross validated error would be selected.

\begin{figure}
    \centering
		\includegraphics[height= 4cm ]{./figures/cv.png}	
    \caption{K-fold Cross Validation}
\end{figure}

However, for penalized cox regression, there is no natural way to carry out cross validation. Since the partial likelihood of the cox model does not give estimates for the baseline, the outcomes can only be compared within the same framework, in terms of their relative risks. This is against the idea of cross validation, which relies on splitting the data into different folds.

\par The most commonly used approach of conducting cross validation in Cox regression is proposed by Verweij elt. (1993). They propose a cross-validated log likelihood as a measure for the model's predictive accuracy.  It is currently implemented in the glmnet package as the "grouped" option. Similarly, the glmnet package also includes the "ungrouped" option where the cross-validated likelihood is calculated only based on the observations from the test set. In this paper, we proposed two alternative ways to carry out cross validation. Instead of cross-validating over partial likelihood, we propose to cross validate over the linear predictors of the regression model and cross validate over the Martingale residual of the Cox model. 
\par We conducted simulation studies to compare how those methods perform in both low dimensional and high dimensional settings with LASSO penalty. We compared the two proposed methods with the cross-validated partial likelihood approaches and traditional information criteria. We showed that the linear predictor approach outperforms other cross validation approaches in various scenarios. We found out that using cross validation to select tuning parameters for LASSO penalized cox regression tends to be a conservative approach in general. Finally, we apply those methods to two genomic data sets with time-to-event outcomes.

\section{Methods}

  \subsection{Cross Validated Likelihood}

Suppose a data set of n observations is split into K folds and the ith fold is left out, then a partial likelihood can be built over the K -1 folds and yields maximum likelihood estimates of the coefficients, denoted by $\hat{\beta}_{-i}$. Verweij elt. defined cross-validated log likelihood of leaving the ith fold out to be \begin{equation}cvl = \sum_{i=1}^{K} \{l(\hat{\beta}_{-i}) - l_{-i}(\hat{\beta}_{-i})\}. \end{equation} When the ith fold is left out, $l(\hat{\beta}_{-i})$ is the log partial likelihood evaluated at $\hat{\beta}_{-i}$ with all k folds of observations. $l_{-i}(\hat{\beta}_{-i})$ is the log partial likelihood evaluated at $\hat{\beta}_{-i}$ with observations from the other K-1 folds.

An alternative way of conducting cross-validation is implemented in the glmnet package. The cross-validated log likelihood is defined as\begin{equation}cvl = \sum_{i=1}^{K} l_{i}(\hat{\beta}_{-i}).\end{equation} Instead of using substraction, $l_{i}(\hat{\beta}_{-i})$ is the partial likelihood evaluated at $\hat{\beta}_{-i}$ with observations from the fold that is left out.

We refer to the first definition as grouped cvl and the second definition as ungrouped cvl. Figure 2 illustrates the ideas of the two methods. The grouped cvl uses observations more efficiently then the ungrouped one. Since only the left-out fold would be used, the ungrouped cvl would not very stable if number of observations is too small.

\begin{figure}
    \centering
		\includegraphics[height= 4cm ]{./figures/02_1.png}
		\includegraphics[height= 4cm ]{./figures/02_2.png}
    \caption{Grouped vs Ungrouped Cross Validated Likelihood}
\end{figure}

  \subsection{Cross Validated Linear Predictors}

Besides cross validating over the partial likelihood, an alternative approach is to cross validate over the linear predictors. The data would still be split into K folds. Suppose ith fold is left out and the other K-1 folds are used as the training set to get the estimates $\hat{\beta}_{-i}$. Then the cross-validated linear predictors would be calculated based on the observations in the test set:  \begin{equation}\hat{\eta}_{-i} = X_{i}\hat{\beta}_{-i}.\end{equation} After repeating this for all K folds, a whole set of linear predictors  $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...  \hat{\eta}_{-K})$ can be obtained. A partial likelihood can be built over this set of linear predictors: \begin{equation} L(\hat{\eta}_{-}) = \prod_{j=1}^{n} \frac{exp (\hat{\eta}_{-j})}{\sum_{ k \in R(t_{j})}exp (\hat{\eta}_{-k})}\end{equation}.

Cross validated error is defined as the log of $L(\hat{\eta}_{-})$. This idea of cross validating over linear predictors is implemented in the package ncvreg. For all the approaches introduced so far, they are all equivalent to each other in linear regressions. But since they are building over different baselines, they are different in cox regression.

   \begin{figure}[h]
    \centering
    \begin{minipage}[b]{0.5\textwidth}
      \centering
		  \includegraphics[height= 3.5cm ]{./figures/03.png}
      \caption{Cross Validated Linear Predictors}
     \end{minipage}
     \begin{minipage}[b]{0.5\textwidth}
      \centering
		  \includegraphics[height= 3.9cm ]{./figures/04.png}
      \caption{Cross Validated Martingale Residuals}
      \end{minipage}	
   \end{figure}

  \subsection{Cross Validated Martingale Residuals}

Since the biggest challenge of conducting cross validation for cox regression is due to the missingness of baseline hazard, we propose a second approach which involves estimating the actual baseline hazard $\hat{\Lambda}_{0}$. Based on the observations in the training set, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method. Then for each observation j in the test set, the Martingale Residual can be calculated: \begin{equation}\hat{M_{j}} = d_{j} -\hat{\Lambda}_{0}(t_{j})e^{\hat{\eta}_{j}},\end{equation} where $d_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation. After repeating this procedure for all K folds, we would obtain a set of matingale residuals for all observations. They are be pulled together by sum of squares as the cross validated error:
\begin{equation}CVE = \sum_{j}\hat{M}_{j}^2\end{equation}

\section{Simulation Studies}

  \subsection{Simulation Settings}

  Simulation studies were conducted to compare how those methods behave relative to each other. We generate data with pre-specified baseline hazard, covariate matrix X, coefficient $\beta$ and censoring mechanism. Both low dimensional and high dimentional scenarios were examined. All cross-validation methods mentioned in Section 2 were applied to the data to select the tuning parameter $\lambda$ and produce $\hat{\beta}$ estimates. Cross-validation were also compared to model selection criteria AIC and BIC.

  Survival times were generated from exponential distribution $h(t) = h_{0} exp(X\beta)$, conditioned on covariates. The entries in the covariate matrix $X_{n \times p}$ were independently generated from Normal (0, 1). True coefficients $\beta$  were assumed to have sparsity: $\beta_{p\times 1} = (\beta_{1},\beta_{2}, ..., \beta_{10}, 0, ..., 0)^{T}$.  Censoring status were generated based on binomial distribution.

  Suppose $\hat{\beta}$ is the coefficients estimated by a fitted cox model. We measure the distance between the fitted model and the true model by mean squared error $MSE = E(\hat{\beta} - \beta)^2$. For each generated data set, the $\lambda$ that has the minimal MSE is chosen as the optimal $\lambda$. Then the $\lamda$s that selected by the cross validation or information criteria would be compared to this optimal $\lambda$. If the $\lambda$ chosen by cross validation is smaller than the optimal $\lambda$, then the cross validation method would be considered liberal. If the $\lambda$ chosen by cross validation is larger than the optimal $\lambda$, then the cross validation method would be considered conservative.

  \subsection{Simulation Results}

    \subsubsection {Simulations in Low Dimensional Settings}

   The simulation experiments were first conducted under lower dimension settings. Number of observations n is set to be 100. Dimension of the data p is set to be 100. 200 replications were used. Censoring percentage, number of folds used in cross-validation, number of non-zero $\beta$s and the magnitude of the $\beta$s were varied in several different simulation scenarios. Patterns and results shown in various simulation scenarios are consistent with one another. Figure 5 and Figure 6 illustrates one simulation scenario, where about 10$\%$ of which were set to be censored and 10 of the $\beta$s were set to be non-zero and 10 folds were used for cross validation. The x - axis of the plots are the magnitude of the non-zero $\beta$ coefficients.

   \begin{figure}[h]
   \centering
    \begin{minipage}[b]{0.4\textwidth}
      \centering
		  \includegraphics[height= 6cm ]{./figures/05.jpeg}
      \caption{ }
     \end{minipage}
     \begin{minipage}[b]{0.4\textwidth}
      \centering
		  \includegraphics[height= 6cm ]{./figures/06.jpeg}
      \caption{ }
      \end{minipage}	
   \end{figure}

    $\lambda$s chosen by different cross-validation methods are compared in Figure 5. All values are from taking the mean of 200 replications. The black line represents the optimal $\lambda$. All four cross validation methods consistently select $\lambda$s larger than the optimal one. Since larger $\lambda$ leads larger penalties and fewer variables are selected, all four cross validation methods seem to be conservative in terms of selecting variables. Ungrouped cross-validated log likelihood and the cross-validated linear predictor seem to be more liberal than the other two methods. The cross-validated martingale residual is the most conservative one and its performance is quite off from the other three methods.

   Mean squared errors are compared in Figure 6. The MSEs of the cross-validation selected models are compared with the MSE of the model given by the optimal $\lambda$. The y-axis represents the ratio of the two. The ungrouped method and the linear predictor approach are outperforming the other methods and are closest to the minimal MSE.

    \subsubsection {Simulations in High Dimensional Setting}

   Simulation studies were also conducted in high dimension scenarios, where the number of observations and dimension of the covariate matrix is more similar to real genetics research. The results are shown in Table 1. Conclusions are close to what is shown in the low dimensional setting. The M- residual approach performs most conservatively. Linear predictor approach and ungrouped approach are relatively more liberal than the other approaches but still slightly more conservative than the optimal $\lambda$.

  \begin{table}[h]
    \centering
			\begin{tabular}{ c | c | c | c}
				CV Methods & $\lambda$ (SD) & MSE Ratio \\
				\hline \hline
				Optimal  & 0.0606 ( 0.0019 )&  / & \\ 
				Grouped  &  0.0895  ( 0.0043 )& 1.598  \\
				Ungrouped & 0.0754 ( 0.0060 )& 1.225 \\
				Linear Pred & 0.0759 ( 0.0058 )& 1.234 \\
				M Residual &0.1283 ( 0.0047 ) & 2.959
			\end{tabular}
			\caption{Simulation Results in High Dimensional Setting}
	\end{table}

  \vspace{20 mm}
    \subsubsection {Comparisons with Information Criteria}
    We also compared cross-validation with information criteria AIC and BIC. ...

   \begin{figure}[h]
    \centering
    \begin{minipage}[b]{0.4\textwidth}
      \centering
		  \includegraphics[height= 6cm ]{./figures/aic_01.jpeg}
      \caption{Comparison to Information Criteria in Lower Dimension}
     \end{minipage}
     \begin{minipage}[b]{0.4\textwidth}
      \centering
		  \includegraphics[height= 6cm ]{./figures/aic_02.jpeg}
      \caption{Comparison to Information Criteria in Higher Dimension}
      \end{minipage}	
   \end{figure}

\section{Application to Real Data}

We also compared those cross validation methods when they are applied to real data sets. Unlike simulated set, we do not know the optimal $\lambda$ in the real data sets, but we can still see how those methods perform relative to each other. The first data set is a study on ovariance cancer from The Cancer Genome Atlas. The second data set is a study on lung cancer (Shedden et al 2004). The outcome of both data sets are time-to-event data.

There are 460 patients and 236 events for the ovarian cancer data set. About 40$\%$ of the observations are censored. The covariate matrix records whether or not there is mutation at the location and has dimension 12376. When we fit the penalized cox model, we first adjusted three clinical variables which is known to have large impact on survival. We first fitted these three covariates into regular cox regression and keep the linear predictors. The selection for the penalty term does not affect the estimation of those clinical variables. The results are illustrated in Table 2. The linear predictor approach has the best predictive accuracy according to AUC.

\begin{table}[h]
\centering
  \begin{tabular}{ c | c | c | c}
				CV Methods & $\lambda$ & $log(\lambda)$ &AUC \\
				\hline \hline
				Grouped  &  0.117 & -2.14 &   0.595\\
				Ungrouped &  0.112 & -2.20 &  0.597\\
				Linear Pred &  0.085 & -2.46 &     0.607\\
				M Residual & 0.171 & -1.76 & 0.529 
	\end{tabular}
  \caption{Tuning Parameter Selected for Ovarian Cancer Data}
\end{table}

In Figure 9, the Cross Validated Error is rescaled and plotted for all four methods. A $\lambda$ will be selected when CVE curve reaches its lower point. The blue line, which represents the linear predictor approach, it has more curvature near its lowest point. It is easier to pick out the minimum point for this blue curve. Hence this approach is better at picking out signals than the other three approaches. If we look at the curve for the cross-validated Martingale Residual, there's almost no signal there.

   \begin{figure}[h]
   \centering
    \begin{minipage}[b]{0.4\textwidth}
      \centering
		  \includegraphics[height= 6cm ]{./figures/07.jpeg}
      \caption{CVE for Ovarian Cancer Data}
     \end{minipage}
     \begin{minipage}[b]{0.4\textwidth}
      \centering
		  \includegraphics[height= 6cm ]{./figures/08.jpeg}
      \caption{CVE for Lung Cancer Data}
      \end{minipage}	
   \end{figure}

In the lung cancer data set that we applied our methods to, there are 442 patients and 236 events. About 50$\%$ of the patients are censored. Again we adjusted three clinical variable before we fit the penalized regression. The results of the analysis are listed in Table 3.

			\begin{table}[h]
			\centering
			\begin{tabular}{ c | c | c | c}
				CV Methods & $\lambda$ & $log(\lambda)$ & AUC \\
				\hline \hline
				Grouped  &   0.119 & -2.12 &  0.608 \\
				Ungrouped &  0.119 & -2.12 &  0.608 \\
				Linear Pred &  0.090 & -2.41 &  0.633\\
				M Residual & 0.203 & -1.59 &  0.564
			\end{tabular}
			\caption{Tuning Parameter Selected for Lung Cancer Data}
			\end{table}

If we look at the CVE curve for all four of those methods, we still see that blue curve has more curvature near its lowest point. But in general, there are more signal in this lung cancer data set compared to the ovarian cancer data set.

\section{Discussion}

\end{document}

### 2017-03-24 cross-validation.tex
share files of the first draft
Added a citation

--- 2017-03-24 cross-validation.Rnw
@@
-\documentclass{article}
+\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage{amssymb,amsmath,natbib,graphicx,enumerate,subcaption,tikz,url,booktabs}
+\usepackage{natbib}
 
@@
 \date{March 24, 2017}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 
-In the Cox model, the hazard function for individual observation is given by \begin{equation}h_{j}(t) = h_{0}(t) exp( X_{j}^{T} \beta),\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{j}^{T} \beta}$ is the relative risk.
+In the Cox model \citep{Cox1972}, the hazard function for individual observation is given by \begin{equation}h_{j}(t) = h_{0}(t) exp( X_{j}^{T} \beta),\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{j}^{T} \beta}$ is the relative risk.
 The estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood \begin{equation}L(\beta) = \prod_{j=1}^{n} \frac{exp ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}exp ( X_{k}^{T} \beta)}.\end{equation} As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard.
@@
 For each generated data set, the $\lambda$ that has the minimal MSE is chosen as the optimal $\lambda$.
-Then the $\lamda$s that selected by the cross validation or information criteria would be compared to this optimal $\lambda$.
+Then the $\lambda$s that selected by the cross validation or information criteria would be compared to this optimal $\lambda$.
 If the $\lambda$ chosen by cross validation is smaller than the optimal $\lambda$, then the cross validation method would be considered liberal.
@@
 
+\bibliographystyle{ims-nourl}
+\bibliography{articles}
+
 \end{document}

### 2017-09-06 cross-validation.tex
Some changes

@@
 \usepackage{amssymb,amsmath,natbib,graphicx,enumerate,subcaption,tikz,url,booktabs}
-\usepackage{natbib}
 
@@
 }
-\date{March 24, 2017}
+\date{ }
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
@@
 Cox propotional hazard regression is one of the most commonly used statistical model for analyzing data with survival outcomes.
-Despite its popularity, the cox model has its own limitations due to the semiparametric nature of the model.
-Since the model only estimates relative risks, it is challenging to evaluate the model's predictive accuracy.
+Despite its popularity, Cox model is limited by its semiparametric nature: it is challenging to evaluate the model's predictive accuracy.
 
-In the Cox model \citep{Cox1972}, the hazard function for individual observation is given by \begin{equation}h_{j}(t) = h_{0}(t) exp( X_{j}^{T} \beta),\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{j}^{T} \beta}$ is the relative risk.
-The estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood \begin{equation}L(\beta) = \prod_{j=1}^{n} \frac{exp ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}exp ( X_{k}^{T} \beta)}.\end{equation} As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard.
+In the Cox model \citep{Cox1975}, the hazard function for an individual observation j is given by \begin{equation}h_{j}(t) = h_{0}(t) \text{exp}( X_{j}^{T} \beta)\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{j}^{T} \beta}$ is the relative risk.
+The estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood \begin{equation}L(\beta) = \prod_{j=1}^{n} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation} As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard.
 Hence the interpretation of the Cox model is only valid in a relative sense.
@@
 Penalized Cox Regression is an extension of the regular cox regression model.
-Instead of dealing with the partial likelihood, estimates of the $\beta$ coefficients are obtained by minizing the objective function \begin{equation}  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{n} log \left \{\frac{exp ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}exp ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta).
-\end{equation} The first part of the objective function is based on the partial likelihood.
-The second part is a penalty term that depends on tuning parameter $\lambda$.
-With LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$, the penalized cox regression extends cox regression to deal with variable selections in high dimensional settings, where the number of variable p $\gg$ n, the number of observation.
-An example of its application would be to study how the overall survival time of a particular cancer is associated with gene expressions.
+Estimates of the $\beta$ coefficients are obtained by minimizing the objective function \begin{equation}  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{n} \text{log} \left \{\frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta).
+\end{equation}, instead of optimizing the partial likelihood.
+The first part of the objective function is based on the partial likelihood.
+The second part is a penalty term that depends on some regularization parameters.
+In this paper, only the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$ was considered.
+The penalized cox regression can be used for analyzing high dimensional data, where the number of covariates p $\gg$ n.
+A common application would be to predict the overall survival time of a particular cancer by certain gene expressions.
 
-To get LASSO estimations for the $\beta$ coefficients, the choice of the tuning parameter $\lambda$ is crutial.
+Selecting the tuning parameter $\lambda$ is crutial to getting LASSO estimation for the $\beta$s.
 As $\lambda$ increases, more $\beta$ coefficients would be set to 0 and fewer variables will be selected by the model.
 When $\lambda$ is too small, the model tends to be redundant.
-And $\lambda$ also should not be set too large.
-In this case, the model would be too conservative to select any variables.
+When $\lambda$ is too large, the model would be too conservative to select any variables.
 
-One common approach for selecting the tuning parameter $\lambda$ for linear regression and logitic regression is via cross validation.
+One common approach for selecting the tuning parameter $\lambda$ for linear regression and logistic regression is via cross validation.
 The idea of K - fold cross validation is illustrated in Figure 1.
@@
 The model would first be built on the training set, then fitted to the test set to obtain cross validation error (CVE).
-To select the most suitable lambda,the CVE would be calculated for each candidate lambda.
-And the one that has the smallest cross validated error would be selected.
+CVE would be calculated for each candidate $\lambda$, then the one that minimizes cross validated error would be selected.
 
@@
 
-However, for penalized cox regression, there is no natural way to carry out cross validation.
-Since the partial likelihood of the cox model does not give estimates for the baseline, the outcomes can only be compared within the same framework, in terms of their relative risks.
+However, for penalized cox regression, it is challenging to carry out cross validation.
+Since the partial likelihood of the cox model does not give estimates for the baseline, it is only meaningful to compare the outcomes within the same framework, which can be interpreted as relative risks.
 This is against the idea of cross validation, which relies on splitting the data into different folds.
@@
 \par The most commonly used approach of conducting cross validation in Cox regression is proposed by Verweij elt.
-(1993).
-They propose a cross-validated log likelihood as a measure for the model's predictive accuracy.
-It is currently implemented in the glmnet package as the "grouped" option.
-Similarly, the glmnet package also includes the "ungrouped" option where the cross-validated likelihood is calculated only based on the observations from the test set.
+\citep{Verweij1993}.
+They proposed a cross-validated log likelihood as a measure for the model's predictive accuracy.
+It is currently implemented in the glmnet package with the "grouped" option.
+Similarly, the glmnet package also includes an "ungrouped" option where the cross-validated likelihood is calculated only based on the observations from the test set.
 In this paper, we proposed two alternative ways to carry out cross validation.
 Instead of cross-validating over partial likelihood, we propose to cross validate over the linear predictors of the regression model and cross validate over the Martingale residual of the Cox model.
+
 \par We conducted simulation studies to compare how those methods perform in both low dimensional and high dimensional settings with LASSO penalty.
@@
 We found out that using cross validation to select tuning parameters for LASSO penalized cox regression tends to be a conservative approach in general.
-Finally, we apply those methods to two genomic data sets with time-to-event outcomes.
+Finally, we apply those methods to two high dimensional data sets with time-to-event outcomes.
 
@@
   \subsection{Cross Validated Likelihood} 
+Suppose a data set of n observations is split into K folds and the ith fold is left out, then a partial likelihood can be built over the K -1 folds and yields maximum likelihood estimates of the coefficients, denoted by $\hat{\beta}_{-i}$.
+An intuitive way to carry out cross validation is to use the partial likelihood on the test data as the cross validated error: \begin{equation}cvl = \sum_{i=1}^{K} l_{i}(\hat{\beta}_{-i})\end{equation} This is implemented in the $\tt{glmnet}$ package as the $\tt{ungrouped}$ option.
 
-Suppose a data set of n observations is split into K folds and the ith fold is left out, then a partial likelihood can be built over the K -1 folds and yields maximum likelihood estimates of the coefficients, denoted by $\hat{\beta}_{-i}$.
-Verweij elt.
-defined cross-validated log likelihood of leaving the ith fold out to be \begin{equation}cvl = \sum_{i=1}^{K} \{l(\hat{\beta}_{-i}) - l_{-i}(\hat{\beta}_{-i})\}.
+\par An alternative way to calculate cross-validated log likelihood was proposed by Verweij elt.
+They defined cross-validated log likelihood of leaving the ith fold out to be \begin{equation}cvl = \sum_{i=1}^{K} \{l(\hat{\beta}_{-i}) - l_{-i}(\hat{\beta}_{-i})\}.
 \end{equation} When the ith fold is left out, $l(\hat{\beta}_{-i})$ is the log partial likelihood evaluated at $\hat{\beta}_{-i}$ with all k folds of observations.
@@
 
-An alternative way of conducting cross-validation is implemented in the glmnet package.
-The cross-validated log likelihood is defined as\begin{equation}cvl = \sum_{i=1}^{K} l_{i}(\hat{\beta}_{-i}).\end{equation} Instead of using substraction, $l_{i}(\hat{\beta}_{-i})$ is the partial likelihood evaluated at $\hat{\beta}_{-i}$ with observations from the fold that is left out.
-
-We refer to the first definition as grouped cvl and the second definition as ungrouped cvl.
+\par We refer to the first definition as ungrouped cross-validated partial likelihood (CVL) and the second definition as grouped cross-validated partial likelihood.
 Figure 2 illustrates the ideas of the two methods.
-The grouped cvl uses observations more efficiently then the ungrouped one.
-Since only the left-out fold would be used, the ungrouped cvl would not very stable if number of observations is too small.
+The grouped CVL uses observations more efficiently than the ungrouped CVL.
+Since only the left-out fold would be used, the ungrouped cvl is unstable in practice when there are only very few observations in some folder.
 
@@
     \subsubsection {Comparisons with Information Criteria}
-    We also compared cross-validation with information criteria AIC and BIC. ...
+    \par We also compared cross-validation with information criteria AIC and BIC.
+Simulation studies were conducted in lower dimension and higher dimension respectively.
+As is illustrated in Figure 7, in lower dimension, AIC tends to perform more liberal and BIC tends to perform more conservative than the optimal $\lambda$.
+While the cross-validation methods are also conservative, most of them are less conservative than BIC.
+In terms of MSE, cross-validation methods yield smaller MSE than the information criteria.
+In higher dimension, both BIC and AIC tend to be more liberal than the optimal $\lambda$ and yield larger MSE.
 
@@
 
-\bibliographystyle{ims-nourl}
+\begin{itemize}
+	\item Explanation for Martingale Residuals
+\end{itemize}
+
+\bibliographystyle{abbrvnat}
+\setcitestyle{round} 
 \bibliography{articles}

### 2017-09-13 cross-validation.Rnw
Add deviance residuals

@@
 Cox propotional hazard regression is one of the most commonly used statistical model for analyzing data with survival outcomes.
-Despite its popularity, the cox model has its own limitations due to the semiparametric nature of the model.
-Since the model only estimates relative risks, it is challenging to evaluate the model's predictive accuracy.
+Despite its popularity, the cox model is also limited by its the semiparametric nature.
+Since the model only yields estimations of relative risks, it is challenging to evaluate the model's predictive accuracy.
 

### 2017-09-13 cross-validation.tex
Add deviance residuals

@@
 In this paper, we proposed two alternative ways to carry out cross validation.
-Instead of cross-validating over partial likelihood, we propose to cross validate over the linear predictors of the regression model and cross validate over the Martingale residual of the Cox model.
+Instead of cross-validating over partial likelihood, we propose to cross validate over the linear predictors of the regression model and cross validate over the deviance residual \citep{Therneau1990}of the Cox model.
 
@@
 		  \includegraphics[height= 3.9cm ]{./figures/04.png}
-      \caption{Cross Validated Martingale Residuals}
+      \caption{Cross Validated Deviance Residuals}
       \end{minipage}	
@@
 
-  \subsection{Cross Validated Martingale Residuals}
+  \subsection{Cross Validated Deviance Residuals}
 
 Since the biggest challenge of conducting cross validation for cox regression is due to the missingness of baseline hazard, we propose a second approach which involves estimating the actual baseline hazard $\hat{\Lambda}_{0}$.
-Based on the observations in the training set, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method.
+As is mentioned in Therneau et al.'s original paper on Martingale Residuals \citep{Therneau1990}, the baseline hazard can be used to compute the Martingale residuals and deviance residuals, which are commonly used for Cox regression model diagnostics.
+
+\par The original idea of calcualting the Martingale and Deviance residuals were integrated into cross validation.
+Based on the observations in the training set, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}.
 Then for each observation j in the test set, the Martingale Residual can be calculated: \begin{equation}\hat{M_{j}} = d_{j} -\hat{\Lambda}_{0}(t_{j})e^{\hat{\eta}_{j}},\end{equation} where $d_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation.
-After repeating this procedure for all K folds, we would obtain a set of matingale residuals for all observations.
-They are be pulled together by sum of squares as the cross validated error:
-\begin{equation}CVE = \sum_{j}\hat{M}_{j}^2\end{equation}
+The deviance residual can be calcualted from the Martingale Residuals: \begin{equation} d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}\end{equation}
+
+After repeating this procedure for all K folds, we would obtain a set of deviance residuals for all observations. They are be pulled together by sum of squares as the cross validated error:
+\begin{equation}CVE = \sum_{j}\hat{d}_{j}^2\end{equation}.
+
+One advantage of using deviance residual is that it is symmetric while Martingale Residual is not.
+However, there does exist a numeric challenge when incorporating deviance residuals for cross validation.
+Since the baseline hazard function $\hat{\Lambda}_{0}$ is a step function from 0, the cumulative hazard at the first time point from 0 is always 0.
+Since deviance residual contains a term $log(\delta_{i} - \hat{M}_{i})$, where $\delta_{i} - \hat{M}_{i} = \hat{\Lambda}_{0}$, this term would be negative infinity for the first time point and lead to a numeric issue for whichever fold that contains the first time point.
+Our solution is to smooth out the step function at the beginning.
 
@@
 Ungrouped cross-validated log likelihood and the cross-validated linear predictor seem to be more liberal than the other two methods.
-The cross-validated martingale residual is the most conservative one and its performance is quite off from the other three methods.
+The cross-validated Deviance residual is the most conservative one and its performance is quite off from the other three methods.
 
@@
 Hence this approach is better at picking out signals than the other three approaches.
-If we look at the curve for the cross-validated Martingale Residual, there's almost no signal there.
+If we look at the curve for the cross-validated Deviance Residual, there's almost no signal there.
 
@@
 
-\begin{itemize}
-	\item Explanation for Martingale Residuals
-\end{itemize}
-
 \bibliographystyle{abbrvnat}

### 2017-09-18 cross-validation.tex
Comments on estimating baseline hazard for deviance resiudals

@@
 
+When invented, both Martingale residuals and Deviance residuals were calculated only for observations that construct the baseline hazard, not for out-of-sample predictions.
+However, in the cross-validation context, residuals are calculated for observations that are not used to construct the baseline hazard.
+It is common to run into situations where some events in the test set occurred after the last event occured in the baseline, where the survival function dropped to 0.
+For those events, they would be treated as censored.
+
 \section{Simulation Studies}

### 2017-10-03 cross-validation.tex
Add graphs for stability
correct index and change figures

@@
 
-In the Cox model \citep{Cox1975}, the hazard function for an individual observation j is given by \begin{equation}h_{j}(t) = h_{0}(t) \text{exp}( X_{j}^{T} \beta)\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{j}^{T} \beta}$ is the relative risk.
-The estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood \begin{equation}L(\beta) = \prod_{j=1}^{n} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation} As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard.
+In the Cox model \citep{Cox1975}, the hazard function for an individual observation i is given by \begin{equation}h_{i}(t) = h_{0}(t) \text{exp}( X_{i}^{T} \beta)\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{i}^{T} \beta}$ is the relative risk.
+Suppose there are n individual observations and m observed failure times, the estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood \begin{equation}L(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation}, where j indexes the observed failure times and $R(t_{j})$ denotes the set of observations at risk at time $t_{j}$.
+As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard.
 Hence the interpretation of the Cox model is only valid in a relative sense.
@@
 Penalized Cox Regression is an extension of the regular cox regression model.
-Estimates of the $\beta$ coefficients are obtained by minimizing the objective function \begin{equation}  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{n} \text{log} \left \{\frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta).
-\end{equation}, instead of optimizing the partial likelihood.
+Estimates of the $\beta$ coefficients are obtained by minimizing the objective function \begin{equation}  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{m} \text{log} \left \{\frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta).
+\end{equation} \citep{Simon2011}.
 The first part of the objective function is based on the partial likelihood.
@@
 \hat{\eta}_{-K})$ can be obtained.
-A partial likelihood can be built over this set of linear predictors: \begin{equation} L(\hat{\eta}_{-}) = \prod_{j=1}^{n} \frac{exp (\hat{\eta}_{-j})}{\sum_{ k \in R(t_{j})}exp (\hat{\eta}_{-k})}\end{equation}.
+A partial likelihood can be built over this set of linear predictors: \begin{equation} L(\hat{\eta}_{-}) = \prod_{j=1}^{m} \frac{exp (\hat{\eta}_{-j})}{\sum_{ k \in R(t_{j})}exp (\hat{\eta}_{-k})}\end{equation}.
 
@@
 Based on the observations in the training set, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}.
-Then for each observation j in the test set, the Martingale Residual can be calculated: \begin{equation}\hat{M_{j}} = d_{j} -\hat{\Lambda}_{0}(t_{j})e^{\hat{\eta}_{j}},\end{equation} where $d_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation.
+Then for each observation in the test set, the Martingale Residual can be calculated: \begin{equation}\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}_{i}},\end{equation} where $\delta_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation.
 The deviance residual can be calcualted from the Martingale Residuals: \begin{equation} d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}\end{equation}
@@
 After repeating this procedure for all K folds, we would obtain a set of deviance residuals for all observations. They are be pulled together by sum of squares as the cross validated error:
-\begin{equation}CVE = \sum_{j}\hat{d}_{j}^2\end{equation}.
+\begin{equation}CVE = \sum_{i}\hat{d}_{i}^2\end{equation}.
 
@@
   \subsection{Simulation Results}
+
+    \subsubsection {Stability}
+   While the ungrouped cross-validated likelihood is an intuitive way to carry out cross-validation, it is quite unstable when it runs into cases where the number of observed events in some fold is really small.
+As is illustrated in Figure 5, cases of undefined cross-validation error would increase when censoring increases or when number of folds increases.
+
+\begin{figure}
+    \centering
+		\includegraphics[height= 8cm ]{./figures/stability2.png}
+    \caption{Comparing Stability}
+\end{figure}
 
@@
 \section{Discussion}
+	\begin{itemize}
+		\item 	
+		\item
+	\end{itemize}
 

### 2018-02-12 cross-validation.tex
making changes for cross validated deviance

@@
 
-In the Cox model \citep{Cox1975}, the hazard function for an individual observation i is given by \begin{equation}h_{i}(t) = h_{0}(t) \text{exp}( X_{i}^{T} \beta)\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{i}^{T} \beta}$ is the relative risk.
-Suppose there are n individual observations and m observed failure times, the estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood \begin{equation}L(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation}, where j indexes the observed failure times and $R(t_{j})$ denotes the set of observations at risk at time $t_{j}$.
+In the Cox model \citep{Cox1975}, the hazard function for an observation i is given by \begin{equation}h_{i}(t) = h_{0}(t) \text{exp}( X_{i}^{T} \beta)\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{i}^{T} \beta}$ is the relative risk.
+Suppose there are n observations and m observed failure times, the estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood \begin{equation}L(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation}, where j indexes the observed failure times and $R(t_{j})$ denotes the set of observations at risk at time $t_{j}$.
 As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard.
@@
 
-Selecting the tuning parameter $\lambda$ is crutial to getting LASSO estimation for the $\beta$s.
-As $\lambda$ increases, more $\beta$ coefficients would be set to 0 and fewer variables will be selected by the model.
+Selecting the tuning parameter $\lambda$ is crutial to getting LASSO estimations for $\beta$.
+As $\lambda$ increases, more coefficients would be set to 0 and fewer variables will be selected by the model.
 When $\lambda$ is too small, the model tends to be redundant.
@@
 
+However, for penalized cox regression, it is challenging to carry out cross validation.
+Since the partial likelihood of the cox model does not give estimates for the baseline, it is not meaningful to compare the coefficients if they are not estimated from different models and different observations.
+Only if the baseline is the same, then the coefficients can be interpreted as relative risks.
+This is against the idea of cross validation, which splits the data and builds models on different subsets.
+
 \begin{figure}
@@
 
-However, for penalized cox regression, it is challenging to carry out cross validation.
-Since the partial likelihood of the cox model does not give estimates for the baseline, it is only meaningful to compare the outcomes within the same framework, which can be interpreted as relative risks.
-This is against the idea of cross validation, which relies on splitting the data into different folds.
-
-\par The most commonly used approach of conducting cross validation in Cox regression is proposed by Verweij elt.
-\citep{Verweij1993}.
-They proposed a cross-validated log likelihood as a measure for the model's predictive accuracy.
-It is currently implemented in the glmnet package with the "grouped" option.
-Similarly, the glmnet package also includes an "ungrouped" option where the cross-validated likelihood is calculated only based on the observations from the test set.
+%ZZZZZZZZZZZZZ
+\par The most intuitive approach to conduct cross validation in Cox regression is i) fit the model to training set then ii) calculate cross-validated partial likelihood based on the observations from the test set as a measure for the model's predictive accuracy.
+As will be discussed in Section 3, this approach is unstable when number of events gets small.
+Verweij elt.
+\citep{Verweij1993} proposed a stablized version of cross-validated log likehood, which has been widely used and implemented in R packages such as $\tt{glmnet}$.
 In this paper, we proposed two alternative ways to carry out cross validation.
@@
     \centering
-    \begin{minipage}[b]{0.5\textwidth}
+    \begin{minipage}[b]{0.45\textwidth}
       \centering
-		  \includegraphics[height= 3.5cm ]{./figures/03.png}
-      \caption{Cross Validated Linear Predictors}
+		  \includegraphics[height= 3.9cm ]{./figures/03.png}
+      \caption{CV Linear Predictors}
      \end{minipage}
-     \begin{minipage}[b]{0.5\textwidth}
+     \begin{minipage}[b]{0.45\textwidth}
       \centering
-		  \includegraphics[height= 3.9cm ]{./figures/04.png}
-      \caption{Cross Validated Deviance Residuals}
+		  \includegraphics[height= 4.1cm ]{./figures/04_2.png}
+      \caption{CV Deviance Residuals}
       \end{minipage}	
@@
   \subsection{Cross Validated Deviance Residuals}
+  %Need to update this section with the new baseline estimation method%
+Since the biggest challenge of conducting cross validation for cox regression is due to the missingness of baseline hazard, we propose a second approach which involves estimating the actual baseline hazard $\hat{\Lambda}_{0}$.
+The sum of squares of Martingale residuals is a natural candicate for evaluating the model performance using baseline hazard.
+Martingale residuals and its normalized form, deviance residuals, were first proposed by \citep{Therneau1990} for Cox regression model diagnostics.
 
-Since the biggest challenge of conducting cross validation for cox regression is due to the missingness of baseline hazard, we propose a second approach which involves estimating the actual baseline hazard $\hat{\Lambda}_{0}$.
-As is mentioned in Therneau et al.'s original paper on Martingale Residuals \citep{Therneau1990}, the baseline hazard can be used to compute the Martingale residuals and deviance residuals, which are commonly used for Cox regression model diagnostics.
+To estimate the baseline hazard, a set of linear predictors $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...
+\hat{\eta}_{-K})$ would be obtained in the same way as the cross validated linear predictor approach.
+$\hat{\eta}_{-i}$ is computed based on $\hat{\beta}_{-i}$s estimated from the training set and observations from the ith test set.
+With $\hat{\eta}_{-}$, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}.
+Then for each observations, the Martingale Residual can be calculated: \begin{equation}\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}_{i}},\end{equation} where $\delta_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation.
+The deviance residual can be calcualted from the Martingale Residuals: \begin{equation} d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}\end{equation} The sum of squares of the deviance residuals, $\sum_{i}\hat{d}_{i}^2$, would be used as the cross validated error.
 
-\par The original idea of calcualting the Martingale and Deviance residuals were integrated into cross validation.
-Based on the observations in the training set, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}.
-Then for each observation in the test set, the Martingale Residual can be calculated: \begin{equation}\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}_{i}},\end{equation} where $\delta_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation.
-The deviance residual can be calcualted from the Martingale Residuals: \begin{equation} d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}\end{equation}
-
-After repeating this procedure for all K folds, we would obtain a set of deviance residuals for all observations. They are be pulled together by sum of squares as the cross validated error:
-\begin{equation}CVE = \sum_{i}\hat{d}_{i}^2\end{equation}.
-
-One advantage of using deviance residual is that it is symmetric while Martingale Residual is not.
-However, there does exist a numeric challenge when incorporating deviance residuals for cross validation.
+As an alternative, the baseline hazard can also be estimated only based on the training set, leaving the test set as out-of-sample observations for calculating the residuals.
+There are two major issues with this approach.
+First, it would be common to run into situations where events in the test set occurred after the last event occured in the baseline, where the survival function dropped to 0.
+Unless those observations are treated as censored, this approach would not work.
+Second, there would be an extra numeric challenge for deviance residuals.
 Since the baseline hazard function $\hat{\Lambda}_{0}$ is a step function from 0, the cumulative hazard at the first time point from 0 is always 0.
 Since deviance residual contains a term $log(\delta_{i} - \hat{M}_{i})$, where $\delta_{i} - \hat{M}_{i} = \hat{\Lambda}_{0}$, this term would be negative infinity for the first time point and lead to a numeric issue for whichever fold that contains the first time point.
-Our solution is to smooth out the step function at the beginning.
-
-When invented, both Martingale residuals and Deviance residuals were calculated only for observations that construct the baseline hazard, not for out-of-sample predictions.
-However, in the cross-validation context, residuals are calculated for observations that are not used to construct the baseline hazard.
-It is common to run into situations where some events in the test set occurred after the last event occured in the baseline, where the survival function dropped to 0.
-For those events, they would be treated as censored.
+Hence extra smoothing for the step function would be needed for this approach to work.
 
 \section{Simulation Studies}
+%reformulate this section with%
+%3.1 comparison between the three methods
+%3.2 discussion about stability issue for ungrouped methods
+%3.3 comparison with AIC
+%3.4 explaining why deviance residuals do not work well
 
@@
 200 replications were used.
-Censoring percentage, number of folds used in cross-validation, number of non-zero $\beta$s and the magnitude of the $\beta$s were varied in several different simulation scenarios.
+Censoring percentage, number of folds used in cross-validation, number of non-zero $\beta$s and the magnitude of the $\beta$000000000s were varied in several different simulation scenarios.
 Patterns and results shown in various simulation scenarios are consistent with one another.

### 2018-02-13 cross-validation.tex
reorganize simulation section

@@
 \section{Simulation Studies}
-%reformulate this section with%
-%3.1 comparison between the three methods
-%3.2 discussion about stability issue for ungrouped methods
-%3.3 comparison with AIC
-%3.4 explaining why deviance residuals do not work well
-
-  \subsection{Simulation Settings}
 
@@
 
-  \subsection{Simulation Results}
-
-    \subsubsection {Stability}
-   While the ungrouped cross-validated likelihood is an intuitive way to carry out cross-validation, it is quite unstable when it runs into cases where the number of observed events in some fold is really small.
-As is illustrated in Figure 5, cases of undefined cross-validation error would increase when censoring increases or when number of folds increases.
-
-\begin{figure}
-    \centering
-		\includegraphics[height= 8cm ]{./figures/stability2.png}
-    \caption{Comparing Stability}
-\end{figure}
-
-    \subsubsection {Simulations in Low Dimensional Settings}
-
+    \subsection {Simulations Comparing Cross-validation Approaches}
+  %need to change figures to include the new cross-validated deviance residuals
    The simulation experiments were first conducted under lower dimension settings.
@@
 200 replications were used.
-Censoring percentage, number of folds used in cross-validation, number of non-zero $\beta$s and the magnitude of the $\beta$000000000s were varied in several different simulation scenarios.
+Censoring percentage, number of folds used in cross-validation, number of non-zero $\beta$s and the magnitude of the $\beta$s were varied in several different simulation scenarios.
 Patterns and results shown in various simulation scenarios are consistent with one another.
@@
 
-    \subsubsection {Simulations in High Dimensional Setting}
-
    Simulation studies were also conducted in high dimension scenarios, where the number of observations and dimension of the covariate matrix is more similar to real genetics research.
@@
 
-  \vspace{20 mm}
-    \subsubsection {Comparisons with Information Criteria}
+ \subsection {Stability}
+   While the ungrouped cross-validated likelihood is an intuitive way to carry out cross-validation, it is quite unstable when it runs into cases where the number of observed events in some fold is really small.
+As is illustrated in Figure 5, cases of undefined cross-validation error would increase when censoring increases or when number of folds increases.
+
+\begin{figure}
+    \centering
+		\includegraphics[height= 8cm ]{./figures/stability2.png}
+    \caption{Comparing Stability}
+\end{figure}
+
+    \subsection {Comparisons with Information Criteria}
+	%need to change the figures to have fewer lines
     \par We also compared cross-validation with information criteria AIC and BIC.
@@
    \end{figure}
+
+ \subsection {Issues with Deviance Residual}
 

### 2018-02-25 simulation/deviance.R
start keeping track of simulation code

@@ -1,0 +1,116 @@
+# cross validation using deviance approach
+
+######################################
+# Functions Produce Sums of Squares: #
+######################################
+
+# Using Kalbfleisch-Prentice's method (discrete estimation)
+loss.deviance.KP <- function (y, eta) 
+{
+  ind <- order(y[, 1])
+  d <- as.numeric(y[ind, 2])
+  # d, eta, and w are all ordered
+  # more than 2 lambda:
+  if (is.matrix(eta)) {
+    w <- exp(eta[ind, , drop = FALSE])
+    val <- apply(w, 2, function(x){ 
+      r <- rev(cumsum(rev(x)))
+      a <- ifelse(d, (1 - x/r)^(1/x), 1)
+      BH0 <- cumsum(1-a)
+      M <- d - BH0*x
+      dev <- sign(M) * sqrt(-2 * (M + ifelse(d == 0, 0, d * 
+                                               log(d - M))))
+      #dev <- (- M - ifelse(d == 0, 0, d * log(d - M)))
+      sum(dev^2)})
+  }
+  else {
+    #construct baseline hazard
+    w <- exp(eta[ind])
+    r <- rev(cumsum(rev(w)))
+    a <- ifelse(d, (1 - w/r)^(1/w), 1)
+    BH0 <- cumsum(1-a)
+    M <- d - BH0*w
+    dev <- sign(M) * sqrt(-2 * (M + ifelse(d == 0, 0, d * 
+                                             log(d - M))))
+    #dev <- (- M - ifelse(d == 0, 0, d * log(d - M)))
+    #dev <- - d - d*log(BH0) - d*eta[ind]
+    val <- sum(dev^2)
+    #val = list(BH0 = BH0, dev = dev)
+  }
+  return(val)
+}
+
+
+###################################################
+# CV functions that takes the baseline estimation #
+###################################################
+cvf.surv <- function (i, XX, y, cv.ind, cv.args) 
+{
+  cv.args$X <- XX[cv.ind != i, , drop = FALSE]
+  cv.args$y <- y[cv.ind != i, ]
+  fit.i <- do.call("ncvsurv", cv.args)
+  X2 <- XX[cv.ind == i, , drop = FALSE]
+  y2 <- y[cv.ind == i, ]
+  nl <- length(fit.i$lambda)
+  yhat <- predict(fit.i, X2)
[…]

### 2018-02-25 simulation/modify_fold.R

@@ -1,0 +1,66 @@
+################################################
+# modify function cv.coxnet in glmnet package: #
+################################################
+
+# so that ungroup will always be ungroup; 
+# won't be automatically changed into grouped
+
+library(glmnet)
+library(R.utils)
+mycv.coxnet <- function (outlist, lambda, x, y, weights, offset, foldid, type.measure, 
+          grouped, keep = FALSE) 
+{
+  typenames = c(deviance = "Partial Likelihood Deviance")
+  if (type.measure == "default") 
+    type.measure = "deviance"
+  if (!match(type.measure, c("deviance"), FALSE)) {
+    warning("Only 'deviance'  available for Cox models; changed to type.measure='deviance'")
+    type.measure = "deviance"
+  }
+  if (!is.null(offset)) {
+    is.offset = TRUE
+    offset = drop(offset)
+  }
+  else is.offset = FALSE
+  nfolds = max(foldid)
+  #if ((length(weights)/nfolds < 10) && !grouped) {
+  #  warning("Option grouped=TRUE enforced for cv.coxnet, since < 3 observations per fold", 
+  #          call. = FALSE)
+  #  grouped = TRUE
+  #}
+  cvraw = matrix(NA, nfolds, length(lambda))
+  mlami = max(sapply(outlist, function(obj) min(obj$lambda)))
+  which_lam = lambda >= mlami
+  for (i in seq(nfolds)) {
+    which = foldid == i
+    fitobj = outlist[[i]]
+    coefmat = predict(fitobj, type = "coeff", s = lambda[which_lam])
+    if (grouped) {
+      plfull = coxnet.deviance(x = x, y = y, offset = offset, 
+                               weights = weights, beta = coefmat)
+      plminusk = coxnet.deviance(x = x[!which, ], y = y[!which, 
+                                                        ], offset = offset[!which], weights = weights[!which], 
+                                 beta = coefmat)
+      cvraw[i, seq(along = plfull)] = plfull - plminusk
+    }
+    else {
+      plk = coxnet.deviance(x = x[which, ], y = y[which, 
+                                                  ], offset = offset[which], weights = weights[which], 
+                            beta = coefmat)
+      cvraw[i, seq(along = plk)] = plk
+    }
+  }
+  status = y[, "status"]
[…]

### 2018-02-25 simulation/scenario_1.R

@@ -1,0 +1,168 @@
+# Scenario 1 (plots the lambda selection and MSE ratio)
+# compare four cross validation methods: 
+#  ungrouped, grouped, linear predictor, deviance residuals
+# varying magnitude of betas at higher dimension (n = 100, p = 1000, nonzero = 10)
+
+
+# load source files
+source("survival_functions.R")
+
+# load modified ungrouped method
+source("modify_fold.R")
+
+# load deviance residuals
+source("deviance.R")
+
+#############################################################################
+# comparison 1 : n = 200, p = 1000, nbeta = 10, censoring = 10% , nfold = 10
+#############################################################################
+# change the magnitude of betas
+n <- 200
+p <- 1000
+c <-  c(0.5,0.75,1,1.25,1.5)
+beta <- matrix(NA,nrow = length(c), ncol = p)
+for (i in 1:length(c)){
+  beta[i,]<- generatebeta(p,nbeta = 10 ,c = c[i])
+}
+N <- 200
+# initiation
+true.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+gp.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+ug.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+ncv.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+dvr.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+
+
+true.mse <- matrix(NA,nrow = N ,ncol = length(c))
+gp.mse <- matrix(NA,nrow = N ,ncol = length(c))
+ug.mse <- matrix(NA,nrow = N ,ncol = length(c))
+ncv.mse <- matrix(NA,nrow = N ,ncol = length(c))
+pd.mse <- matrix(NA,nrow = N ,ncol = length(c))
+dvr.mse <- matrix(NA,nrow = N ,ncol = length(c))
+
+#i <- 1
+#j <- 1
+for(i in 1:N){
+  set.seed(999 + i)
+  foldid <- sample(rep(seq(10),length = n))
+  for (j in 1:length(c)){
[…]

### 2018-02-25 simulation/scenario_1_2.R

@@ -1,0 +1,168 @@
+# Scenario 1 (plots the lambda selection and MSE ratio)
+# compare four cross validation methods: 
+#  ungrouped, grouped, linear predictor, deviance residuals
+# varying magnitude of betas at low dimension (n = 100, p = 10, nonzero = 2)
+
+
+# load source files
+source("survival_functions.R")
+
+# load modified ungrouped method
+source("modify_fold.R")
+
+# load deviance residuals
+source("deviance.R")
+
+#############################################################################
+# comparison 1 : n = 100, p = 10, nbeta = 2, censoring = 10% , nfold = 10
+#############################################################################
+# change the magnitude of betas
+n <- 100
+p <- 10
+c <-  c(0.5,0.75,1,1.25,1.5)
+beta <- matrix(NA,nrow = length(c), ncol = p)
+for (i in 1:length(c)){
+  beta[i,]<- generatebeta(p,nbeta = 2 ,c = c[i])
+}
+N <- 50
+# initiation
+true.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+gp.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+ug.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+ncv.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+dvr.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+
+
+true.mse <- matrix(NA,nrow = N ,ncol = length(c))
+gp.mse <- matrix(NA,nrow = N ,ncol = length(c))
+ug.mse <- matrix(NA,nrow = N ,ncol = length(c))
+ncv.mse <- matrix(NA,nrow = N ,ncol = length(c))
+pd.mse <- matrix(NA,nrow = N ,ncol = length(c))
+dvr.mse <- matrix(NA,nrow = N ,ncol = length(c))
+
+#i <- 1
+#j <- 1
+for(i in 1:N){
+  set.seed(999 + i)
+  foldid <- sample(rep(seq(10),length = n))
+  for (j in 1:length(c)){
[…]

### 2018-02-25 simulation/scenario_1_3.R

@@ -1,0 +1,168 @@
+# Scenario 1 (plots the lambda selection and MSE ratio)
+# compare four cross validation methods: 
+#  ungrouped, grouped, linear predictor, deviance residuals
+# varying magnitude of betas at low dimension (n = 100, p = 100, nonzero = 10)
+
+
+# load source files
+source("survival_functions.R")
+
+# load modified ungrouped method
+source("modify_fold.R")
+
+# load deviance residuals
+source("deviance.R")
+
+#############################################################################
+# comparison 1 : n = 100, p = 100, nbeta = 10, censoring = 10% , nfold = 10
+#############################################################################
+# change the magnitude of betas
+n <- 100
+p <- 100
+c <-  c(0.5,0.75,1,1.25,1.5)
+beta <- matrix(NA,nrow = length(c), ncol = p)
+for (i in 1:length(c)){
+  beta[i,]<- generatebeta(p,nbeta = 10 ,c = c[i])
+}
+N <- 50
+# initiation
+true.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+gp.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+ug.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+ncv.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+dvr.lambda <- matrix(NA,nrow = N ,ncol = length(c))
+
+
+true.mse <- matrix(NA,nrow = N ,ncol = length(c))
+gp.mse <- matrix(NA,nrow = N ,ncol = length(c))
+ug.mse <- matrix(NA,nrow = N ,ncol = length(c))
+ncv.mse <- matrix(NA,nrow = N ,ncol = length(c))
+pd.mse <- matrix(NA,nrow = N ,ncol = length(c))
+dvr.mse <- matrix(NA,nrow = N ,ncol = length(c))
+
+#i <- 1
+#j <- 1
+for(i in 1:N){
+  set.seed(999 + i)
+  foldid <- sample(rep(seq(10),length = n))
+  for (j in 1:length(c)){
[…]

### 2018-02-25 simulation/scenario_2.R

@@ -1,0 +1,285 @@
+# Scenario 1 (plots the lambda selection and MSE ratio)
+# compare four cross validation methods: 
+#  ungrouped, grouped, linear predictor, deviance residuals
+# varying dimension (n = 100, p = 100, nonzero = 10)
+
+# load source files
+source("survival_functions.R")
+
+# load modified ungrouped method
+source("modify_fold.R")
+
+# load deviance residuals
+source("deviance.R")
+
+#############################################################################
+# scenario 2_1 : n = 100, p = 20, nbeta = 10, censoring = 10% , nfold = 10
+#############################################################################
+# change the magnitude of betas
+n <- 100
+p <- 20
+#cpct <- seq(0,0.3,by = 0.1)
+cpct <- 0.1
+beta <- generatebeta(p,nbeta = 10 ,c = 1)
+N <- 200
+
+# initiation
+true.lambda.1 <- matrix(NA,nrow = N ,ncol = length(cpct))
+gp.lambda.1 <- matrix(NA,nrow = N ,ncol = length(cpct))
+ug.lambda.1 <- matrix(NA,nrow = N ,ncol = length(cpct))
+ncv.lambda.1 <- matrix(NA,nrow = N ,ncol = length(cpct))
+dvr.lambda.1 <- matrix(NA,nrow = N ,ncol = length(cpct))
+
+true.mse.1 <- matrix(NA,nrow = N ,ncol = length(cpct))
+gp.mse.1 <- matrix(NA,nrow = N ,ncol = length(cpct))
+ug.mse.1 <- matrix(NA,nrow = N ,ncol = length(cpct))
+ncv.mse.1 <- matrix(NA,nrow = N ,ncol = length(cpct))
+dvr.mse.1 <- matrix(NA,nrow = N ,ncol = length(cpct))
+
+for(i in 1:N){
+  set.seed(999 + i)
+  foldid <- sample(1:10,size=n,replace=TRUE)
+  data <- genSurvData(n=n , beta = beta, cpct = cpct,sd = 1)
+  data$x <- unclass(ncvreg::std(data$x))
+  
[…]

### 2018-02-25 simulation/scenario_3.R

@@ -1,0 +1,238 @@
+##############################################
+# compare the stability of different methods #
+##############################################
+
+# ungrouped
+# grouped
+# linear predictor
+# deviance residuals
+
+# 3.1 vary censoring %
+# 3.2 vary fold numbers
+
+
+# load source files
+source("survival_functions.R")
+
+# load modified ungrouped method
+source("modify_fold.R")
+
+# load deviance residuals
+source("deviance.R")
+
+
+
+#######################
+### vary censoring ####
+#######################
+
+n <- 100
+p <- 100
+cpct <- seq(0,0.5,by = 0.1)
+beta <- generatebeta(p,nbeta = 10 ,c = 1)
+N <- 2
+# initiation
+true.lambda.cen<- matrix(NA,nrow = N ,ncol = length(cpct))
+gp.lambda.cen <- matrix(NA,nrow = N ,ncol = length(cpct))
+ug.lambda.cen <- matrix(NA,nrow = N ,ncol = length(cpct))
+ncv.lambda.cen <- matrix(NA,nrow = N ,ncol = length(cpct))
+dvr.lambda.cen <- matrix(NA,nrow = N ,ncol = length(cpct))
+
+true.mse.cen <- matrix(NA,nrow = N ,ncol = length(cpct))
+gp.mse.cen <- matrix(NA,nrow = N ,ncol = length(cpct))
+ug.mse.cen <- matrix(NA,nrow = N ,ncol = length(cpct))
+ncv.mse.cen <- matrix(NA,nrow = N ,ncol = length(cpct))
+dvr.mse.cen <- matrix(NA,nrow = N ,ncol = length(cpct))
+
+
+for(i in 1:N){
+  set.seed(666 + i)
+  foldid <- sample(1:10,size=n,replace=TRUE)
+  for (j in 1:length(cpct)){
+    data <- genSurvData(n=100 , beta = beta, cpct = cpct[j])
+    f <- try(survncvreg(data$x,data$y,data$beta))
+    cv.gp <- try(cv.glmnet(x = data$x,y = data$y,nfolds = 10,foldid=foldid,lambda = f$lambda,family = "cox"))
[…]

### 2018-02-25 simulation/survival_functions.R

@@ -1,0 +1,179 @@
+################################
+# File that include functions 
+# that are used in simulations to: 
+# 1. generate data
+# 2. run penalized regression models 
+# 3. extract simulation results
+library(glmnet)
+library(survival)
+library(ncvreg)
+
+##############################
+# function that generates beta
+# with p as the length
+# nbeta as non-zero betas
+##############################
+
+generatebeta <- function(p,nbeta,c = 1){
+  pos <- rep(c, ceiling(nbeta/2))
+  neg <- rep(-c, floor(nbeta/2))
+  beta <- c(pos,neg,rep(0,p-nbeta))
+}
+
+#########################################
+# function that generates time-to-event
+# data based on exponential distribution
+#########################################
+genSurvData <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,h = 1 # 70% means 70% of the data are censored; 30% are observed
+){
+  p <- length(beta)
+  ## Generate column vector of coefficients
+  beta <- matrix(beta, ncol = 1)
+  x <- matrix(rnorm(n*p, sd = sd),n,p)
+  #standardize x first
+  x <- ncvreg::std(x)
+  hazard <- h*exp(x %*% beta)
+  ## Calculate survival times
+  time <- rexp(n, hazard)
+  ## observations are censored ("event = rbinom(n, 1, 0.30)")
+  y <- Surv(time = time, event = rbinom(n, 1, 1-cpct))
+  return(list(x=x,y=y,beta = beta))
+}
+
+genSurvData_weibull <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,h = 1,rho = 2 # 70% means 70% of the data are censored; 30% are observed
+){
+  p <- length(beta)
+  ## Generate column vector of coefficients
+  beta <- matrix(beta, ncol = 1)
+  x <- matrix(rnorm(n*p, sd = sd),n,p)
+  #standardize x first
+  x <- ncvreg::std(x)
+  v <- runif(n)
+  time <- (-log(v)/(h*exp(x%*%beta)))^(1/rho)
+  
+  ## observations are censored ("event = rbinom(n, 1, 0.30)")
+  y <- Surv(time = time, event = rbinom(n, 1, 1-cpct))
[…]

### 2018-02-27 cross-validation.tex
add new simulation figures and simulation code

@@
       \centering
-		  \includegraphics[height= 6cm ]{./figures/05.jpeg}
+		  \includegraphics[height= 6cm ]{./figures/05.png}
       \caption{ }
@@
       \centering
-		  \includegraphics[height= 6cm ]{./figures/06.jpeg}
+		  \includegraphics[height= 6cm ]{./figures/06.png}
       \caption{ }
@@
 
+ % \begin{table}[h]
+   % \centering
+%			\begin{tabular}{ c | c | c | c}
+%				CV Methods & $\lambda$ (SD) & MSE Ratio \\
+%				\hline \hline
+%				Optimal  & 0.0606 ( 0.0019 )&  / & \\ 
+%				Grouped  &  0.0895  ( 0.0043 )& 1.598  \\
+%				Ungrouped & 0.0754 ( 0.0060 )& 1.225 \\
+%				Linear Pred & 0.0759 ( 0.0058 )& 1.234 \\
+%				M Residual &0.1283 ( 0.0047 ) & 2.959
+%			\end{tabular}
+%			\caption{Simulation Results in High Dimensional Setting}
+%	\end{table}
+
 \begin{table}[h]
+\small
 \centering
-			\begin{tabular}{ c | c | c | c}
-				CV Methods & $\lambda$ (SD) & MSE Ratio \\
-				\hline \hline
-				Optimal  & 0.0606 ( 0.0019 )&  / & \\ 
-				Grouped  &  0.0895  ( 0.0043 )& 1.598  \\
-				Ungrouped & 0.0754 ( 0.0060 )& 1.225 \\
-				Linear Pred & 0.0759 ( 0.0058 )& 1.234 \\
-				M Residual &0.1283 ( 0.0047 ) & 2.959
+\caption{Simulation Results, Varying n and p}
+\label{my-label}
+\begin{tabular}{llllllllllllllll}
+\hline
+       & \multicolumn{3}{l}{n = 100, p = 20} &  & \multicolumn{3}{l}{n = 100, p = 100} &  & \multicolumn{3}{l}{n = 200, p = 1000} &  & \multicolumn{3}{l}{n = 500, p = 10000} \\
+       & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 20} \\ \cline{2-4} \cline{6-8} \cline{10-12} \cline{14-16}
+       &     &        &MSE       &  &   &       & MSE     &    &    &        & MSE      &  &     &        & MSE      \\
+       & $\lambda$(SD)     & &Ratio       &  & $\lambda$(SD)         & & Ratio     & &$\lambda$(SD)         & & Ratio      & & $\lambda$(SD)         & & Ratio      \\ \cline{2-2} \cline{4-4} \cline{6-6} \cline{8-8} \cline{10-10} \cline{12-12} \cline{14-14} \cline{16-16}
+Optimal &           &        &          &  &           &       &          &  &           &        &          &  &           &        &          \\
+Grouped &           &        &          &  &           &       &          &  &           &        &          &  &           &        &          \\
+Ungrouped &           &        &          &  &           &       &          &  &           &        &          &  &           &        &          \\
+Linear Pred &           &        &          &  &           &       &          &  &           &        &          &  &           &        &          \\
+Deviance &           &        &          &  &           &       &          &  &           &        &          &  &           &        &          \\ \hline
 \end{tabular}
-			\caption{Simulation Results in High Dimensional Setting}
 \end{table}

### 2018-02-27 simulation/scenario_1.R

@@ -81,12 +81,11 @@ for(i in 1:N){
 #apply(na.omit(ncv.lambda),2,mean)
 #apply(na.omit(dvr.lambda),2,mean)
 
-save(true.lambda,gp.lambda,ug.lambda,ncv.lambda,dvr.lambda,
-     gp.mse,ug.mse,ncv.mse,dvr.mse,true.mse,
-     file = "sim_1.RData")
+#save(true.lambda,gp.lambda,ug.lambda,ncv.lambda,dvr.lambda,
+#     gp.mse,ug.mse,ncv.mse,dvr.mse,true.mse,
+#     file = "sim_1.RData")
 
-#load(file = "sim_1.RData")
-
+load(file = "sim_1.RData")
 # lambda
 com_result1 <- matrix(NA,nrow = 5,ncol = length(c))
 com_result1[1,] <- apply(true.lambda,2,mean)
@@ -138,10 +137,10 @@ colnames(mse_ratio) <- c
 rownames(mse_ratio) <- c("grouped","ungrouped","ncvreg","deviance")
 
 #plot lambda
-length(c)
+png(filename= "sim_1_lambda.png")
 maintitle <- expression(paste(lambda," vs the magnitude of non-zero ", beta, "s", sep = ""))
 xlab <- expression(paste("the magnitude of non-zero ", beta, "s", sep = ""))
-plot(x = c, y = com_result1[1,],ylim = c(0.01, 0.3), type = "l",lwd = 2,main = maintitle,
+plot(x = c, y = com_result1[1,],ylim = c(0, 0.20), type = "l",lwd = 2,main = maintitle,
      ylab = expression(lambda),xlab = xlab, axes = FALSE)
 axis(side = 1, at = c,labels = c)
 axis(side = 2,las = 1)
@@ -150,12 +149,13 @@ lines(x = c, y = com_result1[3,],lty = 1,col = 3,lwd = 2)
 lines(x = c, y = com_result1[4,],lty = 1,col = 4,lwd = 2)
 lines(x = c, y = com_result1[5,],lty = 1,col = 6,lwd = 2)
 legend("topright",legend = c("Optimal","Grouped","Ungrouped","Linear Pred","Deviance"),lty =1,col = c("black",2:4,6))
+dev.off()
 
 #plot MSE ratio
-length(c)
+png(filename= "sim_1_mse.png")
[…]

### 2018-02-27 simulation/scenario_1_2.R

@@ -138,10 +138,10 @@ colnames(mse_ratio) <- c
 rownames(mse_ratio) <- c("grouped","ungrouped","ncvreg","deviance")
 
 #plot lambda
-length(c)
+png(filename= "sim_1_2_lambda.png")
 maintitle <- expression(paste(lambda," vs the magnitude of non-zero ", beta, "s", sep = ""))
 xlab <- expression(paste("the magnitude of non-zero ", beta, "s", sep = ""))
-plot(x = c, y = com_result1[1,],ylim = c(0.01, 0.3), type = "l",lwd = 2,main = maintitle,
+plot(x = c, y = com_result1[1,],ylim = c(0, 0.4), type = "l",lwd = 2,main = maintitle,
      ylab = expression(lambda),xlab = xlab, axes = FALSE)
 axis(side = 1, at = c,labels = c)
 axis(side = 2,las = 1)
@@ -150,12 +150,13 @@ lines(x = c, y = com_result1[3,],lty = 1,col = 3,lwd = 2)
 lines(x = c, y = com_result1[4,],lty = 1,col = 4,lwd = 2)
 lines(x = c, y = com_result1[5,],lty = 1,col = 6,lwd = 2)
 legend("topright",legend = c("Optimal","Grouped","Ungrouped","Linear Pred","Deviance"),lty =1,col = c("black",2:4,6))
+dev.off()
 
 #plot MSE ratio
-length(c)
+png(filename= "sim_1_2_mse.png")
 maintitle <- expression(paste("MSE Ratio"," vs the magnitude of non-zero ", beta, "s", sep = ""))
 xlab <- expression(paste("the magnitude of non-zero ", beta, "s", sep = ""))
-plot(x = c, y = mse_ratio[1,],ylim = c(1, 5), type = "l",lwd = 2,main = maintitle,
+plot(x = c, y = mse_ratio[1,],ylim = c(0, 25), type = "l",lwd = 2,main = maintitle,
      ylab = "MSE Ratio",xlab = xlab, axes = FALSE,col = 2)
 axis(side = 1, at = c,labels = c)
 axis(side = 2,las = 1)
@@ -163,6 +164,6 @@ lines(x = c, y = mse_ratio[2,],lty = 1,col = 3,lwd = 2)
 lines(x = c, y = mse_ratio[3,],lty = 1,col = 4,lwd = 2)
[…]

### 2018-02-27 simulation/scenario_1_4.R

@@ -1,7 +1,7 @@
 # Scenario 1 (plots the lambda selection and MSE ratio)
 # compare four cross validation methods: 
 #  ungrouped, grouped, linear predictor, deviance residuals
-# varying magnitude of betas at low dimension (n = 100, p = 100, nonzero = 10)
+# varying magnitude of betas at higher dimension (n = 100, p = 1000, nonzero = 10)
 
 
 # load source files
@@ -14,17 +14,17 @@ source("modify_fold.R")
 source("deviance.R")
 
 #############################################################################
-# comparison 1 : n = 100, p = 100, nbeta = 10, censoring = 10% , nfold = 10
+# comparison 1 : n = 200, p = 1000, nbeta = 10, censoring = 10% , nfold = 10
 #############################################################################
 # change the magnitude of betas
 n <- 100
-p <- 100
+p <- 1000
 c <-  c(0.5,0.75,1,1.25,1.5)
 beta <- matrix(NA,nrow = length(c), ncol = p)
 for (i in 1:length(c)){
   beta[i,]<- generatebeta(p,nbeta = 10 ,c = c[i])
 }
-N <- 50
+N <- 200
 # initiation
 true.lambda <- matrix(NA,nrow = N ,ncol = length(c))
 gp.lambda <- matrix(NA,nrow = N ,ncol = length(c))
@@ -81,12 +81,11 @@ for(i in 1:N){
 #apply(na.omit(ncv.lambda),2,mean)
 #apply(na.omit(dvr.lambda),2,mean)
 
-#save(true.lambda,gp.lambda,ug.lambda,ncv.lambda,dvr.lambda,
-#     gp.mse,ug.mse,ncv.mse,dvr.mse,true.mse,
-#     file = "sim_1_3.RData")
+save(true.lambda,gp.lambda,ug.lambda,ncv.lambda,dvr.lambda,
+     gp.mse,ug.mse,ncv.mse,dvr.mse,true.mse,
+     file = "sim_1_4.RData")
 
-load(file = "sim_1_3.RData")
-
+#load(file = "sim_1_4.RData")
 # lambda
 com_result1 <- matrix(NA,nrow = 5,ncol = length(c))
 com_result1[1,] <- apply(true.lambda,2,mean)
@@ -138,11 +137,11 @@ colnames(mse_ratio) <- c
[…]

### 2018-02-27 simulation/scenario_2.R

@@ -76,9 +76,9 @@ mean(ug.lambda.1)
 mean(ncv.lambda.1)
 mean(dvr.lambda.1)
 
-save(true.lambda.1,gp.lambda.1,ug.lambda.1,ncv.lambda.1,dvr.lambda.1,
-     true.mse.1,gp.mse.1,ug.mse.1,ncv.mse.1,dvr.mse.1,
-     file = "sim_2_1.RData")
+#save(true.lambda.1,gp.lambda.1,ug.lambda.1,ncv.lambda.1,dvr.lambda.1,
+#     true.mse.1,gp.mse.1,ug.mse.1,ncv.mse.1,dvr.mse.1,
+#     file = "sim_2_1.RData")
 
 #############################################################################
 # scenario 2_2 : n = 100, p = 100, nbeta = 10, censoring = 10% , nfold = 10
@@ -144,9 +144,9 @@ mean(ug.lambda.2)
 mean(ncv.lambda.2)
 mean(dvr.lambda.2)
 
-save(true.lambda.2,gp.lambda.2,ug.lambda.2,ncv.lambda.2,dvr.lambda.2,
-     true.mse.2,gp.mse.2,ug.mse.2,ncv.mse.2,dvr.mse.2,
-     file = "sim_2_2.RData")
+#save(true.lambda.2,gp.lambda.2,ug.lambda.2,ncv.lambda.2,dvr.lambda.2,
+#     true.mse.2,gp.mse.2,ug.mse.2,ncv.mse.2,dvr.mse.2,
+#     file = "sim_2_2.RData")
 
 #############################################################################
 # scenario 2_3 : n = 100, p = 1000, nbeta = 10, censoring = 10% , nfold = 10
@@ -280,6 +280,103 @@ mean(ug.lambda.4)
 mean(ncv.lambda.4)
 mean(dvr.lambda.4)
 
-save(true.lambda.4,gp.lambda.4,ug.lambda.4,ncv.lambda.4,dvr.lambda.4,
-     true.mse.4,gp.mse.4,ug.mse.4,ncv.mse.4,dvr.mse.4,
-     file = "sim_2_4.RData")
+#save(true.lambda.4,gp.lambda.4,ug.lambda.4,ncv.lambda.4,dvr.lambda.4,
+#     true.mse.4,gp.mse.4,ug.mse.4,ncv.mse.4,dvr.mse.4,
+#     file = "sim_2_4.RData")
+
[…]

### 2018-02-27 simulation/scenario_3.R

@@ -30,7 +30,7 @@ n <- 100
 p <- 100
 cpct <- seq(0,0.5,by = 0.1)
 beta <- generatebeta(p,nbeta = 10 ,c = 1)
-N <- 2
+N <- 200
 # initiation
 true.lambda.cen<- matrix(NA,nrow = N ,ncol = length(cpct))
 gp.lambda.cen <- matrix(NA,nrow = N ,ncol = length(cpct))
@@ -221,18 +221,21 @@ fold <- rbind(
 )
 fold <- (200-fold)/200
 
-par(mfrow = c(1,2))
+#par(mfrow = c(1,2))
+png(filename= "sim_3_1.png")
 cpct <-seq(0,0.5,by = 0.1)
 plot(x = cpct, y = cen[1,], type = "l", ylim = c(0.5,1), lwd = 2,
      ylab = "Probability of Defined CVE", xlab = "Censored Portion")
 lines(x = cpct, y = cen[2,], col = 1, lwd = 2)
 lines(x = cpct, y = cen[3,], col = 2, lwd = 2)
 legend("bottomleft",legend = c("Linear Predictor","Grouped","Ungrouped"), col = c(1,1,2), lwd = 2)
+dev.off()
 
+png(filename= "sim_3_2.png")
 nfold <- c(5,10,15,20)
 plot(x = nfold, y = fold[1,], type = "l", ylim = c(0.5,1), lwd = 2,
      ylab = "Convergence %", xlab = "Number of Folds")
 lines(x = nfold, y = fold[2,], col = 1, lwd = 2)
 lines(x = nfold, y = fold[3,], col = 2, lwd = 2)
-
+dev.off()
 

### 2018-03-21 simulation/scenario_1_5.R
add figures and tables for real data analysis

@@ -129,13 +129,13 @@ apply(na.omit(ncv.lambda.10),2,mean)
 apply(na.omit(dvr.lambda.10),2,mean)
 
 
-save(true.lambda,gp.lambda,ncv.lambda,dvr.lambda,
-     gp.lambda.10,ug.lambda.10,ncv.lambda.10,dvr.lambda.10,
-     gp.mse,ncv.mse,dvr.mse,true.mse,
-     gp.mse.10,ug.mse.10,ncv.mse.10,dvr.mse.10,
-     file = "sim_1_5.RData")
+#save(true.lambda,gp.lambda,ncv.lambda,dvr.lambda,
+#     gp.lambda.10,ug.lambda.10,ncv.lambda.10,dvr.lambda.10,
+#     gp.mse,ncv.mse,dvr.mse,true.mse,
+#     gp.mse.10,ug.mse.10,ncv.mse.10,dvr.mse.10,
+#     file = "sim_1_5.RData")
 
-#load(file = "sim_1_5.RData")
+load(file = "sim_1_5.RData")
 # lambda
 com_result1 <- matrix(NA,nrow = 8,ncol = length(c))
 com_result1[1,] <- apply(true.lambda,2,mean)

### 2018-03-21 simulation/scenario_2.R

@@ -152,7 +152,7 @@ mean(dvr.lambda.2)
 # scenario 2_3 : n = 100, p = 1000, nbeta = 10, censoring = 10% , nfold = 10
 #############################################################################
 # change the magnitude of betas
-n <- 200
+n <- 100
 p <- 1000
 #cpct <- seq(0,0.3,by = 0.1)
 cpct <- 0.1

### 2018-03-21 cross-validation.tex
Revert "change figures"
This reverts commit 3e81db9eeadf3261a6054f0af237f4a617f40c44.

--- 2018-02-13 cross-validation.tex
@@
 
-\usepackage[margin=1in]{geometry}
+\usepackage[bmargin=0.75in, tmargin =0.75in,lmargin = 0.5in,rmargin = 0.5in]{geometry}
 
@@
 \section*{Abstract}
-Due to the semi-parametric nature of Cox Regression, conducting cross validation for Cox models has always been a challenge.
-While cross validation is a commonly used approach for selecting tuning parameters in penalized regression, little research has been done to study cross validation methods for penalized Cox regression.
+While cross validation is commonly used for selecting tuning parameters in penalized regression, little research has been done to study cross validation methods for penalized Cox regression.
+Due to its semi-parametric nature, conducting cross validation for Cox models has always been a challenge.
 We propose two new cross-validation methods for Cox Regression, and compare them to traditional information criteria as well as a cross-validated partial likelihood approach originally proposed by Verweij et al.
@@
 
-\begin{figure}
-    \centering
-		\includegraphics[height= 4cm ]{./figures/cv.png}	
-    \caption{K-fold Cross Validation}
-\end{figure}
+%\begin{figure}
+ %   \centering
+	%	\includegraphics[height= 4cm ]{./figures/cv.png}	
+    %\caption{K-fold Cross Validation}
+%\end{figure}
 
@@
 
-\begin{figure}
-    \centering
-		\includegraphics[height= 4cm ]{./figures/02_1.png}
-		\includegraphics[height= 4cm ]{./figures/02_2.png}
-    \caption{Grouped vs Ungrouped Cross Validated Likelihood}
-\end{figure}
-
   \subsection{Cross Validated Linear Predictors}
@@
 
-   \begin{figure}[h]
-    \centering
-    \begin{minipage}[b]{0.45\textwidth}
-      \centering
-		  \includegraphics[height= 3.9cm ]{./figures/03.png}
-      \caption{CV Linear Predictors}
-     \end{minipage}
-     \begin{minipage}[b]{0.45\textwidth}
-      \centering
-		  \includegraphics[height= 4.1cm ]{./figures/04_2.png}
-      \caption{CV Deviance Residuals}
-      \end{minipage}	
-   \end{figure}
+  % \begin{figure}[h]
+    %\centering
+    %\begin{minipage}[b]{0.45\textwidth}
+      %\centering
+	%	  \includegraphics[height= 3.9cm ]{./figures/03.png}
+      %\caption{CV Linear Predictors}
+     %\end{minipage}
+     %\begin{minipage}[b]{0.45\textwidth}
+      %\centering
+%		  \includegraphics[height= 4.1cm ]{./figures/04_2.png}
+    %  \caption{CV Deviance Residuals}
+     % \end{minipage}	
+  % \end{figure}
 
@@
 Hence extra smoothing for the step function would be needed for this approach to work.
+
+\begin{figure}
+    \centering
+		\includegraphics[height= 9cm ]{./figures/figure_1.png}
+		%\includegraphics[height= 4cm ]{./figures/02_2.png}
+    \caption{Illustration of K-fold Cross Validation Methods for Cox Regression}
+\end{figure}
 
@@
     \centering
-    \begin{minipage}[b]{0.4\textwidth}
-      \centering
-		  \includegraphics[height= 6cm ]{./figures/05.jpeg}
-      \caption{ }
-     \end{minipage}
-     \begin{minipage}[b]{0.4\textwidth}
-      \centering
-		  \includegraphics[height= 6cm ]{./figures/06.jpeg}
-      \caption{ }
-      \end{minipage}	
+		\includegraphics[height= 7cm ]{./figures/figure_2.png}
+    \caption{Simulation Studies Comparing 4 CV methods in terms of $\lambda$ and MSE ratio. n = 200, p = 1000, magnitude of non-$\beta$s varies from 0.5 to 1.5}
 \end{figure}
@@
 \begin{table}[h]
+\small
 \centering
-			\begin{tabular}{ c | c | c | c}
-				CV Methods & $\lambda$ (SD) & MSE Ratio \\
-				\hline \hline
-				Optimal  & 0.0606 ( 0.0019 )&  / & \\ 
-				Grouped  &  0.0895  ( 0.0043 )& 1.598  \\
-				Ungrouped & 0.0754 ( 0.0060 )& 1.225 \\
-				Linear Pred & 0.0759 ( 0.0058 )& 1.234 \\
-				M Residual &0.1283 ( 0.0047 ) & 2.959
+\caption{Simulations Comparing all four Cross-Validation methods, Varying n and p}
+\begin{tabular}{llllllllllllllll}
+\hline\\[-0.75em]
+       & \multicolumn{3}{l}{n = 100, p = 20} &  & \multicolumn{3}{l}{n = 100, p = 100} &  & \multicolumn{3}{l}{n = 100, p = 1000} &  & \multicolumn{3}{l}{n = 500, p = 10000} \\
+       & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 20} \\ \cline{2-4} \cline{6-8} \cline{10-12} \cline{14-16}
+\hline\\[-0.75em]
+       &     &        &MSE       &  &   &       & MSE     &    &    &        & MSE      &  &     &        & MSE      \\
+       & $\lambda$(SD)     & &Ratio       &  & 	$\lambda$(SD)         & & Ratio     & &	$\lambda$(SD)         & & Ratio      & &	 $\lambda$(SD)         & & Ratio      
+\\ 
+\cline{2-2} \cline{4-4} \cline{6-6} \cline{8-8} \cline{10-10} \cline{12-12} \cline{14-14} \cline{16-16} 
+\\[-0.75em]
+Optimal &0.016 (0.009)&&/&&0.042 (0.003)&&/&&0.056 (0.006)&&/&&0.026 (0.001)&&/\\
+Grouped &0.035 (0.008)&&2.643&&0.074 (0.009)&&2.236&&0.177 (0.054)&&2.278&&0.053 (0.003)&&77.541\\
+Ungrouped &0.030 (0.009)&&2.404&&0.064 (0.010)&&1.801&&0.115 (0.044)&&1.549&&0.041 (0.003)&&56.105\\
+LinearPred &0.033 (0.006)&&2.423&&0.066 (0.009)&&1.850&&0.122 (0.039)&&1.673&&0.041 (0.003)&&59.693\\
+DevResid &0.099 (0.009)&&13.941&&0.108 (0.009)&&4.110&&0.175 (0.052)&&2.273&&0.075 (0.003)&&76.882\\
+\hline
 \end{tabular}
-			\caption{Simulation Results in High Dimensional Setting}
 \end{table}
@@
     \centering
-		\includegraphics[height= 8cm ]{./figures/stability2.png}
+		\includegraphics[height= 7cm ]{./figures/figure_3.png}
     \caption{Comparing Stability}
@@
 
-    \subsection {Comparisons with Information Criteria}
+    \subsection {Results of Leave-One-Out Cross Validation}
 	%need to change the figures to have fewer lines
@@
 
-   \begin{figure}[h]
+\begin{figure}
     \centering
-    \begin{minipage}[b]{0.4\textwidth}
-      \centering
-		  \includegraphics[height= 6cm ]{./figures/aic_01.jpeg}
-      \caption{Comparison to Information Criteria in Lower Dimension}
-     \end{minipage}
-     \begin{minipage}[b]{0.4\textwidth}
-      \centering
-		  \includegraphics[height= 6cm ]{./figures/aic_02.jpeg}
-      \caption{Comparison to Information Criteria in Higher Dimension}
-      \end{minipage}	
+		\includegraphics[height= 8cm ]{./figures/figure_4.png}
+    \caption{LOOCV}
 \end{figure}
 
- \subsection {Issues with Deviance Residual}
+  \subsection {Simulations on Deviance Residuals}
+
+\begin{table}[h]
+	\small
+	\caption{Average of Selected $\lambda$ with Different Baseline Estimation for Deviance Residuals}
+	\centering
+	\begin{tabular}{lllllllll}
+	\hline
+	\\[-0.75em]
+    & Optimal & Linear Pred & KP & KP & Breslow & Breslow & Exact & Exp\\ 
+    &  & & & weighted & & weighted & & \\ \cline{2 - 9} 
+\textbf{True Baseline: $Exponential(1)$} &&&&&&&&\\
+n = 100, p = 10, non-zero $\beta$: 2 & 0.0461 & 0.0665 & 0.1986 & 0.1970 & 0.1908 & 0.1975 & 0.0540 & 0.0539 \\
+n = 100, p = 100, non-zero $\beta$: 20 &0.0555 & 0.1841 & 0.2359 & 0.2201 & 0.2146 & 0.2244 & 0.0805 & 0.0985 \\
+n = 100, p = 1000, non-zero $\beta$: 20 & 0.0268 & 0.0505 & 0.0948 & 0.0798 & 0.0867 & 0.0809 & 0.0288 & 0.0270 \\
+\textbf{True Baseline:} $Weibull$ &&&&&&&& \\
+ n = 100, p = 10, non-zero $\beta$: 2 & 0.0476 & 0.0710 & 0.2004 & 0.1969 & 0.1907 & 0.1907 & N/A & 0.2330 \\ 
+\hline
+\end{tabular}
+
+\end{table}
 
@@
 
-\begin{table}[h]
+\begin{table}[]
 \centering
-  \begin{tabular}{ c | c | c | c}
-				CV Methods & $\lambda$ & $log(\lambda)$ &AUC \\
-				\hline \hline
-				Grouped  &  0.117 & -2.14 &   0.595\\
-				Ungrouped &  0.112 & -2.20 &  0.597\\
-				Linear Pred &  0.085 & -2.46 &     0.607\\
-				M Residual & 0.171 & -1.76 & 0.529 
+\caption{Application to Real Data Sets}
+\label{my-label}
+\begin{tabular}{llllll}
+\hline
+\\[-0.75em]
+ & Ovarian & & & Shedden&  \\
+& $\lambda$ & log($\lambda$) &  &$\lambda$  &  log($\lambda$)   \\ \cline{2-3} \cline{5-6} 
+Grouped &0.1226 & -2.0985 && 0.1215 & -2.1079\\
+Ungrouped & 0.0803 & -2.5222 && 0.0983 & -2.3197\\
+Linear Pred &0.0828 & -2.4919 && 0.1013 & -2.2894\\
+Dev Resid & 0.0992 & -2.3104 && 0.1110 & -2.1987\\ \hline
 \end{tabular}
-  \caption{Tuning Parameter Selected for Ovarian Cancer Data}
 \end{table}
+
+\begin{figure}
+    \centering
+		\includegraphics[height= 8cm ]{./figures/figure_5.png}
+    \caption{Comparing Cross-Validated Error for Different Methods on Real Data Sets}
+\end{figure}
 
@@
 
-   \begin{figure}[h]
-   \centering
-    \begin{minipage}[b]{0.4\textwidth}
-      \centering
-		  \includegraphics[height= 6cm ]{./figures/07.jpeg}
-      \caption{CVE for Ovarian Cancer Data}
-     \end{minipage}
-     \begin{minipage}[b]{0.4\textwidth}
-      \centering
-		  \includegraphics[height= 6cm ]{./figures/08.jpeg}
-      \caption{CVE for Lung Cancer Data}
-      \end{minipage}	
-   \end{figure}
-
 In the lung cancer data set that we applied our methods to, there are 442 patients and 236 events.
@@
 The results of the analysis are listed in Table 3.
-
-			\begin{table}[h]
-			\centering
-			\begin{tabular}{ c | c | c | c}
-				CV Methods & $\lambda$ & $log(\lambda)$ & AUC \\
-				\hline \hline
-				Grouped  &   0.119 & -2.12 &  0.608 \\
-				Ungrouped &  0.119 & -2.12 &  0.608 \\
-				Linear Pred &  0.090 & -2.41 &  0.633\\
-				M Residual & 0.203 & -1.59 &  0.564
-			\end{tabular}
-			\caption{Tuning Parameter Selected for Lung Cancer Data}
-			\end{table}
 

### 2018-03-21 simulation/scenario_1_4.R

@@ -81,11 +81,11 @@ for(i in 1:N){
 #apply(na.omit(ncv.lambda),2,mean)
 #apply(na.omit(dvr.lambda),2,mean)
 
-save(true.lambda,gp.lambda,ug.lambda,ncv.lambda,dvr.lambda,
-     gp.mse,ug.mse,ncv.mse,dvr.mse,true.mse,
-     file = "sim_1_4.RData")
+#save(true.lambda,gp.lambda,ug.lambda,ncv.lambda,dvr.lambda,
+#     gp.mse,ug.mse,ncv.mse,dvr.mse,true.mse,
+#     file = "sim_1_4.RData")
 
-#load(file = "sim_1_4.RData")
+load(file = "sim_1_4.RData")
 # lambda
 com_result1 <- matrix(NA,nrow = 5,ncol = length(c))
 com_result1[1,] <- apply(true.lambda,2,mean)
@@ -140,7 +140,7 @@ rownames(mse_ratio) <- c("grouped","ungrouped","ncvreg","deviance")
 png(filename= "sim_1_4_lambda.png")
 maintitle <- expression(paste(lambda," vs the magnitude of non-zero ", beta, "s", sep = ""))
 xlab <- expression(paste("the magnitude of non-zero ", beta, "s", sep = ""))
-plot(x = c, y = com_result1[1,],ylim = c(0, 0.40), type = "l",lwd = 2,main = maintitle,
+plot(x = c, y = com_result1[1,],ylim = c(0, 0.30), type = "l",lwd = 2,main = maintitle,
      ylab = expression(lambda),xlab = xlab, axes = FALSE,cex = 2)
 axis(side = 1, at = c,labels = c)
 axis(side = 2,las = 1)

### 2018-04-04 cross-validation.tex
Complete draft, with Updated Figures

@@
 \section*{Abstract}
-While cross validation is commonly used for selecting tuning parameters in penalized regression, little research has been done to study cross validation methods for penalized Cox regression.
+While cross validation is commonly used for selecting tuning parameters in penalized regression, little research has been conducted to study cross validation methods for penalized Cox regression.
 Due to its semi-parametric nature, conducting cross validation for Cox models has always been a challenge.
@@
 However, our proposed approach of cross-validating the linear predictors generally offers the best balance of stability and performance.
-We also illustrate these approaches on data from studies of gene expression and progression-free survival in cancer patients.
+We also illustrate these approaches on data from studies of gene expression and overall survival in cancer patients.
 
@@
 In this paper, only the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$ was considered.
-The penalized cox regression can be used for analyzing high dimensional data, where the number of covariates p $\gg$ n.
-A common application would be to predict the overall survival time of a particular cancer by certain gene expressions.
+The LASSO cox regression can be used to build prediction models for survival outcome based on high dimensional data, where the number of covariates p $\gg$ n.
+An example of its application would be to make predictions of cancer patients' the overall survival time with their certain gene expressions.
 
 Selecting the tuning parameter $\lambda$ is crutial to getting LASSO estimations for $\beta$.
+The LASSO penality enforces sparcity to the regression model, so that it carries out an automatic variable selection along the solution path.
 As $\lambda$ increases, more coefficients would be set to 0 and fewer variables will be selected by the model.
@@
 When $\lambda$ is too large, the model would be too conservative to select any variables.
+In either case, model's predictive accuray would be affected.
 
-One common approach for selecting the tuning parameter $\lambda$ for linear regression and logistic regression is via cross validation.
-The idea of K - fold cross validation is illustrated in Figure 1.
+One common approach for selecting the tuning parameter $\lambda$ for linear regression and logistic regression is via the K-fold cross validation.
 The data set would be split into K folds.
@@
 %ZZZZZZZZZZZZZ
-\par The most intuitive approach to conduct cross validation in Cox regression is i) fit the model to training set then ii) calculate cross-validated partial likelihood based on the observations from the test set as a measure for the model's predictive accuracy.
-As will be discussed in Section 3, this approach is unstable when number of events gets small.
-Verweij elt.
-\citep{Verweij1993} proposed a stablized version of cross-validated log likehood, which has been widely used and implemented in R packages such as $\tt{glmnet}$.
-In this paper, we proposed two alternative ways to carry out cross validation.
+\par Currently, there are two common cross-validation approaches for Cox regression, both having its disadvantages.
+The most intuitive approach to conduct cross validation in Cox regression is i) fit the model to training set then ii) calculate cross-validated partial likelihood based on the observations from the test set as a measure for the model's predictive accuracy.
+As will be discussed in Section 2, this approach is unstable when number of events gets small.
+The second approach was propsoed by Verweij et al.
+\citep{Verweij1993}.
+This is a stablized version of cross-validated log likehood, which has been widely used and implemented in R packages such as $\tt{glmnet}$ \citep{glmnet}.
+Although this approach fixes the stability issue, it tends to yield more conservative models in practice.
+We hope to improve cross-validation procedure for Cox regression to achieve a good balance between stability and performance.
+
+\par In this paper, we proposed two alternative ways to carry out cross validation.
 Instead of cross-validating over partial likelihood, we propose to cross validate over the linear predictors of the regression model and cross validate over the deviance residual \citep{Therneau1990}of the Cox model.
-
-\par We conducted simulation studies to compare how those methods perform in both low dimensional and high dimensional settings with LASSO penalty.
-We compared the two proposed methods with the cross-validated partial likelihood approaches and traditional information criteria.
+We conducted simulation studies to compare how those methods perform in both low dimensional and high dimensional settings with LASSO penalty.
+We compared the two proposed methods with the existing cross-validated partial likelihood approaches.
 We showed that the linear predictor approach outperforms other cross validation approaches in various scenarios.
 We found out that using cross validation to select tuning parameters for LASSO penalized cox regression tends to be a conservative approach in general.
-Finally, we apply those methods to two high dimensional data sets with time-to-event outcomes.
+Finally, we apply those methods to two high dimensional data sets from real studies with time-to-event outcomes.
 
@@
 
-\par An alternative way to calculate cross-validated log likelihood was proposed by Verweij elt.
+\par An alternative way to calculate cross-validated log likelihood was proposed by Verweij et al.
 They defined cross-validated log likelihood of leaving the ith fold out to be \begin{equation}cvl = \sum_{i=1}^{K} \{l(\hat{\beta}_{-i}) - l_{-i}(\hat{\beta}_{-i})\}.
@@
 \par We refer to the first definition as ungrouped cross-validated partial likelihood (CVL) and the second definition as grouped cross-validated partial likelihood.
-Figure 2 illustrates the ideas of the two methods.
+The upper panels in Figure 2 illustrates the ideas of the two methods.
 The grouped CVL uses observations more efficiently than the ungrouped CVL.
-Since only the left-out fold would be used, the ungrouped cvl is unstable in practice when there are only very few observations in some folder.
+Since only the left-out fold would be used, the ungrouped CVL is unstable in practice when there are only very few observations in some folder.
 
@@
 Besides cross validating over the partial likelihood, an alternative approach is to cross validate over the linear predictors.
-The data would still be split into K folds.
-Suppose ith fold is left out and the other K-1 folds are used as the training set to get the estimates $\hat{\beta}_{-i}$.
-Then the cross-validated linear predictors would be calculated based on the observations in the test set:  \begin{equation}\hat{\eta}_{-i} = X_{i}\hat{\beta}_{-i}.\end{equation} After repeating this for all K folds, a whole set of linear predictors  $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...
+In the first step, the ith fold would be left out and the other K-1 folds would be used as the training set to get the estimates $\hat{\beta}_{-i}$.
+Then, the cross-validated linear predictors would be calculated based on the observations in the test set:  \begin{equation}\hat{\eta}_{-i} = X_{i}\hat{\beta}_{-i}.\end{equation} After repeating this for all K folds, a whole set of linear predictors  $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...
 \hat{\eta}_{-K})$ can be obtained.
@@
 
-Cross validated error is defined as the log of $L(\hat{\eta}_{-})$.
-This idea of cross validating over linear predictors is implemented in the package ncvreg.
+Cross-validated error is defined as the log of $L(\hat{\eta}_{-})$.
+This idea of cross validating over linear predictors is implemented in the package \texttt{ncvreg} \citep{ncvreg}.
 For all the approaches introduced so far, they are all equivalent to each other in linear regressions.
@@
   %Need to update this section with the new baseline estimation method%
-Since the biggest challenge of conducting cross validation for cox regression is due to the missingness of baseline hazard, we propose a second approach which involves estimating the actual baseline hazard $\hat{\Lambda}_{0}$.
+The fundamental challenge of conducting cross validation for cox regression is that the baseline hazard is not estimated from the model.
+Hence we proposed our second approach which involves estimating the actual baseline hazard $\hat{\Lambda}_{0}$.
 The sum of squares of Martingale residuals is a natural candicate for evaluating the model performance using baseline hazard.
@@
 The deviance residual can be calcualted from the Martingale Residuals: \begin{equation} d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}\end{equation} The sum of squares of the deviance residuals, $\sum_{i}\hat{d}_{i}^2$, would be used as the cross validated error.
+An illustration of this method is drawn in the bottom right panel of Figure 1.
 
@@
   Simulation studies were conducted to compare how those methods behave relative to each other.
-We generate data with pre-specified baseline hazard, covariate matrix X, coefficient $\beta$ and censoring mechanism.
+We generated data with pre-specified baseline hazard, covariate matrix X, coefficient $\beta$ and censoring mechanism.
 Both low dimensional and high dimentional scenarios were examined.
 All cross-validation methods mentioned in Section 2 were applied to the data to select the tuning parameter $\lambda$ and produce $\hat{\beta}$ estimates.
-Cross-validation were also compared to model selection criteria AIC and BIC.
+%Cross-validation were also compared to model selection criteria AIC and BIC.
 
+ The entries in the covariate matrix $X_{n \times p}$ were independently generated from $Normal(0, 1)$.
+True coefficients $\beta$ were assumed to have sparsity: $\beta_{p\times 1} = (\beta_{1},\beta_{2}, ..., \beta_{j}, 0, ..., 0)^{T}$.
+The covariate matrix $X_{n \times p}$ was centered before it was used to generate outcomes.
 Survival times were generated from exponential distribution $h(t) = h_{0} exp(X\beta)$, conditioned on covariates.
-The entries in the covariate matrix $X_{n \times p}$ were independently generated from Normal (0, 1).
-True coefficients $\beta$  were assumed to have sparsity: $\beta_{p\times 1} = (\beta_{1},\beta_{2}, ..., \beta_{10}, 0, ..., 0)^{T}$.
 Censoring status were generated based on binomial distribution.
+All simulations were implemented in R \citep{R}.
 
   Suppose $\hat{\beta}$ is the coefficients estimated by a fitted cox model.
-We measure the distance between the fitted model and the true model by mean squared error $MSE = E(\hat{\beta} - \beta)^2$.
+We quantify the predictive accuracy by measuring the distance between the fitted model and the true model by mean squared error MSE $= E(\hat{\beta} - \beta)^2$.
 For each generated data set, the $\lambda$ that has the minimal MSE is chosen as the optimal $\lambda$.
@@
   %need to change figures to include the new cross-validated deviance residuals
-   The simulation experiments were first conducted under lower dimension settings.
-Number of observations n is set to be 100.
-Dimension of the data p is set to be 100.
-200 replications were used.
-Censoring percentage, number of folds used in cross-validation, number of non-zero $\beta$s and the magnitude of the $\beta$s were varied in several different simulation scenarios.
-Patterns and results shown in various simulation scenarios are consistent with one another.
-Figure 5 and Figure 6 illustrates one simulation scenario, where about 10$\%$ of which were set to be censored and 10 of the $\beta$s were set to be non-zero and 10 folds were used for cross validation.
-The x - axis of the plots are the magnitude of the non-zero $\beta$ coefficients.
+   The simulation experiments were first conducted when the signal in the data was varied.
+In each simulated data set, there are n = 100 observations and p = 1000 covariates.
+10$\%$ of the observations were censored.
+10-fold cross-validation was implemented for all four methods.
+Number of non-zero $\beta$s was 20.
+The signal in the data set was varied by varing the magnitude of the $\beta$s from 0.50 to 1.50.
+For each scenario, 200 replications were used.
 
@@
 		\includegraphics[height= 7cm ]{./figures/figure_2.png}
-    \caption{Simulation Studies Comparing 4 CV methods in terms of $\lambda$ and MSE ratio. n = 200, p = 1000, magnitude of non-$\beta$s varies from 0.5 to 1.5}
+    \caption{Simulation Studies Comparing 4 CV methods in terms of $\lambda$ and MSE ratio. n = 200, p = 1000, 10$\%$ censoring.}
 \end{figure}
 
-    $\lambda$s chosen by different cross-validation methods are compared in Figure 5.
-All values are from taking the mean of 200 replications.
-The black line represents the optimal $\lambda$.
-All four cross validation methods consistently select $\lambda$s larger than the optimal one.
-Since larger $\lambda$ leads larger penalties and fewer variables are selected, all four cross validation methods seem to be conservative in terms of selecting variables.
-Ungrouped cross-validated log likelihood and the cross-validated linear predictor seem to be more liberal than the other two methods.
-The cross-validated Deviance residual is the most conservative one and its performance is quite off from the other three methods.
+\par Results of this simulation were illustrated in Figure 2.
+The x axis of the plots are the magnitude of the non-zero $\beta$ coefficients.
+The left panel compares the $\lambda$ selected by different methods.
+The grouped cross-validated partial likelihood approach, which is currently most widely used approach, consistently selects $\lambda$s that are greater than the optimal one.
+Greater $\lambda$ leads to greater penalties and fewer non-zero variables, hence the grouped approach is conservative in terms of selecting variables.
+While the proposed cross-validated Deviance residual has very close performance as the grouped approach, the ungrouped cross-validated log likelihood and the cross-validated linear predictor were more liberal than the other two methods.
+In general, all four methods tends to act conservatively and selects smaller models.
 
-   Mean squared errors are compared in Figure 6.
+\par The right panel compares the MSE ratios of the model selected.
 The MSEs of the cross-validation selected models are compared with the MSE of the model given by the optimal $\lambda$.
 The y-axis represents the ratio of the two.
-The ungrouped method and the linear predictor approach are outperforming the other methods and are closest to the minimal MSE.
-
-   Simulation studies were also conducted in high dimension scenarios, where the number of observations and dimension of the covariate matrix is more similar to real genetics research.
-The results are shown in Table 1.
-Conclusions are close to what is shown in the low dimensional setting.
-The M- residual approach performs most conservatively.
-Linear predictor approach and ungrouped approach are relatively more liberal than the other approaches but still slightly more conservative than the optimal $\lambda$.
+The ungrouped method has the smallest MSE ratio.
+The linear predictor approach's performance is also very close the ungrouped appraoch.
+Both the grouped approach and the proposed deviance residual approach yield models that have greater MSE, hence are not recommended for use in practice.
 
@@
 \hline\\[-0.75em]
-       & \multicolumn{3}{l}{n = 100, p = 20} &  & \multicolumn{3}{l}{n = 100, p = 100} &  & \multicolumn{3}{l}{n = 100, p = 1000} &  & \multicolumn{3}{l}{n = 500, p = 10000} \\
-       & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 20} \\ \cline{2-4} \cline{6-8} \cline{10-12} \cline{14-16}
+       & \multicolumn{3}{l}{n = 100, p = 20} &  & \multicolumn{3}{l}{n = 100, p = 100} &  & \multicolumn{3}{l}{n = 500, p = 10000} &  & \multicolumn{3}{l}{n = 250, p = 10000} \\
+       & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 20} &  & \multicolumn{3}{l}{$\#$of non-zero = 20} \\ \cline{2-4} \cline{6-8} \cline{10-12} \cline{14-16}
 \hline\\[-0.75em]
        &     &        &MSE       &  &   &       & MSE     &    &    &        & MSE      &  &     &        & MSE      \\
-       & $\lambda$(SD)     & &Ratio       &  & 	$\lambda$(SD)         & & Ratio     & &	$\lambda$(SD)         & & Ratio      & &	 $\lambda$(SD)         & & Ratio      
+       & $\lambda$ (SD)     & &Ratio       &  & 	$\lambda$ (SD)         & & Ratio     & &	$\lambda$ (SD)         & & Ratio      & &	 $\lambda$ (SD)         & & Ratio      
 \\ 
@@
 \\[-0.75em]
-Optimal &0.016 (0.009)&&/&&0.042 (0.003)&&/&&0.056 (0.006)&&/&&0.026 (0.001)&&/\\
-Grouped &0.035 (0.008)&&2.643&&0.074 (0.009)&&2.236&&0.177 (0.054)&&2.278&&0.053 (0.003)&&77.541\\
-Ungrouped &0.030 (0.009)&&2.404&&0.064 (0.010)&&1.801&&0.115 (0.044)&&1.549&&0.041 (0.003)&&56.105\\
-LinearPred &0.033 (0.006)&&2.423&&0.066 (0.009)&&1.850&&0.122 (0.039)&&1.673&&0.041 (0.003)&&59.693\\
-DevResid &0.099 (0.009)&&13.941&&0.108 (0.009)&&4.110&&0.175 (0.052)&&2.273&&0.075 (0.003)&&76.882\\
+Optimal &0.016 (0.009)&&/&&0.042 (0.003)&&/&&0.026 (0.001)&&/&&0.026 (0.001)&&/\\          
+Grouped &0.035 (0.007)&&2.597&&0.073 (0.009)&&2.184&&0.053 (0.003)&&5.627&&0.123 (0.036)&&3.931\\
+Ungrouped &0.030 (0.009)&&2.246&&0.064 (0.01)&&1.815&&0.040 (0.003)&&3.131&&0.070 (0.016)&&2.560\\
+LinearPred &0.034 (0.007)&&2.559&&0.066 (0.008)&&1.880&&0.041 (0.003)&&3.350&&0.074 (0.016)&&2.713\\
+DevResid &0.100 (0.009)&&14.005&&0.108 (0.008)&&4.087&&0.076 (0.003)&&9.133&&0.111 (0.028)&&3.767\\ 
 \hline
@@
 
+\par Simulation studies were also conducted in scenarios with various sample sizes and dimensions.
+The results are shown in Table 1.
+While the ungrouped approach yields the smallest MSE and best predictive accuracy, the Linear Predictor and the Ungrouped approaches perform closely to each other across all settings.
+Grouped approach performs well in low-dimensions or with sufficient samples, while deviance residual approach performs better in high-dimensional settings.
+
  \subsection {Stability}
-   While the ungrouped cross-validated likelihood is an intuitive way to carry out cross-validation, it is quite unstable when it runs into cases where the number of observed events in some fold is really small.
-As is illustrated in Figure 5, cases of undefined cross-validation error would increase when censoring increases or when number of folds increases.
 
-\begin{figure}
+\par In this section, simulations studies were conducted to address the stability issue of the ungrouped cross-validated likelihood.
+The ungrouped approach is an intuitive way to carry out cross-validation.
+It also has the best predictive accuarcy based on the simulation results in previous sections.
+However, one big disadvantage of this approach is that it is unstable when number of observed events in some fold is really small.
+As is illustrated in Figure 3, when censoring increases or when number of folds increases, the model would fail to converge due to insufficient number of events in the test set.
+Hence, the ungrouped cross-validated likelihood should not be used when there is no sufficient number of events.
+
+\begin{figure}[h]
     \centering
@@
     \subsection {Results of Leave-One-Out Cross Validation}
-	%need to change the figures to have fewer lines
-    \par We also compared cross-validation with information criteria AIC and BIC.
+	\par In this section, Leave-One-Out Cross Validation (LOOCV) were studied, in comparison with 10-fold CV.
+When first introduced, the grouped cross-validated partial likelihood was proposed so that LOOCV for Cox regression became feasible.
+It is impossible to carry out LOO cross-validation using the ungrouped approach, because the partial likelihood cannot be built on the one left-out observation.
+While the grouped cross-validated partial likelihood solves the stability issue by aggregating more observations together, it is not an efficient way to evaluate the model fit.
+Simulation studies in Section 3.1 have illustrated that it tends to be conservative in variable selections and selects smaller models in various scenarios.
+This fact can be even more illustrated via the following simulation studies, where both 10-fold and LOOCV were conducted via grouped, linear predictor and deviance residual approaches and 10 fold CV was conducted for the ungrouped approach.
+	\par As is illustrated in Figure 4, LOOCV selected models that have smaller MSE than 10-fold CV for all three CV approaches.
+As is shown in left and middle panels, LOOCV based on deviance residuals and grouped cross-validated partial likelihood still perform more conservatively than the ungrouped approach.
+Even if the grouped approach was proposed to allow LOOCV, the LOOCV based on the grouped approach does not surpass the 10-fold ungrouped approach.
+On the other hand, the Linear Predictors also allows LOOCV, and the LOOCV based on Linear Predictors performs better than the ungrouped approach.
+
+    %\par We also compared cross-validation with information criteria AIC and BIC.
 Simulation studies were conducted in lower dimension and higher dimension respectively.
@@
 
-\begin{figure}
+\begin{figure}[h]
     \centering
@@
 
-  \subsection {Simulations on Deviance Residuals}
+  \subsection {Simulations on Deviance Residuals and Baseline Hazard Estimations}
+\par While the Deviance Residual approach has a reasonable performance in scenarios where dimension is high, it tends to act more conservatively than all the other three approaches in general.
+As is noted by Therneau et al, the sums of squared martingale residuals and deviance residuals do not necessary reflect how good the model fit is \citep{Therneau2000modeling}.
+The difficulty of estimating the baseline hazard seems to play an important role.
+A series of simulations were thus designed to study how baseline hazard estimation affects the deviance residuals' performance as a measurement for model fit.
+
+\par In the following simulation studies, the defined optimal $\lambda$ and the $\lambda$ selected by the Linear Predictor approach were used as benchmarks.
+Various sample sizes and dimensions were varied.
+The survival outcomes were generated both from exponential distribution and Weibull distribution.
+Six different approaches were compared in terms of baseline hazard estimation and their impact on sums of squared deviance residuals.
+Two non-parametric approaches, Kalbfleisch and Prentices' estimator and Breslow's estimator, were used first for estimating the cumulative hazard baseline \citep{Kalbfleisch2011} \citep{breslow1972}.
+As is presented in column 3 and column 5 in Table 2, the sum of squared deviance residuals computed based on both approaches are all very conservative compared to the optimal $\lambda$.
+Their performance is extremely bad when n is large and p is small.
+
+\par Since both of the non-parameteric estimations seemed to perform well at early time points but poorly later, a weighted approach was then applied to emphasize earlier time points.
+Weights are assigned proportional to number of subjects at risk, making earlier time points weighted more than later time points.
+Results of these two approahces are presented in column 4 and 6 in Table 2.
+The attempt to improve the non-parameteric estimations via weighting does not seem to improve the selection too much.
+
+\par In the next step,  the non-parametric baseline hazard estimation was replaced by the exact baseline hazard that was used to generate the outcome.
+Results were listed in the seventh column in Table 2.
+When the baseline was known, the performance of sum of squared deviance residuals actually performed really close to the optimal $\lambda$.
+This simulation suggests that when the baseline hazard was accurately estimated, then the sum of squared deviance residuals gave really good measures for the model fit.
+
+\par In the last step, we replaced the non-parametric baseline hazard estimation with a parametric approach, where we assume the baseline hazard to follow an exponential distribution.
+We derived the estimation for baseline hazard by maximum likelihood approach conditioned on covariates: $\hat{\lambda}_{0} = \frac{\sum d_{i}}{\sum t_{i}e^{\eta_{i}}}$.
+The results are listed in the eighth column in Table 2.
+When this estimation was used for data that were generated from exponential distribution, its performance was as well as when the exact baseline was specified.
+When it was fitted to data that were generated from the Weibull distribution, that is when the model is misspecified, then its performance was even worse than the non-parametric approaches.
+
+\par Hence, getting accurate estimation of the baseline hazard is crucial to whether or not the sum of squared deviance residuals can be a good measurement for model fit.
 
@@
 \end{tabular}
-
 \end{table}
@@
 
-We also compared those cross validation methods when they are applied to real data sets.
-Unlike simulated set, we do not know the optimal $\lambda$ in the real data sets, but we can still see how those methods perform relative to each other.
-The first data set is a study on ovariance cancer from The Cancer Genome Atlas.
-The second data set is a study on lung cancer (Shedden et al 2004).
-The outcome of both data sets are time-to-event data.
+\par In this section, we compared the performances of those cross validation methods when they are applied to data from real studies.
+The first data set is collected from a study on ovariance cancer from The Cancer Genome Atlas.
+The second dataset is a study on lung cancer \citep{shedden2008gene}.
+The outcomes of both data sets are patients' overall survival time.
 
-There are 460 patients and 236 events for the ovarian cancer data set.
+\par In the ovarian cancer data set, there are 460 patients and 236 events .
 About 40$\%$ of the observations are censored.
-The covariate matrix records whether or not there is mutation at the location and has dimension 12376.
-When we fit the penalized cox model, we first adjusted three clinical variables which is known to have large impact on survival.
-We first fitted these three covariates into regular cox regression and keep the linear predictors.
-The selection for the penalty term does not affect the estimation of those clinical variables.
-The results are illustrated in Table 2.
-The linear predictor approach has the best predictive accuracy according to AUC.
+The covariate matrix records whether or not there is mutation at certain gene location and has dimension of 12376.
+Three clinical variables, age, gender and residuals, which were known to have large impact on survival were adjusted in the analysis.
+That is to say, the clinical covariates were first fitted into regular cox regression and only linear predictors were kept so that the selection for the penalty term would not affect the estimation of those clinical variables.
+Cross-validation was used for the selection of the penalty term, which affects the selection of gene mutations.
+The results are illustrated in Table 3.
 
-\begin{table}[]
+\par In the lung cancer data set , researchers collected gene expression data for 22283 genes.
+236 death events were observed for 442 patients.
+About 50$\%$ of the patients are censored.
+Again clinical variables, age, gender, status of getting adjuvant chemotherapy were adjusted before the LASSO cox regression was fitted.
+Cross-validation was again used for the selection of the penalty term, which selects the gene expressions that are associated with patients' overall survival.
+The results of the analysis are also listed in Table 3.
+
+\begin{table}[h]
 \centering
@@
 
-In Figure 9, the Cross Validated Error is rescaled and plotted for all four methods.
+\par How the CV methods perform on real data reflects what is shown in simulation studies.
+The ungrouped CVL is the most liberal approach.
+Deviance Residual and Grouped CVL approaches were both really conversative, while the ungrouped approach was the second liberal to ungrouped CVL.
+In Figure 5, the Cross Validated Error is rescaled and plotted for all four methods.
 A $\lambda$ will be selected when CVE curve reaches its lower point.
-The blue line, which represents the linear predictor approach, it has more curvature near its lowest point.
+The blue line, which represents the linear predictor approach, has more curvature near its lowest point.
 It is easier to pick out the minimum point for this blue curve.
 Hence this approach is better at picking out signals than the other three approaches.
-If we look at the curve for the cross-validated Deviance Residual, there's almost no signal there.
-
-In the lung cancer data set that we applied our methods to, there are 442 patients and 236 events.
-About 50$\%$ of the patients are censored.
-Again we adjusted three clinical variable before we fit the penalized regression.
-The results of the analysis are listed in Table 3.
-
-If we look at the CVE curve for all four of those methods, we still see that blue curve has more curvature near its lowest point.
-But in general, there are more signal in this lung cancer data set compared to the ovarian cancer data set.
+For the grouped cross-validated partial likelihood, there's almost no signal there.
 
 \section{Discussion}
-	\begin{itemize}
-		\item 	
-		\item
-	\end{itemize}
+Four methods of conducting cross-validation for Cox Regression were compared through simulation studies in terms of their predictive accuracy.
+All cross-validation methods tend to select the more conservative model: with greater value of $\lambda$, and fewer variables selected into the model.
+Despite of the stability issue, the ungrouped approach is the most liberal one across all scenarios and yield the least MSE.
+The proposed linear predictor approach performs closely to the ungrouped approach.
+Under LOOCV, the linear predictor approach out-performs the ungrouped approach.
+Both the grouped approach and the proposed Deviance Residual approach perform conservatively, hence not recommended for use in practice.
+Results from real data application reflects results from the simulation studies.
 

### 2018-05-03 cross-validation.tex
Splitting tex into two files; deleting Rnw

--- 2018-02-27 cross-validation.tex
@@
-\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
-\let\hlipl\hlkwb
-
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
-
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+\documentclass{article}
+\usepackage{graphicx}
+\usepackage{color}
 
@@
 \section*{Abstract}
-Due to the semi-parametric nature of Cox Regression, conducting cross validation for Cox models has always been a challenge.
-While cross validation is a commonly used approach for selecting tuning parameters in penalized regression, little research has been done to study cross validation methods for penalized Cox regression.
+While cross validation is commonly used for selecting tuning parameters in penalized regression, little research has been conducted to study cross validation methods for penalized Cox regression.
+Due to its semi-parametric nature, conducting cross validation for Cox models has always been a challenge.
 We propose two new cross-validation methods for Cox Regression, and compare them to traditional information criteria as well as a cross-validated partial likelihood approach originally proposed by Verweij et al.
@@
 However, our proposed approach of cross-validating the linear predictors generally offers the best balance of stability and performance.
-We also illustrate these approaches on data from studies of gene expression and progression-free survival in cancer patients.
+We also illustrate these approaches on data from studies of gene expression and overall survival in cancer patients.
 
-\section{Introduction}
-Cox propotional hazard regression is one of the most commonly used statistical model for analyzing data with survival outcomes.
-Despite its popularity, Cox model is limited by its semiparametric nature: it is challenging to evaluate the model's predictive accuracy.
-
-In the Cox model \citep{Cox1975}, the hazard function for an observation i is given by \begin{equation}h_{i}(t) = h_{0}(t) \text{exp}( X_{i}^{T} \beta)\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{i}^{T} \beta}$ is the relative risk.
-Suppose there are n observations and m observed failure times, the estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood \begin{equation}L(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation}, where j indexes the observed failure times and $R(t_{j})$ denotes the set of observations at risk at time $t_{j}$.
-As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard.
-Hence the interpretation of the Cox model is only valid in a relative sense.
-
-Penalized Cox Regression is an extension of the regular cox regression model.
-Estimates of the $\beta$ coefficients are obtained by minimizing the objective function \begin{equation}  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{m} \text{log} \left \{\frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta).
-\end{equation} \citep{Simon2011}.
-The first part of the objective function is based on the partial likelihood.
-The second part is a penalty term that depends on some regularization parameters.
-In this paper, only the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$ was considered.
-The penalized cox regression can be used for analyzing high dimensional data, where the number of covariates p $\gg$ n.
-A common application would be to predict the overall survival time of a particular cancer by certain gene expressions.
-
-Selecting the tuning parameter $\lambda$ is crutial to getting LASSO estimations for $\beta$.
-As $\lambda$ increases, more coefficients would be set to 0 and fewer variables will be selected by the model.
-When $\lambda$ is too small, the model tends to be redundant.
-When $\lambda$ is too large, the model would be too conservative to select any variables.
-
-One common approach for selecting the tuning parameter $\lambda$ for linear regression and logistic regression is via cross validation.
-The idea of K - fold cross validation is illustrated in Figure 1.
-The data set would be split into K folds.
-One fold would be treated as the test set and the other K - 1 folds as the training set.
-The model would first be built on the training set, then fitted to the test set to obtain cross validation error (CVE).
-CVE would be calculated for each candidate $\lambda$, then the one that minimizes cross validated error would be selected.
-
-However, for penalized cox regression, it is challenging to carry out cross validation.
-Since the partial likelihood of the cox model does not give estimates for the baseline, it is not meaningful to compare the coefficients if they are not estimated from different models and different observations.
-Only if the baseline is the same, then the coefficients can be interpreted as relative risks.
-This is against the idea of cross validation, which splits the data and builds models on different subsets.
-
-\begin{figure}
-    \centering
-		\includegraphics[height= 4cm ]{./figures/cv.png}	
-    \caption{K-fold Cross Validation}
-\end{figure}
-
-%ZZZZZZZZZZZZZ
-\par The most intuitive approach to conduct cross validation in Cox regression is i) fit the model to training set then ii) calculate cross-validated partial likelihood based on the observations from the test set as a measure for the model's predictive accuracy.
-As will be discussed in Section 3, this approach is unstable when number of events gets small.
-Verweij elt.
-\citep{Verweij1993} proposed a stablized version of cross-validated log likehood, which has been widely used and implemented in R packages such as $\tt{glmnet}$.
-In this paper, we proposed two alternative ways to carry out cross validation.
-Instead of cross-validating over partial likelihood, we propose to cross validate over the linear predictors of the regression model and cross validate over the deviance residual \citep{Therneau1990}of the Cox model.
-
-\par We conducted simulation studies to compare how those methods perform in both low dimensional and high dimensional settings with LASSO penalty.
-We compared the two proposed methods with the cross-validated partial likelihood approaches and traditional information criteria.
-We showed that the linear predictor approach outperforms other cross validation approaches in various scenarios.
-We found out that using cross validation to select tuning parameters for LASSO penalized cox regression tends to be a conservative approach in general.
-Finally, we apply those methods to two high dimensional data sets with time-to-event outcomes.
-
-\section{Methods}
-
-  \subsection{Cross Validated Likelihood} 
-Suppose a data set of n observations is split into K folds and the ith fold is left out, then a partial likelihood can be built over the K -1 folds and yields maximum likelihood estimates of the coefficients, denoted by $\hat{\beta}_{-i}$.
-An intuitive way to carry out cross validation is to use the partial likelihood on the test data as the cross validated error: \begin{equation}cvl = \sum_{i=1}^{K} l_{i}(\hat{\beta}_{-i})\end{equation} This is implemented in the $\tt{glmnet}$ package as the $\tt{ungrouped}$ option.
-
-\par An alternative way to calculate cross-validated log likelihood was proposed by Verweij elt.
-They defined cross-validated log likelihood of leaving the ith fold out to be \begin{equation}cvl = \sum_{i=1}^{K} \{l(\hat{\beta}_{-i}) - l_{-i}(\hat{\beta}_{-i})\}.
-\end{equation} When the ith fold is left out, $l(\hat{\beta}_{-i})$ is the log partial likelihood evaluated at $\hat{\beta}_{-i}$ with all k folds of observations.
-$l_{-i}(\hat{\beta}_{-i})$ is the log partial likelihood evaluated at $\hat{\beta}_{-i}$ with observations from the other K-1 folds.
-
-\par We refer to the first definition as ungrouped cross-validated partial likelihood (CVL) and the second definition as grouped cross-validated partial likelihood.
-Figure 2 illustrates the ideas of the two methods.
-The grouped CVL uses observations more efficiently than the ungrouped CVL.
-Since only the left-out fold would be used, the ungrouped cvl is unstable in practice when there are only very few observations in some folder.
-
-\begin{figure}
-    \centering
-		\includegraphics[height= 4cm ]{./figures/02_1.png}
-		\includegraphics[height= 4cm ]{./figures/02_2.png}
-    \caption{Grouped vs Ungrouped Cross Validated Likelihood}
-\end{figure}
-
-  \subsection{Cross Validated Linear Predictors}
-
-Besides cross validating over the partial likelihood, an alternative approach is to cross validate over the linear predictors.
-The data would still be split into K folds.
-Suppose ith fold is left out and the other K-1 folds are used as the training set to get the estimates $\hat{\beta}_{-i}$.
-Then the cross-validated linear predictors would be calculated based on the observations in the test set:  \begin{equation}\hat{\eta}_{-i} = X_{i}\hat{\beta}_{-i}.\end{equation} After repeating this for all K folds, a whole set of linear predictors  $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...
-\hat{\eta}_{-K})$ can be obtained.
-A partial likelihood can be built over this set of linear predictors: \begin{equation} L(\hat{\eta}_{-}) = \prod_{j=1}^{m} \frac{exp (\hat{\eta}_{-j})}{\sum_{ k \in R(t_{j})}exp (\hat{\eta}_{-k})}\end{equation}.
-
-Cross validated error is defined as the log of $L(\hat{\eta}_{-})$.
-This idea of cross validating over linear predictors is implemented in the package ncvreg.
-For all the approaches introduced so far, they are all equivalent to each other in linear regressions.
-But since they are building over different baselines, they are different in cox regression.
-
-   \begin{figure}[h]
-    \centering
-    \begin{minipage}[b]{0.45\textwidth}
-      \centering
-		  \includegraphics[height= 3.9cm ]{./figures/03.png}
-      \caption{CV Linear Predictors}
-     \end{minipage}
-     \begin{minipage}[b]{0.45\textwidth}
-      \centering
-		  \includegraphics[height= 4.1cm ]{./figures/04_2.png}
-      \caption{CV Deviance Residuals}
-      \end{minipage}	
-   \end{figure}
-
-  \subsection{Cross Validated Deviance Residuals}
-  %Need to update this section with the new baseline estimation method%
-Since the biggest challenge of conducting cross validation for cox regression is due to the missingness of baseline hazard, we propose a second approach which involves estimating the actual baseline hazard $\hat{\Lambda}_{0}$.
-The sum of squares of Martingale residuals is a natural candicate for evaluating the model performance using baseline hazard.
-Martingale residuals and its normalized form, deviance residuals, were first proposed by \citep{Therneau1990} for Cox regression model diagnostics.
-
-To estimate the baseline hazard, a set of linear predictors $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...
-\hat{\eta}_{-K})$ would be obtained in the same way as the cross validated linear predictor approach.
-$\hat{\eta}_{-i}$ is computed based on $\hat{\beta}_{-i}$s estimated from the training set and observations from the ith test set.
-With $\hat{\eta}_{-}$, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}.
-Then for each observations, the Martingale Residual can be calculated: \begin{equation}\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}_{i}},\end{equation} where $\delta_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation.
-The deviance residual can be calcualted from the Martingale Residuals: \begin{equation} d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}\end{equation} The sum of squares of the deviance residuals, $\sum_{i}\hat{d}_{i}^2$, would be used as the cross validated error.
-
-As an alternative, the baseline hazard can also be estimated only based on the training set, leaving the test set as out-of-sample observations for calculating the residuals.
-There are two major issues with this approach.
-First, it would be common to run into situations where events in the test set occurred after the last event occured in the baseline, where the survival function dropped to 0.
-Unless those observations are treated as censored, this approach would not work.
-Second, there would be an extra numeric challenge for deviance residuals.
-Since the baseline hazard function $\hat{\Lambda}_{0}$ is a step function from 0, the cumulative hazard at the first time point from 0 is always 0.
-Since deviance residual contains a term $log(\delta_{i} - \hat{M}_{i})$, where $\delta_{i} - \hat{M}_{i} = \hat{\Lambda}_{0}$, this term would be negative infinity for the first time point and lead to a numeric issue for whichever fold that contains the first time point.
-Hence extra smoothing for the step function would be needed for this approach to work.
-
-\section{Simulation Studies}
-
-  Simulation studies were conducted to compare how those methods behave relative to each other.
-We generate data with pre-specified baseline hazard, covariate matrix X, coefficient $\beta$ and censoring mechanism.
-Both low dimensional and high dimentional scenarios were examined.
-All cross-validation methods mentioned in Section 2 were applied to the data to select the tuning parameter $\lambda$ and produce $\hat{\beta}$ estimates.
-Cross-validation were also compared to model selection criteria AIC and BIC.
-
-  Survival times were generated from exponential distribution $h(t) = h_{0} exp(X\beta)$, conditioned on covariates.
-The entries in the covariate matrix $X_{n \times p}$ were independently generated from Normal (0, 1).
-True coefficients $\beta$  were assumed to have sparsity: $\beta_{p\times 1} = (\beta_{1},\beta_{2}, ..., \beta_{10}, 0, ..., 0)^{T}$.
-Censoring status were generated based on binomial distribution.
-
-  Suppose $\hat{\beta}$ is the coefficients estimated by a fitted cox model.
-We measure the distance between the fitted model and the true model by mean squared error $MSE = E(\hat{\beta} - \beta)^2$.
-For each generated data set, the $\lambda$ that has the minimal MSE is chosen as the optimal $\lambda$.
-Then the $\lambda$s that selected by the cross validation or information criteria would be compared to this optimal $\lambda$.
-If the $\lambda$ chosen by cross validation is smaller than the optimal $\lambda$, then the cross validation method would be considered liberal.
-If the $\lambda$ chosen by cross validation is larger than the optimal $\lambda$, then the cross validation method would be considered conservative.
-
-    \subsection {Simulations Comparing Cross-validation Approaches}
-  %need to change figures to include the new cross-validated deviance residuals
-   The simulation experiments were first conducted under lower dimension settings.
-Number of observations n is set to be 100.
-Dimension of the data p is set to be 100.
-200 replications were used.
-Censoring percentage, number of folds used in cross-validation, number of non-zero $\beta$s and the magnitude of the $\beta$s were varied in several different simulation scenarios.
-Patterns and results shown in various simulation scenarios are consistent with one another.
-Figure 5 and Figure 6 illustrates one simulation scenario, where about 10$\%$ of which were set to be censored and 10 of the $\beta$s were set to be non-zero and 10 folds were used for cross validation.
-The x - axis of the plots are the magnitude of the non-zero $\beta$ coefficients.
-
-   \begin{figure}[h]
-   \centering
-    \begin{minipage}[b]{0.4\textwidth}
-      \centering
-		  \includegraphics[height= 6cm ]{./figures/05.png}
-      \caption{ }
-     \end{minipage}
-     \begin{minipage}[b]{0.4\textwidth}
-      \centering
-		  \includegraphics[height= 6cm ]{./figures/06.png}
-      \caption{ }
-      \end{minipage}	
-   \end{figure}
-
-    $\lambda$s chosen by different cross-validation methods are compared in Figure 5.
-All values are from taking the mean of 200 replications.
-The black line represents the optimal $\lambda$.
-All four cross validation methods consistently select $\lambda$s larger than the optimal one.
-Since larger $\lambda$ leads larger penalties and fewer variables are selected, all four cross validation methods seem to be conservative in terms of selecting variables.
-Ungrouped cross-validated log likelihood and the cross-validated linear predictor seem to be more liberal than the other two methods.
-The cross-validated Deviance residual is the most conservative one and its performance is quite off from the other three methods.
-
-   Mean squared errors are compared in Figure 6.
-The MSEs of the cross-validation selected models are compared with the MSE of the model given by the optimal $\lambda$.
-The y-axis represents the ratio of the two.
-The ungrouped method and the linear predictor approach are outperforming the other methods and are closest to the minimal MSE.
-
-   Simulation studies were also conducted in high dimension scenarios, where the number of observations and dimension of the covariate matrix is more similar to real genetics research.
-The results are shown in Table 1.
-Conclusions are close to what is shown in the low dimensional setting.
-The M- residual approach performs most conservatively.
-Linear predictor approach and ungrouped approach are relatively more liberal than the other approaches but still slightly more conservative than the optimal $\lambda$.
-
- % \begin{table}[h]
-   % \centering
-%			\begin{tabular}{ c | c | c | c}
-%				CV Methods & $\lambda$ (SD) & MSE Ratio \\
-%				\hline \hline
-%				Optimal  & 0.0606 ( 0.0019 )&  / & \\ 
-%				Grouped  &  0.0895  ( 0.0043 )& 1.598  \\
-%				Ungrouped & 0.0754 ( 0.0060 )& 1.225 \\
-%				Linear Pred & 0.0759 ( 0.0058 )& 1.234 \\
-%				M Residual &0.1283 ( 0.0047 ) & 2.959
-%			\end{tabular}
-%			\caption{Simulation Results in High Dimensional Setting}
-%	\end{table}
-
-\begin{table}[h]
-\small
-\centering
-\caption{Simulation Results, Varying n and p}
-\label{my-label}
-\begin{tabular}{llllllllllllllll}
-\hline
-       & \multicolumn{3}{l}{n = 100, p = 20} &  & \multicolumn{3}{l}{n = 100, p = 100} &  & \multicolumn{3}{l}{n = 200, p = 1000} &  & \multicolumn{3}{l}{n = 500, p = 10000} \\
-       & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 20} \\ \cline{2-4} \cline{6-8} \cline{10-12} \cline{14-16}
-       &     &        &MSE       &  &   &       & MSE     &    &    &        & MSE      &  &     &        & MSE      \\
-       & $\lambda$(SD)     & &Ratio       &  & $\lambda$(SD)         & & Ratio     & &$\lambda$(SD)         & & Ratio      & & $\lambda$(SD)         & & Ratio      \\ \cline{2-2} \cline{4-4} \cline{6-6} \cline{8-8} \cline{10-10} \cline{12-12} \cline{14-14} \cline{16-16}
-Optimal &           &        &          &  &           &       &          &  &           &        &          &  &           &        &          \\
-Grouped &           &        &          &  &           &       &          &  &           &        &          &  &           &        &          \\
-Ungrouped &           &        &          &  &           &       &          &  &           &        &          &  &           &        &          \\
-Linear Pred &           &        &          &  &           &       &          &  &           &        &          &  &           &        &          \\
-Deviance &           &        &          &  &           &       &          &  &           &        &          &  &           &        &          \\ \hline
-\end{tabular}
-\end{table}
-
- \subsection {Stability}
-   While the ungrouped cross-validated likelihood is an intuitive way to carry out cross-validation, it is quite unstable when it runs into cases where the number of observed events in some fold is really small.
-As is illustrated in Figure 5, cases of undefined cross-validation error would increase when censoring increases or when number of folds increases.
-
-\begin{figure}
-    \centering
-		\includegraphics[height= 8cm ]{./figures/stability2.png}
-    \caption{Comparing Stability}
-\end{figure}
-
-    \subsection {Comparisons with Information Criteria}
-	%need to change the figures to have fewer lines
-    \par We also compared cross-validation with information criteria AIC and BIC.
-Simulation studies were conducted in lower dimension and higher dimension respectively.
-As is illustrated in Figure 7, in lower dimension, AIC tends to perform more liberal and BIC tends to perform more conservative than the optimal $\lambda$.
-While the cross-validation methods are also conservative, most of them are less conservative than BIC.
-In terms of MSE, cross-validation methods yield smaller MSE than the information criteria.
-In higher dimension, both BIC and AIC tend to be more liberal than the optimal $\lambda$ and yield larger MSE.
-
-   \begin{figure}[h]
-    \centering
-    \begin{minipage}[b]{0.4\textwidth}
-      \centering
-		  \includegraphics[height= 6cm ]{./figures/aic_01.jpeg}
-      \caption{Comparison to Information Criteria in Lower Dimension}
-     \end{minipage}
-     \begin{minipage}[b]{0.4\textwidth}
-      \centering
-		  \includegraphics[height= 6cm ]{./figures/aic_02.jpeg}
-      \caption{Comparison to Information Criteria in Higher Dimension}
-      \end{minipage}	
-   \end{figure}
-
- \subsection {Issues with Deviance Residual}
-
-\section{Application to Real Data}
-
-We also compared those cross validation methods when they are applied to real data sets.
-Unlike simulated set, we do not know the optimal $\lambda$ in the real data sets, but we can still see how those methods perform relative to each other.
-The first data set is a study on ovariance cancer from The Cancer Genome Atlas.
-The second data set is a study on lung cancer (Shedden et al 2004).
-The outcome of both data sets are time-to-event data.
-
-There are 460 patients and 236 events for the ovarian cancer data set.
-About 40$\%$ of the observations are censored.
-The covariate matrix records whether or not there is mutation at the location and has dimension 12376.
-When we fit the penalized cox model, we first adjusted three clinical variables which is known to have large impact on survival.
-We first fitted these three covariates into regular cox regression and keep the linear predictors.
-The selection for the penalty term does not affect the estimation of those clinical variables.
-The results are illustrated in Table 2.
-The linear predictor approach has the best predictive accuracy according to AUC.
-
-\begin{table}[h]
-\centering
-  \begin{tabular}{ c | c | c | c}
-				CV Methods & $\lambda$ & $log(\lambda)$ &AUC \\
-				\hline \hline
-				Grouped  &  0.117 & -2.14 &   0.595\\
-				Ungrouped &  0.112 & -2.20 &  0.597\\
-				Linear Pred &  0.085 & -2.46 &     0.607\\
-				M Residual & 0.171 & -1.76 & 0.529 
-	\end{tabular}
-  \caption{Tuning Parameter Selected for Ovarian Cancer Data}
-\end{table}
-
-In Figure 9, the Cross Validated Error is rescaled and plotted for all four methods.
-A $\lambda$ will be selected when CVE curve reaches its lower point.
-The blue line, which represents the linear predictor approach, it has more curvature near its lowest point.
-It is easier to pick out the minimum point for this blue curve.
-Hence this approach is better at picking out signals than the other three approaches.
-If we look at the curve for the cross-validated Deviance Residual, there's almost no signal there.
-
-   \begin{figure}[h]
-   \centering
-    \begin{minipage}[b]{0.4\textwidth}
-      \centering
-		  \includegraphics[height= 6cm ]{./figures/07.jpeg}
-      \caption{CVE for Ovarian Cancer Data}
-     \end{minipage}
-     \begin{minipage}[b]{0.4\textwidth}
-      \centering
-		  \includegraphics[height= 6cm ]{./figures/08.jpeg}
-      \caption{CVE for Lung Cancer Data}
-      \end{minipage}	
-   \end{figure}
-
-In the lung cancer data set that we applied our methods to, there are 442 patients and 236 events.
-About 50$\%$ of the patients are censored.
-Again we adjusted three clinical variable before we fit the penalized regression.
-The results of the analysis are listed in Table 3.
-
-			\begin{table}[h]
-			\centering
-			\begin{tabular}{ c | c | c | c}
-				CV Methods & $\lambda$ & $log(\lambda)$ & AUC \\
-				\hline \hline
-				Grouped  &   0.119 & -2.12 &  0.608 \\
-				Ungrouped &  0.119 & -2.12 &  0.608 \\
-				Linear Pred &  0.090 & -2.41 &  0.633\\
-				M Residual & 0.203 & -1.59 &  0.564
-			\end{tabular}
-			\caption{Tuning Parameter Selected for Lung Cancer Data}
-			\end{table}
-
-If we look at the CVE curve for all four of those methods, we still see that blue curve has more curvature near its lowest point.
-But in general, there are more signal in this lung cancer data set compared to the ovarian cancer data set.
-
-\section{Discussion}
-	\begin{itemize}
-		\item 	
-		\item
-	\end{itemize}
+\input{main}
 

### 2018-05-03 main.tex
Splitting tex into two files; deleting Rnw

\section{Introduction}
Cox propotional hazard regression is one of the most commonly used statistical model for analyzing data with survival outcomes. Despite its popularity, Cox model is limited by its semiparametric nature: it is challenging to evaluate the model's predictive accuracy.

In the Cox model \citep{Cox1975}, the hazard function for an observation i is given by \begin{equation}h_{i}(t) = h_{0}(t) \text{exp}( X_{i}^{T} \beta)\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{i}^{T} \beta}$ is the relative risk. Suppose there are n observations and m observed failure times, the estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood \begin{equation}L(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation}, where j indexes the observed failure times and $R(t_{j})$ denotes the set of observations at risk at time $t_{j}$. As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard. Hence the interpretation of the Cox model is only valid in a relative sense.

Penalized Cox Regression is an extension of the regular cox regression model. Estimates of the $\beta$ coefficients are obtained by minimizing the objective function \begin{equation}  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{m} \text{log} \left \{\frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta). \end{equation} \citep{Simon2011}. The first part of the objective function is based on the partial likelihood. The second part is a penalty term that depends on some regularization parameters. In this paper, only the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$ was considered. The LASSO cox regression can be used to build prediction models for survival outcome based on high dimensional data, where the number of covariates p $\gg$ n. An example of its application would be to make predictions of cancer patients' the overall survival time with their certain gene expressions.

Selecting the tuning parameter $\lambda$ is crutial to getting LASSO estimations for $\beta$. The LASSO penality enforces sparcity to the regression model, so that it carries out an automatic variable selection along the solution path. As $\lambda$ increases, more coefficients would be set to 0 and fewer variables will be selected by the model. When $\lambda$ is too small, the model tends to be redundant. When $\lambda$ is too large, the model would be too conservative to select any variables. In either case, model's predictive accuray would be affected.

One common approach for selecting the tuning parameter $\lambda$ for linear regression and logistic regression is via the K-fold cross validation. The data set would be split into K folds. One fold would be treated as the test set and the other K - 1 folds as the training set. The model would first be built on the training set, then fitted to the test set to obtain cross validation error (CVE).  CVE would be calculated for each candidate $\lambda$, then the one that minimizes cross validated error would be selected.

However, for penalized cox regression, it is challenging to carry out cross validation. Since the partial likelihood of the cox model does not give estimates for the baseline, it is not meaningful to compare the coefficients if they are not estimated from different models and different observations. Only if the baseline is the same, then the coefficients can be interpreted as relative risks. This is against the idea of cross validation, which splits the data and builds models on different subsets.

%\begin{figure}
 %   \centering
	%	\includegraphics[height= 4cm ]{./figures/cv.png}	
    %\caption{K-fold Cross Validation}
%\end{figure}

%ZZZZZZZZZZZZZ
\par Currently, there are two common cross-validation approaches for Cox regression, both having its disadvantages. The most intuitive approach to conduct cross validation in Cox regression is i) fit the model to training set then ii) calculate cross-validated partial likelihood based on the observations from the test set as a measure for the model's predictive accuracy. As will be discussed in Section 2, this approach is unstable when number of events gets small. The second approach was propsoed by Verweij et al. \citep{Verweij1993}. This is a stablized version of cross-validated log likehood, which has been widely used and implemented in R packages such as $\tt{glmnet}$ \citep{glmnet}. Although this approach fixes the stability issue, it tends to yield more conservative models in practice. We hope to improve cross-validation procedure for Cox regression to achieve a good balance between stability and performance.

\par In this paper, we proposed two alternative ways to carry out cross validation. Instead of cross-validating over partial likelihood, we propose to cross validate over the linear predictors of the regression model and cross validate over the deviance residual \citep{Therneau1990}of the Cox model. We conducted simulation studies to compare how those methods perform in both low dimensional and high dimensional settings with LASSO penalty. We compared the two proposed methods with the existing cross-validated partial likelihood approaches. We showed that the linear predictor approach outperforms other cross validation approaches in various scenarios. We found out that using cross validation to select tuning parameters for LASSO penalized cox regression tends to be a conservative approach in general. Finally, we apply those methods to two high dimensional data sets from real studies with time-to-event outcomes.

\section{Methods}

  \subsection{Cross Validated Likelihood} 
Suppose a data set of n observations is split into K folds and the ith fold is left out, then a partial likelihood can be built over the K -1 folds and yields maximum likelihood estimates of the coefficients, denoted by $\hat{\beta}_{-i}$. An intuitive way to carry out cross validation is to use the partial likelihood on the test data as the cross validated error: \begin{equation}cvl = \sum_{i=1}^{K} l_{i}(\hat{\beta}_{-i})\end{equation} This is implemented in the $\tt{glmnet}$ package as the $\tt{ungrouped}$ option.

\par An alternative way to calculate cross-validated log likelihood was proposed by Verweij et al. They defined cross-validated log likelihood of leaving the ith fold out to be \begin{equation}cvl = \sum_{i=1}^{K} \{l(\hat{\beta}_{-i}) - l_{-i}(\hat{\beta}_{-i})\}. \end{equation} When the ith fold is left out, $l(\hat{\beta}_{-i})$ is the log partial likelihood evaluated at $\hat{\beta}_{-i}$ with all k folds of observations. $l_{-i}(\hat{\beta}_{-i})$ is the log partial likelihood evaluated at $\hat{\beta}_{-i}$ with observations from the other K-1 folds.

\par We refer to the first definition as ungrouped cross-validated partial likelihood (CVL) and the second definition as grouped cross-validated partial likelihood. The upper panels in Figure 2 illustrates the ideas of the two methods. The grouped CVL uses observations more efficiently than the ungrouped CVL. Since only the left-out fold would be used, the ungrouped CVL is unstable in practice when there are only very few observations in some folder.

  \subsection{Cross Validated Linear Predictors}

Besides cross validating over the partial likelihood, an alternative approach is to cross validate over the linear predictors. In the first step, the ith fold would be left out and the other K-1 folds would be used as the training set to get the estimates $\hat{\beta}_{-i}$. Then, the cross-validated linear predictors would be calculated based on the observations in the test set:  \begin{equation}\hat{\eta}_{-i} = X_{i}\hat{\beta}_{-i}.\end{equation} After repeating this for all K folds, a whole set of linear predictors  $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...  \hat{\eta}_{-K})$ can be obtained. A partial likelihood can be built over this set of linear predictors: \begin{equation} L(\hat{\eta}_{-}) = \prod_{j=1}^{m} \frac{exp (\hat{\eta}_{-j})}{\sum_{ k \in R(t_{j})}exp (\hat{\eta}_{-k})}\end{equation}.

Cross-validated error is defined as the log of $L(\hat{\eta}_{-})$. This idea of cross validating over linear predictors is implemented in the package \texttt{ncvreg} \citep{ncvreg}. For all the approaches introduced so far, they are all equivalent to each other in linear regressions. But since they are building over different baselines, they are different in cox regression.

  % \begin{figure}[h]
    %\centering
    %\begin{minipage}[b]{0.45\textwidth}
      %\centering
	%	  \includegraphics[height= 3.9cm ]{./figures/03.png}
      %\caption{CV Linear Predictors}
     %\end{minipage}
     %\begin{minipage}[b]{0.45\textwidth}
      %\centering
%		  \includegraphics[height= 4.1cm ]{./figures/04_2.png}
    %  \caption{CV Deviance Residuals}
     % \end{minipage}	
  % \end{figure}

  \subsection{Cross Validated Deviance Residuals}
  %Need to update this section with the new baseline estimation method%
The fundamental challenge of conducting cross validation for cox regression is that the baseline hazard is not estimated from the model. Hence we proposed our second approach which involves estimating the actual baseline hazard $\hat{\Lambda}_{0}$. The sum of squares of Martingale residuals is a natural candicate for evaluating the model performance using baseline hazard. Martingale residuals and its normalized form, deviance residuals, were first proposed by \citep{Therneau1990} for Cox regression model diagnostics.

To estimate the baseline hazard, a set of linear predictors $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...  \hat{\eta}_{-K})$ would be obtained in the same way as the cross validated linear predictor approach. $\hat{\eta}_{-i}$ is computed based on $\hat{\beta}_{-i}$s estimated from the training set and observations from the ith test set. With $\hat{\eta}_{-}$, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}. Then for each observations, the Martingale Residual can be calculated: \begin{equation}\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}_{i}},\end{equation} where $\delta_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation. The deviance residual can be calcualted from the Martingale Residuals: \begin{equation} d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}\end{equation} The sum of squares of the deviance residuals, $\sum_{i}\hat{d}_{i}^2$, would be used as the cross validated error. An illustration of this method is drawn in the bottom right panel of Figure 1.

As an alternative, the baseline hazard can also be estimated only based on the training set, leaving the test set as out-of-sample observations for calculating the residuals. There are two major issues with this approach. First, it would be common to run into situations where events in the test set occurred after the last event occured in the baseline, where the survival function dropped to 0. Unless those observations are treated as censored, this approach would not work. Second, there would be an extra numeric challenge for deviance residuals. Since the baseline hazard function $\hat{\Lambda}_{0}$ is a step function from 0, the cumulative hazard at the first time point from 0 is always 0. Since deviance residual contains a term $log(\delta_{i} - \hat{M}_{i})$, where $\delta_{i} - \hat{M}_{i} = \hat{\Lambda}_{0}$, this term would be negative infinity for the first time point and lead to a numeric issue for whichever fold that contains the first time point. Hence extra smoothing for the step function would be needed for this approach to work.

\begin{figure}
    \centering
		\includegraphics[height= 9cm ]{./figures/figure_1.png}
		%\includegraphics[height= 4cm ]{./figures/02_2.png}
    \caption{Illustration of K-fold Cross Validation Methods for Cox Regression}
\end{figure}

\section{Simulation Studies}

  Simulation studies were conducted to compare how those methods behave relative to each other. We generated data with pre-specified baseline hazard, covariate matrix X, coefficient $\beta$ and censoring mechanism. Both low dimensional and high dimentional scenarios were examined. All cross-validation methods mentioned in Section 2 were applied to the data to select the tuning parameter $\lambda$ and produce $\hat{\beta}$ estimates. %Cross-validation were also compared to model selection criteria AIC and BIC.

 The entries in the covariate matrix $X_{n \times p}$ were independently generated from $Normal(0, 1)$. True coefficients $\beta$ were assumed to have sparsity: $\beta_{p\times 1} = (\beta_{1},\beta_{2}, ..., \beta_{j}, 0, ..., 0)^{T}$. The covariate matrix $X_{n \times p}$ was centered before it was used to generate outcomes. Survival times were generated from exponential distribution $h(t) = h_{0} exp(X\beta)$, conditioned on covariates. Censoring status were generated based on binomial distribution. All simulations were implemented in R \citep{R}.

  Suppose $\hat{\beta}$ is the coefficients estimated by a fitted cox model. We quantify the predictive accuracy by measuring the distance between the fitted model and the true model by mean squared error MSE $= E(\hat{\beta} - \beta)^2$. For each generated data set, the $\lambda$ that has the minimal MSE is chosen as the optimal $\lambda$. Then the $\lambda$s that selected by the cross validation or information criteria would be compared to this optimal $\lambda$. If the $\lambda$ chosen by cross validation is smaller than the optimal $\lambda$, then the cross validation method would be considered liberal. If the $\lambda$ chosen by cross validation is larger than the optimal $\lambda$, then the cross validation method would be considered conservative.

    \subsection {Simulations Comparing Cross-validation Approaches}
  %need to change figures to include the new cross-validated deviance residuals
   The simulation experiments were first conducted when the signal in the data was varied. In each simulated data set, there are n = 100 observations and p = 1000 covariates. 10$\%$ of the observations were censored. 10-fold cross-validation was implemented for all four methods. Number of non-zero $\beta$s was 20. The signal in the data set was varied by varing the magnitude of the $\beta$s from 0.50 to 1.50. For each scenario, 200 replications were used.

\begin{figure}[h]
    \centering
		\includegraphics[height= 7cm ]{./figures/figure_2.png}
    \caption{Simulation Studies Comparing 4 CV methods in terms of $\lambda$ and MSE ratio. n = 200, p = 1000, 10$\%$ censoring.}
\end{figure}

\par Results of this simulation were illustrated in Figure 2.  The x axis of the plots are the magnitude of the non-zero $\beta$ coefficients. The left panel compares the $\lambda$ selected by different methods. The grouped cross-validated partial likelihood approach, which is currently most widely used approach, consistently selects $\lambda$s that are greater than the optimal one.  Greater $\lambda$ leads to greater penalties and fewer non-zero variables, hence the grouped approach is conservative in terms of selecting variables. While the proposed cross-validated Deviance residual has very close performance as the grouped approach, the ungrouped cross-validated log likelihood and the cross-validated linear predictor were more liberal than the other two methods. In general, all four methods tends to act conservatively and selects smaller models.

\par The right panel compares the MSE ratios of the model selected. The MSEs of the cross-validation selected models are compared with the MSE of the model given by the optimal $\lambda$. The y-axis represents the ratio of the two. The ungrouped method has the smallest MSE ratio. The linear predictor approach's performance is also very close the ungrouped appraoch. Both the grouped approach and the proposed deviance residual approach yield models that have greater MSE, hence are not recommended for use in practice.

\begin{table}[h]
\small
\centering
\caption{Simulations Comparing all four Cross-Validation methods, Varying n and p}
\begin{tabular}{llllllllllllllll}
\hline\\[-0.75em]
       & \multicolumn{3}{l}{n = 100, p = 20} &  & \multicolumn{3}{l}{n = 100, p = 100} &  & \multicolumn{3}{l}{n = 500, p = 10000} &  & \multicolumn{3}{l}{n = 250, p = 10000} \\
       & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 20} &  & \multicolumn{3}{l}{$\#$of non-zero = 20} \\ \cline{2-4} \cline{6-8} \cline{10-12} \cline{14-16} 
\hline\\[-0.75em]
       &     &        &MSE       &  &   &       & MSE     &    &    &        & MSE      &  &     &        & MSE      \\
       & $\lambda$ (SD)     & &Ratio       &  & 	$\lambda$ (SD)         & & Ratio     & &	$\lambda$ (SD)         & & Ratio      & &	 $\lambda$ (SD)         & & Ratio      
\\ 
\cline{2-2} \cline{4-4} \cline{6-6} \cline{8-8} \cline{10-10} \cline{12-12} \cline{14-14} \cline{16-16} 
\\[-0.75em]
Optimal &0.016 (0.009)&&/&&0.042 (0.003)&&/&&0.026 (0.001)&&/&&0.026 (0.001)&&/\\          
Grouped &0.035 (0.007)&&2.597&&0.073 (0.009)&&2.184&&0.053 (0.003)&&5.627&&0.123 (0.036)&&3.931\\
Ungrouped &0.030 (0.009)&&2.246&&0.064 (0.01)&&1.815&&0.040 (0.003)&&3.131&&0.070 (0.016)&&2.560\\
LinearPred &0.034 (0.007)&&2.559&&0.066 (0.008)&&1.880&&0.041 (0.003)&&3.350&&0.074 (0.016)&&2.713\\
DevResid &0.100 (0.009)&&14.005&&0.108 (0.008)&&4.087&&0.076 (0.003)&&9.133&&0.111 (0.028)&&3.767\\ 
\hline
\end{tabular}
\end{table}

\par Simulation studies were also conducted in scenarios with various sample sizes and dimensions. The results are shown in Table 1. While the ungrouped approach yields the smallest MSE and best predictive accuracy, the Linear Predictor and the Ungrouped approaches perform closely to each other across all settings. Grouped approach performs well in low-dimensions or with sufficient samples, while deviance residual approach performs better in high-dimensional settings.

 \subsection {Stability}

\par In this section, simulations studies were conducted to address the stability issue of the ungrouped cross-validated likelihood. The ungrouped approach is an intuitive way to carry out cross-validation. It also has the best predictive accuarcy based on the simulation results in previous sections. However, one big disadvantage of this approach is that it is unstable when number of observed events in some fold is really small. As is illustrated in Figure 3, when censoring increases or when number of folds increases, the model would fail to converge due to insufficient number of events in the test set. Hence, the ungrouped cross-validated likelihood should not be used when there is no sufficient number of events.

\begin{figure}[h]
    \centering
		\includegraphics[height= 7cm ]{./figures/figure_3.png}
    \caption{Comparing Stability}
\end{figure}

    \subsection {Results of Leave-One-Out Cross Validation}
	\par In this section, Leave-One-Out Cross Validation (LOOCV) were studied, in comparison with 10-fold CV. When first introduced, the grouped cross-validated partial likelihood was proposed so that LOOCV for Cox regression became feasible. It is impossible to carry out LOO cross-validation using the ungrouped approach, because the partial likelihood cannot be built on the one left-out observation. While the grouped cross-validated partial likelihood solves the stability issue by aggregating more observations together, it is not an efficient way to evaluate the model fit. Simulation studies in Section 3.1 have illustrated that it tends to be conservative in variable selections and selects smaller models in various scenarios. This fact can be even more illustrated via the following simulation studies, where both 10-fold and LOOCV were conducted via grouped, linear predictor and deviance residual approaches and 10 fold CV was conducted for the ungrouped approach.
	\par As is illustrated in Figure 4, LOOCV selected models that have smaller MSE than 10-fold CV for all three CV approaches. As is shown in left and middle panels, LOOCV based on deviance residuals and grouped cross-validated partial likelihood still perform more conservatively than the ungrouped approach. Even if the grouped approach was proposed to allow LOOCV, the LOOCV based on the grouped approach does not surpass the 10-fold ungrouped approach. On the other hand, the Linear Predictors also allows LOOCV, and the LOOCV based on Linear Predictors performs better than the ungrouped approach.

    %\par We also compared cross-validation with information criteria AIC and BIC. Simulation studies were conducted in lower dimension and higher dimension respectively. As is illustrated in Figure 7, in lower dimension, AIC tends to perform more liberal and BIC tends to perform more conservative than the optimal $\lambda$. While the cross-validation methods are also conservative, most of them are less conservative than BIC. In terms of MSE, cross-validation methods yield smaller MSE than the information criteria. In higher dimension, both BIC and AIC tend to be more liberal than the optimal $\lambda$ and yield larger MSE.

\begin{figure}[h]
    \centering
		\includegraphics[height= 8cm ]{./figures/figure_4.png}
    \caption{LOOCV}
\end{figure}

  \subsection {Simulations on Deviance Residuals and Baseline Hazard Estimations}
\par While the Deviance Residual approach has a reasonable performance in scenarios where dimension is high, it tends to act more conservatively than all the other three approaches in general. As is noted by Therneau et al, the sums of squared martingale residuals and deviance residuals do not necessary reflect how good the model fit is \citep{Therneau2000modeling}. The difficulty of estimating the baseline hazard seems to play an important role. A series of simulations were thus designed to study how baseline hazard estimation affects the deviance residuals' performance as a measurement for model fit.

\par In the following simulation studies, the defined optimal $\lambda$ and the $\lambda$ selected by the Linear Predictor approach were used as benchmarks. Various sample sizes and dimensions were varied. The survival outcomes were generated both from exponential distribution and Weibull distribution. Six different approaches were compared in terms of baseline hazard estimation and their impact on sums of squared deviance residuals. Two non-parametric approaches, Kalbfleisch and Prentices' estimator and Breslow's estimator, were used first for estimating the cumulative hazard baseline \citep{Kalbfleisch2011} \citep{breslow1972}.  As is presented in column 3 and column 5 in Table 2, the sum of squared deviance residuals computed based on both approaches are all very conservative compared to the optimal $\lambda$. Their performance is extremely bad when n is large and p is small.

\par Since both of the non-parameteric estimations seemed to perform well at early time points but poorly later, a weighted approach was then applied to emphasize earlier time points. Weights are assigned proportional to number of subjects at risk, making earlier time points weighted more than later time points. Results of these two approahces are presented in column 4 and 6 in Table 2. The attempt to improve the non-parameteric estimations via weighting does not seem to improve the selection too much.

\par In the next step,  the non-parametric baseline hazard estimation was replaced by the exact baseline hazard that was used to generate the outcome. Results were listed in the seventh column in Table 2. When the baseline was known, the performance of sum of squared deviance residuals actually performed really close to the optimal $\lambda$. This simulation suggests that when the baseline hazard was accurately estimated, then the sum of squared deviance residuals gave really good measures for the model fit.

\par In the last step, we replaced the non-parametric baseline hazard estimation with a parametric approach, where we assume the baseline hazard to follow an exponential distribution. We derived the estimation for baseline hazard by maximum likelihood approach conditioned on covariates: $\hat{\lambda}_{0} = \frac{\sum d_{i}}{\sum t_{i}e^{\eta_{i}}}$. The results are listed in the eighth column in Table 2. When this estimation was used for data that were generated from exponential distribution, its performance was as well as when the exact baseline was specified. When it was fitted to data that were generated from the Weibull distribution, that is when the model is misspecified, then its performance was even worse than the non-parametric approaches.

\par Hence, getting accurate estimation of the baseline hazard is crucial to whether or not the sum of squared deviance residuals can be a good measurement for model fit.

\begin{table}[h]
	\small
	\caption{Average of Selected $\lambda$ with Different Baseline Estimation for Deviance Residuals}
	\centering
	\begin{tabular}{lllllllll}
	\hline
	\\[-0.75em]
    & Optimal & Linear Pred & KP & KP & Breslow & Breslow & Exact & Exp\\ 
    &  & & & weighted & & weighted & & \\ \cline{2 - 9} 
\textbf{True Baseline: $Exponential(1)$} &&&&&&&&\\
n = 100, p = 10, non-zero $\beta$: 2 & 0.0461 & 0.0665 & 0.1986 & 0.1970 & 0.1908 & 0.1975 & 0.0540 & 0.0539 \\
n = 100, p = 100, non-zero $\beta$: 20 &0.0555 & 0.1841 & 0.2359 & 0.2201 & 0.2146 & 0.2244 & 0.0805 & 0.0985 \\
n = 100, p = 1000, non-zero $\beta$: 20 & 0.0268 & 0.0505 & 0.0948 & 0.0798 & 0.0867 & 0.0809 & 0.0288 & 0.0270 \\
\textbf{True Baseline:} $Weibull$ &&&&&&&& \\
 n = 100, p = 10, non-zero $\beta$: 2 & 0.0476 & 0.0710 & 0.2004 & 0.1969 & 0.1907 & 0.1907 & N/A & 0.2330 \\ 
\hline
\end{tabular}
\end{table}

\section{Application to Real Data}

\par In this section, we compared the performances of those cross validation methods when they are applied to data from real studies. The first data set is collected from a study on ovariance cancer from The Cancer Genome Atlas. The second dataset is a study on lung cancer \citep{shedden2008gene}. The outcomes of both data sets are patients' overall survival time.

\par In the ovarian cancer data set, there are 460 patients and 236 events . About 40$\%$ of the observations are censored. The covariate matrix records whether or not there is mutation at certain gene location and has dimension of 12376. Three clinical variables, age, gender and residuals, which were known to have large impact on survival were adjusted in the analysis. That is to say, the clinical covariates were first fitted into regular cox regression and only linear predictors were kept so that the selection for the penalty term would not affect the estimation of those clinical variables. Cross-validation was used for the selection of the penalty term, which affects the selection of gene mutations. The results are illustrated in Table 3.

\par In the lung cancer data set , researchers collected gene expression data for 22283 genes. 236 death events were observed for 442 patients. About 50$\%$ of the patients are censored. Again clinical variables, age, gender, status of getting adjuvant chemotherapy were adjusted before the LASSO cox regression was fitted. Cross-validation was again used for the selection of the penalty term, which selects the gene expressions that are associated with patients' overall survival. The results of the analysis are also listed in Table 3.

\begin{table}[h]
\centering
\caption{Application to Real Data Sets}
\label{my-label}
\begin{tabular}{llllll}
\hline
\\[-0.75em]
 & Ovarian & & & Shedden&  \\
& $\lambda$ & log($\lambda$) &  &$\lambda$  &  log($\lambda$)   \\ \cline{2-3} \cline{5-6} 
Grouped &0.1226 & -2.0985 && 0.1215 & -2.1079\\
Ungrouped & 0.0803 & -2.5222 && 0.0983 & -2.3197\\
Linear Pred &0.0828 & -2.4919 && 0.1013 & -2.2894\\
Dev Resid & 0.0992 & -2.3104 && 0.1110 & -2.1987\\ \hline
\end{tabular}
\end{table}

\begin{figure}[h]
    \centering
		\includegraphics[height= 8cm ]{./figures/figure_5.png}
    \caption{Comparing Cross-Validated Error for Different Methods on Real Data Sets}
\end{figure}

\par How the CV methods perform on real data reflects what is shown in simulation studies. The ungrouped CVL is the most liberal approach. Deviance Residual and Grouped CVL approaches were both really conversative, while the ungrouped approach was the second liberal to ungrouped CVL. In Figure 5, the Cross Validated Error is rescaled and plotted for all four methods. A $\lambda$ will be selected when CVE curve reaches its lower point. The blue line, which represents the linear predictor approach, has more curvature near its lowest point. It is easier to pick out the minimum point for this blue curve. Hence this approach is better at picking out signals than the other three approaches. For the grouped cross-validated partial likelihood, there's almost no signal there.

\section{Discussion}
Four methods of conducting cross-validation for Cox Regression were compared through simulation studies in terms of their predictive accuracy. All cross-validation methods tend to select the more conservative model: with greater value of $\lambda$, and fewer variables selected into the model. Despite of the stability issue, the ungrouped approach is the most liberal one across all scenarios and yield the least MSE. The proposed linear predictor approach performs closely to the ungrouped approach. Under LOOCV, the linear predictor approach out-performs the ungrouped approach. Both the grouped approach and the proposed Deviance Residual approach perform conservatively, hence not recommended for use in practice. Results from real data application reflects results from the simulation studies.

### 2018-05-04 cross-validation.tex
Edits to abstract (plus title/authors)

@@
 
-\title{Comparing Cross Validation Methods for Penalized Cox Regression}
-\author{
-}
-\date{ }
+\providecommand{\note}[1]{\textcolor{red}{#1}}
+
+\title{Cross Validation Approaches for Penalized Cox Regression}
+\author{Biyue Dai\\Department of Biostatistics\\University of Iowa
+  \and
+  Patrick Breheny\\Department of Biostatistics\\University of Iowa}
+\date{\today}
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
@@
 \section*{Abstract}
-While cross validation is commonly used for selecting tuning parameters in penalized regression, little research has been conducted to study cross validation methods for penalized Cox regression.
-Due to its semi-parametric nature, conducting cross validation for Cox models has always been a challenge.
-We propose two new cross-validation methods for Cox Regression, and compare them to traditional information criteria as well as a cross-validated partial likelihood approach originally proposed by Verweij et al.
-Our simulation studies show that, in general, cross-validation tends to be conservative (i.e., select smaller models than the ideal choice of tuning parameters) for penalized Cox regression models.
-However, our proposed approach of cross-validating the linear predictors generally offers the best balance of stability and performance.
-We also illustrate these approaches on data from studies of gene expression and overall survival in cancer patients.
-
+Cross validation is commonly used for selecting tuning parameters in penalized regression, but little research has been conducted to study the performance of cross validation in penalized Cox regression.
+Due to its partial likelihood construction, carrying out cross validation for Cox models is not straightforward, and there are several ways one could imagine applying it.
+Here, we propose two new cross-validation methods for Cox regression and compare them to approaches that have been proposed elsewhere.
+\note{as well as to information criteria such as AIC and BIC -- I DON'T THINK THIS IS CURRENTLY IN THE PAPER ANYMORE?}
+Our simulation studies show that in general, all cross-validation approaches tend to be conservative (i.e., select smaller models than the optimal choice of tuning parameters) for penalized Cox regression models.
+However, our proposed approach of cross-validating the linear predictors seems to offer an attractive balance of performance and numerical stability.
+We illustrate these advantages using simulated data as well as real data from two genome-wide studies of survival in cancer patients.
 \input{main}
@@
 \bibliographystyle{abbrvnat}
-\setcitestyle{round} 
+%\setcitestyle{round} 
 \bibliography{articles}

### 2018-05-05 cross-validation.tex
change figure 1,3, add definition of l_{i}

@@
 Here, we propose two new cross-validation methods for Cox regression and compare them to approaches that have been proposed elsewhere.
-\note{as well as to information criteria such as AIC and BIC -- I DON'T THINK THIS IS CURRENTLY IN THE PAPER ANYMORE?}
 Our simulation studies show that in general, all cross-validation approaches tend to be conservative (i.e., select smaller models than the optimal choice of tuning parameters) for penalized Cox regression models.

### 2018-05-05 main.tex
change figure 1,3, add definition of l_{i}

@@
 
-In the Cox model \citep{Cox1975}, the hazard function for an observation i is given by \begin{equation}h_{i}(t) = h_{0}(t) \text{exp}( X_{i}^{T} \beta)\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{i}^{T} \beta}$ is the relative risk.
-Suppose there are n observations and m observed failure times, the estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood \begin{equation}L(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation}, where j indexes the observed failure times and $R(t_{j})$ denotes the set of observations at risk at time $t_{j}$.
+Suppose there are n subjects and m observed failure events.
+Use i for indexing observations, j for indexing events sorted by their failure time.
+For the ith subject, its covarites will be denoted by $X_{i}$; $\delta_{i}$ denotes whether the event is observed or not for this subject.
+$t_{j}$ denotes the failure time for the j th event; $t_{i}$ denotes the time observed for ith subject.
+In the Cox model \citep{Cox1975}, the hazard function for an observation i is given by
+\begin{equation}h_{i}(t) = h_{0}(t) \text{exp}( X_{i}^{T} \beta)\end{equation} 
+where $h_{0}$ is the baseline hazard and $e^{X_{i}^{T} \beta}$ is the relative risk.
+The estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood.
+For each individual subject, its contribution to the partial likelihood is
+\begin{equation} l_{i}(\beta) = \left \{\frac{\text{exp} ( X_{i}^{T} \beta)}{\sum_{ k \in R(t_{i})}\text{exp} ( X_{k}^{T} \beta)}\right \}^{\delta_{i}} \end{equation} \note{introduce notation $l_{i}$} , where $R(t_{i})$ denotes the set of observations at risk at time $t_{i}$.
+
+For all n subjects, the partial likelihood is 
+\begin{equation} L(\beta) =\prod_{i = 1}^{n} l_{i}(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation}.
 As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard.
@@
 \end{figure}
+
+\note{tranpose sign in figure 1 is corrected}
 

### 2018-05-05 simulation/scenario_3.R

@@ -143,7 +143,7 @@ for(j in 1:length(nfold)){
   #foldid <- sample(1:nfold[j],size=n,replace=TRUE)
   for (i in 1:N){
     set.seed(666 + i)
-    data <- genSurvData(n=100 , beta = beta, cpct = 0.2)
+    data <- genSurvData(n=100 , beta = beta, cpct = 0.3) # changed from 0.2 to 0.3
     f <- try(survncvreg(data$x,data$y,data$beta))
     cv.gp <- try(cv.glmnet(x = data$x,y = data$y,nfolds = nfold[j],lambda = f$lambda,family = "cox"))#foldid=foldid,
     cv.ug <- try(cv.glmnet(data$x,data$y,nfolds = nfold[j],lambda = f$lambda,grouped = FALSE,family = "cox"))#foldid=foldid
@@ -151,7 +151,7 @@ for(j in 1:length(nfold)){
     cv.dvr <- try(cv.ncvsurv.KP(X = data$x,y = data$y,nfolds = nfold[j],lambda = f$lambda,penalty = "lasso"))#foldid = foldid
     # true
     true.lambda.fold[i,j] <- f$lambda_t
-    true.mse.cen[i,j] <- min(f$norm)
+    true.mse.fold[i,j] <- min(f$norm)
     
     # gp
     if (typeof(cv.gp) == "character"){

### 2018-05-11 main.tex
Change naming of gp and ungp in text

@@
 \section{Introduction}
-Cox propotional hazard regression is one of the most commonly used statistical model for analyzing data with survival outcomes.
-Despite its popularity, Cox model is limited by its semiparametric nature: it is challenging to evaluate the model's predictive accuracy.
+Cox propotional hazard regression, since its original proposal in 1972, is by far the most common regression approach for analyzing survival data.
+Cox regression is built upon partial likelihood.
+Under the propotional hazard assumption, partial likelihood still preserves most properties of a full likelihood, while the estimation of the baseline can be skipped.
+Despite its popularity, Cox model is limited by this semiparametric nature: it is challenging to evaluate the model's predictive accuracy.
 
@@
 Use i for indexing observations, j for indexing events sorted by their failure time.
-For the ith subject, its covarites will be denoted by $X_{i}$; $\delta_{i}$ denotes whether the event is observed or not for this subject.
-$t_{j}$ denotes the failure time for the j th event; $t_{i}$ denotes the time observed for ith subject.
+For the i th subject, its covarites will be denoted by $X_{i}$; $\delta_{i}$ denotes whether an event is observed or not for subject i.
+$t_{i}$ denotes the time observed for subject i; $t_{j}$ denotes the failure time for the j th event.
 In the Cox model \citep{Cox1975}, the hazard function for an observation i is given by
@@
 where $h_{0}$ is the baseline hazard and $e^{X_{i}^{T} \beta}$ is the relative risk.
-The estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood.
+The estimation of the coefficients $\beta$s is obtained by maximizing the partial likelihood.
 For each individual subject, its contribution to the partial likelihood is
@@
 
-For all n subjects, the partial likelihood is 
+For all n subjects, the partial likelihood is constructed as a product of all individual contributed partial likelihood:
 \begin{equation} L(\beta) =\prod_{i = 1}^{n} l_{i}(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation}.
@@
 
-Selecting the tuning parameter $\lambda$ is crutial to getting LASSO estimations for $\beta$.
-The LASSO penality enforces sparcity to the regression model, so that it carries out an automatic variable selection along the solution path.
+Selecting the tuning parameter $\lambda$ is crucial to getting LASSO estimations for $\beta$.
+The LASSO penalty enforces sparcity to the regression model, so that it carries out an automatic variable selection along the solution path.
 As $\lambda$ increases, more coefficients would be set to 0 and fewer variables will be selected by the model.
@@
 When $\lambda$ is too large, the model would be too conservative to select any variables.
-In either case, model's predictive accuray would be affected.
+In either case, model's predictive accuracy would be affected.
 
@@
 As will be discussed in Section 2, this approach is unstable when number of events gets small.
-The second approach was propsoed by Verweij et al.
+The second approach was proposed by Verweij et al.
 \citep{Verweij1993}.
-This is a stablized version of cross-validated log likehood, which has been widely used and implemented in R packages such as $\tt{glmnet}$ \citep{glmnet}.
+This is a stabilized version of cross-validated log likelihood, which has been widely used and implemented in R packages such as $\tt{glmnet}$ \citep{glmnet}.
 Although this approach fixes the stability issue, it tends to yield more conservative models in practice.
@@
 An intuitive way to carry out cross validation is to use the partial likelihood on the test data as the cross validated error: \begin{equation}cvl = \sum_{i=1}^{K} l_{i}(\hat{\beta}_{-i})\end{equation} This is implemented in the $\tt{glmnet}$ package as the $\tt{ungrouped}$ option.
+We refer to this definition as standard cross-validated partial likelihood.
 
-\par An alternative way to calculate cross-validated log likelihood was proposed by Verweij et al.
+\par An alternative way to calculate cross-validated log likelihood was proposed by Verweij and Van Houwelingen.
 They defined cross-validated log likelihood of leaving the ith fold out to be \begin{equation}cvl = \sum_{i=1}^{K} \{l(\hat{\beta}_{-i}) - l_{-i}(\hat{\beta}_{-i})\}.
@@
 $l_{-i}(\hat{\beta}_{-i})$ is the log partial likelihood evaluated at $\hat{\beta}_{-i}$ with observations from the other K-1 folds.
+We refer to this definition as Verweij and Van Houwelingen's cross-validated partial likelihood.
+This is implemented in the $\tt{glmnet}$ package as the default $\tt{grouped}$ option.
 
-\par We refer to the first definition as ungrouped cross-validated partial likelihood (CVL) and the second definition as grouped cross-validated partial likelihood.
-The upper panels in Figure 2 illustrates the ideas of the two methods.
-The grouped CVL uses observations more efficiently than the ungrouped CVL.
-Since only the left-out fold would be used, the ungrouped CVL is unstable in practice when there are only very few observations in some folder.
+\par The upper panels in Figure 2 illustrates the ideas of the two methods.
+Through substraction, Verweij and Van Houwelingen's approach allows measuring individual observation's contribution to the partial likelihood, so that leave-one-out cross-validation is feasible.
+Since only the left-out fold would be used to construct the partial likelihood, the standard CVL is unstable in practice when there are only very few observations in some folder.
 
@@
 Hence we proposed our second approach which involves estimating the actual baseline hazard $\hat{\Lambda}_{0}$.
-The sum of squares of Martingale residuals is a natural candicate for evaluating the model performance using baseline hazard.
+The sum of squares of Martingale residuals is a natural candidate for evaluating the model performance using baseline hazard.
 Martingale residuals and its normalized form, deviance residuals, were first proposed by \citep{Therneau1990} for Cox regression model diagnostics.
@@
 There are two major issues with this approach.
-First, it would be common to run into situations where events in the test set occurred after the last event occured in the baseline, where the survival function dropped to 0.
+First, it would be common to run into situations where events in the test set occurred after the last event occurred in the baseline, where the survival function dropped to 0.
 Unless those observations are treated as censored, this approach would not work.
@@
 We generated data with pre-specified baseline hazard, covariate matrix X, coefficient $\beta$ and censoring mechanism.
-Both low dimensional and high dimentional scenarios were examined.
+Both low dimensional and high dimensional scenarios were examined.
 All cross-validation methods mentioned in Section 2 were applied to the data to select the tuning parameter $\lambda$ and produce $\hat{\beta}$ estimates.
@@
 Number of non-zero $\beta$s was 20.
-The signal in the data set was varied by varing the magnitude of the $\beta$s from 0.50 to 1.50.
+The signal in the data set was varied by varying the magnitude of the $\beta$s from 0.50 to 1.50.
 For each scenario, 200 replications were used.
@@
 The left panel compares the $\lambda$ selected by different methods.
-The grouped cross-validated partial likelihood approach, which is currently most widely used approach, consistently selects $\lambda$s that are greater than the optimal one.
-Greater $\lambda$ leads to greater penalties and fewer non-zero variables, hence the grouped approach is conservative in terms of selecting variables.
-While the proposed cross-validated Deviance residual has very close performance as the grouped approach, the ungrouped cross-validated log likelihood and the cross-validated linear predictor were more liberal than the other two methods.
+Verweij and Van Houwelingen's cross-validated partial likelihood approach, which is currently most widely used approach, consistently selects $\lambda$s that are greater than the optimal one.
+Greater $\lambda$ leads to greater penalties and fewer non-zero variables, hence Verweij and Van Houwelingen's approach is conservative in terms of selecting variables.
+While the proposed cross-validated Deviance residual has very close performance as Verweij and Van Houwelingen's approach, the standard cross-validated log likelihood and the cross-validated linear predictor were more liberal than the other two methods.
 In general, all four methods tends to act conservatively and selects smaller models.
@@
 The y-axis represents the ratio of the two.
-The ungrouped method has the smallest MSE ratio.
-The linear predictor approach's performance is also very close the ungrouped appraoch.
-Both the grouped approach and the proposed deviance residual approach yield models that have greater MSE, hence are not recommended for use in practice.
+The standard method has the smallest MSE ratio.
+The linear predictor approach's performance is also very close the standard approach.
+Both Verweij and Van Houwelingen's approach and the proposed deviance residual approach yield models that have greater MSE, hence are not recommended for use in practice.
 
@@
 Optimal &0.016 (0.009)&&/&&0.042 (0.003)&&/&&0.026 (0.001)&&/&&0.026 (0.001)&&/\\          
-Grouped &0.035 (0.007)&&2.597&&0.073 (0.009)&&2.184&&0.053 (0.003)&&5.627&&0.123 (0.036)&&3.931\\
-Ungrouped &0.030 (0.009)&&2.246&&0.064 (0.01)&&1.815&&0.040 (0.003)&&3.131&&0.070 (0.016)&&2.560\\
+V $\&$ VH &0.035 (0.007)&&2.597&&0.073 (0.009)&&2.184&&0.053 (0.003)&&5.627&&0.123 (0.036)&&3.931\\
+Standard &0.030 (0.009)&&2.246&&0.064 (0.01)&&1.815&&0.040 (0.003)&&3.131&&0.070 (0.016)&&2.560\\
 LinearPred &0.034 (0.007)&&2.559&&0.066 (0.008)&&1.880&&0.041 (0.003)&&3.350&&0.074 (0.016)&&2.713\\
@@
 The results are shown in Table 1.
-While the ungrouped approach yields the smallest MSE and best predictive accuracy, the Linear Predictor and the Ungrouped approaches perform closely to each other across all settings.
-Grouped approach performs well in low-dimensions or with sufficient samples, while deviance residual approach performs better in high-dimensional settings.
+While the standard approach yields the smallest MSE and best predictive accuracy, the Linear Predictor and the standard approaches perform closely to each other across all settings.
+Verweij and Van Houwelingen's approach performs well in low-dimensions or with sufficient samples, while deviance residual approach performs better in high-dimensional settings.
 
@@
 
-\par In this section, simulations studies were conducted to address the stability issue of the ungrouped cross-validated likelihood.
-The ungrouped approach is an intuitive way to carry out cross-validation.
-It also has the best predictive accuarcy based on the simulation results in previous sections.
+\par In this section, simulations studies were conducted to address the stability issue of the standard cross-validated likelihood.
+The standard approach is an intuitive way to carry out cross-validation.
+It also has the best predictive accuracy based on the simulation results in previous sections.
 However, one big disadvantage of this approach is that it is unstable when number of observed events in some fold is really small.
 As is illustrated in Figure 3, when censoring increases or when number of folds increases, the model would fail to converge due to insufficient number of events in the test set.
-Hence, the ungrouped cross-validated likelihood should not be used when there is no sufficient number of events.
+Hence, the standard cross-validated likelihood should not be used when there is no sufficient number of events.
 
@@
 	\par In this section, Leave-One-Out Cross Validation (LOOCV) were studied, in comparison with 10-fold CV.
-When first introduced, the grouped cross-validated partial likelihood was proposed so that LOOCV for Cox regression became feasible.
-It is impossible to carry out LOO cross-validation using the ungrouped approach, because the partial likelihood cannot be built on the one left-out observation.
-While the grouped cross-validated partial likelihood solves the stability issue by aggregating more observations together, it is not an efficient way to evaluate the model fit.
+When first introduced, Verweij and Van Houwelingen's cross-validated partial likelihood was proposed so that LOOCV for Cox regression became feasible.
+It is impossible to carry out LOO cross-validation using the standard approach, because the partial likelihood cannot be built on the one left-out observation.
+While Verweij and Van Houwelingen's cross-validated partial likelihood solves the stability issue by aggregating more observations together, it is not an efficient way to evaluate the model fit.
 Simulation studies in Section 3.1 have illustrated that it tends to be conservative in variable selections and selects smaller models in various scenarios.
-This fact can be even more illustrated via the following simulation studies, where both 10-fold and LOOCV were conducted via grouped, linear predictor and deviance residual approaches and 10 fold CV was conducted for the ungrouped approach.
+This fact can be even more illustrated via the following simulation studies, where both 10-fold and LOOCV were conducted via Verweij and Van Houwelingen's approach, linear predictor and deviance residual approaches and 10 fold CV was conducted for the standard approach.
 	\par As is illustrated in Figure 4, LOOCV selected models that have smaller MSE than 10-fold CV for all three CV approaches.
-As is shown in left and middle panels, LOOCV based on deviance residuals and grouped cross-validated partial likelihood still perform more conservatively than the ungrouped approach.
-Even if the grouped approach was proposed to allow LOOCV, the LOOCV based on the grouped approach does not surpass the 10-fold ungrouped approach.
-On the other hand, the Linear Predictors also allows LOOCV, and the LOOCV based on Linear Predictors performs better than the ungrouped approach.
+As is shown in left and middle panels, LOOCV based on deviance residuals and Verweij and Van Houwelingen's cross-validated partial likelihood still perform more conservatively than the standard approach.
+Even if Verweij and Van Houwelingen's approach was proposed to allow LOOCV, the LOOCV based on Verweij and Van Houwelingen's approach does not surpass the 10-fold standard approach.
+On the other hand, the Linear Predictors also allows LOOCV, and the LOOCV based on Linear Predictors performs better than the standard approach.
 
@@
 & $\lambda$ & log($\lambda$) &  &$\lambda$  &  log($\lambda$)   \\ \cline{2-3} \cline{5-6} 
-Grouped &0.1226 & -2.0985 && 0.1215 & -2.1079\\
-Ungrouped & 0.0803 & -2.5222 && 0.0983 & -2.3197\\
+V $\&$ VH &0.1226 & -2.0985 && 0.1215 & -2.1079\\
+standard & 0.0803 & -2.5222 && 0.0983 & -2.3197\\
 Linear Pred &0.0828 & -2.4919 && 0.1013 & -2.2894\\
@@
 \par How the CV methods perform on real data reflects what is shown in simulation studies.
-The ungrouped CVL is the most liberal approach.
-Deviance Residual and Grouped CVL approaches were both really conversative, while the ungrouped approach was the second liberal to ungrouped CVL.
+The standard CVL is the most liberal approach.
+Deviance Residual and Verweij and Van Houwelingen's CVL approaches were both really conversative, while the standard approach was the second liberal to standard CVL.
 In Figure 5, the Cross Validated Error is rescaled and plotted for all four methods.
@@
 Hence this approach is better at picking out signals than the other three approaches.
-For the grouped cross-validated partial likelihood, there's almost no signal there.
+For Verweij and Van Houwelingen's cross-validated partial likelihood, there's almost no signal there.
 
@@
 All cross-validation methods tend to select the more conservative model: with greater value of $\lambda$, and fewer variables selected into the model.
-Despite of the stability issue, the ungrouped approach is the most liberal one across all scenarios and yield the least MSE.
-The proposed linear predictor approach performs closely to the ungrouped approach.
-Under LOOCV, the linear predictor approach out-performs the ungrouped approach.
-Both the grouped approach and the proposed Deviance Residual approach perform conservatively, hence not recommended for use in practice.
+Despite of the stability issue, the standard approach is the most liberal one across all scenarios and yield the least MSE.
+The proposed linear predictor approach performs closely to the standard approach.
+Under LOOCV, the linear predictor approach out-performs the standardstandard approach.
+Both Verweij and Van Houwelingen's approach and the proposed Deviance Residual approach perform conservatively, hence not recommended for use in practice.
 Results from real data application reflects results from the simulation studies.

### 2018-05-17 main.tex
Edits to introduction

@@
 \section{Introduction}
-Cox propotional hazard regression, since its original proposal in 1972, is by far the most common regression approach for analyzing survival data.
-Cox regression is built upon partial likelihood.
-Under the propotional hazard assumption, partial likelihood still preserves most properties of a full likelihood, while the estimation of the baseline can be skipped.
-Despite its popularity, Cox model is limited by this semiparametric nature: it is challenging to evaluate the model's predictive accuracy.
+Since its original proposal \citep{Cox1972}, Cox proportional hazards regression has become the most common regression approach for analyzing survival data.
+Cox regression utilizes a partial likelihood construction under an assumption of proportional hazards to estimate the regression coefficients without having to specify the underlying baseline hazard.
+The ability to avoid choosing a specific parametric distribution for the survival time is very attractive, as time-to-event data is often poorly described by fully parametric models, but this semiparametric approach comes at a cost.
 
-Suppose there are n subjects and m observed failure events.
-Use i for indexing observations, j for indexing events sorted by their failure time.
-For the i th subject, its covarites will be denoted by $X_{i}$; $\delta_{i}$ denotes whether an event is observed or not for subject i.
-$t_{i}$ denotes the time observed for subject i; $t_{j}$ denotes the failure time for the j th event.
-In the Cox model \citep{Cox1975}, the hazard function for an observation i is given by
-\begin{equation}h_{i}(t) = h_{0}(t) \text{exp}( X_{i}^{T} \beta)\end{equation} 
-where $h_{0}$ is the baseline hazard and $e^{X_{i}^{T} \beta}$ is the relative risk.
-The estimation of the coefficients $\beta$s is obtained by maximizing the partial likelihood.
-For each individual subject, its contribution to the partial likelihood is
-\begin{equation} l_{i}(\beta) = \left \{\frac{\text{exp} ( X_{i}^{T} \beta)}{\sum_{ k \in R(t_{i})}\text{exp} ( X_{k}^{T} \beta)}\right \}^{\delta_{i}} \end{equation} \note{introduce notation $l_{i}$} , where $R(t_{i})$ denotes the set of observations at risk at time $t_{i}$.
+The Cox regression approach can estimate relative risks, but without estimating the baseline hazard, it cannot make any predictions concerning the absolute failure time for any given individual.
+This poses a challenge to assessing the predictive accuracy of a given Cox regression model.
+The challenge is particularly relevant for penalized Cox regression \citep{Tibshirani1997,Fan2002}, as the assessment of predictive accuracy via cross-validation is the standard method for selecting the regularization parameter and deciding upon a model.
 
-For all n subjects, the partial likelihood is constructed as a product of all individual contributed partial likelihood:
-\begin{equation} L(\beta) =\prod_{i = 1}^{n} l_{i}(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation}.
-As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard.
+Standard cross-validation involves of dividing the data into $K$ folds, then fitting the model on $K-1$ of those folds (the training set) and assessing prediction accuracy on the remaining fold (the testing set).
+For Cox regression, the estimated coefficients allow us to quantify the risk for each subject in the training set relative to other members of the training set, but it is not obvious how to use those estimates to quantify the model's accuracy in the testing set.
+
+One approach is to simply calculate the partial likelihood based on the observations in the test set as a measure for the model's predictive accuracy.
+In doing so, we calculate the risk for each member of the test set relative to other members of the test set.
+One drawback to this approach is that it becomes unstable when the size of the the test set is small.
+In particular, it cannot be applied to leave-one-out cross-validation (LOOCV), as we must have at least two observations in the test set to compare their risk relative to each other.
+
+%In this paper, we briefly review existing approaches for meeting this challenge, then propose two new methods for carrying out cross-validation in Cox regression models.
+
+%As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard.
 Hence the interpretation of the Cox model is only valid in a relative sense.
 
-Penalized Cox Regression is an extension of the regular cox regression model.
-Estimates of the $\beta$ coefficients are obtained by minimizing the objective function \begin{equation}  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{m} \text{log} \left \{\frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta).
-\end{equation} \citep{Simon2011}.
+% NOTE: I don't know that we need this paragraph; cross-validation is pretty common.
+%One common approach for selecting the tuning parameter $\lambda$ for linear regression and logistic regression is via the K-fold cross validation.
+The data set would be split into K folds.
+One fold would be treated as the test set and the other K - 1 folds as the training set.
+The model would first be built on the training set, then fitted to the test set to obtain cross validation error (CVE).
+CVE would be calculated for each candidate $\lambda$, then the one that minimizes cross validated error would be selected.
+
+To overcome this drawback, an alternative approach was proposed by \citet{Verweij1993}.
+Their approach, which we describe in detail in Section~\ref{Sec:cox-cv-existing}, stabilizes the cross-validated log likelihood, enabling its use even when the number of subjects in each fold is small.
+This approach has been widely used and is implemented in R packages such as {\tt glmnet} \citep{glmnet}.
+Although this approach fixes the stability issue, we demonstrate here that in practice, it tends to behave conservatively in terms of model selection.
+
+%We hope to improve cross-validation procedure for Cox regression to achieve a good balance between stability and performance.
+
+In this paper, we propose two alternative ways to carry out cross validation.
+Instead of cross-validating over partial likelihood, we propose to cross validate over the linear predictors of the regression model and cross validate over the deviance residual \citep{Therneau1990}of the Cox model.
+We conducted simulation studies to compare how those methods perform in both low dimensional and high dimensional settings with LASSO penalty.
+We compared the two proposed methods with the existing cross-validated partial likelihood approaches.
+We showed that the linear predictor approach outperforms other cross validation approaches in various scenarios.
+We found out that using cross validation to select tuning parameters for LASSO penalized cox regression tends to be a conservative approach in general.
+Finally, we apply those methods to two high dimensional data sets from real studies with time-to-event outcomes.
+
+\section{Methods}
+
+\note{Move this paragraph to Section 2?} In the Cox model, the hazard function for subject $i$ is given by 
+\begin{equation}
+  h_{i}(t) = h_{0}(t) \text{exp}( X_{i}^{T} \beta),
+\end{equation} 
+where $h_{0}$ is the baseline hazard and $e^{X_i^{T} \beta}$ is the risk for subject $i$ with covariates $X_i$ relative to that baseline.
+The estimation of the coefficients is obtained by maximizing a partial likelihood.
+Letting $t_i$ denote the time on study for subject $i$ and $\delta_{i}$ indicate whether or not an event is observed for subject $i$, each subject's individual contribution to that partial likelihood is
+\begin{equation}
+  l_{i}(\beta) = \left \{\frac{\text{exp} ( X_{i}^{T} \beta)}{\sum_{ k \in R(t_{i})}\text{exp} ( X_{k}^{T} \beta)}\right \}^{\delta_{i}},
+\end{equation}
+where $R(t_{i})$ denotes the set of observations at risk at time $t_{i}$.  The partial likelihood for the entire sample of $n$ subjects is then given by:
+\begin{equation}
+  L(\beta) =\prod_{i = 1}^{n} l_{i}(\beta).
+\end{equation}
+%% \begin{equation}
+%%   L(\beta) =\prod_{i = 1}^{n} l_{i}(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}.
+%% \end{equation}
+
+Penalized Cox Regression is an extension of the regular cox regression model. Estimates of the $\beta$ coefficients are obtained by minimizing the objective function
+\begin{equation}
+  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{m} \text{log} \left \{\frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta).
+\end{equation}
 The first part of the objective function is based on the partial likelihood.
@@
 
-One common approach for selecting the tuning parameter $\lambda$ for linear regression and logistic regression is via the K-fold cross validation.
-The data set would be split into K folds.
-One fold would be treated as the test set and the other K - 1 folds as the training set.
-The model would first be built on the training set, then fitted to the test set to obtain cross validation error (CVE).
-CVE would be calculated for each candidate $\lambda$, then the one that minimizes cross validated error would be selected.
+\subsection{Cross Validated Likelihood} 
+\label{Sec:cox-cv-existing}
 
-However, for penalized cox regression, it is challenging to carry out cross validation.
-Since the partial likelihood of the cox model does not give estimates for the baseline, it is not meaningful to compare the coefficients if they are not estimated from different models and different observations.
-Only if the baseline is the same, then the coefficients can be interpreted as relative risks.
-This is against the idea of cross validation, which splits the data and builds models on different subsets.
-
-%\begin{figure}
- %   \centering
-	%	\includegraphics[height= 4cm ]{./figures/cv.png}	
-    %\caption{K-fold Cross Validation}
-%\end{figure}
-
-%ZZZZZZZZZZZZZ
-\par Currently, there are two common cross-validation approaches for Cox regression, both having its disadvantages.
-The most intuitive approach to conduct cross validation in Cox regression is i) fit the model to training set then ii) calculate cross-validated partial likelihood based on the observations from the test set as a measure for the model's predictive accuracy.
-As will be discussed in Section 2, this approach is unstable when number of events gets small.
-The second approach was proposed by Verweij et al.
-\citep{Verweij1993}.
-This is a stabilized version of cross-validated log likelihood, which has been widely used and implemented in R packages such as $\tt{glmnet}$ \citep{glmnet}.
-Although this approach fixes the stability issue, it tends to yield more conservative models in practice.
-We hope to improve cross-validation procedure for Cox regression to achieve a good balance between stability and performance.
-
-\par In this paper, we proposed two alternative ways to carry out cross validation.
-Instead of cross-validating over partial likelihood, we propose to cross validate over the linear predictors of the regression model and cross validate over the deviance residual \citep{Therneau1990}of the Cox model.
-We conducted simulation studies to compare how those methods perform in both low dimensional and high dimensional settings with LASSO penalty.
-We compared the two proposed methods with the existing cross-validated partial likelihood approaches.
-We showed that the linear predictor approach outperforms other cross validation approaches in various scenarios.
-We found out that using cross validation to select tuning parameters for LASSO penalized cox regression tends to be a conservative approach in general.
-Finally, we apply those methods to two high dimensional data sets from real studies with time-to-event outcomes.
-
-\section{Methods}
-
-  \subsection{Cross Validated Likelihood} 
 Suppose a data set of n observations is split into K folds and the ith fold is left out, then a partial likelihood can be built over the K -1 folds and yields maximum likelihood estimates of the coefficients, denoted by $\hat{\beta}_{-i}$.
@@
 
-\par An alternative way to calculate cross-validated log likelihood was proposed by Verweij and Van Houwelingen.
+An alternative way to calculate cross-validated log likelihood was proposed by Verweij and Van Houwelingen.
 They defined cross-validated log likelihood of leaving the ith fold out to be \begin{equation}cvl = \sum_{i=1}^{K} \{l(\hat{\beta}_{-i}) - l_{-i}(\hat{\beta}_{-i})\}.

### 2018-05-26 main.tex
Removed note
Edit to final paragraph of introduction

@@
 
-%We hope to improve cross-validation procedure for Cox regression to achieve a good balance between stability and performance.
-
-In this paper, we propose two alternative ways to carry out cross validation.
-Instead of cross-validating over partial likelihood, we propose to cross validate over the linear predictors of the regression model and cross validate over the deviance residual \citep{Therneau1990}of the Cox model.
-We conducted simulation studies to compare how those methods perform in both low dimensional and high dimensional settings with LASSO penalty.
-We compared the two proposed methods with the existing cross-validated partial likelihood approaches.
-We showed that the linear predictor approach outperforms other cross validation approaches in various scenarios.
-We found out that using cross validation to select tuning parameters for LASSO penalized cox regression tends to be a conservative approach in general.
-Finally, we apply those methods to two high dimensional data sets from real studies with time-to-event outcomes.
+In this paper, we propose two alternative ways to carry out cross validation for Cox regression.
+Instead of cross-validating over partial likelihood, we propose cross validation over either the linear predictors of the regression model or over the deviance residuals \citep{Therneau1990}.
+Through simulation studies, we compare these proposed methods with existing cross-validation approaches for LASSO penalized Cox regression in both low- and high-dimensional settings.
+We find that all cross validation approaches tend to be conservative, but that the linear predictor approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.
+We conclude by applying both proposed and existing approaches to two high dimensional data sets from real studies with time-to-event outcomes.
 
@@
 
-\note{Move this paragraph to Section 2?} In the Cox model, the hazard function for subject $i$ is given by 
+In the Cox model, the hazard function for subject $i$ is given by 
 \begin{equation}

### 2018-05-26 main.tex
Edits to Ch. 2 intro

@@
 \begin{equation}
-  h_{i}(t) = h_{0}(t) \text{exp}( X_{i}^{T} \beta),
+  h_{i}(t) = h_{0}(t) \exp( X_{i}^{T} \beta),
 \end{equation} 
-where $h_{0}$ is the baseline hazard and $e^{X_i^{T} \beta}$ is the risk for subject $i$ with covariates $X_i$ relative to that baseline.
+where $h_{0}$ is the baseline hazard and $e^{X_i^{T} \beta}$ is the risk for a subject with covariates $X_i$ relative to that baseline.
 The estimation of the coefficients is obtained by maximizing a partial likelihood.
@@
 \begin{equation}
-  l_{i}(\beta) = \left \{\frac{\text{exp} ( X_{i}^{T} \beta)}{\sum_{ k \in R(t_{i})}\text{exp} ( X_{k}^{T} \beta)}\right \}^{\delta_{i}},
+  l_{i}(\beta) = \left \{\frac{\exp ( X_{i}^{T} \beta)}{\sum_{ k \in R(t_{i})}\exp ( X_{k}^{T} \beta)}\right \}^{\delta_{i}},
 \end{equation}
-where $R(t_{i})$ denotes the set of observations at risk at time $t_{i}$.  The partial likelihood for the entire sample of $n$ subjects is then given by:
+where $R(t_{i})$ denotes the set of subjects at risk at time $t_{i}$.  The partial likelihood for the entire sample of $n$ subjects is then given by:
 \begin{equation}
@@
 \end{equation}
-%% \begin{equation}
-%%   L(\beta) =\prod_{i = 1}^{n} l_{i}(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}.
-%% \end{equation}
 
-Penalized Cox Regression is an extension of the regular cox regression model. Estimates of the $\beta$ coefficients are obtained by minimizing the objective function
+Cox regression can be extended by introducing a penalty into the partial likelihood.  In penalized Cox regression, $\beta$ coefficient estimates are obtained by minimizing the objective function
 \begin{equation}
-  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{m} \text{log} \left \{\frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta).
+  Q(\beta) = - \frac{1}{n} \log L(\beta) + P_{\lambda}(\beta),
 \end{equation}
-The first part of the objective function is based on the partial likelihood.
-The second part is a penalty term that depends on some regularization parameters.
-In this paper, only the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$ was considered.
-The LASSO cox regression can be used to build prediction models for survival outcome based on high dimensional data, where the number of covariates p $\gg$ n.
-An example of its application would be to make predictions of cancer patients' the overall survival time with their certain gene expressions.
+where $P_{\lambda}(\beta)$ is a penalty function that depends on a regularization parameter $\lam$.
+In this paper, we focus upon the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$, although the methods we propose can be used with any penalty (indeed, can also be used for model selection in unpenalized Cox regression).
+LASSO cox regression is particularly useful for high dimensional data where the number of covariates p $\gg$ n; for example, in the prediction of cancer patients' overall survival based on genome-wide expression measurements.
 
-Selecting the tuning parameter $\lambda$ is crucial to getting LASSO estimations for $\beta$.
-The LASSO penalty enforces sparcity to the regression model, so that it carries out an automatic variable selection along the solution path.
-As $\lambda$ increases, more coefficients would be set to 0 and fewer variables will be selected by the model.
-When $\lambda$ is too small, the model tends to be redundant.
-When $\lambda$ is too large, the model would be too conservative to select any variables.
-In either case, model's predictive accuracy would be affected.
+Selecting $\lambda$ is critical to LASSO estimation.
+LASSO estimates are sparse, in the sense that some coefficients are estimated to be exactly zero, thereby carrying out automatic variable selection along the solution path.
+At large values of $\lambda$, most or all of the coefficients are 0, but as $\lam$ decreases, more covariates are selected into the model.
+If $\lam$ is too large or too small, the model's predictive accuracy suffers.
 

### 2018-05-26 main.tex
Some edits to 2.1, but we have problems

@@
 In the Cox model, the hazard function for subject $i$ is given by 
-\begin{equation}
+\begin{equation*}
   h_{i}(t) = h_{0}(t) \exp( X_{i}^{T} \beta),
-\end{equation} 
+\end{equation*} 
 where $h_{0}$ is the baseline hazard and $e^{X_i^{T} \beta}$ is the risk for a subject with covariates $X_i$ relative to that baseline.
@@
 Letting $t_i$ denote the time on study for subject $i$ and $\delta_{i}$ indicate whether or not an event is observed for subject $i$, each subject's individual contribution to that partial likelihood is
-\begin{equation}
-  l_{i}(\beta) = \left \{\frac{\exp ( X_{i}^{T} \beta)}{\sum_{ k \in R(t_{i})}\exp ( X_{k}^{T} \beta)}\right \}^{\delta_{i}},
-\end{equation}
+\begin{equation*}
+  L_{i}(\beta) = \left \{\frac{\exp ( X_{i}^{T} \beta)}{\sum_{ k \in R(t_{i})}\exp ( X_{k}^{T} \beta)}\right \}^{\delta_{i}},
+\end{equation*}
 where $R(t_{i})$ denotes the set of subjects at risk at time $t_{i}$.  The partial likelihood for the entire sample of $n$ subjects is then given by:
-\begin{equation}
-  L(\beta) =\prod_{i = 1}^{n} l_{i}(\beta).
-\end{equation}
+\begin{equation*}
+  L(\beta) =\prod_{i = 1}^{n} L_{i}(\beta).
+\end{equation*}
 
@@
 \begin{equation}
+  \label{eq:obj}
   Q(\beta) = - \frac{1}{n} \log L(\beta) + P_{\lambda}(\beta),
@@
 
-\subsection{Cross Validated Likelihood} 
+\subsection{Cross validated partial likelihood} 
 \label{Sec:cox-cv-existing}
 
-Suppose a data set of n observations is split into K folds and the ith fold is left out, then a partial likelihood can be built over the K -1 folds and yields maximum likelihood estimates of the coefficients, denoted by $\hat{\beta}_{-i}$.
-An intuitive way to carry out cross validation is to use the partial likelihood on the test data as the cross validated error: \begin{equation}cvl = \sum_{i=1}^{K} l_{i}(\hat{\beta}_{-i})\end{equation} This is implemented in the $\tt{glmnet}$ package as the $\tt{ungrouped}$ option.
-We refer to this definition as standard cross-validated partial likelihood.
+Suppose a data set of n observations is split into $K$ folds and the $\ith$ fold is left out.
+A partial likelihood can then be built over the remaining $K-1$ folds and used in \eqref{eq:obj} to obtain penalized likelihood estimates, which we denote $\hat{\beta}_{-i}$.
+This manuscript is concerned with the question: how should one use the $\ith$ fold to measure predictive accuracy?
+
+A straightforward approach, which we refer to as \emph{standard cross-validated partial likelihood}, is to use the partial likelihood on the test data as the cross validated error (CVE):
+\note{This section here needs work.
+The equation below is wrong.
+In it, i is used to represent the folds (1 through K) even though $l_i$ has already been defined to represent individuals -- in general, each fold has multiple individuals; only for LOOCV would these be the same.
+We need to introduce new notation here, or a double sum, or something.}
+\begin{equation}
+  \CVE = \sum_{i=1}^{K} l_{i}(\hat{\beta}_{-i}),
+\end{equation}
+where $l_i = \log L_i$.
+This is implemented in the $\tt{glmnet}$ package as the $\tt{ungrouped}$ option.
+\note{Also, I think we should write everything in terms of CVE, not CVL}.
 

### 2018-07-23 main.tex
Change notations in section 2

@@
 
+%%%%%%%%%%%%%
+%%%%%%%%%%%%%
 \section{Methods}
+%%%%%%%%%%%%%
+%%%%%%%%%%%%%
 
@@
 Letting $t_i$ denote the time on study for subject $i$ and $\delta_{i}$ indicate whether or not an event is observed for subject $i$, each subject's individual contribution to that partial likelihood is
+
 \begin{equation*}
@@
 \end{equation*}
+
 where $R(t_{i})$ denotes the set of subjects at risk at time $t_{i}$.  The partial likelihood for the entire sample of $n$ subjects is then given by:
+
 \begin{equation*}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%
+\note{I'm introducing the subgroup notation here, so that we can directly use the subgroup superscript in 2.1}
+
+Since cross-validation methods in general involve partitioning the entire sample into subgroups, superscript G can be used to denote the partial likelihood built for sujects from group G, which is a subset of the entire sample:
+
+\begin{equation*}
+  L^{G}(\beta) =\prod_{i \in G} L^{G}_{i}(\beta).
+\end{equation*}
+
 Cox regression can be extended by introducing a penalty into the partial likelihood.  In penalized Cox regression, $\beta$ coefficient estimates are obtained by minimizing the objective function
+
 \begin{equation}
@@
 \end{equation}
+
 where $P_{\lambda}(\beta)$ is a penalty function that depends on a regularization parameter $\lam$.
@@
 
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Cross validated partial likelihood} 
 \label{Sec:cox-cv-existing}
+%%%%%%%%%%%%%%%%%%%%
 
-Suppose a data set of n observations is split into $K$ folds and the $\ith$ fold is left out.
-A partial likelihood can then be built over the remaining $K-1$ folds and used in \eqref{eq:obj} to obtain penalized likelihood estimates, which we denote $\hat{\beta}_{-i}$.
-This manuscript is concerned with the question: how should one use the $\ith$ fold to measure predictive accuracy?
+Suppose a data set of n observations is split into $K$ folds and the $kth$ fold is left out.
+A partial likelihood $L^{-k}(\beta)$ can then be built over the remaining $K-1$ folds and used in \eqref{eq:obj} to obtain penalized likelihood estimates, which we denote $\hat{\beta}^{-k}$.
+This manuscript is concerned with the question: how should one use the $kth$ fold to measure predictive accuracy?
 
-A straightforward approach, which we refer to as \emph{standard cross-validated partial likelihood}, is to use the partial likelihood on the test data as the cross validated error (CVE):
-\note{This section here needs work.
+A straightforward approach, which we refer to as \emph{standard cross-validated partial likelihood}, is to use negative of the partial likelihood evaluated at $\hat{\beta}^{-k}$ on the test data as the cross validated error (CVE):
+
+%\note{This section here needs work.
 The equation below is wrong.
@@
 We need to introduce new notation here, or a double sum, or something.}
+
 \begin{equation}
-  \CVE = \sum_{i=1}^{K} l_{i}(\hat{\beta}_{-i}),
+  \CVE = - \sum_{k=1}^{K} l^{k}(\hat{\beta}^{-k}),
 \end{equation}
-where $l_i = \log L_i$.
+where $l^{k}(\hat{\beta}^{-k}) = \log L^{k}(\hat{\beta}^{-k}) = \sum \limits_{i \in fold k} log L^{k}_{i}(\hat{\beta}^{-k})$ .
+Greater likelihood usually represents a better fit and better predictive accuracy.
+Hence, the negative value of the partial likelihood can be used as the cross-validated error that should be minimized.
+Models that have better predictive accuracy should have a lower CVE.
 This is implemented in the $\tt{glmnet}$ package as the $\tt{ungrouped}$ option.
-\note{Also, I think we should write everything in terms of CVE, not CVL}.
 
-An alternative way to calculate cross-validated log likelihood was proposed by Verweij and Van Houwelingen.
-They defined cross-validated log likelihood of leaving the ith fold out to be \begin{equation}cvl = \sum_{i=1}^{K} \{l(\hat{\beta}_{-i}) - l_{-i}(\hat{\beta}_{-i})\}.
-\end{equation} When the ith fold is left out, $l(\hat{\beta}_{-i})$ is the log partial likelihood evaluated at $\hat{\beta}_{-i}$ with all k folds of observations.
-$l_{-i}(\hat{\beta}_{-i})$ is the log partial likelihood evaluated at $\hat{\beta}_{-i}$ with observations from the other K-1 folds.
-We refer to this definition as Verweij and Van Houwelingen's cross-validated partial likelihood.
+%\note{Also, I think we should write everything in terms of CVE, not CVL}.
+
+An alternative way to calculate cross-validated error was proposed by Verweij and Van Houwelingen. They defined cross-validated error of leaving the $kth$ fold out to be 
+\begin{equation}
+	CVE = -\sum_{k = 1}^K \left\{ l(\hat{\beta}^{- k})  - l^{-k}(\hat{\beta}^{- k}) \right\}. 
+\end{equation}
+When the $kth$ fold is left out, $l(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on the entire sample.
+$l^{-k}(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on samples from the other K-1 folds.
+We refer to this definition as Verweij and Van Houwelingen's cross-validated error.
+As is explained in the original paper from Verweij and Van Houwelingen, $l(\hat{\beta}^{- k})  - l^{-k}(\hat{\beta}^{- k})$ measures the contribution from the $kth$ fold to the likelihood.
+It measures how well observations from the $kth$ fold can be predicted from other observations.
 This is implemented in the $\tt{glmnet}$ package as the default $\tt{grouped}$ option.
 
+\note {The following paragraph will need to be changed}
 \par The upper panels in Figure 2 illustrates the ideas of the two methods.
@@
 
+%%%%%%%%%%%%%%%%%%%%
   \subsection{Cross Validated Linear Predictors}
+%%%%%%%%%%%%%%%%%%%%
 
 Besides cross validating over the partial likelihood, an alternative approach is to cross validate over the linear predictors.
-In the first step, the ith fold would be left out and the other K-1 folds would be used as the training set to get the estimates $\hat{\beta}_{-i}$.
-Then, the cross-validated linear predictors would be calculated based on the observations in the test set:  \begin{equation}\hat{\eta}_{-i} = X_{i}\hat{\beta}_{-i}.\end{equation} After repeating this for all K folds, a whole set of linear predictors  $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...
-\hat{\eta}_{-K})$ can be obtained.
-A partial likelihood can be built over this set of linear predictors: \begin{equation} L(\hat{\eta}_{-}) = \prod_{j=1}^{m} \frac{exp (\hat{\eta}_{-j})}{\sum_{ k \in R(t_{j})}exp (\hat{\eta}_{-k})}\end{equation}.
+Similar to the two approaches introduced in previous section, the $kth$ fold would be left out and the other K-1 folds would be used as the training set to get fit $\hat{\beta}^{-k}$.
+Then, a cross-validated linear predictor can be calculated for each observation i from the test set fold k:
+	\begin{equation}
+		\hat{\eta}^{-}_{i} = X_{i}\hat{\beta}^{-k} 
+	\end{equation} 
+After repeating this for all K folds, a set of linear predictors  $\hat{\eta}^{-} = ( \hat{\eta}^{-}_{1},  \hat{\eta}^{-}_{2} , ...
+\hat{\eta}^{-}_{n})$ can be obtained for the whole sample.
+A partial likelihood can be built over this set of linear predictors:
+	\begin{equation} 
+	L(\hat{\eta}^{-}) = \prod_{i=1}^{n} \frac{exp (\hat{\eta}^{-}_{i})}{\sum_{ j \in R(t_{i})}exp (\hat{\eta}^{-}_{j})}
+	\end{equation}.
 
-Cross-validated error is defined as the log of $L(\hat{\eta}_{-})$.
+Cross-validated error is defined as $$ CVE = - log L(\hat{\eta}_{-})$$. 
 This idea of cross validating over linear predictors is implemented in the package \texttt{ncvreg} \citep{ncvreg}.
@@
 But since they are building over different baselines, they are different in cox regression.
+
+\par \note {I saw a similar idea of using cross-validated linear predictor in a 2011 paper written by Richard Simon "Using cross-validation to evaluate predictive accuracy of survival risk classifiers based on high-dimensional data".
+The authors were using the same idea to compute time-dependent AUC.
+Shall we cite that paper here?
+-Biyue}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%
   \subsection{Cross Validated Deviance Residuals}
+%%%%%%%%%%%%%%%%%%%%
+
   %Need to update this section with the new baseline estimation method%
@@
 
-To estimate the baseline hazard, a set of linear predictors $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...
-\hat{\eta}_{-K})$ would be obtained in the same way as the cross validated linear predictor approach.
-$\hat{\eta}_{-i}$ is computed based on $\hat{\beta}_{-i}$s estimated from the training set and observations from the ith test set.
-With $\hat{\eta}_{-}$, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}.
-Then for each observations, the Martingale Residual can be calculated: \begin{equation}\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}_{i}},\end{equation} where $\delta_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation.
+To estimate the baseline hazard, a set of linear predictors $\hat{\eta}^{-} = ( \hat{\eta}^{-}_{1},  \hat{\eta}^{-}_{2} , ...
+\hat{\eta}^{-}_{n})$ would be obtained in the same way as the cross validated linear predictor approach.
+With $\hat{\eta}^{-}$, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}.
+Then for each observations, the Martingale Residual can be calculated:
+	\begin{equation}
+	\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}^{-}_{i}},
+	\end{equation} 
+where $\delta_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation.
 The deviance residual can be calcualted from the Martingale Residuals: \begin{equation} d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}\end{equation} The sum of squares of the deviance residuals, $\sum_{i}\hat{d}_{i}^2$, would be used as the cross validated error.
@@
 
+%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%
 \section{Simulation Studies}
+%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     \subsection {Simulations Comparing Cross-validation Approaches}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
  %need to change figures to include the new cross-validated deviance residuals

### 2018-08-21 main.tex
change diagram to pseudo code; add fold notation

@@
 \section{Introduction}
-Since its original proposal \citep{Cox1972}, Cox proportional hazards regression has become the most common regression approach for analyzing survival data.
+\par Since its original proposal \citep{Cox1972}, Cox proportional hazards regression has become the most common regression approach for analyzing survival data.
 Cox regression utilizes a partial likelihood construction under an assumption of proportional hazards to estimate the regression coefficients without having to specify the underlying baseline hazard.
@@
 
-The Cox regression approach can estimate relative risks, but without estimating the baseline hazard, it cannot make any predictions concerning the absolute failure time for any given individual.
+\par The Cox regression approach can estimate relative risks, but without estimating the baseline hazard, it cannot make any predictions concerning the absolute failure time for any given individual.
 This poses a challenge to assessing the predictive accuracy of a given Cox regression model.
@@
 
-Standard cross-validation involves of dividing the data into $K$ folds, then fitting the model on $K-1$ of those folds (the training set) and assessing prediction accuracy on the remaining fold (the testing set).
+\par Standard cross-validation involves of dividing the data into $K$ folds, then fitting the model on $K-1$ of those folds (the training set) and assessing prediction accuracy on the remaining fold (the testing set).
 For Cox regression, the estimated coefficients allow us to quantify the risk for each subject in the training set relative to other members of the training set, but it is not obvious how to use those estimates to quantify the model's accuracy in the testing set.
 
-One approach is to simply calculate the partial likelihood based on the observations in the test set as a measure for the model's predictive accuracy.
+\par One approach is to simply calculate the partial likelihood based on the observations in the test set as a measure for the model's predictive accuracy.
 In doing so, we calculate the risk for each member of the test set relative to other members of the test set.
@@
 
-To overcome this drawback, an alternative approach was proposed by \citet{Verweij1993}.
+\par To overcome this drawback, an alternative approach was proposed by \citet{Verweij1993}.
 Their approach, which we describe in detail in Section~\ref{Sec:cox-cv-existing}, stabilizes the cross-validated log likelihood, enabling its use even when the number of subjects in each fold is small.
@@
 
-In this paper, we propose two alternative ways to carry out cross validation for Cox regression.
+\par In this paper, we propose two alternative ways to carry out cross validation for Cox regression.
 Instead of cross-validating over partial likelihood, we propose cross validation over either the linear predictors of the regression model or over the deviance residuals \citep{Therneau1990}.
@@
 
-In the Cox model, the hazard function for subject $i$ is given by 
+\par In the Cox model, the hazard function for subject $i$ is given by 
 \begin{equation*}
@@
 Letting $t_i$ denote the time on study for subject $i$ and $\delta_{i}$ indicate whether or not an event is observed for subject $i$, each subject's individual contribution to that partial likelihood is
-
 \begin{equation*}
@@
 \end{equation*}
-
 where $R(t_{i})$ denotes the set of subjects at risk at time $t_{i}$.  The partial likelihood for the entire sample of $n$ subjects is then given by:
-
 \begin{equation*}
@@
 \end{equation*}
-
-%%%%%%%%%%%%%%%%%%%%%%%
-\note{I'm introducing the subgroup notation here, so that we can directly use the subgroup superscript in 2.1}
-
-Since cross-validation methods in general involve partitioning the entire sample into subgroups, superscript G can be used to denote the partial likelihood built for sujects from group G, which is a subset of the entire sample:
-
-\begin{equation*}
-  L^{G}(\beta) =\prod_{i \in G} L^{G}_{i}(\beta).
-\end{equation*}
-
 Cox regression can be extended by introducing a penalty into the partial likelihood.  In penalized Cox regression, $\beta$ coefficient estimates are obtained by minimizing the objective function
-
 \begin{equation}
@@
 \end{equation}
-
 where $P_{\lambda}(\beta)$ is a penalty function that depends on a regularization parameter $\lam$.
@@
 
-Selecting $\lambda$ is critical to LASSO estimation.
+\par Selecting $\lambda$ is critical to LASSO estimation.
 LASSO estimates are sparse, in the sense that some coefficients are estimated to be exactly zero, thereby carrying out automatic variable selection along the solution path.
@@
 If $\lam$ is too large or too small, the model's predictive accuracy suffers.
+
+%%%%%%%%%%%%%%%%%%%%%%%
+\note{Relating back to cross validation here}
+\par Suppose a data set $D$ of n observations is partitioned into $K$ folds: $D_{1}, D_{2}, ...
+D_{K}$.
+For a given $k \in \left\{1,2, ...
+, K\right\}$, set $T_{k} = D - D_{k}$ .
+Then, $T_{k}$ is considered as the training set and $D_{k}$ is considered as the test set.
+We use $L^{-k}(\beta)$ to denote the partial likelihood that is built over $T_k$, where set $D_{k}$ is left out.
+$L^{-k}(\beta)$ can be used in \eqref{eq:obj} to obtain penalized likelihood estimates, which we denote by $\hat{\beta}^{-k}$.
+This manuscript is concerned with the question: how should one use the $kth$ fold $D_{k}$ to measure predictive accuracy?
 
@@
 %%%%%%%%%%%%%%%%%%%%
-
-Suppose a data set of n observations is split into $K$ folds and the $kth$ fold is left out.
-A partial likelihood $L^{-k}(\beta)$ can then be built over the remaining $K-1$ folds and used in \eqref{eq:obj} to obtain penalized likelihood estimates, which we denote $\hat{\beta}^{-k}$.
-This manuscript is concerned with the question: how should one use the $kth$ fold to measure predictive accuracy?
-
-A straightforward approach, which we refer to as \emph{standard cross-validated partial likelihood}, is to use negative of the partial likelihood evaluated at $\hat{\beta}^{-k}$ on the test data as the cross validated error (CVE):
-
-%\note{This section here needs work.
-The equation below is wrong.
-In it, i is used to represent the folds (1 through K) even though $l_i$ has already been defined to represent individuals -- in general, each fold has multiple individuals; only for LOOCV would these be the same.
-We need to introduce new notation here, or a double sum, or something.}
-
+\par A straightforward approach, which we refer to as \emph{standard cross-validated partial likelihood}, is to use negative of the partial likelihood evaluated at $\hat{\beta}^{-k}$ on the test data as the cross validated error (CVE):
 \begin{equation}
@@
 \end{equation}
-where $l^{k}(\hat{\beta}^{-k}) = \log L^{k}(\hat{\beta}^{-k}) = \sum \limits_{i \in fold k} log L^{k}_{i}(\hat{\beta}^{-k})$ .
+where $l^{k}(\hat{\beta}^{-k}) = \log L^{k}(\hat{\beta}^{-k}) = \sum \limits_{i \in D_{k}} log L^{k}_{i}(\hat{\beta}^{-k})$ .
 Greater likelihood usually represents a better fit and better predictive accuracy.
@@
 This is implemented in the $\tt{glmnet}$ package as the $\tt{ungrouped}$ option.
-
-%\note{Also, I think we should write everything in terms of CVE, not CVL}.
 
@@
 
-\note {The following paragraph will need to be changed}
-\par The upper panels in Figure 2 illustrates the ideas of the two methods.
+\par Pseudo code Procedure 1 and 2 illustrates the ideas of the two methods in Figure 1.
 Through substraction, Verweij and Van Houwelingen's approach allows measuring individual observation's contribution to the partial likelihood, so that leave-one-out cross-validation is feasible.
-Since only the left-out fold would be used to construct the partial likelihood, the standard CVL is unstable in practice when there are only very few observations in some folder.
+Since only the left-out fold would be used to construct the partial likelihood, the standard CVE is unstable in practice when there are only very few observations in some folder.
 
@@
 	\end{equation} 
-where $\delta_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation.
-The deviance residual can be calcualted from the Martingale Residuals: \begin{equation} d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}\end{equation} The sum of squares of the deviance residuals, $\sum_{i}\hat{d}_{i}^2$, would be used as the cross validated error.
-An illustration of this method is drawn in the bottom right panel of Figure 1.
+where $\delta_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation. The deviance residual can be calcualted from the Martingale Residuals: 
+	\begin{equation} 
+	d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}.
+	\end{equation}
+ The sum of squares of the deviance residuals, $\sum_{i}\hat{d}_{i}^2$, would be used as the cross validated error. An illustration of this method is drawn in the bottom right panel of Figure 1.
 
@@
 
-\begin{figure}
+%\begin{figure}
+%  \centering
+%		\includegraphics[height= 9cm ]{./figures/figure_1.png}
+%		%\includegraphics[height= 4cm ]{./figures/02_2.png}
+%   \caption{Illustration of K-fold Cross Validation Methods for Cox Regression}
+%\end{figure}
+
+\begin{figure}[h]
 	\centering
-		\includegraphics[height= 9cm ]{./figures/figure_1.png}
-		%\includegraphics[height= 4cm ]{./figures/02_2.png}
-    \caption{Illustration of K-fold Cross Validation Methods for Cox Regression}
+	\begin{minipage}[t]{10cm}
+	\begin{algorithm}[H]
+	\caption{Standard Method}
+	\par \For{k in 1 to K } 
+	{
+		 Fit Cox Regression on $T_{k}$: Obtain model estimates $\hat{\beta}^{-k}$\;
+		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D_{k}$: $l^{k}(\hat{\beta}^{-k})$\;		
+	}
+
+	\par $ CVE_{st} = - \sum_{k = 1}^{K} l^{k}(\hat{\beta}^{-k})$
+ 	\end{algorithm}
+
+	\begin{algorithm}[H]
+  	 \caption{Verweij and Van Houwelingen's Method}
+	\par \For{k in 1 to K } 
+	{
+		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
+		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $T_{k}$: $l^{-k}(\hat{\beta}^{-k})$\;
+		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D$: $l(\hat{\beta}^{-k})$\;
+		 Take the difference: $ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) $\;
+	}
+
+	\par $ CVE_{vvh} = - \sum_{k = 1}^{K} \left\{ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) \right\}$
+	\end{algorithm}
+
+	\begin{algorithm}[H]
+    	\caption{Cross-Validated Linear Predictors}
+	\par \For{k in 1 to K } 
+	{
+		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
+		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
+	}
+
+	\par $ CVE_{lp} = - l(\hat{\eta}^{-})$
+	\end{algorithm}
+
+	\begin{algorithm}[H]
+   	\caption{Sums of Squared Deviance Residuals}
+	\par \For{k in 1 to K } 
+	{
+		Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
+		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
+	} 
+	Compute baseline hazard $\hat{\Lambda}_{0}$ based on $\hat{\eta}^{-}$\;
+	$\forall$ $ i$ subject $\in \left\{1,2,...,n\right\}$, compute:
+		\par\hspace{0.5cm} $\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}^{-}_{i}}$;
+		\par\hspace{0.5cm} $d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}$;
+	\par $ CVE_{dr} = \sum_{i = 1}^{n} {d}_{i}^2$ \;
+	\end{algorithm}
+	\end{minipage}
+	\caption{Pseudo Code for Cross-validation Procedures}
 	\end{figure}
@@
 
- %need to change figures to include the new cross-validated deviance residuals
    The simulation experiments were first conducted when the signal in the data was varied.
@@
 	\hline\\[-0.75em]
-       &     &        &MSE       &  &   &       & MSE     &    &    &        & MSE      &  &     &        & MSE      \\
-       & $\lambda$ (SD)     & &Ratio       &  & 	$\lambda$ (SD)         & & Ratio     & &	$\lambda$ (SD)         & & Ratio      & &	 $\lambda$ (SD)         & & Ratio      
+       	& & & MSE & & & & MSE & & & & MSE & & & & MSE \\
+       	& $\lambda$ (SD) & & Ratio & & $\lambda$ (SD) & & Ratio & &	$\lambda$ (SD) & & Ratio & & $\lambda$ (SD) & & Ratio      
 	\\ 
@@
 \par How the CV methods perform on real data reflects what is shown in simulation studies.
-The standard CVL is the most liberal approach.
-Deviance Residual and Verweij and Van Houwelingen's CVL approaches were both really conversative, while the standard approach was the second liberal to standard CVL.
+The standard CVE is the most liberal approach.
+Deviance Residual and Verweij and Van Houwelingen's CVE approaches were both really conversative, while the standard approach was the second liberal to standard CVE.
 In Figure 5, the Cross Validated Error is rescaled and plotted for all four methods.

### 2018-11-06 main.tex
update figures and tables with predictive metrics

@@
 where $P_{\lambda}(\beta)$ is a penalty function that depends on a regularization parameter $\lam$.
-In this paper, we focus upon the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$, although the methods we propose can be used with any penalty (indeed, can also be used for model selection in unpenalized Cox regression).
+
+\par In this paper, we focus upon the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$, although the methods we propose can be used with any penalty (indeed, can also be used for model selection in unpenalized Cox regression).
 LASSO cox regression is particularly useful for high dimensional data where the number of covariates p $\gg$ n; for example, in the prediction of cancer patients' overall survival based on genome-wide expression measurements.
+LASSO estimates yield a sparse model where some coefficients are estimated to be exactly zero, thereby carrying out automatic variable selection along the solution path.
+At large values of $\lambda$, most or all of the coefficients are 0, but as $\lam$ decreases, more covariates are selected into the model.
+Hence selecting $\lambda$ is critical to LASSO estimation in the sense that the model's accuracy suffers if $\lam$ is either too large or too small.
 
-\par Selecting $\lambda$ is critical to LASSO estimation.
-LASSO estimates are sparse, in the sense that some coefficients are estimated to be exactly zero, thereby carrying out automatic variable selection along the solution path.
-At large values of $\lambda$, most or all of the coefficients are 0, but as $\lam$ decreases, more covariates are selected into the model.
-If $\lam$ is too large or too small, the model's predictive accuracy suffers.
-
-%%%%%%%%%%%%%%%%%%%%%%%
-\note{Relating back to cross validation here}
-\par Suppose a data set $D$ of n observations is partitioned into $K$ folds: $D_{1}, D_{2}, ...
+\par Cross-validation is one of the most common methods used to select the appropriate $\lambda$ for penalized regression models.
+Suppose a data set $D$ of n observations is partitioned into $K$ folds: $D_{1}, D_{2}, ...
 D_{K}$.
@@
 $L^{-k}(\beta)$ can be used in \eqref{eq:obj} to obtain penalized likelihood estimates, which we denote by $\hat{\beta}^{-k}$.
-This manuscript is concerned with the question: how should one use the $kth$ fold $D_{k}$ to measure predictive accuracy?
+This manuscript is concerned with the question: how should one use the $kth$ fold $D_{k}$ to measure predictive accuracy based on partial likelihood?
 
@@
 
-An alternative way to calculate cross-validated error was proposed by Verweij and Van Houwelingen. They defined cross-validated error of leaving the $kth$ fold out to be 
+\par An alternative trick to calculate cross-validated error was proposed by Verweij and Van Houwelingen. They defined cross-validated error of leaving the $kth$ fold out to be 
 \begin{equation}
@@
 
-\par Pseudo code Procedure 1 and 2 illustrates the ideas of the two methods in Figure 1.
+\par The diagrams in the top two panels of Figure 1 illustrates the ideas of these two methods.
 Through substraction, Verweij and Van Houwelingen's approach allows measuring individual observation's contribution to the partial likelihood, so that leave-one-out cross-validation is feasible.
@@
 
-Besides cross validating over the partial likelihood, an alternative approach is to cross validate over the linear predictors.
+\par Besides cross validating over the partial likelihood, an alternative approach is to cross validate over the linear predictors.
 Similar to the two approaches introduced in previous section, the $kth$ fold would be left out and the other K-1 folds would be used as the training set to get fit $\hat{\beta}^{-k}$.
@@
 	\begin{equation}
-		\hat{\eta}^{-}_{i} = X_{i}\hat{\beta}^{-k} 
+		\hat{\eta}^{cv}_{i} = X_{i}\hat{\beta}^{-k} 
 	\end{equation} 
-After repeating this for all K folds, a set of linear predictors  $\hat{\eta}^{-} = ( \hat{\eta}^{-}_{1},  \hat{\eta}^{-}_{2} , ...
+After repeating this for all K folds, a set of linear predictors  $\hat{\eta}^{cv} = ( \hat{\eta}^{cv}_{1},  \hat{\eta}^{cv}_{2} , ...
 \hat{\eta}^{-}_{n})$ can be obtained for the whole sample.
@@
 	\begin{equation} 
-	L(\hat{\eta}^{-}) = \prod_{i=1}^{n} \frac{exp (\hat{\eta}^{-}_{i})}{\sum_{ j \in R(t_{i})}exp (\hat{\eta}^{-}_{j})}
+	L(\hat{\eta}^{-}) = \prod_{i=1}^{n} \frac{exp (\hat{\eta}^{cv}_{i})}{\sum_{ j \in R(t_{i})}exp (\hat{\eta}^{cv}_{j})}
 	\end{equation}.
 
-Cross-validated error is defined as $$ CVE = - log L(\hat{\eta}_{-})$$. 
+\par Cross-validated error is defined as $$ CVE = - log L(\hat{\eta}_{cv})$$. 
 This idea of cross validating over linear predictors is implemented in the package \texttt{ncvreg} \citep{ncvreg}.
@@
 
-%\begin{figure}
-%  \centering
-%		\includegraphics[height= 9cm ]{./figures/figure_1.png}
+\begin{figure}
+  \centering
+		\includegraphics[height= 12cm ]{./figures/figure_1_new.png}
 %		%\includegraphics[height= 4cm ]{./figures/02_2.png}
-%   \caption{Illustration of K-fold Cross Validation Methods for Cox Regression}
-%\end{figure}
-
-\begin{figure}[h]
-	\centering
-	\begin{minipage}[t]{10cm}
-	\begin{algorithm}[H]
-	\caption{Standard Method}
-	\par \For{k in 1 to K } 
-	{
-		 Fit Cox Regression on $T_{k}$: Obtain model estimates $\hat{\beta}^{-k}$\;
-		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D_{k}$: $l^{k}(\hat{\beta}^{-k})$\;		
-	}
-
-	\par $ CVE_{st} = - \sum_{k = 1}^{K} l^{k}(\hat{\beta}^{-k})$
- 	\end{algorithm}
-
-	\begin{algorithm}[H]
-  	 \caption{Verweij and Van Houwelingen's Method}
-	\par \For{k in 1 to K } 
-	{
-		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
-		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $T_{k}$: $l^{-k}(\hat{\beta}^{-k})$\;
-		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D$: $l(\hat{\beta}^{-k})$\;
-		 Take the difference: $ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) $\;
-	}
-
-	\par $ CVE_{vvh} = - \sum_{k = 1}^{K} \left\{ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) \right\}$
-	\end{algorithm}
-
-	\begin{algorithm}[H]
-    	\caption{Cross-Validated Linear Predictors}
-	\par \For{k in 1 to K } 
-	{
-		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
-		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
-	}
-
-	\par $ CVE_{lp} = - l(\hat{\eta}^{-})$
-	\end{algorithm}
-
-	\begin{algorithm}[H]
-   	\caption{Sums of Squared Deviance Residuals}
-	\par \For{k in 1 to K } 
-	{
-		Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
-		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
-	} 
-	Compute baseline hazard $\hat{\Lambda}_{0}$ based on $\hat{\eta}^{-}$\;
-	$\forall$ $ i$ subject $\in \left\{1,2,...,n\right\}$, compute:
-		\par\hspace{0.5cm} $\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}^{-}_{i}}$;
-		\par\hspace{0.5cm} $d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}$;
-	\par $ CVE_{dr} = \sum_{i = 1}^{n} {d}_{i}^2$ \;
-	\end{algorithm}
-	\end{minipage}
-	\caption{Pseudo Code for Cross-validation Procedures}
+   \caption{Illustration of K-fold Cross Validation Methods for Cox Regression}
 \end{figure}
@@
    The simulation experiments were first conducted when the signal in the data was varied.
-In each simulated data set, there are n = 100 observations and p = 1000 covariates.
+In each simulated data set, there are n = 150 observations and p = 1000 covariates.
 10$\%$ of the observations were censored.
 10-fold cross-validation was implemented for all four methods.
-Number of non-zero $\beta$s was 20.
-The signal in the data set was varied by varying the magnitude of the $\beta$s from 0.50 to 1.50.
+Number of non-zero $\beta$s was 10.
+The signal in the data set was varied by varying the magnitude of the $\beta$s from 0.40 to 0.9.
 For each scenario, 200 replications were used.
@@
     \centering
-		\includegraphics[height= 7cm ]{./figures/figure_2.png}
-    \caption{Simulation Studies Comparing 4 CV methods in terms of $\lambda$ and MSE ratio. n = 200, p = 1000, 10$\%$ censoring.}
+		\includegraphics[height= 7cm ]{./figures/figure_2_new.png}
+    \caption{Simulation Studies Comparing 4 CV methods in terms of $\lambda$ and MSE ratio. n = 150, p = 1000, 10$\%$ censoring.}
 \end{figure}
@@
 
-\begin{table}[h]
-	\small
+\begin{table}
+
+\setlength{\tabcolsep}{3pt}
+
+\caption{\label{tab:} Simulation Studies that Compare All Four Cross-Validation Methods Varying Dimension and Signal }
 \centering
-	\caption{Simulations Comparing all four Cross-Validation methods, Varying n and p}
-	\begin{tabular}{llllllllllllllll}
-	\hline\\[-0.75em]
-       	& \multicolumn{3}{l}{n = 100, p = 20} &  & \multicolumn{3}{l}{n = 100, p = 100} &  & \multicolumn{3}{l}{n = 500, p = 10000} &  & \multicolumn{3}{l}{n = 250, p = 10000} \\
-       	& \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 20} &  & \multicolumn{3}{l}{$\#$of non-zero = 20} \\ \cline{2-4} \cline{6-8} \cline{10-12} \cline{14-16}
-	\hline\\[-0.75em]
-       	& & & MSE & & & & MSE & & & & MSE & & & & MSE \\
-       	& $\lambda$ (SD) & & Ratio & & $\lambda$ (SD) & & Ratio & &	$\lambda$ (SD) & & Ratio & & $\lambda$ (SD) & & Ratio      
-	\\ 
-	\cline{2-2} \cline{4-4} \cline{6-6} \cline{8-8} \cline{10-10} \cline{12-12} \cline{14-14} \cline{16-16} 
-	\\[-0.75em]
-	Optimal &0.016 (0.009)&&/&&0.042 (0.003)&&/&&0.026 (0.001)&&/&&0.026 (0.001)&&/\\          
-	V $\&$ VH &0.035 (0.007)&&2.597&&0.073 (0.009)&&2.184&&0.053 (0.003)&&5.627&&0.123 (0.036)&&3.931\\
-	Standard &0.030 (0.009)&&2.246&&0.064 (0.01)&&1.815&&0.040 (0.003)&&3.131&&0.070 (0.016)&&2.560\\
-	LinearPred &0.034 (0.007)&&2.559&&0.066 (0.008)&&1.880&&0.041 (0.003)&&3.350&&0.074 (0.016)&&2.713\\
-	DevResid &0.100 (0.009)&&14.005&&0.108 (0.008)&&4.087&&0.076 (0.003)&&9.133&&0.111 (0.028)&&3.767\\ 
-	\hline
+\begin{tabular}[t]{cccccccccc}
+\toprule
+ n& p& $\#$ Non-Zero & Signal & Method & $\lambda$ & MSR Ratio &Brier Score & Kullback Leibler & C Index\\
+\midrule
+150 & 50 & 5 & Weak & V VH  & 0.095 (0.028) & 1.323 (0.443) & 0.242 (0.013) & 0.678 (0.029) & 0.617 (0.034) \\
+      &    &   &     & Standard  & 0.087 (0.030) & 1.323 (0.431) & 0.242 (0.013) & 0.678 (0.029) & 0.618 (0.032) \\
+      &    &   &     & LinearPred  & 0.088 (0.025) & 1.285 (0.419) & 0.241 (0.013) & 0.677 (0.028) & 0.620 (0.029) \\
+      &    &   &     & DevResid  & 0.132 (0.032) & 1.697 (0.716) & 0.248 (0.014) & 0.689 (0.030) & 0.601 (0.044) \\
+\addlinespace
+150 & 50 & 5 & Strong & V VH  & 0.070 (0.011) & 1.373 (0.362) & 0.177 (0.012) & 0.527 (0.030) & 0.750 (0.014) \\
+    &    &   &     & Standard  & 0.066 (0.012) & 1.313 (0.358) & 0.176 (0.012) & 0.526 (0.029) & 0.749 (0.014) \\
+    &    &   &     & LinearPred  & 0.067 (0.011) & 1.319 (0.335) & 0.176 (0.012) & 0.526 (0.030) & 0.750 (0.014) \\
+    &    &   &     & DevResid  & 0.110 (0.012) & 2.569 (0.826) & 0.187 (0.014) & 0.554 (0.033) & 0.750 (0.018) \\
+\addlinespace
+300 & 10000 & 20 & Weak & V VH  & 0.157 (0.021) & 1.079 (0.063) & 0.258 (0.008) & 0.711 (0.017) & 0.548 (0.037) \\
+       &    &   &     & Standard  & 0.137 (0.030) & 1.053 (0.053) & 0.255 (0.010) & 0.705 (0.021) & 0.557 (0.036) \\
+       &    &   &     & LinearPred  & 0.139 (0.029) & 1.054 (0.058) & 0.256 (0.009) & 0.705 (0.020) & 0.556 (0.036) \\
+       &    &   &     & DevResid  & 0.157 (0.024) & 1.078 (0.063) & 0.258 (0.008) & 0.710 (0.017) & 0.548 (0.037) \\
+\addlinespace
+ 500 & 10000 & 20 & Moderate & V VH  & 0.098 (0.012) & 1.298 (0.131) & 0.236 (0.014) & 0.665 (0.029) & 0.64 (0.036) \\
+        &    &   &     & Standard  & 0.080 (0.014) & 1.134 (0.11) & 0.226 (0.014) & 0.641 (0.031) & 0.651 (0.033) \\
+        &    &   &     & LinearPred  & 0.080 (0.012) & 1.139 (0.101) & 0.226 (0.014) & 0.642 (0.030) & 0.652 (0.031) \\
+        &    &   &     & DevResid  & 0.101 (0.015) & 1.317 (0.147) & 0.238 (0.014) & 0.667 (0.029) & 0.637 (0.040) \\
+\bottomrule
 \end{tabular}
@@
 
-    \subsection {Results of Leave-One-Out Cross Validation}
+    \subsection {Leave-One-Out Cross Validation}
 	\par In this section, Leave-One-Out Cross Validation (LOOCV) were studied, in comparison with 10-fold CV.
@@
 
-  \subsection {Simulations on Deviance Residuals and Baseline Hazard Estimations}
-\par While the Deviance Residual approach has a reasonable performance in scenarios where dimension is high, it tends to act more conservatively than all the other three approaches in general.
-As is noted by Therneau et al, the sums of squared martingale residuals and deviance residuals do not necessary reflect how good the model fit is \citep{Therneau2000modeling}.
-The difficulty of estimating the baseline hazard seems to play an important role.
-A series of simulations were thus designed to study how baseline hazard estimation affects the deviance residuals' performance as a measurement for model fit.
-
-\par In the following simulation studies, the defined optimal $\lambda$ and the $\lambda$ selected by the Linear Predictor approach were used as benchmarks.
-Various sample sizes and dimensions were varied.
-The survival outcomes were generated both from exponential distribution and Weibull distribution.
-Six different approaches were compared in terms of baseline hazard estimation and their impact on sums of squared deviance residuals.
-Two non-parametric approaches, Kalbfleisch and Prentices' estimator and Breslow's estimator, were used first for estimating the cumulative hazard baseline \citep{Kalbfleisch2011} \citep{breslow1972}.
-As is presented in column 3 and column 5 in Table 2, the sum of squared deviance residuals computed based on both approaches are all very conservative compared to the optimal $\lambda$.
-Their performance is extremely bad when n is large and p is small.
-
-\par Since both of the non-parameteric estimations seemed to perform well at early time points but poorly later, a weighted approach was then applied to emphasize earlier time points.
-Weights are assigned proportional to number of subjects at risk, making earlier time points weighted more than later time points.
-Results of these two approahces are presented in column 4 and 6 in Table 2.
-The attempt to improve the non-parameteric estimations via weighting does not seem to improve the selection too much.
-
-\par In the next step,  the non-parametric baseline hazard estimation was replaced by the exact baseline hazard that was used to generate the outcome.
-Results were listed in the seventh column in Table 2.
-When the baseline was known, the performance of sum of squared deviance residuals actually performed really close to the optimal $\lambda$.
-This simulation suggests that when the baseline hazard was accurately estimated, then the sum of squared deviance residuals gave really good measures for the model fit.
-
-\par In the last step, we replaced the non-parametric baseline hazard estimation with a parametric approach, where we assume the baseline hazard to follow an exponential distribution.
-We derived the estimation for baseline hazard by maximum likelihood approach conditioned on covariates: $\hat{\lambda}_{0} = \frac{\sum d_{i}}{\sum t_{i}e^{\eta_{i}}}$.
-The results are listed in the eighth column in Table 2.
-When this estimation was used for data that were generated from exponential distribution, its performance was as well as when the exact baseline was specified.
-When it was fitted to data that were generated from the Weibull distribution, that is when the model is misspecified, then its performance was even worse than the non-parametric approaches.
-
-\par Hence, getting accurate estimation of the baseline hazard is crucial to whether or not the sum of squared deviance residuals can be a good measurement for model fit.
-
-\begin{table}[h]
-	\small
-	\caption{Average of Selected $\lambda$ with Different Baseline Estimation for Deviance Residuals}
-	\centering
-	\begin{tabular}{lllllllll}
-	\hline
-	\\[-0.75em]
-    & Optimal & Linear Pred & KP & KP & Breslow & Breslow & Exact & Exp\\ 
-    &  & & & weighted & & weighted & & \\ \cline{2 - 9} 
-	\textbf{True Baseline: $Exponential(1)$} &&&&&&&&\\
-	n = 100, p = 10, non-zero $\beta$: 2 & 0.0461 & 0.0665 & 0.1986 & 0.1970 & 0.1908 & 0.1975 & 0.0540 & 0.0539 \\
-	n = 100, p = 100, non-zero $\beta$: 20 &0.0555 & 0.1841 & 0.2359 & 0.2201 & 0.2146 & 0.2244 & 0.0805 & 0.0985 \\
-	n = 100, p = 1000, non-zero $\beta$: 20 & 0.0268 & 0.0505 & 0.0948 & 0.0798 & 0.0867 & 0.0809 & 0.0288 & 0.0270 \\
-	\textbf{True Baseline:} $Weibull$ &&&&&&&& \\
- 	n = 100, p = 10, non-zero $\beta$: 2 & 0.0476 & 0.0710 & 0.2004 & 0.1969 & 0.1907 & 0.1907 & N/A & 0.2330 \\ 
-	\hline
-\end{tabular}
-\end{table}
+  %\subsection {Simulations on Deviance Residuals and Baseline Hazard Estimations}
 
@@
 \section{Discussion}
-Four methods of conducting cross-validation for Cox Regression were compared through simulation studies in terms of their predictive accuracy.
+
+\par Four methods of conducting cross-validation for Cox Regression were compared through simulation studies in terms of their predictive accuracy.
 All cross-validation methods tend to select the more conservative model: with greater value of $\lambda$, and fewer variables selected into the model.
@@
 Results from real data application reflects results from the simulation studies.
+
+% Discussion on Deviance Residual
+
+\par While the Deviance Residual approach has a reasonable performance in scenarios where dimension is high, it tends to act more conservatively than all the other three approaches in general.
+As is noted by Therneau et al, the sums of squared martingale residuals and deviance residuals do not necessary reflect how good the model fit is \citep{Therneau2000modeling}.
+The difficulty of estimating the baseline hazard seems to play an important role.
+A series of simulations were thus designed to study how baseline hazard estimation affects the deviance residuals' performance as a measurement for model fit.
+
+\par In the following simulation studies, the defined optimal $\lambda$ and the $\lambda$ selected by the Linear Predictor approach were used as benchmarks.
+Various sample sizes and dimensions were varied.
+The survival outcomes were generated both from exponential distribution and Weibull distribution.
+Six different approaches were compared in terms of baseline hazard estimation and their impact on sums of squared deviance residuals.
+Two non-parametric approaches, Kalbfleisch and Prentices' estimator and Breslow's estimator, were used first for estimating the cumulative hazard baseline \citep{Kalbfleisch2011} \citep{breslow1972}.
+As is presented in column 3 and column 5 in Table 2, the sum of squared deviance residuals computed based on both approaches are all very conservative compared to the optimal $\lambda$.
+Their performance is extremely bad when n is large and p is small.
+
+\par Since both of the non-parameteric estimations seemed to perform well at early time points but poorly later, a weighted approach was then applied to emphasize earlier time points.
+Weights are assigned proportional to number of subjects at risk, making earlier time points weighted more than later time points.
+Results of these two approahces are presented in column 4 and 6 in Table 2.
+The attempt to improve the non-parameteric estimations via weighting does not seem to improve the selection too much.
+
+\par In the next step,  the non-parametric baseline hazard estimation was replaced by the exact baseline hazard that was used to generate the outcome.
+Results were listed in the seventh column in Table 2.
+When the baseline was known, the performance of sum of squared deviance residuals actually performed really close to the optimal $\lambda$.
+This simulation suggests that when the baseline hazard was accurately estimated, then the sum of squared deviance residuals gave really good measures for the model fit.
+
+\par In the last step, we replaced the non-parametric baseline hazard estimation with a parametric approach, where we assume the baseline hazard to follow an exponential distribution.
+We derived the estimation for baseline hazard by maximum likelihood approach conditioned on covariates: $\hat{\lambda}_{0} = \frac{\sum d_{i}}{\sum t_{i}e^{\eta_{i}}}$.
+The results are listed in the eighth column in Table 2.
+When this estimation was used for data that were generated from exponential distribution, its performance was as well as when the exact baseline was specified.
+When it was fitted to data that were generated from the Weibull distribution, that is when the model is misspecified, then its performance was even worse than the non-parametric approaches.
+
+\par Hence, getting accurate estimation of the baseline hazard is crucial to whether or not the sum of squared deviance residuals can be a good measurement for model fit.
+
+\begin{table}[h]
+	\small
+	\caption{Average of Selected $\lambda$ with Different Baseline Estimation for Deviance Residuals}
+	\centering
+	\begin{tabular}{lllllllll}
+	\hline
+	\\[-0.75em]
+    & Optimal & Linear Pred & KP & KP & Breslow & Breslow & Exact & Exp\\ 
+    &  & & & weighted & & weighted & & \\ \cline{2 - 9} 
+	\textbf{True Baseline: $Exponential(1)$} &&&&&&&&\\
+	n = 100, p = 10, non-zero $\beta$: 2 & 0.0461 & 0.0665 & 0.1986 & 0.1970 & 0.1908 & 0.1975 & 0.0540 & 0.0539 \\
+	n = 100, p = 100, non-zero $\beta$: 20 &0.0555 & 0.1841 & 0.2359 & 0.2201 & 0.2146 & 0.2244 & 0.0805 & 0.0985 \\
+	n = 100, p = 1000, non-zero $\beta$: 20 & 0.0268 & 0.0505 & 0.0948 & 0.0798 & 0.0867 & 0.0809 & 0.0288 & 0.0270 \\
+	\textbf{True Baseline:} $Weibull$ &&&&&&&& \\
+ 	n = 100, p = 10, non-zero $\beta$: 2 & 0.0476 & 0.0710 & 0.2004 & 0.1969 & 0.1907 & 0.1907 & N/A & 0.2330 \\ 
+	\hline
+\end{tabular}
+\end{table}
+
+\section{Appendix}
+
+\begin{figure}[h]
+	\centering
+	\begin{minipage}[t]{10cm}
+	\begin{algorithm}[H]
+	\caption{Standard Method}
+	\par \For{k in 1 to K } 
+	{
+		 Fit Cox Regression on $T_{k}$: Obtain model estimates $\hat{\beta}^{-k}$\;
+		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D_{k}$: $l^{k}(\hat{\beta}^{-k})$\;		
+	}
+
+	\par $ CVE_{st} = - \sum_{k = 1}^{K} l^{k}(\hat{\beta}^{-k})$
+ 	\end{algorithm}
+
+	\begin{algorithm}[H]
+  	 \caption{Verweij and Van Houwelingen's Method}
+	\par \For{k in 1 to K } 
+	{
+		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
+		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $T_{k}$: $l^{-k}(\hat{\beta}^{-k})$\;
+		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D$: $l(\hat{\beta}^{-k})$\;
+		 Take the difference: $ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) $\;
+	}
+
+	\par $ CVE_{vvh} = - \sum_{k = 1}^{K} \left\{ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) \right\}$
+	\end{algorithm}
+
+	\begin{algorithm}[H]
+    	\caption{Cross-Validated Linear Predictors}
+	\par \For{k in 1 to K } 
+	{
+		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
+		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
+	}
+
+	\par $ CVE_{lp} = - l(\hat{\eta}^{-})$
+	\end{algorithm}
+
+	\begin{algorithm}[H]
+   	\caption{Sums of Squared Deviance Residuals}
+	\par \For{k in 1 to K } 
+	{
+		Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
+		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
+	} 
+	Compute baseline hazard $\hat{\Lambda}_{0}$ based on $\hat{\eta}^{-}$\;
+	$\forall$ $ i$ subject $\in \left\{1,2,...,n\right\}$, compute:
+		\par\hspace{0.5cm} $\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}^{-}_{i}}$;
+		\par\hspace{0.5cm} $d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}$;
+	\par $ CVE_{dr} = \sum_{i = 1}^{n} {d}_{i}^2$ \;
+	\end{algorithm}
+	\end{minipage}
+	\caption{Pseudo Code for Cross-validation Procedures}
+	\end{figure}

### 2018-11-07 main.tex
revised main.tex till section 3.2

@@
 For a given $k \in \left\{1,2, ...
-, K\right\}$, set $T_{k} = D - D_{k}$ .
-Then, $T_{k}$ is considered as the training set and $D_{k}$ is considered as the test set.
-We use $L^{-k}(\beta)$ to denote the partial likelihood that is built over $T_k$, where set $D_{k}$ is left out.
-$L^{-k}(\beta)$ can be used in \eqref{eq:obj} to obtain penalized likelihood estimates, which we denote by $\hat{\beta}^{-k}$.
+, K\right\}$, set $T_{k} = D - D_{k}$.
+Then, we consider $T_{k}$ as the training set and $D_{k}$ as the test set.
 This manuscript is concerned with the question: how should one use the $kth$ fold $D_{k}$ to measure predictive accuracy based on partial likelihood?
+In the following sections, we use $L^{-k}$ to denote the partial likelihood when it is built over data in $T_k$; $L^{k}$ when it is built over data in $D_k$; and $L$ when it is built over the entire data set $D$.
+The log partial likelihood is denoted by lower case $l$.
+For individual subject's contribution to the likelihood, we use $L_i$, where $i$ denotes the individual.
+Particularly of our interest, if $L^{-k}(\beta)$ is used in \eqref{eq:obj} to obtain penalized likelihood estimates using data in $T_k$, we use $\hat{\beta}^{-k}$ to denote the penalized estimates that are obtained.
 
@@
 %%%%%%%%%%%%%%%%%%%%
-\par A straightforward approach, which we refer to as \emph{standard cross-validated partial likelihood}, is to use negative of the partial likelihood evaluated at $\hat{\beta}^{-k}$ on the test data as the cross validated error (CVE):
+\par A straightforward approach to measure a model's predictive accuracy in $D_k$ is to evaluate the log partial likelihood at $\hat{\beta}^{-k}$ directly upon data in $D_k$.
+Greater likelihood usually represents a better fit and better predictive accuracy.
+Hence, the negative value of the partial likelihood can be used as the cross-validated error.
+We define the cross validated error (CVE) by pooling the negative log-likelihood from all K folds together:
 \begin{equation}
+  \label{eq:standard}
   \CVE = - \sum_{k=1}^{K} l^{k}(\hat{\beta}^{-k}),
 \end{equation}
-where $l^{k}(\hat{\beta}^{-k}) = \log L^{k}(\hat{\beta}^{-k}) = \sum \limits_{i \in D_{k}} log L^{k}_{i}(\hat{\beta}^{-k})$ .
-Greater likelihood usually represents a better fit and better predictive accuracy.
-Hence, the negative value of the partial likelihood can be used as the cross-validated error that should be minimized.
-Models that have better predictive accuracy should have a lower CVE.
+where $l^{k}(\hat{\beta}^{-k}) = \sum \limits_{i \in D_{k}} log L^{k}_{i}(\hat{\beta}^{-k})$.
+We refer this approach as \emph{standard cross-validated partial likelihood}.
 This is implemented in the $\tt{glmnet}$ package as the $\tt{ungrouped}$ option.
 
-\par An alternative trick to calculate cross-validated error was proposed by Verweij and Van Houwelingen. They defined cross-validated error of leaving the $kth$ fold out to be 
+\par The standard approach is the same trick that is used to conduct cross-validation for linear regression or logistic regression.
+One pitfall of this approach is that there may not be sufficient observations in $D_k$ to build up the risk set for partial likelihood.
+An alternative trick that arises to address this issue was proposed by Verweij and Van Houwelingen.
+They defined cross-validated error to be
 \begin{equation}
+\label{eq:VVH}
 	CVE = -\sum_{k = 1}^K \left\{ l(\hat{\beta}^{- k})  - l^{-k}(\hat{\beta}^{- k}) \right\}. 
 \end{equation}
-When the $kth$ fold is left out, $l(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on the entire sample.
-$l^{-k}(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on samples from the other K-1 folds.
-We refer to this definition as Verweij and Van Houwelingen's cross-validated error.
-As is explained in the original paper from Verweij and Van Houwelingen, $l(\hat{\beta}^{- k})  - l^{-k}(\hat{\beta}^{- k})$ measures the contribution from the $kth$ fold to the likelihood.
-It measures how well observations from the $kth$ fold can be predicted from other observations.
-This is implemented in the $\tt{glmnet}$ package as the default $\tt{grouped}$ option.
+$l(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on the entire data $D$ and $l^{-k}(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on the training set$T_k$.
+There should be enough observations in $T_k$ to build up the risk set as long as the entire data has a reasonable sample size.
 
-\par The diagrams in the top two panels of Figure 1 illustrates the ideas of these two methods.
-Through substraction, Verweij and Van Houwelingen's approach allows measuring individual observation's contribution to the partial likelihood, so that leave-one-out cross-validation is feasible.
-Since only the left-out fold would be used to construct the partial likelihood, the standard CVE is unstable in practice when there are only very few observations in some folder.
+\par We refer to this definition as \emph{Verweij and Van Houwelingen's cross-validated partial likelihood}.
+It was originally proposed to as a trick to conduct Leave One Out Cross-Validation for partial likelihood, which is impossible with the standard approach.
+The standard approach and Verweij and Van Houwelingen's approach both try to measure the contributions from $D_k$ to the partial likelihood.
+For many other likelihoods, say logistic regression or linear regression, these two quantities, $l(\hat{\beta}^{- k})  - l^{-k}(\hat{\beta}^{- k})$ and  $l^{k}(\hat{\beta}^{-k})$, are in fact equivalent to each other.
+Since its original proposal, Verweij and Van Houwelingen's cross-validated error has been widely implemented as a tool to select tuning parameters for predictive models, especially regression-based models.
+It is implemented in the $\tt{glmnet}$ package as the default $\tt{grouped}$ option for penalized Cox regression.
+Other R packages that used this trick include but not limited to $\tt{CoxRidge}$, $\tt{fastcox}$, $\tt{SGL}$, $\tt{CoxBoost}$, and $\tt{mboost}$.
+\note{not sure whether we should mention these packages here.
+citations will be added later}
 
@@
 
-\par Besides cross validating over the partial likelihood, an alternative approach is to cross validate over the linear predictors.
-Similar to the two approaches introduced in previous section, the $kth$ fold would be left out and the other K-1 folds would be used as the training set to get fit $\hat{\beta}^{-k}$.
-Then, a cross-validated linear predictor can be calculated for each observation i from the test set fold k:
+\par Instead of directly using partial likelihood to measure model's predictive accuracy over $D_k$, an alternative method is to obtain linear predictors for observations in $D_k$.
+We refer to this approach as \emph{cross-validated linear predictors}.
+To be more specific, we would first obtain $\hat{\beta}^{-k}$ from training set $T_{k}$; then the cross-validated linear predictors can be calculated for each observation $i$ from the test set $D_k$:
 	\begin{equation}
@@
 	\end{equation} 
-After repeating this for all K folds, a set of linear predictors  $\hat{\eta}^{cv} = ( \hat{\eta}^{cv}_{1},  \hat{\eta}^{cv}_{2} , ...
-\hat{\eta}^{-}_{n})$ can be obtained for the whole sample.
-A partial likelihood can be built over this set of linear predictors:
+After repeating this for all K folds, a complete set of cross-validated linear predictors $\hat{\eta}^{cv} = ( \hat{\eta}^{cv}_{1},  \hat{\eta}^{cv}_{2} , ...
+\hat{\eta}^{cv}_{n})$ for the whole sample is obtained.
+A partial likelihood can then be built over this set of linear predictors to evaluate the predictive accuracy of the model:
 	\begin{equation} 
-	L(\hat{\eta}^{-}) = \prod_{i=1}^{n} \frac{exp (\hat{\eta}^{cv}_{i})}{\sum_{ j \in R(t_{i})}exp (\hat{\eta}^{cv}_{j})}
+	L(\hat{\eta}^{cv}) = \prod_{i=1}^{n} \frac{exp (\hat{\eta}^{cv}_{i})}{\sum_{ j \in R(t_{i})}exp (\hat{\eta}^{cv}_{j})}
 	\end{equation}.
-
-\par Cross-validated error is defined as $$ CVE = - log L(\hat{\eta}_{cv})$$. 
-This idea of cross validating over linear predictors is implemented in the package \texttt{ncvreg} \citep{ncvreg}.
-For all the approaches introduced so far, they are all equivalent to each other in linear regressions.
-But since they are building over different baselines, they are different in cox regression.
-
-\par \note {I saw a similar idea of using cross-validated linear predictor in a 2011 paper written by Richard Simon "Using cross-validation to evaluate predictive accuracy of survival risk classifiers based on high-dimensional data".
-The authors were using the same idea to compute time-dependent AUC.
-Shall we cite that paper here?
--Biyue}
+We define the cross-validated error evaluated with cross-validated linear preditors to be $$ CVE = - log L(\hat{\eta}^{cv}).$$ 
+\par This idea of obtaining linear predictors over test sets, then constructing the partial likelihood after pooling all linear predictors together is implemented in R package \texttt{ncvreg} \citep{ncvreg}.
+Before being applied to calculating partial likelihood, such approach also appeared in literatures that aimed to conduct cross-validation for statistics that can hardly be subsetted, such as AUC for both logistic regression and Cox regression \citep{Parker2007}\citep{Simon2011a}\citep{Subramanian2011}.
 
@@
 The fundamental challenge of conducting cross validation for cox regression is that the baseline hazard is not estimated from the model.
-Hence we proposed our second approach which involves estimating the actual baseline hazard $\hat{\Lambda}_{0}$.
-The sum of squares of Martingale residuals is a natural candidate for evaluating the model performance using baseline hazard.
-Martingale residuals and its normalized form, deviance residuals, were first proposed by \citep{Therneau1990} for Cox regression model diagnostics.
+Hence we proposed our next approach based on getting the baseline hazard $\hat{\Lambda}_{0}$.
+We would first obtain cross-validated linear predictors $\hat{\eta}^{cv}$ from $D_k$ in the same way as the previous approach.
+With $\hat{\eta}^{cv}$,  a cross-validated version of baseline hazard $\hat{\Lambda}^{cv}_{0}$ that uses all observations can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}.
 
-To estimate the baseline hazard, a set of linear predictors $\hat{\eta}^{-} = ( \hat{\eta}^{-}_{1},  \hat{\eta}^{-}_{2} , ...
-\hat{\eta}^{-}_{n})$ would be obtained in the same way as the cross validated linear predictor approach.
-With $\hat{\eta}^{-}$, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}.
-Then for each observations, the Martingale Residual can be calculated:
+As an analogue to the residual sum of squaires in linear regression, we considered using the sum of squared Martingale residuals, preferably its normalized form, deviance residuals, to measure the model's predictive accuracy \citep{Therneau1990}.
+First proposed for model diagnosis, deviance residuals have also been used as a predictive accuracy measure in survival tree models \citep{Therneau2018}.
+Given that the baseline hazard and linear predictors are obtained via cross-validation, we can first obtain the cross-validated Martingale residual:
 	\begin{equation}
-	\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}^{-}_{i}},
+	\hat{M^{cv}_{i}} = \delta_{i} -\hat{\Lambda}^{cv}_{0}(t_{i})e^{\hat{\eta}^{cv}_{i}},
 	\end{equation} 
-where $\delta_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation. The deviance residual can be calcualted from the Martingale Residuals: 
+where $\delta_{i}$ is the status of the $ith$ observation and $t_{i}$ is the time component of the $ith$ observation.
+The cross-validated deviance residuals can then be derived from the Martingale residuals:
 	\begin{equation} 
-	d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}.
+	d_{i} = sgn(\hat{M}^{cv}_{i})(-2(\hat{M}^{cv}_{i} + \delta_{i}log(\delta_{i} - \hat{M}^{cv}_{i})))^{1/2}.
 	\end{equation}
- The sum of squares of the deviance residuals, $\sum_{i}\hat{d}_{i}^2$, would be used as the cross validated error. An illustration of this method is drawn in the bottom right panel of Figure 1.
+The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$ are used as the cross validated error. We refer this approach as the \emph{cross-validated deviance residuals}.
 
-As an alternative, the baseline hazard can also be estimated only based on the training set, leaving the test set as out-of-sample observations for calculating the residuals.
+A more straightforward way to compute the cross-validated deviance residuals is to estimate the baseline hazard using $T_k$, leaving $D_k$ as out-of-sample observations to calculate the residuals.
 There are two major issues with this approach.
-First, it would be common to run into situations where events in the test set occurred after the last event occurred in the baseline, where the survival function dropped to 0.
-Unless those observations are treated as censored, this approach would not work.
-Second, there would be an extra numeric challenge for deviance residuals.
-Since the baseline hazard function $\hat{\Lambda}_{0}$ is a step function from 0, the cumulative hazard at the first time point from 0 is always 0.
-Since deviance residual contains a term $log(\delta_{i} - \hat{M}_{i})$, where $\delta_{i} - \hat{M}_{i} = \hat{\Lambda}_{0}$, this term would be negative infinity for the first time point and lead to a numeric issue for whichever fold that contains the first time point.
-Hence extra smoothing for the step function would be needed for this approach to work.
+First, it is possible that some events in $D_k$ occurred after the last event in $T_k$.
+In this case, the survival function would drop to 0.
+Secondly, there is a numeric challenge for computing the term $log(\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}_{i}})$ in deviance residuals.
+Since the baseline hazard function is a step function from 0, the cumulative hazard $\hat{\Lambda}_{0}$ at the first time point from 0 is always 0.
+For whichever $D_k$ that contains the first time point, $log(\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}_{i}})$ would be negative infinity for that observation and lead to a numeric issue.
+To solve the first issue, events that occur later than the last event in $T_k$ should be censored.
+To solve the second issue, the step function of the baseline hazard needs to be smoothed.
+Both issues do not exist when the baseline hazard is built over the cross-validated linear predictors using all samples.
+\note{this paragraph may be optional, and can also be moved to discussion}
 
@@
 
-\note{tranpose sign in figure 1 is corrected}
-
 %%%%%%%%%%%%%%%%%%%%
@@
 
-  Simulation studies were conducted to compare how those methods behave relative to each other.
-We generated data with pre-specified baseline hazard, covariate matrix X, coefficient $\beta$ and censoring mechanism.
-Both low dimensional and high dimensional scenarios were examined.
-All cross-validation methods mentioned in Section 2 were applied to the data to select the tuning parameter $\lambda$ and produce $\hat{\beta}$ estimates.
-%Cross-validation were also compared to model selection criteria AIC and BIC.
-
- The entries in the covariate matrix $X_{n \times p}$ were independently generated from $Normal(0, 1)$.
-True coefficients $\beta$ were assumed to have sparsity: $\beta_{p\times 1} = (\beta_{1},\beta_{2}, ..., \beta_{j}, 0, ..., 0)^{T}$.
-The covariate matrix $X_{n \times p}$ was centered before it was used to generate outcomes.
-Survival times were generated from exponential distribution $h(t) = h_{0} exp(X\beta)$, conditioned on covariates.
-Censoring status were generated based on binomial distribution.
+Simulation studies were conducted to compare how those methods behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for LASSO penalized regression.
+Survival times were generated from exponential distribution $h(t) = h_{0} exp(X\beta)$, where $h_{0}$ is the prespecified baseline; $X$ is the covariate matrix generated from independent $Normal(0, 1)$ with $n$ observations and $p$ covariates; and $\beta$, the coefficients of the true model, is a vector whose majority entries are set to be 0 to guarentee sparsity.
+Censoring status were generated based on an independent binomial distribution.
 All simulations were implemented in R \citep{R}.
 
-  Suppose $\hat{\beta}$ is the coefficients estimated by a fitted cox model.
-We quantify the predictive accuracy by measuring the distance between the fitted model and the true model by mean squared error MSE $= E(\hat{\beta} - \beta)^2$.
-For each generated data set, the $\lambda$ that has the minimal MSE is chosen as the optimal $\lambda$.
-Then the $\lambda$s that selected by the cross validation or information criteria would be compared to this optimal $\lambda$.
-If the $\lambda$ chosen by cross validation is smaller than the optimal $\lambda$, then the cross validation method would be considered liberal.
-If the $\lambda$ chosen by cross validation is larger than the optimal $\lambda$, then the cross validation method would be considered conservative.
+We first studied the accuracy of the models via several in-sample and out-of-sample measures.
+We varied the sample size, dimension, censoring percentage and signals of the simulated data sets to examine how those factors would affect the model selected via those cross-validation approaches.
+We also conducted a simulation study to examine the stability issue of the standard approach.
+Finally, we extended the comparison from K-fold cross-validation to Leave One Out Cross-validation (LOOCV).
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \subsection {Simulations Comparing Cross-validation Approaches}
+    \subsection {Simulations Comparing Predictive Accuracy}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\par The first sets of simulation studies were conducted to study the predictive accuarcy of the model selected by the forementioned cross-validation approaches.
+We compared the predictive accuracy of the selected model via several distinct criteria.
+
+\par We first measure the distance between the fitted model and the true model via mean squared error of the coefficients.
+Suppose $\hat{\beta}(\lambda)$ is the set of coefficients estimated by a cox model with LASSO penalty at $\lambda$.
+Mean squared error $MSE$ is defined to be
+\begin{equation}
+E(\hat{\beta}(\lambda) - \beta)^2
+\end{equation}
+For each generated data set, the $\lambda$ that has the minimal MSE is chosen as the optimal $\lambda$, which we denote by $\lambda^{opt}$.
+Then the models selected by the cross validation methods with $\lambda^{cv}$ would be compared to the model that corresponds to $\lambda^{opt}$.
+We would take the ratio between $MSE$ of $\hat{\beta}(\lambda^{cv})$ and $MSE$ of $\hat{\beta}(\lambda^{optimal})$ for each cross-validation selected $\lambda^{cv}$ in each generated data set.
+The results would be averaged over N replications.
+
+\par We also measure the predictive accuracy via Brier score, which can be interpreted as the mean squared error of the prediction \citep{VanHouwelingen2011}.
+Suppose $\hat{S}(t_0|\lambda,x)$ is the prediction of survival probability of an individual beyond time $t_0$ with covariate $x$ based on model $\hat{\beta}(\lambda)$.
+Suppose the actual observation is defined as $y = \mathbb{1}(\left\{ t > t_{0}\right\})$, which is an indicator of whether the event is observed after $t_0$ or not.
+Then the Brier score is defined to be:
+\begin{equation}
+	Brier(y, \hat{S}(t_0|\lambda,x)) = (y - \hat{S}(t_0|\lambda,x))^2
+\end{equation}
+In our simulations, for each generated data set, we generated another independent data set that has 1000 observations without censoring.
+We computed Brier Scores upon all individuals in this independent test set where $t_0$ is set to be the median survival time for all four models selected by cross-validation.
+Lower Brier Scores means smaller prediction error, thus indicating better prediction accuracy.
+
+\par The third measure we used is the Kullback-Leibler score, which measures the log-likelihood of the prediction model evaluated at the observations \citep{VanHouwelingen2011}.
+With the same notation defined in the previous paragraph, the Kullback-Leibler score is defined to be:
+\begin{equation}
+	KL(y, \hat{S}(t_0|\lambda,x)) = -\left\{ ylog(\hat{S}(t_0|\lambda,x) + (1 - y)log(1 - \hat{S}(t_0|\lambda,x)) \right\}
+\end{equation}
+We computed KL score upon individuals in the independent test set at their median survival time for all four models selected by cross-validation.
+Results were then averaged over the N replications.
+Lower KL Scores means smaller discrepency between the selected model and the true model, thus indicating better prediction accuracy.
+
+\par While the MSE, Brier Score and Kullback-Leibler Score are good measures of the model, they are not very interpretable.
+Hence, we included Harrell's C Index as the last measure of the predictive accuracy\citep{HarrellJr1984}.
+The C Index is a rank-based statistic that measures the concordance between the linear predictor of the selected model and the observed outcome.
+Suppose we arbitrarily take two observation out of the data.
+If the observation that has a greater prognostic score also has a greater risk or shorter survival time until the event occur, then this would be considered as a concordant pair.
+C Index estimates of proportion of concordance pairs in the data set.
+It has a range between 0.5 and 1.
+Where C Index of 0.5 means the prediction is as as good as flipping a coin and Index of 1 means perfect concordance.
+We also computed Harrell's C Index upon the independent test set for all four models selected by cross-validation and averaged over all replications.
 
@@
 Number of non-zero $\beta$s was 10.
-The signal in the data set was varied by varying the magnitude of the $\beta$s from 0.40 to 0.9.
+The signal in the data set was varied by varying the magnitude of the $\beta$s from 0.4 to 0.9.
 For each scenario, 200 replications were used.
@@
 
-\par Results of this simulation were illustrated in Figure 2.
-The x axis of the plots are the magnitude of the non-zero $\beta$ coefficients.
-The left panel compares the $\lambda$ selected by different methods.
+%The left panel compares the $\lambda$ selected by different methods.
 Verweij and Van Houwelingen's cross-validated partial likelihood approach, which is currently most widely used approach, consistently selects $\lambda$s that are greater than the optimal one.
@@
 
-\par The right panel compares the MSE ratios of the model selected.
-The MSEs of the cross-validation selected models are compared with the MSE of the model given by the optimal $\lambda$.
-The y-axis represents the ratio of the two.
+\par Results of the simulation were illustrated in Figure 2.
+The x axis of the plots are the magnitude of the non-zero $\beta$ coefficients.
+The left panel compares the MSE ratios of the models selected.
+The y-axis represents the MSE ratios between the cross-validation selected models and the optimal model.
 The standard method has the smallest MSE ratio.
 The linear predictor approach's performance is also very close the standard approach.
-Both Verweij and Van Houwelingen's approach and the proposed deviance residual approach yield models that have greater MSE, hence are not recommended for use in practice.
+Both Verweij and Van Houwelingen's approach and the proposed deviance residual approach yield models that have greater MSE.
+The middle panel compares the Brier scores of the models selected.
+The y-axis represents the Brier score.
+The standard method and the linear predictor approach have lower brier score across all levels of signals.
+Both Verweij and Van Houwelingen's approach and the proposed deviance residual approach yield models that have greater Brier scores.
+The left panel compares Harrell's C of the models selected.
+The y-axis represents Harrell's C, which increases from around 0.6 to 0.8 as the coefficent size increases.
+The standard method and the linear predictor approach have slightly higher Harrell's C across all levels of signals, indicating better prediction accuracy.
+Overall, all four methods have very close performance .
 
-\begin{table}
+\begin{table}[h]
 
@@
 \toprule
- n& p& $\#$ Non-Zero & Signal & Method & $\lambda$ & MSR Ratio &Brier Score & Kullback Leibler & C Index\\
+ n& p& $\#$ Non-Zero & Signal & Method & $\lambda$ & MSE Ratio &Brier Score & Kullback Leibler & C Index\\
 \midrule
-150 & 50 & 5 & Weak & V VH  & 0.095 (0.028) & 1.323 (0.443) & 0.242 (0.013) & 0.678 (0.029) & 0.617 (0.034) \\
+150 & 50 & 5 & Weak & V$\&$ VH  & 0.095 (0.028) & 1.323 (0.443) & 0.242 (0.013) & 0.678 (0.029) & 0.617 (0.034) \\
       &    &   &     & Standard  & 0.087 (0.030) & 1.323 (0.431) & 0.242 (0.013) & 0.678 (0.029) & 0.618 (0.032) \\
@@
 \addlinespace
-150 & 50 & 5 & Strong & V VH  & 0.070 (0.011) & 1.373 (0.362) & 0.177 (0.012) & 0.527 (0.030) & 0.750 (0.014) \\
+150 & 50 & 5 & Strong & V$\&$VH  & 0.070 (0.011) & 1.373 (0.362) & 0.177 (0.012) & 0.527 (0.030) & 0.750 (0.014) \\
     &    &   &     & Standard  & 0.066 (0.012) & 1.313 (0.358) & 0.176 (0.012) & 0.526 (0.029) & 0.749 (0.014) \\
@@
 \addlinespace
-300 & 10000 & 20 & Weak & V VH  & 0.157 (0.021) & 1.079 (0.063) & 0.258 (0.008) & 0.711 (0.017) & 0.548 (0.037) \\
-       &    &   &     & Standard  & 0.137 (0.030) & 1.053 (0.053) & 0.255 (0.010) & 0.705 (0.021) & 0.557 (0.036) \\
-       &    &   &     & LinearPred  & 0.139 (0.029) & 1.054 (0.058) & 0.256 (0.009) & 0.705 (0.020) & 0.556 (0.036) \\
-       &    &   &     & DevResid  & 0.157 (0.024) & 1.078 (0.063) & 0.258 (0.008) & 0.710 (0.017) & 0.548 (0.037) \\
+300 & 10000 & 20 & Weak & V$\&$VH  & 0.139 (0.021) & 1.353 (0.192) & 0.244 (0.015) & 0.681 (0.032) & 0.612 (0.051) \\
+      &    &   &     & Standard  & 0.105 (0.027) & 1.206 (0.150) & 0.228 (0.021) & 0.646 (0.045) & 0.639 (0.050) \\
+      &    &   &     & LinearPred  & 0.107 (0.026) & 1.218 (0.159) & 0.229 (0.021) & 0.648 (0.045) & 0.638 (0.049) \\
+      &    &   &     & DevResid  & 0.139 (0.027) & 1.341 (0.198) & 0.243 (0.018) & 0.678 (0.038) & 0.610 (0.058) \\
 \addlinespace
- 500 & 10000 & 20 & Moderate & V VH  & 0.098 (0.012) & 1.298 (0.131) & 0.236 (0.014) & 0.665 (0.029) & 0.64 (0.036) \\
-        &    &   &     & Standard  & 0.080 (0.014) & 1.134 (0.11) & 0.226 (0.014) & 0.641 (0.031) & 0.651 (0.033) \\
-        &    &   &     & LinearPred  & 0.080 (0.012) & 1.139 (0.101) & 0.226 (0.014) & 0.642 (0.030) & 0.652 (0.031) \\
-        &    &   &     & DevResid  & 0.101 (0.015) & 1.317 (0.147) & 0.238 (0.014) & 0.667 (0.029) & 0.637 (0.040) \\
+500 & 10000 & 20 & Strong & V$\&$VH  & 0.072 (0.005) & 1.837 (0.184) & 0.162 (0.015) & 0.495 (0.036) & 0.784 (0.021) \\
+      &    &   &     & Standard  & 0.056 (0.005) & 1.368 (0.146) & 0.153 (0.011) & 0.465 (0.029) & 0.782 (0.017) \\
+      &    &   &     & LinearPred  & 0.057 (0.005) & 1.410 (0.150) & 0.153 (0.011) & 0.466 (0.029) & 0.783 (0.017) \\
+      &    &   &     & DevResid  & 0.076 (0.005) & 1.985 (0.227) & 0.167 (0.015) & 0.508 (0.035) & 0.782 (0.022) \\
 \bottomrule
@@
 
-\par Simulation studies were also conducted in scenarios with various sample sizes and dimensions.
+\par Additional simulation studies were conducted for other four scenarios, which a combination of different dimensions and signal in the data.
 The results are shown in Table 1.
-While the standard approach yields the smallest MSE and best predictive accuracy, the Linear Predictor and the standard approaches perform closely to each other across all settings.
-Verweij and Van Houwelingen's approach performs well in low-dimensions or with sufficient samples, while deviance residual approach performs better in high-dimensional settings.
+More specifically, we examined low dimension (p = 50) with weak signal, low dimension with strong signal, high dimension (p = 10000) with weak signal and high dimensional with strong signal.
+Overall, the standard approach and linear predictor approach tend to select smaller $\lambda$ and select more variables than the other two approaches.
+Among all four of them, the Deviance residual approahc is the most conservative one.
+
+\par Scenarios with higher signal and lower dimensions are considered as easier scenarios to select the most predictive model.
+As is shown in the results, at p = 10000 with weak signal, which is the hardest scenario, the standard approach and the linear predictor approach have the greatest advantage.
+They out-performed the other two methods in all predictive measures.
+The differences in performance are neglegible at low dimension or when there is strong signal in the data.
 
@@
 \par In this section, simulations studies were conducted to address the stability issue of the standard cross-validated likelihood.
-The standard approach is an intuitive way to carry out cross-validation.
-It also has the best predictive accuracy based on the simulation results in previous sections.
-However, one big disadvantage of this approach is that it is unstable when number of observed events in some fold is really small.
-As is illustrated in Figure 3, when censoring increases or when number of folds increases, the model would fail to converge due to insufficient number of events in the test set.
-Hence, the standard cross-validated likelihood should not be used when there is no sufficient number of events.
+The standard approach is the most straighforward among all four methods that are presented in this paper.
+It also has the best predictive accuracy based on the simulation results in the previous section.
+However, this approach is much less widely used due to its unstability.
+When number of observed events in $D_k$ is really small, there would be insufficient number of events to construct the risk set for the partial likelihood.
+We thus conducted a simulation study varying the number of folds and censoring to see when the method would start to have convergence issue.
+In both studies, we set the number of observations to be n = 100.
+When we varied the censoring percentage, we fixed the number of folds K = 10; when we varied the number of folds, we fixed the censoring to be 10$\%$.
+\par As is illustrated in Figure 3,  the standard approach start to have converge issue fairly quickly when censoring increases or when the number of folds increases.
+While the linear predictor approach performs as well as the standard approach in terms of prediction accuracy, the linear predictor approach is also as stable as the Verweij and Van Houwelingen's approach.
 

### 2018-11-12 main.tex
oracle MSE and CV-AUC

@@
 \end{equation}
-For each generated data set, the $\lambda$ that has the minimal MSE is chosen as the optimal $\lambda$, which we denote by $\lambda^{opt}$.
+For each generated data set, an oracle model would be fitted using Cox regression only upon variables that are set as non-zero in the data generating process.
+The resulted estimates would be denoted by $\hat{\beta}^{oracle}$.
+Then the models selected by the cross validation methods with $\lambda^{cv}$ would be compared to the oracle model via
+\begin{equation}
+log\left\{ \frac{MSE(\hat{\beta}(\lambda^{cv}))}{MSE(\hat{\beta}^{oracle})} \right\}
+\end{equation}
+for each cross-validation selected $\lambda^{cv}$ in each generated data set. The results would be averaged over N replications.
+
+%For each generated data set, the $\lambda$ that has the minimal MSE is chosen as the optimal $\lambda$, which we denote by $\lambda^{opt}$.
 Then the models selected by the cross validation methods with $\lambda^{cv}$ would be compared to the model that corresponds to $\lambda^{opt}$.
@@
 \toprule
- n& p& $\#$ Non-Zero & Signal & Method & $\lambda$ & MSE Ratio &Brier Score & Kullback Leibler & C Index\\
+ n& p& $\#$ Non-Zero & Signal & Method & $\lambda$ & log(MSE Ratio) &Brier Score & Kullback Leibler & C Index\\
 \midrule
-150 & 50 & 5 & Weak & V$\&$ VH  & 0.095 (0.028) & 1.323 (0.443) & 0.242 (0.013) & 0.678 (0.029) & 0.617 (0.034) \\
-      &    &   &     & Standard  & 0.087 (0.030) & 1.323 (0.431) & 0.242 (0.013) & 0.678 (0.029) & 0.618 (0.032) \\
-      &    &   &     & LinearPred  & 0.088 (0.025) & 1.285 (0.419) & 0.241 (0.013) & 0.677 (0.028) & 0.620 (0.029) \\
-      &    &   &     & DevResid  & 0.132 (0.032) & 1.697 (0.716) & 0.248 (0.014) & 0.689 (0.030) & 0.601 (0.044) \\
+150 & 50 & 5 & 0.3 & V VH  & 0.095 (0.028) & 1.457 (0.954) & 0.242 (0.013) & 0.678 (0.029) & 0.617 (0.034) \\
+    &    &   &     & Standard  & 0.087 (0.030) & 1.459 (0.922) & 0.242 (0.013) & 0.678 (0.029) & 0.618 (0.032) \\
+    &    &   &     & LinearPred  & 0.088 (0.025) & 1.433 (0.927) & 0.241 (0.013) & 0.677 (0.028) & 0.620 (0.029) \\
+    &    &   &     & DevResid  & 0.132 (0.032) & 1.676 (1.019) & 0.248 (0.014) & 0.689 (0.030) & 0.601 (0.044) \\
 \addlinespace
-150 & 50 & 5 & Strong & V$\&$VH  & 0.070 (0.011) & 1.373 (0.362) & 0.177 (0.012) & 0.527 (0.030) & 0.750 (0.014) \\
-    &    &   &     & Standard  & 0.066 (0.012) & 1.313 (0.358) & 0.176 (0.012) & 0.526 (0.029) & 0.749 (0.014) \\
-    &    &   &     & LinearPred  & 0.067 (0.011) & 1.319 (0.335) & 0.176 (0.012) & 0.526 (0.030) & 0.750 (0.014) \\
-    &    &   &     & DevResid  & 0.110 (0.012) & 2.569 (0.826) & 0.187 (0.014) & 0.554 (0.033) & 0.750 (0.018) \\
+150 & 50 & 5 & 0.6 & V VH  & 0.070 (0.011) & 1.733 (0.922) & 0.177 (0.012) & 0.527 (0.030) & 0.750 (0.014) \\
+    &    &   &     & Standard  & 0.066 (0.012) & 1.688 (0.909) & 0.176 (0.012) & 0.526 (0.029) & 0.749 (0.014) \\
+    &    &   &     & LinearPred  & 0.067 (0.011) & 1.695 (0.925) & 0.176 (0.012) & 0.526 (0.030) & 0.750 (0.014) \\
+    &    &   &     & DevResid  & 0.110 (0.012) & 2.339 (0.868) & 0.187 (0.014) & 0.554 (0.033) & 0.750 (0.018) \\
 \addlinespace
-300 & 10000 & 20 & Weak & V$\&$VH  & 0.139 (0.021) & 1.353 (0.192) & 0.244 (0.015) & 0.681 (0.032) & 0.612 (0.051) \\
-      &    &   &     & Standard  & 0.105 (0.027) & 1.206 (0.150) & 0.228 (0.021) & 0.646 (0.045) & 0.639 (0.050) \\
-      &    &   &     & LinearPred  & 0.107 (0.026) & 1.218 (0.159) & 0.229 (0.021) & 0.648 (0.045) & 0.638 (0.049) \\
-      &    &   &     & DevResid  & 0.139 (0.027) & 1.341 (0.198) & 0.243 (0.018) & 0.678 (0.038) & 0.610 (0.058) \\
+300 & 10000 & 20 & low & V VH  & 0.139 (0.021) & 3.463 (0.438) & 0.244 (0.015) & 0.681 (0.032) & 0.612 (0.051) \\
+    &    &   &     & Standard  & 0.105 (0.027) & 3.350 (0.458) & 0.228 (0.021) & 0.646 (0.045) & 0.639 (0.05) \\
+    &    &   &     & LinearPred  & 0.107 (0.026) & 3.360 (0.462) & 0.229 (0.021) & 0.648 (0.045) & 0.638 (0.049) \\
+    &    &   &     & DevResid  & 0.139 (0.027) & 3.454 (0.447) & 0.243 (0.018) & 0.678 (0.038) & 0.610 (0.058) \\
 \addlinespace
-500 & 10000 & 20 & Strong & V$\&$VH  & 0.072 (0.005) & 1.837 (0.184) & 0.162 (0.015) & 0.495 (0.036) & 0.784 (0.021) \\
-      &    &   &     & Standard  & 0.056 (0.005) & 1.368 (0.146) & 0.153 (0.011) & 0.465 (0.029) & 0.782 (0.017) \\
-      &    &   &     & LinearPred  & 0.057 (0.005) & 1.410 (0.150) & 0.153 (0.011) & 0.466 (0.029) & 0.783 (0.017) \\
-      &    &   &     & DevResid  & 0.076 (0.005) & 1.985 (0.227) & 0.167 (0.015) & 0.508 (0.035) & 0.782 (0.022) \\
+500 & 10000 & 20 & strong & V VH  & 0.072 (0.005) & 3.469 (0.480) & 0.162 (0.015) & 0.495 (0.036) & 0.784 (0.021) \\
+    &    &   &     & Standard  & 0.056 (0.005) & 3.174 (0.504) & 0.153 (0.011) & 0.465 (0.029) & 0.782 (0.017) \\
+    &    &   &     & LinearPred  & 0.057 (0.005) & 3.204 (0.498) & 0.153 (0.011) & 0.466 (0.029) & 0.783 (0.017) \\
+    &    &   &     & DevResid  & 0.076 (0.005) & 3.545 (0.467) & 0.167 (0.015) & 0.508 (0.035) & 0.782 (0.022) \\
 \bottomrule
@@
   %\subsection {Simulations on Deviance Residuals and Baseline Hazard Estimations}
+
+ \subsection {Cross-Validated C Index}
+
+\begin{figure}[h]
+    \centering
+		\includegraphics[height= 15cm ]{./figures/figure_auc.png}
+    \caption{Cross-Validated AUC}
+\end{figure}
 
@@
 
-\par While the Deviance Residual approach has a reasonable performance in scenarios where dimension is high, it tends to act more conservatively than all the other three approaches in general.
+%\par While the Deviance Residual approach has a reasonable performance in scenarios where dimension is high, it tends to act more conservatively than all the other three approaches in general.
 As is noted by Therneau et al, the sums of squared martingale residuals and deviance residuals do not necessary reflect how good the model fit is \citep{Therneau2000modeling}.
@@
 
-\par In the following simulation studies, the defined optimal $\lambda$ and the $\lambda$ selected by the Linear Predictor approach were used as benchmarks.
+%\par In the following simulation studies, the defined optimal $\lambda$ and the $\lambda$ selected by the Linear Predictor approach were used as benchmarks.
 Various sample sizes and dimensions were varied.
@@
 
-\par Since both of the non-parameteric estimations seemed to perform well at early time points but poorly later, a weighted approach was then applied to emphasize earlier time points.
+%\par Since both of the non-parameteric estimations seemed to perform well at early time points but poorly later, a weighted approach was then applied to emphasize earlier time points.
 Weights are assigned proportional to number of subjects at risk, making earlier time points weighted more than later time points.
@@
 
-\par In the next step,  the non-parametric baseline hazard estimation was replaced by the exact baseline hazard that was used to generate the outcome.
+%\par In the next step,  the non-parametric baseline hazard estimation was replaced by the exact baseline hazard that was used to generate the outcome.
 Results were listed in the seventh column in Table 2.
@@
 
-\par In the last step, we replaced the non-parametric baseline hazard estimation with a parametric approach, where we assume the baseline hazard to follow an exponential distribution.
+%\par In the last step, we replaced the non-parametric baseline hazard estimation with a parametric approach, where we assume the baseline hazard to follow an exponential distribution.
 We derived the estimation for baseline hazard by maximum likelihood approach conditioned on covariates: $\hat{\lambda}_{0} = \frac{\sum d_{i}}{\sum t_{i}e^{\eta_{i}}}$.
@@
 
-\par Hence, getting accurate estimation of the baseline hazard is crucial to whether or not the sum of squared deviance residuals can be a good measurement for model fit.
+%\par Hence, getting accurate estimation of the baseline hazard is crucial to whether or not the sum of squared deviance residuals can be a good measurement for model fit.
 
-\begin{table}[h]
-	\small
-	\caption{Average of Selected $\lambda$ with Different Baseline Estimation for Deviance Residuals}
-	\centering
-	\begin{tabular}{lllllllll}
-	\hline
-	\\[-0.75em]
-    & Optimal & Linear Pred & KP & KP & Breslow & Breslow & Exact & Exp\\ 
-    &  & & & weighted & & weighted & & \\ \cline{2 - 9} 
-	\textbf{True Baseline: $Exponential(1)$} &&&&&&&&\\
-	n = 100, p = 10, non-zero $\beta$: 2 & 0.0461 & 0.0665 & 0.1986 & 0.1970 & 0.1908 & 0.1975 & 0.0540 & 0.0539 \\
-	n = 100, p = 100, non-zero $\beta$: 20 &0.0555 & 0.1841 & 0.2359 & 0.2201 & 0.2146 & 0.2244 & 0.0805 & 0.0985 \\
-	n = 100, p = 1000, non-zero $\beta$: 20 & 0.0268 & 0.0505 & 0.0948 & 0.0798 & 0.0867 & 0.0809 & 0.0288 & 0.0270 \\
-	\textbf{True Baseline:} $Weibull$ &&&&&&&& \\
- 	n = 100, p = 10, non-zero $\beta$: 2 & 0.0476 & 0.0710 & 0.2004 & 0.1969 & 0.1907 & 0.1907 & N/A & 0.2330 \\ 
-	\hline
-\end{tabular}
-\end{table}
+%\begin{table}[h]
+%	\small
+%	\caption{Average of Selected $\lambda$ with Different Baseline Estimation for Deviance Residuals}
+	%\centering
+	%\begin{tabular}{lllllllll}
+	%\hline
+	%\\[-0.75em]
+    %& Optimal & Linear Pred & KP & KP & Breslow & Breslow & Exact & Exp\\ 
+    %&  & & & weighted & & weighted & & \\ \cline{2 - 9} 
+	%\textbf{True Baseline: $Exponential(1)$} &&&&&&&&\\
+	%n = 100, p = 10, non-zero $\beta$: 2 & 0.0461 & 0.0665 & 0.1986 & 0.1970 & 0.1908 & 0.1975 & 0.0540 & 0.0539 \\
+	%n = 100, p = 100, non-zero $\beta$: 20 &0.0555 & 0.1841 & 0.2359 & 0.2201 & 0.2146 & 0.2244 & 0.0805 & 0.0985 \\
+	%n = 100, p = 1000, non-zero $\beta$: 20 & 0.0268 & 0.0505 & 0.0948 & 0.0798 & 0.0867 & 0.0809 & 0.0288 & 0.0270 \\
+	%\textbf{True Baseline:} $Weibull$ &&&&&&&& \\
+ 	%n = 100, p = 10, non-zero $\beta$: 2 & 0.0476 & 0.0710 & 0.2004 & 0.1969 & 0.1907 & 0.1907 & N/A & 0.2330 \\ 
+	%\hline
+%\end{tabular}
+%\end{table}
 

### 2018-11-14 main.tex
some wording/citation changes

@@
 This manuscript is concerned with the question: how should one use the $kth$ fold $D_{k}$ to measure predictive accuracy based on partial likelihood?
-In the following sections, we use $L^{-k}$ to denote the partial likelihood when it is built over data in $T_k$; $L^{k}$ when it is built over data in $D_k$; and $L$ when it is built over the entire data set $D$.
+In the following sections, we use $L^{-k}$ to denote the partial likelihood that is built over data in $T_k$; use $L^{k}$ for partial likelihood built over data in $D_k$; and $L$ for partial likelihood built over the entire data set $D$.
 The log partial likelihood is denoted by lower case $l$.
-For individual subject's contribution to the likelihood, we use $L_i$, where $i$ denotes the individual.
+Individual subject's contribution to the likelihood is denoted by $L_i$, where $i$ denotes the individual.
 Particularly of our interest, if $L^{-k}(\beta)$ is used in \eqref{eq:obj} to obtain penalized likelihood estimates using data in $T_k$, we use $\hat{\beta}^{-k}$ to denote the penalized estimates that are obtained.
@@
 Hence we proposed our next approach based on getting the baseline hazard $\hat{\Lambda}_{0}$.
-We would first obtain cross-validated linear predictors $\hat{\eta}^{cv}$ from $D_k$ in the same way as the previous approach.
+We would first obtain cross-validated linear predictors $\hat{\eta}^{cv}$ in the same way as the previous approach.
 With $\hat{\eta}^{cv}$,  a cross-validated version of baseline hazard $\hat{\Lambda}^{cv}_{0}$ that uses all observations can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}.
 
-As an analogue to the residual sum of squaires in linear regression, we considered using the sum of squared Martingale residuals, preferably its normalized form, deviance residuals, to measure the model's predictive accuracy \citep{Therneau1990}.
+As an analogue to the residual sum of squaires in linear regression, we considered using the sum of squared deviance residuals, the normalized form of Martingale residuals, to measure the model's predictive accuracy \citep{Therneau1990}.
 First proposed for model diagnosis, deviance residuals have also been used as a predictive accuracy measure in survival tree models \citep{Therneau2018}.
@@
 \end{equation}
-for each cross-validation selected $\lambda^{cv}$ in each generated data set. The results would be averaged over N replications.
+for each cross-validation selected $\lambda^{cv}$ in each generated data set. The results would be averaged over N replications. We take $log$ of the MSE ratio to correct the right skewness.
 
@@
 C Index estimates of proportion of concordance pairs in the data set.
-It has a range between 0.5 and 1.
-Where C Index of 0.5 means the prediction is as as good as flipping a coin and Index of 1 means perfect concordance.
+C Index of 0.5 means the prediction is as good as flipping a coin and C Index of 1 means perfect concordance.
 We also computed Harrell's C Index upon the independent test set for all four models selected by cross-validation and averaged over all replications.
@@
 The simulation experiments were first conducted when the signal in the data was varied.
-In each simulated data set, there are n = 150 observations and p = 1000 covariates.
+In each simulated data set, there are n = 120 observations and p = 1000 covariates.
 10$\%$ of the observations were censored.
@@
 The x axis of the plots are the magnitude of the non-zero $\beta$ coefficients.
-The left panel compares the MSE ratios of the models selected.
-The y-axis represents the MSE ratios between the cross-validation selected models and the optimal model.
+The left panel compares the MSE ratios of selected models.
+The y-axis represents the MSE ratios between the cross-validation selected models and the oracle model.
 The standard method has the smallest MSE ratio.
@@
 \midrule
-150 & 50 & 5 & 0.3 & V VH  & 0.095 (0.028) & 1.457 (0.954) & 0.242 (0.013) & 0.678 (0.029) & 0.617 (0.034) \\
+150 & 50 & 5 & low & V VH  & 0.095 (0.028) & 1.457 (0.954) & 0.242 (0.013) & 0.678 (0.029) & 0.617 (0.034) \\
     &    &   &     & Standard  & 0.087 (0.030) & 1.459 (0.922) & 0.242 (0.013) & 0.678 (0.029) & 0.618 (0.032) \\
@@
 \addlinespace
-150 & 50 & 5 & 0.6 & V VH  & 0.070 (0.011) & 1.733 (0.922) & 0.177 (0.012) & 0.527 (0.030) & 0.750 (0.014) \\
+150 & 50 & 5 & strong & V VH  & 0.070 (0.011) & 1.733 (0.922) & 0.177 (0.012) & 0.527 (0.030) & 0.750 (0.014) \\
     &    &   &     & Standard  & 0.066 (0.012) & 1.688 (0.909) & 0.176 (0.012) & 0.526 (0.029) & 0.749 (0.014) \\
@@
 
-\par Additional simulation studies were conducted for other four scenarios, which a combination of different dimensions and signal in the data.
+\par Additional simulation studies were conducted for other four scenarios, which are a combination of different dimensions and signal in the data.
 The results are shown in Table 1.
-More specifically, we examined low dimension (p = 50) with weak signal, low dimension with strong signal, high dimension (p = 10000) with weak signal and high dimensional with strong signal.
+More specifically, we examined low dimension (p = 50) with weak signal, low dimension with strong signal, high dimension (p = 10000) with weak signal and high dimension with strong signal.
+Censoring was set to be 30$\%$.
 Overall, the standard approach and linear predictor approach tend to select smaller $\lambda$ and select more variables than the other two approaches.
@@
 
-\par In this section, simulations studies were conducted to address the stability issue of the standard cross-validated likelihood.
+\par In this section, simulation studies were conducted to address the stability issue of the standard cross-validated likelihood.
 The standard approach is the most straighforward among all four methods that are presented in this paper.
@@
 When we varied the censoring percentage, we fixed the number of folds K = 10; when we varied the number of folds, we fixed the censoring to be 10$\%$.
+N = 200 replications were used to estimate the convergence probability.
 \par As is illustrated in Figure 3,  the standard approach start to have converge issue fairly quickly when censoring increases or when the number of folds increases.
@@
     \subsection {Leave-One-Out Cross Validation}
-	\par In this section, Leave-One-Out Cross Validation (LOOCV) were studied, in comparison with 10-fold CV.
+	\par In this section, we compared Leave-One-Out Cross Validation (LOOCV) with 10-fold cross validation.
 When first introduced, Verweij and Van Houwelingen's cross-validated partial likelihood was proposed so that LOOCV for Cox regression became feasible.
-It is impossible to carry out LOO cross-validation using the standard approach, because the partial likelihood cannot be built on the one left-out observation.
-While Verweij and Van Houwelingen's cross-validated partial likelihood solves the stability issue by aggregating more observations together, it is not an efficient way to evaluate the model fit.
-Simulation studies in Section 3.1 have illustrated that it tends to be conservative in variable selections and selects smaller models in various scenarios.
-This fact can be even more illustrated via the following simulation studies, where both 10-fold and LOOCV were conducted via Verweij and Van Houwelingen's approach, linear predictor and deviance residual approaches and 10 fold CV was conducted for the standard approach.
-	\par As is illustrated in Figure 4, LOOCV selected models that have smaller MSE than 10-fold CV for all three CV approaches.
+While Verweij and Van Houwelingen's cross-validated partial likelihood solves the stability issue by aggregating more observations together, it does not seem to be an efficient way to evaluate the model fit based on results in Section 3.1.
+It tends to be conservative in variable selections and selects fewer variables in various scenarios.
+We further examined Verweij and Van Houwelingen's cross-validated partial likelihood under Leave-One-Out Cross Validation.
+In the following simulation studies, we used 10 fold cross-validated standard approach as our baseline for comparison.
+Then we conducted both 10-fold CV and LOOCV for Verweij and Van Houwelingen's approach, linear predictor approach and deviance residual approach.
+	\par Results of the simulation experiment is illustrated in Figure 4.For all three methods, LOOCV selected models that have smaller MSE than 10-fold CV for all three CV approaches.
 As is shown in left and middle panels, LOOCV based on deviance residuals and Verweij and Van Houwelingen's cross-validated partial likelihood still perform more conservatively than the standard approach.
 Even if Verweij and Van Houwelingen's approach was proposed to allow LOOCV, the LOOCV based on Verweij and Van Houwelingen's approach does not surpass the 10-fold standard approach.
-On the other hand, the Linear Predictors also allows LOOCV, and the LOOCV based on Linear Predictors performs better than the standard approach.
 
@@
  \subsection {Cross-Validated C Index}
+\par In this section, we conducted simulation experiments to compare likelihood-based cross validation against AUC-based cross validation.
+While the focus of this manuscript is to study valid cross-validation methods for partial-likelihood-based models such as penalized Cox rgression, the methods proposed in this paper are not applicable to some other survival machine learning models that do not involve constructing a partial likelihood.
+In those cases, the cross-validated C Index, also called cross-validated AUC, is one of the most popular options \citep{Subramanian2011} \citep{Simon2011a}.
+As is proposed by Simon et al., cross-validated AUC is computed also based on cross-validated linear predictors or some other types of prognostic factors.
+Thus, it is of our curiosity to see which one has a better performance when those two approaches, likelihood-based and AUC-based, are both available after we obtain the cross-validated linear predictors.
+
+\par We conducted simulations at both low dimension (p = 50) and high dimension (p = 1000), varying the signal in the data.
+Results of the simulations are illustrated in Figure 5.
+In general, the likelihood-based approach is more stable than the AUC-based cross validation approach.
+For MSE ratio and Brier Score, the likelihood approach tends to select more predictive models than the AUC-based approach.
+For the actual $\lambda$, it is interesting that the AUC-based approach is quite liberal and selected smaller $\lambda$s, which would yield models that contain more variables.
 

### 2018-11-29 cross-validation.tex
Edits to Ch. 2 intro
Some edits to 2.1, but we have problems
change diagram to pseudo code; add fold notation
revised till section 3.2
Edits to abstract

@@
 \usepackage{amssymb,amsmath,natbib,graphicx,enumerate,subcaption,tikz,url,booktabs}
-
 \usepackage[bmargin=0.75in, tmargin =0.75in,lmargin = 0.5in,rmargin = 0.5in]{geometry}
 
+\usepackage[ruled]{algorithm2e}
+\usepackage{algpseudocode}
+\renewcommand{\algorithmcfname}{Procedure}
+\usepackage{bbold}
+
+\newcommand{\ith}{i^\textrm{th}}
+\newcommand{\CVE}{\textrm{CVE}}
 \newcommand{\logit}{\mbox{logit}}
@@
 
+\newcommand{\lam}{\lambda}
 \newcommand{\bmu}{\boldsymbol\mu}
@@
 \section*{Abstract}
-Cross validation is commonly used for selecting tuning parameters in penalized regression, but little research has been conducted to study the performance of cross validation in penalized Cox regression.
-Due to its partial likelihood construction, carrying out cross validation for Cox models is not straightforward, and there are several ways one could imagine applying it.
+Cross validation is commonly used for selecting tuning parameters in penalized regression, but its use in penalized Cox regression models has received relatively little attention in the literature.
+Due to its partial likelihood construction, carrying out cross validation for Cox models is not straightforward, and there are several potential approaches for implementation.
 Here, we propose two new cross-validation methods for Cox regression and compare them to approaches that have been proposed elsewhere.
-Our simulation studies show that in general, all cross-validation approaches tend to be conservative (i.e., select smaller models than the optimal choice of tuning parameters) for penalized Cox regression models.
-However, our proposed approach of cross-validating the linear predictors seems to offer an attractive balance of performance and numerical stability.
+\note{IS THIS STILL TRUE?
+WE SHOULD EDIT THIS SENTENCE: Our simulation studies show that in general, all cross-validation approaches tend to be conservative (i.e., select smaller models than the optimal choice of tuning parameters) for penalized Cox regression models.} However, our proposed approach of cross-validating the linear predictors seems to offer an attractive balance of performance and numerical stability.
 We illustrate these advantages using simulated data as well as real data from two genome-wide studies of survival in cancer patients.

### 2018-11-29 main.tex
Edits to introduction

@@
 Cox regression utilizes a partial likelihood construction under an assumption of proportional hazards to estimate the regression coefficients without having to specify the underlying baseline hazard.
-The ability to avoid choosing a specific parametric distribution for the survival time is very attractive, as time-to-event data is often poorly described by fully parametric models, but this semiparametric approach comes at a cost.
+The ability to avoid choosing a specific parametric distribution for the survival time is very attractive, as time-to-event data are often poorly described by fully parametric models.
 
-\par The Cox regression approach can estimate relative risks, but without estimating the baseline hazard, it cannot make any predictions concerning the absolute failure time for any given individual.
+This semiparametric flexibility, however, comes at a cost.
+The Cox regression approach can estimate relative risks, but without estimating the baseline hazard, it does not make any predictions concerning the absolute failure time for any given individual.
 This poses a challenge to assessing the predictive accuracy of a given Cox regression model.
@@
 
-\par Standard cross-validation involves of dividing the data into $K$ folds, then fitting the model on $K-1$ of those folds (the training set) and assessing prediction accuracy on the remaining fold (the testing set).
+\par Standard cross-validation involves dividing the data into $K$ folds, then fitting the model on $K-1$ of those folds (the training set) and assessing prediction accuracy on the remaining fold (the testing set).
+This process is then repeated, with each fold serving as the testing set exactly once.
 For Cox regression, the estimated coefficients allow us to quantify the risk for each subject in the training set relative to other members of the training set, but it is not obvious how to use those estimates to quantify the model's accuracy in the testing set.
@@
 \par In this paper, we propose two alternative ways to carry out cross validation for Cox regression.
-Instead of cross-validating over partial likelihood, we propose cross validation over either the linear predictors of the regression model or over the deviance residuals \citep{Therneau1990}.
+Instead of applying cross-validation to the partial likelihood directly, we propose applying cross validation to either the linear predictors of the regression model or to the deviance residuals \citep{Therneau1990}.
 Through simulation studies, we compare these proposed methods with existing cross-validation approaches for LASSO penalized Cox regression in both low- and high-dimensional settings.
-We find that all cross validation approaches tend to be conservative, but that the linear predictor approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.
+\note{STILL TRUE??
+We find that all cross validation approaches tend to be conservative,} but that the linear predictor approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.
 We conclude by applying both proposed and existing approaches to two high dimensional data sets from real studies with time-to-event outcomes.

### 2018-12-05 main.tex
Edits to beginning of section 2, leaving off at VVH

@@
 Letting $t_i$ denote the time on study for subject $i$ and $\delta_{i}$ indicate whether or not an event is observed for subject $i$, each subject's individual contribution to that partial likelihood is
-\begin{equation*}
+\begin{equation}
+  \label{eq:cox-pl-subj}
   L_{i}(\beta) = \left \{\frac{\exp ( X_{i}^{T} \beta)}{\sum_{ k \in R(t_{i})}\exp ( X_{k}^{T} \beta)}\right \}^{\delta_{i}},
-\end{equation*}
+\end{equation}
 where $R(t_{i})$ denotes the set of subjects at risk at time $t_{i}$.  The partial likelihood for the entire sample of $n$ subjects is then given by:
-\begin{equation*}
+\begin{equation}
+  \label{eq:cox-pl-sum}
   L(\beta) =\prod_{i = 1}^{n} L_{i}(\beta).
-\end{equation*}
+\end{equation}
 Cox regression can be extended by introducing a penalty into the partial likelihood.  In penalized Cox regression, $\beta$ coefficient estimates are obtained by minimizing the objective function
@@
 
-\par In this paper, we focus upon the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$, although the methods we propose can be used with any penalty (indeed, can also be used for model selection in unpenalized Cox regression).
-LASSO cox regression is particularly useful for high dimensional data where the number of covariates p $\gg$ n; for example, in the prediction of cancer patients' overall survival based on genome-wide expression measurements.
-LASSO estimates yield a sparse model where some coefficients are estimated to be exactly zero, thereby carrying out automatic variable selection along the solution path.
-At large values of $\lambda$, most or all of the coefficients are 0, but as $\lam$ decreases, more covariates are selected into the model.
-Hence selecting $\lambda$ is critical to LASSO estimation in the sense that the model's accuracy suffers if $\lam$ is either too large or too small.
+\par In this paper, we focus upon the LASSO penalty $P_{\lambda}(\beta) = \lambda \sum_{j} |\beta_{j}|$, although the methods we analyze can be used with any penalty as well as with model selection in unpenalized Cox regression (best subset selection).
+LASSO-penalized cox regression is particularly useful for high dimensional data where the number of covariates $p > n$; for example, in the prediction of overall survival in cancer patients based on genome-wide expression measurements.
+LASSO estimates yield a sparse model where some coefficients are estimated to be exactly zero.
+This sparsity pattern changes as we vary $\lam$: at large values of $\lambda$, most or all of the coefficients are 0, but as $\lam$ decreases, more covariates are selected.
+Selecting $\lambda$ is critical to LASSO estimation in the sense that the model's accuracy suffers if $\lam$ is either too large or too small.
 
 \par Cross-validation is one of the most common methods used to select the appropriate $\lambda$ for penalized regression models.
-Suppose a data set $D$ of n observations is partitioned into $K$ folds: $D_{1}, D_{2}, ...
-D_{K}$.
-For a given $k \in \left\{1,2, ...
-, K\right\}$, set $T_{k} = D - D_{k}$.
-Then, we consider $T_{k}$ as the training set and $D_{k}$ as the test set.
+Suppose a data set $D$ of n observations is partitioned into $K$ folds: $D_{1}, D_{2}, \ldots, D_{K}$.
+For a given $k \in \left\{1,2,\ldots, K\right\}$, let $T_{k} = D - D_{k}$ denote the training set.
 This manuscript is concerned with the question: how should one use the $kth$ fold $D_{k}$ to measure predictive accuracy based on partial likelihood?
-In the following sections, we use $L^{-k}$ to denote the partial likelihood that is built over data in $T_k$; use $L^{k}$ for partial likelihood built over data in $D_k$; and $L$ for partial likelihood built over the entire data set $D$.
-The log partial likelihood is denoted by lower case $l$.
-Individual subject's contribution to the likelihood is denoted by $L_i$, where $i$ denotes the individual.
-Particularly of our interest, if $L^{-k}(\beta)$ is used in \eqref{eq:obj} to obtain penalized likelihood estimates using data in $T_k$, we use $\hat{\beta}^{-k}$ to denote the penalized estimates that are obtained.
+In particular, the partial likelihood \eqref{eq:cox-pl-subj} involves a risk set -- which observations should be used to form that risk set?
+In the following sections, let $L^{-k}$ denote the partial likelihood built over data in $T_k$, let $L^{k}$ denote the partial likelihood built over data in $D_k$, and $L$ denote the partial likelihood built over the entire data set $D$.
+The log partial likelihood is denoted by $\ell$, with $\ell^k$ and $\ell^{-k}$ defined similarly.
+Finally, let $\hat{\beta}^{-k}$ denote the penalized estimates that are obtained by using $L^{-k}(\beta)$ in \eqref{eq:obj}.
 
@@
 %%%%%%%%%%%%%%%%%%%%
-\par A straightforward approach to measure a model's predictive accuracy in $D_k$ is to evaluate the log partial likelihood at $\hat{\beta}^{-k}$ directly upon data in $D_k$.
-Greater likelihood usually represents a better fit and better predictive accuracy.
-Hence, the negative value of the partial likelihood can be used as the cross-validated error.
-We define the cross validated error (CVE) by pooling the negative log-likelihood from all K folds together:
+
+\par A direct approach to measure a model's predictive accuracy in $D_k$ is to evaluate the log partial likelihood at $\hat{\beta}^{-k}$ using the data in $D_k$.
+Typically, cross-validation error (CVE) is measured by the total deviance across all $K$ folds, which (up to a constant) is given by:
 \begin{equation}
   \label{eq:standard}
-  \CVE = - \sum_{k=1}^{K} l^{k}(\hat{\beta}^{-k}),
+  \CVE = -2 \sum_{k=1}^{K} \ell^{k}(\hat{\beta}^{-k}).
 \end{equation}
-where $l^{k}(\hat{\beta}^{-k}) = \sum \limits_{i \in D_{k}} log L^{k}_{i}(\hat{\beta}^{-k})$.
+%where $\ell^{k}(\hat{\beta}^{-k}) = \sum \limits_{i \in D_{k}} \log L^{k}_{i}(\hat{\beta}^{-k})$.
 We refer this approach as \emph{standard cross-validated partial likelihood}.
-This is implemented in the $\tt{glmnet}$ package as the $\tt{ungrouped}$ option.
+This is implemented in the $\tt{glmnet}$ package as the \verb|"ungrouped"| option.
+\note{I added a -2 here so that CVE corresponds to deviance; this is what's done in glmnet/ncvreg}.
 
-\par The standard approach is the same trick that is used to conduct cross-validation for linear regression or logistic regression.
-One pitfall of this approach is that there may not be sufficient observations in $D_k$ to build up the risk set for partial likelihood.
+\par The standard approach is by far the most common way of conducting cross-validation for models such as linear regression and logistic regression.
+However, the standard approach can be problematic for Cox regression in the sense that that there may not be sufficient observations in $D_k$ to build up the risk set for partial likelihood.
 An alternative trick that arises to address this issue was proposed by Verweij and Van Houwelingen.

### 2019-01-23 main.tex
try moving to git.com
more simulations on Brier score; update Shedden part

@@
     \centering
-		\includegraphics[height= 7cm ]{./figures/figure_3.png}
+		\includegraphics[height= 7cm ]{./figures/figure_3_new.png}
     \caption{Comparing Stability}
@@
 
+\begin{figure}[h]
+    \centering
+		\includegraphics[height= 15cm ]{./figures/figure_auc_extended.png}
+    \caption{Cross-Validated AUC}
+\end{figure}
+
 \section{Application to Real Data}
 
-\par In this section, we compared the performances of those cross validation methods when they are applied to data from real studies.
-The first data set is collected from a study on ovariance cancer from The Cancer Genome Atlas.
-The second dataset is a study on lung cancer \citep{shedden2008gene}.
-The outcomes of both data sets are patients' overall survival time.
-
-\par In the ovarian cancer data set, there are 460 patients and 236 events .
-About 40$\%$ of the observations are censored.
-The covariate matrix records whether or not there is mutation at certain gene location and has dimension of 12376.
-Three clinical variables, age, gender and residuals, which were known to have large impact on survival were adjusted in the analysis.
-That is to say, the clinical covariates were first fitted into regular cox regression and only linear predictors were kept so that the selection for the penalty term would not affect the estimation of those clinical variables.
-Cross-validation was used for the selection of the penalty term, which affects the selection of gene mutations.
-The results are illustrated in Table 3.
-
-\par In the lung cancer data set , researchers collected gene expression data for 22283 genes.
-236 death events were observed for 442 patients.
-About 50$\%$ of the patients are censored.
-Again clinical variables, age, gender, status of getting adjuvant chemotherapy were adjusted before the LASSO cox regression was fitted.
+\par In this section, we compared the performances of those cross validation methods when they are applied to data from real lung cancer study.
+\cite{shedden2008gene} conducted a large, retrospective study to validate prognostic gene expression signatures as predictors of overall survival in lung cancer patients, They performed gene expression profiling of \textit{442} non-small-cell lung cancer (NSCLC) patients.
+The data set contains gene expression measures from \textit{22,283} genes and basic clinical variables such as stage, age and gender.
+236 death events were observed and about 50$\%$ of the patients are censored.
+The clinical variables stage, age and gender were known to have a larger prognostic impact, thus they were fitted in the model without penalization.
 Cross-validation was again used for the selection of the penalty term, which selects the gene expressions that are associated with patients' overall survival.
 The results of the analysis are also listed in Table 3.
-
-\begin{table}[h]
-\centering
-\caption{Application to Real Data Sets}
-\label{my-label}
-\begin{tabular}{llllll}
-\hline
-\\[-0.75em]
- & Ovarian & & & Shedden&  \\
-& $\lambda$ & log($\lambda$) &  &$\lambda$  &  log($\lambda$)   \\ \cline{2-3} \cline{5-6} 
-V $\&$ VH &0.1226 & -2.0985 && 0.1215 & -2.1079\\
-standard & 0.0803 & -2.5222 && 0.0983 & -2.3197\\
-Linear Pred &0.0828 & -2.4919 && 0.1013 & -2.2894\\
-Dev Resid & 0.0992 & -2.3104 && 0.1110 & -2.1987\\ \hline
-\end{tabular}
-\end{table}
-
-\begin{figure}[h]
-    \centering
-		\includegraphics[height= 8cm ]{./figures/figure_5.png}
-    \caption{Comparing Cross-Validated Error for Different Methods on Real Data Sets}
-\end{figure}
 
@@
 For Verweij and Van Houwelingen's cross-validated partial likelihood, there's almost no signal there.
+
+\begin{figure}[h]
+    \centering
+		\includegraphics[height= 8cm ]{./figures/shedden.jpg}
+    \caption{Comparing Cross-Validated Error for Different Methods on Real Data Sets}
+\end{figure}
+
+\begin{table}[h]
+\centering
+\caption{Application to Real Data Sets}
+\label{my-label}
+\begin{tabular}{lccccc}
+\toprule
+						& $\lambda$ & log($\lambda$) & CV AUC & Brier Score \\ 
+						\cline{2-5} \\ 
+						V $\&$ VH& 0.123 & -2.097 & 0.640 & 0.246\\
+						standard & 0.102 & -2.278 & 0.643& 0.239\\
+						Linear Pred &0.085 & -2.460 & 0.651&  0.227\\
+						Dev Resid & 0.127& -2.066 & 0.639& 0.246\\ 
+						\bottomrule
+\end{tabular}
+\end{table}
 

### 2019-02-20 main.tex
Edits to Section 2

@@
 \par The standard approach is by far the most common way of conducting cross-validation for models such as linear regression and logistic regression.
-However, the standard approach can be problematic for Cox regression in the sense that that there may not be sufficient observations in $D_k$ to build up the risk set for partial likelihood.
-An alternative trick that arises to address this issue was proposed by Verweij and Van Houwelingen.
-They defined cross-validated error to be
+However, the standard approach can be problematic for Cox regression in the sense that that there may not be enough observations in $D_k$ to build up the risk set for partial likelihood.
+For example, the standard approach cannot work with leave-one-out cross-validation, since $\ell_k$ would either be zero or undefined for all folds.
+
+To address this issue, \citet{Verweij1993} proposed an alternative method for cross-validation in Cox models:
 \begin{equation}
 \label{eq:VVH}
-	CVE = -\sum_{k = 1}^K \left\{ l(\hat{\beta}^{- k})  - l^{-k}(\hat{\beta}^{- k}) \right\}. 
+	\CVE = -\sum_{k = 1}^K \left\{ l(\hat{\beta}^{- k})  - l^{-k}(\hat{\beta}^{- k}) \right\}. 
 \end{equation}
-$l(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on the entire data $D$ and $l^{-k}(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on the training set$T_k$.
-There should be enough observations in $T_k$ to build up the risk set as long as the entire data has a reasonable sample size.
-
-\par We refer to this definition as \emph{Verweij and Van Houwelingen's cross-validated partial likelihood}.
-It was originally proposed to as a trick to conduct Leave One Out Cross-Validation for partial likelihood, which is impossible with the standard approach.
-The standard approach and Verweij and Van Houwelingen's approach both try to measure the contributions from $D_k$ to the partial likelihood.
-For many other likelihoods, say logistic regression or linear regression, these two quantities, $l(\hat{\beta}^{- k})  - l^{-k}(\hat{\beta}^{- k})$ and  $l^{k}(\hat{\beta}^{-k})$, are in fact equivalent to each other.
-Since its original proposal, Verweij and Van Houwelingen's cross-validated error has been widely implemented as a tool to select tuning parameters for predictive models, especially regression-based models.
-It is implemented in the $\tt{glmnet}$ package as the default $\tt{grouped}$ option for penalized Cox regression.
-Other R packages that used this trick include but not limited to $\tt{CoxRidge}$, $\tt{fastcox}$, $\tt{SGL}$, $\tt{CoxBoost}$, and $\tt{mboost}$.
-\note{not sure whether we should mention these packages here.
-citations will be added later}
+Here, $l(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ using the entire data set $D$ and $l^{-k}(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on \note{FIX!
+the training set $T_k$.
+There should be enough observations in $T_k$ to build up the risk set as long as the entire data has a reasonable sample size.}  We refer to this cross-validation method as the Verweij and Van Houwelingen (V\&VH) approach.
+For many models, such as logistic regression or linear regression, these quantities $l(\hat{\beta}^{- k})  - l^{-k}(\hat{\beta}^{- k})$ and  $l^{k}(\hat{\beta}^{-k})$ are equivalent to each other -- in other words, the standard and V\&VH approaches agree.
+However, Verweij and Van Houwelingen's approach is more stable for Cox regression as there is always a large number of observations in the risk sets its calculations are based on.
+Since its proposal, the V\&VH approach has been widely implemented as a tool for cross-validation is Cox models.
+For example, it is used by the R packages {\tt CoxRidge}, {\tt fastcox}, {\tt SGL}, {\tt CoxBoost}, {\tt mboost}, and {\tt glmnet} packages.
+In {\tt glmnet}, it is the default choice for penalized Cox regression and is referred to as the {\tt grouped} option in the package's syntax.\note{PB: I like the mention of these packages, but they need citations.}
 
@@
 
-\par Instead of directly using partial likelihood to measure model's predictive accuracy over $D_k$, an alternative method is to obtain linear predictors for observations in $D_k$.
-We refer to this approach as \emph{cross-validated linear predictors}.
+\par Both of the methods in Section~\ref{Sec:cox-cv-existing} consist of adding together partial likelihood measures from each fold.
+An alternative method is to obtain linear predictors for each fold, then combine these linear predictors to calculate a partial likelihood.
+We refer to this method as the \emph{cross-validated linear predictors} approach.
 To be more specific, we would first obtain $\hat{\beta}^{-k}$ from training set $T_{k}$; then the cross-validated linear predictors can be calculated for each observation $i$ from the test set $D_k$:
 \begin{equation}
+  \label{eq:cv-lp}
   \hat{\eta}^{cv}_{i} = X_{i}\hat{\beta}^{-k}
@@
 	\begin{equation} 
-	L(\hat{\eta}^{cv}) = \prod_{i=1}^{n} \frac{exp (\hat{\eta}^{cv}_{i})}{\sum_{ j \in R(t_{i})}exp (\hat{\eta}^{cv}_{j})}
-	\end{equation}.
-We define the cross-validated error evaluated with cross-validated linear preditors to be $$ CVE = - log L(\hat{\eta}^{cv}).$$ 
+	L(\hat{\eta}^{cv}) = \prod_{i=1}^{n} \frac{exp (\hat{\eta}^{cv}_{i})}{\sum_{ j \in R(t_{i})}exp (\hat{\eta}^{cv}_{j})}.
+	\end{equation}
+\note{BIYUE: Throughout, use $\log$ instead of $log$, $\exp$ instead of $exp$, and $\ell$ instead of $l$.}
+We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - log L(\hat{\eta}^{cv}).$$
+
 \par This idea of obtaining linear predictors over test sets, then constructing the partial likelihood after pooling all linear predictors together is implemented in R package \texttt{ncvreg} \citep{ncvreg}.
-Before being applied to calculating partial likelihood, such approach also appeared in literatures that aimed to conduct cross-validation for statistics that can hardly be subsetted, such as AUC for both logistic regression and Cox regression \citep{Parker2007}\citep{Simon2011a}\citep{Subramanian2011}.
+This idea has also appeared in the cross-validation literature for quantities that cannot be evaluated on only a subset of the data, such as AUC in logistic and Cox regression models \citep{Parker2007,Simon2011a,Subramanian2011}.
 
@@
 
-  %Need to update this section with the new baseline estimation method%
-The fundamental challenge of conducting cross validation for cox regression is that the baseline hazard is not estimated from the model.
-Hence we proposed our next approach based on getting the baseline hazard $\hat{\Lambda}_{0}$.
-We would first obtain cross-validated linear predictors $\hat{\eta}^{cv}$ in the same way as the previous approach.
-With $\hat{\eta}^{cv}$,  a cross-validated version of baseline hazard $\hat{\Lambda}^{cv}_{0}$ that uses all observations can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}.
+The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model.
+Hence, another alternative approach to cross-validation would be to obtain an estimate $\hat{\Lambda}_{0}$ for the baseline hazard and use this in cross-validation.
+There are several potential approaches to estimating $\Lambda_0$; in this manuscript, we chose to obtain the cross-validated linear predictors $\hat{\eta}^{cv}$ as in \eqref{eq:cv-lp}, then estimate $\Lambda_0$ via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}, treating these linear predictors as fixed inputs.\note{Maybe this isn't critical, but I still think we could just use $\hat{\Lambda}_0$ from the full data fit \ldots that estimator just makes more sense to me}
 
-As an analogue to the residual sum of squaires in linear regression, we considered using the sum of squared deviance residuals, the normalized form of Martingale residuals, to measure the model's predictive accuracy \citep{Therneau1990}.
-First proposed for model diagnosis, deviance residuals have also been used as a predictive accuracy measure in survival tree models \citep{Therneau2018}.
+As an analogue to the residual sum of squares in linear regression, we considered using the sum of squared deviance residuals, the normalized form of Martingale residuals, to measure the model's predictive accuracy \citep{Therneau1990}.
+First proposed for model diagnostics, deviance residuals have also been used as a predictive accuracy measure in recursive partitioning models for survival data \citep{Therneau2018}.
 Given that the baseline hazard and linear predictors are obtained via cross-validation, we can first obtain the cross-validated Martingale residual:
@@
 	\end{equation} 
-where $\delta_{i}$ is the status of the $ith$ observation and $t_{i}$ is the time component of the $ith$ observation.
 The cross-validated deviance residuals can then be derived from the Martingale residuals: 
 	\begin{equation} 
-	d_{i} = sgn(\hat{M}^{cv}_{i})(-2(\hat{M}^{cv}_{i} + \delta_{i}log(\delta_{i} - \hat{M}^{cv}_{i})))^{1/2}.
+	d_{i} = sgn(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}log(\delta_{i} - \hat{M}^{cv}_{i}))}.
 	\end{equation}
-The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$ are used as the cross validated error. We refer this approach as the \emph{cross-validated deviance residuals}.
+The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$ are used as the cross validated error. We refer to this method as the \emph{cross-validated deviance residuals} approach.
 
-A more straightforward way to compute the cross-validated deviance residuals is to estimate the baseline hazard using $T_k$, leaving $D_k$ as out-of-sample observations to calculate the residuals.
+%A more straightforward way to compute the cross-validated deviance residuals is to estimate the baseline hazard using $T_k$, leaving $D_k$ as out-of-sample observations to calculate the residuals.
 There are two major issues with this approach.

### 2019-02-23 main.tex
add package citations; change l,exp and log; a few more comments in blue

@@
 \note{STILL TRUE??
-We find that all cross validation approaches tend to be conservative,} but that the linear predictor approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.
+We find that all cross validation approaches tend to be conservative,} \bnote{We should discuss this and the sentence in abstract at our next meeting.
+They are conservative only in terms of selecting the model that has the best preditive performance; they do not miss selecting variables in the true support}but that the linear predictor approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.
 We conclude by applying both proposed and existing approaches to two high dimensional data sets from real studies with time-to-event outcomes.
@@
 This is implemented in the $\tt{glmnet}$ package as the \verb|"ungrouped"| option.
-\note{I added a -2 here so that CVE corresponds to deviance; this is what's done in glmnet/ncvreg}.
+%\note{I added a -2 here so that CVE corresponds to deviance; this is what's done in glmnet/ncvreg}.
 
@@
 \label{eq:VVH}
-	\CVE = -\sum_{k = 1}^K \left\{ l(\hat{\beta}^{- k})  - l^{-k}(\hat{\beta}^{- k}) \right\}. 
+	\CVE = -\sum_{k = 1}^K \left\{ \ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k}) \right\}. 
 \end{equation}
-Here, $l(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ using the entire data set $D$ and $l^{-k}(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on \note{FIX!
+\bnote{we should add 2 here as well to be consistent with the notation in previous paragraph?}
+Here, $\ell(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ using the entire data set $D$ and $\ell^{-k}(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on %\note{FIX!}
 the training set $T_k$.
-There should be enough observations in $T_k$ to build up the risk set as long as the entire data has a reasonable sample size.}  We refer to this cross-validation method as the Verweij and Van Houwelingen (V\&VH) approach.
-For many models, such as logistic regression or linear regression, these quantities $l(\hat{\beta}^{- k})  - l^{-k}(\hat{\beta}^{- k})$ and  $l^{k}(\hat{\beta}^{-k})$ are equivalent to each other -- in other words, the standard and V\&VH approaches agree.
+There should be enough observations in $T_k$ to build up the risk set as long as the entire data has a reasonable sample size.
+We refer to this cross-validation method as the Verweij and Van Houwelingen (V\&VH) approach.
+For many models, such as logistic regression or linear regression, these quantities $\ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k})$ and  $\ell^{k}(\hat{\beta}^{-k})$ are equivalent to each other -- in other words, the standard and V\&VH approaches agree.
 However, Verweij and Van Houwelingen's approach is more stable for Cox regression as there is always a large number of observations in the risk sets its calculations are based on.
 Since its proposal, the V\&VH approach has been widely implemented as a tool for cross-validation is Cox models.
-For example, it is used by the R packages {\tt CoxRidge}, {\tt fastcox}, {\tt SGL}, {\tt CoxBoost}, {\tt mboost}, and {\tt glmnet} packages.
-In {\tt glmnet}, it is the default choice for penalized Cox regression and is referred to as the {\tt grouped} option in the package's syntax.\note{PB: I like the mention of these packages, but they need citations.}
+For example, it is used by the R packages {\tt CoxRidge} \citep{CoxRidge}, {\tt fastcox} \citep{fastcox}, {\tt SGL}\citep{SGL} , {\tt CoxBoost}\citep{CoxBoost}, {\tt mboost}\citep{mboost}, and {\tt glmnet}\citep{glmnet}.
+In {\tt glmnet}, it is the default choice for penalized Cox regression and is referred to as the {\tt grouped} option in the package's syntax.%\note{PB: I like the mention of these packages, but they need citations.}%\bnote{citation added}
 
@@
 	\begin{equation} 
-	L(\hat{\eta}^{cv}) = \prod_{i=1}^{n} \frac{exp (\hat{\eta}^{cv}_{i})}{\sum_{ j \in R(t_{i})}exp (\hat{\eta}^{cv}_{j})}.
+	L(\hat{\eta}^{cv}) = \prod_{i=1}^{n} \frac{\exp (\hat{\eta}^{cv}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}^{cv}_{j})}.
 	\end{equation}
-\note{BIYUE: Throughout, use $\log$ instead of $log$, $\exp$ instead of $exp$, and $\ell$ instead of $l$.}
-We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - log L(\hat{\eta}^{cv}).$$
+%\note{BIYUE: Throughout, use $\log$ instead of $log$, $\exp$ instead of $exp$, and $\ell$ instead of $l$.}
+We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - \log L(\hat{\eta}^{cv}).$$
 
@@
 Hence, another alternative approach to cross-validation would be to obtain an estimate $\hat{\Lambda}_{0}$ for the baseline hazard and use this in cross-validation.
-There are several potential approaches to estimating $\Lambda_0$; in this manuscript, we chose to obtain the cross-validated linear predictors $\hat{\eta}^{cv}$ as in \eqref{eq:cv-lp}, then estimate $\Lambda_0$ via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}, treating these linear predictors as fixed inputs.\note{Maybe this isn't critical, but I still think we could just use $\hat{\Lambda}_0$ from the full data fit \ldots that estimator just makes more sense to me}
+There are several potential approaches to estimating $\Lambda_0$; in this manuscript, we chose to obtain the cross-validated linear predictors $\hat{\eta}^{cv}$ as in \eqref{eq:cv-lp}, then estimate $\Lambda_0$ via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}, treating these linear predictors as fixed inputs.\note{Maybe this isn't critical, but I still think we could just use $\hat{\Lambda}_0$ from the full data fit \ldots that estimator just makes more sense to me}\bnote{Biyue is taking a look at the baseline now.
+Discuss this at next meeting}
 
@@
 	\begin{equation} 
-	d_{i} = sgn(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}log(\delta_{i} - \hat{M}^{cv}_{i}))}.
+	d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}.
 	\end{equation}
@@
 Simulation studies were conducted to compare how those methods behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for LASSO penalized regression.
-Survival times were generated from exponential distribution $h(t) = h_{0} exp(X\beta)$, where $h_{0}$ is the prespecified baseline; $X$ is the covariate matrix generated from independent $Normal(0, 1)$ with $n$ observations and $p$ covariates; and $\beta$, the coefficients of the true model, is a vector whose majority entries are set to be 0 to guarentee sparsity.
+Survival times were generated from exponential distribution $h(t) = h_{0} \exp(X\beta)$, where $h_{0}$ is the prespecified baseline; $X$ is the covariate matrix generated from independent $\text{Normal}(0, 1)$ with $n$ observations and $p$ covariates; and $\beta$, the coefficients of the true model, is a vector whose majority entries are set to be 0 to guarentee sparsity.
 Censoring status were generated based on an independent binomial distribution.
@@
 \begin{equation}
-log\left\{ \frac{MSE(\hat{\beta}(\lambda^{cv}))}{MSE(\hat{\beta}^{oracle})} \right\}
+\log\left\{ \frac{MSE(\hat{\beta}(\lambda^{\text{cv}}))}{MSE(\hat{\beta}^{\text{oracle}})} \right\}
 \end{equation}
@@
 \begin{equation}
-	Brier(y, \hat{S}(t_0|\lambda,x)) = (y - \hat{S}(t_0|\lambda,x))^2
+	\text{Brier}(y, \hat{S}(t_0|\lambda,x)) = (y - \hat{S}(t_0|\lambda,x))^2
 \end{equation}
@@
 \begin{equation}
-	KL(y, \hat{S}(t_0|\lambda,x)) = -\left\{ ylog(\hat{S}(t_0|\lambda,x) + (1 - y)log(1 - \hat{S}(t_0|\lambda,x)) \right\}
+	\text{KL}(y, \hat{S}(t_0|\lambda,x)) = -\left\{ y\log(\hat{S}(t_0|\lambda,x) + (1 - y)\log(1 - \hat{S}(t_0|\lambda,x)) \right\}
 \end{equation}

### 2019-02-27 main.tex
Edits to 3.1

@@
 The ability to avoid choosing a specific parametric distribution for the survival time is very attractive, as time-to-event data are often poorly described by fully parametric models.
+
+\note{What's with all the \textbackslash par commands?  They don't seem to do anything.}
 
@@
 \section{Methods}
+\label{Sec:methods}
 %%%%%%%%%%%%%
@@
 \label{eq:VVH}
-	\CVE = -\sum_{k = 1}^K \left\{ \ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k}) \right\}. 
+	\CVE = -2\sum_{k = 1}^K \left\{ \ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k}) \right\}. 
 \end{equation}
-\bnote{we should add 2 here as well to be consistent with the notation in previous paragraph?}
-Here, $\ell(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ using the entire data set $D$ and $\ell^{-k}(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on %\note{FIX!}
+Here, $\ell(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ using the entire data set $D$ and $\ell^{-k}(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on
 the training set $T_k$.
-There should be enough observations in $T_k$ to build up the risk set as long as the entire data has a reasonable sample size.
+By avoiding the construction of a partial likelihood on the test set, this approach ensures that the risk set is always sufficiently large and the partial likelihood well-defined.
 We refer to this cross-validation method as the Verweij and Van Houwelingen (V\&VH) approach.
-For many models, such as logistic regression or linear regression, these quantities $\ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k})$ and  $\ell^{k}(\hat{\beta}^{-k})$ are equivalent to each other -- in other words, the standard and V\&VH approaches agree.
+For many models, such as logistic regression or linear regression, the quantities $\ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k})$ and  $\ell^{k}(\hat{\beta}^{-k})$ are equivalent to each other -- in other words, the standard and V\&VH approaches agree.
 However, Verweij and Van Houwelingen's approach is more stable for Cox regression as there is always a large number of observations in the risk sets its calculations are based on.
@@
 For example, it is used by the R packages {\tt CoxRidge} \citep{CoxRidge}, {\tt fastcox} \citep{fastcox}, {\tt SGL}\citep{SGL} , {\tt CoxBoost}\citep{CoxBoost}, {\tt mboost}\citep{mboost}, and {\tt glmnet}\citep{glmnet}.
-In {\tt glmnet}, it is the default choice for penalized Cox regression and is referred to as the {\tt grouped} option in the package's syntax.%\note{PB: I like the mention of these packages, but they need citations.}%\bnote{citation added}
+In {\tt glmnet}, it is the default choice for penalized Cox regression and is referred to as the {\tt grouped} option in the package's syntax.
 
@@
 	\end{equation}
-%\note{BIYUE: Throughout, use $\log$ instead of $log$, $\exp$ instead of $exp$, and $\ell$ instead of $l$.}
-We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - \log L(\hat{\eta}^{cv}).$$
+We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - 2 \log L(\hat{\eta}^{cv}).$$
 
@@
 This idea has also appeared in the cross-validation literature for quantities that cannot be evaluated on only a subset of the data, such as AUC in logistic and Cox regression models \citep{Parker2007,Simon2011a,Subramanian2011}.
-
-  % \begin{figure}[h]
-    %\centering
-    %\begin{minipage}[b]{0.45\textwidth}
-      %\centering
-	%	  \includegraphics[height= 3.9cm ]{./figures/03.png}
-      %\caption{CV Linear Predictors}
-     %\end{minipage}
-     %\begin{minipage}[b]{0.45\textwidth}
-      %\centering
-%		  \includegraphics[height= 4.1cm ]{./figures/04_2.png}
-    %  \caption{CV Deviance Residuals}
-     % \end{minipage}	
-  % \end{figure}
 
@@
   \includegraphics[height= 12cm ]{./figures/figure_1_new.png}
-%		%\includegraphics[height= 4cm ]{./figures/02_2.png}
   \caption{Illustration of K-fold Cross Validation Methods for Cox Regression}
@@
 Simulation studies were conducted to compare how those methods behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for LASSO penalized regression.
-Survival times were generated from exponential distribution $h(t) = h_{0} \exp(X\beta)$, where $h_{0}$ is the prespecified baseline; $X$ is the covariate matrix generated from independent $\text{Normal}(0, 1)$ with $n$ observations and $p$ covariates; and $\beta$, the coefficients of the true model, is a vector whose majority entries are set to be 0 to guarentee sparsity.
-Censoring status were generated based on an independent binomial distribution.
+Survival times were generated from exponential distribution $h(t) = h_{0} \exp(X\beta)$, where $h_{0}$ is the prespecified baseline\note{huh?
+so what is $h_0$?}; the covariate matrix $X$ with $n$ observations and $p$ covariates is generated from independent $\text{Normal}(0, 1)$ samples; and the coefficient vector $\beta$ is sparse, with most entries equal to zero.
+Censoring status indicators were drawn independently from the binomial distribution.
 All simulations were implemented in R \citep{R}.
-
-We first studied the accuracy of the models via several in-sample and out-of-sample measures.
-We varied the sample size, dimension, censoring percentage and signals of the simulated data sets to examine how those factors would affect the model selected via those cross-validation approaches.
-We also conducted a simulation study to examine the stability issue of the standard approach.
-Finally, we extended the comparison from K-fold cross-validation to Leave One Out Cross-validation (LOOCV).
+We varied the sample size, dimension, censoring percentage and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\par The first sets of simulation studies were conducted to study the predictive accuarcy of the model selected by the forementioned cross-validation approaches.
-We compared the predictive accuracy of the selected model via several distinct criteria.
 
-\par We first measure the distance between the fitted model and the true model via mean squared error of the coefficients.
+Our first simulation study concerns the accuracy of the cross-validation approaches introduces in Section~\ref{Sec:methods}.  Accuracy was assesed via several criteria.
+
+Estimation accuracy was measured using the mean squared error of the coefficients ($\MSE$).
 Suppose $\hat{\beta}(\lambda)$ is the set of coefficients estimated by a cox model with LASSO penalty at $\lambda$.
@@
 \begin{equation}
-E(\hat{\beta}(\lambda) - \beta)^2
+E(\hat{\beta}(\lambda) - \beta)^2.
 \end{equation}
-For each generated data set, an oracle model would be fitted using Cox regression only upon variables that are set as non-zero in the data generating process.
-The resulted estimates would be denoted by $\hat{\beta}^{oracle}$.
-Then the models selected by the cross validation methods with $\lambda^{cv}$ would be compared to the oracle model via
+\note{this is the equation for a scalar MSE, not a vector}
+
+Throughout, we communicate the estimation accuracy in terms of relative MSE, as compared to the oracle model.
+For each generated data set, an oracle model was fit using Cox regression only including the variables with non-zero coefficients.
+Letting $\hat{\beta}^{oracle}$ denote the resulting estimates, the relative MSE, on the log scale, is given by
 \begin{equation}
-\log\left\{ \frac{MSE(\hat{\beta}(\lambda^{\text{cv}}))}{MSE(\hat{\beta}^{\text{oracle}})} \right\}
+\log\left\{ \frac{\MSE(\hat{\beta}(\lambda^{\text{cv}}))}{\MSE(\hat{\beta}^{\text{oracle}})} \right\},
 \end{equation}
-for each cross-validation selected $\lambda^{cv}$ in each generated data set. The results would be averaged over N replications. We take $log$ of the MSE ratio to correct the right skewness.
+where $\lambda^{cv}$ denotes the value of $\lambda$ selected by cross-validation.   These results are then averaged over N replications; the $\log$ of the ratio is taken to correct for skewness.
 
@@
 
-\par We also measure the predictive accuracy via Brier score, which can be interpreted as the mean squared error of the prediction \citep{VanHouwelingen2011}.
-Suppose $\hat{S}(t_0|\lambda,x)$ is the prediction of survival probability of an individual beyond time $t_0$ with covariate $x$ based on model $\hat{\beta}(\lambda)$.
-Suppose the actual observation is defined as $y = \mathbb{1}(\left\{ t > t_{0}\right\})$, which is an indicator of whether the event is observed after $t_0$ or not.
-Then the Brier score is defined to be:
+\par We also measure the predictive accuracy via the mean squared prediction error, or Brier score \citep{VanHouwelingen2011}.
+Let $\hat{S}(t_0|\lambda,x)$ denote the predicted probability for an individual with covariates $x$ to survive beyond time $t_0$, and let $y = \mathbb{1}\left\{ t > t_{0}\right\}$ indicate whether the individual actually survived beyond time $t_0$.
+Then the Brier score is defined to be
 \begin{equation}
-	\text{Brier}(y, \hat{S}(t_0|\lambda,x)) = (y - \hat{S}(t_0|\lambda,x))^2
+\text{Brier}(y, \hat{S}(t_0|\lambda,x)) = (y - \hat{S}(t_0|\lambda,x))^2.
 \end{equation}
-In our simulations, for each generated data set, we generated another independent data set that has 1000 observations without censoring.
-We computed Brier Scores upon all individuals in this independent test set where $t_0$ is set to be the median survival time for all four models selected by cross-validation.
-Lower Brier Scores means smaller prediction error, thus indicating better prediction accuracy.
+In our simulation study, for each data set, we generated another independent data set with 1000 uncensored observations.
+We computed Brier scores for all of the cross-validation methods for all individuals in this test set, setting $t_0$ to the median survival time.
 
 \par The third measure we used is the Kullback-Leibler score, which measures the log-likelihood of the prediction model evaluated at the observations \citep{VanHouwelingen2011}.
-With the same notation defined in the previous paragraph, the Kullback-Leibler score is defined to be:
+With the same notation defined in the previous paragraph, the Kullback-Leibler score is defined to be
 \begin{equation}
-	\text{KL}(y, \hat{S}(t_0|\lambda,x)) = -\left\{ y\log(\hat{S}(t_0|\lambda,x) + (1 - y)\log(1 - \hat{S}(t_0|\lambda,x)) \right\}
+	\text{KL}(y, \hat{S}(t_0|\lambda,x)) = -\left\{ y\log(\hat{S}(t_0|\lambda,x) + (1 - y)\log(1 - \hat{S}(t_0|\lambda,x)) \right\}.
 \end{equation}
-We computed KL score upon individuals in the independent test set at their median survival time for all four models selected by cross-validation.
+As for the Brier score, KL scores were computed for all CV approaches at the median survival time.
 Results were then averaged over the N replications.
-Lower KL Scores means smaller discrepency between the selected model and the true model, thus indicating better prediction accuracy.
+For both Brier and KL scores, a score represents smaller prediction error, thus indicating better prediction accuracy.
 
-\par While the MSE, Brier Score and Kullback-Leibler Score are good measures of the model, they are not very interpretable.
-Hence, we included Harrell's C Index as the last measure of the predictive accuracy\citep{HarrellJr1984}.
+\par The final metric we included in our study was Harrell's C Index \citep{HarrellJr1984}.
 The C Index is a rank-based statistic that measures the concordance between the linear predictor of the selected model and the observed outcome.
-Suppose we arbitrarily take two observation out of the data.
-If the observation that has a greater prognostic score also has a greater risk or shorter survival time until the event occur, then this would be considered as a concordant pair.
-C Index estimates of proportion of concordance pairs in the data set.
-C Index of 0.5 means the prediction is as good as flipping a coin and C Index of 1 means perfect concordance.
-We also computed Harrell's C Index upon the independent test set for all four models selected by cross-validation and averaged over all replications.
+Suppose we arbitrarily select two individuals; if the individual with the higher predicted risk also died first, then this would be considered as a concordant pair.
+The C Index considers all possible such pairs and estimates of proportion of those pairs that are concordant.
+A C Index of 0.5 means the prediction is no better than flipping a coin and C Index of 1 means perfect concordance.
+We computed Harrell's C Index using the independent test set of 1000 individuals described above.
 
-The simulation experiments were first conducted when the signal in the data was varied.
-In each simulated data set, there are n = 120 observations and p = 1000 covariates.
-10$\%$ of the observations were censored.
-10-fold cross-validation was implemented for all four methods.
+For each simulated data set, there were $n = 120$ observations, $p = 1000$ covariates.
+and 10\% of the observations were censored.
+For each of the four methods described in Section~\ref{Sec:methods}, 10-fold cross-validation was implemented.
 Number of non-zero $\beta$s was 10.
 The signal in the data set was varied by varying the magnitude of the $\beta$s from 0.4 to 0.9.
-For each scenario, 200 replications were used.
+\note{Would it be better here to descibe the $\beta$ values explicitly, with an equation?} For each scenario, 200 replications were used.
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \centering
   \includegraphics[height= 7cm ]{./figures/figure_2_new.png}
-    \caption{Simulation Studies Comparing 4 CV methods in terms of $\lambda$ and MSE ratio. n = 150, p = 1000, 10$\%$ censoring.}
+  \caption{\label{Fig:mse-brier-c} Simulation comparing 4 CV methods in terms of $\lambda$ and MSE ratio (relative to the oracle model) \note{UPDATE CAPTION}. n = 150, p = 1000, 10$\%$ censoring.}
 \end{figure}
@@
 
-\par Results of the simulation were illustrated in Figure 2.
-The x axis of the plots are the magnitude of the non-zero $\beta$ coefficients.
-The left panel compares the MSE ratios of selected models.
-The y-axis represents the MSE ratios between the cross-validation selected models and the oracle model.
-The standard method has the smallest MSE ratio.
-The linear predictor approach's performance is also very close the standard approach.
-Both Verweij and Van Houwelingen's approach and the proposed deviance residual approach yield models that have greater MSE.
-The middle panel compares the Brier scores of the models selected.
-The y-axis represents the Brier score.
-The standard method and the linear predictor approach have lower brier score across all levels of signals.
-Both Verweij and Van Houwelingen's approach and the proposed deviance residual approach yield models that have greater Brier scores.
-The left panel compares Harrell's C of the models selected.
-The y-axis represents Harrell's C, which increases from around 0.6 to 0.8 as the coefficent size increases.
-The standard method and the linear predictor approach have slightly higher Harrell's C across all levels of signals, indicating better prediction accuracy.
-Overall, all four methods have very close performance .
+\par Figure~\ref{Fig:mse-brier-c} illustrates the results of this simulation study.
+In each panel, the horizontal axis is the magnitude of the non-zero $\beta$ coefficients.
+For all four metrics, the overall pattern is the same (KL scores not shown in Figure~\ref{Fig:mse-brier-c} due to space constraints, but they were very similar to Brier scores).
+Although the four cross-validation methods performed relatively similarly, the standard and linear predictor approaches had the best performance across all metrics: lower MSE, better Brier scores, and higher concordance.
 
-\begin{table}[h]
+\begin{table}[htb]
 
@@
 
-\caption{\label{tab:} Simulation Studies that Compare All Four Cross-Validation Methods Varying Dimension and Signal }
+\caption{\label{Tab:sim} Simulation comparing four cross-validation methods with varying dimension and signal strength}
 \centering
@@
 
-\par Additional simulation studies were conducted for other four scenarios, which are a combination of different dimensions and signal in the data.
-The results are shown in Table 1.
-More specifically, we examined low dimension (p = 50) with weak signal, low dimension with strong signal, high dimension (p = 10000) with weak signal and high dimension with strong signal.
-Censoring was set to be 30$\%$.
-Overall, the standard approach and linear predictor approach tend to select smaller $\lambda$ and select more variables than the other two approaches.
-Among all four of them, the Deviance residual approahc is the most conservative one.
+Table~\ref{Tab:sim} displays the results of additional simulations carried out as we varied the dimension of the data set along with the signal strength.
+Specifically, we examined the following settings: low dimension (p = 50) with weak signal, low dimension with strong signal, high dimension (p = 10000) with weak signal, and high dimension with strong signal.
+In all settings, the censoring rate was 30\%.
 
-\par Scenarios with higher signal and lower dimensions are considered as easier scenarios to select the most predictive model.
-As is shown in the results, at p = 10000 with weak signal, which is the hardest scenario, the standard approach and the linear predictor approach have the greatest advantage.
-They out-performed the other two methods in all predictive measures.
-The differences in performance are neglegible at low dimension or when there is strong signal in the data.
+As in Figure~\ref{Fig:mse-brier-c}, the standard and linear predictor approaches outperform the deviance residual and $V\&VH$ approaches.
+The improvement is particularly noticeable in the most difficult setting: high-dimensional data with weak signal.
+Table~\ref{Tab:sim} also provides some insight into why the deviance residual and $V\&VH$ approaches perform poorly relative to standard CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and selecting fewer variables.
+\note{if we're going to say something about ``selecting fewer variables'', should Table 1 include the average model size?}
 

### 2019-03-06 main.tex
clean/reply to some previous comments

@@
 
-\note{What's with all the \textbackslash par commands?  They don't seem to do anything.}
+%\note{What's with all the \textbackslash par commands?  They don't seem to do anything.}
 
@@
 Through simulation studies, we compare these proposed methods with existing cross-validation approaches for LASSO penalized Cox regression in both low- and high-dimensional settings.
-\note{STILL TRUE??
+%\note{STILL TRUE??
 We find that all cross validation approaches tend to be conservative,} \bnote{We should discuss this and the sentence in abstract at our next meeting.
-They are conservative only in terms of selecting the model that has the best preditive performance; they do not miss selecting variables in the true support}but that the linear predictor approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.
+They are conservative only in terms of selecting the model that has the best preditive performance; they do not miss selecting variables in the true support}
+The linear predictor approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.
 We conclude by applying both proposed and existing approaches to two high dimensional data sets from real studies with time-to-event outcomes.
@@
 Hence, another alternative approach to cross-validation would be to obtain an estimate $\hat{\Lambda}_{0}$ for the baseline hazard and use this in cross-validation.
-There are several potential approaches to estimating $\Lambda_0$; in this manuscript, we chose to obtain the cross-validated linear predictors $\hat{\eta}^{cv}$ as in \eqref{eq:cv-lp}, then estimate $\Lambda_0$ via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}, treating these linear predictors as fixed inputs.\note{Maybe this isn't critical, but I still think we could just use $\hat{\Lambda}_0$ from the full data fit \ldots that estimator just makes more sense to me}\bnote{Biyue is taking a look at the baseline now.
-Discuss this at next meeting}
+There are several potential approaches to estimating $\Lambda_0$; in this manuscript, we chose to obtain the cross-validated linear predictors $\hat{\eta}^{cv}$ as in \eqref{eq:cv-lp}, then estimate $\Lambda_0$ via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}, treating these linear predictors as fixed inputs.\note{Maybe this isn't critical, but I still think we could just use $\hat{\Lambda}_0$ from the full data fit \ldots that estimator just makes more sense to me}\bnote{They do act a bit different in simulations.
+The baseline built from the full data performed worse.
+Discuss this next meeting}
 
@@
 Survival times were generated from exponential distribution $h(t) = h_{0} \exp(X\beta)$, where $h_{0}$ is the prespecified baseline\note{huh?
-so what is $h_0$?}; the covariate matrix $X$ with $n$ observations and $p$ covariates is generated from independent $\text{Normal}(0, 1)$ samples; and the coefficient vector $\beta$ is sparse, with most entries equal to zero.
+so what is $h_0$?} \bnote{$h_0$ is 1 in all simulations.
+Basically, I first generated $X$, then the linear predictor $X\beta$; Then I compute the rate parameter from $h_0exp(X\beta)$ and inserted this to the \texttt{rexp} function to generate the survival times}; the covariate matrix $X$ with $n$ observations and $p$ covariates is generated from independent $\text{Normal}(0, 1)$ samples; and the coefficient vector $\beta$ is sparse, with most entries equal to zero.
 Censoring status indicators were drawn independently from the binomial distribution.
@@
 Mean squared error $MSE$ is defined to be
+\bnote{
 \begin{equation}
-E(\hat{\beta}(\lambda) - \beta)^2.
+E\| \hat{\beta}(\lambda) - \beta \| ^2.
 \end{equation}
+}
 \note{this is the equation for a scalar MSE, not a vector}
@@
 Table~\ref{Tab:sim} also provides some insight into why the deviance residual and $V\&VH$ approaches perform poorly relative to standard CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and selecting fewer variables.
-\note{if we're going to say something about ``selecting fewer variables'', should Table 1 include the average model size?}
+\note{if we're going to say something about ``selecting fewer variables'', should Table 1 include the average model size?} \bnote{I'm afraid that the average model size would be confusing to the readers.
+The most conversative model would have the smallest model size, and would be closest to the true model size}
 
@@
 
-\begin{figure}[h]
-    \centering
-		\includegraphics[height= 15cm ]{./figures/figure_auc_extended.png}
-    \caption{Cross-Validated AUC}
-\end{figure}
+%\begin{figure}[h]
+%    \centering
+%		\includegraphics[height= 15cm ]{./figures/figure_auc_extended.png}
+%    \caption{Cross-Validated AUC}
+%\end{figure}
 

### 2019-03-06 main.tex
Edits to 3.2 and 3.3

@@
 \subsection {Simulations Comparing Predictive Accuracy}
+\label{Sec:accuracy}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection {Stability}
+\label{Sec:stability}
 
-\par In this section, simulation studies were conducted to address the stability issue of the standard cross-validated likelihood.
-The standard approach is the most straighforward among all four methods that are presented in this paper.
-It also has the best predictive accuracy based on the simulation results in the previous section.
-However, this approach is much less widely used due to its unstability.
-When number of observed events in $D_k$ is really small, there would be insufficient number of events to construct the risk set for the partial likelihood.
-We thus conducted a simulation study varying the number of folds and censoring to see when the method would start to have convergence issue.
-In both studies, we set the number of observations to be n = 100.
-When we varied the censoring percentage, we fixed the number of folds K = 10; when we varied the number of folds, we fixed the censoring to be 10$\%$.
-N = 200 replications were used to estimate the convergence probability.
-\par As is illustrated in Figure 3,  the standard approach start to have converge issue fairly quickly when censoring increases or when the number of folds increases.
-While the linear predictor approach performs as well as the standard approach in terms of prediction accuracy, the linear predictor approach is also as stable as the Verweij and Van Houwelingen's approach.
+\par In this section, we present a small simulation study to illustrate the numerical stability issues faced by standard cross-validation in the partial likelihood setting.
+The standard approach is both straightforward and enjoys very good predictive accuracy, as seen in Section~\ref{Sec:accuracy}.
+The main drawback to its widespread use is that the number of uncensored observations in the test set may be insufficient to construct a well-defined partial likelihood, leaving the overall estimate of cross-validation error undefined as well.
+
+To illustrate how common this problem is, and what factors it depends on, we conducted a simulation study varying the number of folds and censoring rate.
+Throughout we fixed the number of observations at $n = 100$.
+When varying the censoring rate, we fixed the number of folds at $K = 10$; when varying the number of folds, we fixed the censoring rate at 10\%.
+We generated 200 independent data sets according to the mechanism described in Section~\ref{Sec:accuracy} and recorded whether or not standard cross-validation was able to be carried out (i.e., whether the partial likelihood was well-defined for all folds).
+
+\note{did we make an effort to balance censoring across folds here?  we should comment on this}
+
+\note{also, hmm, censoring rate isn't actually fixed at, say, 30\%...eh, maybe it's OK}
+
+\par As illustrated in Figure~\ref{Fig:stability}, at $n=100$ the standard approach can be problematic even at fairly modest censoring rates.
+For example, standard CV cannot be used with over 25\% of data sets when half of all observations are censored.
+Furthermore, even if the censoring rate is just 10\%, standard CV strongly limits the number of folds that can be included in the cross-validation process, becoming very unstable when the number of folds exceeds 10.
 
@@
   \includegraphics[height= 7cm ]{./figures/figure_3_new.png}
-    \caption{Comparing Stability}
+  \caption{\label{Fig:stability}Comparing Stability \note{This plot needs better y axis labels and caption; also, these are not percents.}}
 \end{figure}
@@
 \subsection {Leave-One-Out Cross Validation}
-	\par In this section, we compared Leave-One-Out Cross Validation (LOOCV) with 10-fold cross validation.
-When first introduced, Verweij and Van Houwelingen's cross-validated partial likelihood was proposed so that LOOCV for Cox regression became feasible.
-While Verweij and Van Houwelingen's cross-validated partial likelihood solves the stability issue by aggregating more observations together, it does not seem to be an efficient way to evaluate the model fit based on results in Section 3.1.
-It tends to be conservative in variable selections and selects fewer variables in various scenarios.
-We further examined Verweij and Van Houwelingen's cross-validated partial likelihood under Leave-One-Out Cross Validation.
-In the following simulation studies, we used 10 fold cross-validated standard approach as our baseline for comparison.
-Then we conducted both 10-fold CV and LOOCV for Verweij and Van Houwelingen's approach, linear predictor approach and deviance residual approach.
-	\par Results of the simulation experiment is illustrated in Figure 4.For all three methods, LOOCV selected models that have smaller MSE than 10-fold CV for all three CV approaches.
-As is shown in left and middle panels, LOOCV based on deviance residuals and Verweij and Van Houwelingen's cross-validated partial likelihood still perform more conservatively than the standard approach.
-Even if Verweij and Van Houwelingen's approach was proposed to allow LOOCV, the LOOCV based on Verweij and Van Houwelingen's approach does not surpass the 10-fold standard approach.
+\label{Sec:loocv}
 
-    %\par We also compared cross-validation with information criteria AIC and BIC.
-Simulation studies were conducted in lower dimension and higher dimension respectively.
-As is illustrated in Figure 7, in lower dimension, AIC tends to perform more liberal and BIC tends to perform more conservative than the optimal $\lambda$.
-While the cross-validation methods are also conservative, most of them are less conservative than BIC.
-In terms of MSE, cross-validation methods yield smaller MSE than the information criteria.
-In higher dimension, both BIC and AIC tend to be more liberal than the optimal $\lambda$ and yield larger MSE.
+As discussed earlier, standard cross-validation cannot be used to carry out leave-one-out cross validation (LOOCV) for Cox regression models.
+The three other approaches (V\&VH, linear predictor, and deviance residual), however, are not subject to the stability issues described in Section~\ref{Sec:stability}, and can be used to implement LOOCV.
+Here, we compare these three approaches for carrying out LOOCV, and also compare LOOCV to 10-fold CV.
+
+For this study, we conducted both 10-fold CV and LOOCV for the V\&VH, linear predictor, and deviance residual approach, with standard 10-fold cross-validation used as a baseline for comparison.
+These results are shown in Figure~\ref{Fig:loocv}.
+\note{FYI, never write \texttt{Figure 4}; always use \texttt{\textbackslash ref}} For all three methods, LOOCV selected models with a smaller MSE than 10-fold CV.
+For the V\&VH and deviance residual approaches, however, LOOCV was still was inferior to standard 10-fold CV.
+As described in Section~\ref{Sec:stability}, these approaches tend to be conservative, selecting a larger-than-optimal value of the regularization parameter; LOOCV helps with this problem, but does not alleviate it entirely.
+
+For the linear predictor approach, on the other hand, \ldots \note{we need to talk about the linear predictor approach too!}
 
@@
   \includegraphics[height= 8cm ]{./figures/figure_4.png}
-    \caption{LOOCV}
+  \caption{\label{Fig:loocv}LOOCV\note{Needs a better caption; also, should y-axis be log mse ratio?}}
 \end{figure}
 
-  %\subsection {Simulations on Deviance Residuals and Baseline Hazard Estimations}
+\subsection {Cross-Validated C Index \note{Are we removing this section?}}
 
- \subsection {Cross-Validated C Index}
 \par In this section, we conducted simulation experiments to compare likelihood-based cross validation against AUC-based cross validation.

### 2019-03-06 main.tex
Edits/comments: Section 3.4

@@
 \subsection{Cross Validated Linear Predictors}
+\label{Sec:linear-predictor}
 %%%%%%%%%%%%%%%%%%%%
@@
 
-As in Figure~\ref{Fig:mse-brier-c}, the standard and linear predictor approaches outperform the deviance residual and $V\&VH$ approaches.
+As in Figure~\ref{Fig:mse-brier-c}, the standard and linear predictor approaches outperform the deviance residual and V\&VH approaches.
 The improvement is particularly noticeable in the most difficult setting: high-dimensional data with weak signal.
-Table~\ref{Tab:sim} also provides some insight into why the deviance residual and $V\&VH$ approaches perform poorly relative to standard CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and selecting fewer variables.
+Table~\ref{Tab:sim} also provides some insight into why the deviance residual and V\&VH approaches perform poorly relative to standard CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and selecting fewer variables.
 \note{if we're going to say something about ``selecting fewer variables'', should Table 1 include the average model size?} \bnote{I'm afraid that the average model size would be confusing to the readers.
@@
 
-\subsection {Cross-Validated C Index \note{Are we removing this section?}}
+\subsection {Cross-Validated C Index}
 
-\par In this section, we conducted simulation experiments to compare likelihood-based cross validation against AUC-based cross validation.
-While the focus of this manuscript is to study valid cross-validation methods for partial-likelihood-based models such as penalized Cox rgression, the methods proposed in this paper are not applicable to some other survival machine learning models that do not involve constructing a partial likelihood.
-In those cases, the cross-validated C Index, also called cross-validated AUC, is one of the most popular options \citep{Subramanian2011} \citep{Simon2011a}.
-As is proposed by Simon et al., cross-validated AUC is computed also based on cross-validated linear predictors or some other types of prognostic factors.
-Thus, it is of our curiosity to see which one has a better performance when those two approaches, likelihood-based and AUC-based, are both available after we obtain the cross-validated linear predictors.
+\par All of the cross-validation approaches described in Section~\ref{Sec:methods} use the Cox partial likelihood, in one way or another, as the criterion upon which model selection is based.
+In this section, we compare these likelihood-based approaches against an alternative approach of cross validation using concordance, also known as the cross-validated AUC (CV-AUC).
+This is a widely used approach to cross-validation with time-to-event outcomes, particularly for machine learning models that are not likelihood-based \citep{Subramanian2011,Simon2011a}.
+This approach typically proceeds by constructing cross-validated linear predictors, as in Section~\ref{Sec:linear-predictor}, but then using those predictors to calculate concordance as opposed to partial likelihood.
 
-\par We conducted simulations at both low dimension (p = 50) and high dimension (p = 1000), varying the signal in the data.
-Results of the simulations are illustrated in Figure 5.
-In general, the likelihood-based approach is more stable than the AUC-based cross validation approach.
-For MSE ratio and Brier Score, the likelihood approach tends to select more predictive models than the AUC-based approach.
-For the actual $\lambda$, it is interesting that the AUC-based approach is quite liberal and selected smaller $\lambda$s, which would yield models that contain more variables.
+\par We conducted simulations in both low dimensional ($p = 50$) and high dimensional ($p = 1000$) settings, varying the signal in the data \note{as in Section 3.1?} \note{is $n=100$ here?}.
+Results of the simulations are illustrated in Figure 5 \note{make this into a reference}.
+We used the linear predictor approach (CV-LP) here as the representative approach for likelihood-based cross-validation.
 
@@
   \includegraphics[height= 15cm ]{./figures/figure_auc.png}
-    \caption{Cross-Validated AUC}
+  \caption{Cross-Validated AUC \note{Needs a more complete caption}}
 \end{figure}
 
-%\begin{figure}[h]
-%    \centering
-%		\includegraphics[height= 15cm ]{./figures/figure_auc_extended.png}
-%    \caption{Cross-Validated AUC}
-%\end{figure}
+Several observations can be drawn from Figure~\ref{5}.
+First, likelihood-based CV tends to select more accurate models, as judged by both mean squared error and Brier score.
+Second, likelihood-based CV is also more stable than the AUC-based CV, in the sense that it tends to select a narrower range of models, with more consistent performance, than CV-AUC.
+Lastly, CV-AUC is consistently more liberal than likelihood-based CV, selecting models with smaller values of the regularization parameter $\lam$, and therefore preferring models with larger coefficient values and more nonzero coefficients.
 

### 2019-03-13 main.tex
2.1, 3.1, figure 2 and figure 5

@@
 %%%%%%%%%%%%%%%%%%%%
+\bnote{I rewrote this section}
 
 The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model.
-Hence, another alternative approach to cross-validation would be to obtain an estimate $\hat{\Lambda}_{0}$ for the baseline hazard and use this in cross-validation.
-There are several potential approaches to estimating $\Lambda_0$; in this manuscript, we chose to obtain the cross-validated linear predictors $\hat{\eta}^{cv}$ as in \eqref{eq:cv-lp}, then estimate $\Lambda_0$ via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}, treating these linear predictors as fixed inputs.\note{Maybe this isn't critical, but I still think we could just use $\hat{\Lambda}_0$ from the full data fit \ldots that estimator just makes more sense to me}\bnote{They do act a bit different in simulations.
-The baseline built from the full data performed worse.
-Discuss this next meeting}
+Hence, another alternative approach to cross-validation would be to obtain an estimate $\hat{\Lambda}_{0}$ for the cumulative baseline hazard and use it to quantify the cross-validation error.
+The simplest way to estimate $\Lambda_0$ is via the Nelson-Aalen Estimator \citep{aalen1978} \citep{nelson1969}:
+$$
+	 \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk at time }t^{-}_j}
+$$.
+Based on $\hat{\Lambda}_{0}(t)$, we can obtain a cumulative hazard for each individual by including their covariates: $$  \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\hat\beta)$$.
+To incorporate the cross-validation procedure, we would first obtain $\hat{\beta}^{-k}$ from the training set $T_k$.
+For each $i \in D_k$,
 
-As an analogue to the residual sum of squares in linear regression, we considered using the sum of squared deviance residuals, the normalized form of Martingale residuals, to measure the model's predictive accuracy \citep{Therneau1990}.
+\begin{equation}
+ \label{eq:cv-bhz}
+  \hat{\Lambda}^{cv}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\hat\beta^{-k})
+\end{equation}
+, where the original $\hat{\beta}$ from the full fit is now replaced by $\hat{\beta}^{-k}$.
+
+A natural candidate that incorporates the cumulative hazard into a form of a loss function is via the deviance residual, which is the normalized form of the Martingale residual \citep{Therneau1990}.
 First proposed for model diagnostics, deviance residuals have also been used as a predictive accuracy measure in recursive partitioning models for survival data \citep{Therneau2018}.
-Given that the baseline hazard and linear predictors are obtained via cross-validation, we can first obtain the cross-validated Martingale residual:
+Given \eqref{eq:cv-bhz}, we can first obtain the cross-validated Martingale residual:
 	\begin{equation}
-	\hat{M^{cv}_{i}} = \delta_{i} -\hat{\Lambda}^{cv}_{0}(t_{i})e^{\hat{\eta}^{cv}_{i}},
+	\hat{M^{cv}_{i}} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\hat\beta^{-k}),
 	\end{equation} 
@@
 	\end{equation}
-The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$ are used as the cross validated error. We refer to this method as the \emph{cross-validated deviance residuals} approach.
+The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$ are used as the cross validated error as an analogy to the residual sums of square for linear regression.
+We refer to this method as the \emph{cross-validated deviance residuals} approach.
+
+Our baseline estimation in the \emph{cross-validated deviance residuals} slightly deviates from the baseline estimation used in the original version of the Martingale and Deviance residuals \citep{Therneau1990}.
+Instead of using the Nelson-Aalen Estimator, the baseline cumulative hazard in the original Martingale and Deviance residuals is estimated via the Breslow's Estimator \citep{breslow1972}, which adjusts covariates of the fitted model in the baseline estimation.
+Here we provide a heuristic justification for our modification.
+The deviance residual measures the difference between the fitted model against the saturated model, which is considered as a perfect fit to the data.
+In the Cox Regression context, the baseline cumulative hazard is part of the saturated model.
+With the Breslow's Estimator or other estimators that involve the covariates, this saturated model would change everytime when a new model is fitted.
+This is not an issue when these two residuals are used in the model diagnositics context.
+However, in the model selection context, the sum of squared deviance residuals would have difficulty to tell which model is the best fit for the data if the Breslow's Esimator is used.
+\bnote{(This is not the case for Brier Score and Kullback Leibler Score.
+We should still use the covariate adjusted survival function for these two, because hey are comparing the predicted probability of survival against the observed probability of survival)}
 
@@
 Simulation studies were conducted to compare how those methods behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for LASSO penalized regression.
-Survival times were generated from exponential distribution $h(t) = h_{0} \exp(X\beta)$, where $h_{0}$ is the prespecified baseline\note{huh?
-so what is $h_0$?} \bnote{$h_0$ is 1 in all simulations.
-Basically, I first generated $X$, then the linear predictor $X\beta$; Then I compute the rate parameter from $h_0exp(X\beta)$ and inserted this to the \texttt{rexp} function to generate the survival times}; the covariate matrix $X$ with $n$ observations and $p$ covariates is generated from independent $\text{Normal}(0, 1)$ samples; and the coefficient vector $\beta$ is sparse, with most entries equal to zero.
-Censoring status indicators were drawn independently from the binomial distribution.
+\bnote{Covariate matrix $X$ with $n$ observations and $p$ covariates is generated from independent $\text{Normal}(0, 1)$ samples; and coefficient vector $\beta$ is sparse, with most entries equal to zero.
+Given $X$ and $\beta$, survival times were generated from exponential distribution with rate parameter set to be $\exp(X\beta)$.} Censoring status indicators were drawn independently from the binomial distribution.
 All simulations were implemented in R \citep{R}.
@@
 Mean squared error $MSE$ is defined to be
-\bnote{
+%\bnote{
 \begin{equation}
@@
 \end{equation}
-}
-\note{this is the equation for a scalar MSE, not a vector}
+%}
+%\note{this is the equation for a scalar MSE, not a vector}
 
@@
 \end{equation}
-where $\lambda^{cv}$ denotes the value of $\lambda$ selected by cross-validation.   These results are then averaged over N replications; the $\log$ of the ratio is taken to correct for skewness.
+where $\lambda^{cv}$ denotes the value of $\lambda$ selected by cross-validation.
+These results are then averaged over N replications; the $\log$ of the ratio is taken to correct for skewness.
+\bnote{clarification: MSE is estimated by avevaring across N replications respectively for $\hat{\beta}(\lambda^{cv})$ and $\hat{\beta}^{oracle}$.
+The ratio and $\log$ were taken afterwards }
 
@@
 
-For each simulated data set, there were $n = 120$ observations, $p = 1000$ covariates.
-and 10\% of the observations were censored.
+\bnote{I rewrote this paragraph} For each simulated data set, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates.
+The covariate vector $\beta$ is assumed to have sparsity, where $\beta_j = 0$ for $j \in \{11,12,..,1000\}$.
+Then for $j \in \{1,2, ..., 10 \}$, $\beta_j = s$, where $s$ is a prespecified constant that finetunes the signal strength in the simulated data set.
+In the first simulation study, we varied $s$ from 0.4 to 0.9.
+The survival outcomes have 10$\%$ expected censoring.
 For each of the four methods described in Section~\ref{Sec:methods}, 10-fold cross-validation was implemented.
-Number of non-zero $\beta$s was 10.
-The signal in the data set was varied by varying the magnitude of the $\beta$s from 0.4 to 0.9.
-\note{Would it be better here to descibe the $\beta$ values explicitly, with an equation?} For each scenario, 200 replications were used.
+$N = 200$ replications were used for each scenario.
 
@@
   \centering
-  \includegraphics[height= 7cm ]{./figures/figure_2_new.png}
-  \caption{\label{Fig:mse-brier-c} Simulation comparing 4 CV methods in terms of $\lambda$ and MSE ratio (relative to the oracle model) \note{UPDATE CAPTION}. n = 150, p = 1000, 10$\%$ censoring.}
+  \includegraphics[height= 7cm ]{./manuscript_figure/figure_2_new.png}
+  \caption{\label{Fig:mse-brier-c} Simulation comparing 4 CV methods.
+The horizontal axis in all three plots is $s$, which represents the signal strength in the data.
+The value of $s$ is varied from 0.4 to 0.9.
+$\log(\text{MSE ratio})$, relative to the oracle modle is plotted in the left panel.
+Out-of-sample Brier Scores is plotted in the middle.
+Out-of-sample Harrell's C Index is plotted in the right panel.
+For each simulated data set, n = 120, p = 1000.
+Expected censoring percentage is 10$\%$.
+\bnote{behavior of the deviance residual approach has slightly changed}}
 \end{figure}
@@
 Table~\ref{Tab:sim} also provides some insight into why the deviance residual and V\&VH approaches perform poorly relative to standard CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and selecting fewer variables.
-\note{if we're going to say something about ``selecting fewer variables'', should Table 1 include the average model size?} \bnote{I'm afraid that the average model size would be confusing to the readers.
+%\note{if we're going to say something about ``selecting fewer variables'', should Table 1 include the average model size?} \bnote{I'm afraid that the average model size would be confusing to the readers.
 The most conversative model would have the smallest model size, and would be closest to the true model size}
@@
 
-\subsection {Leave-One-Out Cross Validation}
+\subsection {Leave-One-Out Cross Validation \bnote{Updating Simulation}}
 \label{Sec:loocv}
@@
   \centering
-  \includegraphics[height= 15cm ]{./figures/figure_auc.png}
-  \caption{Cross-Validated AUC \note{Needs a more complete caption}}
+  \includegraphics[height= 15cm ]{./manuscript_figure/figure_auc.png}
+  \caption{\bnote{caption edited}Comparing Partial Likelihood and C Index both built from Cross-validated Linear Predictors.
+The horizontal axis represents the strength of signal $s$ in the data.
+The left figures are under the low dimension scenarios where p = 50 and n = 150.
+The right figures are under the high dimension scenarios where p = 1000 and n = 120.
+The expected percentages of censoring for both scenarios are 30 $\%$.
+The top row are box plots of the selected tuning parameter$\lambda$.
+The mid row plots the $\log(\text{SE Ratio})$ \bnote{(it just occured to me, this may not exactly be the log of MSE ratio.
+because to get MSE, we need to take average over all replications.
+But this figure is plotting each replication...)}, relative to the oracle model.
+The bottom row plots the Brier Scores.
+}
 \end{figure}

### 2019-03-13 manuscript_code/figure_2_git.R

@@ -1,0 +1,299 @@
+#------------------------------
+# Self-contained R code
+# to reproduce simulation results
+# in Figure 2 of the Cross Validation 
+# Manuscript
+# scenario: increasing signal
+# in n = 120, p = 1000
+#------------------------------
+
+library(survival)
+library(ncvreg)
+
+#-------------------------------------------------------------------------
+# functions
+
+generatebeta <- function(p,nbeta,c = 1){
+  pos <- rep(c, ceiling(nbeta/2))
+  neg <- rep(-c, floor(nbeta/2))
+  beta <- c(pos,neg,rep(0,p-nbeta))
+}
+
+genSurvData <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,h = 1 # 70% means 70% of the data are censored; 30% are observed
+){
+  p <- length(beta)
+  ## Generate column vector of coefficients
+  beta <- matrix(beta, ncol = 1)
+  x <- matrix(rnorm(n*p, sd = sd),n,p)
+  #standardize x first
+  x <- ncvreg::std(x)
+  hazard <- h*exp(x %*% beta)
+  ## Calculate survival times
+  time <- rexp(n, hazard)
+  ## observations are censored ("event = rbinom(n, 1, 0.30)")
+  y <- Surv(time = time, event = rbinom(n, 1, 1-cpct))
+  return(list(x=x,y=y,beta = beta))
+}
+
+
+# Brier score for selected lambda
+brierSurv_lambda <- function(data,time,fit,lambda) {
+  predicted <- apply(data$x,1, 
+                     function(x)
+                     {predict(fit, x,type = "survival",
+                              lambda = lambda)(time)})
+  is_obs_after <- data$y[,1] > time
+  mean((is_obs_after - predicted)^2)
+}
+
+# Kullback-Leibler for selected lambda
+KL_lambda <- function(data,time,fit,lambda) {
+  predicted <- apply(data$x,1, 
+                     function(x)
+                     {predict(fit, x,type = "survival",
+                              lambda = lambda)(time)})
+  is_obs_after <- data$y[,1] > time
[…]

### 2019-03-13 manuscript_code/figure_5_git.R

@@ -1,0 +1,276 @@
+#------------------------------
+# Self-contained R code
+# to reproduce part of the simulations
+# in Figure 5 of the Cross Validation 
+# Manuscript:
+# Explored low dimension p =50
+#------------------------------
+
+library(survival)
+library(ncvreg)
+
+#-------------------------------------------------------------------------
+# functions
+
+generatebeta <- function(p,nbeta,c = 1){
+  pos <- rep(c, ceiling(nbeta/2))
+  neg <- rep(-c, floor(nbeta/2))
+  beta <- c(pos,neg,rep(0,p-nbeta))
+}
+
+genSurvData <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,h = 1 # 70% means 70% of the data are censored; 30% are observed
+){
+  p <- length(beta)
+  ## Generate column vector of coefficients
+  beta <- matrix(beta, ncol = 1)
+  x <- matrix(rnorm(n*p, sd = sd),n,p)
+  #standardize x first
+  x <- ncvreg::std(x)
+  hazard <- h*exp(x %*% beta)
+  ## Calculate survival times
+  time <- rexp(n, hazard)
+  ## observations are censored ("event = rbinom(n, 1, 0.30)")
+  y <- Surv(time = time, event = rbinom(n, 1, 1-cpct))
+  return(list(x=x,y=y,beta = beta))
+}
+
+# Brier score for selected lambda
+brierSurv_lambda <- function(data,time,fit,lambda) {
+  predicted <- apply(data$x,1, 
+                     function(x)
+                     {predict(fit, x,type = "survival",
+                              lambda = lambda)(time)})
+  is_obs_after <- data$y[,1] > time
+  mean((is_obs_after - predicted)^2)
+}
+
+
+#-------------------------------------------------------------
+N <- 200
+n <- 150
+p <- 50
+c <- c(0.2,0.3,0.4,0.5,0.6,0.7)
+nbeta <- 5
+nfold <- 10
+foldid <- numeric(n)
+
+beta <- matrix(NA,nrow = length(c), ncol = p)
+for (i in 1:length(c)){
+  beta[i,]<- generatebeta(p,nbeta = nbeta ,c = c[i])
+}
+
+# initiation
+ncv <- array(NA,c(3,N,length(c)))
[…]

### 2019-03-13 main.tex
update table 1 and figure 4

@@
 \midrule
-150 & 50 & 5 & low & V VH  & 0.095 (0.028) & 1.457 (0.954) & 0.242 (0.013) & 0.678 (0.029) & 0.617 (0.034) \\
-    &    &   &     & Standard  & 0.087 (0.030) & 1.459 (0.922) & 0.242 (0.013) & 0.678 (0.029) & 0.618 (0.032) \\
-    &    &   &     & LinearPred  & 0.088 (0.025) & 1.433 (0.927) & 0.241 (0.013) & 0.677 (0.028) & 0.620 (0.029) \\
-    &    &   &     & DevResid  & 0.132 (0.032) & 1.676 (1.019) & 0.248 (0.014) & 0.689 (0.030) & 0.601 (0.044) \\
+150 & 50 & 5 & weak & V VH  & 0.098 (0.030) & 1.561 (0.851) & 0.245 (0.014) & 0.685 (0.030) & 0.610 (0.035) \\
+    &    &   &     & Standard  & 0.088 (0.025) & 1.531 (0.838) & 0.244 (0.013) & 0.683 (0.029) & 0.614 (0.029) \\
+    &    &   &     & LinearPred  & 0.092 (0.029) & 1.543 (0.846) & 0.245 (0.014) & 0.684 (0.030) & 0.612 (0.034) \\
+    &    &   &     & DevResid  & 0.098 (0.029) & 1.547 (0.848) & 0.245 (0.014) & 0.684 (0.030) & 0.611 (0.035) \\
 \addlinespace
-150 & 50 & 5 & strong & V VH  & 0.070 (0.011) & 1.733 (0.922) & 0.177 (0.012) & 0.527 (0.030) & 0.750 (0.014) \\
-    &    &   &     & Standard  & 0.066 (0.012) & 1.688 (0.909) & 0.176 (0.012) & 0.526 (0.029) & 0.749 (0.014) \\
-    &    &   &     & LinearPred  & 0.067 (0.011) & 1.695 (0.925) & 0.176 (0.012) & 0.526 (0.030) & 0.750 (0.014) \\
-    &    &   &     & DevResid  & 0.110 (0.012) & 2.339 (0.868) & 0.187 (0.014) & 0.554 (0.033) & 0.750 (0.018) \\
+150 & 50 & 5 & strong & V VH  & 0.071 (0.010) & 1.586 (0.989) & 0.176 (0.011) & 0.526 (0.027) & 0.750 (0.012) \\
+    &    &   &     & Standard  & 0.066 (0.011) & 1.544 (0.971) & 0.176 (0.011) & 0.526 (0.028) & 0.749 (0.013) \\
+    &    &   &     & LinearPred  & 0.067 (0.010) & 1.540 (0.981) & 0.176 (0.011) & 0.525 (0.028) & 0.750 (0.012) \\
+    &    &   &     & DevResid  & 0.094 (0.011) & 1.995 (0.841) & 0.181 (0.011) & 0.539 (0.026) & 0.751 (0.013) \\
 \addlinespace
-300 & 10000 & 20 & low & V VH  & 0.139 (0.021) & 3.463 (0.438) & 0.244 (0.015) & 0.681 (0.032) & 0.612 (0.051) \\
-    &    &   &     & Standard  & 0.105 (0.027) & 3.350 (0.458) & 0.228 (0.021) & 0.646 (0.045) & 0.639 (0.05) \\
-    &    &   &     & LinearPred  & 0.107 (0.026) & 3.360 (0.462) & 0.229 (0.021) & 0.648 (0.045) & 0.638 (0.049) \\
-    &    &   &     & DevResid  & 0.139 (0.027) & 3.454 (0.447) & 0.243 (0.018) & 0.678 (0.038) & 0.610 (0.058) \\
+400 & 10000 & 20 & weak & V VH  & 0.126 (0.019) & 2.958 (0.416) & 0.251 (0.012) & 0.696 (0.026) & 0.589 (0.044) \\
+    &    &   &     & Standard  & 0.103 (0.022) & 2.880 (0.437) & 0.244 (0.015) & 0.680 (0.031) & 0.603 (0.039) \\
+    &    &   &     & LinearPred  & 0.105 (0.023) & 2.886 (0.437) & 0.244 (0.015) & 0.681 (0.031) & 0.602 (0.039) \\
+    &    &   &     & DevResid  & 0.107 (0.022) & 2.895 (0.433) & 0.245 (0.014) & 0.683 (0.03) & 0.602 (0.039) \\
 \addlinespace
-500 & 10000 & 20 & strong & V VH  & 0.072 (0.005) & 3.469 (0.480) & 0.162 (0.015) & 0.495 (0.036) & 0.784 (0.021) \\
-    &    &   &     & Standard  & 0.056 (0.005) & 3.174 (0.504) & 0.153 (0.011) & 0.465 (0.029) & 0.782 (0.017) \\
-    &    &   &     & LinearPred  & 0.057 (0.005) & 3.204 (0.498) & 0.153 (0.011) & 0.466 (0.029) & 0.783 (0.017) \\
-    &    &   &     & DevResid  & 0.076 (0.005) & 3.545 (0.467) & 0.167 (0.015) & 0.508 (0.035) & 0.782 (0.022) \\
+400 & 10000 & 20 & strong & V VH  & 0.084 (0.012) & 3.799 (0.548) & 0.176 (0.03) & 0.528 (0.069) & 0.763 (0.048) \\
+    &    &   &     & Standard  & 0.059 (0.007) & 3.455 (0.578) & 0.154 (0.019) & 0.468 (0.049) & 0.780 (0.031) \\
+    &    &   &     & LinearPred  & 0.060 (0.007) & 3.479 (0.574) & 0.155 (0.019) & 0.470 (0.049) & 0.780 (0.030) \\
+    &    &   &     & DevResid  & 0.076 (0.018) & 3.708 (0.518) & 0.166 (0.024) & 0.502 (0.057) & 0.772 (0.049) \\
 \bottomrule
@@
 \label{Sec:stability}
-
+  \bnote{still working on a new simulation}
 \par In this section, we present a small simulation study to illustrate the numerical stability issues faced by standard cross-validation in the partial likelihood setting.
@@
 
-\subsection {Leave-One-Out Cross Validation \bnote{Updating Simulation}}
+\subsection {Leave-One-Out Cross Validation}
 \label{Sec:loocv}
 
+\bnote{Simulation Updated, still need to work on writing}
 As discussed earlier, standard cross-validation cannot be used to carry out leave-one-out cross validation (LOOCV) for Cox regression models.
@@
   \centering
-  \includegraphics[height= 8cm ]{./figures/figure_4.png}
+  \includegraphics[height= 8cm ]{./manuscript_figure/figure_4.png}
   \caption{\label{Fig:loocv}LOOCV\note{Needs a better caption; also, should y-axis be log mse ratio?}}

### 2019-03-13 manuscript_code/figure_4_git.R

@@ -1,0 +1,273 @@
+#------------------------------
+# Self-contained R code
+# to reproduce simulation results
+# in Figure 4: LOOCV
+# scenario: increasing signal
+# in n = 120, p = 1000
+#------------------------------
+
+library(survival)
+library(ncvreg)
+
+#-------------------------------------------------------------------------
+# functions
+
+generatebeta <- function(p,nbeta,c = 1){
+  pos <- rep(c, ceiling(nbeta/2))
+  neg <- rep(-c, floor(nbeta/2))
+  beta <- c(pos,neg,rep(0,p-nbeta))
+}
+
+genSurvData <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,h = 1 # 70% means 70% of the data are censored; 30% are observed
+){
+  p <- length(beta)
+  ## Generate column vector of coefficients
+  beta <- matrix(beta, ncol = 1)
+  x <- matrix(rnorm(n*p, sd = sd),n,p)
+  #standardize x first
+  x <- ncvreg::std(x)
+  hazard <- h*exp(x %*% beta)
+  ## Calculate survival times
+  time <- rexp(n, hazard)
+  ## observations are censored ("event = rbinom(n, 1, 0.30)")
+  y <- Surv(time = time, event = rbinom(n, 1, 1-cpct))
+  return(list(x=x,y=y,beta = beta))
+}
+
+
+#-------------------------------------------------------------------------
+# scenarios: 
+n <- 120
+p <- 1000
+N <- 200
+c <- c(0.6,0.8,1,1.2)
+nbeta <- 10
+nfold <- 10
+beta <- matrix(NA,nrow = length(c), ncol = p)
+for (i in 1:length(c)){
+  beta[i,]<- generatebeta(p,nbeta = nbeta ,c = c[i])
+}
+
+# initiation
+gp <- array(NA,c(2,N,length(c)))
+ug <- array(NA,c(2,N,length(c)))
+ncv <- array(NA,c(2,N,length(c)))
+dvr <- array(NA,c(2,N,length(c)))
+gp.loo <- array(NA,c(2,N,length(c)))
+ncv.loo <- array(NA,c(2,N,length(c)))
+dvr.loo <- array(NA,c(2,N,length(c)))
+oracle.mse <- matrix(NA, nrow = N, ncol = length(c))
+
+foldid <- numeric(n)
+
+
+for(i in 1:N){
+  set.seed(999 + i)
+  for (j in 1:length(c)){
[…]

### 2019-03-13 manuscript_code/table_1_git.R

@@ -1,0 +1,501 @@
+#------------------------------
+# Self-contained R code
+# to reproduce part of the simulations
+# in Table 1 of the Cross Validation 
+# Manuscript:
+# Explored 2 Scenarios here: 
+# low dim low signal
+# low dim high signal
+#------------------------------
+
+library(survival)
+library(ncvreg)
+
+#-------------------------------------------------------------------------
+# functions
+
+generatebeta <- function(p,nbeta,c = 1){
+  pos <- rep(c, ceiling(nbeta/2))
+  neg <- rep(-c, floor(nbeta/2))
+  beta <- c(pos,neg,rep(0,p-nbeta))
+}
+
+genSurvData <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,h = 1 # 70% means 70% of the data are censored; 30% are observed
+){
+  p <- length(beta)
+  ## Generate column vector of coefficients
+  beta <- matrix(beta, ncol = 1)
+  x <- matrix(rnorm(n*p, sd = sd),n,p)
+  #standardize x first
+  x <- ncvreg::std(x)
+  hazard <- h*exp(x %*% beta)
+  ## Calculate survival times
+  time <- rexp(n, hazard)
+  ## observations are censored ("event = rbinom(n, 1, 0.30)")
+  y <- Surv(time = time, event = rbinom(n, 1, 1-cpct))
+  return(list(x=x,y=y,beta = beta))
+}
+
+
+# Brier score for selected lambda
+brierSurv_lambda <- function(data,time,fit,lambda) {
+  predicted <- apply(data$x,1, 
+                     function(x)
+                     {predict(fit, x,type = "survival",
+                              lambda = lambda)(time)})
+  is_obs_after <- data$y[,1] > time
+  mean((is_obs_after - predicted)^2)
+}
+
+# Kullback-Leibler for selected lambda
+KL_lambda <- function(data,time,fit,lambda) {
+  predicted <- apply(data$x,1, 
+                     function(x)
+                     {predict(fit, x,type = "survival",
+                              lambda = lambda)(time)})
[…]

### 2019-03-20 main.tex
Edits to 2.3

@@
 %%%%%%%%%%%%%%%%%%%%
-\bnote{I rewrote this section}
 
 The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model.
-Hence, another alternative approach to cross-validation would be to obtain an estimate $\hat{\Lambda}_{0}$ for the cumulative baseline hazard and use it to quantify the cross-validation error.
-The simplest way to estimate $\Lambda_0$ is via the Nelson-Aalen Estimator \citep{aalen1978} \citep{nelson1969}:
-$$
-	 \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk at time }t^{-}_j}
-$$.
-Based on $\hat{\Lambda}_{0}(t)$, we can obtain a cumulative hazard for each individual by including their covariates: $$  \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\hat\beta)$$.
-To incorporate the cross-validation procedure, we would first obtain $\hat{\beta}^{-k}$ from the training set $T_k$.
-For each $i \in D_k$,
+Hence, another approach to cross-validation would be to include an extra step of estimating the baseline hazard and using it to quantify the cross-validation error.
+The cumulative baseline hazard $\Lambda_0$ is typically estimated via the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
+\begin{equation}
+  \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk at time }t^{-}_j}.
+\end{equation}
+\note{need to define $t^{-}_j$.}
+The Cox model implies the following relationship between the baseline cumulative hazard and the cumulative hazard for individual $i$:
+\begin{equation}
+  \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\hat\beta).
+\end{equation}
 
+The baseline hazard could be incorporated into cross-validation in a variety of ways.
+Here, we consider the following approach: for fold $k$, we obtain $\hat{\beta}^{-k}$ from the training set $T_k$ and then for each $i$ in the test set $D_k$,
 \begin{equation}
  \label{eq:cv-bhz}
-  \hat{\Lambda}^{cv}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\hat\beta^{-k})
+  \hat{\Lambda}^{cv}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\hat\beta^{-k}).
 \end{equation}
-, where the original $\hat{\beta}$ from the full fit is now replaced by $\hat{\beta}^{-k}$.
+Note that this uses an estimate of the baseline hazard coming from the full data under the null model, but cross-validation has been applied to the estimation of the regression coefficients $\hat{\beta}$; we will discuss the rationale for this estimator shortly.
 
-A natural candidate that incorporates the cumulative hazard into a form of a loss function is via the deviance residual, which is the normalized form of the Martingale residual \citep{Therneau1990}.
-First proposed for model diagnostics, deviance residuals have also been used as a predictive accuracy measure in recursive partitioning models for survival data \citep{Therneau2018}.
+A natural candidate for incorporating the cumulative hazard into a loss function is the deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}.
+First proposed for model diagnostics, deviance residuals have also been occasionally been used as a predictive accuracy measure \citep{Therneau2018}.
 Given \eqref{eq:cv-bhz}, we can first obtain the cross-validated Martingale residual:
 \begin{equation}
-	\hat{M^{cv}_{i}} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\hat\beta^{-k}),
+  \hat{M^{cv}_{i}} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\hat\beta^{-k}).
 \end{equation}
@@
 
-Our baseline estimation in the \emph{cross-validated deviance residuals} slightly deviates from the baseline estimation used in the original version of the Martingale and Deviance residuals \citep{Therneau1990}.
-Instead of using the Nelson-Aalen Estimator, the baseline cumulative hazard in the original Martingale and Deviance residuals is estimated via the Breslow's Estimator \citep{breslow1972}, which adjusts covariates of the fitted model in the baseline estimation.
-Here we provide a heuristic justification for our modification.
-The deviance residual measures the difference between the fitted model against the saturated model, which is considered as a perfect fit to the data.
-In the Cox Regression context, the baseline cumulative hazard is part of the saturated model.
-With the Breslow's Estimator or other estimators that involve the covariates, this saturated model would change everytime when a new model is fitted.
-This is not an issue when these two residuals are used in the model diagnositics context.
-However, in the model selection context, the sum of squared deviance residuals would have difficulty to tell which model is the best fit for the data if the Breslow's Esimator is used.
-\bnote{(This is not the case for Brier Score and Kullback Leibler Score.
-We should still use the covariate adjusted survival function for these two, because hey are comparing the predicted probability of survival against the observed probability of survival)}
+The baseline hazard estimate we describe here differs from the one typically used in constructing deviance residuals \citep{Therneau1990}.
+Instead of the Nelson-Aalen estimator, deviance residuals are typically calculated based on a baseline hazard estimate that has been adjusted for the covariates in the model such as the one described in \citet{breslow1972}.
+In the context of cross-validation, however, this approach is problematic.
+Deviance residuals measure the difference between the fitted model and a saturated model; in Cox regression, this saturated model depends on the baseline hazard.
+Thus, a covariate-adjusted baseline hazard would mean that each fold is compared to a different saturated model.
+For this reason, we consider it important that the baseline hazard remains constant across folds when calculating cross-validated deviance residuals; this intuition is borne out by simulations involving various other approaches \note{supplement?
+otherwise, this sentence is a little tricky}.
+Note that this is not the case for the Brier Score or Kullback Leibler Score (described in Section~\ref{Sec:accuracy}), which involve absolute loss functions, not comparisons with a saturated model.
 

### 2019-03-20 main.tex
Some edits to Section 3

@@
 
-Simulation studies were conducted to compare how those methods behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for LASSO penalized regression.
-\bnote{Covariate matrix $X$ with $n$ observations and $p$ covariates is generated from independent $\text{Normal}(0, 1)$ samples; and coefficient vector $\beta$ is sparse, with most entries equal to zero.
-Given $X$ and $\beta$, survival times were generated from exponential distribution with rate parameter set to be $\exp(X\beta)$.} Censoring status indicators were drawn independently from the binomial distribution.
+Simulation studies were conducted to compare how the methods described in Section~\ref{Sec:methods} behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for LASSO penalized regression.
+For these simulation studies, the covariate matrix $X$, consisting of $n$ observations and $p$ covariates, was generated from independent $\text{Normal}(0, 1)$ samples.
+Given $X$ and a coefficient vector $\beta$, survival times were generated from exponential distribution with rate parameter $\exp(X\beta)$.
+Censoring status indicators were drawn independently from the binomial distribution.
 All simulations were implemented in R \citep{R}.
@@
 \bnote{clarification: MSE is estimated by avevaring across N replications respectively for $\hat{\beta}(\lambda^{cv})$ and $\hat{\beta}^{oracle}$.
-The ratio and $\log$ were taken afterwards }
+The ratio and $\log$ were taken afterwards } \note{I'm not sure why this is a note; shouldn't you rewrite the sentence?}
 
@@
 \bnote{I rewrote this paragraph} For each simulated data set, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates.
-The covariate vector $\beta$ is assumed to have sparsity, where $\beta_j = 0$ for $j \in \{11,12,..,1000\}$.
-Then for $j \in \{1,2, ..., 10 \}$, $\beta_j = s$, where $s$ is a prespecified constant that finetunes the signal strength in the simulated data set.
+The covariate vector $\beta$ is assumed to be sparse, with $\beta_j = 0$ for $j \in \{11,12,..,1000\}$ and $\beta_j = s$ for $j \in \{1,2, ..., 10 \}$, where $s$ is a prespecified constant that parameterizes the signal strength in the simulated data set.
 In the first simulation study, we varied $s$ from 0.4 to 0.9.
@@
 $\log(\text{MSE ratio})$, relative to the oracle modle is plotted in the left panel.
-Out-of-sample Brier Scores is plotted in the middle.
-Out-of-sample Harrell's C Index is plotted in the right panel.
+Out-of-sample Brier scores are plotted in the middle.
+Out-of-sample Harrell's C index is plotted in the right panel.
 For each simulated data set, n = 120, p = 1000.
 Expected censoring percentage is 10$\%$.
-\bnote{behavior of the deviance residual approach has slightly changed}}
+\note{I really think the x axis should be Signal (s) or Signal strength (s)}}
 \end{figure}

### 2019-03-20 main.tex
Some edits to Section 4

@@
 
-\par In this section, we compared the performances of those cross validation methods when they are applied to data from real lung cancer study.
-\cite{shedden2008gene} conducted a large, retrospective study to validate prognostic gene expression signatures as predictors of overall survival in lung cancer patients, They performed gene expression profiling of \textit{442} non-small-cell lung cancer (NSCLC) patients.
-The data set contains gene expression measures from \textit{22,283} genes and basic clinical variables such as stage, age and gender.
-236 death events were observed and about 50$\%$ of the patients are censored.
+\par In this section, we apply the cross validation methods described in Section~\ref{Sec:methods} to data from a study of gene expression and survival in patients with lung cancer.
+\citet{shedden2008gene} conducted a large, retrospective study to validate prognostic gene expression signatures as predictors of overall survival in lung cancer patients.
+They gathered expression profiles of \textit{22,283} genes for \textit{442} \note{is there some reason 442 is in italics?
+other numbers aren't} non-small-cell lung cancer (NSCLC) patients along with a number of clinical variables such as stage, age and gender.
+\note{Did you only adjust for stage, age, and gender, or did you adjust for smoking, etc., as well?
+If not, maybe take out ``such as'', which implies that there are other variables you're not mentioning.} About 50\% of the patients died during the course of the study (236 events), with the rest censored.
 The clinical variables stage, age and gender were known to have a larger prognostic impact, thus they were fitted in the model without penalization.

### 2019-04-02 main.tex
edit 2.3, 3.1, change x label for all figures

@@
 \begin{equation}
-  \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk at time }t^{-}_j}.
+  \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}.
 \end{equation}
-\note{need to define $t^{-}_j$.}
+\bnote{I write out the $t^{-}_j$ definition directly into the fomula so that we don't need to define $t^{-}_j$ separately.}
 The Cox model implies the following relationship between the baseline cumulative hazard and the cumulative hazard for individual $i$:
@@
 A natural candidate for incorporating the cumulative hazard into a loss function is the deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}.
-First proposed for model diagnostics, deviance residuals have also been occasionally been used as a predictive accuracy measure \citep{Therneau2018}.
+First proposed for model diagnostics, deviance residuals have also been occasionally used as a predictive accuracy measure \citep{Therneau2018}.
 Given \eqref{eq:cv-bhz}, we can first obtain the cross-validated Martingale residual:
@@
 For this reason, we consider it important that the baseline hazard remains constant across folds when calculating cross-validated deviance residuals; this intuition is borne out by simulations involving various other approaches \note{supplement?
-otherwise, this sentence is a little tricky}.
+otherwise, this sentence is a little tricky}\bnote{sure}.
 Note that this is not the case for the Brier Score or Kullback Leibler Score (described in Section~\ref{Sec:accuracy}), which involve absolute loss functions, not comparisons with a saturated model.
@@
   \centering
-  \includegraphics[height= 12cm ]{./figures/figure_1_new.png}
+  \includegraphics[height= 12cm ]{./manuscript_figure/figure_1_new.png}
   \caption{Illustration of K-fold Cross Validation Methods for Cox Regression}
@@
 where $\lambda^{cv}$ denotes the value of $\lambda$ selected by cross-validation.
-These results are then averaged over N replications; the $\log$ of the ratio is taken to correct for skewness.
-\bnote{clarification: MSE is estimated by avevaring across N replications respectively for $\hat{\beta}(\lambda^{cv})$ and $\hat{\beta}^{oracle}$.
-The ratio and $\log$ were taken afterwards } \note{I'm not sure why this is a note; shouldn't you rewrite the sentence?}
-
-%For each generated data set, the $\lambda$ that has the minimal MSE is chosen as the optimal $\lambda$, which we denote by $\lambda^{opt}$.
-Then the models selected by the cross validation methods with $\lambda^{cv}$ would be compared to the model that corresponds to $\lambda^{opt}$.
-We would take the ratio between $MSE$ of $\hat{\beta}(\lambda^{cv})$ and $MSE$ of $\hat{\beta}(\lambda^{optimal})$ for each cross-validation selected $\lambda^{cv}$ in each generated data set.
-The results would be averaged over N replications.
+MSE is estimated by avevaring over N replications respectively for $\hat{\beta}(\lambda^{cv})$ and $\hat{\beta}^{oracle}$.
+The ratio is taken afterwards.
+The $\log$ is taken to correct for skewness.
+%\note{I'm not sure why this is a note; shouldn't you rewrite the sentence?}
 
@@
 
-\bnote{I rewrote this paragraph} For each simulated data set, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates.
+ For each simulated data set, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates.
 The covariate vector $\beta$ is assumed to be sparse, with $\beta_j = 0$ for $j \in \{11,12,..,1000\}$ and $\beta_j = s$ for $j \in \{1,2, ..., 10 \}$, where $s$ is a prespecified constant that parameterizes the signal strength in the simulated data set.

### 2019-04-02 manuscript_code/figure_2_git.R

@@ -257,7 +257,7 @@ df$method <- factor(df$method,levels = c("V & VH","Standard","LinearPred","DevRe
 lambda_p <- ggplot(data = df, aes(x = coef,y = lambda,group = method))+
   geom_line(aes(color=method),size=1.2,stat = 'identity')+
   geom_point(aes(color=method),size=3)+ theme_gray()+
-  labs(x = "s", 
+  labs(x = "signal strength (s)", 
        y = expression(lambda))+
   theme(legend.position = "none")
 mse_p <- ggplot(data = df, aes(x = coef,y = mse,group = method))+
@@ -265,7 +265,7 @@ mse_p <- ggplot(data = df, aes(x = coef,y = mse,group = method))+
   geom_point(aes(color=method),size=3)+
   theme_gray()+
   scale_colour_discrete(drop=TRUE,limits = levels(df$method))+
-  labs(x = "s", 
+  labs(x = "signal strength (s)", 
        y = "log(MSE Ratio)") +
   theme(legend.position = "none")
 Brier_p <- ggplot(data = df, aes(x = coef,y = Brier,group = method))+
@@ -273,7 +273,7 @@ Brier_p <- ggplot(data = df, aes(x = coef,y = Brier,group = method))+
   geom_point(aes(color=method),size=3)+
   theme_gray()+
   scale_colour_discrete(drop=TRUE,limits = levels(df$method))+
-  labs(x = "s", 
+  labs(x = "signal strength (s)", 
        y = "Brier Score") +
   theme(legend.position = "none")
 KL_p <- ggplot(data = df, aes(x = coef,y = KL,group = method))+
@@ -281,7 +281,7 @@ KL_p <- ggplot(data = df, aes(x = coef,y = KL,group = method))+
   geom_point(aes(color=method),size=3)+
   theme_gray()+
   scale_colour_discrete(drop=TRUE,limits = levels(df$method))+
-  labs(x = "s", 
+  labs(x = "signal strength (s)", 
        y = "Kullback Leibler")+
   theme(legend.position = "none")
 CIndex <- ggplot(data = df, aes(x = coef,y = CIndex,group = method))+
@@ -289,7 +289,7 @@ CIndex <- ggplot(data = df, aes(x = coef,y = CIndex,group = method))+
   geom_point(aes(color=method),size=3)+
   theme_gray()+
   scale_colour_discrete(drop=TRUE,limits = levels(df$method))+
-  labs(x = "s" , 
[…]

### 2019-04-02 manuscript_code/figure_4_git.R

@@ -263,7 +263,7 @@ p <- ggplot(df, aes(x=coef, y=mse_ratio, colour = methods)) +
   geom_line(aes(color = type), stat="identity",size = 1.2)+
   geom_point(aes(color = type), stat="identity",size = 3)+
   facet_grid(.~methods)+
-  labs(x = "s",
+  labs(x = "signal strength (s)",
          #expression(paste("Magnitude of Non-Zero ", beta, "s", sep = ""))
        y = "log(MSE Ratio)")+
   theme_gray()+

### 2019-04-02 manuscript_code/figure_5_git.R

@@ -270,7 +270,7 @@ df_combined$Type <- factor(df_combined$Type,levels = c("lambda","log(MSE Ratio)"
 auc_plot <- ggplot(df_combined, aes(Coef, Value,fill=Method)) + 
   geom_boxplot() + 
   facet_grid(Type ~ dim,scales = "free") +
-  labs(x = expression(paste("Magnitude of Non-Zero ", beta, "s", sep = ""))) +
+  labs(x = "signal strength (s)") +
   theme_grey()
 
 auc_plot

### 2019-04-03 main.tex
edit 3.1

@@
 Expected censoring percentage is 10$\%$.
-\note{I really think the x axis should be Signal (s) or Signal strength (s)}}
+%\note{I really think the x axis should be Signal (s) or Signal strength (s)}
+}
 \end{figure}
@@
 \label{Sec:stability}
-  \bnote{still working on a new simulation}
-\par In this section, we present a small simulation study to illustrate the numerical stability issues faced by standard cross-validation in the partial likelihood setting.
-The standard approach is both straightforward and enjoys very good predictive accuracy, as seen in Section~\ref{Sec:accuracy}.
-The main drawback to its widespread use is that the number of uncensored observations in the test set may be insufficient to construct a well-defined partial likelihood, leaving the overall estimate of cross-validation error undefined as well.
+ \bnote{I updated the writing and figure~\ref{Fig:stability} for this section}
 
-To illustrate how common this problem is, and what factors it depends on, we conducted a simulation study varying the number of folds and censoring rate.
-Throughout we fixed the number of observations at $n = 100$.
-When varying the censoring rate, we fixed the number of folds at $K = 10$; when varying the number of folds, we fixed the censoring rate at 10\%.
-We generated 200 independent data sets according to the mechanism described in Section~\ref{Sec:accuracy} and recorded whether or not standard cross-validation was able to be carried out (i.e., whether the partial likelihood was well-defined for all folds).
+The standard cross-validation in the partial likelihood setting is both straightforward and enjoys very good predictive accuracy, as seen in Section~\ref{Sec:accuracy}.
+However there are two main stability drawbacks to its widespread use.
+The following simulation studies illustrated those issues faced by the standard approach.
 
-\note{did we make an effort to balance censoring across folds here?  we should comment on this}
+The first stability issue with the standard approach is that the number of uncensored observations in the test set may be insufficient to construct a well-defined partial likelihood, leaving the overall estimate of cross-validation error undefined as well.
+When computing the partial likelihood, only individuals whose events are observed will contribute to the likelihood.
+If there is no event observed in a particular fold, then the partial likelihood for that fold is undefined.
+To illustrate how common this problem is, we conducted a simulation study.
+We generated 1000 independent data sets of $n = 100$ observations with fixed censoring rate.
+For each data set, we conducted a 10-fold cross-validation and recorded whether or not there was at least one event in each fold (i.e., whether the partial likelihood was well-defined for all folds).
+We repeated the same mechanism for scenarios with different censoring rates, ranging from $40\%$ to $80\%$.
+As illustrated in the left panel in Figure~\ref{Fig:stability}, at $n=100$ the standard approach can be problematic even at modest censoring rates.
+For example, standard CV cannot be used with over 20\% of data sets when $70\%$ of all observations are censored.
 
-\note{also, hmm, censoring rate isn't actually fixed at, say, 30\%...eh, maybe it's OK}
+One way could be used to get around the undefined partial likelihood for the standard approach is to balance the censoring status across all folds when making fold assignments.
+That is to make sure the number of events observed is about the same across all folds.
+Unless the number of the folds exceeds the number of observed events, such as the leave-one-out cross-validation, this approach will guarantee the partial likelihood in each fold is well-defined for the standard approach, thus the overall cross-validated error is well-defined.
 
-\par As illustrated in Figure~\ref{Fig:stability}, at $n=100$ the standard approach can be problematic even at fairly modest censoring rates.
-For example, standard CV cannot be used with over 25\% of data sets when half of all observations are censored.
-Furthermore, even if the censoring rate is just 10\%, standard CV strongly limits the number of folds that can be included in the cross-validation process, becoming very unstable when the number of folds exceeds 10.
+Even if the partial likelihood could be well-defined with balanced fold assignment, there is another drawback of the standard approach: it tends to have more uncertainty in its selection.
+The predictive performance of the model selected by this approach has greaterr variation than the other approaches, especially when there are only a few events available to construct the partial likelihood.
+
+We designed the following simulation study to illustrate this issue.
+We generated data sets with $n = 120$ observations.
+The censoring rate is fixed exactly at $50\%$, thus there are 60 events observed among all observations.
+For each data set, we conducted the balanced fold assiggment with number of folds set to be 10, 12, 15, 20, 30 and 60.
+When the number of folds was set to be 10, there were exactly 6 events in each fold; when the number of folds was set to be 60, there were one event in each fold.
+After each fold assignment, we computed the cross-validated error using all four approaches described in Section~\ref{Sec:accuracy}.
+We recorded the model that minimizes the cross-validated error for each approach and computed its squared error loss.
+We replicated the whole process for $N = 200$ time.
+In the end, instead of looking at the mean as in Section~\ref{Sec:accuracy}, we computed the standard deviation of the squared error loss for each approach.
+
+The results of this simulation are illustrated in the right panel in Figure~\ref{Fig:stability}.
+While all other three approaches are not affected by the number of folds and number of events in each fold, the variability of the model performance has drastically changed for the standard approach.
+When there are sufficient number of events in each fold, the standard approach perform pretty close to the other three approaches.
+However, when there is only one or two events in each fold, the standard approach has much more variability in its predictive performance.
+Even if we can still use the standard approach when there are few number of events, there is a bigger uncertainty in terms of how well the predictive performance would be.
+
+Based on the two simulations illustrated in this section, we would make the following recommendations: i) always balance the censoring status when conducting fold assignments for the standard approach, ii) avoid using the standard approach in conjuction with high number of folds, iii) avoid using the standard approach when there are few number of events observed.
 
@@
   \centering
-  \includegraphics[height= 7cm ]{./figures/figure_3_new.png}
-  \caption{\label{Fig:stability}Comparing Stability \note{This plot needs better y axis labels and caption; also, these are not percents.}}
+  \includegraphics[height= 7cm ]{./manuscript_figure/figure_3.png}
+  \caption{\label{Fig:stability} Simulations illustrating the stability issues with the standard approach.
+The figure on the left is generated with $n = 100$ observations.
+The horizontal axis is the censoring rate.
+The vertical axis is the proportion of undefined partial likelihood when the fold assignment is not balanced.
+The figure on the right is generated with $n = 120$.
+The fold assignment is balanced in this simulation so that the number of events in each fold is the same.
+The horizontal axis is the number of events per fold.
+The vertical axis is the standard deviation of the squared error loss.
+All four methods were compared in this simulation.
+The standard approach would have much greater variability when there are only few events in the fold.}
 \end{figure}

### 2019-04-03 main.tex
edit 3.3

@@
 
- For each simulated data set, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates.
-The covariate vector $\beta$ is assumed to be sparse, with $\beta_j = 0$ for $j \in \{11,12,..,1000\}$ and $\beta_j = s$ for $j \in \{1,2, ..., 10 \}$, where $s$ is a prespecified constant that parameterizes the signal strength in the simulated data set.
-In the first simulation study, we varied $s$ from 0.4 to 0.9.
-The survival outcomes have 10$\%$ expected censoring.
-For each of the four methods described in Section~\ref{Sec:methods}, 10-fold cross-validation was implemented.
-$N = 200$ replications were used for each scenario.
-
 \begin{figure}[ht]
@@
 
-%The left panel compares the $\lambda$ selected by different methods.
-Verweij and Van Houwelingen's cross-validated partial likelihood approach, which is currently most widely used approach, consistently selects $\lambda$s that are greater than the optimal one.
-Greater $\lambda$ leads to greater penalties and fewer non-zero variables, hence Verweij and Van Houwelingen's approach is conservative in terms of selecting variables.
-While the proposed cross-validated Deviance residual has very close performance as Verweij and Van Houwelingen's approach, the standard cross-validated log likelihood and the cross-validated linear predictor were more liberal than the other two methods.
-In general, all four methods tends to act conservatively and selects smaller models.
+ For each simulated data set, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates.
+The covariate vector $\beta$ is assumed to be sparse, with $\beta_j = 0$ for $j \in \{11,12,..,1000\}$ and $\beta_j = s$ for $j \in \{1,2, ..., 10 \}$, where $s$ is a prespecified constant that parameterizes the signal strength in the simulated data set.
+In the first simulation study, we varied $s$ from 0.4 to 0.9.
+The survival outcomes have 10$\%$ expected censoring.
+For each of the four methods described in Section~\ref{Sec:methods}, 10-fold cross-validation was implemented.
+$N = 200$ replications were used for each scenario.
 
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[height= 7cm ]{./manuscript_figure/figure_3.png}
+  \caption{\label{Fig:stability} Simulations illustrating the stability issues with the standard approach.
+The figure on the left is generated with $n = 100$ observations.
+The horizontal axis is the censoring rate.
+The vertical axis is the proportion of undefined partial likelihood when the fold assignment is not balanced.
+The figure on the right is generated with $n = 120$.
+The fold assignment is balanced in this simulation so that the number of events in each fold is the same.
+The horizontal axis is the number of events per fold.
+The vertical axis is the standard deviation of the squared error loss.
+All four methods were compared in this simulation.
+The standard approach would have much greater variability when there are only few events in the fold.}
+\end{figure}
+
 One way could be used to get around the undefined partial likelihood for the standard approach is to balance the censoring status across all folds when making fold assignments.
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[height= 7cm ]{./manuscript_figure/figure_3.png}
-  \caption{\label{Fig:stability} Simulations illustrating the stability issues with the standard approach.
-The figure on the left is generated with $n = 100$ observations.
-The horizontal axis is the censoring rate.
-The vertical axis is the proportion of undefined partial likelihood when the fold assignment is not balanced.
-The figure on the right is generated with $n = 120$.
-The fold assignment is balanced in this simulation so that the number of events in each fold is the same.
-The horizontal axis is the number of events per fold.
-The vertical axis is the standard deviation of the squared error loss.
-All four methods were compared in this simulation.
-The standard approach would have much greater variability when there are only few events in the fold.}
-\end{figure}
-
 \subsection {Leave-One-Out Cross Validation}
@@
 
-\bnote{Simulation Updated, still need to work on writing}
+\bnote{I edited this section and the figure caption}
+
 As discussed earlier, standard cross-validation cannot be used to carry out leave-one-out cross validation (LOOCV) for Cox regression models.
@@
 
-For this study, we conducted both 10-fold CV and LOOCV for the V\&VH, linear predictor, and deviance residual approach, with standard 10-fold cross-validation used as a baseline for comparison.
-These results are shown in Figure~\ref{Fig:loocv}.
-\note{FYI, never write \texttt{Figure 4}; always use \texttt{\textbackslash ref}} For all three methods, LOOCV selected models with a smaller MSE than 10-fold CV.
-For the V\&VH and deviance residual approaches, however, LOOCV was still was inferior to standard 10-fold CV.
-As described in Section~\ref{Sec:stability}, these approaches tend to be conservative, selecting a larger-than-optimal value of the regularization parameter; LOOCV helps with this problem, but does not alleviate it entirely.
-
-For the linear predictor approach, on the other hand, \ldots \note{we need to talk about the linear predictor approach too!}
-
 \begin{figure}[h]
@@
   \includegraphics[height= 8cm ]{./manuscript_figure/figure_4.png}
-  \caption{\label{Fig:loocv}LOOCV\note{Needs a better caption; also, should y-axis be log mse ratio?}}
+  \caption{\label{Fig:loocv}Simulation comparing LOO and 10-fold CV across three different approaches.
+With $N = 200$ replications, the data sets generated have dimension $n = 120$ and $p = 1000$.
+The horizontal axis is the signal strength.
+The verticle axis is the $\log$(MSE ratio).
+In each panel, the LOO version and 10-fold version of the same approach are plotted.
+The 10-fold standard approach is plotted as the blue line in all three panels as a baseline for comparison.
+}
 \end{figure}
+
+For this simulation experiment, we generated datasets with $n = 120$ and $p = 1000$, using the same data generating procedure as describted in Section~\ref{Sec:accuracy}.
+For each of the V\&VH, linear predictor, and deviance residual approaches, we conducted both 10-fold CV and LOOCV.
+We also conducted the standard 10-fold cross-validation as a baseline for comparison.
+We recorded the model that minimizes the cross-validated error  for each approach.
+After $N = 200$ replications, $\log$ of the MSE ratio was computed using the same definition described in Section~\ref{Sec:accuracy}.
+
+Results of the simulation are shown in Figure~\ref{Fig:loocv}.
+The panels illustrate the results for each CV approach, one with the 10-fold version and one with the leave-one-out version.
+The 10-fold standard approach was plotted in all three panels as a baseline for comparison.
+For all three approaches, the Leave-One-Out version selected models with a smaller $\log(\text{MSE ratio})$ than the 10-fold version.
+For the V\&VH and deviance residual approaches, however, LOOCV was still inferior to standard 10-fold CV.
+As described in Section~\ref{Sec:accuracy}, these approaches tend to be conservative, selecting a larger-than-optimal value of the regularization parameter; LOOCV helps with this problem, but does not alleviate it entirely.
+
+For the linear predictor approach, on the other hand, the LOOCV version surpasses the standard 10-fold CV by having smaller $\log(\text{MSE ratio})$ across all signal strength levels.
+At the same time, even the 10-fold version of the linear predictor approach also outperforms the leave-one-out version of the V$\&$VH approach and the deviance residual approach.
+That is to say, even though the leave-one-cross validation improves the predictive accuracy for V$\&$VH and deviance residual approach with a higher computation cost, the magnitude of improvement in predictive performance is smaller than switching to the linear predictor approach.
 

### 2019-04-10 main.tex
meeting 4/10/2019

@@
 \end{equation}
-\bnote{I write out the $t^{-}_j$ definition directly into the fomula so that we don't need to define $t^{-}_j$ separately.}
 The Cox model implies the following relationship between the baseline cumulative hazard and the cumulative hazard for individual $i$:
@@
 For all four metrics, the overall pattern is the same (KL scores not shown in Figure~\ref{Fig:mse-brier-c} due to space constraints, but they were very similar to Brier scores).
-Although the four cross-validation methods performed relatively similarly, the standard and linear predictor approaches had the best performance across all metrics: lower MSE, better Brier scores, and higher concordance.
+Although the four cross-validation methods performed relatively similarly, the standard and linear predictor approaches had the best performance across all metrics: lower MSE, better Brier scores, and higher Harrell's C Index.
 
@@
 \toprule
- n& p& $\#$ Non-Zero & Signal & Method & $\lambda$ & log(MSE Ratio) &Brier Score & Kullback Leibler & C Index\\
+ n& p& $\#$ Non-Zero & Signal & Method & $\lambda$ & log(MSE Ratio) &Brier Score & Kullback Leibler & Harrell's C Index\\
 \midrule
@@
 If there is no event observed in a particular fold, then the partial likelihood for that fold is undefined.
+
 To illustrate how common this problem is, we conducted a simulation study.
@@
 
-We designed the following simulation study to illustrate this issue.
+The following simulation study was designed to illustrate this issue.
 We generated data sets with $n = 120$ observations.
@@
 
+The 60-fold scenario in the last simulation study was also a corner example where there is only one event and one censored observation in each fold.
+In this scenario, it is common to run into cases where the uncensored event happens after the other event is censored.
+In this situation, the risk set at the time right before the event occurred would only contain this observation itself.
+Thus its contribution to the log partial likelihood would be 0, which is well-defined but not informative.
+This situation can happen regardless of the fold size and censoring rate.
+But it would matter more for the standard approach when there are fewer events in the fold.
+
 Based on the two simulations illustrated in this section, we would make the following recommendations: i) always balance the censoring status when conducting fold assignments for the standard approach, ii) avoid using the standard approach in conjuction with high number of folds, iii) avoid using the standard approach when there are few number of events observed.
@@
 As discussed earlier, standard cross-validation cannot be used to carry out leave-one-out cross validation (LOOCV) for Cox regression models.
-The three other approaches (V\&VH, linear predictor, and deviance residual), however, are not subject to the stability issues described in Section~\ref{Sec:stability}, and can be used to implement LOOCV.
+The other three approaches (V\&VH, linear predictor, and deviance residual), however, are not subject to the stability issues described in Section~\ref{Sec:stability}, and can be used to implement LOOCV.
 Here, we compare these three approaches for carrying out LOOCV, and also compare LOOCV to 10-fold CV.
@@
 
-For this simulation experiment, we generated datasets with $n = 120$ and $p = 1000$, using the same data generating procedure as describted in Section~\ref{Sec:accuracy}.
+For this simulation experiment, we generated datasets with $n = 120$ and $p = 1000$, using the same data generating procedure as described in Section~\ref{Sec:accuracy}.
 For each of the V\&VH, linear predictor, and deviance residual approaches, we conducted both 10-fold CV and LOOCV.
@@
 
-For the linear predictor approach, on the other hand, the LOOCV version surpasses the standard 10-fold CV by having smaller $\log(\text{MSE ratio})$ across all signal strength levels.
+On the other hand, the LOOCV version of the linear predictor approach surpasses the standard 10-fold CV by having smaller $\log(\text{MSE ratio})$ across all signal strength levels.
 At the same time, even the 10-fold version of the linear predictor approach also outperforms the leave-one-out version of the V$\&$VH approach and the deviance residual approach.
-That is to say, even though the leave-one-cross validation improves the predictive accuracy for V$\&$VH and deviance residual approach with a higher computation cost, the magnitude of improvement in predictive performance is smaller than switching to the linear predictor approach.
+Even though the leave-one-out cross validation improves the predictive accuracy for V$\&$VH and deviance residual approach, the magnitude of improvement in predictive performance is smaller than switching to the linear predictor approach.
 
 \subsection {Cross-Validated C Index}
+\bnote{unified the naming of C index in this manuscript.
+Use Harrell's C Index or C Index instead of other alternatives.
+This is the default concordance statistic that the survival package would return.}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[height= 15cm ]{./manuscript_figure/figure_auc.png}
+  \caption{\label{Fig:cvauc}Comparing Partial Likelihood and Harrell's C Index both built from Cross-validated Linear Predictors.
+The horizontal axis represents the strength of signal $s$ in the data.
+The left figures are under the low dimension scenarios where p = 50 and n = 150.
+The right figures are under the high dimension scenarios where p = 1000 and n = 120.
+The expected percentages of censoring for both scenarios are 30 $\%$.
+The top row are box plots of the selected tuning parameter$\lambda$.
+The mid row plots the $\log(\text{relative Squared Error Loss})$, with respect to the oracle model.
+The bottom row plots the Brier Scores.
+}
+\end{figure}
 
@@
 \par We conducted simulations in both low dimensional ($p = 50$) and high dimensional ($p = 1000$) settings, varying the signal in the data \note{as in Section 3.1?} \note{is $n=100$ here?}.
-Results of the simulations are illustrated in Figure 5 \note{make this into a reference}.
+Results of the simulations are illustrated in Figure~\ref{Fig:cvauc}.
 We used the linear predictor approach (CV-LP) here as the representative approach for likelihood-based cross-validation.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[height= 15cm ]{./manuscript_figure/figure_auc.png}
-  \caption{\bnote{caption edited}Comparing Partial Likelihood and C Index both built from Cross-validated Linear Predictors.
-The horizontal axis represents the strength of signal $s$ in the data.
-The left figures are under the low dimension scenarios where p = 50 and n = 150.
-The right figures are under the high dimension scenarios where p = 1000 and n = 120.
-The expected percentages of censoring for both scenarios are 30 $\%$.
-The top row are box plots of the selected tuning parameter$\lambda$.
-The mid row plots the $\log(\text{SE Ratio})$ \bnote{(it just occured to me, this may not exactly be the log of MSE ratio.
-because to get MSE, we need to take average over all replications.
-But this figure is plotting each replication...)}, relative to the oracle model.
-The bottom row plots the Brier Scores.
-}
-\end{figure}
 
@@
 \toprule
-						& $\lambda$ & log($\lambda$) & CV AUC & Brier Score \\ 
+						& $\lambda$ & log($\lambda$) & CV C-Index & Brier Score \\ 
 						\cline{2-5} \\ 

### 2019-04-10 simulation/minimum_working_example.R
minimum working example

@@ -1,0 +1,84 @@
+#------------------------------  
+# a minimal working example
+# to validate cv.ncvreg 
+#------------------------------
+library(survival)
+library(ncvreg)
+
+# generate data
+beta <- c(rep(1,5),rep(0,20))
+n <- 100
+p <- length(beta)
+x <- matrix(rnorm(n*p, sd = 1),n,p)
+x <- ncvreg::std(x)
+hazard <- exp(x %*% beta)
+time <- rexp(n, hazard)
+y <- Surv(time = time, event = rbinom(n, 1, 0.7))
+
+data <- list(y=y,x=x)
+
+# inital fit
+
+fit <- ncvsurv(data$x,data$y,penalty = "lasso")
+lambda <- fit$lambda
+
+# cross-validation
+nfold <- 10
+nevent <- sum(data$y[,2])
+sde <- sqrt(.Machine$double.eps)
+foldid <- numeric(n)
+foldid[data$y[,2] == 1] <-  ceiling(sample(1:nevent)/(nevent + sde) * nfold)
+foldid[data$y[,2] == 0] <-  ceiling(sample(1:(n-nevent))/(n - nevent + sde) * nfold)
+
+# use cv.ncvsurv:
+cv_ncv <- cv.ncvsurv(data$x,data$y,penalty = "lasso",
+                     fold = foldid, lambda = lambda,
+                     returnY = TRUE,returnX=TRUE)
+cv_ncv$fit$X[1:10,1:10]
+data$x[1:10,1:10]
+
+sort(cv_ncv$fit$X[,1])
+sort(data$x[,1])
+
+# # use Biyue's code:
+# eta.cv <- matrix(nrow = n, ncol = length(lambda))
+# l_diff <- matrix(nrow = nfold , ncol = length(lambda))
+# l_test <- matrix(nrow = nfold, ncol = length(lambda))
+# 
+# ind <- order(data$y[,1])
+# d <- data$y[ind,2]
+# 
+# for(f in 1:nfold){
+#   ifold <- foldid == f
+#   cvfit <- ncvsurv(cv_ncv$fit$X[!ifold,], data$y[!ifold,],
+#                    penalty = "lasso", lambda = lambda)
+#   cvlambda <- cvfit$lambda
+#   beta_train <- cvfit$beta
+#   
+#   # linear predictors
[…]

### 2019-04-17 main.tex
elasticnet for shedden data

@@
 \par All of the cross-validation approaches described in Section~\ref{Sec:methods} use the Cox partial likelihood, in one way or another, as the criterion upon which model selection is based.
-In this section, we compare these likelihood-based approaches against an alternative approach of cross validation using concordance, also known as the cross-validated AUC (CV-AUC).
-This is a widely used approach to cross-validation with time-to-event outcomes, particularly for machine learning models that are not likelihood-based \citep{Subramanian2011,Simon2011a}.
+In this section, we compare these likelihood-based approaches against an alternative approach of cross validation using concordance, also known as the cross-validated C Index.
+This is a widely used approach for conducting cross-validation with time-to-event outcomes, particularly for machine learning models \citep{Subramanian2011,Simon2011a}.
 This approach typically proceeds by constructing cross-validated linear predictors, as in Section~\ref{Sec:linear-predictor}, but then using those predictors to calculate concordance as opposed to partial likelihood.
 
-\par We conducted simulations in both low dimensional ($p = 50$) and high dimensional ($p = 1000$) settings, varying the signal in the data \note{as in Section 3.1?} \note{is $n=100$ here?}.
+\par We conducted simulations in both low dimensional ($p = 50$, $n = 150$) and high dimensional ($p = 1000$, $n = 120$) settings.
+For each of the scenario, we generated data from a spectrum of signal strength as in Section~\ref{Sec:accuracy}.
 Results of the simulations are illustrated in Figure~\ref{Fig:cvauc}.
@@
 
-Several observations can be drawn from Figure~\ref{5}.
+Several observations can be drawn from Figure~\ref{Fig:cvauc}.
 First, likelihood-based CV tends to select more accurate models, as judged by both mean squared error and Brier score.
@@
 Lastly, CV-AUC is consistently more liberal than likelihood-based CV, selecting models with smaller values of the regularization parameter $\lam$, and therefore preferring models with larger coefficient values and more nonzero coefficients.
+
+\bnote{(END)}
 

### 2019-05-01 main.tex
Added biological interpretation to real data analysis

@@
 
-\par In this section, we apply the cross validation methods described in Section~\ref{Sec:methods} to data from a study of gene expression and survival in patients with lung cancer.
+In this section, we apply the cross validation methods described in Section~\ref{Sec:methods} to data from a study of gene expression and survival in patients with lung cancer.
 \citet{shedden2008gene} conducted a large, retrospective study to validate prognostic gene expression signatures as predictors of overall survival in lung cancer patients.
@@
 If not, maybe take out ``such as'', which implies that there are other variables you're not mentioning.} About 50\% of the patients died during the course of the study (236 events), with the rest censored.
-The clinical variables stage, age and gender were known to have a larger prognostic impact, thus they were fitted in the model without penalization.
-Cross-validation was again used for the selection of the penalty term, which selects the gene expressions that are associated with patients' overall survival.
-The results of the analysis are also listed in Table 3.
+The clinical variables stage, age and gender were included in the model without penalization, as it is reasonable to assume that all three have an effect on survival.
 
-\par How the CV methods perform on real data reflects what is shown in simulation studies.
+For this analysis, we fit elastic net models to the data in addition to lasso models, as these models often offer improvement over the lasso when many correlated genes are involved \citep{Zou2005}.
+We used cross-validation to select the regularization parameter $\lam$ as well as the elastic net tuning parameter $\alpha$, where $\alpha=1$ corresponds to the lasso and $\alpha=0$ corresponds to ridge regression.
+\note{More on elastic net, justify why we focus on $\alpha=0.3$}.
+
+\note{This paragraph will probably need to change based on new results.}  How the CV methods perform on real data reflects what is shown in simulation studies.
 The standard CVE is the most liberal approach.
@@
   \includegraphics[height= 8cm ]{./figures/shedden.jpg}
-    \caption{Comparing Cross-Validated Error for Different Methods on Real Data Sets}
+  \caption{Comparing Cross-Validated Error for Different Methods on Real Data Sets \note{Change to $\alpha=0.3$?}}
 \end{figure}
@@
 \centering
-\caption{Application to Real Data Sets}
+\caption{Application to Real Data Sets\note{Change to $\alpha=0.3$?}}
 \label{my-label}
@@
 \end{table}
+
+For the elastic net model with $\lam$ and $\alpha$ selected by the cross-validated linear predictor approach, 81 genes were selected.
+Of those, 78 were well-annotated.
+We carried out an enrichment analysis of these 78 genes using ToppGene \citep{Chen2009c}; the five most significantly enriched molecular functions are listed in Table~\ref{Tab:go}.
+
+\begin{table}[ht]
+\centering
+\caption{\label{Tab:go} Five most significantly enriched molecular functions among genes selected by the linear predictor method with the elastic net ($\alpha=0.3$).
+The five terms have a false discovery rate of 3\% using the Benjamini-Hochberg method.
+\note{I'm not sure whether we should include Selected and Annotated or not...if we do, we'll have to describe them better.}}
+\begin{tabular}{llrrr}
+\toprule
+ID & Name & p & Selected & Annotated\\
+\midrule
+GO:0008134 & transcription factor binding & 0.00015 & 10 & 584\\
+GO:0035257 & nuclear hormone receptor binding & 0.00057 & 5 & 161\\
+GO:0019869 & chloride channel inhibitor activity & 0.00061 & 2 & 9\\
+GO:0003700 & DNA-binding transcription factor activity & 0.00068 & 14 & 1264\\
+GO:0005102 & signaling receptor binding & 0.00083 & 16 & 1601\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Overall, the results are consistent with our understanding of cancer: many of the genes selected by the elastic net model are involved in signaling and the binding of transcription factors and receptors.
+These are precisely the pathways that tend to be disrupted in cancer.
+Chloride channels have also recently been implicated in several cancers as well as therapeutic targets \citep{Peretti2015}.
+
+Of particular interest is the fact that three of the selected genes (PIK3R1, PDPK1, and ITGB1) are involved in the PTEN pathway.
+The PTEN gene is an important tumor suppressor; mutations in PTEN are among the most common genetic changes found in human cancers and have been linked to cancers of the lung, breast, prostate, brain, and more.
+Time-to-event analyses like this one of gene expression data provide valuable insights for developing gene signatures that can help to identify patients with more severe disruptions to important pathways such as PTEN.
+These signatures thereby inform prognosis and impact treatment decisions.
 

### 2019-05-08 main.tex
update data analysis section

@@
 \subsection {Cross-Validated C Index}
+\label{Sec:CIndex}
 \bnote{unified the naming of C index in this manuscript.
@@
 
-\bnote{(END)}
-
 \section{Application to Real Data}
 
-In this section, we apply the cross validation methods described in Section~\ref{Sec:methods} to data from a study of gene expression and survival in patients with lung cancer.
+In this section, we demonstrate the usage and performance of the aforementioned cross validation methods when fitting penalized regression models on data from a study of gene expression and survival in patients with lung cancer.
 \citet{shedden2008gene} conducted a large, retrospective study to validate prognostic gene expression signatures as predictors of overall survival in lung cancer patients.
-They gathered expression profiles of \textit{22,283} genes for \textit{442} \note{is there some reason 442 is in italics?
-other numbers aren't} non-small-cell lung cancer (NSCLC) patients along with a number of clinical variables such as stage, age and gender.
-\note{Did you only adjust for stage, age, and gender, or did you adjust for smoking, etc., as well?
-If not, maybe take out ``such as'', which implies that there are other variables you're not mentioning.} About 50\% of the patients died during the course of the study (236 events), with the rest censored.
+They gathered expression profiles of 22,283 genes for 442 non-small-cell lung cancer (NSCLC) patients along with some additional clinical and demographic data.
+About $50\%$ of the patients died during the course of the study (236 events), with the rest censored.
 The clinical variables stage, age and gender were included in the model without penalization, as it is reasonable to assume that all three have an effect on survival.
 
+\begin{table}[h]
+\centering
+\begin{tabular}{ccccccccccc}
+\toprule
+ & \multicolumn{2}{c}{$\alpha = 0.3$ } & \multicolumn{2}{c}{$\alpha = 0.5$} & \multicolumn{2}{c}{$\alpha = 0.7$} & \multicolumn{2}{c}{$\alpha = 0.9$} & \multicolumn{2}{c}{$\alpha = 1$} \\
+ &      $\lambda$      &   CVE         &   $\lambda$        &     CVE       &      $\lambda$      &     CVE       &    $\lambda$        &    CVE        &     $\lambda$      &    CVE       \\
+
+\hline\\
+
+ V$\&$VH  & 0.363 &  296.32  &  0.228   &  296.35  & 0.163 &  296.34  &  0.133  &   296.31  &  0.119  &  296.30  \\
+ Standard & 0.301 & 145.17  & 0.218  &  145.38 & 0.163  & 145.37 &  0.127 &  145.35 & 0.114   & 145.34 \\
+ Linear Pred & 0.288 & 10.661  & 0.181  & 10.673 & 0.163 & 10.673    & 0.127  & 10.672  &    0.114 &  10.672  \\
+ Dev Resid & 0.316 & 571.40 & 0.228  &  571.96 & 0.163 & 571.95  & 0.133  &   571.79 &  0.119  &  571.70  \\
+\bottomrule
+\end{tabular}
+\caption{\bnote{Not sure whether to include this table or not}}
+\end{table}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{lcccccc}
+  \toprule
+  & $\lambda$ & log($\lambda$) & $\#$ of Genes Selected & CV C-Index & Brier Score \\ 
+\cline{2-6} \\ 
+V $\&$ VH& 0.363 & -1.013 &  & 0.639 & 0.243\\
+standard & 0.301 & -1.200 &  & 0.647& 0.232\\
+Linear Pred &0.288 & -1.246 &  & 0.650&  0.229\\
+Dev Resid & 0.316 & -1.153 &  & 0.645& 0.235\\ 
+\bottomrule
+\end{tabular}
+\caption{\label{tb:shedden}Results of elasticnet model with $\alpha = 0.3$ on the Lung Cancer Data Set.
+The linear Predictor approach picked the smallest $\lambda$.
+The corresponding model yields the highest CV C-Index and smallest Brier Score evaluated at median survvial time \bnote{I got different number of genes for LP approach...}}
+\end{table}
+
 For this analysis, we fit elastic net models to the data in addition to lasso models, as these models often offer improvement over the lasso when many correlated genes are involved \citep{Zou2005}.
-We used cross-validation to select the regularization parameter $\lam$ as well as the elastic net tuning parameter $\alpha$, where $\alpha=1$ corresponds to the lasso and $\alpha=0$ corresponds to ridge regression.
-\note{More on elastic net, justify why we focus on $\alpha=0.3$}.
+Besides the regularization parameter $\lam$, the elastic net model has an extra tuning parameter $\alpha$, where $\alpha=1$ corresponds to lasso and $\alpha=0$ corresponds to ridge regression.
+For each $\alpha = 0.3, 0.5, 0.7, 0.9 \text{ and } 1$, we selected a regularization parameter $\lam$ that has the lowest cross-validation error.
+The standard approach, linear predictor approach and deviance approach all reach the lowest cross-validation error at $\alpha = 0.3$.
 
-\note{This paragraph will probably need to change based on new results.}  How the CV methods perform on real data reflects what is shown in simulation studies.
-The standard CVE is the most liberal approach.
-Deviance Residual and Verweij and Van Houwelingen's CVE approaches were both really conversative, while the standard approach was the second liberal to standard CVE.
-In Figure 5, the Cross Validated Error is rescaled and plotted for all four methods.
-A $\lambda$ will be selected when CVE curve reaches its lower point.
-The blue line, which represents the linear predictor approach, has more curvature near its lowest point.
-It is easier to pick out the minimum point for this blue curve.
-Hence this approach is better at picking out signals than the other three approaches.
-For Verweij and Van Houwelingen's cross-validated partial likelihood, there's almost no signal there.
+We report the predictive performance for all four models at $\alpha = 0.3$ and their corresponding $\lambda$s in Table~\ref{tb:shedden}.
+For each model, we computed the cross-validated C-Index and Brier Scores at predicting the median survival time.
+How the CV methods perform on real data reflects what is shown in simulation studies.
+The standard approach and the linear predictor approach, selecting smaller $\lambda$ values, were the more liberal, while Verweij and Van Houwelingen's CVE approaches were both more conversative.
+With the highest C Index and lowest Brier Score, the linear predictor approach has best predictive performance among all four methods.
+The same patterns are repeated when $\alpha$ is at different values.
+
+We compared the CVE along the solution path for different methods.
+In Figure~\ref{fig:CVE}, the Cross Validated Error is rescaled and plotted for all four methods.
+A $\lambda$ is selected when CVE curve reaches its lower point.
+The blue line and green line, which represent the linear predictor approach and standard approah, have more curvature near theirs lowest point, making it is easier to pick out the minimum point.
 
@@
   \centering
-  \includegraphics[height= 8cm ]{./figures/shedden.jpg}
-  \caption{Comparing Cross-Validated Error for Different Methods on Real Data Sets \note{Change to $\alpha=0.3$?}}
+  \includegraphics[height= 8cm ]{./manuscript_figure/shedden.jpg}
+  \caption{\label{fig:CVE}Rescaled Cross Validated Errors at $\alpha = 0.3$.
+The horizonal axis is $\log(\lambda)$.
+The vertical axis is the cross-validated error rescaled to a range from 0 to 1.
+After the rescaling, the Cross Validated Error of four different methods can be plotted together for comparison.
+}
 \end{figure}
-
-\begin{table}[h]
-\centering
-\caption{Application to Real Data Sets\note{Change to $\alpha=0.3$?}}
-\label{my-label}
-\begin{tabular}{lccccc}
-  \toprule
-  & $\lambda$ & log($\lambda$) & CV C-Index & Brier Score \\ 
-						\cline{2-5} \\ 
-						V $\&$ VH& 0.123 & -2.097 & 0.640 & 0.246\\
-						standard & 0.102 & -2.278 & 0.643& 0.239\\
-						Linear Pred &0.085 & -2.460 & 0.651&  0.227\\
-						Dev Resid & 0.127& -2.066 & 0.639& 0.246\\ 
-						\bottomrule
-\end{tabular}
-\end{table}
 
@@
 
+\bnote{(END)}
 \section{Discussion}

### 2019-05-14 cross-validation.tex
add package citations; change l,exp and log; a few more comments in blue
Edits to 3.1
clean/reply to some previous comments
edit application and discussion section

@@
 \newcommand{\CVE}{\textrm{CVE}}
+\newcommand{\MSE}{\textrm{MSE}}
 \newcommand{\logit}{\mbox{logit}}
@@
 \providecommand{\note}[1]{\textcolor{red}{#1}}
+\providecommand{\bnote}[1]{\textcolor{blue}{#1}}
 
@@
 Here, we propose two new cross-validation methods for Cox regression and compare them to approaches that have been proposed elsewhere.
-\note{IS THIS STILL TRUE?
-WE SHOULD EDIT THIS SENTENCE: Our simulation studies show that in general, all cross-validation approaches tend to be conservative (i.e., select smaller models than the optimal choice of tuning parameters) for penalized Cox regression models.} However, our proposed approach of cross-validating the linear predictors seems to offer an attractive balance of performance and numerical stability.
+%\note{IS THIS STILL TRUE?
+WE SHOULD EDIT THIS SENTENCE: Our simulation studies show that in general, all cross-validation approaches tend to be conservative (i.e., select smaller models than the optimal choice of tuning parameters) for penalized Cox regression models.}
+Our proposed approach of cross-validating the linear predictors seems to offer an attractive balance of performance and numerical stability.
 We illustrate these advantages using simulated data as well as real data from two genome-wide studies of survival in cancer patients.

### 2019-05-14 main.tex
Fixing number of genes (85 -> 81)
edit application and discussion section

@@
 This semiparametric flexibility, however, comes at a cost.
-The Cox regression approach can estimate relative risks, but without estimating the baseline hazard, it does not make any predictions concerning the absolute failure time for any given individual.
+The Cox regression model can estimate relative risks, but without estimating the baseline hazard, it does not make any predictions concerning the absolute failure time for any given individual.
 This poses a challenge to assessing the predictive accuracy of a given Cox regression model.
@@
 The estimation of the coefficients is obtained by maximizing a partial likelihood.
-Letting $t_i$ denote the time on study for subject $i$ and $\delta_{i}$ indicate whether or not an event is observed for subject $i$, each subject's individual contribution to that partial likelihood is
+Letting $t_i$ denote the time of being in the study for subject $i$ and $\delta_{i}$ indicate whether or not an event is observed for subject $i$, each subject's individual contribution to that partial likelihood is
 \begin{equation}
@@
 
-\par Both of the methods in Section~\ref{Sec:cox-cv-existing} consist of adding together partial likelihood measures from each fold.
+\par Both of the methods in Section~\ref{Sec:cox-cv-existing} consist of adding together log partial likelihood measures from each fold.
 An alternative method is to obtain linear predictors for each fold, then combine these linear predictors to calculate a partial likelihood.
@@
 All simulations were implemented in R \citep{R}.
-We varied the sample size, dimension, censoring percentage and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation.
 
@@
 
-Our first simulation study concerns the accuracy of the cross-validation approaches introduces in Section~\ref{Sec:methods}.  Accuracy was assesed via several criteria.
+Our first simulation study concerns the accuracy of the cross-validation approaches introduces in Section~\ref{Sec:methods}.
+We varied the sample size, dimension, censoring percentage and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation.
+Accuracy was assesed via several criteria.
 
@@
 Results were then averaged over the N replications.
-For both Brier and KL scores, a score represents smaller prediction error, thus indicating better prediction accuracy.
+For both Brier and KL scores, a smaller score represents smaller prediction error, thus indicating better prediction accuracy.
 
@@
 \label{Sec:stability}
- \bnote{I updated the writing and figure~\ref{Fig:stability} for this section}
 
@@
 The censoring rate is fixed exactly at $50\%$, thus there are 60 events observed among all observations.
-For each data set, we conducted the balanced fold assiggment with number of folds set to be 10, 12, 15, 20, 30 and 60.
-When the number of folds was set to be 10, there were exactly 6 events in each fold; when the number of folds was set to be 60, there were one event in each fold.
+For each data set, we conducted the balanced fold assignment with number of folds set to be 10, 12, 15, 20, 30 and 60.
+When the number of folds was set to be 10, there were exactly 6 events in each fold; when the number of folds was set to be 60, there was exactly one event in each fold.
 After each fold assignment, we computed the cross-validated error using all four approaches described in Section~\ref{Sec:accuracy}.
@@
 We replicated the whole process for $N = 200$ time.
-In the end, instead of looking at the mean as in Section~\ref{Sec:accuracy}, we computed the standard deviation of the squared error loss for each approach.
+In the end, we computed the standard deviation of the squared error loss for each approach.
 
@@
 \label{Sec:loocv}
-
-\bnote{I edited this section and the figure caption}
 
@@
 \label{Sec:CIndex}
-\bnote{unified the naming of C index in this manuscript.
-Use Harrell's C Index or C Index instead of other alternatives.
-This is the default concordance statistic that the survival package would return.}
 
@@
 
-In this section, we demonstrate the usage and performance of the aforementioned cross validation methods when fitting penalized regression models on data from a study of gene expression and survival in patients with lung cancer.
+In this section, we demonstrate the usage and performance of the aforementioned cross-validation methods when fitting penalized regression models on data from a study of gene expression and survival in patients with lung cancer.
 \citet{shedden2008gene} conducted a large, retrospective study to validate prognostic gene expression signatures as predictors of overall survival in lung cancer patients.
@@
 About $50\%$ of the patients died during the course of the study (236 events), with the rest censored.
+
+For this analysis, we fit elastic net models to the data in addition to lasso models, as these models often offer improvement over the lasso when many correlated genes are involved \citep{Zou2005}.
 The clinical variables stage, age and gender were included in the model without penalization, as it is reasonable to assume that all three have an effect on survival.
+Besides the regularization parameter $\lam$, the elastic net model has an extra tuning parameter $\alpha$, where $\alpha=1$ corresponds to lasso and $\alpha=0$ corresponds to ridge regression.
+For each $\alpha = 0.3, 0.5, 0.7, 0.9 \text{ and } 1$, we selected a regularization parameter $\lam$ that has the lowest cross-validation error for each of the four cross-validation methods described in Section~\ref{Sec:methods}.
+We reported the cross-validated C-Index and Brier Scores of thoses models in Table~\ref{tb:shedden}.
+With the highest C Index and lowest Brier Score, the model selected by the linear predictor approach at $\alpha = 0.3$ has best predictive performance among all of them.
 
-\begin{table}[h]
+\begin{table}[ht]
 \centering
-\begin{tabular}{ccccccccccc}
-\toprule
- & \multicolumn{2}{c}{$\alpha = 0.3$ } & \multicolumn{2}{c}{$\alpha = 0.5$} & \multicolumn{2}{c}{$\alpha = 0.7$} & \multicolumn{2}{c}{$\alpha = 0.9$} & \multicolumn{2}{c}{$\alpha = 1$} \\
- &      $\lambda$      &   CVE         &   $\lambda$        &     CVE       &      $\lambda$      &     CVE       &    $\lambda$        &    CVE        &     $\lambda$      &    CVE       \\
-
-\hline\\
-
- V$\&$VH  & 0.363 &  296.32  &  0.228   &  296.35  & 0.163 &  296.34  &  0.133  &   296.31  &  0.119  &  296.30  \\
- Standard & 0.301 & 145.17  & 0.218  &  145.38 & 0.163  & 145.37 &  0.127 &  145.35 & 0.114   & 145.34 \\
- Linear Pred & 0.288 & 10.661  & 0.181  & 10.673 & 0.163 & 10.673    & 0.127  & 10.672  &    0.114 &  10.672  \\
- Dev Resid & 0.316 & 571.40 & 0.228  &  571.96 & 0.163 & 571.95  & 0.133  &   571.79 &  0.119  &  571.70  \\
-\bottomrule
+\caption{\label{tb:shedden} The C Index and Brier Score of all elasticnet models selected by cross-validation methods }
+\begin{tabular}{lccccccccccc}
+\hline
+  & \multicolumn{5}{c}{Cross-validated C Index} & \multicolumn{5}{c}{Brier Score} \\ 
+  & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ &  
+ & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-6} \cline{8-12} 
+V $\&$ VH      & 0.639 & 0.637 & 0.638 & 0.637 & 0.637 &  & 0.243 & 0.244 & 0.244 & 0.245 & 0.245 \\
+Standard       & 0.647 & 0.639 & 0.638 & 0.638 & 0.639 &  & 0.233 & 0.243 & 0.244 & 0.244 & 0.244 \\
+Linear Pred    & \textbf{0.650} & 0.647 & 0.638 & 0.638 & 0.639 &  & \textbf{0.229} & 0.232 & 0.244 & 0.244 & 0.244 \\
+Deviance Resid & 0.645 & 0.637 & 0.638 & 0.637 & 0.637 &  & 0.235 & 0.244 & 0.244 & 0.245 & 0.245 \\ \hline
 \end{tabular}
-\caption{\bnote{Not sure whether to include this table or not}}
 \end{table}
 
-\begin{table}[h]
-\centering
-\begin{tabular}{lcccccc}
-  \toprule
-  & $\lambda$ & log($\lambda$) & $\#$ of Genes Selected & CV C-Index & Brier Score \\ 
-\cline{2-6} \\ 
-V $\&$ VH& 0.363 & -1.013 &  & 0.639 & 0.243\\
-standard & 0.301 & -1.200 &  & 0.647& 0.232\\
-Linear Pred &0.288 & -1.246 &  & 0.650&  0.229\\
-Dev Resid & 0.316 & -1.153 &  & 0.645& 0.235\\ 
-\bottomrule
-\end{tabular}
-\caption{\label{tb:shedden}Results of elasticnet model with $\alpha = 0.3$ on the Lung Cancer Data Set.
-The linear Predictor approach picked the smallest $\lambda$.
-The corresponding model yields the highest CV C-Index and smallest Brier Score evaluated at median survvial time \bnote{I got different number of genes for LP approach...}}
-\end{table}
-
-For this analysis, we fit elastic net models to the data in addition to lasso models, as these models often offer improvement over the lasso when many correlated genes are involved \citep{Zou2005}.
-Besides the regularization parameter $\lam$, the elastic net model has an extra tuning parameter $\alpha$, where $\alpha=1$ corresponds to lasso and $\alpha=0$ corresponds to ridge regression.
-For each $\alpha = 0.3, 0.5, 0.7, 0.9 \text{ and } 1$, we selected a regularization parameter $\lam$ that has the lowest cross-validation error.
-The standard approach, linear predictor approach and deviance approach all reach the lowest cross-validation error at $\alpha = 0.3$.
-
-We report the predictive performance for all four models at $\alpha = 0.3$ and their corresponding $\lambda$s in Table~\ref{tb:shedden}.
-For each model, we computed the cross-validated C-Index and Brier Scores at predicting the median survival time.
-How the CV methods perform on real data reflects what is shown in simulation studies.
-The standard approach and the linear predictor approach, selecting smaller $\lambda$ values, were the more liberal, while Verweij and Van Houwelingen's CVE approaches were both more conversative.
-With the highest C Index and lowest Brier Score, the linear predictor approach has best predictive performance among all four methods.
-The same patterns are repeated when $\alpha$ is at different values.
-
-We compared the CVE along the solution path for different methods.
-In Figure~\ref{fig:CVE}, the Cross Validated Error is rescaled and plotted for all four methods.
-A $\lambda$ is selected when CVE curve reaches its lower point.
-The blue line and green line, which represent the linear predictor approach and standard approah, have more curvature near theirs lowest point, making it is easier to pick out the minimum point.
+For each fixed $\alpha$, we compared the cross-validated errors(CVE) from the four cross-validation methods along the solution path.
+At $\alpha = 0.3$, CVEs are rescaled and plotted for all four met  hods at the same spectrum of $\lambda$ values In Figure~\ref{fig:CVE}.
+A $\lambda$ is selected when a CVE curve reaches its lower point.
+The blue line and the green line, which represent the CVE curves of linear predictor approach and standard approah, have more curvature near theirs lowest point, making it easier to pick out the minimum points.
+The selected $\lambda$ values, represented by the vertical dashed lines, and their corresponding models, reflected what is shown in simulation studies.
+The linear predictor approach, selecting smallest $\lambda$ values, was most liberal and selected 85 genes.
+While the standard approach and deviance residual approach selected 68 genes and 60 genes respectively, Verweij and Van Houwelingen's approach only selected 34 genes and performed most conservatively on this data set.
+Even though we only illustrated the models at $\alpha = 0.3$, same patterns are repeated when $\alpha$ is fixed at other values.
 
@@
   \centering
-  \includegraphics[height= 8cm ]{./manuscript_figure/shedden.jpg}
+  \includegraphics[height= 8cm ]{./manuscript_figure/shedden.png}
   \caption{\label{fig:CVE}Rescaled Cross Validated Errors at $\alpha = 0.3$.
@@
 After the rescaling, the Cross Validated Error of four different methods can be plotted together for comparison.
-}
+The vertical dashed lines represent the $\lambda$ selected by a specific method.
+The linear predictor approach selected the smallest $\lambda$ and Verweij and Van Houwelingen's approach selected the greatest $\lambda$}
 \end{figure}
 
-For the elastic net model with $\lam$ and $\alpha$ selected by the cross-validated linear predictor approach, 81 genes were selected.
-Of those, 78 were well-annotated.
+We carried out further data analysis based on the elastic net model with $\lam = 0.650$ and $\alpha = 0.3$, selected by the cross-validated linear predictor approach.
+Of the 85 genes that were selected by this model, 78 were well-annotated.
 We carried out an enrichment analysis of these 78 genes using ToppGene \citep{Chen2009c}; the five most significantly enriched molecular functions are listed in Table~\ref{Tab:go}.
@@
 
-\bnote{(END)}
 \section{Discussion}
 
-\par Four methods of conducting cross-validation for Cox Regression were compared through simulation studies in terms of their predictive accuracy.
-All cross-validation methods tend to select the more conservative model: with greater value of $\lambda$, and fewer variables selected into the model.
-Despite of the stability issue, the standard approach is the most liberal one across all scenarios and yield the least MSE.
-The proposed linear predictor approach performs closely to the standard approach.
-Under LOOCV, the linear predictor approach out-performs the standardstandard approach.
-Both Verweij and Van Houwelingen's approach and the proposed Deviance Residual approach perform conservatively, hence not recommended for use in practice.
-Results from real data application reflects results from the simulation studies.
+Conducting cross-validation is a crutial step in fitting penalized Cox regression model.
+In this paper, we proposed two intuitive ways to conduct cross-validation for Cox Regression: cross-validated linear predictor and cross-validated deviance residuals.
+We conducted a series of simulation experiments to study and compare their performance with two existing approaches, the standard cross-validation and Verweij and Van Houwelingen's cross-validated partial likelihood, in selecting tuning parameters when fitting penalized regression models.
+
+Despite the fact Verweij and Van Houwelingen's cross-validated partial likelihood has been widely implemented, its predictive performance seems to suffer, particularly when the dimension of the data grows and signals become weak.
+As is illustrated in both simulation studies and real data example, Verweij and Van Houwelingen's approach tends to perform conversatively, selecting fewer genes than all other models.
+In the lung cancer data example, the number of genes selected by other methods almost doubled the number of genes selected by Verweij and Van Houwelingen's approach.
+From an application standpoint, using Verweij and Van Houwelingen's approach in high-dimensional datasets may lead to more false negatives and miss important true findings.
+
+While the standard approach was the most liberal one across all simulation scenarios and yielded the best predictive performance, this approach limited by its stability issues which are carefully discussed in Section~\ref{Sec:stability}.
+The cross-validated linear predictor approach performs closely to the standard approach and demostrates an attractive balance of predictive performance and numerical stability.
+We would recommend using the proposed linear predictor approach in practice for selecting regularization parameters in penalized regression.
 
@@
 
-\section{Appendix}
+%\section{Appendix}
 
-\begin{figure}[h]
-	\centering
-	\begin{minipage}[t]{10cm}
-	\begin{algorithm}[H]
-	\caption{Standard Method}
-	\par \For{k in 1 to K } 
-	{
-		 Fit Cox Regression on $T_{k}$: Obtain model estimates $\hat{\beta}^{-k}$\;
-		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D_{k}$: $l^{k}(\hat{\beta}^{-k})$\;		
-	}
-
-	\par $ CVE_{st} = - \sum_{k = 1}^{K} l^{k}(\hat{\beta}^{-k})$
- 	\end{algorithm}
-
-	\begin{algorithm}[H]
-  	 \caption{Verweij and Van Houwelingen's Method}
-	\par \For{k in 1 to K } 
-	{
-		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
-		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $T_{k}$: $l^{-k}(\hat{\beta}^{-k})$\;
-		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D$: $l(\hat{\beta}^{-k})$\;
-		 Take the difference: $ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) $\;
-	}
-
-	\par $ CVE_{vvh} = - \sum_{k = 1}^{K} \left\{ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) \right\}$
-	\end{algorithm}
-
-	\begin{algorithm}[H]
-    	\caption{Cross-Validated Linear Predictors}
-	\par \For{k in 1 to K } 
-	{
-		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
-		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
-	}
-
-	\par $ CVE_{lp} = - l(\hat{\eta}^{-})$
-	\end{algorithm}
-
-	\begin{algorithm}[H]
-   	\caption{Sums of Squared Deviance Residuals}
-	\par \For{k in 1 to K } 
-	{
-		Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
-		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
-	} 
-	Compute baseline hazard $\hat{\Lambda}_{0}$ based on $\hat{\eta}^{-}$\;
-	$\forall$ $ i$ subject $\in \left\{1,2,...,n\right\}$, compute:
-		\par\hspace{0.5cm} $\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}^{-}_{i}}$;
-		\par\hspace{0.5cm} $d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}$;
-	\par $ CVE_{dr} = \sum_{i = 1}^{n} {d}_{i}^2$ \;
-	\end{algorithm}
-	\end{minipage}
-	\caption{Pseudo Code for Cross-validation Procedures}
-	\end{figure}
+%\begin{figure}[h]
+%	\centering
+%	\begin{minipage}[t]{10cm}
+%	\begin{algorithm}[H]
+%	\caption{Standard Method}
+%	\par \For{k in 1 to K } 
+%	{
+%		 Fit Cox Regression on $T_{k}$: Obtain model estimates $\hat{\beta}^{-k}$\;
+%		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D_{k}$: $l^{k}(\hat{\beta}^{-k})$\;		
+%	} 
+%	
+%	\par $ CVE_{st} = - \sum_{k = 1}^{K} l^{k}(\hat{\beta}^{-k})$
+% 	\end{algorithm}
+%
+%	\begin{algorithm}[H]
+%  	 \caption{Verweij and Van Houwelingen's Method}
+%	\par \For{k in 1 to K } 
+%	{
+%		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
+%		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $T_{k}$: $l^{-k}(\hat{\beta}^{-k})$\;
+%		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D$: $l(\hat{\beta}^{-k})$\;
+%		 Take the difference: $ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) $\;
+%	} 
+%
+%	\par $ CVE_{vvh} = - \sum_{k = 1}^{K} \left\{ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) \right\}$
+%	\end{algorithm}
+%
+%	\begin{algorithm}[H]
+%    	\caption{Cross-Validated Linear Predictors}
+%	\par \For{k in 1 to K } 
+%	{
+%		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
+%		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
+%	} 
+%
+%	\par $ CVE_{lp} = - l(\hat{\eta}^{-})$
+%	\end{algorithm}
+%	
+%	\begin{algorithm}[H]
+%   	\caption{Sums of Squared Deviance Residuals}
+%	\par \For{k in 1 to K } 
+%	{
+%		Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
+%		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
+%	} 
+%	Compute baseline hazard $\hat{\Lambda}_{0}$ based on $\hat{\eta}^{-}$\;
+%	$\forall$ $ i$ subject $\in \left\{1,2,...,n\right\}$, compute:
+%		\par\hspace{0.5cm} $\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}^{-}_{i}}$;
+%		\par\hspace{0.5cm} $d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}$;
+%	\par $ CVE_{dr} = \sum_{i = 1}^{n} {d}_{i}^2$ \;
+%	\end{algorithm}
+%	\end{minipage}
+%	\caption{Pseudo Code for Cross-validation Procedures}
+%	\end{figure}

### 2019-05-14 manuscript_code/shedden_elasticnet_git.R
add R code for Shedden data, update figure

@@ -1,0 +1,256 @@
+#------------------------
+#
+# Self contained R code
+# to analyze the Shedden Lung Cancer data set
+# with "MCP" penalty
+#
+#------------------------
+
+#------------------------
+# load/define functions that will be used in the analysis
+library(survival)
+library(ncvreg)
+library(glmnet)
+
+# Brier score for selected lambda
+
+brierSurv_lambda <- function(data,time,fit,lambda) {
+  predicted <- apply(data$x,1, 
+                     function(x)
+                     {predict(fit, x,type = "survival",
+                              lambda = lambda)(time)})
+  is_obs_after <- data$y[,1] > time
+  mean((is_obs_after - predicted)^2)
+}
+
+#------------------------
+# load data
+
+# Shedden Lung Cancer Data
+load(file = "Shedden2008.RData")
+
+# censor S by 72 month
+plot(survfit(S ~ 1), xlab = "month", ylab = "Survival Probability")
+abline(h = 0.5, col = "grey", lty = 2)
+
+# clean data
+ncolX <- ncol(X)
+n <- nrow(X)
+Z <- as.data.frame(Z)
+
+data_low <- data.frame(S,Z)
+data_low$Age <- as.numeric(data_low$Age)
+levels(data_low$Sex) <- c(0,1)
+data_low$Sex <- as.numeric(data_low$Sex)
+data_low$AdjChemo <- as.numeric(data_low$AdjChemo == "Yes")
+
+x <- as.matrix(data.frame(data_low[,c(2,3,5)],X))
+x <- ncvreg::std(x)
+data <- list(y=S,x=x)
+
+# prespecify fold id
+set.seed(123456)
+nfold <- 10
+nevent <- sum(data$y[,2])
+sde <- sqrt(.Machine$double.eps)
+foldid <- numeric(0)
+foldid[data$y[,2] == 1] <-  ceiling(sample(1:nevent)/(nevent + sde) * nfold)
+foldid[data$y[,2] == 0] <-  ceiling(sample(1:(n-nevent))/(n - nevent + sde) * nfold)
+
+
+Alpha <- c(1,0.9,0.7,0.5,0.3)
+A <- length(Alpha)
+
+#------------------------------------------------------------------
+for(a in 1:A){
+  # initial fit
+    alpha <- Alpha[a]
+    fit <-  glmnet(data$x, data$y,family = "cox",alpha = alpha,
[…]

### 2019-05-17 main.tex
Editing real data section (in progress)

@@
 The ability to avoid choosing a specific parametric distribution for the survival time is very attractive, as time-to-event data are often poorly described by fully parametric models.
-
-%\note{What's with all the \textbackslash par commands?  They don't seem to do anything.}
 
@@
 A C Index of 0.5 means the prediction is no better than flipping a coin and C Index of 1 means perfect concordance.
-We computed Harrell's C Index using the independent test set of 1000 individuals described above.
+We computed the C Index using the independent test set of 1000 individuals described above.
 
@@
 Out-of-sample Brier scores are plotted in the middle.
-Out-of-sample Harrell's C index is plotted in the right panel.
+Out-of-sample C index is plotted in the right panel.
 For each simulated data set, n = 120, p = 1000.
@@
 For all four metrics, the overall pattern is the same (KL scores not shown in Figure~\ref{Fig:mse-brier-c} due to space constraints, but they were very similar to Brier scores).
-Although the four cross-validation methods performed relatively similarly, the standard and linear predictor approaches had the best performance across all metrics: lower MSE, better Brier scores, and higher Harrell's C Index.
+Although the four cross-validation methods performed relatively similarly, the standard and linear predictor approaches had the best performance across all metrics: lower MSE, better Brier scores, and higher C Index.
 
@@
 \toprule
- n& p& $\#$ Non-Zero & Signal & Method & $\lambda$ & log(MSE Ratio) &Brier Score & Kullback Leibler & Harrell's C Index\\
+ n& p& $\#$ Non-Zero & Signal & Method & $\lambda$ & log(MSE Ratio) &Brier Score & Kullback Leibler & C Index\\
 \midrule
@@
   \includegraphics[height= 15cm ]{./manuscript_figure/figure_auc.png}
-  \caption{\label{Fig:cvauc}Comparing Partial Likelihood and Harrell's C Index both built from Cross-validated Linear Predictors.
+  \caption{\label{Fig:cvauc}Comparing Partial Likelihood and C Index both built from Cross-validated Linear Predictors.
 The horizontal axis represents the strength of signal $s$ in the data.
@@
 
-For this analysis, we fit elastic net models to the data in addition to lasso models, as these models often offer improvement over the lasso when many correlated genes are involved \citep{Zou2005}.
+For this analysis, we fit elastic net models to the data in addition to lasso models, as these models often improve upon the lasso when many correlated genes are involved \citep{Zou2005}.
 The clinical variables stage, age and gender were included in the model without penalization, as it is reasonable to assume that all three have an effect on survival.
-Besides the regularization parameter $\lam$, the elastic net model has an extra tuning parameter $\alpha$, where $\alpha=1$ corresponds to lasso and $\alpha=0$ corresponds to ridge regression.
-For each $\alpha = 0.3, 0.5, 0.7, 0.9 \text{ and } 1$, we selected a regularization parameter $\lam$ that has the lowest cross-validation error for each of the four cross-validation methods described in Section~\ref{Sec:methods}.
-We reported the cross-validated C-Index and Brier Scores of thoses models in Table~\ref{tb:shedden}.
-With the highest C Index and lowest Brier Score, the model selected by the linear predictor approach at $\alpha = 0.3$ has best predictive performance among all of them.
+In addition to the regularization parameter $\lam$, the elastic net model has an extra tuning parameter $\alpha$, where $\alpha=1$ corresponds to lasso and $\alpha=0$ corresponds to ridge regression.
+For $\alpha \in \{0.3, 0.5, 0.7, 0.9, 1\}$, we selected a regularization parameter $\lam$ that has the lowest cross-validation error for each of the four cross-validation methods described in Section~\ref{Sec:methods}.
+Since the cross-validation metrics for each method are on different scales and thus not directly comparable, Table~\ref{tb:shedden} reports the cross-validated C-Index and Brier Scores for the CV-selected model at each value of $\alpha$.
+As shown in the table, the elastic net model with $\alpha=0.3$ and $\lam$ selected by the cross-validated linear predictor approach has the best performance in terms of C Index and Brier Score.
 
@@
 
-For each fixed $\alpha$, we compared the cross-validated errors(CVE) from the four cross-validation methods along the solution path.
-At $\alpha = 0.3$, CVEs are rescaled and plotted for all four met  hods at the same spectrum of $\lambda$ values In Figure~\ref{fig:CVE}.
-A $\lambda$ is selected when a CVE curve reaches its lower point.
-The blue line and the green line, which represent the CVE curves of linear predictor approach and standard approah, have more curvature near theirs lowest point, making it easier to pick out the minimum points.
+For each fixed $\alpha$, we compared the cross-validation errors (CVE) from the four cross-validation methods along the solution path.
+At $\alpha = 0.3$, CVEs are rescaled and plotted for all four methods along the same grid of $\lambda$ values in Figure~\ref{fig:CVE}.
+The selected $\lambda$ values, where a given CVE curve reaches its lowest point, are indicated with dotted lines.
+The CVE curves for the standard and linear predictor approaches have greater curvature near their minima, and thus offer more clear support for the selected value of $\lam$.
+In comparison, the deviance residual and V\&VH approaches are flatter near their minima, supporting a considerably wider range of $\lam$ values and leading to greater model uncertainty.
 The selected $\lambda$ values, represented by the vertical dashed lines, and their corresponding models, reflected what is shown in simulation studies.

### 2019-05-20 main.tex
Edits to Sections 4 and 5

@@
 \section{Introduction}
+
+\note{I hate to change the name again, but what do you think about calling it ``basic'' cross-validation instead of ``standard''?
+My concern is that it's not really standard in the sense of ``the thing most people do''; I think most people use V\&VH since that's the default in glmnet.}
+
 \par Since its original proposal \citep{Cox1972}, Cox proportional hazards regression has become the most common regression approach for analyzing survival data.
@@
 Thus, a covariate-adjusted baseline hazard would mean that each fold is compared to a different saturated model.
-For this reason, we consider it important that the baseline hazard remains constant across folds when calculating cross-validated deviance residuals; this intuition is borne out by simulations involving various other approaches \note{supplement?
-otherwise, this sentence is a little tricky}\bnote{sure}.
+For this reason, it is important that the baseline hazard remains constant across folds when calculating cross-validated deviance residuals; this intuition is borne out by simulations involving various other approaches (see Supporting Information) \note{do we actually have a supporting information document written up?}.
 Note that this is not the case for the Brier Score or Kullback Leibler Score (described in Section~\ref{Sec:accuracy}), which involve absolute loss functions, not comparisons with a saturated model.
@@
 For $\alpha \in \{0.3, 0.5, 0.7, 0.9, 1\}$, we selected a regularization parameter $\lam$ that has the lowest cross-validation error for each of the four cross-validation methods described in Section~\ref{Sec:methods}.
-Since the cross-validation metrics for each method are on different scales and thus not directly comparable, Table~\ref{tb:shedden} reports the cross-validated C-Index and Brier Scores for the CV-selected model at each value of $\alpha$.
-As shown in the table, the elastic net model with $\alpha=0.3$ and $\lam$ selected by the cross-validated linear predictor approach has the best performance in terms of C Index and Brier Score.
+Although each method can be used to select a pair of $\alpha$ and $\lam$ values, the methods are on different scales and cannot be directly compared with each other.
+For this reason, Table~\ref{tb:shedden} reports the cross-validated C-Index and Brier Scores for each method's choice of $\lam$ at each value of $\alpha$.
+As shown in the table, the elastic net model with $\lam$ and $\alpha=0.3$ selected by the cross-validated linear predictor approach has the best performance in terms of C Index and Brier Score.
 
@@
 In comparison, the deviance residual and V\&VH approaches are flatter near their minima, supporting a considerably wider range of $\lam$ values and leading to greater model uncertainty.
-The selected $\lambda$ values, represented by the vertical dashed lines, and their corresponding models, reflected what is shown in simulation studies.
-The linear predictor approach, selecting smallest $\lambda$ values, was most liberal and selected 85 genes.
-While the standard approach and deviance residual approach selected 68 genes and 60 genes respectively, Verweij and Van Houwelingen's approach only selected 34 genes and performed most conservatively on this data set.
-Even though we only illustrated the models at $\alpha = 0.3$, same patterns are repeated when $\alpha$ is fixed at other values.
+The selected $\lambda$ values, represented by the vertical dashed lines, and their corresponding models, reflect what was observed in the simulation studies.
+The linear predictor approach was most liberal and selected 85 genes.
+While the standard approach and deviance residual approach selected 68 genes and 60 genes respectively, Verweij and Van Houwelingen's approach was most conservative, selecting only 34 genes.
+Similar patterns were observed for other $\alpha$ values.
 
@@
 The horizonal axis is $\log(\lambda)$.
-The vertical axis is the cross-validated error rescaled to a range from 0 to 1.
-After the rescaling, the Cross Validated Error of four different methods can be plotted together for comparison.
+The vertical axis is the cross-validated error rescaled to a range from 0 to 1 in order to allow plotting of the four methods on a common axis.
 The vertical dashed lines represent the $\lambda$ selected by a specific method.
-The linear predictor approach selected the smallest $\lambda$ and Verweij and Van Houwelingen's approach selected the greatest $\lambda$}
+The linear predictor approach selected the smallest $\lambda$ and Verweij and Van Houwelingen's approach selected the greatest $\lambda$.}
 \end{figure}
 
-We carried out further data analysis based on the elastic net model with $\lam = 0.650$ and $\alpha = 0.3$, selected by the cross-validated linear predictor approach.
-Of the 85 genes that were selected by this model, 78 were well-annotated.
+As mentioned earlier, for the elastic net model with tuning parameters chosen by cross-validated linear predictors ($\lam = 0.650$, $\alpha = 0.3$), 85 genes were selected, of which 78 were well-annotated.
 We carried out an enrichment analysis of these 78 genes using ToppGene \citep{Chen2009c}; the five most significantly enriched molecular functions are listed in Table~\ref{Tab:go}.
+Overall, the results are consistent with our understanding of cancer: many of the genes selected by the elastic net model are involved in signaling and the binding of transcription factors and receptors.
+These are precisely the pathways that tend to be disrupted in cancer.
+Chloride channels have also recently been implicated in several cancers as therapeutic targets \citep{Peretti2015}.
 
@@
 The five terms have a false discovery rate of 3\% using the Benjamini-Hochberg method.
-\note{I'm not sure whether we should include Selected and Annotated or not...if we do, we'll have to describe them better.}}
+Enrichment: frequency of the annotation among the selected genes divided by its frequency in the genome as a whole.}
 \begin{tabular}{llrrr}
 \toprule
-ID & Name & p & Selected & Annotated\\
+ID & Name & Selected & Enrichment & p \\
 \midrule
-GO:0008134 & transcription factor binding & 0.00015 & 10 & 584\\
-GO:0035257 & nuclear hormone receptor binding & 0.00057 & 5 & 161\\
-GO:0019869 & chloride channel inhibitor activity & 0.00061 & 2 & 9\\
-GO:0003700 & DNA-binding transcription factor activity & 0.00068 & 14 & 1264\\
-GO:0005102 & signaling receptor binding & 0.00083 & 16 & 1601\\
+GO:0008134 & transcription factor binding & 10 & 4.1 & 0.00015 \\
+GO:0035257 & nuclear hormone receptor binding & 5 & 7.4 & 0.00057 \\
+GO:0019869 & chloride channel inhibitor activity & 2 & 53.2 & 0.00061 \\
+GO:0003700 & DNA-binding transcription factor activity & 14 & 2.6 & 0.00068 \\
+GO:0005102 & signaling receptor binding & 16 & 2.4 & 0.00083 \\
 \bottomrule
@@
 
-Overall, the results are consistent with our understanding of cancer: many of the genes selected by the elastic net model are involved in signaling and the binding of transcription factors and receptors.
-These are precisely the pathways that tend to be disrupted in cancer.
-Chloride channels have also recently been implicated in several cancers as well as therapeutic targets \citep{Peretti2015}.
-
 Of particular interest is the fact that three of the selected genes (PIK3R1, PDPK1, and ITGB1) are involved in the PTEN pathway.
 The PTEN gene is an important tumor suppressor; mutations in PTEN are among the most common genetic changes found in human cancers and have been linked to cancers of the lung, breast, prostate, brain, and more.
-Time-to-event analyses like this one of gene expression data provide valuable insights for developing gene signatures that can help to identify patients with more severe disruptions to important pathways such as PTEN.
-These signatures thereby inform prognosis and impact treatment decisions.
+Time-to-event analyses like this one of gene expression data provide valuable insights for developing gene signatures that can help to identify patients with more severe disruptions to important pathways such as PTEN, thereby informing prognosis and impacting treatment decisions.
 
@@
 
-Conducting cross-validation is a crutial step in fitting penalized Cox regression model.
-In this paper, we proposed two intuitive ways to conduct cross-validation for Cox Regression: cross-validated linear predictor and cross-validated deviance residuals.
+Cross-validation is a crutial step in fitting penalized Cox regression models.
+In this paper, we propose two cross-validation methods for Cox Regression: cross-validated linear predictors and cross-validated deviance residuals.
 We conducted a series of simulation experiments to study and compare their performance with two existing approaches, the standard cross-validation and Verweij and Van Houwelingen's cross-validated partial likelihood, in selecting tuning parameters when fitting penalized regression models.
 
-Despite the fact Verweij and Van Houwelingen's cross-validated partial likelihood has been widely implemented, its predictive performance seems to suffer, particularly when the dimension of the data grows and signals become weak.
-As is illustrated in both simulation studies and real data example, Verweij and Van Houwelingen's approach tends to perform conversatively, selecting fewer genes than all other models.
+Despite the fact Verweij and Van Houwelingen's cross-validated partial likelihood has been widely implemented, its predictive performance appears suboptimal, particularly when the dimension of the data is large and/or signals are weak.
+As illustrated in both simulation studies and real data example, Verweij and Van Houwelingen's approach tends to perform conservatively, selecting fewer features than all other models.
 In the lung cancer data example, the number of genes selected by other methods almost doubled the number of genes selected by Verweij and Van Houwelingen's approach.
-From an application standpoint, using Verweij and Van Houwelingen's approach in high-dimensional datasets may lead to more false negatives and miss important true findings.
+From an application standpoint, using Verweij and Van Houwelingen's approach in high-dimensional applications may lead to more false negatives and miss important true findings.
 
-While the standard approach was the most liberal one across all simulation scenarios and yielded the best predictive performance, this approach limited by its stability issues which are carefully discussed in Section~\ref{Sec:stability}.
-The cross-validated linear predictor approach performs closely to the standard approach and demostrates an attractive balance of predictive performance and numerical stability.
-We would recommend using the proposed linear predictor approach in practice for selecting regularization parameters in penalized regression.
+While the standard approach performed well in terms of predictive performance, this approach suffers from a lack of numerical stability as discussed in Section~\ref{Sec:stability}.
+The cross-validated linear predictor approach performed very similarly to the standard approach in terms of predictive performance, but without any problems due to numerical stability.
+We recommend using the proposed linear predictor approach in practice for selecting regularization parameters in penalized Cox regression.
 

### 2019-05-21 cross-validation.tex
Edits to 1, 2, and 3.1

@@
 \newcommand{\Diag}{\mbox{Diag}}
+\providecommand{\Tr}{^{\scriptscriptstyle\top}}
 
@@
 Here, we propose two new cross-validation methods for Cox regression and compare them to approaches that have been proposed elsewhere.
-%\note{IS THIS STILL TRUE?
-WE SHOULD EDIT THIS SENTENCE: Our simulation studies show that in general, all cross-validation approaches tend to be conservative (i.e., select smaller models than the optimal choice of tuning parameters) for penalized Cox regression models.}
 Our proposed approach of cross-validating the linear predictors seems to offer an attractive balance of performance and numerical stability.
-We illustrate these advantages using simulated data as well as real data from two genome-wide studies of survival in cancer patients.
+We illustrate these advantages using simulated data as well as using them to analyze data from a high-dimensional study of survival in lung cancer patients.
 \input{main}

### 2019-05-21 main.tex
Edits to 1, 2, and 3.1

@@
 This semiparametric flexibility, however, comes at a cost.
-The Cox regression model can estimate relative risks, but without estimating the baseline hazard, it does not make any predictions concerning the absolute failure time for any given individual.
+The Cox regression model can estimate relative risks, but without estimating the baseline hazard, it does not predict the absolute failure time for any given individual.
 This poses a challenge to assessing the predictive accuracy of a given Cox regression model.
@@
 
-%In this paper, we briefly review existing approaches for meeting this challenge, then propose two new methods for carrying out cross-validation in Cox regression models.
-
-%As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard.
-Hence the interpretation of the Cox model is only valid in a relative sense.
-
-% NOTE: I don't know that we need this paragraph; cross-validation is pretty common.
-%One common approach for selecting the tuning parameter $\lambda$ for linear regression and logistic regression is via the K-fold cross validation.
-The data set would be split into K folds.
-One fold would be treated as the test set and the other K - 1 folds as the training set.
-The model would first be built on the training set, then fitted to the test set to obtain cross validation error (CVE).
-CVE would be calculated for each candidate $\lambda$, then the one that minimizes cross validated error would be selected.
-
 \par To overcome this drawback, an alternative approach was proposed by \citet{Verweij1993}.
 Their approach, which we describe in detail in Section~\ref{Sec:cox-cv-existing}, stabilizes the cross-validated log likelihood, enabling its use even when the number of subjects in each fold is small.
-This approach has been widely used and is implemented in R packages such as {\tt glmnet} \citep{glmnet}.
+This approach has been widely used and is the default approach to cross-validation in many software programs, such as the R package {\tt glmnet} \citep{glmnet}.
 Although this approach fixes the stability issue, we demonstrate here that in practice, it tends to behave conservatively in terms of model selection.
@@
 Through simulation studies, we compare these proposed methods with existing cross-validation approaches for LASSO penalized Cox regression in both low- and high-dimensional settings.
-%\note{STILL TRUE??
-We find that all cross validation approaches tend to be conservative,} \bnote{We should discuss this and the sentence in abstract at our next meeting.
-They are conservative only in terms of selecting the model that has the best preditive performance; they do not miss selecting variables in the true support}
 The linear predictor approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.
-We conclude by applying both proposed and existing approaches to two high dimensional data sets from real studies with time-to-event outcomes.
+We conclude by applying both proposed and existing approaches to a high-dimensional study of survival in lung cancer patients.
 
@@
 \begin{equation*}
-  h_{i}(t) = h_{0}(t) \exp( X_{i}^{T} \beta),
+  h_{i}(t) = h_{0}(t) \exp( X_{i}\Tr\beta),
 \end{equation*} 
-where $h_{0}$ is the baseline hazard and $e^{X_i^{T} \beta}$ is the risk for a subject with covariates $X_i$ relative to that baseline.
+where $h_{0}$ is the baseline hazard and $\exp(X_i\Tr\beta)$ is the risk for a subject with covariates $X_i$ relative to that baseline.
 The estimation of the coefficients is obtained by maximizing a partial likelihood.
-Letting $t_i$ denote the time of being in the study for subject $i$ and $\delta_{i}$ indicate whether or not an event is observed for subject $i$, each subject's individual contribution to that partial likelihood is
+Letting $t_i$ denote the time on study for subject $i$ and $\delta_{i}$ indicate whether or not an event is observed for subject $i$, each subject's individual contribution to that partial likelihood is
 \begin{equation}
   \label{eq:cox-pl-subj}
-  L_{i}(\beta) = \left \{\frac{\exp ( X_{i}^{T} \beta)}{\sum_{ k \in R(t_{i})}\exp ( X_{k}^{T} \beta)}\right \}^{\delta_{i}},
+  L_{i}(\beta) = \left \{\frac{\exp ( X_{i}\Tr\beta)}{\sum_{ k \in R(t_{i})}\exp ( X_{k}\Tr \beta)}\right \}^{\delta_{i}},
 \end{equation}
@@
 
-\par In this paper, we focus upon the LASSO penalty $P_{\lambda}(\beta) = \lambda \sum_{j} |\beta_{j}|$, although the methods we analyze can be used with any penalty as well as with model selection in unpenalized Cox regression (best subset selection).
-LASSO-penalized cox regression is particularly useful for high dimensional data where the number of covariates $p > n$; for example, in the prediction of overall survival in cancer patients based on genome-wide expression measurements.
-LASSO estimates yield a sparse model where some coefficients are estimated to be exactly zero.
+\par In this paper, we focus upon the LASSO penalty $P_{\lambda}(\beta) = \lambda \sum_{j} |\beta_{j}|$, although the methods we analyze can be used with any penalty as well as with model selection in unpenalized Cox regression.
+LASSO-penalized cox regression is particularly useful for high dimensional data where the number of covariates $p > n$, such as predicting overall survival in cancer patients based on genome-wide expression measurements.
+LASSO estimates yield a sparse model where some coefficients are exactly zero.
 This sparsity pattern changes as we vary $\lam$: at large values of $\lambda$, most or all of the coefficients are 0, but as $\lam$ decreases, more covariates are selected.
@@
 
-\par Cross-validation is one of the most common methods used to select the appropriate $\lambda$ for penalized regression models.
+\par Cross-validation is the most common method for selecting $\lambda$ in penalized regression models.
 Suppose a data set $D$ of n observations is partitioned into $K$ folds: $D_{1}, D_{2}, \ldots, D_{K}$.
@@
 In particular, the partial likelihood \eqref{eq:cox-pl-subj} involves a risk set -- which observations should be used to form that risk set?
-In the following sections, let $L^{-k}$ denote the partial likelihood built over data in $T_k$, let $L^{k}$ denote the partial likelihood built over data in $D_k$, and $L$ denote the partial likelihood built over the entire data set $D$.
+In the following sections, let $L^{-k}$ denote the partial likelihood built using data in $T_k$, let $L^{k}$ denote the partial likelihood built using data in $D_k$, and $L$ denote the partial likelihood built using the entire data set $D$.
 The log partial likelihood is denoted by $\ell$, with $\ell^k$ and $\ell^{-k}$ defined similarly.
 Finally, let $\hat{\beta}^{-k}$ denote the penalized estimates that are obtained by using $L^{-k}(\beta)$ in \eqref{eq:obj}.
+Figure~\ref{Fig:methods} diagrams the relationships between the four cross-validation methods we will describe in this section.
 
@@
 \par A direct approach to measure a model's predictive accuracy in $D_k$ is to evaluate the log partial likelihood at $\hat{\beta}^{-k}$ using the data in $D_k$.
-Typically, cross-validation error (CVE) is measured by the total deviance across all $K$ folds, which (up to a constant) is given by:
+Cross-validation error (CVE) is then measured by the total deviance across all $K$ folds, which (up to a constant) is given by:
 \begin{equation}
   \label{eq:standard}
-  \CVE = -2 \sum_{k=1}^{K} \ell^{k}(\hat{\beta}^{-k}).
+  \CVE = -2 \sum_{k=1}^{K} \ell^{k}(\hat{\beta}^{-k});
 \end{equation}
-%where $\ell^{k}(\hat{\beta}^{-k}) = \sum \limits_{i \in D_{k}} \log L^{k}_{i}(\hat{\beta}^{-k})$.
-We refer this approach as \emph{standard cross-validated partial likelihood}.
-This is implemented in the $\tt{glmnet}$ package as the \verb|"ungrouped"| option.
-%\note{I added a -2 here so that CVE corresponds to deviance; this is what's done in glmnet/ncvreg}.
+the factor of 2 is arbitrary, but often used so that CVE estimates the deviance of the model.
+We refer this approach as \emph{standard cross-validated partial likelihood}. This is implemented in the $\tt{glmnet}$ package as the \verb|"ungrouped"| option.
 
@@
 However, the standard approach can be problematic for Cox regression in the sense that that there may not be enough observations in $D_k$ to build up the risk set for partial likelihood.
-For example, the standard approach cannot work with leave-one-out cross-validation, since $\ell_k$ would either be zero or undefined for all folds.
+For example, the standard approach cannot work with leave-one-out cross-validation, since for all folds, $\ell_k$ would either be zero or undefined.
 
@@
 Since its proposal, the V\&VH approach has been widely implemented as a tool for cross-validation is Cox models.
-For example, it is used by the R packages {\tt CoxRidge} \citep{CoxRidge}, {\tt fastcox} \citep{fastcox}, {\tt SGL}\citep{SGL} , {\tt CoxBoost}\citep{CoxBoost}, {\tt mboost}\citep{mboost}, and {\tt glmnet}\citep{glmnet}.
-In {\tt glmnet}, it is the default choice for penalized Cox regression and is referred to as the {\tt grouped} option in the package's syntax.
+For example, it is used by the R packages {\tt CoxRidge} \citep{CoxRidge}, {\tt fastcox} \citep{fastcox}, {\tt SGL} \citep{SGL} , {\tt CoxBoost} \citep{CoxBoost}, {\tt mboost} \citep{mboost}, and {\tt glmnet} \citep{glmnet}.
+In {\tt glmnet}, it is the default choice for penalized Cox regression and is referred to as the \verb|"grouped"| option in the package's syntax.
+
+\begin{figure}
+  \centering
+  \includegraphics[height= 12cm ]{./manuscript_figure/figure_1_new.png}
+  \caption{\label{Fig:methods} Illustration of $K$-fold cross validation methods for Cox regression.}
+\end{figure}
 
@@
   \label{eq:cv-lp}
-  \hat{\eta}^{cv}_{i} = X_{i}\hat{\beta}^{-k}
+  \hat{\eta}^{cv}_{i} = X_{i}\Tr\hat{\beta}^{-k}
 \end{equation} 
@@
 \hat{\eta}^{cv}_{n})$ for the whole sample is obtained.
-A partial likelihood can then be built over this set of linear predictors to evaluate the predictive accuracy of the model:
+A partial likelihood can then be built using this set of linear predictors to evaluate the predictive accuracy of the model:
 	\begin{equation} 
@@
 
-\par This idea of obtaining linear predictors over test sets, then constructing the partial likelihood after pooling all linear predictors together is implemented in R package \texttt{ncvreg} \citep{ncvreg}.
+\par This idea of obtaining linear predictors using test sets, then constructing the partial likelihood after pooling all linear predictors together is implemented in R package \texttt{ncvreg} \citep{ncvreg}.
 This idea has also appeared in the cross-validation literature for quantities that cannot be evaluated on only a subset of the data, such as AUC in logistic and Cox regression models \citep{Parker2007,Simon2011a,Subramanian2011}.
@@
 Hence, another approach to cross-validation would be to include an extra step of estimating the baseline hazard and using it to quantify the cross-validation error.
-The cumulative baseline hazard $\Lambda_0$ is typically estimated via the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
+The standard approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
 \begin{equation}
@@
 \begin{equation}
-  \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\hat\beta).
+  \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta).
 \end{equation}
@@
  \label{eq:cv-bhz}
-  \hat{\Lambda}^{cv}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\hat\beta^{-k}).
+  \hat{\Lambda}^{cv}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta^{-k}).
 \end{equation}
@@
 \begin{equation}
-  \hat{M^{cv}_{i}} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\hat\beta^{-k}).
+  \hat{M^{cv}_{i}} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}).
 \end{equation}
@@
 \end{equation}
-The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$ are used as the cross validated error as an analogy to the residual sums of square for linear regression.
+The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross validated error, analogously to using the residual sum of squares in linear regression.
 We refer to this method as the \emph{cross-validated deviance residuals} approach.
@@
 The baseline hazard estimate we describe here differs from the one typically used in constructing deviance residuals \citep{Therneau1990}.
-Instead of the Nelson-Aalen estimator, deviance residuals are typically calculated based on a baseline hazard estimate that has been adjusted for the covariates in the model such as the one described in \citet{breslow1972}.
+Instead of the Nelson-Aalen estimator, deviance residuals are typically calculated based on a baseline hazard estimate that has been adjusted for the covariates in the model \citep{breslow1972}.
 In the context of cross-validation, however, this approach is problematic.
@@
 Thus, a covariate-adjusted baseline hazard would mean that each fold is compared to a different saturated model.
-For this reason, it is important that the baseline hazard remains constant across folds when calculating cross-validated deviance residuals; this intuition is borne out by simulations involving various other approaches (see Supporting Information) \note{do we actually have a supporting information document written up?}.
+For this reason, it is important that the baseline hazard remains constant across folds when calculating cross-validated deviance residuals; this intuition is borne out by simulations involving various other possible ways of constructing cross-validated deviance residuals (see Supporting Information).
 Note that this is not the case for the Brier Score or Kullback Leibler Score (described in Section~\ref{Sec:accuracy}), which involve absolute loss functions, not comparisons with a saturated model.
-
-%A more straightforward way to compute the cross-validated deviance residuals is to estimate the baseline hazard using $T_k$, leaving $D_k$ as out-of-sample observations to calculate the residuals.
-There are two major issues with this approach.
-First, it is possible that some events in $D_k$ occurred after the last event in $T_k$.
-In this case, the survival function would drop to 0.
-Secondly, there is a numeric challenge for computing the term $log(\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}_{i}})$ in deviance residuals.
-Since the baseline hazard function is a step function from 0, the cumulative hazard $\hat{\Lambda}_{0}$ at the first time point from 0 is always 0.
-For whichever $D_k$ that contains the first time point, $log(\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}_{i}})$ would be negative infinity for that observation and lead to a numeric issue.
-To solve the first issue, events that occur later than the last event in $T_k$ should be censored.
-To solve the second issue, the step function of the baseline hazard needs to be smoothed.
-Both issues do not exist when the baseline hazard is built over the cross-validated linear predictors using all samples.
-\note{this paragraph may be optional, and can also be moved to discussion}
-
-\begin{figure}
-  \centering
-  \includegraphics[height= 12cm ]{./manuscript_figure/figure_1_new.png}
-  \caption{Illustration of K-fold Cross Validation Methods for Cox Regression}
-\end{figure}
 
@@
 
-Our first simulation study concerns the accuracy of the cross-validation approaches introduces in Section~\ref{Sec:methods}.
-We varied the sample size, dimension, censoring percentage and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation.
+In this section, we consider the accuracy of the cross-validation approaches introduced in Section~\ref{Sec:methods}.
+We varied the sample size, dimension, censoring percentage \note{(we don't really vary censoring percentage, do we?)}, and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation.
 Accuracy was assesed via several criteria.
@@
 Estimation accuracy was measured using the mean squared error of the coefficients ($\MSE$).
-Suppose $\hat{\beta}(\lambda)$ is the set of coefficients estimated by a cox model with LASSO penalty at $\lambda$.
-Mean squared error $MSE$ is defined to be
-%\bnote{
+Suppose $\hat{\beta}(\lambda)$ is the vector of coefficients estimated by a Cox model with LASSO penalty at $\lambda$, and $\beta^*$ the value of $\beta$ used the generate the data.
+Mean squared error is defined to be
 \begin{equation}
-E\| \hat{\beta}(\lambda) - \beta \| ^2.
+\MSE = E\| \hat{\beta}(\lambda) - \beta^* \| ^2.
 \end{equation}
-%}
-%\note{this is the equation for a scalar MSE, not a vector}
 
@@
 where $\lambda^{cv}$ denotes the value of $\lambda$ selected by cross-validation.
-MSE is estimated by avevaring over N replications respectively for $\hat{\beta}(\lambda^{cv})$ and $\hat{\beta}^{oracle}$.
-The ratio is taken afterwards.
+MSE is estimated by averaging over $N$ replications respectively for $\hat{\beta}(\lambda^{cv})$ and $\hat{\beta}^{oracle}$, with the ratio taken afterwards.
 The $\log$ is taken to correct for skewness.
-%\note{I'm not sure why this is a note; shouldn't you rewrite the sentence?}
 
@@
 Expected censoring percentage is 10$\%$.
-%\note{I really think the x axis should be Signal (s) or Signal strength (s)}
 }
@@
 
- For each simulated data set, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates.
-The covariate vector $\beta$ is assumed to be sparse, with $\beta_j = 0$ for $j \in \{11,12,..,1000\}$ and $\beta_j = s$ for $j \in \{1,2, ..., 10 \}$, where $s$ is a prespecified constant that parameterizes the signal strength in the simulated data set.
-In the first simulation study, we varied $s$ from 0.4 to 0.9.
+For our first simulation study, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates.
+The covariate vector $\beta$ is assumed to be sparse, with $\beta_j = 0$ for $j \in \{11, 12, \ldots, 1000\}$ and $\beta_j = s$ for $j \in \{1, 2, \ldots, 10 \}$, where $s$ is a prespecified constant that specifies the signal strength in the simulated data set.
+Here, we vary $s$ from 0.4 to 0.9.
 The survival outcomes have 10$\%$ expected censoring.
@@
 $N = 200$ replications were used for each scenario.
-
-\par Figure~\ref{Fig:mse-brier-c} illustrates the results of this simulation study.
-In each panel, the horizontal axis is the magnitude of the non-zero $\beta$ coefficients.
+Figure~\ref{Fig:mse-brier-c} illustrates the results of this simulation study.
 For all four metrics, the overall pattern is the same (KL scores not shown in Figure~\ref{Fig:mse-brier-c} due to space constraints, but they were very similar to Brier scores).
 Although the four cross-validation methods performed relatively similarly, the standard and linear predictor approaches had the best performance across all metrics: lower MSE, better Brier scores, and higher C Index.
+
+\note{Just curious: I would imagine that standard/basic cross-validation does best when there is least censoring, as in Section 3.2.
+Is this true?
+In other words, does Figure 2 make basic cross-validation look better than it often is, because the censoring rate is so small, just 10\%?}
 
@@
 In all settings, the censoring rate was 30\%.
+\note{What is s for the strong and weak cases?}
 
@@
 The improvement is particularly noticeable in the most difficult setting: high-dimensional data with weak signal.
-Table~\ref{Tab:sim} also provides some insight into why the deviance residual and V\&VH approaches perform poorly relative to standard CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and selecting fewer variables.
-%\note{if we're going to say something about ``selecting fewer variables'', should Table 1 include the average model size?} \bnote{I'm afraid that the average model size would be confusing to the readers.
-The most conversative model would have the smallest model size, and would be closest to the true model size}
+Table~\ref{Tab:sim} also provides some insight into why the deviance residual and V\&VH approaches perform poorly relative to standard CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and thus, selecting fewer variables.
 

### 2019-05-22 manuscript_code/figure_3_git.R
add simulation code for figure 3

@@ -1,0 +1,264 @@
+#------------------------------
+# Self-contained R code
+# to reproduce simulations
+# in Figure 3 of the Cross Validation 
+# Manuscript:
+# Explore Stability of the Standard
+# CV Method
+#------------------------------
+
+# functions
+library(survival)
+library(ncvreg)
+
+generatebeta <- function(p,nbeta,c = 1){
+  pos <- rep(c, ceiling(nbeta/2))
+  neg <- rep(-c, floor(nbeta/2))
+  beta <- c(pos,neg,rep(0,p-nbeta))
+}
+
+genSurvData <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,h = 1 # 70% means 70% of the data are censored; 30% are observed
+){
+  p <- length(beta)
+  ## Generate column vector of coefficients
+  beta <- matrix(beta, ncol = 1)
+  x <- matrix(rnorm(n*p, sd = sd),n,p)
+  #standardize x first
+  x <- ncvreg::std(x)
+  hazard <- h*exp(x %*% beta)
+  ## Calculate survival times
+  time <- rexp(n, hazard)
+  
+  ## observations are censored
+  id <- sample(n,round(cpct * n))
+  event <- rep(1,n)
+  event[id] <- 0
+  
+  y <- Surv(time = time, event = event)
+  return(list(x=x,y=y,beta = beta))
+}
+
+
+
+# increase number of folds (plot on the right)
+#----------------------------------------------------------------------------------------
+
+n <- 120
+p <- 1000
+cpct <- 0.5
+nbeta <- 10
+beta <- generatebeta(p,nbeta = nbeta ,c = 0.5)
+N <- 200
+Nfold <- c(10,12,15,20,30,60)
+
+# initiation
+gp <- array(NA,c(2,N,length(Nfold)))
+ug <- array(NA,c(2,N,length(Nfold)))
+ncv <- array(NA,c(2,N,length(Nfold)))
+dvr <- array(NA,c(2,N,length(Nfold)))
+oracle.mse <- matrix(NA, nrow = N, ncol = length(Nfold))
+foldid <- numeric(n)
+
+for(i in 1:N){
+  set.seed(666 + i)
+  
+  for (j in 1:length(Nfold)){
+    nfold <- Nfold[j]
+    # generate data
[…]

### 2019-05-22 supplementary.tex
add supplementary material and R code

\documentclass{article}
\usepackage{graphicx}
\usepackage{color}

\usepackage{alltt}
\usepackage{amssymb,amsmath,natbib,graphicx,enumerate,subcaption,tikz,url,booktabs}
\usepackage[bmargin=0.75in, tmargin =0.75in,lmargin = 0.5in,rmargin = 0.5in]{geometry}

\usepackage{dsfont}
\usepackage[ruled]{algorithm2e}
\usepackage{algpseudocode}
\renewcommand{\algorithmcfname}{Procedure}
\usepackage{bbold}

\newcommand{\ith}{i^\textrm{th}}
\newcommand{\CVE}{\textrm{CVE}}
\newcommand{\MSE}{\textrm{MSE}}
\newcommand{\logit}{\mbox{logit}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\ones}{\mathbbm{1}}
\newcommand{\indic}[1]{\boldsymbol{1}_{\{ #1 \}}}
\newcommand{\mvec}{\mbox{vec}}
\newcommand{\cov}{\mbox{Cov}}
\newcommand{\eqdist}{\overset{{\cal D}}{=}}
\newcommand{\const}{\mbox{const}}
\newcommand{\iid}{\overset{iid}{\sim}}
\newcommand{\diag}{\mbox{{\bf diag}}}
\newcommand{\Diag}{\mbox{Diag}}
\providecommand{\Tr}{^{\scriptscriptstyle\top}}

\newcommand{\lam}{\lambda}
\newcommand{\bmu}{\boldsymbol\mu}
\newcommand{\balpha}{\boldsymbol\alpha}
\newcommand{\br}{\boldsymbol{r}}
\newcommand{\bs}{\boldsymbol{s}}
\newcommand{\by}{\boldsymbol{y}}
\newcommand{\bbeta}{\boldsymbol{\beta}}
\newcommand{\btheta}{\boldsymbol{\theta}}
\newcommand{\btau}{\boldsymbol{\tau}}
\newcommand{\bx}{\boldsymbol{x}}
\newcommand{\beps}{\boldsymbol\epsilon}

\newcommand{\cX}{{\cal X}}
\newcommand{\cA}{{\cal A}}
\newcommand{\cAmin}{{\cal A}^{\mathsmaller{-}}}

\newcommand{\vm}[1]{\mbox{vec}^\mathsmaller{-}\hspace{-0.25pc}\left(#1\right)}
\newcommand{\Imin}{I^\mathsmaller{-}}
\newcommand{\Itri}{{I^\mathsmaller{\triangle}}}

\newcommand{\deriv}[1]{\noindent {\it Derivation:} #1 $\hfill\square$}
\newcommand{\pbe}[1]{{\cal P}_{\epsilon,#1}}
\newcommand{\be}{{\cal B}_{\epsilon}}

\newcommand{\adjSet}{\{A_{k_\ell}\}_{\ell=1}^K}
\newcommand{\abSet}{\{\alpha_{k_\ell},\bbeta_{k_\ell}\}_{\ell=1}^K}
\newcommand{\bZ}{{\bf Z}}

\providecommand{\note}[1]{\textcolor{red}{#1}}
\providecommand{\bnote}[1]{\textcolor{blue}{#1}}

\title{Supplementary Material\\
\textit{Cross Validation Approaches for Penalized Cox Regression}}

\author{Biyue Dai\\Department of Biostatistics\\University of Iowa
  \and
  Patrick Breheny\\Department of Biostatistics\\University of Iowa}
\date{\today}

\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

\begin{document}
\maketitle

\section{Cross-Validated Deviance Residuals}

The Cox model implies the following relationship between the baseline cumulative hazard and the cumulative hazard for individual $i$:
\begin{equation}
  \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta).
\end{equation}
Consider the following approach: for fold $k$, we obtain $\hat{\beta}^{-k}$ from the training set $T_k$ and then for each $i$ in the test set $D_k$, 
\begin{equation}
 \label{eq:cv-bhz}
  \hat{\Lambda}^{cv}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta^{-k}).
\end{equation}

A natural candidate for incorporating the cumulative hazard into a loss function is the deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}.  Given \eqref{eq:cv-bhz}, we can first obtain the cross-validated Martingale residual: 
\begin{equation}
  \hat{M^{cv}_{i}} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}).
\end{equation}
The cross-validated deviance residuals can then be derived from the Martingale residuals: 
\begin{equation} 
  d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}.
\end{equation}
The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross validated error, analogously to using the residual sum of squares in linear regression. We refer to this method as the \emph{cross-validated deviance residuals} approach.

Deviance residuals are typically calculated based on a baseline hazard estimate that has been adjusted for the covariates in the model. In the context of cross-validation, however, this approach is problematic. Deviance residuals measure the difference between the fitted model and a saturated model; in Cox regression, this saturated model depends on the baseline hazard. Thus, a covariate-adjusted baseline hazard would mean that each fold is compared to a different saturated model. For this reason, it is important that the baseline hazard remains constant across folds when calculating cross-validated deviance residuals; this intuition is borne out by simulations involving various other possible ways of constructing cross-validated deviance residuals.

\section{Baseline Estimations}

	\subsection{Nelson-Aalen Estimator}
	 The simplest approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
\begin{equation}
  \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}.
\end{equation}
	The Nelson-Aalen Estimator does not involve any information from the covariates. It only depends on the failure times.

	\subsection{Kalbfleisch and Prentice's Approach}					
	\cite{Kalbfleisch2011} proposed a cumulative hazard estimator that incorporates covariates. Assuming only a single failure occurs at $t_i$, based on the fitted values of $\hat{\bbeta}$, we can first compute linear predictors $\hat{\eta}_i$ at each observed failure time. Based on $\hat{\eta}_i$, we can compute a discrete hazard component at each observed failure time):
\begin{equation}
1 - \hat{\alpha}_{i} = 1 - \left\{ 1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}exp (\hat{\eta}_{j}} \right\}^{exp(-\hat{\eta}_{i})},
\end{equation}
and aggregate them together into the cumulative hazard function:
\begin{equation} 
\hat{\Lambda}_0(t) = \sum(1 - \hat{\alpha}_i)\mathds{1}(t_{i} \leq t),
\end{equation}

\section{Simulation Comparison}
The following simulation illustrates how the performance of cross-validated deviance residuals would vary based on different baseline estimation. We generated data using the same mechanism as we described in Section 3.1 in the paper. In each simulated data set, there were $n = 150$ observations, $p = 1000$ number of feature and 10 non-zero features with signal strength $s$. Censoring proportion is fixed at $10\%$. For each scenario, we varied $s$ from 0.4 to 0.9 and 200 replications were used. We compared the cross-validated deviance residuals constructed by Nelson-Aalen estimator and the cross-validated deviance residuals constructed by Kalbfleisch and Prentice's estimator. We also included cross-validated linear predictor approach (introduced in Section 2.3 in the paper) for comparison. The results of the simulation is listed in the following table:

\begin{table}[h]
\centering
\begin{tabular}{lrrrrrr}
\toprule
  & 0.4 & 0.5 & 0.6 & 0.7 & 0.8 & 0.9\\
\midrule
\vspace{1mm}
 & \multicolumn{5}{c}{$\lambda$} & \\ 
\vspace{1mm}
CVDR - Nelson Aalen & 0.1531 & 0.1346 & 0.1251 & 0.1217 & 0.1213 & 0.1210\\
CVDR - Kalbfleisch and Prentice& 0.2534 & 0.2275 & 0.2084 & 0.1767 & 0.1588 & 0.1414\\
CVLP & 0.1441 & 0.1214 & 0.1078 & 0.0981 & 0.0914 & 0.0849\\
%Oracle &  0.1090 & 0.0941 & 0.0762 & 0.0615 & 0.0567 & 0.0516\\
\vspace{1mm}
 & \multicolumn{5}{c}{$\log($MSE$)$} & \\ 
\vspace{1mm}

CVDR - Nelson Aalen & 2.4038 & 2.5107 & 2.8289 & 3.0084 & 3.1723 & 3.3463\\
CVDR - Kalbfleisch and Prentice& 2.7075 & 2.9397 & 3.2592 & 3.3596 & 3.4228 & 3.4934\\
CVLP &2.3684 & 2.4089 & 2.6595 & 2.7173 & 2.7761 & 2.8264\\
\bottomrule
\end{tabular}
\end{table}

Top half of the table listed different $\lambda$ values selected by the method. With the Kalbfleisch and Prentice baseline, the cross-validated deviance residual always selected larger $\lambda$ values and very conservative models. With the covariate-adjusted baseline, the deviance residuals seem to have stronger preference for models closer to the null model. This phenomenon is also observed by the original authors of the deviance residuals in classical Cox regression applications. As they attempted to use the covariate-adjusted deviance residuals and Martingale residuals to evaluate model fit, they observed that "the no-covariate model has the smallest sum of squared and the best model, the one with all five predictors, has the largest" \citep{Therneau2000modeling}.

The bottom half of the table listed $\log($MSE$)$ of the models selected by cross-validation approaches compared to the oracle model. The definition of $\log($MSE$)$ is described in more details in Section 3.1 in the paper. Throughout different signal strength, CVDR with Kalbfleisch and Prentice baseline had the largest $\log($MSE$)$, indicating worst model estimation among all methods. With Nelson-Aalen estimator, the CVDR approach had better performance and performed closer to CVLP at weaker signals. As the signal strength increases, the performance of CVDR with Nelson-Aalen estimator also declined.

Overall, our simulation results support the use of Nelson-Aalen estimator in estimating the baseline hazard for the cross-validated deviance residual approach.

\bibliographystyle{abbrvnat}

\bibliography{articles}

\end{document}

### 2019-05-22 manuscript_code/supplementary_git.R

@@ -1,0 +1,152 @@
+#--------------------------------------
+# R code that generates the simluation 
+# in the supplementary material:
+# compares different baseline construction
+# for CV deviance residuals
+#--------------------------------------
+
+library(ncvreg)
+library(survival)
+
+generatebeta <- function(p,nbeta,c = 1){
+  pos <- rep(c, ceiling(nbeta/2))
+  neg <- rep(-c, floor(nbeta/2))
+  beta <- c(pos,neg,rep(0,p-nbeta))
+}
+
+genSurvData <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,h = 1 # 70% means 70% of the data are censored; 30% are observed
+){
+  p <- length(beta)
+  ## Generate column vector of coefficients
+  beta <- matrix(beta, ncol = 1)
+  x <- matrix(rnorm(n*p, sd = sd),n,p)
+  #standardize x first
+  x <- ncvreg::std(x)
+  hazard <- h*exp(x %*% beta)
+  ## Calculate survival times
+  time <- rexp(n, hazard)
+  ## observations are censored ("event = rbinom(n, 1, 0.30)")
+  y <- Surv(time = time, event = rbinom(n, 1, 1-cpct))
+  return(list(x=x,y=y,beta = beta))
+}
+
+#---------------------------------------------------------------------------------------------------------
+# simulate data
+
+n <- 150
+p <- 1000
+c <- 4:9/10
+rep <- 200
+
+result_lambda <- array(NA,dim = c(5,length(c),rep),dimnames = list(c("NA","Full","CV","LP","Oracle"),
+                                                                     as.character(c),
+                                                                     NULL))
+result_MSER <- array(NA,dim = c(5,length(c),rep),dimnames = list(c("NA","Full","CV","LP","Oracle"),
+                                                                 as.character(c),
+                                                                 NULL))
+set.seed(1)
+for(R in 1:rep){
+  for(j in 1:length(c)){
+    beta <- generatebeta(p,nbeta = 10 ,c = c[j])
+    data <- genSurvData(n=n , beta = beta, cpct = 0.1, sd = 1)
+
+    #initial fit
[…]

### 2019-05-22 main.tex
Edits to Section 3.2

@@
 
-The standard cross-validation in the partial likelihood setting is both straightforward and enjoys very good predictive accuracy, as seen in Section~\ref{Sec:accuracy}.
-However there are two main stability drawbacks to its widespread use.
-The following simulation studies illustrated those issues faced by the standard approach.
+Given that basic cross-validation is both straightforward and demonstrates strong predictive accuracy, as seen in Section~\ref{Sec:accuracy}, one might wonder why it is not more widely used for Cox regression.
+The reason is that it suffers from several drawbacks with respect to numerical stability.
 
-The first stability issue with the standard approach is that the number of uncensored observations in the test set may be insufficient to construct a well-defined partial likelihood, leaving the overall estimate of cross-validation error undefined as well.
+First, the number of uncensored observations in the test set may be insufficient to construct a well-defined partial likelihood for standard cross-validation, leaving the overall estimate of cross-validation error undefined as well.
 When computing the partial likelihood, only individuals whose events are observed will contribute to the likelihood.
-If there is no event observed in a particular fold, then the partial likelihood for that fold is undefined.
+If there are no events observed in a particular fold, then the partial likelihood for that fold is undefined.
 
-To illustrate how common this problem is, we conducted a simulation study.
-We generated 1000 independent data sets of $n = 100$ observations with fixed censoring rate.
-For each data set, we conducted a 10-fold cross-validation and recorded whether or not there was at least one event in each fold (i.e., whether the partial likelihood was well-defined for all folds).
+To illustrate how common this problem is, we generated 1000 independent data sets of $n = 100$ observations with a fixed censoring rate.
+For each data set, we conducted 10-fold cross-validation and recorded whether or not there was at least one event in each fold (i.e., whether the partial likelihood was well-defined for all folds).
 We repeated the same mechanism for scenarios with different censoring rates, ranging from $40\%$ to $80\%$.
-As illustrated in the left panel in Figure~\ref{Fig:stability}, at $n=100$ the standard approach can be problematic even at modest censoring rates.
-For example, standard CV cannot be used with over 20\% of data sets when $70\%$ of all observations are censored.
+As illustrated in the left panel in Figure~\ref{Fig:stability}, at $n=100$ the standard approach can be problematic even at moderate censoring rates.
+For example, standard CV cannot be used with over 20\% of data sets when $70\%$ of observations are censored.
 
@@
   \caption{\label{Fig:stability} Simulations illustrating the stability issues with the standard approach.
-The figure on the left is generated with $n = 100$ observations.
-The horizontal axis is the censoring rate.
-The vertical axis is the proportion of undefined partial likelihood when the fold assignment is not balanced.
-The figure on the right is generated with $n = 120$.
-The fold assignment is balanced in this simulation so that the number of events in each fold is the same.
-The horizontal axis is the number of events per fold.
-The vertical axis is the standard deviation of the squared error loss.
-All four methods were compared in this simulation.
-The standard approach would have much greater variability when there are only few events in the fold.}
+For the figure on the left, $n = 100$, fold assignments are unbalanced, and the vertical axis is the proportion of undefined partial likelihoods.
+For the figure on the right, $n = 120$, fold assignments are balanced so that the number of events in each fold is the same, and the vertical axis is the standard deviation of the squared error loss.
+Basic cross-validation is much more variability than other approaches when there are only few events in each fold.}
 \end{figure}
 
-One way could be used to get around the undefined partial likelihood for the standard approach is to balance the censoring status across all folds when making fold assignments.
-That is to make sure the number of events observed is about the same across all folds.
-Unless the number of the folds exceeds the number of observed events, such as the leave-one-out cross-validation, this approach will guarantee the partial likelihood in each fold is well-defined for the standard approach, thus the overall cross-validated error is well-defined.
+One approach for alleviating this issue is to balance the censoring status across all folds when making fold assignments.
+That is, to ensure the number of events observed is the same (or within 1) across all folds.
+Unless the number of the folds exceeds the number of observed events, such as in leave-one-out cross-validation, this approach will guarantee the partial likelihood in each fold is well-defined for the standard approach, thus the overall cross-validated error is well-defined.
 
-Even if the partial likelihood could be well-defined with balanced fold assignment, there is another drawback of the standard approach: it tends to have more uncertainty in its selection.
-The predictive performance of the model selected by this approach has greaterr variation than the other approaches, especially when there are only a few events available to construct the partial likelihood.
-
-The following simulation study was designed to illustrate this issue.
-We generated data sets with $n = 120$ observations.
-The censoring rate is fixed exactly at $50\%$, thus there are 60 events observed among all observations.
-For each data set, we conducted the balanced fold assignment with number of folds set to be 10, 12, 15, 20, 30 and 60.
-When the number of folds was set to be 10, there were exactly 6 events in each fold; when the number of folds was set to be 60, there was exactly one event in each fold.
+In this case, however, even though the partial likelihood is well-defined, basic cross-validation suffers from extreme variability.
+To illustrate, we generated data sets with $n = 120$ observations and a fixed censoring rate of $50\%$, so that there are 60 observed events.
+For each data set, we carried out cross-validation using balanced fold assignments with the number of folds set to be 10, 12, 15, 20, 30 and 60.
+When the number of folds was set to be 10, there were exactly 6 events in each fold; when the number of folds was set to be 60, there was exactly one event in each fold, and so on.
 After each fold assignment, we computed the cross-validated error using all four approaches described in Section~\ref{Sec:accuracy}.
-We recorded the model that minimizes the cross-validated error for each approach and computed its squared error loss.
+We recorded the model that minimizes the cross-validation error for each approach and computed its squared error loss.
 We replicated the whole process for $N = 200$ time.
 In the end, we computed the standard deviation of the squared error loss for each approach.
+\note{Does this phenomenon show up in the MSE too?}
 
-The results of this simulation are illustrated in the right panel in Figure~\ref{Fig:stability}.
-While all other three approaches are not affected by the number of folds and number of events in each fold, the variability of the model performance has drastically changed for the standard approach.
-When there are sufficient number of events in each fold, the standard approach perform pretty close to the other three approaches.
-However, when there is only one or two events in each fold, the standard approach has much more variability in its predictive performance.
-Even if we can still use the standard approach when there are few number of events, there is a bigger uncertainty in terms of how well the predictive performance would be.
+The results of this simulation are illustrated in the right panel of Figure~\ref{Fig:stability}.
+While the other three approaches are not affected by the number of folds and number of events in each fold, the variability of basic cross-validation increases dramatically as the number of events per fold decreases.
+When there are sufficient number of events in each fold, the standard approach is similar to the other three approaches.
+However, when there are only one or two events in each fold, the standard approach can be up to 10 times more variable than other methods.
+Thus, although the approach can be used to calculate CVE, there is considerable uncertainty surrounding its choice of the best model.
 
-The 60-fold scenario in the last simulation study was also a corner example where there is only one event and one censored observation in each fold.
-In this scenario, it is common to run into cases where the uncensored event happens after the other event is censored.
-In this situation, the risk set at the time right before the event occurred would only contain this observation itself.
-Thus its contribution to the log partial likelihood would be 0, which is well-defined but not informative.
-This situation can happen regardless of the fold size and censoring rate.
-But it would matter more for the standard approach when there are fewer events in the fold.
+To shed some light on why the variability of basic cross-validation explodes in the 60-fold scenario, note that in this case, there is exactly one event and one censored observation in each fold.
+In many folds, the uncensored event happens after the other observation has been censored.
+In this situation, the risk set at the time immediately prior to the event contains only a single observation.
+Its contribution to the log partial likelihood is therefore 0, which is well-defined but not informative.
+This situation can happen regardless of the fold size and censoring rate, but it happens far more often as the number of events per fold decreases.
 
-Based on the two simulations illustrated in this section, we would make the following recommendations: i) always balance the censoring status when conducting fold assignments for the standard approach, ii) avoid using the standard approach in conjuction with high number of folds, iii) avoid using the standard approach when there are few number of events observed.
+\note{Should we have this paragraph?
+Seems to contradict our conclusion, in which our recommendation is to use the linear predictor approach.}  Based on the two simulations illustrated in this section, we would make the following recommendations: i) always balance the censoring status when conducting fold assignments for the standard approach, ii) avoid using the standard approach in conjuction with high number of folds, iii) avoid using the standard approach when there are few number of events observed.
 

### 2019-05-22 main.tex
Edits to 3.3

@@
 Expected censoring percentage is 10$\%$.
-}
+\note{For the sake of consistency, Signal should be capitalized in the x-axis label above.}}
 \end{figure}
@@
 With $N = 200$ replications, the data sets generated have dimension $n = 120$ and $p = 1000$.
-The horizontal axis is the signal strength.
-The verticle axis is the $\log$(MSE ratio).
+The vertical axis is the $\log$(MSE ratio).
 In each panel, the LOO version and 10-fold version of the same approach are plotted.
-The 10-fold standard approach is plotted as the blue line in all three panels as a baseline for comparison.
-}
+In all three panels, 10-fold standard cross-validation is included as a baseline for comparison (blue line).
+\note{Again, Signal should be capitalized.}}
 \end{figure}
@@
 
-On the other hand, the LOOCV version of the linear predictor approach surpasses the standard 10-fold CV by having smaller $\log(\text{MSE ratio})$ across all signal strength levels.
-At the same time, even the 10-fold version of the linear predictor approach also outperforms the leave-one-out version of the V$\&$VH approach and the deviance residual approach.
-Even though the leave-one-out cross validation improves the predictive accuracy for V$\&$VH and deviance residual approach, the magnitude of improvement in predictive performance is smaller than switching to the linear predictor approach.
+On the other hand, the LOOCV version of the linear predictor approach surpasses standard 10-fold CV, with a smaller $\log(\text{MSE ratio})$ across all signal strength levels.
+At the same time, even the 10-fold version of the linear predictor approach outperforms the leave-one-out version of the V$\&$VH approach and the deviance residual approach.
+In other words, although adopting a leave-one-out approach improves the predictive accuracy for V$\&$VH and deviance residual cross-validation, the magnitude of improvement in predictive performance is smaller than the improvement gained from switching to the linear predictor approach.
 

### 2019-05-23 main.tex
Very minor edits to 3.4
change to basic; minor edits to 3.2

@@
 \section{Introduction}
-
-\note{I hate to change the name again, but what do you think about calling it ``basic'' cross-validation instead of ``standard''?
-My concern is that it's not really standard in the sense of ``the thing most people do''; I think most people use V\&VH since that's the default in glmnet.}
 
@@
 
-\par One approach is to simply calculate the partial likelihood based on the observations in the test set as a measure for the model's predictive accuracy.
+\par One basic approach is to simply calculate the partial likelihood based on the observations in the test set as a measure for the model's predictive accuracy.
 In doing so, we calculate the risk for each member of the test set relative to other members of the test set.
@@
 \begin{equation}
-  \label{eq:standard}
+  \label{eq:basic}
   \CVE = -2 \sum_{k=1}^{K} \ell^{k}(\hat{\beta}^{-k});
@@
 the factor of 2 is arbitrary, but often used so that CVE estimates the deviance of the model.
-We refer this approach as \emph{standard cross-validated partial likelihood}. This is implemented in the $\tt{glmnet}$ package as the \verb|"ungrouped"| option.
+We refer this approach as \emph{basic cross-validated partial likelihood}. This is implemented in the $\tt{glmnet}$ package as the \verb|"ungrouped"| option.
 
-\par The standard approach is by far the most common way of conducting cross-validation for models such as linear regression and logistic regression.
-However, the standard approach can be problematic for Cox regression in the sense that that there may not be enough observations in $D_k$ to build up the risk set for partial likelihood.
-For example, the standard approach cannot work with leave-one-out cross-validation, since for all folds, $\ell_k$ would either be zero or undefined.
+\par The basic approach is by far the most common way of conducting cross-validation for models such as linear regression and logistic regression.
+However, the basic approach can be problematic for Cox regression in the sense that that there may not be enough observations in $D_k$ to build up the risk set for partial likelihood.
+For example, the basic approach cannot work with leave-one-out cross-validation, since for all folds, $\ell_k$ would either be zero or undefined.
 
@@
 We refer to this cross-validation method as the Verweij and Van Houwelingen (V\&VH) approach.
-For many models, such as logistic regression or linear regression, the quantities $\ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k})$ and  $\ell^{k}(\hat{\beta}^{-k})$ are equivalent to each other -- in other words, the standard and V\&VH approaches agree.
+For many models, such as logistic regression or linear regression, the quantities $\ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k})$ and  $\ell^{k}(\hat{\beta}^{-k})$ are equivalent to each other -- in other words, the basic and V\&VH approaches agree.
 However, Verweij and Van Houwelingen's approach is more stable for Cox regression as there is always a large number of observations in the risk sets its calculations are based on.
@@
 Hence, another approach to cross-validation would be to include an extra step of estimating the baseline hazard and using it to quantify the cross-validation error.
-The standard approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
+The basic approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
 \begin{equation}
@@
 In this section, we consider the accuracy of the cross-validation approaches introduced in Section~\ref{Sec:methods}.
-We varied the sample size, dimension, censoring percentage \note{(we don't really vary censoring percentage, do we?)}, and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation.
+We varied the sample size, dimension, censoring percentage, and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation.
 Accuracy was assesed via several criteria.
@@
 For each simulated data set, n = 120, p = 1000.
-Expected censoring percentage is 10$\%$.
-\note{For the sake of consistency, Signal should be capitalized in the x-axis label above.}}
+Expected censoring percentage is 10$\%$.}
 \end{figure}
@@
 For all four metrics, the overall pattern is the same (KL scores not shown in Figure~\ref{Fig:mse-brier-c} due to space constraints, but they were very similar to Brier scores).
-Although the four cross-validation methods performed relatively similarly, the standard and linear predictor approaches had the best performance across all metrics: lower MSE, better Brier scores, and higher C Index.
+Although the four cross-validation methods performed relatively similarly, the basic and linear predictor approaches had the best performance across all metrics: lower MSE, better Brier scores, and higher C Index.
 
-\note{Just curious: I would imagine that standard/basic cross-validation does best when there is least censoring, as in Section 3.2.
-Is this true?
-In other words, does Figure 2 make basic cross-validation look better than it often is, because the censoring rate is so small, just 10\%?}
-
-\begin{table}[htb]
+\begin{table}[!htb]
 
@@
 \midrule
-150 & 50 & 5 & weak & V VH  & 0.098 (0.030) & 1.561 (0.851) & 0.245 (0.014) & 0.685 (0.030) & 0.610 (0.035) \\
-    &    &   &     & Standard  & 0.088 (0.025) & 1.531 (0.838) & 0.244 (0.013) & 0.683 (0.029) & 0.614 (0.029) \\
+150 & 50 & 5 & weak & V\&VH  & 0.098 (0.030) & 1.561 (0.851) & 0.245 (0.014) & 0.685 (0.030) & 0.610 (0.035) \\
+    &    &   &     & Basic  & 0.088 (0.025) & 1.531 (0.838) & 0.244 (0.013) & 0.683 (0.029) & 0.614 (0.029) \\
     &    &   &     & LinearPred  & 0.092 (0.029) & 1.543 (0.846) & 0.245 (0.014) & 0.684 (0.030) & 0.612 (0.034) \\
@@
 \addlinespace
-150 & 50 & 5 & strong & V VH  & 0.071 (0.010) & 1.586 (0.989) & 0.176 (0.011) & 0.526 (0.027) & 0.750 (0.012) \\
-    &    &   &     & Standard  & 0.066 (0.011) & 1.544 (0.971) & 0.176 (0.011) & 0.526 (0.028) & 0.749 (0.013) \\
+150 & 50 & 5 & strong & V\&VH  & 0.071 (0.010) & 1.586 (0.989) & 0.176 (0.011) & 0.526 (0.027) & 0.750 (0.012) \\
+    &    &   &     & Basic  & 0.066 (0.011) & 1.544 (0.971) & 0.176 (0.011) & 0.526 (0.028) & 0.749 (0.013) \\
     &    &   &     & LinearPred  & 0.067 (0.010) & 1.540 (0.981) & 0.176 (0.011) & 0.525 (0.028) & 0.750 (0.012) \\
@@
 \addlinespace
-400 & 10000 & 20 & weak & V VH  & 0.126 (0.019) & 2.958 (0.416) & 0.251 (0.012) & 0.696 (0.026) & 0.589 (0.044) \\
-    &    &   &     & Standard  & 0.103 (0.022) & 2.880 (0.437) & 0.244 (0.015) & 0.680 (0.031) & 0.603 (0.039) \\
+400 & 10000 & 20 & weak & V\&VH  & 0.126 (0.019) & 2.958 (0.416) & 0.251 (0.012) & 0.696 (0.026) & 0.589 (0.044) \\
+    &    &   &     & Basic  & 0.103 (0.022) & 2.880 (0.437) & 0.244 (0.015) & 0.680 (0.031) & 0.603 (0.039) \\
     &    &   &     & LinearPred  & 0.105 (0.023) & 2.886 (0.437) & 0.244 (0.015) & 0.681 (0.031) & 0.602 (0.039) \\
@@
 \addlinespace
-400 & 10000 & 20 & strong & V VH  & 0.084 (0.012) & 3.799 (0.548) & 0.176 (0.03) & 0.528 (0.069) & 0.763 (0.048) \\
-    &    &   &     & Standard  & 0.059 (0.007) & 3.455 (0.578) & 0.154 (0.019) & 0.468 (0.049) & 0.780 (0.031) \\
+400 & 10000 & 20 & strong & V\&VH  & 0.084 (0.012) & 3.799 (0.548) & 0.176 (0.03) & 0.528 (0.069) & 0.763 (0.048) \\
+    &    &   &     & Basic  & 0.059 (0.007) & 3.455 (0.578) & 0.154 (0.019) & 0.468 (0.049) & 0.780 (0.031) \\
     &    &   &     & LinearPred  & 0.060 (0.007) & 3.479 (0.574) & 0.155 (0.019) & 0.470 (0.049) & 0.780 (0.030) \\
@@
 In all settings, the censoring rate was 30\%.
-\note{What is s for the strong and weak cases?}
+For weak cases, signal strength $s$ was set to be 0.3; for strong cases, $s$ was set to be 0.6.
 
-As in Figure~\ref{Fig:mse-brier-c}, the standard and linear predictor approaches outperform the deviance residual and V\&VH approaches.
+As in Figure~\ref{Fig:mse-brier-c}, the basic and linear predictor approaches outperform the deviance residual and V\&VH approaches.
 The improvement is particularly noticeable in the most difficult setting: high-dimensional data with weak signal.
-Table~\ref{Tab:sim} also provides some insight into why the deviance residual and V\&VH approaches perform poorly relative to standard CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and thus, selecting fewer variables.
+Table~\ref{Tab:sim} also provides some insight into why the deviance residual and V\&VH approaches perform poorly relative to basic CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and thus, selecting fewer variables.
 
@@
 
-First, the number of uncensored observations in the test set may be insufficient to construct a well-defined partial likelihood for standard cross-validation, leaving the overall estimate of cross-validation error undefined as well.
+First, the number of uncensored observations in the test set may be insufficient to construct a well-defined partial likelihood for basic cross-validation, leaving the overall estimate of cross-validation error undefined as well.
 When computing the partial likelihood, only individuals whose events are observed will contribute to the likelihood.
@@
 We repeated the same mechanism for scenarios with different censoring rates, ranging from $40\%$ to $80\%$.
-As illustrated in the left panel in Figure~\ref{Fig:stability}, at $n=100$ the standard approach can be problematic even at moderate censoring rates.
-For example, standard CV cannot be used with over 20\% of data sets when $70\%$ of observations are censored.
+As illustrated in the left panel in Figure~\ref{Fig:stability}, at $n=100$ the basic approach can be problematic even at moderate censoring rates.
+For example, basic CV cannot be used with over 20\% of data sets when $70\%$ of observations are censored.
 
-\begin{figure}[h]
+\begin{figure}[!htb]
   \centering
   \includegraphics[height= 7cm ]{./manuscript_figure/figure_3.png}
-  \caption{\label{Fig:stability} Simulations illustrating the stability issues with the standard approach.
+  \caption{\label{Fig:stability} Simulations illustrating the stability issues with the basic approach.
 For the figure on the left, $n = 100$, fold assignments are unbalanced, and the vertical axis is the proportion of undefined partial likelihoods.
-For the figure on the right, $n = 120$, fold assignments are balanced so that the number of events in each fold is the same, and the vertical axis is the standard deviation of the squared error loss.
+For the figure on the right, $n = 120$, fold assignments are balanced so that the number of events in each fold is the same, and the vertical axis is the basic deviation of the squared error loss.
 Basic cross-validation is much more variability than other approaches when there are only few events in each fold.}
@@
 That is, to ensure the number of events observed is the same (or within 1) across all folds.
-Unless the number of the folds exceeds the number of observed events, such as in leave-one-out cross-validation, this approach will guarantee the partial likelihood in each fold is well-defined for the standard approach, thus the overall cross-validated error is well-defined.
+Unless the number of the folds exceeds the number of observed events, such as in leave-one-out cross-validation, this approach will guarantee the partial likelihood in each fold is well-defined for the basic approach, thus the overall cross-validated error is well-defined.
 
@@
 In the end, we computed the standard deviation of the squared error loss for each approach.
-\note{Does this phenomenon show up in the MSE too?}
 
@@
 While the other three approaches are not affected by the number of folds and number of events in each fold, the variability of basic cross-validation increases dramatically as the number of events per fold decreases.
-When there are sufficient number of events in each fold, the standard approach is similar to the other three approaches.
-However, when there are only one or two events in each fold, the standard approach can be up to 10 times more variable than other methods.
+When there are sufficient number of events in each fold, the basic approach is similar to the other three approaches.
+However, when there are only one or two events in each fold, the basic approach can be up to 10 times more variable than other methods.
 Thus, although the approach can be used to calculate CVE, there is considerable uncertainty surrounding its choice of the best model.
+The same phenomenon is also observed for the mean squared error.
+When there are only few events in each fold, MSE of the basic approach is affected and increases with the increasing uncertainty.
 
@@
 
-\note{Should we have this paragraph?
-Seems to contradict our conclusion, in which our recommendation is to use the linear predictor approach.}  Based on the two simulations illustrated in this section, we would make the following recommendations: i) always balance the censoring status when conducting fold assignments for the standard approach, ii) avoid using the standard approach in conjuction with high number of folds, iii) avoid using the standard approach when there are few number of events observed.
+Based on the simulation results illustrated in this section, we suggest that one should proceed with caution if they want to use the basic approach: i) always balance the censoring status when conducting fold assignments for the basic approach, ii) avoid using the basic approach in conjuction with high number of folds, iii) avoid using the basic approach when there are few number of events observed.
+Overall we would recommend using the linear predictor approach than the basic approach in practice.
 
@@
 
-As discussed earlier, standard cross-validation cannot be used to carry out leave-one-out cross validation (LOOCV) for Cox regression models.
+As discussed earlier, basic cross-validation cannot be used to carry out leave-one-out cross validation (LOOCV) for Cox regression models.
 The other three approaches (V\&VH, linear predictor, and deviance residual), however, are not subject to the stability issues described in Section~\ref{Sec:stability}, and can be used to implement LOOCV.
@@
 
-\begin{figure}[h]
+\begin{figure}[!htb]
   \centering
@@
 In each panel, the LOO version and 10-fold version of the same approach are plotted.
-In all three panels, 10-fold standard cross-validation is included as a baseline for comparison (blue line).
-\note{Again, Signal should be capitalized.}}
+In all three panels, 10-fold basic cross-validation is included as a baseline for comparison (blue line).}
 \end{figure}
@@
 For each of the V\&VH, linear predictor, and deviance residual approaches, we conducted both 10-fold CV and LOOCV.
-We also conducted the standard 10-fold cross-validation as a baseline for comparison.
+We also conducted the basic 10-fold cross-validation as a baseline for comparison.
 We recorded the model that minimizes the cross-validated error  for each approach.
@@
 The panels illustrate the results for each CV approach, one with the 10-fold version and one with the leave-one-out version.
-The 10-fold standard approach was plotted in all three panels as a baseline for comparison.
+The 10-fold basic approach was plotted in all three panels as a baseline for comparison.
 For all three approaches, the Leave-One-Out version selected models with a smaller $\log(\text{MSE ratio})$ than the 10-fold version.
-For the V\&VH and deviance residual approaches, however, LOOCV was still inferior to standard 10-fold CV.
+For the V\&VH and deviance residual approaches, however, LOOCV was still inferior to basic 10-fold CV.
 As described in Section~\ref{Sec:accuracy}, these approaches tend to be conservative, selecting a larger-than-optimal value of the regularization parameter; LOOCV helps with this problem, but does not alleviate it entirely.
 
-On the other hand, the LOOCV version of the linear predictor approach surpasses standard 10-fold CV, with a smaller $\log(\text{MSE ratio})$ across all signal strength levels.
+On the other hand, the LOOCV version of the linear predictor approach surpasses basic 10-fold CV, with a smaller $\log(\text{MSE ratio})$ across all signal strength levels.
 At the same time, even the 10-fold version of the linear predictor approach outperforms the leave-one-out version of the V$\&$VH approach and the deviance residual approach.
@@
 
-\begin{figure}[h]
+\par All of the cross-validation approaches described in Section~\ref{Sec:methods} use the Cox partial likelihood, in one way or another, as the criterion upon which model selection is based.
+In this section, we compare these likelihood-based approaches against an alternative approach of cross validation using concordance, also known as the cross-validated C Index.
+This is a widely used approach for conducting cross-validation with time-to-event outcomes, particularly for machine learning models \citep{Subramanian2011,Simon2011a}.
+This approach typically proceeds by constructing cross-validated linear predictors, as in Section~\ref{Sec:linear-predictor}, but then using those predictors to calculate concordance as opposed to partial likelihood.
+
+\par We conducted simulations in both low dimensional ($p = 50$, $n = 150$) and high dimensional ($p = 1000$, $n = 120$) settings.
+For each scenario, we generated data from a spectrum of signal strength as in Section~\ref{Sec:accuracy}.
+Results of the simulations are illustrated in Figure~\ref{Fig:cvauc}.
+We used the linear predictor approach (CV-LP) here as the representative approach for likelihood-based cross-validation.
+
+\begin{figure}[!htb]
   \centering
@@
 \end{figure}
-
-\par All of the cross-validation approaches described in Section~\ref{Sec:methods} use the Cox partial likelihood, in one way or another, as the criterion upon which model selection is based.
-In this section, we compare these likelihood-based approaches against an alternative approach of cross validation using concordance, also known as the cross-validated C Index.
-This is a widely used approach for conducting cross-validation with time-to-event outcomes, particularly for machine learning models \citep{Subramanian2011,Simon2011a}.
-This approach typically proceeds by constructing cross-validated linear predictors, as in Section~\ref{Sec:linear-predictor}, but then using those predictors to calculate concordance as opposed to partial likelihood.
-
-\par We conducted simulations in both low dimensional ($p = 50$, $n = 150$) and high dimensional ($p = 1000$, $n = 120$) settings.
-For each of the scenario, we generated data from a spectrum of signal strength as in Section~\ref{Sec:accuracy}.
-Results of the simulations are illustrated in Figure~\ref{Fig:cvauc}.
-We used the linear predictor approach (CV-LP) here as the representative approach for likelihood-based cross-validation.
 
@@
 V $\&$ VH      & 0.639 & 0.637 & 0.638 & 0.637 & 0.637 &  & 0.243 & 0.244 & 0.244 & 0.245 & 0.245 \\
-Standard       & 0.647 & 0.639 & 0.638 & 0.638 & 0.639 &  & 0.233 & 0.243 & 0.244 & 0.244 & 0.244 \\
+Basic       & 0.647 & 0.639 & 0.638 & 0.638 & 0.639 &  & 0.233 & 0.243 & 0.244 & 0.244 & 0.244 \\
 Linear Pred    & \textbf{0.650} & 0.647 & 0.638 & 0.638 & 0.639 &  & \textbf{0.229} & 0.232 & 0.244 & 0.244 & 0.244 \\
@@
 The selected $\lambda$ values, where a given CVE curve reaches its lowest point, are indicated with dotted lines.
-The CVE curves for the standard and linear predictor approaches have greater curvature near their minima, and thus offer more clear support for the selected value of $\lam$.
+The CVE curves for the basic and linear predictor approaches have greater curvature near their minima, and thus offer more clear support for the selected value of $\lam$.
 In comparison, the deviance residual and V\&VH approaches are flatter near their minima, supporting a considerably wider range of $\lam$ values and leading to greater model uncertainty.
@@
 The linear predictor approach was most liberal and selected 85 genes.
-While the standard approach and deviance residual approach selected 68 genes and 60 genes respectively, Verweij and Van Houwelingen's approach was most conservative, selecting only 34 genes.
+While the basic approach and deviance residual approach selected 68 genes and 60 genes respectively, Verweij and Van Houwelingen's approach was most conservative, selecting only 34 genes.
 Similar patterns were observed for other $\alpha$ values.
@@
 In this paper, we propose two cross-validation methods for Cox Regression: cross-validated linear predictors and cross-validated deviance residuals.
-We conducted a series of simulation experiments to study and compare their performance with two existing approaches, the standard cross-validation and Verweij and Van Houwelingen's cross-validated partial likelihood, in selecting tuning parameters when fitting penalized regression models.
+We conducted a series of simulation experiments to study and compare their performance with two existing approaches, the basic cross-validation and Verweij and Van Houwelingen's cross-validated partial likelihood, in selecting tuning parameters when fitting penalized regression models.
 
@@
 
-While the standard approach performed well in terms of predictive performance, this approach suffers from a lack of numerical stability as discussed in Section~\ref{Sec:stability}.
-The cross-validated linear predictor approach performed very similarly to the standard approach in terms of predictive performance, but without any problems due to numerical stability.
+While the basic approach performed well in terms of predictive performance, this approach suffers from a lack of numerical stability as discussed in Section~\ref{Sec:stability}.
+The cross-validated linear predictor approach performed very similarly to the basic approach in terms of predictive performance, but without any problems due to numerical stability.
 We recommend using the proposed linear predictor approach in practice for selecting regularization parameters in penalized Cox regression.

### 2019-05-23 manuscript_code/shedden_elasticnet.R
change shedden data source

@@ -1,0 +1,283 @@
+#------------------------
+#
+# Self contained R code
+# to analyze the Shedden Lung Cancer data set
+# with "MCP" penalty
+#
+#------------------------
+
+#------------------------
+# load/define functions that will be used in the analysis
+library(survival)
+library(ncvreg)
+library(glmnet)
+
+# Brier score for selected lambda
+
+brierSurv_lambda <- function(data,time,fit,lambda) {
+  predicted <- apply(data$x,1, 
+                     function(x)
+                     {predict(fit, x,type = "survival",
+                              lambda = lambda)(time)})
+  is_obs_after <- data$y[,1] > time
+  mean((is_obs_after - predicted)^2)
+}
+
+#------------------------
+# load data
+
+# Shedden Lung Cancer Data
+load(file = "H:\\Courses\\cross_validation\\simulation\\data\\Shedden2008.RData")
+
+# censor S by 72 month
+plot(survfit(S ~ 1), xlab = "month", ylab = "Survival Probability")
+abline(h = 0.5, col = "grey", lty = 2)
+
+# clean data
+ncolX <- ncol(X)
+n <- nrow(X)
+Z <- as.data.frame(Z)
+
+data_low <- data.frame(S,Z)
+data_low$Age <- as.numeric(data_low$Age)
+levels(data_low$Sex) <- c(0,1)
+data_low$Sex <- as.numeric(data_low$Sex)
+data_low$AdjChemo <- as.numeric(data_low$AdjChemo == "Yes")
+
+x <- as.matrix(data.frame(data_low[,c(2,3,5)],X))
+x <- ncvreg::std(x)
+data <- list(y=S,x=x)
+
+# prespecify fold id
+set.seed(123456)
+nfold <- 10
+nevent <- sum(data$y[,2])
+sde <- sqrt(.Machine$double.eps)
+foldid <- numeric(0)
+foldid[data$y[,2] == 1] <-  ceiling(sample(1:nevent)/(nevent + sde) * nfold)
+foldid[data$y[,2] == 0] <-  ceiling(sample(1:(n-nevent))/(n - nevent + sde) * nfold)
+
+
+Alpha <- c(1,0.9,0.7,0.5,0.3)
+A <- length(Alpha)
+
+#------------------------------------------------------------------
+for(a in 1:A){
+  # initial fit
+    alpha <- Alpha[a]
[…]

### 2019-05-23 manuscript_code/shedden_elasticnet.R
push the right shedden file

@@ -1,283 +1,0 @@
-#------------------------
-#
-# Self contained R code
-# to analyze the Shedden Lung Cancer data set
-# with "MCP" penalty
-#
-#------------------------
-
-#------------------------
-# load/define functions that will be used in the analysis
-library(survival)
-library(ncvreg)
-library(glmnet)
-
-# Brier score for selected lambda
-
-brierSurv_lambda <- function(data,time,fit,lambda) {
-  predicted <- apply(data$x,1, 
-                     function(x)
-                     {predict(fit, x,type = "survival",
-                              lambda = lambda)(time)})
-  is_obs_after <- data$y[,1] > time
-  mean((is_obs_after - predicted)^2)
-}
-
-#------------------------
-# load data
-
-# Shedden Lung Cancer Data
-load(file = "H:\\Courses\\cross_validation\\simulation\\data\\Shedden2008.RData")
-
-# censor S by 72 month
-plot(survfit(S ~ 1), xlab = "month", ylab = "Survival Probability")
-abline(h = 0.5, col = "grey", lty = 2)
-
-# clean data
-ncolX <- ncol(X)
-n <- nrow(X)
-Z <- as.data.frame(Z)
-
-data_low <- data.frame(S,Z)
-data_low$Age <- as.numeric(data_low$Age)
-levels(data_low$Sex) <- c(0,1)
-data_low$Sex <- as.numeric(data_low$Sex)
-data_low$AdjChemo <- as.numeric(data_low$AdjChemo == "Yes")
-
-x <- as.matrix(data.frame(data_low[,c(2,3,5)],X))
-x <- ncvreg::std(x)
-data <- list(y=S,x=x)
-
-# prespecify fold id
-set.seed(123456)
-nfold <- 10
-nevent <- sum(data$y[,2])
-sde <- sqrt(.Machine$double.eps)
-foldid <- numeric(0)
-foldid[data$y[,2] == 1] <-  ceiling(sample(1:nevent)/(nevent + sde) * nfold)
-foldid[data$y[,2] == 0] <-  ceiling(sample(1:(n-nevent))/(n - nevent + sde) * nfold)
-
-
-Alpha <- c(1,0.9,0.7,0.5,0.3)
-A <- length(Alpha)
-
-#------------------------------------------------------------------
-for(a in 1:A){
-  # initial fit
-    alpha <- Alpha[a]
[…]

### 2019-05-23 manuscript_code/shedden_elasticnet_git.R

@@ -27,7 +27,10 @@ brierSurv_lambda <- function(data,time,fit,lambda) {
 # load data
 
 # Shedden Lung Cancer Data
-load(file = "Shedden2008.RData")
+data <- readRDS(url("https://s3.amazonaws.com/pbreheny-data-sets/Shedden2008.rds"))
+X <- data$X
+S <- data$S
+Z <- data$Z
 
 # censor S by 72 month
 plot(survfit(S ~ 1), xlab = "month", ylab = "Survival Probability")

### 2019-05-24 main.tex
Submitted to Statistics in Medicine

@@
   \centering
-  \includegraphics[height= 12cm ]{./manuscript_figure/figure_1_new.png}
+  \includegraphics[height= 12cm ]{figure_1_new.png}
   \caption{\label{Fig:methods} Illustration of $K$-fold cross validation methods for Cox regression.}
@@
 Hence, another approach to cross-validation would be to include an extra step of estimating the baseline hazard and using it to quantify the cross-validation error.
-The basic approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
+The standard approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
 \begin{equation}
@@
   \centering
-  \includegraphics[height= 7cm ]{./manuscript_figure/figure_2_new.png}
+  \includegraphics[height= 7cm ]{figure_2_new.png}
   \caption{\label{Fig:mse-brier-c} Simulation comparing 4 CV methods.
@@
   \centering
-  \includegraphics[height= 7cm ]{./manuscript_figure/figure_3.png}
+  \includegraphics[height= 7cm ]{figure_3.png}
   \caption{\label{Fig:stability} Simulations illustrating the stability issues with the basic approach.
 For the figure on the left, $n = 100$, fold assignments are unbalanced, and the vertical axis is the proportion of undefined partial likelihoods.
-For the figure on the right, $n = 120$, fold assignments are balanced so that the number of events in each fold is the same, and the vertical axis is the basic deviation of the squared error loss.
-Basic cross-validation is much more variability than other approaches when there are only few events in each fold.}
+For the figure on the right, $n = 120$, fold assignments are balanced so that the number of events in each fold is the same, and the vertical axis is the standard deviation of the squared error loss.
+Basic cross-validation is much more variable than other approaches when the number of events in each fold is small.}
 \end{figure}
@@
 While the other three approaches are not affected by the number of folds and number of events in each fold, the variability of basic cross-validation increases dramatically as the number of events per fold decreases.
-When there are sufficient number of events in each fold, the basic approach is similar to the other three approaches.
-However, when there are only one or two events in each fold, the basic approach can be up to 10 times more variable than other methods.
-Thus, although the approach can be used to calculate CVE, there is considerable uncertainty surrounding its choice of the best model.
-The same phenomenon is also observed for the mean squared error.
-When there are only few events in each fold, MSE of the basic approach is affected and increases with the increasing uncertainty.
+When there is a sufficient number of events in each fold, the basic approach is similar to the other three approaches.
+However, when there are only one or two events in each fold, the basic approach can be up to 10 times more variable than other methods, with a corresponding increase in the mean squared error.
+In these situations, therefore, even though the basic approach is defined and CVE can be calculated, its performance is very poor.
 
@@
 
-Based on the simulation results illustrated in this section, we suggest that one should proceed with caution if they want to use the basic approach: i) always balance the censoring status when conducting fold assignments for the basic approach, ii) avoid using the basic approach in conjuction with high number of folds, iii) avoid using the basic approach when there are few number of events observed.
-Overall we would recommend using the linear predictor approach than the basic approach in practice.
+Based on the simulation results illustrated in this section, we suggest that one should proceed with caution if they want to use the basic approach: (i) it is best to balance the censoring status when creating fold assignments for the basic approach, (ii) avoid using the basic approach in conjunction with a high number of folds, and (iii) avoid using the basic approach when the number of observed events is small.
+Alternatively, use the linear predictor approach, which has similar performance without the stability issues.
 
@@
   \centering
-  \includegraphics[height= 8cm ]{./manuscript_figure/figure_4.png}
+  \includegraphics[height= 8cm ]{figure_4.png}
   \caption{\label{Fig:loocv}Simulation comparing LOO and 10-fold CV across three different approaches.
@@
 For each of the V\&VH, linear predictor, and deviance residual approaches, we conducted both 10-fold CV and LOOCV.
-We also conducted the basic 10-fold cross-validation as a baseline for comparison.
+We also conducted basic 10-fold cross-validation as a baseline for comparison.
 We recorded the model that minimizes the cross-validated error  for each approach.
@@
   \centering
-  \includegraphics[height= 15cm ]{./manuscript_figure/figure_auc.png}
+  \includegraphics[height= 15cm ]{figure_auc.png}
   \caption{\label{Fig:cvauc}Comparing Partial Likelihood and C Index both built from Cross-validated Linear Predictors.
@@
 The expected percentages of censoring for both scenarios are 30 $\%$.
-The top row are box plots of the selected tuning parameter$\lambda$.
+The top row are box plots of the selected tuning parameter $\lambda$.
 The mid row plots the $\log(\text{relative Squared Error Loss})$, with respect to the oracle model.
@@
   \centering
-  \includegraphics[height= 8cm ]{./manuscript_figure/shedden.png}
+  \includegraphics[height= 8cm ]{shedden.png}
   \caption{\label{fig:CVE}Rescaled Cross Validated Errors at $\alpha = 0.3$.
@@
 In this paper, we propose two cross-validation methods for Cox Regression: cross-validated linear predictors and cross-validated deviance residuals.
-We conducted a series of simulation experiments to study and compare their performance with two existing approaches, the basic cross-validation and Verweij and Van Houwelingen's cross-validated partial likelihood, in selecting tuning parameters when fitting penalized regression models.
+We conducted a series of simulation experiments to study and compare their performance with two existing approaches, basic cross-validation and Verweij and Van Houwelingen's cross-validated partial likelihood, in selecting tuning parameters when fitting penalized regression models.
 

### 2019-05-24 sub/sim/cover.tex
Submitted to Statistics in Medicine

\documentclass{uiletter}
\usepackage{hyperref}
\hypersetup{colorlinks=true, citecolor=black, urlcolor=cyan}

\begin{document}

\StartLetter{}{Dear editors of {\em Statistics in Medicine},}

We are submitting the manuscript ``Cross validation approaches for penalized Cox regression'' for your consideration to be published in {\em Statistics in Medicine}.

Using cross-validation (CV) for model selection is fairly straightforward in most models.  However, for models based on partial likelihood, such as Cox regression, we obtain predictions only of relative risk, not absolute risk, thereby complicating the cross-validation paradigm.  This work proposes a new approach for CV in these models and demonstrates that it achieves a combination of performance and stability that existing methods do not match.  We believe this is an accessible article of clear practical importance, and would be interesting to readers of {\em Statistics in Medicine}.

Code to reproduce all results in the manuscript, including the raw data for the gene expression analysis, is available at \url{https://github.com/biydai/Cross-Validation}.  This repository is currently private, but will be made public upon acceptance of this manuscript.

This work has not been previously published and is not under review elsewhere.  Thank you very much for your consideration.

\EndLetter

\end{document}

### 2019-05-24 sub/sim/coxcv.tex
Submitted to Statistics in Medicine

--- 2019-05-21 cross-validation.tex
@@
-\documentclass{article}
-\usepackage{graphicx}
+\documentclass[AMA,STIX1COL]{WileyNJD-v2}
+\articletype{Research Article}
+\raggedbottom
+
 \usepackage{color}
-
 \usepackage{alltt}
-\usepackage{amssymb,amsmath,natbib,graphicx,enumerate,subcaption,tikz,url,booktabs}
-\usepackage[bmargin=0.75in, tmargin =0.75in,lmargin = 0.5in,rmargin = 0.5in]{geometry}
-
-\usepackage[ruled]{algorithm2e}
-\usepackage{algpseudocode}
-\renewcommand{\algorithmcfname}{Procedure}
+\usepackage{amssymb,amsmath,graphicx,enumerate,subcaption,booktabs}
+%\usepackage[bmargin=0.75in, tmargin =0.75in,lmargin = 0.5in,rmargin = 0.5in]{geometry}
+\graphicspath{{../../manuscript_figure/}}
+\usepackage{natbib}
 \usepackage{bbold}
+\hypersetup{colorlinks=true, citecolor=black, urlcolor=cyan}
 
@@
 
+\begin{document}
+
 \title{Cross Validation Approaches for Penalized Cox Regression}
-\author{Biyue Dai\\Department of Biostatistics\\University of Iowa
-  \and
-  Patrick Breheny\\Department of Biostatistics\\University of Iowa}
-\date{\today}
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\begin{document}
-\maketitle
 
-\section*{Abstract}
-Cross validation is commonly used for selecting tuning parameters in penalized regression, but its use in penalized Cox regression models has received relatively little attention in the literature.
+\author[1]{Biyue Dai}
+
+\author[1]{Patrick Breheny}
+
+\authormark{Dai and Breheny}
+
+\address[1]{\orgdiv{Department of Biostatistics}, \orgname{University of Iowa}, \orgaddress{\state{Iowa}, \country{USA}}}
+
+\corres{Patrick Breheny \\ Department of Biostatistics \\ 145 N Riverside Dr, N336 CPHB \\ Iowa City, IA 52242 \\ \email{patrick-breheny@uiowa.edu}}
+
+\abstract[Summary]{Cross validation is commonly used for selecting tuning parameters in penalized regression, but its use in penalized Cox regression models has received relatively little attention in the literature.
 Due to its partial likelihood construction, carrying out cross validation for Cox models is not straightforward, and there are several potential approaches for implementation.
@@
 Our proposed approach of cross-validating the linear predictors seems to offer an attractive balance of performance and numerical stability.
-We illustrate these advantages using simulated data as well as using them to analyze data from a high-dimensional study of survival in lung cancer patients.
+We illustrate these advantages using simulated data as well as using them to analyze data from a high-dimensional study of survival in lung cancer patients.}
+
+\keywords{Model selection, Cross validation, Cox regression, Lasso, Penalized regression}
+
+\jnlcitation{\cname{%
+\author{Dai B} and
+\author{Breheny P}} (\cyear{2019}),
+\ctitle{Cross Validation Approaches for Penalized Cox Regression}, \cjournal{Statistics in Medicine}, \cvol{xxxx; xx:xx-xx}.}
+
+\maketitle
+
 \input{main}
 
-\bibliographystyle{abbrvnat}
-%\setcitestyle{round} 
-\bibliography{articles}
+\subsection*{Data and code availability}
+
+Code to reproduce all results in this manuscript, including the raw data for the gene expression analysis, is available at \url{https://github.com/biydai/Cross-Validation}.
+NOTE: This repository is currently private, but will be made public upon acceptance of this manuscript.
+
+\subsection*{Supporting information}
+
+A short document supporting our approach for estimating $\Lambda_0$ in the deviance residual method is provided as supporting information.
+
+\bibliographystyle{WileyNJD-AMA}
+\bibliography{../../articles}
 

### 2019-11-26 sub/smimr/coxcv.tex
Starting new submission folder for Statistical Methods in Medical Research

--- 2019-05-21 cross-validation.tex
@@
 \usepackage{color}
+\graphicspath{{../../manuscript_figure/}}
+\usepackage[doublespacing]{setspace}  % Required for SMIMR review
 
 \usepackage{alltt}
-\usepackage{amssymb,amsmath,natbib,graphicx,enumerate,subcaption,tikz,url,booktabs}
-\usepackage[bmargin=0.75in, tmargin =0.75in,lmargin = 0.5in,rmargin = 0.5in]{geometry}
+\usepackage{amssymb,amsmath,graphicx,enumerate,subcaption,tikz,url,booktabs}
+\usepackage[bmargin=5cm, tmargin=5cm, lmargin=3cm, rmargin=3cm]{geometry}
+\usepackage{hyperref}
+\hypersetup{colorlinks=true, citecolor=black, urlcolor=cyan}
 
-\usepackage[ruled]{algorithm2e}
-\usepackage{algpseudocode}
-\renewcommand{\algorithmcfname}{Procedure}
+\usepackage{natbib}
 \usepackage{bbold}
@@
 
-\bibliographystyle{abbrvnat}
-%\setcitestyle{round} 
+\bibliographystyle{SageH}
 \bibliography{articles}

### 2019-12-06 main.tex
Starting new submission folder for Statistical Methods in Medical Research
fix table width, add subsection on power

@@
 	\begin{equation} 
-	L(\hat{\eta}^{cv}) = \prod_{i=1}^{n} \frac{\exp (\hat{\eta}^{cv}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}^{cv}_{j})}.
+	L(\hat{\eta}^{cv}) = \prod_{i=1}^{n} \left\{ \frac{\exp (\hat{\eta}^{cv}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}^{cv}_{j})} \right\} ^{\delta_i}.
 	\end{equation}
@@
   \centering
-  \includegraphics[height= 7cm ]{figure_2_new.png}
+  \includegraphics[width=\textwidth]{figure_2_new.png}
   \caption{\label{Fig:mse-brier-c} Simulation comparing 4 CV methods.
@@
 
-\caption{\label{Tab:sim} Simulation comparing four cross-validation methods with varying dimension and signal strength}
+\caption{\label{Tab:sim} Comparing the predictive accuracy of four cross-validation methods under scenarios with varying dimension and signal strength}
 \centering
-\begin{tabular}[t]{cccccccccc}
+\begin{tabular}[t]{lccccc}
 \toprule
- n& p& $\#$ Non-Zero & Signal & Method & $\lambda$ & log(MSE Ratio) &Brier Score & Kullback Leibler & C Index\\
+ Method & $\lambda$ & log(MSE Ratio) &Brier Score & Kullback Leibler & C Index\\
 \midrule
-150 & 50 & 5 & weak & V\&VH  & 0.098 (0.030) & 1.561 (0.851) & 0.245 (0.014) & 0.685 (0.030) & 0.610 (0.035) \\
-    &    &   &     & Basic  & 0.088 (0.025) & 1.531 (0.838) & 0.244 (0.013) & 0.683 (0.029) & 0.614 (0.029) \\
-    &    &   &     & LinearPred  & 0.092 (0.029) & 1.543 (0.846) & 0.245 (0.014) & 0.684 (0.030) & 0.612 (0.034) \\
-    &    &   &     & DevResid  & 0.098 (0.029) & 1.547 (0.848) & 0.245 (0.014) & 0.684 (0.030) & 0.611 (0.035) \\
+&\multicolumn{5}{c}{\textbf{Scenario 1}: n = 150, p = 50, 5 non-zero features, weak signal}\\
+\cline{2-6}
+V\&VH  & 0.098 (0.030) & 1.561 (0.851) & 0.245 (0.014) & 0.685 (0.030) & 0.610 (0.035) \\
+Basic  & 0.088 (0.025) & 1.531 (0.838) & 0.244 (0.013) & 0.683 (0.029) & 0.614 (0.029) \\
+LinearPred  & 0.092 (0.029) & 1.543 (0.846) & 0.245 (0.014) & 0.684 (0.030) & 0.612 (0.034) \\
+DevResid  & 0.098 (0.029) & 1.547 (0.848) & 0.245 (0.014) & 0.684 (0.030) & 0.611 (0.035) \\
 \addlinespace
-150 & 50 & 5 & strong & V\&VH  & 0.071 (0.010) & 1.586 (0.989) & 0.176 (0.011) & 0.526 (0.027) & 0.750 (0.012) \\
-    &    &   &     & Basic  & 0.066 (0.011) & 1.544 (0.971) & 0.176 (0.011) & 0.526 (0.028) & 0.749 (0.013) \\
-    &    &   &     & LinearPred  & 0.067 (0.010) & 1.540 (0.981) & 0.176 (0.011) & 0.525 (0.028) & 0.750 (0.012) \\
-    &    &   &     & DevResid  & 0.094 (0.011) & 1.995 (0.841) & 0.181 (0.011) & 0.539 (0.026) & 0.751 (0.013) \\
+&\multicolumn{5}{c}{\textbf{Scenario 2}: n = 150, p = 50, 5 non-zero features, strong signal}\\
+\cline{2-6}
+V\&VH  & 0.071 (0.010) & 1.586 (0.989) & 0.176 (0.011) & 0.526 (0.027) & 0.750 (0.012) \\
+Basic  & 0.066 (0.011) & 1.544 (0.971) & 0.176 (0.011) & 0.526 (0.028) & 0.749 (0.013) \\
+LinearPred  & 0.067 (0.010) & 1.540 (0.981) & 0.176 (0.011) & 0.525 (0.028) & 0.750 (0.012) \\
+DevResid  & 0.094 (0.011) & 1.995 (0.841) & 0.181 (0.011) & 0.539 (0.026) & 0.751 (0.013) \\
 \addlinespace
-400 & 10000 & 20 & weak & V\&VH  & 0.126 (0.019) & 2.958 (0.416) & 0.251 (0.012) & 0.696 (0.026) & 0.589 (0.044) \\
-    &    &   &     & Basic  & 0.103 (0.022) & 2.880 (0.437) & 0.244 (0.015) & 0.680 (0.031) & 0.603 (0.039) \\
-    &    &   &     & LinearPred  & 0.105 (0.023) & 2.886 (0.437) & 0.244 (0.015) & 0.681 (0.031) & 0.602 (0.039) \\
-    &    &   &     & DevResid  & 0.107 (0.022) & 2.895 (0.433) & 0.245 (0.014) & 0.683 (0.03) & 0.602 (0.039) \\
+&\multicolumn{5}{c}{\textbf{Scenario 3}: n = 400, p = 10000, 20 non-zero features, weak signal}\\
+\cline{2-6}
+V\&VH  & 0.126 (0.019) & 2.958 (0.416) & 0.251 (0.012) & 0.696 (0.026) & 0.589 (0.044) \\
+Basic  & 0.103 (0.022) & 2.880 (0.437) & 0.244 (0.015) & 0.680 (0.031) & 0.603 (0.039) \\
+LinearPred  & 0.105 (0.023) & 2.886 (0.437) & 0.244 (0.015) & 0.681 (0.031) & 0.602 (0.039) \\
+DevResid  & 0.107 (0.022) & 2.895 (0.433) & 0.245 (0.014) & 0.683 (0.03) & 0.602 (0.039) \\
 \addlinespace
-400 & 10000 & 20 & strong & V\&VH  & 0.084 (0.012) & 3.799 (0.548) & 0.176 (0.03) & 0.528 (0.069) & 0.763 (0.048) \\
-    &    &   &     & Basic  & 0.059 (0.007) & 3.455 (0.578) & 0.154 (0.019) & 0.468 (0.049) & 0.780 (0.031) \\
-    &    &   &     & LinearPred  & 0.060 (0.007) & 3.479 (0.574) & 0.155 (0.019) & 0.470 (0.049) & 0.780 (0.030) \\
-    &    &   &     & DevResid  & 0.076 (0.018) & 3.708 (0.518) & 0.166 (0.024) & 0.502 (0.057) & 0.772 (0.049) \\
+&\multicolumn{5}{c}{\textbf{Scenario 4}: n = 400, p = 10000, 20 non-zero features, weak signal}\\
+\cline{2-6}
+V\&VH  & 0.084 (0.012) & 3.799 (0.548) & 0.176 (0.03) & 0.528 (0.069) & 0.763 (0.048) \\
+Basic  & 0.059 (0.007) & 3.455 (0.578) & 0.154 (0.019) & 0.468 (0.049) & 0.780 (0.031) \\
+LinearPred  & 0.060 (0.007) & 3.479 (0.574) & 0.155 (0.019) & 0.470 (0.049) & 0.780 (0.030) \\
+DevResid  & 0.076 (0.018) & 3.708 (0.518) & 0.166 (0.024) & 0.502 (0.057) & 0.772 (0.049) \\
 \bottomrule
@@
 Table~\ref{Tab:sim} also provides some insight into why the deviance residual and V\&VH approaches perform poorly relative to basic CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and thus, selecting fewer variables.
+
+\subsection{Power}
+\label{Sec:power}
+While cross-validation is mainly a measure of the model's predictive accuracy, it is not sufficient to only compare those methods based on predictive accuracy.
+Another important criterion is that how many important variables are correctly selected by the model.
+The following simulations are conducted to compare the methods in this aspect.
+
+In Table~\ref{Tab:power}, we computed the average number of true positives selected by each cross-validation approach.
+True positives are features that are selected in the fitted model and their true $\beta$ coefficients are non-zero in the data-generating model.
+The simulation scenarios vary from low dimension, with 150 observations and 50 features, to high dimension, with 400 observations and 10000 features.
+We generate data with both weak signal strength and strong signal strength.
+Across all simulation scenarios, features are independent from each other.
+The results of the simulation is presented in Table~\ref{Tab:power}.
+
+\begin{table}[!htb]
+\setlength{\tabcolsep}{3pt}
+\caption{\label{Tab:power} Number of non-zero features (true positives) selected by each cross-validation method, under scenarios with varying dimension and signal strength.}
+\centering
+\begin{tabular}[t]{l cc c cc c cc}
+\toprule
+&\multicolumn{2}{c}{n = 150, p = 50} & & \multicolumn{2}{c}{n = 300, p = 1000} & & \multicolumn{2}{c}{n = 400, p = 10000} \\
+&\multicolumn{2}{c}{$\#$ non-zero features: 5} & & \multicolumn{2}{c}{$\#$ non-zero features: 20} & & \multicolumn{2}{c}{$\#$ non-zero features: 20} \\
+\cline{2-3} \cline{5-6} \cline{8-9}
+& weak signal & strong signal & & weak signal & strong signal && weak signal & strong signal\\
+V\&VH        & 3.87 & 5& & 10.96 & 19.61 & & 6.67 & 18.38\\
+Basic          & 4.06 & 5& & 13.68 & 19.96 & & 9.58 & 19.87\\
+LinearPred  & 3.99 & 5& & 13.46 & 19.95 & & 9.40 & 19.85\\
+DevResid    & 3.90 & 5& & 12.96 & 19.38 & & 8.98 & 19.05\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+When the signal strength in the data is weak, it is more difficult to pick up the signal and identify the non-zero features.
+Hence there is a big difference among the numbers of true positives selected by different cross-validation approaches, as is illustrated in Table~\ref{Tab:power}.
+In these scenarios, the basic cross-validation approach and linear predictor approach are more powerful, while the V$\&$VH approach always has the least number of true positive features.
+Especially in the high dimensional cases, the V$\&$VH approach select about three features fewer than the other approaches.
+In practice, missing important findings is a real concern when applying those methods.
 
@@
   \centering
-  \includegraphics[height= 7cm ]{figure_3.png}
+  \includegraphics[width=\textwidth]{figure_3.png}
   \caption{\label{Fig:stability} Simulations illustrating the stability issues with the basic approach.
@@
   \centering
-  \includegraphics[height= 8cm ]{figure_4.png}
+  \includegraphics[width=\textwidth]{figure_4.png}
   \caption{\label{Fig:loocv}Simulation comparing LOO and 10-fold CV across three different approaches.
@@
 \caption{\label{tb:shedden} The C Index and Brier Score of all elasticnet models selected by cross-validation methods }
-\begin{tabular}{lccccccccccc}
+\begin{tabular}{lccccc}
 \hline
-  & \multicolumn{5}{c}{Cross-validated C Index} & \multicolumn{5}{c}{Brier Score} \\ 
-  & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ &  
- & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-6} \cline{8-12} 
-V $\&$ VH      & 0.639 & 0.637 & 0.638 & 0.637 & 0.637 &  & 0.243 & 0.244 & 0.244 & 0.245 & 0.245 \\
-Basic       & 0.647 & 0.639 & 0.638 & 0.638 & 0.639 &  & 0.233 & 0.243 & 0.244 & 0.244 & 0.244 \\
-Linear Pred    & \textbf{0.650} & 0.647 & 0.638 & 0.638 & 0.639 &  & \textbf{0.229} & 0.232 & 0.244 & 0.244 & 0.244 \\
-Deviance Resid & 0.645 & 0.637 & 0.638 & 0.637 & 0.637 &  & 0.235 & 0.244 & 0.244 & 0.245 & 0.245 \\ \hline
+  & \multicolumn{5}{c}{Cross-validated C Index} \\ 
+  & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-6}
+V $\&$ VH      & 0.639 & 0.637 & 0.638 & 0.637 & 0.637\\
+Basic       & 0.647 & 0.639 & 0.638 & 0.638 & 0.639 \\
+Linear Pred    & \textbf{0.650} & 0.647 & 0.638 & 0.638 & 0.639  \\
+Deviance Resid & 0.645 & 0.637 & 0.638 & 0.637 & 0.637 \\ 
+\addlinespace
+  & \multicolumn{5}{c}{Brier Score} \\ 
+  & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-6}
+V $\&$ VH  & 0.243 & 0.244 & 0.244 & 0.245 & 0.245 \\
+Basic  & 0.233 & 0.243 & 0.244 & 0.244 & 0.244 \\
+Linear Pred & \textbf{0.229} & 0.232 & 0.244 & 0.244 & 0.244 \\
+Deviance Resid & 0.235 & 0.244 & 0.244 & 0.245 & 0.245 \\ 
+\hline
 \end{tabular}

### 2019-12-06 main.tex
add correlated case to supporting materials

@@
 
-\subsection{Power}
+\bnote{Simulations with features that has more complicated correlation structures were conducted as well.
+Similar patterns were observed as in Figure~\ref{Fig:mse-brier-c} and Table~\ref{Tab:sim}.
+Results of these simulations are provided in the supporting document.}
+
+\subsection{\bnote{Power}}
 \label{Sec:power}
+
 While cross-validation is mainly a measure of the model's predictive accuracy, it is not sufficient to only compare those methods based on predictive accuracy.

### 2019-12-06 supplementary.tex
Submitted to Statistics in Medicine
add correlated case to supporting materials

@@
 
-\title{Supplementary Material\\
+\title{Supporting Information\\
 \textit{Cross Validation Approaches for Penalized Cox Regression}}
@@
 
-\section{Cross-Validated Deviance Residuals}
-
+\section{Baseline Estimation for Cross-Validated Deviance Residuals}
+\subsection{Cross-Validated Deviance Residuals}
 The Cox model implies the following relationship between the baseline cumulative hazard and the cumulative hazard for individual $i$:
@@
 
-\section{Baseline Estimations}
+\subsection{Baseline Estimations}
 
-	\subsection{Nelson-Aalen Estimator}
+	\subsubsection{Nelson-Aalen Estimator}
 	 The simplest approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
@@
 
-	\subsection{Kalbfleisch and Prentice's Approach}					
+	\subsubsection{Kalbfleisch and Prentice's Approach}					
 	\cite{Kalbfleisch2011} proposed a cumulative hazard estimator that incorporates covariates.
@@
 
-\section{Simulation Comparison}
+\subsection{Simulation Comparison}
 The following simulation illustrates how the performance of cross-validated deviance residuals would vary based on different baseline estimation.
@@
 
+\section{Simulation with Correlated Features}
+
+Simulations in this section are conducted to examine how correlations among features can impact the performance of the cross-validation approaches.
+
+The data generating mechanism is similar to simulations presented in Figure 2 in the manuscript, except that each non-zero feature is correlated with 10 null features.
+We considered two types of correlation structures: exchangeable and autoregressive.
+We simulated data with a spectrum of correlation, from 0 to 0.8.
+Results of the simulation is presented in Figure~\ref{Fig:ex} for exchangeable correlation structure and Figure~\ref{Fig:auto} for autoregressive correlation structure.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{manuscript_figure/figure_2_ex.png}
+  \caption{\label{Fig:ex} The horizontal axis in all three plots is the correlation among the features.
+The value of $s$ is varied from 0.4 to 0.9.
+$\log(\text{MSE ratio})$, relative to the oracle modle is plotted in the left panel.
+Out-of-sample Brier scores are plotted in the middle.
+Out-of-sample C index is plotted in the right panel.
+For each simulated data set, n = 120, p = 1000.
+Expected censoring percentage is 10$\%$.}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{manuscript_figure/figure_2_auto.png}
+  \caption{\label{Fig:auto} The horizontal axis in all three plots is the correlation among the features.
+The value of $s$ is varied from 0.4 to 0.9.
+$\log(\text{MSE ratio})$, relative to the oracle modle is plotted in the left panel.
+Out-of-sample Brier scores are plotted in the middle.
+Out-of-sample C index is plotted in the right panel.
+For each simulated data set, n = 120, p = 1000.
+Expected censoring percentage is 10$\%$.}
+\end{figure}
+
+As is shown in the figures, across different correlation structure and strength of the correlation, the basic approach and the linear predictor appraoch had the best performance of having lower MSE, smaller Brier scores and higher C Index, whereas the V$\&$VH approach has the worst performance.
+
 \bibliographystyle{abbrvnat}
-
 \bibliography{articles}

### 2019-12-13 manuscript_code/figure_2_auto.R
update R code for additional sim (correlated; power)

@@ -1,0 +1,390 @@
+#------------------------------
+# Self-contained R code
+# to reproduce simulation results
+# for the correlated case in Figure 2 
+# of the Cross Validation Manuscript
+# scenario: increasing signal
+# in n = 120, p = 1000
+#------------------------------
+
+library(survival)
+library(ncvreg)
+
+#-------------------------------------------------------------------------
+# functions
+genS <- function(p, rho, corr) {
+  if (corr=='exchangeable') {
+    S <- matrix(rho, p, p) + (1-rho)*diag(p)
+  } else if (corr=='autoregressive') {
+    RHO <- matrix(rho^(0:(p-1)), p, p, byrow=TRUE)
+    S <- Matrix::bandSparse(p, k=0:(p-1), diagonals=RHO, symmetric=TRUE)
+  }
+  S
+}
+
+genX <- function(n, p, S) {
+  R <- chol(S)
+  as.matrix(matrix(rnorm(n*p), n, p) %*% R)
+}
+
+genSurvData <- function(n=100, p=60, a=6, b=2, rho=0.5, noise=c('exchangeable', 'autoregressive'),
+                        rho.noise=0, beta = 0.5,
+                        h = 1,
+                        cpct = 0.70,# 70% means 70% of the data are censored; 30% are observed
+                        corr = c("exchangeable","autoregressive")
+){
+  noise <- match.arg(noise)
+  corr <- match.arg(corr)
+  K <- b + 1
+  
+  # Gen X, S
+  sigmaList <- vector('list', a+1)
+  if(corr == "exchangeable"){
+    for (i in 1:a) {
+    sigmaList[[i]] <- matrix(rho, K, K) + (1-rho)*diag(K)
+    }
+  }
+  if(corr == "autoregressive"){
+    for (i in 1:a) {
+      RHO <- matrix(rho^(0:(K-1)), K, K, byrow=TRUE)
+      sigmaList[[i]] <- Matrix::bandSparse(K, k=0:(K-1), diagonals=RHO, symmetric=TRUE)
+    }
+  }
+  sigmaList[[a+1]] <- genS(p-K*a, rho.noise, noise)
+  S <- Matrix::.bdiag(sigmaList)
+  X <- genX(n, p, S)
+  
+  # Gen beta
+    bb <- beta*(c(-1,1)[(1:a)%%2+1])
+    beta <- numeric(p)
+    beta[((1:a)-1)*K+1] <- bb
+  
[…]

### 2019-12-13 manuscript_code/figure_2_ex.R

@@ -1,0 +1,389 @@
+#------------------------------
+# Self-contained R code
+# to reproduce simulation results
+# for the correlated case in Figure 2 
+# of the Cross Validation Manuscript
+# scenario: increasing signal
+# in n = 120, p = 1000
+#------------------------------
+
+library(survival)
+library(ncvreg)
+
+#-------------------------------------------------------------------------
+# functions
+genS <- function(p, rho, corr) {
+  if (corr=='exchangeable') {
+    S <- matrix(rho, p, p) + (1-rho)*diag(p)
+  } else if (corr=='autoregressive') {
+    RHO <- matrix(rho^(0:(p-1)), p, p, byrow=TRUE)
+    S <- Matrix::bandSparse(p, k=0:(p-1), diagonals=RHO, symmetric=TRUE)
+  }
+  S
+}
+
+genX <- function(n, p, S) {
+  R <- chol(S)
+  as.matrix(matrix(rnorm(n*p), n, p) %*% R)
+}
+
+genSurvData <- function(n=100, p=60, a=6, b=2, rho=0.5, noise=c('exchangeable', 'autoregressive'),
+                        rho.noise=0, beta = 0.5,
+                        h = 1,
+                        cpct = 0.70,# 70% means 70% of the data are censored; 30% are observed
+                        corr = c("exchangeable","autoregressive")
+){
+  noise <- match.arg(noise)
+  corr <- match.arg(corr)
+  K <- b + 1
+  
+  # Gen X, S
+  sigmaList <- vector('list', a+1)
+  if(corr == "exchangeable"){
+    for (i in 1:a) {
+      sigmaList[[i]] <- matrix(rho, K, K) + (1-rho)*diag(K)
+    }
+  }
+  if(corr == "autoregressive"){
+    for (i in 1:a) {
+      RHO <- matrix(rho^(0:(K-1)), K, K, byrow=TRUE)
+      sigmaList[[i]] <- Matrix::bandSparse(K, k=0:(K-1), diagonals=RHO, symmetric=TRUE)
+    }
+  }
+  sigmaList[[a+1]] <- genS(p-K*a, rho.noise, noise)
+  S <- Matrix::.bdiag(sigmaList)
+  X <- genX(n, p, S)
+  
+  # Gen beta
+  bb <- beta*(c(-1,1)[(1:a)%%2+1])
+  beta <- numeric(p)
+  beta[((1:a)-1)*K+1] <- bb
+  
[…]

### 2019-12-13 manuscript_code/power_table/combine_results.R

@@ -1,0 +1,135 @@
+setwd("A:\\cv_manuscript_code\\fdr_power")
+
+library(abind)
+
+gp_combined <- NULL
+dvr_combined <- NULL
+ncv_combined <- NULL
+ug_combined <- NULL
+
+#--------------------------------------------------------------
+# low dimension
+#--------------------------------------------------------------
+
+load("low_fdr.RData")
+
+results <- list()
+for(i in 1:4){
+  results[[i]] <- rbind(
+    apply(gp[i,,],2,mean),
+    apply(ncv[i,,],2,mean),
+    apply(ug[i,,],2,mean),
+    apply(dvr[i,,],2,mean)
+  )
+  rownames(results[[i]]) <- c("VVH","LP","ST","DVR")
+  colnames(results[[i]]) <- c(0.3,0.6)
+}
+names(results) <- c("lambda",
+                    "MSE",
+                    "tp",
+                    "fdr")
+
+results$lambda
+results$MSE
+results$tp
+results$fdr
+
+results_low <- results
+
+#--------------------------------------------------------------
+# middle dimension
+#--------------------------------------------------------------
+
+gp_combined <- NULL
+dvr_combined <- NULL
+ncv_combined <- NULL
+ug_combined <- NULL
+
+for(i in 1:20){
+  
+  load(paste("mid_fdr_",i,".RData",sep=""))
+  gp_1 <- gp
+  gp_combined <- abind(gp_combined,gp_1,along = 2)
+
+  dvr_1 <- dvr
+  dvr_combined <- abind(dvr_combined,dvr_1,along = 2)
+  
+  ncv_1 <- ncv
+  ncv_combined <- abind(ncv_combined,ncv_1,along = 2)
+  
+  ug_1 <- ug
+  ug_combined <- abind(ug_combined,ug_1,along = 2)
+  
+}
+
+results <- list()
+for(i in 1:4){
+  results[[i]] <- rbind(
+    apply(gp_combined[i,,],2,mean),
+    apply(ncv_combined[i,,],2,mean),
+    apply(ug_combined[i,,],2,mean),
+    apply(dvr_combined[i,,],2,mean)
+  )
+  rownames(results[[i]]) <- c("VVH","LP","ST","DVR")
+  colnames(results[[i]]) <- c(0.2,0.6)
+}
+names(results) <- c("lambda",
+                    "MSE",
+                    "tp",
+                    "fdr")
+
+results$lambda
[…]

### 2019-12-13 manuscript_code/power_table/power_high_dim.R

@@ -1,0 +1,211 @@
+#------------------------------
+# Self-contained R code
+# to reproduce part of the simulations
+# in Table 1 of the Cross Validation 
+# Manuscript:
+# Explored 2 Scenarios here: 
+# low dim low signal
+# low dim high signal
+#------------------------------
+
+library(survival)
+library(ncvreg)
+
+args <- as.numeric(commandArgs(TRUE)[1])
+
+#-------------------------------------------------------------------------
+# functions
+
+generatebeta <- function(p,nbeta,c = 1){
+  pos <- rep(c, ceiling(nbeta/2))
+  neg <- rep(-c, floor(nbeta/2))
+  beta <- c(pos,neg,rep(0,p-nbeta))
+}
+
+genSurvData <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,h = 1 # 70% means 70% of the data are censored; 30% are observed
+){
+  p <- length(beta)
+  ## Generate column vector of coefficients
+  beta <- matrix(beta, ncol = 1)
+  x <- matrix(rnorm(n*p, sd = sd),n,p)
+  #standardize x first
+  x <- ncvreg::std(x)
+  hazard <- h*exp(x %*% beta)
+  ## Calculate survival times
+  time <- rexp(n, hazard)
+  ## observations are censored ("event = rbinom(n, 1, 0.30)")
+  y <- Surv(time = time, event = rbinom(n, 1, 1-cpct))
+  return(list(x=x,y=y,beta = beta))
+}
+
+
+#-------------------------------------------------------------------------
+# scenarios: 
+N <- 10
+p <- 10000
+n <- 400
+c <- c(0.3,0.6)
+nbeta <- 20
+
+nfold <- 10
+beta <- matrix(NA,nrow = length(c), ncol = p)
+for (i in 1:length(c)){
+  beta[i,]<- generatebeta(p,nbeta = nbeta ,c = c[i])
+}
+
+# initiation
+gp <- array(NA,c(4,N,length(c)))
+ug <- array(NA,c(4,N,length(c)))
+ncv <- array(NA,c(4,N,length(c)))
+dvr <- array(NA,c(4,N,length(c)))
+oracle.mse <- matrix(NA, nrow = N, ncol = length(c))
+foldid <- numeric(n)
+
+set.seed(999 + args*100)
+for(i in 1:N){
[…]

### 2019-12-13 manuscript_code/power_table/power_low_dim.R

@@ -1,0 +1,210 @@
+#------------------------------
+# Self-contained R code
+# to reproduce part of the simulations
+# in Table 1 of the Cross Validation 
+# Manuscript:
+# Explored 2 Scenarios here: 
+# low dim low signal
+# low dim high signal
+#------------------------------
+
+library(survival)
+library(ncvreg)
+
+#-------------------------------------------------------------------------
+# functions
+
+generatebeta <- function(p,nbeta,c = 1){
+  pos <- rep(c, ceiling(nbeta/2))
+  neg <- rep(-c, floor(nbeta/2))
+  beta <- c(pos,neg,rep(0,p-nbeta))
+}
+
+genSurvData <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,h = 1 # 70% means 70% of the data are censored; 30% are observed
+){
+  p <- length(beta)
+  ## Generate column vector of coefficients
+  beta <- matrix(beta, ncol = 1)
+  x <- matrix(rnorm(n*p, sd = sd),n,p)
+  #standardize x first
+  x <- ncvreg::std(x)
+  hazard <- h*exp(x %*% beta)
+  ## Calculate survival times
+  time <- rexp(n, hazard)
+  ## observations are censored ("event = rbinom(n, 1, 0.30)")
+  y <- Surv(time = time, event = rbinom(n, 1, 1-cpct))
+  return(list(x=x,y=y,beta = beta))
+}
+
+
+#-------------------------------------------------------------------------
+# scenarios: 
+N <- 200
+p <- 50
+n <- 150
+c <- c(0.3,0.6)
+nbeta <- 5
+
+nfold <- 10
+beta <- matrix(NA,nrow = length(c), ncol = p)
+for (i in 1:length(c)){
+  beta[i,]<- generatebeta(p,nbeta = nbeta ,c = c[i])
+}
+
+# initiation
+gp <- array(NA,c(4,N,length(c)))
+ug <- array(NA,c(4,N,length(c)))
+ncv <- array(NA,c(4,N,length(c)))
+dvr <- array(NA,c(4,N,length(c)))
+oracle.mse <- matrix(NA, nrow = N, ncol = length(c))
+foldid <- numeric(n)
+
+
+for(i in 1:N){
+  set.seed(999 + i)
+  for (j in 1:length(c)){
+    # generate data
[…]

### 2019-12-13 manuscript_code/power_table/power_mid_dim.R

@@ -1,0 +1,211 @@
+#------------------------------
+# Self-contained R code
+# to reproduce part of the simulations
+# in Table 1 of the Cross Validation 
+# Manuscript:
+# Explored 2 Scenarios here: 
+# low dim low signal
+# low dim high signal
+#------------------------------
+
+library(survival)
+library(ncvreg)
+
+args <- as.numeric(commandArgs(TRUE)[1])
+
+#-------------------------------------------------------------------------
+# functions
+
+generatebeta <- function(p,nbeta,c = 1){
+  pos <- rep(c, ceiling(nbeta/2))
+  neg <- rep(-c, floor(nbeta/2))
+  beta <- c(pos,neg,rep(0,p-nbeta))
+}
+
+genSurvData <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,h = 1 # 70% means 70% of the data are censored; 30% are observed
+){
+  p <- length(beta)
+  ## Generate column vector of coefficients
+  beta <- matrix(beta, ncol = 1)
+  x <- matrix(rnorm(n*p, sd = sd),n,p)
+  #standardize x first
+  x <- ncvreg::std(x)
+  hazard <- h*exp(x %*% beta)
+  ## Calculate survival times
+  time <- rexp(n, hazard)
+  ## observations are censored ("event = rbinom(n, 1, 0.30)")
+  y <- Surv(time = time, event = rbinom(n, 1, 1-cpct))
+  return(list(x=x,y=y,beta = beta))
+}
+
+
+#-------------------------------------------------------------------------
+# scenarios: 
+N <- 10
+p <- 1000
+n <- 300
+c <- c(0.3,0.6)
+nbeta <- 20
+
+nfold <- 10
+beta <- matrix(NA,nrow = length(c), ncol = p)
+for (i in 1:length(c)){
+  beta[i,]<- generatebeta(p,nbeta = nbeta ,c = c[i])
+}
+
+# initiation
+gp <- array(NA,c(4,N,length(c)))
+ug <- array(NA,c(4,N,length(c)))
+ncv <- array(NA,c(4,N,length(c)))
+dvr <- array(NA,c(4,N,length(c)))
+oracle.mse <- matrix(NA, nrow = N, ncol = length(c))
+foldid <- numeric(n)
+
+set.seed(999 + args*100)
+for(i in 1:N){
[…]

### 2019-12-16 sub/smimr/scrub-url.R
scrub-url.R to fix bib file for SMIMR submission

@@ -1,0 +1,6 @@
+bib <- readLines('../../articles.bib')
+bib <- bib[!grepl('eprint', bib)]
+url <- grepl("url\\s*=", bib)
+rproj <- grepl("R-project\\.org", bib)
+bib <- bib[!url | rproj]
+writeLines(bib, 'articles.bib')

### 2019-12-16 main.tex
Edits to deviance residuals section

@@
 
-The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model.
-Hence, another approach to cross-validation would be to include an extra step of estimating the baseline hazard and using it to quantify the cross-validation error.
-The standard approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
+\note{Biyue, I commented most of this section out, rather than deleting it, because it would probably make sense to move some of this to the supplement.
+Can you please handle migrating this content to the supplement so that all the relevant equations/notation are defined?}
+
+The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model, which prevents us from calculating prediction errors (residuals).
+However, if one takes additional steps to estimate the baseline hazard in addition to estimating the coefficients, then such residuals can be calculated.
+The deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}, is a natural candidate for quantifying prediction error to use in quantifying the cross-validation error.
+Originally proposed for model diagnostics, deviance residuals have also been occasionally used as a predictive accuracy measure \citep{Therneau2018}.
+Given an estimate of the baseline cumulative hazard function, $\hat{\Lambda}_0(t)$, the cross-validated Martingale residual is given by
 \begin{equation}
-  \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}.
+  m^{cv}_{i} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}),
 \end{equation}
-The Cox model implies the following relationship between the baseline cumulative hazard and the cumulative hazard for individual $i$:
+and the cross-validated deviance residuals by
 \begin{equation} 
-  \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta).
+  d_{i} = \text{sgn}(m^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}.
 \end{equation}
-
-The baseline hazard could be incorporated into cross-validation in a variety of ways.
-Here, we consider the following approach: for fold $k$, we obtain $\hat{\beta}^{-k}$ from the training set $T_k$ and then for each $i$ in the test set $D_k$,
-\begin{equation}
- \label{eq:cv-bhz}
-  \hat{\Lambda}^{cv}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta^{-k}).
-\end{equation}
-Note that this uses an estimate of the baseline hazard coming from the full data under the null model, but cross-validation has been applied to the estimation of the regression coefficients $\hat{\beta}$; we will discuss the rationale for this estimator shortly.
-
-A natural candidate for incorporating the cumulative hazard into a loss function is the deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}.
-First proposed for model diagnostics, deviance residuals have also been occasionally used as a predictive accuracy measure \citep{Therneau2018}.
-Given \eqref{eq:cv-bhz}, we can first obtain the cross-validated Martingale residual:
-\begin{equation}
-  \hat{M^{cv}_{i}} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}).
-\end{equation}
-The cross-validated deviance residuals can then be derived from the Martingale residuals: 
-\begin{equation} 
-  d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}.
-\end{equation}
-The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross validated error, analogously to using the residual sum of squares in linear regression.
+The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross validated error, analogous to using the residual sum of squares in linear regression.
 We refer to this method as the \emph{cross-validated deviance residuals} approach.
 
-The baseline hazard estimate we describe here differs from the one typically used in constructing deviance residuals \citep{Therneau1990}.
+In order to use this approach, of course, one must estimate the baseline hazard, and there are a variety of ways this may be accomplished.
+We considered several approaches and found them to give more or less similar results, so in this manuscript, we present the baseline hazard estimation method that performed the best overall as the ``cross-validated deviance residuals'' approach.
+Details of this method, the other methods we considered, and simulation results comparing these methods, are available in the supplemental materials.
+
+%% The standard approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
+%% \begin{equation}
+%%   \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}.
+%% \end{equation}
+%% The Cox model implies the following relationship between the baseline cumulative hazard and the cumulative hazard for individual $i$:
+%% \begin{equation}
+%%   \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta).
+%% \end{equation}
+
+%% The baseline hazard could be incorporated into cross-validation in a variety of ways.
+Here, we consider the following approach: for fold $k$, we obtain $\hat{\beta}^{-k}$ from the training set $T_k$ and then for each $i$ in the test set $D_k$,
+%% \begin{equation}
+%%  \label{eq:cv-bhz}
+%%   \hat{\Lambda}^{cv}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta^{-k}).
+%% \end{equation}
+%% Note that this uses an estimate of the baseline hazard coming from the full data under the null model, but cross-validation has been applied to the estimation of the regression coefficients $\hat{\beta}$; we will discuss the rationale for this estimator shortly.
+
+%% The baseline hazard estimate we describe here differs from the one typically used in constructing deviance residuals \citep{Therneau1990}.
 Instead of the Nelson-Aalen estimator, deviance residuals are typically calculated based on a baseline hazard estimate that has been adjusted for the covariates in the model \citep{breslow1972}.

### 2019-12-17 main.tex
Added a few sentences on standardization

@@
 where $P_{\lambda}(\beta)$ is a penalty function that depends on a regularization parameter $\lam$.
+Throughout, we will assume without loss of generality that the covariates have been standardized such that $\sum_i x_{ij} = 0$ and $\tfrac{1}{n}\sum_i x_{ij}^2=1$.
+This is standard practice in penalized regression to ensure that the penalization has the same effect for all covariates and is invariant to changes of location and scale.
+The estimates are then converted back to the original scale after the model has been fit.
+Centering the predictors is also standard in Cox regression to ensure that the baseline hazard is invariant to location shifts of the covariates.
 
@@
 	\begin{equation} 
-	L(\hat{\eta}^{cv}) = \prod_{i=1}^{n} \left\{ \frac{\exp (\hat{\eta}^{cv}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}^{cv}_{j})} \right\} ^{\delta_i}.
+	L(\hat{\eta}^{cv}) = \prod_{i=1}^{n} \left\{ \frac{\exp (\hat{\eta}^{cv}_{i})}{\sum_{k \in R(t_{i})}\exp (\hat{\eta}^{cv}_k)} \right\} ^{\delta_i}.
 	\end{equation}

### 2019-12-19 supplementary.tex
update supplement simulation; still working on writing

@@
 
+	\subsubsection{Breslow's Approach}
+
 	\subsubsection{Kalbfleisch and Prentice's Approach}					
@@
 \vspace{1mm}
-CVDR - Nelson Aalen & 0.1531 & 0.1346 & 0.1251 & 0.1217 & 0.1213 & 0.1210\\
-CVDR - Kalbfleisch and Prentice& 0.2534 & 0.2275 & 0.2084 & 0.1767 & 0.1588 & 0.1414\\
-CVLP & 0.1441 & 0.1214 & 0.1078 & 0.0981 & 0.0914 & 0.0849\\
+Nelson Aalen & 0.152 & 0.134 & 0.127 & 0.121 & 0.121 & 0.121\\
+Breslow & 0.258 & 0.247 & 0.228 & 0.199 & 0.178 & 0.165\\
+Kalbfleisch and Prentice & 0.249 & 0.234 & 0.210 & 0.180 & 0.155 & 0.144\\
+%CVLP & 0.1441 & 0.1214 & 0.1078 & 0.0981 & 0.0914 & 0.0849\\
 %Oracle &  0.1090 & 0.0941 & 0.0762 & 0.0615 & 0.0567 & 0.0516\\
+\vspace{0.5mm}\\
+ & \multicolumn{5}{c}{$\log($MSE Ratio$)$} & \\ 
 \vspace{1mm}
- & \multicolumn{5}{c}{$\log($MSE$)$} & \\ 
-\vspace{1mm}
-
-CVDR - Nelson Aalen & 2.4038 & 2.5107 & 2.8289 & 3.0084 & 3.1723 & 3.3463\\
-CVDR - Kalbfleisch and Prentice& 2.7075 & 2.9397 & 3.2592 & 3.3596 & 3.4228 & 3.4934\\
-CVLP &2.3684 & 2.4089 & 2.6595 & 2.7173 & 2.7761 & 2.8264\\
+Nelson Aalen & 2.431 & 2.624 & 2.806 & 2.983 & 3.186 & 3.346 \\
+Breslow & 2.729 & 3.074 & 3.301 & 3.449 & 3.562 & 3.632 \\
+Kalbfleisch and Prentice & 2.712 & 3.034 & 3.232 & 3.355 & 3.427 & 3.508 \\
+%CVLP &2.3684 & 2.4089 & 2.6595 & 2.7173 & 2.7761 & 2.8264\\
 \bottomrule
@@
 
-The bottom half of the table listed $\log($MSE$)$ of the models selected by cross-validation approaches compared to the oracle model.
-The definition of $\log($MSE$)$ is described in more details in Section 3.1 in the paper.
-Throughout different signal strength, CVDR with Kalbfleisch and Prentice baseline had the largest $\log($MSE$)$, indicating worst model estimation among all methods.
+The bottom half of the table listed $\log($MSE Ratio$)$ of the models selected by cross-validation approaches compared to the oracle model.
+The definition of $\log($MSE Ratio$)$ is described in more details in Section 3.1 in the paper.
+Throughout different signal strength, CVDR with Kalbfleisch and Prentice baseline had the largest $\log($MSE Ratio$)$, indicating worst model estimation among all methods.
 With Nelson-Aalen estimator, the CVDR approach had better performance and performed closer to CVLP at weaker signals.

### 2019-12-19 manuscript_code/supplementary_git.R

@@ -38,10 +38,10 @@ p <- 1000
 c <- 4:9/10
 rep <- 200
 
-result_lambda <- array(NA,dim = c(5,length(c),rep),dimnames = list(c("NA","Full","CV","LP","Oracle"),
+result_lambda <- array(NA,dim = c(6,length(c),rep),dimnames = list(c("NA","Full","Full_breslow","CV","LP","Oracle"),
                                                                      as.character(c),
                                                                      NULL))
-result_MSER <- array(NA,dim = c(5,length(c),rep),dimnames = list(c("NA","Full","CV","LP","Oracle"),
+result_MSER <- array(NA,dim = c(6,length(c),rep),dimnames = list(c("NA","Full","Full_breslow","CV","LP","Oracle"),
                                                                  as.character(c),
                                                                  NULL))
 set.seed(1)
@@ -115,6 +115,22 @@ for(R in 1:rep){
                                                log(d - M))))
       sum(dev^2)})
     
+    #-------------------------------------------------------
+    # with eta from full fit: KP
+    BH_full_breslow <- apply(w, 2, function(x){ 
+      r <- rev(cumsum(rev(x)))
+      h <- d/r
+      cumsum(h)
+    }
+    )
+    BH_full_breslow <- BH_full_breslow[,1:ncol(w.cv)]
+    
+    val_full_breslow <- apply(w.cv*BH_full_breslow, 2, function(x){ 
+      M <- d - x
+      dev <- sign(M) * sqrt(-2 * (M + ifelse(d == 0, 0, d * 
+                                               log(d - M))))
+      sum(dev^2)})
+    
     
     #-------------------------------------------------------
     # with cross-validated eta (old KP)
@@ -138,7 +154,7 @@ for(R in 1:rep){
 
 #-------------------------------------------------------
 # extract results
-  id <- c(which.min(val_NA), which.min(val_full), which.min(val_cv), which.min(pl),
+  id <- c(which.min(val_NA), which.min(val_full), which.min(val_full_breslow),which.min(val_cv), which.min(pl),
           which.min(log_MSE_ratio))
   result_lambda[,j,R] <- lambda[id]
     

### 2020-03-25 main.tex
Revisions to deviance residuals section

@@
 A partial likelihood can then be built using this set of linear predictors to evaluate the predictive accuracy of the model:
-	\begin{equation} 
+\begin{equation*}
   L(\hat{\eta}^{cv}) = \prod_{i=1}^{n} \left\{ \frac{\exp (\hat{\eta}^{cv}_{i})}{\sum_{k \in R(t_{i})}\exp (\hat{\eta}^{cv}_k)} \right\} ^{\delta_i}.
-	\end{equation}
+\end{equation*}
 We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - 2 \log L(\hat{\eta}^{cv}).$$
@@
 
-\note{Biyue, I commented most of this section out, rather than deleting it, because it would probably make sense to move some of this to the supplement.
-Can you please handle migrating this content to the supplement so that all the relevant equations/notation are defined?}
-
 The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model, which prevents us from calculating prediction errors (residuals).
@@
 Originally proposed for model diagnostics, deviance residuals have also been occasionally used as a predictive accuracy measure \citep{Therneau2018}.
-Given an estimate of the baseline cumulative hazard function, $\hat{\Lambda}_0(t)$, the cross-validated Martingale residual is given by
-\begin{equation}
-  m^{cv}_{i} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}),
-\end{equation}
-and the cross-validated deviance residuals by
-\begin{equation} 
-  d_{i} = \text{sgn}(m^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}.
-\end{equation}
+In the Cox model, the Martingale residuals $\{m_i\}_{i=1}^n$ and deviance residuals $\{d_i\}_{i=1}^n$ are given by
+\begin{align}
+  \label{eq:res-m}
+  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(X_i\Tr\beta), \\
+  \label{eq:res-d}
+  d_{i} &= \text{sign}(m_{i})\sqrt{-2(m_i + \delta_{i}\log(\delta_{i} - m_i))}.  
+\end{align}
+where $H_0$ is the cumulative baseline hazard.
+
+In order to obtain a cross-validated version of these residuals, we would replace $\beta$ in the above equations with $\hat{\beta}^{-k}$.
+However, it is not obvious how to obtain a cross-validated version of $H_0$, or if it is even necessary to do so.
+We experimented with three versions of cross-validated deviance residuals, each with a different choice for $H_0$: (1) The Nelson-Aalen estimator, which does not involve the covariates and is therefore the same across cross-validation folds, (2) the Breslow estimator \citep{Breslow1972}, and (3) the Kalbfleisch-Prentice estimator \citep{Kalbfleisch2011}.
+Details for the three approaches are given in the supplementary materials, along with the results of a simulation study comparing them.
+Because the Nelson-Aalen baseline hazard performed best in all situations, we use this estimate to construct the cross-validated deviance residuals throughout the rest of the paper.
+
 The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross validated error, analogous to using the residual sum of squares in linear regression.
 We refer to this method as the \emph{cross-validated deviance residuals} approach.
-
-In order to use this approach, of course, one must estimate the baseline hazard, and there are a variety of ways this may be accomplished.
-We considered several approaches and found them to give more or less similar results, so in this manuscript, we present the baseline hazard estimation method that performed the best overall as the ``cross-validated deviance residuals'' approach.
-Details of this method, the other methods we considered, and simulation results comparing these methods, are available in the supplemental materials.
-
-%% The standard approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
-%% \begin{equation}
-%%   \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}.
-%% \end{equation}
-%% The Cox model implies the following relationship between the baseline cumulative hazard and the cumulative hazard for individual $i$:
-%% \begin{equation}
-%%   \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta).
-%% \end{equation}
-
-%% The baseline hazard could be incorporated into cross-validation in a variety of ways.
-Here, we consider the following approach: for fold $k$, we obtain $\hat{\beta}^{-k}$ from the training set $T_k$ and then for each $i$ in the test set $D_k$,
-%% \begin{equation}
-%%  \label{eq:cv-bhz}
-%%   \hat{\Lambda}^{cv}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta^{-k}).
-%% \end{equation}
-%% Note that this uses an estimate of the baseline hazard coming from the full data under the null model, but cross-validation has been applied to the estimation of the regression coefficients $\hat{\beta}$; we will discuss the rationale for this estimator shortly.
-
-%% The baseline hazard estimate we describe here differs from the one typically used in constructing deviance residuals \citep{Therneau1990}.
-Instead of the Nelson-Aalen estimator, deviance residuals are typically calculated based on a baseline hazard estimate that has been adjusted for the covariates in the model \citep{breslow1972}.
-In the context of cross-validation, however, this approach is problematic.
-Deviance residuals measure the difference between the fitted model and a saturated model; in Cox regression, this saturated model depends on the baseline hazard.
-Thus, a covariate-adjusted baseline hazard would mean that each fold is compared to a different saturated model.
-For this reason, it is important that the baseline hazard remains constant across folds when calculating cross-validated deviance residuals; this intuition is borne out by simulations involving various other possible ways of constructing cross-validated deviance residuals (see Supporting Information).
-Note that this is not the case for the Brier Score or Kullback Leibler Score (described in Section~\ref{Sec:accuracy}), which involve absolute loss functions, not comparisons with a saturated model.
 

### 2020-03-30 main.tex
baseline info in supplementary

@@
 In doing so, we calculate the risk for each member of the test set relative to other members of the test set.
-One drawback to this approach is that it becomes unstable when the size of the the test set is small.
+One drawback to this approach is that it becomes unstable when the size of the test set is small.
 In particular, it cannot be applied to leave-one-out cross-validation (LOOCV), as we must have at least two observations in the test set to compare their risk relative to each other.
@@
 
-\bnote{Simulations with features that has more complicated correlation structures were conducted as well.
+\bnote{More simulations were conducted to study scenarios where features have more complicated correlation structures.
 Similar patterns were observed as in Figure~\ref{Fig:mse-brier-c} and Table~\ref{Tab:sim}.
-Results of these simulations are provided in the supporting document.}
+Results of these simulations are provided in the supplementary material.}
 

### 2020-03-30 supplementary.tex
baseline info in supplementary

@@
 \begin{equation}
-  \hat{M^{cv}_{i}} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}).
+  \hat{M}^{cv}_{i} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}).
 \end{equation}
@@
 \end{equation}
-	The Nelson-Aalen Estimator does not involve any information from the covariates. It only depends on the failure times.
+Under the assumption that there is no ties among failure times, the numerator in the summation is either 0 or 1.
+The Nelson-Aalen Estimator does not involve any information from the covariates.
+It only depends on the failure times.
+When we use the Nelson-Aalen Estimator for the cross-validated deviance residuals, we follow the below algorithm:
+\begin{algorithm}
+\SetAlgoLined
+Use all observations' time and status to compute $\hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}$\;
+\For{each regularization parameter $\lambda$}{
+ \For{each fold $k = 1,2,...,K$}{
+  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
+  	\For{each obervation $i$ in the $k$th fold}{
+  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
+  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
+  	}
+ }
+ Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
+ }
+ \caption{Compute Cross-validated Deviance Residuals Based on Nelson-Aalen Estimator}
+\end{algorithm}
 
@@
 
+In the original proposal of the Martingale Residuals \citep{Therneau1990}, Breslow's estimator \citep{Breslow1974} was the prefered approach for estimating the cumulative baseline hazard.
+Under the assumption that there is no ties among the failure times, the Breslow's estimator is defined as:
+\begin{equation} 
+\hat{\Lambda}_0(t) = \sum_{t_{i} \leq t}\frac{\delta(t_i)}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})}.
+\end{equation}
+Covariates of the fitted model are incorporated in the denominator of the summation. When we use Breslow's estimator for the cross-validated deviance residuals, we follow the below algorithm:
+\begin{algorithm}
+\SetAlgoLined
+\For{each regularization parameter $\lambda$}{
+ Fit the model using all observations' time and status \;
+ Obtain linear predictor $\hat{\eta}$ from the fitted model\;
+ Compute $\hat{\Lambda}_0(t) = \sum_{t_{i} \leq t}\frac{\delta(t_i)}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})}$\;
+ \For{each fold $k = 1,2,...,K$}{
+  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
+  	\For{each obervation $i$ in the $k$th fold}{
+  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
+  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
+  	}
+ }
+ Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
+ }
+ \caption{Compute Cross-validated Deviance Residuals Based on Breslow's Approach}
+\end{algorithm}
+
 \subsubsection{Kalbfleisch and Prentice's Approach}
+
 	\cite{Kalbfleisch2011} proposed a cumulative hazard estimator that incorporates covariates.
@@
 \begin{equation}
-1 - \hat{\alpha}_{i} = 1 - \left\{ 1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}exp (\hat{\eta}_{j}} \right\}^{exp(-\hat{\eta}_{i})},
+1 - \hat{\alpha}_{i} = 1 - \left( 1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})} \right)^{\exp(-\hat{\eta}_{i})},
 \end{equation}
@@
 \begin{equation} 
-\hat{\Lambda}_0(t) = \sum(1 - \hat{\alpha}_i)\mathds{1}(t_{i} \leq t),
+\hat{\Lambda}_0(t) = \sum_{t_{i} \leq t}(1 - \hat{\alpha}_i),
 \end{equation}
+When we use the Kalbfleisch and Prentice's Estimator for the cross-validated deviance residuals, we follow the below algorithm:
+\begin{algorithm}
+\SetAlgoLined
+\For{each regularization parameter $\lambda$}{
+ Fit the model using all observations' time and status \;
+ Obtain linear predictor $\hat{\eta}$ from the fitted model\;
+ Compute $\hat{\Lambda}_0(t) = \sum_{t_{i} \leq t} \left\{1 -  \left(1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})} \right) ^{\exp(-\hat{\eta}_{i})}\right\}$\;
+ \For{each fold $k = 1,2,...,K$}{
+  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
+  	\For{each obervation $i$ in the $k$th fold}{
+  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
+  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
+  	}
+ }
+ Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
+ }
+ \caption{Compute Cross-validated Deviance Residuals Based on Kalbfleisch and Prentice's Estimator}
+\end{algorithm}
 

### 2020-05-06 supplementary.tex
Edits to deviance residuals (section and supplement)

--- 2019-12-19 supplementary.tex
@@
 \section{Baseline Estimation for Cross-Validated Deviance Residuals}
-\subsection{Cross-Validated Deviance Residuals}
-The Cox model implies the following relationship between the baseline cumulative hazard and the cumulative hazard for individual $i$:
-\begin{equation}
-  \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta).
-\end{equation}
-Consider the following approach: for fold $k$, we obtain $\hat{\beta}^{-k}$ from the training set $T_k$ and then for each $i$ in the test set $D_k$, 
-\begin{equation}
- \label{eq:cv-bhz}
-  \hat{\Lambda}^{cv}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta^{-k}).
-\end{equation}
 
-A natural candidate for incorporating the cumulative hazard into a loss function is the deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}.
-Given \eqref{eq:cv-bhz}, we can first obtain the cross-validated Martingale residual:
-\begin{equation}
-  \hat{M^{cv}_{i}} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}).
-\end{equation}
-The cross-validated deviance residuals can then be derived from the Martingale residuals: 
-\begin{equation} 
-  d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}.
-\end{equation}
-The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross validated error, analogously to using the residual sum of squares in linear regression.
-We refer to this method as the \emph{cross-validated deviance residuals} approach.
+As described in the manuscript, the Martingale residuals $\{m_i\}_{i=1}^n$ and deviance residuals $\{d_i\}_{i=1}^n$ for the Cox model are given by
+\begin{align*}
+  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(X_i\Tr\beta) \\
+  d_{i} &= \text{sign}(m_{i})\sqrt{-2(m_i + \delta_{i}\log(\delta_{i} - m_i))},
+\end{align*}
+where $H_0$ is the cumulative baseline hazard.  Our cross-validated versions of these residuals use
+\begin{align*}
+  m_{i}^{cv} &= \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}) \\
+  d_{i}^{cv} &= \text{sign}(m_{i}^{cv})\sqrt{-2(m_i^{cv} + \delta_{i}\log(\delta_{i} - m_i^{cv}))},
+\end{align*}
+for $i$ in the test set $D_k$, with $\hat{\beta}^{-k}$ obtained from the training set $T_k$.
+The sum of squared cross-validated deviance residuals, $\sum_{i}(d_{i}^{cv})^2$, are then used as the cross validated error, analogously to using the residual sum of squares in linear regression.
+In the manuscript, this method is referred to as the \emph{cross-validated deviance residuals} approach.
 
@@
 	\subsubsection{Nelson-Aalen Estimator}
-	 The simplest approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
+	 The simplest approach to estimating the cumulative baseline hazard, $H_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
 \begin{equation}
-  \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}.
+  \hat{H}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}.
 \end{equation}
-	The Nelson-Aalen Estimator does not involve any information from the covariates. It only depends on the failure times.
+Under the assumption that there is no ties among failure times, the numerator in the summation is either 0 or 1.
+The Nelson-Aalen Estimator depends only on the failure times and does not involve any information from the covariates.
+Procedure~\ref{alg:na} provides an explicit description of this algorithm.
+
+\begin{algorithm}
+\SetAlgoLined
+Use all observations' time and status to compute $\hat{H}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}$\;
+\For{each regularization parameter $\lambda$}{
+ \For{each fold $k = 1,2,...,K$}{
+  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
+  	\For{each obervation $i$ in the $k$th fold}{
+  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
+  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
+  	}
+ }
+ Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
+ }
+ \caption{\label{alg:na} Compute cross-validated deviance residuals based on the Nelson-Aalen estimator}
+\end{algorithm}
 
@@
 
+In the original proposal of the Martingale Residuals \citep{Therneau1990}, Breslow's estimator \citep{Breslow1974} was the prefered approach for estimating the cumulative baseline hazard.
+Under the assumption that there is no ties among the failure times, the Breslow's estimator is defined as:
+\begin{equation} 
+\hat{H}_0(t) = \sum_{t_{i} \leq t}\frac{\delta(t_i)}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})}.
+\end{equation}
+Covariates of the fitted model are incorporated in the denominator of the summation.
+Procedure~\ref{alg:br} provides an explicit description of this algorithm.
+\begin{algorithm}
+\SetAlgoLined
+\For{each regularization parameter $\lambda$}{
+ Fit the model using all observations' time and status \;
+ Obtain linear predictor $\hat{\eta}$ from the fitted model\;
+ Compute $\hat{H}_0(t) = \sum_{t_{i} \leq t}\frac{\delta(t_i)}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})}$\;
+ \For{each fold $k = 1,2,...,K$}{
+  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
+  	\For{each obervation $i$ in the $k$th fold}{
+  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
+  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
+  	}
+ }
+ Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
+ }
+ \caption{\label{alg:br} Compute cross-validated deviance residuals based on the Breslow estimator}
+\end{algorithm}
+
 \subsubsection{Kalbfleisch and Prentice's Approach}
+
 \cite{Kalbfleisch2011} proposed a cumulative hazard estimator that incorporates covariates.
@@
 \begin{equation} 
-\hat{\Lambda}_0(t) = \sum(1 - \hat{\alpha}_i)\mathds{1}(t_{i} \leq t),
+\hat{H}_0(t) = \sum_{t_{i} \leq t}(1 - \hat{\alpha}_i),
 \end{equation}
+Procedure~\ref{alg:kp} provides an explicit description of this algorithm.
+\begin{algorithm}
+\SetAlgoLined
+\For{each regularization parameter $\lambda$}{
+ Fit the model using all observations' time and status \;
+ Obtain linear predictor $\hat{\eta}$ from the fitted model\;
+ Compute $\hat{H}_0(t) = \sum_{t_{i} \leq t} \left\{1 -  \left(1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})} \right) ^{\exp(-\hat{\eta}_{i})}\right\}$\;
+ \For{each fold $k = 1,2,...,K$}{
+  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
+  	\For{each obervation $i$ in the $k$th fold}{
+  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
+  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
+  	}
+ }
+ Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
+ }
+ \caption{\label{alg:kp} Compute cross-validated deviance residuals based on the Kalbfleisch and Prentice estimator}
+\end{algorithm}
 

### 2020-05-18 supplementary.tex
Edit supplementary procedure

@@
 Covariates of the fitted model are incorporated in the denominator of the summation.
-Procedure~\ref{alg:br} provides an explicit description of this algorithm.
+Procedure~\ref{alg:brkp} includes an explicit description of this algorithm.
 \begin{algorithm}
@@
  Obtain linear predictor $\hat{\eta}$ from the fitted model\;
- Compute $\hat{H}_0(t) = \sum_{t_{i} \leq t}\frac{\delta(t_i)}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})}$\;
+ \uIf{Use the Breslow's estimator}{
+ $\hat{H}_0(t) = \sum_{t_{i} \leq t}\frac{\delta(t_i)}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})}$;
+ }\ElseIf{Use the Kalbfleisch and Prentice estimator}{
+  $\hat{H}_0(t) = \sum_{t_{i} \leq t} \left\{1 -  \left(1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})} \right) ^{\exp(-\hat{\eta}_{i})}\right\}$\;
+ }
  \For{each fold $k = 1,2,...,K$}{
@@
  }
- \caption{\label{alg:br} Compute cross-validated deviance residuals based on the Breslow estimator}
+ \caption{\label{alg:brkp} Compute cross-validated deviance residuals based on the Breslow estimator and the Kalbfleisch and Prentice estimator}
 \end{algorithm}
@@
 \end{equation}
-Procedure~\ref{alg:kp} provides an explicit description of this algorithm.
-\begin{algorithm}
-\SetAlgoLined
-\For{each regularization parameter $\lambda$}{
- Fit the model using all observations' time and status \;
- Obtain linear predictor $\hat{\eta}$ from the fitted model\;
- Compute $\hat{H}_0(t) = \sum_{t_{i} \leq t} \left\{1 -  \left(1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})} \right) ^{\exp(-\hat{\eta}_{i})}\right\}$\;
- \For{each fold $k = 1,2,...,K$}{
-  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
-  	\For{each obervation $i$ in the $k$th fold}{
-  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
-  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
-  	}
- }
- Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
- }
- \caption{\label{alg:kp} Compute cross-validated deviance residuals based on the Kalbfleisch and Prentice estimator}
-\end{algorithm}
+Procedure~\ref{alg:brkp} includes an explicit description of this algorithm.
 

### 2021-12-04 main.tex
Edits to deviance residuals (section and supplement)
Edits to paper and citations following Dec 3 meeting

@@
 \par In this paper, we focus upon the LASSO penalty $P_{\lambda}(\beta) = \lambda \sum_{j} |\beta_{j}|$, although the methods we analyze can be used with any penalty as well as with model selection in unpenalized Cox regression.
-LASSO-penalized cox regression is particularly useful for high dimensional data where the number of covariates $p > n$, such as predicting overall survival in cancer patients based on genome-wide expression measurements.
+LASSO-penalized Cox regression is particularly useful for high dimensional data where the number of covariates $p > n$, such as predicting overall survival in cancer patients based on genome-wide expression measurements.
 LASSO estimates yield a sparse model where some coefficients are exactly zero.
@@
 For a given $k \in \left\{1,2,\ldots, K\right\}$, let $T_{k} = D - D_{k}$ denote the training set.
-This manuscript is concerned with the question: how should one use the $kth$ fold $D_{k}$ to measure predictive accuracy based on partial likelihood?
+This manuscript is concerned with the question: how should one use the $k$th fold $D_{k}$ to measure predictive accuracy based on partial likelihood?
 In particular, the partial likelihood \eqref{eq:cox-pl-subj} involves a risk set -- which observations should be used to form that risk set?
@@
 the factor of 2 is arbitrary, but often used so that CVE estimates the deviance of the model.
-We refer this approach as \emph{basic cross-validated partial likelihood}. This is implemented in the $\tt{glmnet}$ package as the \verb|"ungrouped"| option.
+We refer this approach as \emph{basic cross-validated partial likelihood}. This is implemented in the $\tt{glmnet}$ package as the \texttt{grouped=FALSE} option.
 
@@
 For example, it is used by the R packages {\tt CoxRidge} \citep{CoxRidge}, {\tt fastcox} \citep{fastcox}, {\tt SGL} \citep{SGL} , {\tt CoxBoost} \citep{CoxBoost}, {\tt mboost} \citep{mboost}, and {\tt glmnet} \citep{glmnet}.
-In {\tt glmnet}, it is the default choice for penalized Cox regression and is referred to as the \verb|"grouped"| option in the package's syntax.
+In {\tt glmnet}, it is the default choice for penalized Cox regression (\texttt{grouped=TRUE} in the package's syntax).
 
@@
 %%%%%%%%%%%%%%%%%%%%
-\subsection{Cross Validated Linear Predictors}
+\subsection{Cross validated linear predictors}
 \label{Sec:linear-predictor}
@@
   \label{eq:cv-lp}
-  \hat{\eta}^{cv}_{i} = X_{i}\Tr\hat{\beta}^{-k}
+  \hat{\eta}^{\text{cv}}_{i} = X_{i}\Tr\hat{\beta}^{-k}
 \end{equation} 
-After repeating this for all K folds, a complete set of cross-validated linear predictors $\hat{\eta}^{cv} = ( \hat{\eta}^{cv}_{1},  \hat{\eta}^{cv}_{2} , ...
-\hat{\eta}^{cv}_{n})$ for the whole sample is obtained.
+After repeating this for all K folds, a complete set of cross-validated linear predictors $\hat{\eta}^{\text{cv}} = ( \hat{\eta}^{\text{cv}}_{1},  \hat{\eta}^{\text{cv}}_{2} , ...
+\hat{\eta}^{\text{cv}}_{n})$ for the whole sample is obtained.
 A partial likelihood can then be built using this set of linear predictors to evaluate the predictive accuracy of the model:
 \begin{equation*}
-  L(\hat{\eta}^{cv}) = \prod_{i=1}^{n} \left\{ \frac{\exp (\hat{\eta}^{cv}_{i})}{\sum_{k \in R(t_{i})}\exp (\hat{\eta}^{cv}_k)} \right\} ^{\delta_i}.
+  L(\hat{\eta}^{\text{cv}}) = \prod_{i=1}^{n} \left\{ \frac{\exp (\hat{\eta}^{\text{cv}}_{i})}{\sum_{k \in R(t_{i})}\exp (\hat{\eta}^{\text{cv}}_k)} \right\} ^{\delta_i}.
 \end{equation*}
-We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - 2 \log L(\hat{\eta}^{cv}).$$
+We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - 2 \log L(\hat{\eta}^{\text{cv}}).$$
 
@@
 %%%%%%%%%%%%%%%%%%%%
-\subsection{Cross Validated Deviance Residuals}
+\subsection{Cross validated deviance residuals}
 %%%%%%%%%%%%%%%%%%%%
@@
 In order to obtain a cross-validated version of these residuals, we would replace $\beta$ in the above equations with $\hat{\beta}^{-k}$.
-However, it is not obvious how to obtain a cross-validated version of $H_0$, or if it is even necessary to do so.
+However, it is not obvious how to obtain a cross-validated version of $H_0$, or to what extent this is necessary.
 We experimented with three versions of cross-validated deviance residuals, each with a different choice for $H_0$: (1) The Nelson-Aalen estimator, which does not involve the covariates and is therefore the same across cross-validation folds, (2) the Breslow estimator \citep{Breslow1972}, and (3) the Kalbfleisch-Prentice estimator \citep{Kalbfleisch2011}.
@@
 %%%%%%%%%%%%%%%%%%%%
-\section{Simulation Studies}
+\section{Simulation studies}
 %%%%%%%%%%%%%%%%%%%%
@@
 For these simulation studies, the covariate matrix $X$, consisting of $n$ observations and $p$ covariates, was generated from independent $\text{Normal}(0, 1)$ samples.
-Given $X$ and a coefficient vector $\beta$, survival times were generated from exponential distribution with rate parameter $\exp(X\beta)$.
+Given $X_i$ and a coefficient vector $\beta$, survival times were generated from an exponential distribution with rate parameter $\exp(X_i \Tr \beta)$.
 Censoring status indicators were drawn independently from the binomial distribution.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection {Simulations Comparing Predictive Accuracy}
+\subsection{Simulations comparing predictive accuracy}
 \label{Sec:accuracy}
@@
 
-Throughout, we communicate the estimation accuracy in terms of relative MSE, as compared to the oracle model.
-For each generated data set, an oracle model was fit using Cox regression only including the variables with non-zero coefficients.
-Letting $\hat{\beta}^{oracle}$ denote the resulting estimates, the relative MSE, on the log scale, is given by
+Throughout, we communicate the estimation accuracy in terms of relative MSE, as compared to the oracle
+model.
+For each generated data set, an oracle model was fit using Cox regression including only the true features (i.e., those with non-zero coefficients).
+Letting $\hat{\beta}^{\text{oracle}}$ denote the resulting estimates, the relative MSE, on the log scale, is given by
 \begin{equation}
@@
 \end{equation}
-where $\lambda^{cv}$ denotes the value of $\lambda$ selected by cross-validation.
-MSE is estimated by averaging over $N$ replications respectively for $\hat{\beta}(\lambda^{cv})$ and $\hat{\beta}^{oracle}$, with the ratio taken afterwards.
+where $\lambda^{\text{cv}}$ denotes the value of $\lambda$ selected by cross-validation.
+MSE is estimated by averaging over $N$ replications respectively for $\hat{\beta}(\lambda^{\text{cv}})$ and $\hat{\beta}^{\text{oracle}}$, with the ratio taken afterwards.
 The $\log$ is taken to correct for skewness.
@@
 
-\bnote{More simulations were conducted to study scenarios where features have more complicated correlation structures.
-Similar patterns were observed as in Figure~\ref{Fig:mse-brier-c} and Table~\ref{Tab:sim}.
-Results of these simulations are provided in the supplementary material.}
+The patterns seen in Figure~\ref{Fig:mse-brier-c} and Table~\ref{Tab:sim} are similar when features are correlated (supplementary material).
 
-\subsection{\bnote{Power}}
+\subsection{Power to identify important features}
 \label{Sec:power}
 
-While cross-validation is mainly a measure of the model's predictive accuracy, it is not sufficient to only compare those methods based on predictive accuracy.
-Another important criterion is that how many important variables are correctly selected by the model.
-The following simulations are conducted to compare the methods in this aspect.
+A significant scientific goal of survival modeling is often to identify features that are associated with increased or decreased risk.
+By choosing a larger value of the regularization parameter $\lam$ and stopping too early on the coefficient path, the V\&VH and deviance residual methods risk failing to identify important features.
+In this section, we quantify this loss.
 
-In Table~\ref{Tab:power}, we computed the average number of true positives selected by each cross-validation approach.
-True positives are features that are selected in the fitted model and their true $\beta$ coefficients are non-zero in the data-generating model.
+Table~\ref{Tab:power} presents the average number of true positives selected by each cross-validation approach.
+True positives are features with non-zero $\beta$ coefficients in the true data-generating model that were also selected by the model.
 The simulation scenarios vary from low dimension, with 150 observations and 50 features, to high dimension, with 400 observations and 10000 features.
-We generate data with both weak signal strength and strong signal strength.
-Across all simulation scenarios, features are independent from each other.
-The results of the simulation is presented in Table~\ref{Tab:power}.
+We generate data with both weak and strong signal strength \note{Need to be more specific here; what is $s$?
+Include either here or in the caption to Table 2}.
+Here, all features are independent.
 
@@
 
-When the signal strength in the data is weak, it is more difficult to pick up the signal and identify the non-zero features.
-Hence there is a big difference among the numbers of true positives selected by different cross-validation approaches, as is illustrated in Table~\ref{Tab:power}.
-In these scenarios, the basic cross-validation approach and linear predictor approach are more powerful, while the V$\&$VH approach always has the least number of true positive features.
-Especially in the high dimensional cases, the V$\&$VH approach select about three features fewer than the other approaches.
-In practice, missing important findings is a real concern when applying those methods.
+When the signal strength is weaker, identify the true features becomes more challenging.
+Thus, while all cross-validation approaches perform reasonably well with high signal, there is a meaningful difference among them when signal is lower.
+In these scenarios, the basic and linear predictor approach are more powerful, while the V\&VH approach selects the fewest true features.
+This is most apparent in the high dimensional setting, where the V\&VH approach select approximately three features fewer than the best approaches.
+Missing multiple important findings is an important concern when applying this method in practice.
 
@@
 On the other hand, the LOOCV version of the linear predictor approach surpasses basic 10-fold CV, with a smaller $\log(\text{MSE ratio})$ across all signal strength levels.
-At the same time, even the 10-fold version of the linear predictor approach outperforms the leave-one-out version of the V$\&$VH approach and the deviance residual approach.
-In other words, although adopting a leave-one-out approach improves the predictive accuracy for V$\&$VH and deviance residual cross-validation, the magnitude of improvement in predictive performance is smaller than the improvement gained from switching to the linear predictor approach.
+At the same time, even the 10-fold version of the linear predictor approach outperforms the leave-one-out version of the V\&VH approach and the deviance residual approach.
+In other words, although adopting a leave-one-out approach improves the predictive accuracy for V\&VH and deviance residual cross-validation, the magnitude of improvement in predictive performance is smaller than the improvement gained from switching to the linear predictor approach.
 
@@
 In this section, we compare these likelihood-based approaches against an alternative approach of cross validation using concordance, also known as the cross-validated C Index.
-This is a widely used approach for conducting cross-validation with time-to-event outcomes, particularly for machine learning models \citep{Subramanian2011,Simon2011a}.
+This is a widely used approach for conducting cross-validation with time-to-event outcomes, particularly for machine learning models \citep{Subramanian2011,Simon2011a}, and was recently added as an option in the \texttt{glmnet} package (\texttt{type.measure="C"}).
 This approach typically proceeds by constructing cross-validated linear predictors, as in Section~\ref{Sec:linear-predictor}, but then using those predictors to calculate concordance as opposed to partial likelihood.
@@
   & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-6}
-V $\&$ VH      & 0.639 & 0.637 & 0.638 & 0.637 & 0.637\\
+V \& VH      & 0.639 & 0.637 & 0.638 & 0.637 & 0.637\\
 Basic       & 0.647 & 0.639 & 0.638 & 0.638 & 0.639 \\
@@
   & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-6}
-V $\&$ VH  & 0.243 & 0.244 & 0.244 & 0.245 & 0.245 \\
+V \& VH  & 0.243 & 0.244 & 0.244 & 0.245 & 0.245 \\
 Basic  & 0.233 & 0.243 & 0.244 & 0.244 & 0.244 \\

### 2021-12-05 main.tex
comment out supplementary par and add s in Sec 3.2

@@
 The simulation scenarios vary from low dimension, with 150 observations and 50 features, to high dimension, with 400 observations and 10000 features.
-We generate data with both weak and strong signal strength \note{Need to be more specific here; what is $s$?
-Include either here or in the caption to Table 2}.
+We generate data with both weak and strong signal strength, where we set $s$, the value of non-zero $\beta$ coefficients, to be 0.3 and 0.6 respectively.
 Here, all features are independent.
@@
 \setlength{\tabcolsep}{3pt}
-\caption{\label{Tab:power} Number of non-zero features (true positives) selected by each cross-validation method, under scenarios with varying dimension and signal strength.}
+\caption{\label{Tab:power} Number of non-zero features (true positives) selected by each cross-validation method, under scenarios with varying dimension and signal strength.
+For weak and strong signal strength, $s$ is set to be 0.3 and 0.6 respectively.}
 \centering

### 2021-12-05 supplementary.tex
comment out supplementary par and add s in Sec 3.2

@@
 
-Deviance residuals are typically calculated based on a baseline hazard estimate that has been adjusted for the covariates in the model.
+%Deviance residuals are typically calculated based on a baseline hazard estimate that has been adjusted for the covariates in the model.
 In the context of cross-validation, however, this approach is problematic.

### 2022-01-07 main.tex
Submitted to SMMR

@@
 A significant scientific goal of survival modeling is often to identify features that are associated with increased or decreased risk.
-By choosing a larger value of the regularization parameter $\lam$ and stopping too early on the coefficient path, the V\&VH and deviance residual methods risk failing to identify important features.
-In this section, we quantify this loss.
+By choosing a larger value of the regularization parameter $\lam$ and stopping too early on the coefficient path, the V\&VH and deviance residual methods may fail to identify important features.
+In this section, we quantify this loss of power.
 
-Table~\ref{Tab:power} presents the average number of true positives selected by each cross-validation approach.
-True positives are features with non-zero $\beta$ coefficients in the true data-generating model that were also selected by the model.
-The simulation scenarios vary from low dimension, with 150 observations and 50 features, to high dimension, with 400 observations and 10000 features.
-We generate data with both weak and strong signal strength, where we set $s$, the value of non-zero $\beta$ coefficients, to be 0.3 and 0.6 respectively.
+Table~\ref{Tab:power} presents the average number of true positives identified by each cross-validation approach.
+True positives are features with non-zero $\beta$ coefficients in the true data-generating mechanism that were also selected by the model.
+The simulation scenarios vary from low dimensional, with 150 observations and 50 features, to high dimensional, with 400 observations and 10000 features.
+We generate data with both weak (nonzero $\beta$ coefficients equal to 0.3) and strong (nonzero $\beta$ = 0.6) signal strength.
 Here, all features are independent.
@@
 \caption{\label{Tab:power} Number of non-zero features (true positives) selected by each cross-validation method, under scenarios with varying dimension and signal strength.
-For weak and strong signal strength, $s$ is set to be 0.3 and 0.6 respectively.}
+For weak and strong signal strength, nonzero coefficients are set to 0.3 and 0.6 respectively.}
 \centering
@@
 
-When the signal strength is weaker, identify the true features becomes more challenging.
+When the signal strength is weak, identifying the true features becomes more challenging.
 Thus, while all cross-validation approaches perform reasonably well with high signal, there is a meaningful difference among them when signal is lower.
@@
 
+\afterpage{\clearpage}
+
 Several observations can be drawn from Figure~\ref{Fig:cvauc}.

### 2022-01-07 sub/smimr/coxcv.tex
Fixing citation style for SMIMR submission
Submitted to SMMR

@@
 \usepackage{color}
-\graphicspath{{../../manuscript_figure/}}
 \usepackage[doublespacing]{setspace}  % Required for SMIMR review
@@
 \hypersetup{colorlinks=true, citecolor=black, urlcolor=cyan}
+\usepackage{afterpage}
 
-\usepackage{natbib}
+\usepackage[numbers]{natbib}
 \usepackage{bbold}
@@
 We illustrate these advantages using simulated data as well as using them to analyze data from a high-dimensional study of survival in lung cancer patients.
-\input{main}
+\section{Introduction}
 
-\bibliographystyle{SageH}
-\bibliography{articles}
+\par Since its original proposal \citep{Cox1972}, Cox proportional hazards regression has become the most common regression approach for analyzing survival data.
+Cox regression utilizes a partial likelihood construction under an assumption of proportional hazards to estimate the regression coefficients without having to specify the underlying baseline hazard.
+The ability to avoid choosing a specific parametric distribution for the survival time is very attractive, as time-to-event data are often poorly described by fully parametric models.
+
+This semiparametric flexibility, however, comes at a cost.
+The Cox regression model can estimate relative risks, but without estimating the baseline hazard, it does not predict the absolute failure time for any given individual.
+This poses a challenge to assessing the predictive accuracy of a given Cox regression model.
+The challenge is particularly relevant for penalized Cox regression \citep{Tibshirani1997,Fan2002}, as the assessment of predictive accuracy via cross-validation is the standard method for selecting the regularization parameter and deciding upon a model.
+
+\par Standard cross-validation involves dividing the data into $K$ folds, then fitting the model on $K-1$ of those folds (the training set) and assessing prediction accuracy on the remaining fold (the testing set).
+This process is then repeated, with each fold serving as the testing set exactly once.
+For Cox regression, the estimated coefficients allow us to quantify the risk for each subject in the training set relative to other members of the training set, but it is not obvious how to use those estimates to quantify the model's accuracy in the testing set.
+
+\par One basic approach is to simply calculate the partial likelihood based on the observations in the test set as a measure for the model's predictive accuracy.
+In doing so, we calculate the risk for each member of the test set relative to other members of the test set.
+One drawback to this approach is that it becomes unstable when the size of the test set is small.
+In particular, it cannot be applied to leave-one-out cross-validation (LOOCV), as we must have at least two observations in the test set to compare their risk relative to each other.
+
+\par To overcome this drawback, an alternative approach was proposed by Verweij and Van Houwelingen\citep{Verweij1993}.
+Their approach, which we describe in detail in Section~\ref{Sec:cox-cv-existing}, stabilizes the cross-validated log likelihood, enabling its use even when the number of subjects in each fold is small.
+This approach has been widely used and is the default approach to cross-validation in many software programs, such as the R package {\tt glmnet} \citep{glmnet}.
+Although this approach fixes the stability issue, we demonstrate here that in practice, it tends to behave conservatively in terms of model selection.
+
+\par In this paper, we propose two alternative ways to carry out cross validation for Cox regression.
+Instead of applying cross-validation to the partial likelihood directly, we propose applying cross validation to either the linear predictors of the regression model or to the deviance residuals \citep{Therneau1990}.
+Through simulation studies, we compare these proposed methods with existing cross-validation approaches for LASSO penalized Cox regression in both low- and high-dimensional settings.
+The linear predictor approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.
+We conclude by applying both proposed and existing approaches to a high-dimensional study of survival in lung cancer patients.
+
+%%%%%%%%%%%%%
+%%%%%%%%%%%%%
+\section{Methods}
+\label{Sec:methods}
+%%%%%%%%%%%%%
+%%%%%%%%%%%%%
+
+\par In the Cox model, the hazard function for subject $i$ is given by 
+\begin{equation*}
+  h_{i}(t) = h_{0}(t) \exp( X_{i}\Tr\beta),
+\end{equation*} 
+where $h_{0}$ is the baseline hazard and $\exp(X_i\Tr\beta)$ is the risk for a subject with covariates $X_i$ relative to that baseline.
+The estimation of the coefficients is obtained by maximizing a partial likelihood.
+Letting $t_i$ denote the time on study for subject $i$ and $\delta_{i}$ indicate whether or not an event is observed for subject $i$, each subject's individual contribution to that partial likelihood is
+\begin{equation}
+  \label{eq:cox-pl-subj}
+  L_{i}(\beta) = \left \{\frac{\exp ( X_{i}\Tr\beta)}{\sum_{ k \in R(t_{i})}\exp ( X_{k}\Tr \beta)}\right \}^{\delta_{i}},
+\end{equation}
+where $R(t_{i})$ denotes the set of subjects at risk at time $t_{i}$.  The partial likelihood for the entire sample of $n$ subjects is then given by:
+\begin{equation}
+  \label{eq:cox-pl-sum}
+  L(\beta) =\prod_{i = 1}^{n} L_{i}(\beta).
+\end{equation}
+Cox regression can be extended by introducing a penalty into the partial likelihood.  In penalized Cox regression, $\beta$ coefficient estimates are obtained by minimizing the objective function
+\begin{equation}
+  \label{eq:obj}
+  Q(\beta) = - \frac{1}{n} \log L(\beta) + P_{\lambda}(\beta),
+\end{equation}
+where $P_{\lambda}(\beta)$ is a penalty function that depends on a regularization parameter $\lam$.
+Throughout, we will assume without loss of generality that the covariates have been standardized such that $\sum_i x_{ij} = 0$ and $\tfrac{1}{n}\sum_i x_{ij}^2=1$.
+This is standard practice in penalized regression to ensure that the penalization has the same effect for all covariates and is invariant to changes of location and scale.
+The estimates are then converted back to the original scale after the model has been fit.
+Centering the predictors is also standard in Cox regression to ensure that the baseline hazard is invariant to location shifts of the covariates.
+
+\par In this paper, we focus upon the LASSO penalty $P_{\lambda}(\beta) = \lambda \sum_{j} |\beta_{j}|$, although the methods we analyze can be used with any penalty as well as with model selection in unpenalized Cox regression.
+LASSO-penalized Cox regression is particularly useful for high dimensional data where the number of covariates $p > n$, such as predicting overall survival in cancer patients based on genome-wide expression measurements.
+LASSO estimates yield a sparse model where some coefficients are exactly zero.
+This sparsity pattern changes as we vary $\lam$: at large values of $\lambda$, most or all of the coefficients are 0, but as $\lam$ decreases, more covariates are selected.
+Selecting $\lambda$ is critical to LASSO estimation in the sense that the model's accuracy suffers if $\lam$ is either too large or too small.
+
+\par Cross-validation is the most common method for selecting $\lambda$ in penalized regression models.
+Suppose a data set $D$ of n observations is partitioned into $K$ folds: $D_{1}, D_{2}, \ldots, D_{K}$.
+For a given $k \in \left\{1,2,\ldots, K\right\}$, let $T_{k} = D - D_{k}$ denote the training set.
+This manuscript is concerned with the question: how should one use the $k$th fold $D_{k}$ to measure predictive accuracy based on partial likelihood?
+In particular, the partial likelihood \eqref{eq:cox-pl-subj} involves a risk set -- which observations should be used to form that risk set?
+In the following sections, let $L^{-k}$ denote the partial likelihood built using data in $T_k$, let $L^{k}$ denote the partial likelihood built using data in $D_k$, and $L$ denote the partial likelihood built using the entire data set $D$.
+The log partial likelihood is denoted by $\ell$, with $\ell^k$ and $\ell^{-k}$ defined similarly.
+Finally, let $\hat{\beta}^{-k}$ denote the penalized estimates that are obtained by using $L^{-k}(\beta)$ in \eqref{eq:obj}.
+Figure~\ref{Fig:methods} diagrams the relationships between the four cross-validation methods we will describe in this section.
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Cross validated partial likelihood} 
+\label{Sec:cox-cv-existing}
+%%%%%%%%%%%%%%%%%%%%
+
+\par A direct approach to measure a model's predictive accuracy in $D_k$ is to evaluate the log partial likelihood at $\hat{\beta}^{-k}$ using the data in $D_k$.
+Cross-validation error (CVE) is then measured by the total deviance across all $K$ folds, which (up to a constant) is given by:
+\begin{equation}
+  \label{eq:basic}
+  \CVE = -2 \sum_{k=1}^{K} \ell^{k}(\hat{\beta}^{-k});
+\end{equation}
+the factor of 2 is arbitrary, but often used so that CVE estimates the deviance of the model.
+We refer this approach as \emph{basic cross-validated partial likelihood}. This is implemented in the $\tt{glmnet}$ package as the \texttt{grouped=FALSE} option.
+
+\par The basic approach is by far the most common way of conducting cross-validation for models such as linear regression and logistic regression.
+However, the basic approach can be problematic for Cox regression in the sense that that there may not be enough observations in $D_k$ to build up the risk set for partial likelihood.
+For example, the basic approach cannot work with leave-one-out cross-validation, since for all folds, $\ell_k$ would either be zero or undefined.
+
+To address this issue, Verweij and Van Houwelingen\citep{Verweij1993} proposed an alternative method for cross-validation in Cox models:
+\begin{equation}
+  \label{eq:VVH}
+  \CVE = -2\sum_{k = 1}^K \left\{ \ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k}) \right\}. 
+\end{equation}
+Here, $\ell(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ using the entire data set $D$ and $\ell^{-k}(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on
+the training set $T_k$.
+By avoiding the construction of a partial likelihood on the test set, this approach ensures that the risk set is always sufficiently large and the partial likelihood well-defined.
+We refer to this cross-validation method as the Verweij and Van Houwelingen (V\&VH) approach.
+For many models, such as logistic regression or linear regression, the quantities $\ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k})$ and  $\ell^{k}(\hat{\beta}^{-k})$ are equivalent to each other -- in other words, the basic and V\&VH approaches agree.
+However, Verweij and Van Houwelingen's approach is more stable for Cox regression as there is always a large number of observations in the risk sets its calculations are based on.
+Since its proposal, the V\&VH approach has been widely implemented as a tool for cross-validation is Cox models.
+For example, it is used by the R packages {\tt CoxRidge} \citep{CoxRidge}, {\tt fastcox} \citep{fastcox}, {\tt SGL} \citep{SGL} , {\tt CoxBoost} \citep{CoxBoost}, {\tt mboost} \citep{mboost}, and {\tt glmnet} \citep{glmnet}.
+In {\tt glmnet}, it is the default choice for penalized Cox regression (\texttt{grouped=TRUE} in the package's syntax).
+
+\begin{figure}
+  \centering
+  \includegraphics[height= 12cm ]{figure_1_new.png}
+  \caption{\label{Fig:methods} Illustration of $K$-fold cross validation methods for Cox regression.}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Cross validated linear predictors}
+\label{Sec:linear-predictor}
+%%%%%%%%%%%%%%%%%%%%
+
+\par Both of the methods in Section~\ref{Sec:cox-cv-existing} consist of adding together log partial likelihood measures from each fold.
+An alternative method is to obtain linear predictors for each fold, then combine these linear predictors to calculate a partial likelihood.
+We refer to this method as the \emph{cross-validated linear predictors} approach.
+To be more specific, we would first obtain $\hat{\beta}^{-k}$ from training set $T_{k}$; then the cross-validated linear predictors can be calculated for each observation $i$ from the test set $D_k$:
+\begin{equation}
+  \label{eq:cv-lp}
+  \hat{\eta}^{\text{cv}}_{i} = X_{i}\Tr\hat{\beta}^{-k}
+\end{equation} 
+After repeating this for all K folds, a complete set of cross-validated linear predictors $\hat{\eta}^{\text{cv}} = ( \hat{\eta}^{\text{cv}}_{1},  \hat{\eta}^{\text{cv}}_{2} , ...
+\hat{\eta}^{\text{cv}}_{n})$ for the whole sample is obtained.
+A partial likelihood can then be built using this set of linear predictors to evaluate the predictive accuracy of the model:
+\begin{equation*}
+  L(\hat{\eta}^{\text{cv}}) = \prod_{i=1}^{n} \left\{ \frac{\exp (\hat{\eta}^{\text{cv}}_{i})}{\sum_{k \in R(t_{i})}\exp (\hat{\eta}^{\text{cv}}_k)} \right\} ^{\delta_i}.
+\end{equation*}
+We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - 2 \log L(\hat{\eta}^{\text{cv}}).$$
+
+\par This idea of obtaining linear predictors using test sets, then constructing the partial likelihood after pooling all linear predictors together is implemented in R package \texttt{ncvreg} \citep{ncvreg}.
+This idea has also appeared in the cross-validation literature for quantities that cannot be evaluated on only a subset of the data, such as AUC in logistic and Cox regression models \citep{Parker2007,Simon2011a,Subramanian2011}.
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Cross validated deviance residuals}
+%%%%%%%%%%%%%%%%%%%%
+
+The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model, which prevents us from calculating prediction errors (residuals).
+However, if one takes additional steps to estimate the baseline hazard in addition to estimating the coefficients, then such residuals can be calculated.
+The deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}, is a natural candidate for quantifying prediction error to use in quantifying the cross-validation error.
+Originally proposed for model diagnostics, deviance residuals have also been occasionally used as a predictive accuracy measure \citep{Therneau2018}.
+In the Cox model, the Martingale residuals $\{m_i\}_{i=1}^n$ and deviance residuals $\{d_i\}_{i=1}^n$ are given by
+\begin{align}
+  \label{eq:res-m}
+  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(X_i\Tr\beta), \\
+  \label{eq:res-d}
+  d_{i} &= \text{sign}(m_{i})\sqrt{-2(m_i + \delta_{i}\log(\delta_{i} - m_i))}.  
+\end{align}
+where $H_0$ is the cumulative baseline hazard.
+
+In order to obtain a cross-validated version of these residuals, we would replace $\beta$ in the above equations with $\hat{\beta}^{-k}$.
+However, it is not obvious how to obtain a cross-validated version of $H_0$, or to what extent this is necessary.
+We experimented with three versions of cross-validated deviance residuals, each with a different choice for $H_0$: (1) The Nelson-Aalen estimator, which does not involve the covariates and is therefore the same across cross-validation folds, (2) the Breslow estimator \citep{Breslow1972}, and (3) the Kalbfleisch-Prentice estimator \citep{Kalbfleisch2011}.
+Details for the three approaches are given in the supplementary materials, along with the results of a simulation study comparing them.
+Because the Nelson-Aalen baseline hazard performed best in all situations, we use this estimate to construct the cross-validated deviance residuals throughout the rest of the paper.
+
+The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross validated error, analogous to using the residual sum of squares in linear regression.
+We refer to this method as the \emph{cross-validated deviance residuals} approach.
+
+%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%
+\section{Simulation studies}
+%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%
+
+Simulation studies were conducted to compare how the methods described in Section~\ref{Sec:methods} behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for LASSO penalized regression.
+For these simulation studies, the covariate matrix $X$, consisting of $n$ observations and $p$ covariates, was generated from independent $\text{Normal}(0, 1)$ samples.
+Given $X_i$ and a coefficient vector $\beta$, survival times were generated from an exponential distribution with rate parameter $\exp(X_i \Tr \beta)$.
+Censoring status indicators were drawn independently from the binomial distribution.
+All simulations were implemented in R \citep{R}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Simulations comparing predictive accuracy}
+\label{Sec:accuracy}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this section, we consider the accuracy of the cross-validation approaches introduced in Section~\ref{Sec:methods}.
+We varied the sample size, dimension, censoring percentage, and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation.
+Accuracy was assesed via several criteria.
+
+Estimation accuracy was measured using the mean squared error of the coefficients ($\MSE$).
+Suppose $\hat{\beta}(\lambda)$ is the vector of coefficients estimated by a Cox model with LASSO penalty at $\lambda$, and $\beta^*$ the value of $\beta$ used the generate the data.
+Mean squared error is defined to be
+\begin{equation}
+\MSE = E\| \hat{\beta}(\lambda) - \beta^* \| ^2.
+\end{equation}
+
+Throughout, we communicate the estimation accuracy in terms of relative MSE, as compared to the oracle
+model.
+For each generated data set, an oracle model was fit using Cox regression including only the true features (i.e., those with non-zero coefficients).
+Letting $\hat{\beta}^{\text{oracle}}$ denote the resulting estimates, the relative MSE, on the log scale, is given by
+\begin{equation}
+\log\left\{ \frac{\MSE(\hat{\beta}(\lambda^{\text{cv}}))}{\MSE(\hat{\beta}^{\text{oracle}})} \right\},
+\end{equation}
+where $\lambda^{\text{cv}}$ denotes the value of $\lambda$ selected by cross-validation.
+MSE is estimated by averaging over $N$ replications respectively for $\hat{\beta}(\lambda^{\text{cv}})$ and $\hat{\beta}^{\text{oracle}}$, with the ratio taken afterwards.
+The $\log$ is taken to correct for skewness.
+
+\par We also measure the predictive accuracy via the mean squared prediction error, or Brier score \citep{VanHouwelingen2011}.
+Let $\hat{S}(t_0|\lambda,x)$ denote the predicted probability for an individual with covariates $x$ to survive beyond time $t_0$, and let $y = \mathbb{1}\left\{ t > t_{0}\right\}$ indicate whether the individual actually survived beyond time $t_0$.
+Then the Brier score is defined to be
+\begin{equation}
+\text{Brier}(y, \hat{S}(t_0|\lambda,x)) = (y - \hat{S}(t_0|\lambda,x))^2.
+\end{equation}
+In our simulation study, for each data set, we generated another independent data set with 1000 uncensored observations.
+We computed Brier scores for all of the cross-validation methods for all individuals in this test set, setting $t_0$ to the median survival time.
+
+\par The third measure we used is the Kullback-Leibler score, which measures the log-likelihood of the prediction model evaluated at the observations \citep{VanHouwelingen2011}.
+With the same notation defined in the previous paragraph, the Kullback-Leibler score is defined to be
+\begin{equation}
+	\text{KL}(y, \hat{S}(t_0|\lambda,x)) = -\left\{ y\log(\hat{S}(t_0|\lambda,x) + (1 - y)\log(1 - \hat{S}(t_0|\lambda,x)) \right\}.
+\end{equation}
+As for the Brier score, KL scores were computed for all CV approaches at the median survival time.
+Results were then averaged over the N replications.
+For both Brier and KL scores, a smaller score represents smaller prediction error, thus indicating better prediction accuracy.
+
+\par The final metric we included in our study was Harrell's C Index \citep{HarrellJr1984}.
+The C Index is a rank-based statistic that measures the concordance between the linear predictor of the selected model and the observed outcome.
+Suppose we arbitrarily select two individuals; if the individual with the higher predicted risk also died first, then this would be considered as a concordant pair.
+The C Index considers all possible such pairs and estimates of proportion of those pairs that are concordant.
+A C Index of 0.5 means the prediction is no better than flipping a coin and C Index of 1 means perfect concordance.
+We computed the C Index using the independent test set of 1000 individuals described above.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{figure_2_new.png}
+  \caption{\label{Fig:mse-brier-c} Simulation comparing 4 CV methods.
+The horizontal axis in all three plots is $s$, which represents the signal strength in the data.
+The value of $s$ is varied from 0.4 to 0.9.
+$\log(\text{MSE ratio})$, relative to the oracle modle is plotted in the left panel.
+Out-of-sample Brier scores are plotted in the middle.
+Out-of-sample C index is plotted in the right panel.
+For each simulated data set, n = 120, p = 1000.
+Expected censoring percentage is 10$\%$.}
+\end{figure}
+
+For our first simulation study, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates.
+The covariate vector $\beta$ is assumed to be sparse, with $\beta_j = 0$ for $j \in \{11, 12, \ldots, 1000\}$ and $\beta_j = s$ for $j \in \{1, 2, \ldots, 10 \}$, where $s$ is a prespecified constant that specifies the signal strength in the simulated data set.
+Here, we vary $s$ from 0.4 to 0.9.
+The survival outcomes have 10$\%$ expected censoring.
+For each of the four methods described in Section~\ref{Sec:methods}, 10-fold cross-validation was implemented.
+$N = 200$ replications were used for each scenario.
+Figure~\ref{Fig:mse-brier-c} illustrates the results of this simulation study.
+For all four metrics, the overall pattern is the same (KL scores not shown in Figure~\ref{Fig:mse-brier-c} due to space constraints, but they were very similar to Brier scores).
+Although the four cross-validation methods performed relatively similarly, the basic and linear predictor approaches had the best performance across all metrics: lower MSE, better Brier scores, and higher C Index.
+
+\begin{table}[!htb]
+
+\setlength{\tabcolsep}{3pt}
+
+\caption{\label{Tab:sim} Comparing the predictive accuracy of four cross-validation methods under scenarios with varying dimension and signal strength}
+\centering
+\begin{tabular}[t]{lccccc}
+\toprule
+ Method & $\lambda$ & log(MSE Ratio) &Brier Score & Kullback Leibler & C Index\\
+\midrule
+&\multicolumn{5}{c}{\textbf{Scenario 1}: n = 150, p = 50, 5 non-zero features, weak signal}\\
+\cline{2-6}
+V\&VH  & 0.098 (0.030) & 1.561 (0.851) & 0.245 (0.014) & 0.685 (0.030) & 0.610 (0.035) \\
+Basic  & 0.088 (0.025) & 1.531 (0.838) & 0.244 (0.013) & 0.683 (0.029) & 0.614 (0.029) \\
+LinearPred  & 0.092 (0.029) & 1.543 (0.846) & 0.245 (0.014) & 0.684 (0.030) & 0.612 (0.034) \\
+DevResid  & 0.098 (0.029) & 1.547 (0.848) & 0.245 (0.014) & 0.684 (0.030) & 0.611 (0.035) \\
+\addlinespace
+&\multicolumn{5}{c}{\textbf{Scenario 2}: n = 150, p = 50, 5 non-zero features, strong signal}\\
+\cline{2-6}
+V\&VH  & 0.071 (0.010) & 1.586 (0.989) & 0.176 (0.011) & 0.526 (0.027) & 0.750 (0.012) \\
+Basic  & 0.066 (0.011) & 1.544 (0.971) & 0.176 (0.011) & 0.526 (0.028) & 0.749 (0.013) \\
+LinearPred  & 0.067 (0.010) & 1.540 (0.981) & 0.176 (0.011) & 0.525 (0.028) & 0.750 (0.012) \\
+DevResid  & 0.094 (0.011) & 1.995 (0.841) & 0.181 (0.011) & 0.539 (0.026) & 0.751 (0.013) \\
+\addlinespace
+&\multicolumn{5}{c}{\textbf{Scenario 3}: n = 400, p = 10000, 20 non-zero features, weak signal}\\
+\cline{2-6}
+V\&VH  & 0.126 (0.019) & 2.958 (0.416) & 0.251 (0.012) & 0.696 (0.026) & 0.589 (0.044) \\
+Basic  & 0.103 (0.022) & 2.880 (0.437) & 0.244 (0.015) & 0.680 (0.031) & 0.603 (0.039) \\
+LinearPred  & 0.105 (0.023) & 2.886 (0.437) & 0.244 (0.015) & 0.681 (0.031) & 0.602 (0.039) \\
+DevResid  & 0.107 (0.022) & 2.895 (0.433) & 0.245 (0.014) & 0.683 (0.03) & 0.602 (0.039) \\
+\addlinespace
+&\multicolumn{5}{c}{\textbf{Scenario 4}: n = 400, p = 10000, 20 non-zero features, weak signal}\\
+\cline{2-6}
+V\&VH  & 0.084 (0.012) & 3.799 (0.548) & 0.176 (0.03) & 0.528 (0.069) & 0.763 (0.048) \\
+Basic  & 0.059 (0.007) & 3.455 (0.578) & 0.154 (0.019) & 0.468 (0.049) & 0.780 (0.031) \\
+LinearPred  & 0.060 (0.007) & 3.479 (0.574) & 0.155 (0.019) & 0.470 (0.049) & 0.780 (0.030) \\
+DevResid  & 0.076 (0.018) & 3.708 (0.518) & 0.166 (0.024) & 0.502 (0.057) & 0.772 (0.049) \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Table~\ref{Tab:sim} displays the results of additional simulations carried out as we varied the dimension of the data set along with the signal strength.
+Specifically, we examined the following settings: low dimension (p = 50) with weak signal, low dimension with strong signal, high dimension (p = 10000) with weak signal, and high dimension with strong signal.
+In all settings, the censoring rate was 30\%.
+For weak cases, signal strength $s$ was set to be 0.3; for strong cases, $s$ was set to be 0.6.
+
+As in Figure~\ref{Fig:mse-brier-c}, the basic and linear predictor approaches outperform the deviance residual and V\&VH approaches.
+The improvement is particularly noticeable in the most difficult setting: high-dimensional data with weak signal.
+Table~\ref{Tab:sim} also provides some insight into why the deviance residual and V\&VH approaches perform poorly relative to basic CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and thus, selecting fewer variables.
+
+The patterns seen in Figure~\ref{Fig:mse-brier-c} and Table~\ref{Tab:sim} are similar when features are correlated (supplementary material).
+
+\subsection{Power to identify important features}
+\label{Sec:power}
+
+A significant scientific goal of survival modeling is often to identify features that are associated with increased or decreased risk.
+By choosing a larger value of the regularization parameter $\lam$ and stopping too early on the coefficient path, the V\&VH and deviance residual methods may fail to identify important features.
+In this section, we quantify this loss of power.
+
+Table~\ref{Tab:power} presents the average number of true positives identified by each cross-validation approach.
+True positives are features with non-zero $\beta$ coefficients in the true data-generating mechanism that were also selected by the model.
+The simulation scenarios vary from low dimensional, with 150 observations and 50 features, to high dimensional, with 400 observations and 10000 features.
+We generate data with both weak (nonzero $\beta$ coefficients equal to 0.3) and strong (nonzero $\beta$ = 0.6) signal strength.
+Here, all features are independent.
+
+\begin{table}[!htb]
+\setlength{\tabcolsep}{3pt}
+\caption{\label{Tab:power} Number of non-zero features (true positives) selected by each cross-validation method, under scenarios with varying dimension and signal strength.
+For weak and strong signal strength, nonzero coefficients are set to 0.3 and 0.6 respectively.}
+\centering
+\begin{tabular}[t]{l cc c cc c cc}
+\toprule
+&\multicolumn{2}{c}{n = 150, p = 50} & & \multicolumn{2}{c}{n = 300, p = 1000} & & \multicolumn{2}{c}{n = 400, p = 10000} \\
+&\multicolumn{2}{c}{$\#$ non-zero features: 5} & & \multicolumn{2}{c}{$\#$ non-zero features: 20} & & \multicolumn{2}{c}{$\#$ non-zero features: 20} \\
+\cline{2-3} \cline{5-6} \cline{8-9}
+& weak signal & strong signal & & weak signal & strong signal && weak signal & strong signal\\
+V\&VH        & 3.87 & 5& & 10.96 & 19.61 & & 6.67 & 18.38\\
+Basic          & 4.06 & 5& & 13.68 & 19.96 & & 9.58 & 19.87\\
+LinearPred  & 3.99 & 5& & 13.46 & 19.95 & & 9.40 & 19.85\\
+DevResid    & 3.90 & 5& & 12.96 & 19.38 & & 8.98 & 19.05\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+When the signal strength is weak, identifying the true features becomes more challenging.
+Thus, while all cross-validation approaches perform reasonably well with high signal, there is a meaningful difference among them when signal is lower.
+In these scenarios, the basic and linear predictor approach are more powerful, while the V\&VH approach selects the fewest true features.
+This is most apparent in the high dimensional setting, where the V\&VH approach select approximately three features fewer than the best approaches.
+Missing multiple important findings is an important concern when applying this method in practice.
+
+\subsection {Stability}
+\label{Sec:stability}
+
+Given that basic cross-validation is both straightforward and demonstrates strong predictive accuracy, as seen in Section~\ref{Sec:accuracy}, one might wonder why it is not more widely used for Cox regression.
+The reason is that it suffers from several drawbacks with respect to numerical stability.
+
+First, the number of uncensored observations in the test set may be insufficient to construct a well-defined partial likelihood for basic cross-validation, leaving the overall estimate of cross-validation error undefined as well.
+When computing the partial likelihood, only individuals whose events are observed will contribute to the likelihood.
+If there are no events observed in a particular fold, then the partial likelihood for that fold is undefined.
+
+To illustrate how common this problem is, we generated 1000 independent data sets of $n = 100$ observations with a fixed censoring rate.
+For each data set, we conducted 10-fold cross-validation and recorded whether or not there was at least one event in each fold (i.e., whether the partial likelihood was well-defined for all folds).
+We repeated the same mechanism for scenarios with different censoring rates, ranging from $40\%$ to $80\%$.
+As illustrated in the left panel in Figure~\ref{Fig:stability}, at $n=100$ the basic approach can be problematic even at moderate censoring rates.
+For example, basic CV cannot be used with over 20\% of data sets when $70\%$ of observations are censored.
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figure_3.png}
+  \caption{\label{Fig:stability} Simulations illustrating the stability issues with the basic approach.
+For the figure on the left, $n = 100$, fold assignments are unbalanced, and the vertical axis is the proportion of undefined partial likelihoods.
+For the figure on the right, $n = 120$, fold assignments are balanced so that the number of events in each fold is the same, and the vertical axis is the standard deviation of the squared error loss.
+Basic cross-validation is much more variable than other approaches when the number of events in each fold is small.}
+\end{figure}
+
+One approach for alleviating this issue is to balance the censoring status across all folds when making fold assignments.
+That is, to ensure the number of events observed is the same (or within 1) across all folds.
+Unless the number of the folds exceeds the number of observed events, such as in leave-one-out cross-validation, this approach will guarantee the partial likelihood in each fold is well-defined for the basic approach, thus the overall cross-validated error is well-defined.
+
+In this case, however, even though the partial likelihood is well-defined, basic cross-validation suffers from extreme variability.
+To illustrate, we generated data sets with $n = 120$ observations and a fixed censoring rate of $50\%$, so that there are 60 observed events.
+For each data set, we carried out cross-validation using balanced fold assignments with the number of folds set to be 10, 12, 15, 20, 30 and 60.
+When the number of folds was set to be 10, there were exactly 6 events in each fold; when the number of folds was set to be 60, there was exactly one event in each fold, and so on.
+After each fold assignment, we computed the cross-validated error using all four approaches described in Section~\ref{Sec:accuracy}.
+We recorded the model that minimizes the cross-validation error for each approach and computed its squared error loss.
+We replicated the whole process for $N = 200$ time.
+In the end, we computed the standard deviation of the squared error loss for each approach.
+
+The results of this simulation are illustrated in the right panel of Figure~\ref{Fig:stability}.
+While the other three approaches are not affected by the number of folds and number of events in each fold, the variability of basic cross-validation increases dramatically as the number of events per fold decreases.
+When there is a sufficient number of events in each fold, the basic approach is similar to the other three approaches.
+However, when there are only one or two events in each fold, the basic approach can be up to 10 times more variable than other methods, with a corresponding increase in the mean squared error.
+In these situations, therefore, even though the basic approach is defined and CVE can be calculated, its performance is very poor.
+
+To shed some light on why the variability of basic cross-validation explodes in the 60-fold scenario, note that in this case, there is exactly one event and one censored observation in each fold.
+In many folds, the uncensored event happens after the other observation has been censored.
+In this situation, the risk set at the time immediately prior to the event contains only a single observation.
+Its contribution to the log partial likelihood is therefore 0, which is well-defined but not informative.
+This situation can happen regardless of the fold size and censoring rate, but it happens far more often as the number of events per fold decreases.
+
+Based on the simulation results illustrated in this section, we suggest that one should proceed with caution if they want to use the basic approach: (i) it is best to balance the censoring status when creating fold assignments for the basic approach, (ii) avoid using the basic approach in conjunction with a high number of folds, and (iii) avoid using the basic approach when the number of observed events is small.
+Alternatively, use the linear predictor approach, which has similar performance without the stability issues.
+
+\subsection {Leave-One-Out Cross Validation}
+\label{Sec:loocv}
+
+As discussed earlier, basic cross-validation cannot be used to carry out leave-one-out cross validation (LOOCV) for Cox regression models.
+The other three approaches (V\&VH, linear predictor, and deviance residual), however, are not subject to the stability issues described in Section~\ref{Sec:stability}, and can be used to implement LOOCV.
+Here, we compare these three approaches for carrying out LOOCV, and also compare LOOCV to 10-fold CV.
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figure_4.png}
+  \caption{\label{Fig:loocv}Simulation comparing LOO and 10-fold CV across three different approaches.
+With $N = 200$ replications, the data sets generated have dimension $n = 120$ and $p = 1000$.
+The vertical axis is the $\log$(MSE ratio).
+In each panel, the LOO version and 10-fold version of the same approach are plotted.
+In all three panels, 10-fold basic cross-validation is included as a baseline for comparison (blue line).}
+\end{figure}
+
+For this simulation experiment, we generated datasets with $n = 120$ and $p = 1000$, using the same data generating procedure as described in Section~\ref{Sec:accuracy}.
+For each of the V\&VH, linear predictor, and deviance residual approaches, we conducted both 10-fold CV and LOOCV.
+We also conducted basic 10-fold cross-validation as a baseline for comparison.
+We recorded the model that minimizes the cross-validated error  for each approach.
+After $N = 200$ replications, $\log$ of the MSE ratio was computed using the same definition described in Section~\ref{Sec:accuracy}.
+
+Results of the simulation are shown in Figure~\ref{Fig:loocv}.
+The panels illustrate the results for each CV approach, one with the 10-fold version and one with the leave-one-out version.
+The 10-fold basic approach was plotted in all three panels as a baseline for comparison.
+For all three approaches, the Leave-One-Out version selected models with a smaller $\log(\text{MSE ratio})$ than the 10-fold version.
+For the V\&VH and deviance residual approaches, however, LOOCV was still inferior to basic 10-fold CV.
+As described in Section~\ref{Sec:accuracy}, these approaches tend to be conservative, selecting a larger-than-optimal value of the regularization parameter; LOOCV helps with this problem, but does not alleviate it entirely.
+
+On the other hand, the LOOCV version of the linear predictor approach surpasses basic 10-fold CV, with a smaller $\log(\text{MSE ratio})$ across all signal strength levels.
+At the same time, even the 10-fold version of the linear predictor approach outperforms the leave-one-out version of the V\&VH approach and the deviance residual approach.
+In other words, although adopting a leave-one-out approach improves the predictive accuracy for V\&VH and deviance residual cross-validation, the magnitude of improvement in predictive performance is smaller than the improvement gained from switching to the linear predictor approach.
+
+\subsection {Cross-Validated C Index}
+\label{Sec:CIndex}
+
+\par All of the cross-validation approaches described in Section~\ref{Sec:methods} use the Cox partial likelihood, in one way or another, as the criterion upon which model selection is based.
+In this section, we compare these likelihood-based approaches against an alternative approach of cross validation using concordance, also known as the cross-validated C Index.
+This is a widely used approach for conducting cross-validation with time-to-event outcomes, particularly for machine learning models \citep{Subramanian2011,Simon2011a}, and was recently added as an option in the \texttt{glmnet} package (\texttt{type.measure="C"}).
+This approach typically proceeds by constructing cross-validated linear predictors, as in Section~\ref{Sec:linear-predictor}, but then using those predictors to calculate concordance as opposed to partial likelihood.
+
+\par We conducted simulations in both low dimensional ($p = 50$, $n = 150$) and high dimensional ($p = 1000$, $n = 120$) settings.
+For each scenario, we generated data from a spectrum of signal strength as in Section~\ref{Sec:accuracy}.
+Results of the simulations are illustrated in Figure~\ref{Fig:cvauc}.
+We used the linear predictor approach (CV-LP) here as the representative approach for likelihood-based cross-validation.
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[height= 14cm ]{figure_auc.png}
+  \caption{\label{Fig:cvauc}Comparing Partial Likelihood and C Index both built from Cross-validated Linear Predictors.
+The horizontal axis represents the strength of signal $s$ in the data.
+The left figures are under the low dimension scenarios where p = 50 and n = 150.
+The right figures are under the high dimension scenarios where p = 1000 and n = 120.
+The expected percentages of censoring for both scenarios are 30 $\%$.
+The top row are box plots of the selected tuning parameter $\lambda$.
+The mid row plots the $\log(\text{relative Squared Error Loss})$, with respect to the oracle model.
+The bottom row plots the Brier Scores.
+}
+\end{figure}
+
+\afterpage{\clearpage}
+
+Several observations can be drawn from Figure~\ref{Fig:cvauc}.
+First, likelihood-based CV tends to select more accurate models, as judged by both mean squared error and Brier score.
+Second, likelihood-based CV is also more stable than the AUC-based CV, in the sense that it tends to select a narrower range of models, with more consistent performance, than CV-AUC.
+Lastly, CV-AUC is consistently more liberal than likelihood-based CV, selecting models with smaller values of the regularization parameter $\lam$, and therefore preferring models with larger coefficient values and more nonzero coefficients.
+
+\section{Application to Real Data}
+
+In this section, we demonstrate the usage and performance of the aforementioned cross-validation methods when fitting penalized regression models on data from a study of gene expression and survival in patients with lung cancer.
+Shedden et al.\citep{shedden2008gene} conducted a large, retrospective study to validate prognostic gene expression signatures as predictors of overall survival in lung cancer patients.
+They gathered expression profiles of 22,283 genes for 442 non-small-cell lung cancer (NSCLC) patients along with some additional clinical and demographic data.
+About $50\%$ of the patients died during the course of the study (236 events), with the rest censored.
+
+For this analysis, we fit elastic net models to the data in addition to lasso models, as these models often improve upon the lasso when many correlated genes are involved \citep{Zou2005}.
+The clinical variables stage, age and gender were included in the model without penalization, as it is reasonable to assume that all three have an effect on survival.
+In addition to the regularization parameter $\lam$, the elastic net model has an extra tuning parameter $\alpha$, where $\alpha=1$ corresponds to lasso and $\alpha=0$ corresponds to ridge regression.
+For $\alpha \in \{0.3, 0.5, 0.7, 0.9, 1\}$, we selected a regularization parameter $\lam$ that has the lowest cross-validation error for each of the four cross-validation methods described in Section~\ref{Sec:methods}.
+Although each method can be used to select a pair of $\alpha$ and $\lam$ values, the methods are on different scales and cannot be directly compared with each other.
+For this reason, Table~\ref{tb:shedden} reports the cross-validated C-Index and Brier Scores for each method's choice of $\lam$ at each value of $\alpha$.
+As shown in the table, the elastic net model with $\lam$ and $\alpha=0.3$ selected by the cross-validated linear predictor approach has the best performance in terms of C Index and Brier Score.
+
+\begin{table}[ht]
+\centering
+\caption{\label{tb:shedden} The C Index and Brier Score of all elasticnet models selected by cross-validation methods }
+\begin{tabular}{lccccc}
+\hline
+  & \multicolumn{5}{c}{Cross-validated C Index} \\ 
+  & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-6}
+V \& VH      & 0.639 & 0.637 & 0.638 & 0.637 & 0.637\\
+Basic       & 0.647 & 0.639 & 0.638 & 0.638 & 0.639 \\
+Linear Pred    & \textbf{0.650} & 0.647 & 0.638 & 0.638 & 0.639  \\
+Deviance Resid & 0.645 & 0.637 & 0.638 & 0.637 & 0.637 \\ 
+\addlinespace
+  & \multicolumn{5}{c}{Brier Score} \\ 
+  & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-6}
+V \& VH  & 0.243 & 0.244 & 0.244 & 0.245 & 0.245 \\
+Basic  & 0.233 & 0.243 & 0.244 & 0.244 & 0.244 \\
+Linear Pred & \textbf{0.229} & 0.232 & 0.244 & 0.244 & 0.244 \\
+Deviance Resid & 0.235 & 0.244 & 0.244 & 0.245 & 0.245 \\ 
+\hline
+\end{tabular}
+\end{table}
+
+For each fixed $\alpha$, we compared the cross-validation errors (CVE) from the four cross-validation methods along the solution path.
+At $\alpha = 0.3$, CVEs are rescaled and plotted for all four methods along the same grid of $\lambda$ values in Figure~\ref{fig:CVE}.
+The selected $\lambda$ values, where a given CVE curve reaches its lowest point, are indicated with dotted lines.
+The CVE curves for the basic and linear predictor approaches have greater curvature near their minima, and thus offer more clear support for the selected value of $\lam$.
+In comparison, the deviance residual and V\&VH approaches are flatter near their minima, supporting a considerably wider range of $\lam$ values and leading to greater model uncertainty.
+The selected $\lambda$ values, represented by the vertical dashed lines, and their corresponding models, reflect what was observed in the simulation studies.
+The linear predictor approach was most liberal and selected 85 genes.
+While the basic approach and deviance residual approach selected 68 genes and 60 genes respectively, Verweij and Van Houwelingen's approach was most conservative, selecting only 34 genes.
+Similar patterns were observed for other $\alpha$ values.
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[height= 8cm ]{shedden.png}
+  \caption{\label{fig:CVE}Rescaled Cross Validated Errors at $\alpha = 0.3$.
+The horizonal axis is $\log(\lambda)$.
+The vertical axis is the cross-validated error rescaled to a range from 0 to 1 in order to allow plotting of the four methods on a common axis.
+The vertical dashed lines represent the $\lambda$ selected by a specific method.
+The linear predictor approach selected the smallest $\lambda$ and Verweij and Van Houwelingen's approach selected the greatest $\lambda$.}
+\end{figure}
+
+As mentioned earlier, for the elastic net model with tuning parameters chosen by cross-validated linear predictors ($\lam = 0.650$, $\alpha = 0.3$), 85 genes were selected, of which 78 were well-annotated.
+We carried out an enrichment analysis of these 78 genes using ToppGene \citep{Chen2009c}; the five most significantly enriched molecular functions are listed in Table~\ref{Tab:go}.
+Overall, the results are consistent with our understanding of cancer: many of the genes selected by the elastic net model are involved in signaling and the binding of transcription factors and receptors.
+These are precisely the pathways that tend to be disrupted in cancer.
+Chloride channels have also recently been implicated in several cancers as therapeutic targets \citep{Peretti2015}.
+
+\begin{table}[ht]
+\centering
+\caption{\label{Tab:go} Five most significantly enriched molecular functions among genes selected by the linear predictor method with the elastic net ($\alpha=0.3$).
+The five terms have a false discovery rate of 3\% using the Benjamini-Hochberg method.
+Enrichment: frequency of the annotation among the selected genes divided by its frequency in the genome as a whole.}
+\begin{tabular}{llrrr}
+\toprule
+ID & Name & Selected & Enrichment & p \\
+\midrule
+GO:0008134 & transcription factor binding & 10 & 4.1 & 0.00015 \\
+GO:0035257 & nuclear hormone receptor binding & 5 & 7.4 & 0.00057 \\
+GO:0019869 & chloride channel inhibitor activity & 2 & 53.2 & 0.00061 \\
+GO:0003700 & DNA-binding transcription factor activity & 14 & 2.6 & 0.00068 \\
+GO:0005102 & signaling receptor binding & 16 & 2.4 & 0.00083 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Of particular interest is the fact that three of the selected genes (PIK3R1, PDPK1, and ITGB1) are involved in the PTEN pathway.
+The PTEN gene is an important tumor suppressor; mutations in PTEN are among the most common genetic changes found in human cancers and have been linked to cancers of the lung, breast, prostate, brain, and more.
+Time-to-event analyses like this one of gene expression data provide valuable insights for developing gene signatures that can help to identify patients with more severe disruptions to important pathways such as PTEN, thereby informing prognosis and impacting treatment decisions.
+
+\section{Discussion}
+
+Cross-validation is a crutial step in fitting penalized Cox regression models.
+In this paper, we propose two cross-validation methods for Cox Regression: cross-validated linear predictors and cross-validated deviance residuals.
+We conducted a series of simulation experiments to study and compare their performance with two existing approaches, basic cross-validation and Verweij and Van Houwelingen's cross-validated partial likelihood, in selecting tuning parameters when fitting penalized regression models.
+
+Despite the fact Verweij and Van Houwelingen's cross-validated partial likelihood has been widely implemented, its predictive performance appears suboptimal, particularly when the dimension of the data is large and/or signals are weak.
+As illustrated in both simulation studies and real data example, Verweij and Van Houwelingen's approach tends to perform conservatively, selecting fewer features than all other models.
+In the lung cancer data example, the number of genes selected by other methods almost doubled the number of genes selected by Verweij and Van Houwelingen's approach.
+From an application standpoint, using Verweij and Van Houwelingen's approach in high-dimensional applications may lead to more false negatives and miss important true findings.
+
+While the basic approach performed well in terms of predictive performance, this approach suffers from a lack of numerical stability as discussed in Section~\ref{Sec:stability}.
+The cross-validated linear predictor approach performed very similarly to the basic approach in terms of predictive performance, but without any problems due to numerical stability.
+We recommend using the proposed linear predictor approach in practice for selecting regularization parameters in penalized Cox regression.
+
+% Discussion on Deviance Residual
+
+%\par While the Deviance Residual approach has a reasonable performance in scenarios where dimension is high, it tends to act more conservatively than all the other three approaches in general.
+As is noted by Therneau et al, the sums of squared martingale residuals and deviance residuals do not necessary reflect how good the model fit is \citep{Therneau2000modeling}.
+The difficulty of estimating the baseline hazard seems to play an important role.
+A series of simulations were thus designed to study how baseline hazard estimation affects the deviance residuals' performance as a measurement for model fit.
+
+%\par In the following simulation studies, the defined optimal $\lambda$ and the $\lambda$ selected by the Linear Predictor approach were used as benchmarks.
+Various sample sizes and dimensions were varied.
+The survival outcomes were generated both from exponential distribution and Weibull distribution.
+Six different approaches were compared in terms of baseline hazard estimation and their impact on sums of squared deviance residuals.
+Two non-parametric approaches, Kalbfleisch and Prentices' estimator and Breslow's estimator, were used first for estimating the cumulative hazard baseline \citep{Kalbfleisch2011} \citep{breslow1972}.
+As is presented in column 3 and column 5 in Table 2, the sum of squared deviance residuals computed based on both approaches are all very conservative compared to the optimal $\lambda$.
+Their performance is extremely bad when n is large and p is small.
+
+%\par Since both of the non-parameteric estimations seemed to perform well at early time points but poorly later, a weighted approach was then applied to emphasize earlier time points.
+Weights are assigned proportional to number of subjects at risk, making earlier time points weighted more than later time points.
+Results of these two approahces are presented in column 4 and 6 in Table 2.
+The attempt to improve the non-parameteric estimations via weighting does not seem to improve the selection too much.
+
+%\par In the next step,  the non-parametric baseline hazard estimation was replaced by the exact baseline hazard that was used to generate the outcome.
+Results were listed in the seventh column in Table 2.
+When the baseline was known, the performance of sum of squared deviance residuals actually performed really close to the optimal $\lambda$.
+This simulation suggests that when the baseline hazard was accurately estimated, then the sum of squared deviance residuals gave really good measures for the model fit.
+
+%\par In the last step, we replaced the non-parametric baseline hazard estimation with a parametric approach, where we assume the baseline hazard to follow an exponential distribution.
+We derived the estimation for baseline hazard by maximum likelihood approach conditioned on covariates: $\hat{\lambda}_{0} = \frac{\sum d_{i}}{\sum t_{i}e^{\eta_{i}}}$.
+The results are listed in the eighth column in Table 2.
+When this estimation was used for data that were generated from exponential distribution, its performance was as well as when the exact baseline was specified.
+When it was fitted to data that were generated from the Weibull distribution, that is when the model is misspecified, then its performance was even worse than the non-parametric approaches.
+
+%\par Hence, getting accurate estimation of the baseline hazard is crucial to whether or not the sum of squared deviance residuals can be a good measurement for model fit.
+
+%\begin{table}[h]
+%	\small
+%	\caption{Average of Selected $\lambda$ with Different Baseline Estimation for Deviance Residuals}
+	%\centering
+	%\begin{tabular}{lllllllll}
+	%\hline
+	%\\[-0.75em]
+    %& Optimal & Linear Pred & KP & KP & Breslow & Breslow & Exact & Exp\\ 
+    %&  & & & weighted & & weighted & & \\ \cline{2 - 9} 
+	%\textbf{True Baseline: $Exponential(1)$} &&&&&&&&\\
+	%n = 100, p = 10, non-zero $\beta$: 2 & 0.0461 & 0.0665 & 0.1986 & 0.1970 & 0.1908 & 0.1975 & 0.0540 & 0.0539 \\
+	%n = 100, p = 100, non-zero $\beta$: 20 &0.0555 & 0.1841 & 0.2359 & 0.2201 & 0.2146 & 0.2244 & 0.0805 & 0.0985 \\
+	%n = 100, p = 1000, non-zero $\beta$: 20 & 0.0268 & 0.0505 & 0.0948 & 0.0798 & 0.0867 & 0.0809 & 0.0288 & 0.0270 \\
+	%\textbf{True Baseline:} $Weibull$ &&&&&&&& \\
+ 	%n = 100, p = 10, non-zero $\beta$: 2 & 0.0476 & 0.0710 & 0.2004 & 0.1969 & 0.1907 & 0.1907 & N/A & 0.2330 \\ 
+	%\hline
+%\end{tabular}
+%\end{table}
+
+%\section{Appendix}
+
+%\begin{figure}[h]
+%	\centering
+%	\begin{minipage}[t]{10cm}
+%	\begin{algorithm}[H]
+%	\caption{Standard Method}
+%	\par \For{k in 1 to K } 
+%	{
+%		 Fit Cox Regression on $T_{k}$: Obtain model estimates $\hat{\beta}^{-k}$\;
+%		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D_{k}$: $l^{k}(\hat{\beta}^{-k})$\;		
+%	} 
+%	
+%	\par $ CVE_{st} = - \sum_{k = 1}^{K} l^{k}(\hat{\beta}^{-k})$
+% 	\end{algorithm}
+%
+%	\begin{algorithm}[H]
+%  	 \caption{Verweij and Van Houwelingen's Method}
+%	\par \For{k in 1 to K } 
+%	{
+%		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
+%		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $T_{k}$: $l^{-k}(\hat{\beta}^{-k})$\;
+%		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D$: $l(\hat{\beta}^{-k})$\;
+%		 Take the difference: $ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) $\;
+%	} 
+%
+%	\par $ CVE_{vvh} = - \sum_{k = 1}^{K} \left\{ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) \right\}$
+%	\end{algorithm}
+%
+%	\begin{algorithm}[H]
+%    	\caption{Cross-Validated Linear Predictors}
+%	\par \For{k in 1 to K } 
+%	{
+%		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
+%		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
+%	} 
+%
+%	\par $ CVE_{lp} = - l(\hat{\eta}^{-})$
+%	\end{algorithm}
+%	
+%	\begin{algorithm}[H]
+%   	\caption{Sums of Squared Deviance Residuals}
+%	\par \For{k in 1 to K } 
+%	{
+%		Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
+%		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
+%	} 
+%	Compute baseline hazard $\hat{\Lambda}_{0}$ based on $\hat{\eta}^{-}$\;
+%	$\forall$ $ i$ subject $\in \left\{1,2,...,n\right\}$, compute:
+%		\par\hspace{0.5cm} $\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}^{-}_{i}}$;
+%		\par\hspace{0.5cm} $d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}$;
+%	\par $ CVE_{dr} = \sum_{i = 1}^{n} {d}_{i}^2$ \;
+%	\end{algorithm}
+%	\end{minipage}
+%	\caption{Pseudo Code for Cross-validation Procedures}
+%	\end{figure}
+
+\bibliographystyle{SageV}
+\begin{thebibliography}{10}
+\providecommand{\url}[1]{\texttt{#1}}
+\providecommand{\urlprefix}{URL }
+\expandafter\ifx\csname urlstyle\endcsname\relax
+  \providecommand{\doi}[1]{DOI:\discretionary{}{}{}#1}\else
+  \providecommand{\doi}{DOI:\discretionary{}{}{}\begingroup
+  \urlstyle{rm}\Url}\fi
+\providecommand{\eprint}[2][]{\url{#2}}
+
+\bibitem{Cox1972}
+Cox DR.
+\newblock Regression models and life-tables.
+\newblock \emph{Journal of the Royal Statistical Society: Series B
+  (Methodological)} 1972; 34(2): 187--202.
+
+\bibitem{Tibshirani1997}
+Tibshirani R.
+\newblock The lasso method for variable selection in the {Cox} model.
+\newblock \emph{Statistics in Medicine} 1997; 16(4): 385--395.
+
+\bibitem{Fan2002}
+Fan J and Li R.
+\newblock Variable selection for {Cox's} proportional hazards model and frailty
+  model.
+\newblock \emph{The Annals of Statistics} 2002; 30(1): 74--99.
+
+\bibitem{Verweij1993}
+Verweij PJM and Van~Houwelingen HC.
+\newblock Cross-validation in survival analysis.
+\newblock \emph{Statistics in Medicine} 1993; 12(24): 2305--2314.
+
+\bibitem{glmnet}
+Friedman J, Hastie T and Tibshirani R.
+\newblock Regularization paths for generalized linear models via coordinate
+  descent.
+\newblock \emph{Journal of Statistical Software} 2010; 33(1): 1--22.
+
+\bibitem{Therneau1990}
+Therneau TM, Grambsch PM and Fleming TR.
+\newblock Martingale-based residuals for survival models.
+\newblock \emph{Biometrika} 1990; 77(1): 147--160.
+
+\bibitem{CoxRidge}
+Perperoglou A.
+\newblock \emph{{CoxRidge: Cox} models with dynamic ridge penalties}, 2015.
+\newblock \urlprefix\url{https://CRAN.R-project.org/package=CoxRidge}.
+\newblock R package version 0.9.2.
+
+\bibitem{fastcox}
+Yang Y and Zou H.
+\newblock \emph{Fastcox: {L}asso and elastic-net penalized {Cox's} regression
+  in high dimensions models using the cocktail algorithm}, 2017.
+\newblock \urlprefix\url{https://CRAN.R-project.org/package=fastcox}.
+\newblock R package version 1.1.3.
+
+\bibitem{SGL}
+Simon N, Friedman J, Hastie T et~al.
+\newblock \emph{{SGL: Fit} a {GLM} (or {Cox} model) with a combination of lasso
+  and group lasso regularization}, 2018.
+\newblock \urlprefix\url{https://CRAN.R-project.org/package=SGL}.
+\newblock R package version 1.2.
+
+\bibitem{CoxBoost}
+Binder H.
+\newblock \emph{{CoxBoost: Cox} models by likelihood based boosting for a
+  single survival endpoint or competing risks}, 2013.
+\newblock \urlprefix\url{https://CRAN.R-project.org/package=CoxBoost}.
+\newblock R package version 1.4.
+
+\bibitem{mboost}
+Hothorn T, B\"uhlmann P, Kneib T et~al.
+\newblock \emph{{mboost}: {M}odel-based boosting}, 2018.
+\newblock \urlprefix\url{https://CRAN.R-project.org/package=mboost}.
+\newblock {R} package version 2.9-1.
+
+\bibitem{ncvreg}
+Breheny P and Huang J.
+\newblock Coordinate descent algorithms for nonconvex penalized regression,
+  with applications to biological feature selection.
+\newblock \emph{Annals of Applied Statistics} 2011; 5(1): 232--253.
+
+\bibitem{Parker2007}
+Parker BJ, G\"unter S and Bedo J.
+\newblock Stratification bias in low signal microarray studies.
+\newblock \emph{BMC Bioinformatics} 2007; 8: 326.
+
+\bibitem{Simon2011a}
+Simon RM, Subramanian J, Li MC et~al.
+\newblock Using cross-validation to evaluate predictive accuracy of survival
+  risk classifiers based on high-dimensional data.
+\newblock \emph{Briefings in Bioinformatics} 2011; 12(3): 203--214.
+\newblock \doi{10.1093/bib/bbr001}.
+
+\bibitem{Subramanian2011}
+Subramanian J and Simon R.
+\newblock An evaluation of resampling methods for assessment of survival risk
+  prediction in high-dimensional settings.
+\newblock \emph{Statistics in Medicine} 2011; 30(6): 642--653.
+\newblock \doi{10.1002/sim.4106}.
+
+\bibitem{Therneau2018}
+Therneau T, Atkinson B and Ripley B.
+\newblock \emph{{rpart: Recursive} partitioning and regression trees}, 2018.
+\newblock \urlprefix\url{https://CRAN.R-project.org/package=rpart}.
+
+\bibitem{Breslow1972}
+Breslow N.
+\newblock Discussion following ``{R}egression models and life-tables''.
+\newblock \emph{Journal of the Royal Statistical Society Series B
+  (Methodological)} 1972; 34(2): 187--220.
+
+\bibitem{Kalbfleisch2011}
+Kalbfleisch JD and Prentice RL.
+\newblock \emph{The statistical analysis of failure time data}.
+\newblock Wiley Series in Probability and Statistics, John Wiley \& Sons, 2011.
+
+\bibitem{R}
+{R Core Team}.
+\newblock \emph{{R: A} language and environment for statistical computing}.
+\newblock R Foundation for Statistical Computing, Vienna, Austria, 2017.
+\newblock \urlprefix\url{https://www.R-project.org/}.
+
+\bibitem{VanHouwelingen2011}
+Van~Houwelingen H and Putter H.
+\newblock \emph{Dynamic prediction in clinical survival analysis}.
+\newblock CRC Press, 2011.
+
+\bibitem{HarrellJr1984}
+Harrell FE, Lee KL, Califf RM et~al.
+\newblock Regression modelling strategies for improved prognostic prediction.
+\newblock \emph{Statistics in medicine} 1984; 3(2): 143--152.
+
+\bibitem{shedden2008gene}
+Shedden K, Taylor JMG, Enkemann SA et~al.
+\newblock Gene expression--based survival prediction in lung adenocarcinoma: a
+  multi-site, blinded validation study.
+\newblock \emph{Nature Medicine} 2008; 14(8): 822.
+
+\bibitem{Zou2005}
+Zou H and Hastie T.
+\newblock Regularization and variable selection via the elastic net.
+\newblock \emph{Journal of the Royal Statistical Society: Series B
+  (Methodological)} 2005; 67: 301--320.
+
+\bibitem{Chen2009c}
+Chen J, Bardes EE, Aronow BJ et~al.
+\newblock {ToppGene Suite} for gene list enrichment analysis and candidate gene
+  prioritization.
+\newblock \emph{Nucleic Acids Research} 2009; 37: 305--311.
+\newblock \doi{10.1093/nar/gkp427}.
+
+\bibitem{Peretti2015}
+Peretti M, Angelini M, Savalli N et~al.
+\newblock Chloride channels in cancer: {F}ocus on chloride intracellular
+  channel 1 and 4 ({CLIC1} and {CLIC4}) proteins in tumor development and as
+  novel therapeutic targets.
+\newblock \emph{Biochimica et Biophysica Acta (BBA) - Biomembranes} 2015;
+  1848(10, Part B): 2523--2531.
+\newblock \doi{10.1016/j.bbamem.2014.12.012}.
+
+\end{thebibliography}
 

### 2022-01-07 supplementary.tex
Submitted to SMMR

@@
 \documentclass{article}
+
 \usepackage{graphicx}
+\graphicspath{{manuscript_figure/}}
 \usepackage{color}
@@
 \end{equation}
-Under the assumption that there is no ties among failure times, the numerator in the summation is either 0 or 1.
+Under the assumption that there are no ties among failure times, the numerator in the summation is always either 0 or 1.
 The Nelson-Aalen Estimator depends only on the failure times and does not involve any information from the covariates.
@@
 \subsection{Simulation Comparison}
-The following simulation illustrates how the performance of cross-validated deviance residuals would vary based on different baseline estimation.
+
+The following simulation illustrates how the performance of cross-validated deviance residuals varies based on different approaches to baseline estimation.
 We generated data using the same mechanism as we described in Section 3.1 in the paper.
@@
 For each scenario, we varied $s$ from 0.4 to 0.9 and 200 replications were used.
-We compared the cross-validated deviance residuals constructed by Nelson-Aalen estimator and the cross-validated deviance residuals constructed by Kalbfleisch and Prentice's estimator.
-We also included cross-validated linear predictor approach (introduced in Section 2.3 in the paper) for comparison.
-The results of the simulation is listed in the following table:
+We compared the cross-validated deviance residuals where baseline estimation used the Nelson-Aalen estimator, the Breslow estimator, and the Kalbfleisch and Prentice estimator.
+%We also included the cross-validated linear predictor approach (introduced in Section 2.3 in the paper) for comparison.
+The results of the simulation are given in the following table:
 
@@
 \toprule
+ & \multicolumn{6}{c}{$s$} \\ 
  & 0.4 & 0.5 & 0.6 & 0.7 & 0.8 & 0.9\\
@@
 
-Top half of the table listed different $\lambda$ values selected by the method.
-With the Kalbfleisch and Prentice baseline, the cross-validated deviance residual always selected larger $\lambda$ values and very conservative models.
+The top half of the table reports the average $\lambda$ values selected by each method.
+With the Breslow and Kalbfleisch-Prentice approaches, the cross-validated deviance residual always selected larger $\lambda$ values and thus, more conservative models.
 With the covariate-adjusted baseline, the deviance residuals seem to have stronger preference for models closer to the null model.
-This phenomenon is also observed by the original authors of the deviance residuals in classical Cox regression applications.
-As they attempted to use the covariate-adjusted deviance residuals and Martingale residuals to evaluate model fit, they observed that "the no-covariate model has the smallest sum of squared and the best model, the one with all five predictors, has the largest" \citep{Therneau2000modeling}.
+This phenomenon was also observed by the original authors of the deviance residuals in classical Cox regression applications.
+As they attempted to use the covariate-adjusted deviance residuals and Martingale residuals to evaluate model fit, they observed that "the no-covariate model has the smallest sum of squares and the best model, the one with all five predictors, has the largest" \citep{Therneau2000modeling}.
 
@@
 The definition of $\log($MSE Ratio$)$ is described in more details in Section 3.1 in the paper.
-Throughout different signal strength, CVDR with Kalbfleisch and Prentice baseline had the largest $\log($MSE Ratio$)$, indicating worst model estimation among all methods.
-With Nelson-Aalen estimator, the CVDR approach had better performance and performed closer to CVLP at weaker signals.
-As the signal strength increases, the performance of CVDR with Nelson-Aalen estimator also declined.
-
-Overall, our simulation results support the use of Nelson-Aalen estimator in estimating the baseline hazard for the cross-validated deviance residual approach.
+Across all signal strengths, CVDR with Nelson-Aalen baseline estimation had the smallest $\log($MSE Ratio$)$, indicating best model estimation compared to Breslow and Kalbfleisch-Prentice.
+Overall, our simulation results support the use of Nelson-Aalen estimator in estimating the baseline hazard for the cross-validated deviance residual approach, and was therefore employed in the manuscript.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{manuscript_figure/figure_2_ex.png}
+  \includegraphics[width=\textwidth]{figure_2_ex.png}
   \caption{\label{Fig:ex} The horizontal axis in all three plots is the correlation among the features.
@@
   \centering
-  \includegraphics[width=\textwidth]{manuscript_figure/figure_2_auto.png}
+  \includegraphics[width=\textwidth]{figure_2_auto.png}
   \caption{\label{Fig:auto} The horizontal axis in all three plots is the correlation among the features.
@@
 
-\bibliographystyle{abbrvnat}
+\bibliographystyle{ims-nourl}
 \bibliography{articles}

### 2022-08-10 sub/smimr/1/cover.tex
SMM: Referee comments

--- 2019-05-24 sub/sim/cover.tex
@@
 
-\StartLetter{}{Dear editors of {\em Statistics in Medicine},}
+\StartLetter{}{Dear editors of {\em Statistical Methods in Medical Research},}
 
-We are submitting the manuscript ``Cross validation approaches for penalized Cox regression'' for your consideration to be published in {\em Statistics in Medicine}.
+We are submitting the manuscript ``Cross validation approaches for penalized Cox regression'' for your consideration to be published in {\em Statistical Methods in Medical Research}.
 
@@
 This work proposes a new approach for CV in these models and demonstrates that it achieves a combination of performance and stability that existing methods do not match.
-We believe this is an accessible article of clear practical importance, and would be interesting to readers of {\em Statistics in Medicine}.
+We believe this is an accessible article of clear practical importance, and would be interesting to readers of {\em Statistical Methods in Medical Research}.
 

### 2022-08-14 sub/smimr/response.tex
start a response file for SMM comments

\documentclass[12]{article}
\usepackage[utf8]{inputenc}
\usepackage[a4paper,margin = 1in]{geometry}
\usepackage{enumitem}
\usepackage{color}

\newcommand{\re}{\textbf{Response: }}
\newcommand\bd[1]{{\color{blue}#1}}

\begin{document}

\section*{Response to Reviewers}
\subsection*{Manuscript ID SMM-22-0007: Cross Validation Approaches for
Penalized Cox Regression}

The authors wish to thank the reviewers for their careful reviews and thoughtful comments.

\subsection*{Response to Reviewer 1's Comments}

The authors compare four different cross-validation options for the important problem of penalized Cox regression, including two new proposed methods. The methodology is nicely described, and the illustration in Figure 1 should prove helpful to readers new to this area. The interpretations of the simulation results are helpful, and a good choice for the application dataset was made. The take-home message of the usefulness of the linear predictor approach was clearly made based on the work performed. However, there are a number of important questions and comments to mention here that should be addressed:

\begin{enumerate}[align = left]
\item On p.4, roughly line 34, you have the statement “although the methods we analyze can be used with any penalty”.  This statement should be further elaborated, providing some tangible examples (including beyond the straightforward elastic net approach considered in the application analysis).

\re \bd{should be easy to address. follow the suggestion to include a few examples of different penalties}

\item Though you aim to make various considerations in your simulation study, the impact of this work may very well be limited given consideration of only exponential survival distribution and independent censoring. 
    \begin{enumerate}
        \item Extensions into time-varying covariates, and, preferably, joint modeling, would be useful as well. At least the first two items (more flexible survival distributions, and dependent censoring) should be explored in your work, while perhaps extensions into joint modeling to accommodate one or more longitudinal predictors could be discussed as future work in your Discussion.

        \re \bd{selection among a few time-varying covariates (reviewer says one or two...) isn't really within the scope of the common application of penalized models. This could be an extension of using cross-validation methods, but doesn't really fit here... Maybe the reviewer means how the inference for the time-varying covariate would be different based on the other covariates that are selected? This is an estimation problem, also not within the scope of this work. Maybe in the application example we can look into whether we can include one time-varying covariate to illustrate?}

        \item Regarding distributional form, the exponential form here is quite limiting, and even common choices for a more complex distributional form, such as Weibull and Gompertz, might be considered too limiting. See, for example, Harden and Kropko (2019).

        \re \bd{ for Cox models, as long as the underlying assumption for proportional hazard doesn't change, then the underlying distribution (I believe neither Weibull nor Gompertze violates proportional hazard assumption) used to simulate the data shouldn't matter}

    \end{enumerate}

\item Speaking of your Discussion section, there is no mention of any limitations of the current work. Aside from the joint modeling comment above, you should also talk about violations of linearity, and of non-proportional hazards. That is, you need to provide a number of additional sentences, and references, as warranted, to talk about how the current contributions could ultimately be made more generalizable to various survival dataset characteristics, and not just those with a very simplifying structure. Some discussion on how to think about survival datasets with rare events should be commented on as well.

\re \bd{Easy to address. To do: add following topics to Discussion: violation of linearity, non-proportional hazards, rare events}

\item On p.10, second sentence first paragraph of Section 3.1, should mention that the specifics of the different conditions considered (in the simulation study) are described below.

\re \bd{Easy to address. To do: add more descriptions in that paragraph}

\item In Section 3.1, p.11, first line (and in subsequent uses below), I would suggest adding a $t_0$ index to $y$, since the definition of $y$ is a direct function of $t_0$.

\re \bd{Easy to address. To do: update notation in manuscript}

\item And only looking at one $t_0$ in your simulation seems limiting here (for the Brier and KL score calculations). Though if just one choice is used it makes sense to use the median, explain why only one choice of $t_0$ is considered.  This would be especially true if you consider other more complicated survival distributions.

\re \bd{looks like the reviewer really want to see more $t_0$ for these criteria. There is a global version for Brier and KL score that take into account all time points or look at interval of times. Maybe mention this Discussion as a supplement?}

\item Was confused on a point made in the second to final sentence on p.12. It says “In all settings, the censoring rate was 30$\%$”. Is this just set at 30$\%$ for the additional simulations, noting that you just described a setting in the prior paragraph with 10$\%$ expected censoring. Please clarify, including in the text.

\re \bd{Easy to address. To do: clarify more in the writing}

\item There is an interesting semantic point to raise. You comment on the approaches taking/accepting fewer variables as conservative and, thus, those accepting more variables as liberal. However, an argument can be made that a method making fewer decisions, i.e., not throwing out as many variables as another, as being more cautious/conservative. It’s worth elaborating on this point.

\re \bd{To do: address this in Discussion}

\item In the application section, you mention fitting elastic net models to the gene expression dataset, since elastic net does a better job than traditional Lasso when modeling many correlated genes. However, there is evidence that newer Lasso approaches (e.g., Precision Lasso: Wang et al, 2019, Bioinformatics) outperform elastic net models in this setting, so you need to justify why only the elastic net is considered as the only alternative to traditional Lasso here. Perhaps you can mention potential alternatives to elastic net as an improvement in Section 4, and include this as an area of future focus in the Discussion.

\re \bd{similar to first comment. easy to address in the writing}

\item Regarding elastic net, it was not justified why other alpha values (beyond 0.3, 0.5, 0.7, and 0.9) are not considered in the analysis, such as a denser grid between 0.1 and 0.5, particularly after seeing the results (in Table 3) from the sparse selection of alpha values used. More alpha values should be considered.

\re \bd{maybe the easiest way is to use more alpha values in this example?}

\item There are several typos sprinkled throughout the paper, so a careful proofreading is required. Some examples include: (i) Section 3.1, second paragraph, where the phrase “the value of $\beta$ used the generate the data” requires editing. (ii) Table 1, Scenario 4, would seem to be for
strong signal, not weak. (iii) p.22, “more clear” should be changed to “clearer”.

\re \bd{to do: make edits as suggested}

\end{enumerate}

\subsection*{Response to Reviewer 2's Comments}
In this manuscript, the authors proposed two new cross-validation methods for Cox proportional hazards regression.

\begin{enumerate}[align = left]
\item Throughout the paper, the authors mainly (almost only) advocate for use of one of the proposed methods, “cross validated linear predictors”. Also based on numerical studies, I see no point in presenting the other proposed method, “cross validated deviance residual”. If the authors want to present both proposed methods, the cross validated deviance residual method should be better highlighted in numerical studies and Discussion for its advantages, which is not sufficiently illustrated in the current version of paper.

\re \bd{maybe one advantage is that the deviance residuals can be applied to a more wide range of survival models other than the regression approaches. say, for a random forest model with survival outcome, they can't really use the linear predictor approach right?}

\item Presentation of simulation studies is quite confusing so should be streamlined. Specifically, data characteristics such as sample size and censoring rate are not consistent across simulation scenarios considered for each of different evaluation metrics in Section 3.

    \begin{enumerate}
    \item I suggest presenting all the simulation scenarios varying different data characteristics at the beginning of Section 3 and discuss the comparative performance of the methods using each of model evaluation metrics in the following subsections. Addressing this comment will require additional computations for Sections 3.3-3.5

    \re \bd{I think there are definitely values to present a different variety of simulation scenarios. but maybe we should re-structure the writing a bit more}

    \item Considering the censoring rate of 50$\%$ in the application, I would consider scenarios with a higher censoring rate (say 50$\%$) than what was considered in Sections 3.1 and 3.2 (10$\%$ and 30$\%$). This will also help justify authors’ argument in P17L34.

    \re \bd{I remember looking at higher censorings before. the conclusion doesn't change with the censoring proportion. We can include this in the writing, rather than including more simulations}

    \item Table 2 - Why only present the number of true positives here? I suggest presenting other operating characteristics such as TPR, NPR, PPV, etc... Then the results might help better justify the findings from the application. For example, depending on the results, referring to the simulation results could strengthen the argument given at the end of the second paragraph in Discussion (which is quite speculative in my opinion).

    \re \bd{one argument would be we are evaluting the predictive performance. This comment can be addressed jointly with Reviewer 1's comment 8}

\end{enumerate}

\item Did the authors make the software available for the proposed approaches?

\re \bd{of course}

\item Many places require major/minor clarification and corrections:
\bd{To do: make those edits in the manuscript}
\begin{itemize}[align = left]
\item[-]  P3L33: define beta

\re

\item[-] Distinguish between scalars and vectors by using bold letters for vectors. e.g. beta

\re

\item[-] Figure 1: Since “Folder 1” is highlighted for each panel, I think the figure corresponds to the specific case with k=1. If so, k should be replaced by 1 in expressions.

\re

\item[-]  P8L5: I would not use “k” in the expression because k is used for k-th “fold” in this paper.

\re

\item[-]  Equation (6) and P8L10: I found the definition of $\eta^{CV}$ confusing. Do we not need some step for pooling cross-validated linear predictors over all K folds?

\re

\item[-]  Table 1: what are the numbers in parentheses?

\re

\item[-] Figure 4 caption: LOO is not defined

\re

\item[-]  Figure 4 caption: (blue line) -> (green line)?

\re

\item[-]  Figure 4 legend: The “Method” can mean two different things in this paper – either (four different CV methods) or (10 fold CV, LOOCV). Therefore, the legend is very confusing because “Basic” here actually means basic CV with 10 folds. Although the text in the paper clarifies this, tables and figures need to be understandable by themselves without reference to the text.

\re

\item[-]  Table 4, what are “Selected” and “P” here?

\re

\end{itemize}
\end{enumerate}

\subsection*{References}

Harden JJ, Kropko J (2019). Simulating duration data for the Cox model, Political Science Research and Methods, 7(4): 921-928. https://doi.org/10.1017/psrm.2018.19

Wang H, Lengerich BJ, Aragam B, Xing EP (2019). Precision Lasso: accounting for correlations and linear dependencies in high-dimensional genomic data. Bioinformatics, 35(7), 1181-1187. https://doi.org/10.1093/bioinformatics/bty750

\end{document}

### 2022-10-13 sub/smimr/response.tex
updates to response, supplementary and main

@@
 \usepackage{color}
+\usepackage{amssymb,amsmath}
 
@@
 \newcommand\bd[1]{{\color{blue}#1}}
+\newcommand{\bbeta}{\boldsymbol{\beta}}
+\newcommand{\boldeta}{\boldsymbol\eta}
 
@@
 
-\re \bd{should be easy to address. follow the suggestion to include a few examples of different penalties}
+\re Thank you for this comment.
+We provided a few examples of penalty functions, such as the elastic net penalty, smoothly clipped absolute deviations (SCAD) penalty, and minimax concave penalty (MCP) on page 4.
+We also included a paragraph for more general application in the Discussion section.
 
@@
 
-        \re \bd{selection among a few time-varying covariates (reviewer says one or two...) isn't really within the scope of the common application of penalized models.
-This could be an extension of using cross-validation methods, but doesn't really fit here...
-Maybe the reviewer means how the inference for the time-varying covariate would be different based on the other covariates that are selected?
-This is an estimation problem, also not within the scope of this work.
-Maybe in the application example we can look into whether we can include one time-varying covariate to illustrate?}
+        \re Thank you for making this comment.
+This work aims to make comparisons among cross-validation methods, which are more commonly used in the model selection context.
+In the Discussion section, we included a paragraph to state that the methods in this manuscript can be applied to any model selection process that involves the cox partial likelihood.
+While cross-validation and model selection play a key role for fitting penalized cox regression models, the process of fitting time-varying covariates and joint modeling with longitudinal covariates do not always involve model selection, therefore this extension is out of the main focus of this work.
 
@@
 
-        \re \bd{ for Cox models, as long as the underlying assumption for proportional hazard doesn't change, then the underlying distribution (I believe neither Weibull nor Gompertze violates proportional hazard assumption) used to simulate the data shouldn't matter}
+        \re Thank you for this comment.
+We included additional simulation results in the supplementary material where the baseline hazard is generated from Weibull and Gompertz distribution.
+The overall conclusions do not change for both of these more complex distributional form.
 
@@
 
-\re \bd{Easy to address. To do: add following topics to Discussion: violation of linearity, non-proportional hazards, rare events}
+\re Thank you for the comments.
+The violation of linearity is usually handled via selecting or fitting a non-linear functional form for the covariates.
+The approaches that are discussed here are built upon Cox partial likelihood, which do not deal with non-proportional hazards and rare events, therefore those topics are out of the scope of this manuscript.
+We included sentences in the Discussion section to address those points.
 
@@
 
-\re \bd{Easy to address. To do: add more descriptions in that paragraph}
+\re Thank you for the suggestion. We added the sentence to the first paragraph of Section 3.1.
 
@@
 
-\re \bd{Easy to address. To do: update notation in manuscript}
+\re Thank you for the suggestion. We added $t_0$ index to $y$ notation.
 
@@
 
-\re \bd{looks like the reviewer really want to see more $t_0$ for these criteria.
-There is a global version for Brier and KL score that take into account all time points or look at interval of times.
-Maybe mention this Discussion as a supplement?}
+\re Thank you for this comment.
+We provided additional simulation results in the supplement with more choices of $t_0$ with Exponential, Weibull and Gompertz distributions.
+The overall conclusions do not change with various $t_0$.
 
@@
 
-\re \bd{Easy to address. To do: clarify more in the writing}
+\re Thank you for catching this. We clarified in the writing that this censoring rate only applies to settings in Table 1.
 
@@
 
-\re \bd{similar to first comment. easy to address in the writing}
+\re Thank you for the suggeston. We included the precision lasso method in the discussion section.
 
@@
 
-\re \bd{to do: make edits as suggested}
+\re Thank you for catching those. We made edits in the manuscript.
 
@@
 
-    \re \bd{I remember looking at higher censorings before. the conclusion doesn't change with the censoring proportion. We can include this in the writing, rather than including more simulations}
+    \re We explored the impact of censoring in Section 3.1, varying from 40$\%$ and to 80$\%$.
+The censoring proportion has the greatest impact on the stability for the basic method, but does not impact the overall conclusions for the other methods.
 
@@
 
-\re \bd{of course}
+\re The cross-validated linear predictor approach is implemented in the \texttt{ncvreg} package.
+The basic approach and the Verweij and Van Houwelingen's approach are available with the \texttt{glmnet} package.
+\bd{are we submitting the R codes as supplementary material?}
 
 \item Many places require major/minor clarification and corrections:
-\bd{To do: make those edits in the manuscript}
+\bd{the following points are all edited in the manuscript}
 \begin{itemize}[align = left]
@@
 
-\re
+\re Thank you for the suggestion. We included definition for $\bbeta$ in this paragraph when first introducing the notation.
 
-\item[-] Distinguish between scalars and vectors by using bold letters for vectors. e.g. beta
+\item[-] Distinguish between scalars and vectors by using bold letters for vectors
 
-\re
+\re Thank you for the suggestion. We updated the manuscript by using bold betters for vectors.
 
@@
 
-\re
+\re Thank you for this careful observation.
+In the figure, we used "Folder 1" for the simplicity of drawing.
+However, replacing $k$ with $1$ in the expressions in the figure may cause confusion to the readers.
+Therefore, we added description in the caption of Figure 1 to clarify that this illustration represents $k = 1$.
 
@@
 
-\re
+\re Thank you for the suggestion. We replaced $k$ with $\tau$ in that formula
 
-\item[-]  Equation (6) and P8L10: I found the definition of $\eta^{CV}$ confusing. Do we not need some step for pooling cross-validated linear predictors over all K folds?
+\item[-]  Equation (6) and P8L10: I found the definition of $\hat{\eta}^{\text{cv}}$ confusing. Do we not need some step for pooling cross-validated linear predictors over all K folds?
 
-\re
+\re No additional steps are needed to pool the cross-validated linear predictors over all K folds.
+Please refer to the linear predictor diagram in Figure 1.
+Suppose each fold has $n_k$ number of observations and the complete data have $n = \sum_{k=1}^{K}  n_k$ observations.
+The cross-validated linear predictor of the $k$th folder $\hat{\boldeta}^{\text{cv}}_k$ is a vector of length $n_k$ and the complete set of cross-validated linear predictors $\hat{\boldeta}^{\text{cv}}$ has a length of $n$.
+We also updated the vector $\boldeta$ by using bold.
 
@@
 
-\re
+\re The numbers in the parentheses are standard deviation across all simulated samples, which is a metric for simulation errors.
 
@@
 
-\re
+\re Thank you for the suggestion. We updated the caption to include defition for LOO.
 
-\item[-]  Figure 4 caption: (blue line) -> (green line)?
+\item[-]  Figure 4 caption: (blue line) should be (green line)?
 
-\re
+\re Thank you for catching this. We corrected this in the caption.
 
@@
 
-\re
+\re Thank you for the suggestion. We edited the caption of Figure 4 to clarify the legends.
 
@@
 
-\re
+\re The column "Selected" represents the number of genes that are selected by the linear predictor cross-validation approach; each row represents the numbder of genes that fall into that function category.
+$p$ is the p-value.
+\bd{is the p-value adjusted?
+update $p$ with "p-value"?}
 

### 2022-10-13 main.tex
updates to response, supplementary and main

@@
 \begin{equation*}
-  h_{i}(t) = h_{0}(t) \exp( X_{i}\Tr\beta),
+  h_{i}(t) = h_{0}(t) \exp( X_{i}\Tr\bbeta),
 \end{equation*} 
-where $h_{0}$ is the baseline hazard and $\exp(X_i\Tr\beta)$ is the risk for a subject with covariates $X_i$ relative to that baseline.
+where $\bbeta$ is the coefficient vector for the model and $h_{0}$ is the baseline hazard and $\exp(X_i\Tr\bbeta)$ is the risk for a subject with covariates $X_i$ relative to that baseline.
 The estimation of the coefficients is obtained by maximizing a partial likelihood.
@@
   \label{eq:cox-pl-subj}
-  L_{i}(\beta) = \left \{\frac{\exp ( X_{i}\Tr\beta)}{\sum_{ k \in R(t_{i})}\exp ( X_{k}\Tr \beta)}\right \}^{\delta_{i}},
+  L_{i}(\bbeta) = \left \{\frac{\exp ( X_{i}\Tr\bbeta)}{\sum_{ k \in R(t_{i})}\exp ( X_{k}\Tr \bbeta)}\right \}^{\delta_{i}},
 \end{equation}
@@
   \label{eq:cox-pl-sum}
-  L(\beta) =\prod_{i = 1}^{n} L_{i}(\beta).
+  L(\bbeta) =\prod_{i = 1}^{n} L_{i}(\bbeta).
 \end{equation}
-Cox regression can be extended by introducing a penalty into the partial likelihood.  In penalized Cox regression, $\beta$ coefficient estimates are obtained by minimizing the objective function
+Cox regression can be extended by introducing a penalty into the partial likelihood.  In penalized Cox regression, $\bbeta$ coefficient estimates are obtained by minimizing the objective function
 \begin{equation}
   \label{eq:obj}
-  Q(\beta) = - \frac{1}{n} \log L(\beta) + P_{\lambda}(\beta),
+  Q(\bbeta) = - \frac{1}{n} \log L(\bbeta) + P_{\lambda}(\bbeta),
 \end{equation}
-where $P_{\lambda}(\beta)$ is a penalty function that depends on a regularization parameter $\lam$.
+where $P_{\lambda}(\bbeta)$ is a penalty function that depends on a regularization parameter $\lam$.
 Throughout, we will assume without loss of generality that the covariates have been standardized such that $\sum_i x_{ij} = 0$ and $\tfrac{1}{n}\sum_i x_{ij}^2=1$.
@@
 
-\par In this paper, we focus upon the LASSO penalty $P_{\lambda}(\beta) = \lambda \sum_{j} |\beta_{j}|$, although the methods we analyze can be used with any penalty as well as with model selection in unpenalized Cox regression.
+\par In this paper, we focus upon the LASSO penalty $P_{\lambda}(\bbeta) = \lambda \sum_{j} |\beta_{j}|$, although the methods that we analyze not only can be used for penalized Cox regression with any penalty, such as elastic net \citep{Zou2005}, smoothly clipped absolute deviations (SCAD) penalty \citep{FanLi2001}, and minimax concave penalty (MCP) \citep{Zhang2010}, but are also applicable to model selection in unpenalized Cox regression.
 LASSO-penalized Cox regression is particularly useful for high dimensional data where the number of covariates $p > n$, such as predicting overall survival in cancer patients based on genome-wide expression measurements.
@@
 The log partial likelihood is denoted by $\ell$, with $\ell^k$ and $\ell^{-k}$ defined similarly.
-Finally, let $\hat{\beta}^{-k}$ denote the penalized estimates that are obtained by using $L^{-k}(\beta)$ in \eqref{eq:obj}.
+Finally, let $\hat{\bbeta}^{-k}$ denote the penalized estimates that are obtained by using $L^{-k}(\bbeta)$ in \eqref{eq:obj}.
 Figure~\ref{Fig:methods} diagrams the relationships between the four cross-validation methods we will describe in this section.
@@
 
-\par A direct approach to measure a model's predictive accuracy in $D_k$ is to evaluate the log partial likelihood at $\hat{\beta}^{-k}$ using the data in $D_k$.
+\par A direct approach to measure a model's predictive accuracy in $D_k$ is to evaluate the log partial likelihood at $\hat{\bbeta}^{-k}$ using the data in $D_k$.
 Cross-validation error (CVE) is then measured by the total deviance across all $K$ folds, which (up to a constant) is given by:
@@
   \label{eq:basic}
-  \CVE = -2 \sum_{k=1}^{K} \ell^{k}(\hat{\beta}^{-k});
+  \CVE = -2 \sum_{k=1}^{K} \ell^{k}(\hat{\bbeta}^{-k});
 \end{equation}
@@
   \label{eq:VVH}
-  \CVE = -2\sum_{k = 1}^K \left\{ \ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k}) \right\}. 
+  \CVE = -2\sum_{k = 1}^K \left\{ \ell(\hat{\bbeta}^{- k})  - \ell^{-k}(\hat{\bbeta}^{- k}) \right\}. 
 \end{equation}
-Here, $\ell(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ using the entire data set $D$ and $\ell^{-k}(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on
+Here, $\ell(\hat{\bbeta}^{-k})$ is the log partial likelihood evaluated at $\hat{\bbeta}^{-k}$ using the entire data set $D$ and $\ell^{-k}(\hat{\bbeta}^{-k})$ is the log partial likelihood evaluated at $\hat{\bbeta}^{-k}$ on
 the training set $T_k$.
@@
 We refer to this cross-validation method as the Verweij and Van Houwelingen (V\&VH) approach.
-For many models, such as logistic regression or linear regression, the quantities $\ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k})$ and  $\ell^{k}(\hat{\beta}^{-k})$ are equivalent to each other -- in other words, the basic and V\&VH approaches agree.
+For many models, such as logistic regression or linear regression, the quantities $\ell(\hat{\bbeta}^{- k})  - \ell^{-k}(\hat{\bbeta}^{- k})$ and  $\ell^{k}(\hat{\bbeta}^{-k})$ are equivalent to each other -- in other words, the basic and V\&VH approaches agree.
 However, Verweij and Van Houwelingen's approach is more stable for Cox regression as there is always a large number of observations in the risk sets its calculations are based on.
@@
   \includegraphics[height= 12cm ]{figure_1_new.png}
-  \caption{\label{Fig:methods} Illustration of $K$-fold cross validation methods for Cox regression.}
+  \caption{\label{Fig:methods} Illustration of $K$-fold cross validation methods for Cox regression when $k = 1$}
 \end{figure}
@@
 We refer to this method as the \emph{cross-validated linear predictors} approach.
-To be more specific, we would first obtain $\hat{\beta}^{-k}$ from training set $T_{k}$; then the cross-validated linear predictors can be calculated for each observation $i$ from the test set $D_k$:
+To be more specific, we would first obtain $\hat{\bbeta}^{-k}$ from training set $T_{k}$; then the cross-validated linear predictors can be calculated for each observation $i$ from the test set $D_k$:
 \begin{equation}
   \label{eq:cv-lp}
-  \hat{\eta}^{\text{cv}}_{i} = X_{i}\Tr\hat{\beta}^{-k}
+  \hat{\eta}^{\text{cv}}_{i} = X_{i}\Tr\hat{\bbeta}^{-k}
 \end{equation} 
-After repeating this for all K folds, a complete set of cross-validated linear predictors $\hat{\eta}^{\text{cv}} = ( \hat{\eta}^{\text{cv}}_{1},  \hat{\eta}^{\text{cv}}_{2} , ...
+After repeating this for all K folds, a complete set of cross-validated linear predictors $\hat{\boldeta}^{\text{cv}} = ( \hat{\eta}^{\text{cv}}_{1},  \hat{\eta}^{\text{cv}}_{2} , ...
 \hat{\eta}^{\text{cv}}_{n})$ for the whole sample is obtained.
@@
 \begin{equation*}
-  L(\hat{\eta}^{\text{cv}}) = \prod_{i=1}^{n} \left\{ \frac{\exp (\hat{\eta}^{\text{cv}}_{i})}{\sum_{k \in R(t_{i})}\exp (\hat{\eta}^{\text{cv}}_k)} \right\} ^{\delta_i}.
+  L(\hat{\boldeta}^{\text{cv}}) = \prod_{i=1}^{n} \left\{ \frac{\exp (\hat{\eta}^{\text{cv}}_{i})}{\sum_{\tau \in R(t_{i})}\exp (\hat{\eta}^{\text{cv}}_\tau)} \right\} ^{\delta_i}.
 \end{equation*}
-We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - 2 \log L(\hat{\eta}^{\text{cv}}).$$
+We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - 2 \log L(\hat{\boldeta}^{\text{cv}}).$$
 
@@
   \label{eq:res-m}
-  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(X_i\Tr\beta), \\
+  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(X_i\Tr\bbeta), \\
   \label{eq:res-d}
@@
 
-In order to obtain a cross-validated version of these residuals, we would replace $\beta$ in the above equations with $\hat{\beta}^{-k}$.
+In order to obtain a cross-validated version of these residuals, we would replace $\bbeta$ in the above equations with $\hat{\bbeta}^{-k}$.
 However, it is not obvious how to obtain a cross-validated version of $H_0$, or to what extent this is necessary.
@@
 For these simulation studies, the covariate matrix $X$, consisting of $n$ observations and $p$ covariates, was generated from independent $\text{Normal}(0, 1)$ samples.
-Given $X_i$ and a coefficient vector $\beta$, survival times were generated from an exponential distribution with rate parameter $\exp(X_i \Tr \beta)$.
+Given $X_i$ and a coefficient vector $\bbeta$, survival times were generated from an exponential distribution with rate parameter $\exp(X_i \Tr \bbeta)$.
 Censoring status indicators were drawn independently from the binomial distribution.
@@
 Accuracy was assesed via several criteria.
+The specifics of the different conditions considered in each simulation experiment are described below.
 
 Estimation accuracy was measured using the mean squared error of the coefficients ($\MSE$).
-Suppose $\hat{\beta}(\lambda)$ is the vector of coefficients estimated by a Cox model with LASSO penalty at $\lambda$, and $\beta^*$ the value of $\beta$ used the generate the data.
+We define $\hat{\bbeta}(\lambda)$ as the vector of coefficients estimated by a Cox model with LASSO penalty at $\lambda$, and define $\bbeta^*$ as the value of $\bbeta$ that is used the generate the data.
 Mean squared error is defined to be
 \begin{equation}
-\MSE = E\| \hat{\beta}(\lambda) - \beta^* \| ^2.
+\MSE = E\| \hat{\bbeta}(\lambda) - \bbeta^* \| ^2.
 \end{equation}
@@
 For each generated data set, an oracle model was fit using Cox regression including only the true features (i.e., those with non-zero coefficients).
-Letting $\hat{\beta}^{\text{oracle}}$ denote the resulting estimates, the relative MSE, on the log scale, is given by
+Letting $\hat{\bbeta}^{\text{oracle}}$ denote the resulting estimates, the relative MSE, on the log scale, is given by
 \begin{equation}
-\log\left\{ \frac{\MSE(\hat{\beta}(\lambda^{\text{cv}}))}{\MSE(\hat{\beta}^{\text{oracle}})} \right\},
+\log\left\{ \frac{\MSE(\hat{\bbeta}(\lambda^{\text{cv}}))}{\MSE(\hat{\bbeta}^{\text{oracle}})} \right\},
 \end{equation}
 where $\lambda^{\text{cv}}$ denotes the value of $\lambda$ selected by cross-validation.
-MSE is estimated by averaging over $N$ replications respectively for $\hat{\beta}(\lambda^{\text{cv}})$ and $\hat{\beta}^{\text{oracle}}$, with the ratio taken afterwards.
+MSE is estimated by averaging over $N$ replications respectively for $\hat{\bbeta}(\lambda^{\text{cv}})$ and $\hat{\bbeta}^{\text{oracle}}$, with the ratio taken afterwards.
 The $\log$ is taken to correct for skewness.
@@
 \par We also measure the predictive accuracy via the mean squared prediction error, or Brier score \citep{VanHouwelingen2011}.
-Let $\hat{S}(t_0|\lambda,x)$ denote the predicted probability for an individual with covariates $x$ to survive beyond time $t_0$, and let $y = \mathbb{1}\left\{ t > t_{0}\right\}$ indicate whether the individual actually survived beyond time $t_0$.
+Let $\hat{S}(t_0|\lambda,x)$ denote the predicted probability for an individual with covariates $x$ to survive beyond time $t_0$, and let $y(t_0) = \mathbb{1}\left\{ t > t_{0}\right\}$ indicate whether the individual actually survived beyond time $t_0$.
 Then the Brier score is defined to be
 \begin{equation}
-\text{Brier}(y, \hat{S}(t_0|\lambda,x)) = (y - \hat{S}(t_0|\lambda,x))^2.
+\text{Brier}(y(t_0), \hat{S}(t_0|\lambda,x)) = (y(t_0) - \hat{S}(t_0|\lambda,x))^2.
 \end{equation}
@@
 \begin{equation}
-	\text{KL}(y, \hat{S}(t_0|\lambda,x)) = -\left\{ y\log(\hat{S}(t_0|\lambda,x) + (1 - y)\log(1 - \hat{S}(t_0|\lambda,x)) \right\}.
+	\text{KL}(y(t_0), \hat{S}(t_0|\lambda,x)) = -\left\{ y(t_0)\log(\hat{S}(t_0|\lambda,x) + (1 - y(t_0))\log(1 - \hat{S}(t_0|\lambda,x)) \right\}.
 \end{equation}
@@
 For our first simulation study, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates.
-The covariate vector $\beta$ is assumed to be sparse, with $\beta_j = 0$ for $j \in \{11, 12, \ldots, 1000\}$ and $\beta_j = s$ for $j \in \{1, 2, \ldots, 10 \}$, where $s$ is a prespecified constant that specifies the signal strength in the simulated data set.
+The covariate vector $\bbeta$ is assumed to be sparse, with $\beta_j = 0$ for $j \in \{11, 12, \ldots, 1000\}$ and $\beta_j = s$ for $j \in \{1, 2, \ldots, 10 \}$, where $s$ is a prespecified constant that specifies the signal strength in the simulated data set.
 Here, we vary $s$ from 0.4 to 0.9.
@@
 \addlinespace
-&\multicolumn{5}{c}{\textbf{Scenario 4}: n = 400, p = 10000, 20 non-zero features, weak signal}\\
+&\multicolumn{5}{c}{\textbf{Scenario 4}: n = 400, p = 10000, 20 non-zero features, strong signal}\\
 \cline{2-6}
@@
 Specifically, we examined the following settings: low dimension (p = 50) with weak signal, low dimension with strong signal, high dimension (p = 10000) with weak signal, and high dimension with strong signal.
-In all settings, the censoring rate was 30\%.
+In all settings presented in Table~\ref{Tab:sim}, the censoring rate was 30\%.
 For weak cases, signal strength $s$ was set to be 0.3; for strong cases, $s$ was set to be 0.6.
+More simulation scenarios are explored using Weibull and Gompertz distribution to generate the baseline hazard, and are evaluated at various $t_0$ for the Brier Score and Kullback Leibler score.
+The details of the simulation setup and results are included in the supplementary material.
+The overall findings of those simulation experiments are consistent with the findings in Table~\ref{Tab:sim}.
 
@@
   \includegraphics[width=\textwidth]{figure_4.png}
-  \caption{\label{Fig:loocv}Simulation comparing LOO and 10-fold CV across three different approaches.
+  \caption{\label{Fig:loocv}Simulation comparing LOO (Leave-One-Out) and 10-fold CV across three approaches (V\&VH, linear predictor, and deviance residual).
 With $N = 200$ replications, the data sets generated have dimension $n = 120$ and $p = 1000$.
 The vertical axis is the $\log$(MSE ratio).
+Each panel represents one approach.
 In each panel, the LOO version and 10-fold version of the same approach are plotted.
-In all three panels, 10-fold basic cross-validation is included as a baseline for comparison (blue line).}
+In all three panels, the 10-fold basic cross-validation is included as a baseline for comparison (green line).}
 \end{figure}
@@
 \centering
-\caption{\label{tb:shedden} The C Index and Brier Score of all elasticnet models selected by cross-validation methods }
+\caption{\label{tb:shedden} The C Index and Brier Score of all elastic net models selected by cross-validation methods }
 \begin{tabular}{lccccc}
@@
 The selected $\lambda$ values, where a given CVE curve reaches its lowest point, are indicated with dotted lines.
-The CVE curves for the basic and linear predictor approaches have greater curvature near their minima, and thus offer more clear support for the selected value of $\lam$.
+The CVE curves for the basic and linear predictor approaches have greater curvature near their minima, and thus offer clearer support for the selected value of $\lam$.
 In comparison, the deviance residual and V\&VH approaches are flatter near their minima, supporting a considerably wider range of $\lam$ values and leading to greater model uncertainty.
@@
 We recommend using the proposed linear predictor approach in practice for selecting regularization parameters in penalized Cox regression.
+
+Although we focused upon the penalized Cox regression with a LASSO penalty in this work, the methods that we analyzed are applicable in any context that involves conducting model selection for survival models that are built upon the Cox partial likelihood, such fitting penalized Cox regression with other penalty functions (elastic net \citep{Zou2005}, SCAD\citep{FanLi2001}, MCP\citep{Zhang2010}, precision lasso\citep{Wang2018}) and penalty structures (group selection, bi-level selection), selecting functional forms of covariates in unpenalized Cox regression, model selection for unpenalized Cox regression that involve time-varying covariates and longitudinal covariates.
+Since all the cross-validation approaches that are discussed in this manuscript depend on the Cox partial likelihood, they are subject to limitations such as non-proportional hazards and rare events that are tied to the Cox proportional hazard model.
 

### 2022-10-13 supplementary.tex
updates to response, supplementary and main

@@
 \usepackage{bbold}
+\usepackage{pdflscape}
 
@@
 
-\begin{table}[h]
+\begin{table}[!htb]
 \centering
+\caption{\label{Tab:base} Baseline Estimation Approaches for Deviation Residual}
 \begin{tabular}{lrrrrrr}
@@
 
+\begin{landscape}
+
+\section{Simulations with Different Time Points for Brier Score and Kullbeck-Liebler Score}
+
+In the main manuscript, only median survival time was used to calcualate the Brier Score and Kullbeck-Liebler Score.
+Additional simulations are presented here to accomodate more time points.
+More specifically, the $25$ and $75$ percentile of all survival times are selected for each simulated dataset.
+The overall conclusion remains the same when Brier Score and Kullbeck-Liebler Score are evaluated at different time points.
+
+\begin{table}[h]
+
+\setlength{\tabcolsep}{3pt}
+
+\caption{\label{Tab:exp} Exponential Distribution}
+\centering
+\begin{tabular}[t]{lccccccccc}
+\toprule
+ Method & $\lambda$ & log(MSE Ratio) & Brier (25$\%$) & Brier(Median) & Brier(75$\%$) & KL (25$\%$)& KL (Median) & KL (75$\%$) & C Index\\
+\midrule
+&\multicolumn{9}{c}{\textbf{Scenario 1}: n = 150, p = 50, 5 non-zero features, weak signal}\\
+\cline{2-10}
+V VH  & 0.102 (0.030) & 1.537 (0.880) & 0.184 (0.007) & 0.245 (0.014) & 0.183 (0.014) & 0.556 (0.023) & 0.683 (0.030) & 0.547 (0.034) & 0.609 (0.043) \\    
+Standard  & 0.093 (0.033) & 1.537 (0.876) & 0.184 (0.007) & 0.244 (0.014) & 0.183 (0.014) & 0.556 (0.023) & 0.683 (0.031) & 0.546 (0.035) & 0.611 (0.039) \\    
+LinearPred  & 0.095 (0.032) & 1.521 (0.877) & 0.184 (0.007) & 0.244 (0.014) & 0.182 (0.014) & 0.555 (0.023) & 0.682 (0.030) & 0.545 (0.034) & 0.610 (0.041) \\    
+DevResid  & 0.103 (0.029) & 1.534 (0.870) & 0.184 (0.007) & 0.245 (0.014) & 0.183 (0.014) & 0.556 (0.023) & 0.683 (0.029) & 0.546 (0.034) & 0.610 (0.041) \\
+\addlinespace
+&\multicolumn{9}{c}{\textbf{Scenario 2}: n = 150, p = 50, 5 non-zero features, strong signal}\\
+\cline{2-10}
+ V VH  & 0.070 (0.011) & 1.754 (0.882) & 0.147 (0.010) & 0.176 (0.012) & 0.128 (0.009) & 0.454 (0.025) & 0.526 (0.030) & 0.396 (0.027) & 0.750 (0.014) \\  
+ Standard  & 0.066 (0.011) & 1.707 (0.879) & 0.147 (0.009) & 0.176 (0.012) & 0.129 (0.010) & 0.453 (0.024) & 0.525 (0.030) & 0.396 (0.028) & 0.749 (0.014) \\   
+ LinearPred  & 0.066 (0.010) & 1.709 (0.875) & 0.147 (0.009) & 0.176 (0.012) & 0.128 (0.009) & 0.453 (0.024) & 0.525 (0.029) & 0.395 (0.027) & 0.749 (0.014) \\  
+ DevResid  & 0.094 (0.011) & 2.123 (0.804) & 0.151 (0.010) & 0.180 (0.013) & 0.130 (0.010) & 0.463 (0.024) & 0.538 (0.030) & 0.406 (0.028) & 0.751 (0.016) \\
+\addlinespace
+&\multicolumn{9}{c}{\textbf{Scenario 3}: n = 400, p = 10000, 20 non-zero features, weak signal}\\
+\cline{2-10}
+ V VH  & 0.126 (0.019) & 2.927 (0.376) & 0.188 (0.005) & 0.252 (0.010) & 0.191 (0.011) & 0.564 (0.016) & 0.697 (0.022) & 0.569 (0.025) & 0.587 (0.042) \\
+ Standard  & 0.103 (0.021) & 2.846 (0.403) & 0.184 (0.007) & 0.243 (0.013) & 0.182 (0.012) & 0.554 (0.019) & 0.679 (0.028) & 0.547 (0.031) & 0.604 (0.037) \\
+ LinearPred  & 0.105 (0.021) & 2.853 (0.399) & 0.185 (0.006) & 0.244 (0.013) & 0.183 (0.012) & 0.555 (0.019) & 0.681 (0.028) & 0.548 (0.030) & 0.603 (0.038) \\
+ DevResid  & 0.107 (0.020) & 2.863 (0.393) & 0.185 (0.006) & 0.245 (0.012) & 0.183 (0.012) & 0.556 (0.018) & 0.682 (0.026) & 0.550 (0.029) & 0.603 (0.038) \\
+\addlinespace
+&\multicolumn{9}{c}{\textbf{Scenario 4}: n = 400, p = 10000, 20 non-zero features, strong signal}\\
+\cline{2-10}
+ V VH  & 0.083 (0.011) & 3.741 (0.527) & 0.145 (0.017) & 0.173 (0.027) & 0.127 (0.019) & 0.443 (0.046) & 0.52 (0.064) & 0.395 (0.055) & 0.768 (0.043) \\
+ Standard  & 0.060 (0.008) & 3.421 (0.545) & 0.131 (0.014) & 0.154 (0.018) & 0.117 (0.013) & 0.404 (0.037) & 0.467 (0.046) & 0.365 (0.040) & 0.781 (0.027) \\
+ LinearPred  & 0.061 (0.007) & 3.444 (0.538) & 0.131 (0.014) & 0.154 (0.018) & 0.117 (0.013) & 0.405 (0.037) & 0.468 (0.046) & 0.363 (0.039) & 0.781 (0.028) \\ 
+ DevResid  & 0.077 (0.021) & 3.668 (0.512) & 0.140 (0.015) & 0.165 (0.024) & 0.123 (0.019) & 0.429 (0.040) & 0.501 (0.057) & 0.380 (0.053) & 0.772 (0.053) \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\section{Simulations with Weibull and Gompertz Distribution}
+
+In the main manuscript, the baseline hazard for all simulation experiments are generated using the exponential distribution for simplicity.
+More simulation results are presented in the supplement for scenarios where Weibull and Gompertz distributions are used.
+
+For the Weibull distribution, the following parameterization is used for the hazard function:
+$$h(t) = \frac{a}{b}(\frac{t}{b})^{a - 1}\exp{(x^{T}\beta)}, $$
+where $a$ represents the shape parameter and $b$ represents the scale parameter. This is same as the parameterization used in R.
+
+%
+%\begin{table}[h]
+%
+%\setlength{\tabcolsep}{3pt}
+%
+%\caption{\label{Tab:wb1} Weibull Distribution, scale = 1, shape = 0.85}
+%\centering
+%\begin{tabular}[t]{lccccccccc}
+%\toprule
+% Method & $\lambda$ & log(MSE Ratio) & Brier (25$\%$) & Brier(Median) & Brier(75$\%$) & KL (25$\%$)& KL (Median) & KL (75$\%$) & C Index\\
+%\midrule
+%&\multicolumn{9}{c}{\textbf{Scenario 1}: n = 150, p = 50, 5 non-zero features, weak signal}\\
+%\cline{2-10}
+%V VH  & 0.096 (0.026) & 1.502 (0.89) & 0.185 (0.006) & 0.244 (0.015) & 0.182 (0.014) & 0.558 (0.02) & 0.681 (0.032) & 0.545 (0.034) & 0.614 (0.03) \\ 
+%Standard  & 0.089 (0.03) & 1.479 (0.873) & 0.185 (0.007) & 0.243 (0.015) & 0.182 (0.013) & 0.557 (0.021) & 0.681 (0.032) & 0.544 (0.033) & 0.615 (0.031) \\ 
+%LinearPred  & 0.09 (0.027) & 1.482 (0.871) & 0.185 (0.007) & 0.243 (0.014) & 0.182 (0.014) & 0.558 (0.02) & 0.681 (0.031) & 0.544 (0.033) & 0.615 (0.03) \\  
+%DevResid  & 0.098 (0.025) & 1.503 (0.869) & 0.185 (0.006) & 0.244 (0.015) & 0.182 (0.014) & 0.558 (0.02) & 0.681 (0.031) & 0.545 (0.033) & 0.615 (0.03) \\
+%\addlinespace
+%&\multicolumn{9}{c}{\textbf{Scenario 2}: n = 150, p = 50, 5 non-zero features, strong signal}\\
+%\cline{2-10}
+% V VH  & 0.07 (0.011) & 1.733 (0.83) & 0.147 (0.009) & 0.176 (0.012) & 0.128 (0.009) & 0.453 (0.023) & 0.525 (0.028) & 0.395 (0.026) & 0.75 (0.013) \\   
+%Standard  & 0.066 (0.011) & 1.693 (0.832) & 0.147 (0.009) & 0.176 (0.011) & 0.128 (0.009) & 0.452 (0.023) & 0.525 (0.028) & 0.395 (0.026) & 0.749 (0.014) \\ 
+%LinearPred  & 0.067 (0.01) & 1.699 (0.834) & 0.147 (0.009) & 0.176 (0.011) & 0.128 (0.009) & 0.452 (0.023) & 0.525 (0.028) & 0.395 (0.026) & 0.749 (0.014) \\    
+%DevResid  & 0.096 (0.013) & 2.14 (0.799) & 0.151 (0.009) & 0.181 (0.012) & 0.13 (0.009) & 0.463 (0.022) & 0.539 (0.028) & 0.406 (0.026) & 0.751 (0.013) \\
+%\addlinespace
+%&\multicolumn{9}{c}{\textbf{Scenario 3}: n = 400, p = 10000, 20 non-zero features, weak signal}\\
+%\cline{2-10}
+%V VH  & 0.126 (0.019) & 2.948 (0.442) & 0.188 (0.005) & 0.252 (0.011) & 0.191 (0.011) & 0.564 (0.016) & 0.696 (0.023) & 0.569 (0.026) & 0.585 (0.043) \\   
+%Standard  & 0.103 (0.024) & 2.866 (0.465) & 0.184 (0.006) & 0.243 (0.014) & 0.183 (0.012) & 0.555 (0.019) & 0.68 (0.029) & 0.548 (0.03) & 0.6 (0.041) \\  
+%LinearPred  & 0.104 (0.024) & 2.869 (0.464) & 0.185 (0.006) & 0.244 (0.014) & 0.183 (0.012) & 0.556 (0.019) & 0.68 (0.029) & 0.549 (0.03) & 0.6 (0.041) \\   
+%DevResid  & 0.107 (0.023) & 2.88 (0.458) & 0.185 (0.006) & 0.244 (0.013) & 0.184 (0.012) & 0.556 (0.019) & 0.682 (0.028) & 0.55 (0.03) & 0.599 (0.041) \\
+%\addlinespace
+%&\multicolumn{9}{c}{\textbf{Scenario 4}: n = 400, p = 10000, 20 non-zero features, strong signal}\\
+%\cline{2-10}
+% V VH  & 0.082 (0.012) & 3.742 (0.551) & 0.146 (0.02) & 0.173 (0.029) & 0.126 (0.02) & 0.447 (0.051) & 0.519 (0.068) & 0.393 (0.058) & 0.766 (0.047) \\  
+%Standard  & 0.059 (0.009) & 3.417 (0.58) & 0.133 (0.015) & 0.154 (0.02) & 0.117 (0.014) & 0.409 (0.041) & 0.467 (0.05) & 0.366 (0.044) & 0.78 (0.03) \\   
+%LinearPred  & 0.06 (0.009) & 3.437 (0.581) & 0.133 (0.015) & 0.154 (0.02) & 0.117 (0.014) & 0.41 (0.042) & 0.468 (0.051) & 0.365 (0.043) & 0.78 (0.031) \\
+%DevResid  & 0.084 (0.034) & 3.704 (0.528) & 0.144 (0.02) & 0.17 (0.033) & 0.126 (0.026) & 0.441 (0.053) & 0.511 (0.075) & 0.39 (0.071) & 0.758 (0.079) \\
+%\bottomrule
+%\end{tabular}
+%\end{table}
+%
+
+\begin{table}[h]
+\setlength{\tabcolsep}{3pt}
+\caption{\label{Tab:wb2} Weibull Distribution, scale = 1, shape = 1.5}
+\centering
+\begin{tabular}[t]{lccccccccc}
+\toprule
+ Method & $\lambda$ & log(MSE Ratio) & Brier (25$\%$) & Brier(Median) & Brier(75$\%$) & KL (25$\%$)& KL (Median) & KL (75$\%$) & C Index\\
+\midrule
+&\multicolumn{9}{c}{\textbf{Scenario 1}: n = 150, p = 50, 5 non-zero features, weak signal}\\
+\cline{2-10}
+V VH  & 0.096 (0.026) & 1.502 (0.89) & 0.185 (0.006) & 0.244 (0.015) & 0.182 (0.014) & 0.558 (0.02) & 0.681 (0.032) & 0.545 (0.034) & 0.614 (0.03) \\
+Standard  & 0.089 (0.03) & 1.479 (0.873) & 0.184 (0.007) & 0.243 (0.015) & 0.182 (0.013) & 0.557 (0.021) & 0.681 (0.032) & 0.544 (0.033) & 0.615 (0.031) \\
+LinearPred  & 0.09 (0.027) & 1.482 (0.871) & 0.185 (0.007) & 0.243 (0.014) & 0.182 (0.014) & 0.558 (0.02) & 0.681 (0.031) & 0.544 (0.033) & 0.615 (0.03) \\
+DevResid  & 0.098 (0.025) & 1.503 (0.869) & 0.185 (0.006) & 0.244 (0.015) & 0.182 (0.014) & 0.558 (0.02) & 0.681 (0.031) & 0.545 (0.033) & 0.615 (0.03) \\
+\addlinespace
+&\multicolumn{9}{c}{\textbf{Scenario 2}: n = 150, p = 50, 5 non-zero features, strong signal}\\
+\cline{2-10}
+ V VH  & 0.07 (0.011) & 1.733 (0.83) & 0.147 (0.009) & 0.176 (0.012) & 0.128 (0.009) & 0.453 (0.023) & 0.525 (0.028) & 0.395 (0.026) & 0.75 (0.013) \\ 
+Standard  & 0.066 (0.011) & 1.693 (0.832) & 0.147 (0.009) & 0.176 (0.011) & 0.128 (0.009) & 0.452 (0.023) & 0.525 (0.028) & 0.395 (0.026) & 0.749 (0.014) \\
+ LinearPred  & 0.067 (0.01) & 1.699 (0.834) & 0.147 (0.009) & 0.176 (0.011) & 0.128 (0.009) & 0.452 (0.023) & 0.525 (0.028) & 0.395 (0.026) & 0.749 (0.014) \\
+ DevResid  & 0.096 (0.013) & 2.14 (0.799) & 0.151 (0.009) & 0.181 (0.012) & 0.13 (0.009) & 0.463 (0.022) & 0.539 (0.028) & 0.406 (0.026) & 0.751 (0.013) \\
+\addlinespace
+&\multicolumn{9}{c}{\textbf{Scenario 3}: n = 400, p = 10000, 20 non-zero features, weak signal}\\
+\cline{2-10}
+V VH  & 0.126 (0.019) & 2.948 (0.442) & 0.188 (0.005) & 0.252 (0.011) & 0.191 (0.011) & 0.564 (0.016) & 0.696 (0.023) & 0.569 (0.026) & 0.585 (0.043) \\ 
+Standard  & 0.103 (0.024) & 2.866 (0.465) & 0.184 (0.006) & 0.243 (0.014) & 0.183 (0.012) & 0.555 (0.019) & 0.68 (0.029) & 0.548 (0.03) & 0.6 (0.041) \\ 
+LinearPred  & 0.104 (0.024) & 2.869 (0.464) & 0.185 (0.006) & 0.244 (0.014) & 0.183 (0.012) & 0.556 (0.019) & 0.68 (0.029) & 0.549 (0.03) & 0.6 (0.041) \\   
+ DevResid  & 0.107 (0.023) & 2.88 (0.458) & 0.185 (0.006) & 0.244 (0.013) & 0.184 (0.012) & 0.556 (0.019) & 0.682 (0.028) & 0.55 (0.03) & 0.599 (0.041) \\
+\addlinespace
+&\multicolumn{9}{c}{\textbf{Scenario 4}: n = 400, p = 10000, 20 non-zero features, strong signal}\\
+\cline{2-10}
+ V VH  & 0.082 (0.012) & 3.74 (0.554) & 0.145 (0.019) & 0.173 (0.029) & 0.127 (0.02) & 0.443 (0.05) & 0.519 (0.068) & 0.395 (0.058) & 0.766 (0.047) \\  
+Standard  & 0.059 (0.009) & 3.414 (0.582) & 0.131 (0.015) & 0.154 (0.02) & 0.117 (0.014) & 0.405 (0.041) & 0.467 (0.05) & 0.367 (0.044) & 0.78 (0.03) \\   
+LinearPred  & 0.06 (0.009) & 3.435 (0.583) & 0.132 (0.015) & 0.154 (0.02) & 0.117 (0.014) & 0.406 (0.041) & 0.468 (0.051) & 0.366 (0.043) & 0.78 (0.031) \\   
+ DevResid  & 0.074 (0.008) & 3.66 (0.51) & 0.139 (0.014) & 0.163 (0.021) & 0.121 (0.015) & 0.426 (0.038) & 0.494 (0.049) & 0.375 (0.042) & 0.777 (0.036) \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+For the Gompertz distribution, the following parameterization is used for the hazard function:
+$$h(t) = a e^{b t}e^{x^{T}\beta}, $$
+where $a$ represents the scale parameter and $b$ represents the rate parameter.
+
+The overall conclusions from the simulation experiments remain the same.
+This is no surprise as both Weibull and Gompertz distribution follow the proportional hazard assumption, which is the underlying assumption for Cox Regression.
+
+\begin{table}[h]
+
+\setlength{\tabcolsep}{3pt}
+
+\caption{\label{Tab:sim} Gompertz Distribution, rate = 2, shape = 1}
+\centering
+\begin{tabular}[t]{lccccccccc}
+\toprule
+ Method & $\lambda$ & log(MSE Ratio) & Brier (25$\%$) & Brier(Median) & Brier(75$\%$) & KL (25$\%$)& KL (Median) & KL (75$\%$) & C Index\\
+\midrule
+&\multicolumn{9}{c}{\textbf{Scenario 1}: n = 150, p = 50, 5 non-zero features, weak signal}\\
+\cline{2-10}
+V VH  & 0.095 (0.028) & 1.464 (0.831) & 0.185 (0.007) & 0.242 (0.015) & 0.183 (0.014) & 0.558 (0.021) & 0.679 (0.032) & 0.546 (0.034) & 0.612 (0.036) \\
+Standard  & 0.089 (0.03) & 1.472 (0.817) & 0.184 (0.007) & 0.242 (0.015) & 0.183 (0.014) & 0.558 (0.022) & 0.679 (0.033) & 0.547 (0.035) & 0.612 (0.036) \\
+LinearPred  & 0.088 (0.026) & 1.444 (0.819) & 0.184 (0.007) & 0.242 (0.015) & 0.182 (0.013) & 0.557 (0.022) & 0.678 (0.032) & 0.545 (0.034) & 0.614 (0.034) \\
+DevResid  & 0.097 (0.026) & 1.449 (0.83) & 0.185 (0.007) & 0.242 (0.014) & 0.182 (0.013) & 0.557 (0.021) & 0.678 (0.031) & 0.545 (0.032) & 0.613 (0.035) \\
+\addlinespace
+&\multicolumn{9}{c}{\textbf{Scenario 2}: n = 150, p = 50, 5 non-zero features, strong signal}\\
+\cline{2-10}
+V VH  & 0.071 (0.011) & 1.68 (0.895) & 0.147 (0.009) & 0.177 (0.013) & 0.129 (0.01) & 0.452 (0.024) & 0.527 (0.032) & 0.398 (0.029) & 0.749 (0.015) \\ 
+Standard  & 0.066 (0.011) & 1.65 (0.865) & 0.146 (0.009) & 0.177 (0.012) & 0.129 (0.01) & 0.451 (0.023) & 0.526 (0.031) & 0.398 (0.028) & 0.748 (0.015) \\ 
+LinearPred  & 0.067 (0.011) & 1.65 (0.875) & 0.146 (0.009) & 0.177 (0.013) & 0.129 (0.01) & 0.451 (0.024) & 0.526 (0.032) & 0.398 (0.03) & 0.748 (0.015) \\
+ DevResid  & 0.095 (0.012) & 2.049 (0.817) & 0.151 (0.009) & 0.181 (0.014) & 0.131 (0.01) & 0.461 (0.022) & 0.539 (0.031) & 0.407 (0.03) & 0.75 (0.016) \\
+\addlinespace
+&\multicolumn{9}{c}{\textbf{Scenario 3}: n = 400, p = 10000, 20 non-zero features, weak signal}\\
+\cline{2-10}
+V VH  & 0.127 (0.019) & 2.973 (0.441) & 0.188 (0.005) & 0.252 (0.011) & 0.191 (0.012) & 0.563 (0.015) & 0.697 (0.023) & 0.569 (0.028) & 0.585 (0.043) \\ 
+Standard  & 0.109 (0.025) & 2.91 (0.469) & 0.185 (0.007) & 0.245 (0.014) & 0.185 (0.013) & 0.556 (0.019) & 0.683 (0.029) & 0.553 (0.034) & 0.598 (0.041) \\
+LinearPred  & 0.108 (0.025) & 2.909 (0.467) & 0.185 (0.007) & 0.245 (0.014) & 0.185 (0.013) & 0.556 (0.019) & 0.683 (0.029) & 0.553 (0.034) & 0.598 (0.041) \\  
+DevResid  & 0.111 (0.024) & 2.919 (0.464) & 0.185 (0.006) & 0.246 (0.014) & 0.185 (0.013) & 0.557 (0.019) & 0.685 (0.028) & 0.554 (0.033) & 0.596 (0.041) \\
+\addlinespace
+&\multicolumn{9}{c}{\textbf{Scenario 4}: n = 400, p = 10000, 20 non-zero features, strong signal}\\
+\cline{2-10}
+V VH  & 0.082 (0.011) & 3.78 (0.559) & 0.144 (0.018) & 0.173 (0.027) & 0.127 (0.019) & 0.442 (0.048) & 0.519 (0.064) & 0.395 (0.055) & 0.769 (0.042) \\ 
+Standard  & 0.059 (0.007) & 3.447 (0.589) & 0.13 (0.014) & 0.152 (0.017) & 0.116 (0.012) & 0.401 (0.038) & 0.464 (0.044) & 0.364 (0.037) & 0.783 (0.026) \\    
+ LinearPred  & 0.061 (0.007) & 3.473 (0.583) & 0.131 (0.014) & 0.153 (0.017) & 0.116 (0.012) & 0.403 (0.039) & 0.465 (0.045) & 0.363 (0.037) & 0.783 (0.027) \\ 
+ DevResid  & 0.075 (0.013) & 3.697 (0.535) & 0.138 (0.014) & 0.163 (0.019) & 0.121 (0.014) & 0.425 (0.036) & 0.495 (0.045) & 0.376 (0.039) & 0.778 (0.036) \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+%\begin{table}[!htb]
+%
+%\setlength{\tabcolsep}{3pt}
+%
+%\caption{\label{Tab:sim} Gompertz Distribution, scale = 1, shape = 1.5}
+%\centering
+%\begin{tabular}[t]{lccccccccc}
+%\toprule
+% Method & $\lambda$ & log(MSE Ratio) & Brier (25$\%$) & Brier(Median) & Brier(75$\%$) & KL (25$\%$)& KL (Median) & KL (75$\%$) & C Index\\
+%\midrule
+%&\multicolumn{9}{c}{\textbf{Scenario 1}: n = 150, p = 50, 5 non-zero features, weak signal}\\
+%\cline{2-10}
+% V VH  & 0.095 (0.028) & 1.464 (0.831) & 0.185 (0.007) & 0.242 (0.015) & 0.183 (0.014) & 0.558 (0.021) & 0.679 (0.032) & 0.546 (0.034) & 0.612 (0.036) \\ 
+% Standard  & 0.089 (0.03) & 1.472 (0.817) & 0.184 (0.007) & 0.242 (0.015) & 0.183 (0.014) & 0.558 (0.022) & 0.679 (0.033) & 0.547 (0.035) & 0.612 (0.036) \\ 
+% LinearPred  & 0.088 (0.026) & 1.444 (0.819) & 0.184 (0.007) & 0.242 (0.015) & 0.182 (0.013) & 0.557 (0.022) & 0.678 (0.032) & 0.545 (0.034) & 0.614 (0.034) \\   
+% DevResid  & 0.097 (0.026) & 1.449 (0.83) & 0.185 (0.007) & 0.242 (0.014) & 0.182 (0.013) & 0.557 (0.021) & 0.678 (0.031) & 0.545 (0.032) & 0.613 (0.035) \\
+%\addlinespace
+%&\multicolumn{9}{c}{\textbf{Scenario 2}: n = 150, p = 50, 5 non-zero features, strong signal}\\
+%\cline{2-10}
+% V VH  & 0.071 (0.011) & 1.68 (0.895) & 0.147 (0.009) & 0.177 (0.013) & 0.129 (0.01) & 0.452 (0.024) & 0.527 (0.032) & 0.398 (0.029) & 0.749 (0.015) \\ 
+%Standard  & 0.066 (0.011) & 1.65 (0.865) & 0.146 (0.009) & 0.177 (0.012) & 0.129 (0.01) & 0.451 (0.023) & 0.526 (0.031) & 0.398 (0.028) & 0.748 (0.015) \\ 
+% LinearPred  & 0.067 (0.011) & 1.65 (0.875) & 0.146 (0.009) & 0.177 (0.013) & 0.129 (0.01) & 0.451 (0.024) & 0.526 (0.032) & 0.398 (0.03) & 0.748 (0.015) \\   
+%DevResid  & 0.095 (0.012) & 2.049 (0.817) & 0.151 (0.009) & 0.181 (0.014) & 0.131 (0.01) & 0.461 (0.022) & 0.539 (0.031) & 0.407 (0.03) & 0.75 (0.016) \\
+%\addlinespace
+%&\multicolumn{9}{c}{\textbf{Scenario 3}: n = 400, p = 10000, 20 non-zero features, weak signal}\\
+%\cline{2-10}
+% V VH  & 0.127 (0.019) & 2.973 (0.441) & 0.188 (0.005) & 0.252 (0.011) & 0.191 (0.012) & 0.563 (0.015) & 0.697 (0.023) & 0.569 (0.028) & 0.585 (0.043) \\  
+%Standard  & 0.109 (0.025) & 2.91 (0.469) & 0.185 (0.007) & 0.245 (0.014) & 0.185 (0.013) & 0.556 (0.019) & 0.683 (0.029) & 0.553 (0.034) & 0.598 (0.041) \\  
+% LinearPred  & 0.108 (0.025) & 2.909 (0.467) & 0.185 (0.007) & 0.245 (0.014) & 0.185 (0.013) & 0.556 (0.019) & 0.683 (0.029) & 0.553 (0.034) & 0.598 (0.041) \\  
+% DevResid  & 0.111 (0.024) & 2.919 (0.464) & 0.185 (0.006) & 0.246 (0.014) & 0.185 (0.013) & 0.557 (0.019) & 0.685 (0.028) & 0.554 (0.033) & 0.596 (0.041) \\
+%\addlinespace
+%&\multicolumn{9}{c}{\textbf{Scenario 4}: n = 400, p = 10000, 20 non-zero features, strong signal}\\
+%\cline{2-10}
+% V VH  & 0.082 (0.011) & 3.78 (0.559) & 0.144 (0.018) & 0.173 (0.027) & 0.127 (0.019) & 0.442 (0.048) & 0.519 (0.064) & 0.395 (0.055) & 0.769 (0.042) \\  
+%Standard  & 0.059 (0.007) & 3.447 (0.589) & 0.13 (0.014) & 0.152 (0.017) & 0.116 (0.012) & 0.401 (0.038) & 0.464 (0.044) & 0.364 (0.037) & 0.783 (0.026) \\ 
+%LinearPred  & 0.061 (0.007) & 3.473 (0.583) & 0.131 (0.014) & 0.153 (0.017) & 0.116 (0.012) & 0.403 (0.039) & 0.465 (0.045) & 0.363 (0.037) & 0.783 (0.027) \\  
+%DevResid  & 0.075 (0.013) & 3.697 (0.535) & 0.138 (0.014) & 0.163 (0.019) & 0.121 (0.014) & 0.425 (0.036) & 0.495 (0.045) & 0.376 (0.039) & 0.778 (0.036) \\
+%\bottomrule
+%\end{tabular}
+%\end{table}
+
+\end{landscape}
+
 \bibliographystyle{ims-nourl}

### 2022-10-27 cross-validation.tex
orgnize citation
Submitted to Statistics in Medicine
Submitted to arXiv
Edits to abstract and intro

@@
 \usepackage{alltt}
-\usepackage{amssymb,amsmath,natbib,graphicx,enumerate,subcaption,tikz,url,booktabs}
+\usepackage{amssymb,amsmath,graphicx,enumerate,subcaption,tikz,url,booktabs}
 \usepackage[bmargin=0.75in, tmargin =0.75in,lmargin = 0.5in,rmargin = 0.5in]{geometry}
+\graphicspath{{manuscript_figure/}}
+\usepackage{hyperref}
+\hypersetup{colorlinks=true, citecolor=black, urlcolor=cyan}
+\usepackage{afterpage}
 
-\usepackage[ruled]{algorithm2e}
-\usepackage{algpseudocode}
-\renewcommand{\algorithmcfname}{Procedure}
+\usepackage{natbib}
 \usepackage{bbold}
@@
 \newcommand{\beps}{\boldsymbol\epsilon}
+\newcommand{\boldeta}{\boldsymbol\eta}
 
@@
 \section*{Abstract}
-Cross validation is commonly used for selecting tuning parameters in penalized regression, but its use in penalized Cox regression models has received relatively little attention in the literature.
+Cross validation is the most common way of selecting tuning parameters in penalized regression, but its use in penalized Cox regression models has received relatively little attention in the literature.
 Due to its partial likelihood construction, carrying out cross validation for Cox models is not straightforward, and there are several potential approaches for implementation.
-Here, we propose two new cross-validation methods for Cox regression and compare them to approaches that have been proposed elsewhere.
-Our proposed approach of cross-validating the linear predictors seems to offer an attractive balance of performance and numerical stability.
-We illustrate these advantages using simulated data as well as using them to analyze data from a high-dimensional study of survival in lung cancer patients.
+Here, we propose a new approach based on cross-validating the linear predictors of the Cox model and compare it to approaches that have been proposed elsewhere.
+We show that the proposed approach offers an attractive balance of performance and numerical stability, and illustrate these advantages using simulated data as well as analyzing a high-dimensional study of gene expression and survival in lung cancer patients.
 \input{main}
 
-\bibliographystyle{abbrvnat}
-%\setcitestyle{round} 
+\bibliographystyle{ims-nourl}
 \bibliography{articles}

### 2022-10-27 main.tex
Edits to abstract and intro

@@
 
-\par Since its original proposal \citep{Cox1972}, Cox proportional hazards regression has become the most common regression approach for analyzing survival data.
+Since its original proposal \citep{Cox1972}, Cox proportional hazards regression has become the most common regression approach for analyzing survival data.
 Cox regression utilizes a partial likelihood construction under an assumption of proportional hazards to estimate the regression coefficients without having to specify the underlying baseline hazard.
@@
 
-\par Standard cross-validation involves dividing the data into $K$ folds, then fitting the model on $K-1$ of those folds (the training set) and assessing prediction accuracy on the remaining fold (the testing set).
+Standard cross-validation involves dividing the data into $K$ folds, then fitting the model on $K-1$ of those folds (the training set) and assessing prediction accuracy on the remaining fold (the testing set).
 This process is then repeated, with each fold serving as the testing set exactly once.
@@
 
-\par One basic approach is to simply calculate the partial likelihood based on the observations in the test set as a measure for the model's predictive accuracy.
+One basic approach is to simply calculate the partial likelihood based on the observations in the test set as a measure for the model's predictive accuracy.
 In doing so, we calculate the risk for each member of the test set relative to other members of the test set.
@@
 
-\par To overcome this drawback, an alternative approach was proposed by \citet{Verweij1993}.
+To overcome this drawback, an alternative approach was proposed by \citet{Verweij1993}.
 Their approach, which we describe in detail in Section~\ref{Sec:cox-cv-existing}, stabilizes the cross-validated log likelihood, enabling its use even when the number of subjects in each fold is small.
 This approach has been widely used and is the default approach to cross-validation in many software programs, such as the R package {\tt glmnet} \citep{glmnet}.
-Although this approach fixes the stability issue, we demonstrate here that in practice, it tends to behave conservatively in terms of model selection.
+Although this approach fixes the stability issue, we demonstrate here that in practice, it tends to behave conservatively in terms of model selection (throughout, we use the term ``conservative'' to mean a tendency to select a model with few predictors, or even the null model).
 
-\par In this paper, we propose two alternative ways to carry out cross validation for Cox regression.
-Instead of applying cross-validation to the partial likelihood directly, we propose applying cross validation to either the linear predictors of the regression model or to the deviance residuals \citep{Therneau1990}.
+Another approach, more common in the random survival forest literature than the penalized regression literature, is to calculate deviance residuals \citep{Therneau1990}.
+As noted earlier, if we do not estimate the baseline hazard, then we cannot predict absolute failure time.
+However, if we combine Cox regression with an estimate of the baseline hazard, such as the Nelson-Aalen estimator \citep{nelson1969, aalen1978}, then we can obtain residuals based on deviance.
+These residuals can then be used like conventional residuals to estimate cross-validation error.
+However, we find this approach also to be rather conservative.
+
+In this paper, we propose an alternative way to carry out cross validation for Cox regression.
+Instead of applying cross-validation to the partial likelihood directly, we propose applying cross validation to the linear predictors of the regression model.
 Through simulation studies, we compare these proposed methods with existing cross-validation approaches for LASSO penalized Cox regression in both low- and high-dimensional settings.
-The linear predictor approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.
-We conclude by applying both proposed and existing approaches to a high-dimensional study of survival in lung cancer patients.
+We find that the linear predictor approach avoids the conservatism of the deviance residual and \citet{Verweij1993} approaches as well as the numerical instability of ordinary cross-validation.
+The proposed approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.
+We conclude by applying both proposed and existing approaches to a high-dimensional study of gene expression and survival in lung cancer patients.
 
@@
 
-\par In the Cox model, the hazard function for subject $i$ is given by 
+In the Cox model, the hazard function for subject $i$ is given by 
 \begin{equation*}
@@
 
-\par In this paper, we focus upon the LASSO penalty $P_{\lambda}(\bbeta) = \lambda \sum_{j} |\beta_{j}|$, although the methods that we analyze not only can be used for penalized Cox regression with any penalty, such as elastic net \citep{Zou2005}, smoothly clipped absolute deviations (SCAD) penalty \citep{FanLi2001}, and minimax concave penalty (MCP) \citep{Zhang2010}, but are also applicable to model selection in unpenalized Cox regression.
+In this paper, we focus upon the LASSO penalty $P_{\lambda}(\bbeta) = \lambda \sum_{j} |\beta_{j}|$, although the methods that we analyze not only can be used for penalized Cox regression with any penalty, such as elastic net \citep{Zou2005}, smoothly clipped absolute deviations (SCAD) penalty \citep{FanLi2001}, and minimax concave penalty (MCP) \citep{Zhang2010}, but are also applicable to model selection in unpenalized Cox regression.
 LASSO-penalized Cox regression is particularly useful for high dimensional data where the number of covariates $p > n$, such as predicting overall survival in cancer patients based on genome-wide expression measurements.
@@
 
-\par Cross-validation is the most common method for selecting $\lambda$ in penalized regression models.
+Cross-validation is the most common method for selecting $\lambda$ in penalized regression models.
 Suppose a data set $D$ of n observations is partitioned into $K$ folds: $D_{1}, D_{2}, \ldots, D_{K}$.
@@
 
-\par A direct approach to measure a model's predictive accuracy in $D_k$ is to evaluate the log partial likelihood at $\hat{\bbeta}^{-k}$ using the data in $D_k$.
+A direct approach to measure a model's predictive accuracy in $D_k$ is to evaluate the log partial likelihood at $\hat{\bbeta}^{-k}$ using the data in $D_k$.
 Cross-validation error (CVE) is then measured by the total deviance across all $K$ folds, which (up to a constant) is given by:
@@
 
-\par The basic approach is by far the most common way of conducting cross-validation for models such as linear regression and logistic regression.
+The basic approach is by far the most common way of conducting cross-validation for models such as linear regression and logistic regression.
 However, the basic approach can be problematic for Cox regression in the sense that that there may not be enough observations in $D_k$ to build up the risk set for partial likelihood.
@@
 
-\par Both of the methods in Section~\ref{Sec:cox-cv-existing} consist of adding together log partial likelihood measures from each fold.
+Both of the methods in Section~\ref{Sec:cox-cv-existing} consist of adding together log partial likelihood measures from each fold.
 An alternative method is to obtain linear predictors for each fold, then combine these linear predictors to calculate a partial likelihood.
@@
 
-\par This idea of obtaining linear predictors using test sets, then constructing the partial likelihood after pooling all linear predictors together is implemented in R package \texttt{ncvreg} \citep{ncvreg}.
+This idea of obtaining linear predictors using test sets, then constructing the partial likelihood after pooling all linear predictors together is implemented in R package \texttt{ncvreg} \citep{ncvreg}.
 This idea has also appeared in the cross-validation literature for quantities that cannot be evaluated on only a subset of the data, such as AUC in logistic and Cox regression models \citep{Parker2007,Simon2011a,Subramanian2011}.
@@
 
-\par We also measure the predictive accuracy via the mean squared prediction error, or Brier score \citep{VanHouwelingen2011}.
+We also measure the predictive accuracy via the mean squared prediction error, or Brier score \citep{VanHouwelingen2011}.
 Let $\hat{S}(t_0|\lambda,x)$ denote the predicted probability for an individual with covariates $x$ to survive beyond time $t_0$, and let $y(t_0) = \mathbb{1}\left\{ t > t_{0}\right\}$ indicate whether the individual actually survived beyond time $t_0$.
@@
 
-\par The third measure we used is the Kullback-Leibler score, which measures the log-likelihood of the prediction model evaluated at the observations \citep{VanHouwelingen2011}.
+The third measure we used is the Kullback-Leibler score, which measures the log-likelihood of the prediction model evaluated at the observations \citep{VanHouwelingen2011}.
 With the same notation defined in the previous paragraph, the Kullback-Leibler score is defined to be
@@
 
-\par The final metric we included in our study was Harrell's C Index \citep{HarrellJr1984}.
+The final metric we included in our study was Harrell's C Index \citep{HarrellJr1984}.
 The C Index is a rank-based statistic that measures the concordance between the linear predictor of the selected model and the observed outcome.
@@
 
-\par All of the cross-validation approaches described in Section~\ref{Sec:methods} use the Cox partial likelihood, in one way or another, as the criterion upon which model selection is based.
+All of the cross-validation approaches described in Section~\ref{Sec:methods} use the Cox partial likelihood, in one way or another, as the criterion upon which model selection is based.
 In this section, we compare these likelihood-based approaches against an alternative approach of cross validation using concordance, also known as the cross-validated C Index.
@@
 
-\par We conducted simulations in both low dimensional ($p = 50$, $n = 150$) and high dimensional ($p = 1000$, $n = 120$) settings.
+We conducted simulations in both low dimensional ($p = 50$, $n = 150$) and high dimensional ($p = 1000$, $n = 120$) settings.
 For each scenario, we generated data from a spectrum of signal strength as in Section~\ref{Sec:accuracy}.
@@
 
-%\par While the Deviance Residual approach has a reasonable performance in scenarios where dimension is high, it tends to act more conservatively than all the other three approaches in general.
+%While the Deviance Residual approach has a reasonable performance in scenarios where dimension is high, it tends to act more conservatively than all the other three approaches in general.
 As is noted by Therneau et al, the sums of squared martingale residuals and deviance residuals do not necessary reflect how good the model fit is \citep{Therneau2000modeling}.
@@
 
-%\par In the following simulation studies, the defined optimal $\lambda$ and the $\lambda$ selected by the Linear Predictor approach were used as benchmarks.
+%In the following simulation studies, the defined optimal $\lambda$ and the $\lambda$ selected by the Linear Predictor approach were used as benchmarks.
 Various sample sizes and dimensions were varied.
@@
 
-%\par Since both of the non-parameteric estimations seemed to perform well at early time points but poorly later, a weighted approach was then applied to emphasize earlier time points.
+%Since both of the non-parameteric estimations seemed to perform well at early time points but poorly later, a weighted approach was then applied to emphasize earlier time points.
 Weights are assigned proportional to number of subjects at risk, making earlier time points weighted more than later time points.
@@
 
-%\par In the next step,  the non-parametric baseline hazard estimation was replaced by the exact baseline hazard that was used to generate the outcome.
+%In the next step,  the non-parametric baseline hazard estimation was replaced by the exact baseline hazard that was used to generate the outcome.
 Results were listed in the seventh column in Table 2.
@@
 
-%\par In the last step, we replaced the non-parametric baseline hazard estimation with a parametric approach, where we assume the baseline hazard to follow an exponential distribution.
+%In the last step, we replaced the non-parametric baseline hazard estimation with a parametric approach, where we assume the baseline hazard to follow an exponential distribution.
 We derived the estimation for baseline hazard by maximum likelihood approach conditioned on covariates: $\hat{\lambda}_{0} = \frac{\sum d_{i}}{\sum t_{i}e^{\eta_{i}}}$.
@@
 
-%\par Hence, getting accurate estimation of the baseline hazard is crucial to whether or not the sum of squared deviance residuals can be a good measurement for model fit.
+%Hence, getting accurate estimation of the baseline hazard is crucial to whether or not the sum of squared deviance residuals can be a good measurement for model fit.
 
@@
 %	\caption{Standard Method}
-%	\par \For{k in 1 to K } 
+%	\For{k in 1 to K } 
 %	{
@@
 %	
-%	\par $ CVE_{st} = - \sum_{k = 1}^{K} l^{k}(\hat{\beta}^{-k})$
+%	$ CVE_{st} = - \sum_{k = 1}^{K} l^{k}(\hat{\beta}^{-k})$
 % 	\end{algorithm}
@@
 %  	 \caption{Verweij and Van Houwelingen's Method}
-%	\par \For{k in 1 to K } 
+%	\For{k in 1 to K } 
 %	{
@@
 %
-%	\par $ CVE_{vvh} = - \sum_{k = 1}^{K} \left\{ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) \right\}$
+%	$ CVE_{vvh} = - \sum_{k = 1}^{K} \left\{ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) \right\}$
 %	\end{algorithm}
@@
 %    	\caption{Cross-Validated Linear Predictors}
-%	\par \For{k in 1 to K } 
+%	\For{k in 1 to K } 
 %	{
@@
 %
-%	\par $ CVE_{lp} = - l(\hat{\eta}^{-})$
+%	$ CVE_{lp} = - l(\hat{\eta}^{-})$
 %	\end{algorithm}
@@
 %   	\caption{Sums of Squared Deviance Residuals}
-%	\par \For{k in 1 to K } 
+%	\For{k in 1 to K } 
 %	{
@@
 %		\par\hspace{0.5cm} $d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}$;
-%	\par $ CVE_{dr} = \sum_{i = 1}^{n} {d}_{i}^2$ \;
+%	$ CVE_{dr} = \sum_{i = 1}^{n} {d}_{i}^2$ \;
 %	\end{algorithm}

### 2022-10-27 sub/smimr/1/smimr.tex
updates to response, supplementary and main
Edits to abstract and intro

--- 2019-11-26 sub/smimr/coxcv.tex
@@
 \section*{Abstract}
-Cross validation is commonly used for selecting tuning parameters in penalized regression, but its use in penalized Cox regression models has received relatively little attention in the literature.
+Cross validation is the most common way of selecting tuning parameters in penalized regression, but its use in penalized Cox regression models has received relatively little attention in the literature.
 Due to its partial likelihood construction, carrying out cross validation for Cox models is not straightforward, and there are several potential approaches for implementation.
-Here, we propose two new cross-validation methods for Cox regression and compare them to approaches that have been proposed elsewhere.
-Our proposed approach of cross-validating the linear predictors seems to offer an attractive balance of performance and numerical stability.
-We illustrate these advantages using simulated data as well as using them to analyze data from a high-dimensional study of survival in lung cancer patients.
+Here, we propose a new approach based on cross-validating the linear predictors of the Cox model and compare it to approaches that have been proposed elsewhere.
+We show that the proposed approach offers an attractive balance of performance and numerical stability, and illustrate these advantages using simulated data as well as analyzing a high-dimensional study of gene expression and survival in lung cancer patients.
 \input{main}
 
-\bibliographystyle{SageH}
+\bibliographystyle{SageV}
 \bibliography{articles}

### 2022-10-27 main.tex
Edits to Section 2

@@
 
-Another approach, more common in the random survival forest literature than the penalized regression literature, is to calculate deviance residuals \citep{Therneau1990}.
+Another approach, more common in the random survival forest literature than the penalized regression literature, is to calculate deviance residuals.
 As noted earlier, if we do not estimate the baseline hazard, then we cannot predict absolute failure time.
@@
 %%%%%%%%%%%%%%%%%%%%
+\subsection{Cross validated deviance residuals}
+%%%%%%%%%%%%%%%%%%%%
+
+The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model, which prevents us from calculating prediction errors (residuals).
+However, if one takes additional steps to estimate the baseline hazard in addition to estimating the regression coefficients, then such residuals can be calculated.
+The deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}, is a natural candidate for quantifying prediction error to use in quantifying the cross-validation error.
+Originally proposed for model diagnostics, deviance residuals have also been used as a predictive accuracy measure, particularly in the random survival forest literature \citep{LeBlanc1992, Ishwaran2004}.
+In the Cox model, the Martingale residuals $\{m_i\}_{i=1}^n$ and deviance residuals $\{d_i\}_{i=1}^n$ are given by
+\begin{align}
+  \label{eq:res-m}
+  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(X_i\Tr\bbeta), \\
+  \label{eq:res-d}
+  d_{i} &= \text{sign}(m_{i})\sqrt{-2(m_i + \delta_{i}\log(\delta_{i} - m_i))}.  
+\end{align}
+where $H_0$ is the cumulative baseline hazard.
+
+In order to obtain a cross-validated version of these residuals, we would replace $\bbeta$ in the above equations with $\hat{\bbeta}^{-k}$.
+Typically the Nelson-Aalen estimator is used to estimate $H_0$, although other approaches are also possible, such as the Breslow and Kalbfleisch-Prentice estimators \citep{Breslow1972, Kalbfleisch2011}.
+These are described in greater detail in the supplementary materials, along with the results of a simulation study comparing them.
+The Nelson-Aalen estimate of the baseline hazard performed best in these simulations, and also seems to be the most common choice in the literature, so we use this estimate to construct cross-validated deviance residuals throughout the rest of the paper.
+
+The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross validated error, analogous to using the residual sum of squares in linear regression.
+We refer to this method as the \emph{cross-validated deviance residuals} approach.
+
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Cross validated linear predictors}
@@
 
-Both of the methods in Section~\ref{Sec:cox-cv-existing} consist of adding together log partial likelihood measures from each fold.
+Both of the cross-validation approaches in Section~\ref{Sec:cox-cv-existing} consist of adding together log partial likelihood measures from each fold.
 An alternative method is to obtain linear predictors for each fold, then combine these linear predictors to calculate a partial likelihood.
@@
 This idea has also appeared in the cross-validation literature for quantities that cannot be evaluated on only a subset of the data, such as AUC in logistic and Cox regression models \citep{Parker2007,Simon2011a,Subramanian2011}.
-
-%%%%%%%%%%%%%%%%%%%%
-\subsection{Cross validated deviance residuals}
-%%%%%%%%%%%%%%%%%%%%
-
-The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model, which prevents us from calculating prediction errors (residuals).
-However, if one takes additional steps to estimate the baseline hazard in addition to estimating the coefficients, then such residuals can be calculated.
-The deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}, is a natural candidate for quantifying prediction error to use in quantifying the cross-validation error.
-Originally proposed for model diagnostics, deviance residuals have also been occasionally used as a predictive accuracy measure \citep{Therneau2018}.
-In the Cox model, the Martingale residuals $\{m_i\}_{i=1}^n$ and deviance residuals $\{d_i\}_{i=1}^n$ are given by
-\begin{align}
-  \label{eq:res-m}
-  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(X_i\Tr\bbeta), \\
-  \label{eq:res-d}
-  d_{i} &= \text{sign}(m_{i})\sqrt{-2(m_i + \delta_{i}\log(\delta_{i} - m_i))}.  
-\end{align}
-where $H_0$ is the cumulative baseline hazard.
-
-In order to obtain a cross-validated version of these residuals, we would replace $\bbeta$ in the above equations with $\hat{\bbeta}^{-k}$.
-However, it is not obvious how to obtain a cross-validated version of $H_0$, or to what extent this is necessary.
-We experimented with three versions of cross-validated deviance residuals, each with a different choice for $H_0$: (1) The Nelson-Aalen estimator, which does not involve the covariates and is therefore the same across cross-validation folds, (2) the Breslow estimator \citep{Breslow1972}, and (3) the Kalbfleisch-Prentice estimator \citep{Kalbfleisch2011}.
-Details for the three approaches are given in the supplementary materials, along with the results of a simulation study comparing them.
-Because the Nelson-Aalen baseline hazard performed best in all situations, we use this estimate to construct the cross-validated deviance residuals throughout the rest of the paper.
-
-The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross validated error, analogous to using the residual sum of squares in linear regression.
-We refer to this method as the \emph{cross-validated deviance residuals} approach.
 

### 2022-10-27 sub/smimr/response.tex
Edits to responses

@@
 \newcommand{\re}{\textbf{Response: }}
+\newcommand\pb[1]{{\color{red}#1}}
 \newcommand\bd[1]{{\color{blue}#1}}
@@
 
-\section*{Response to Reviewers}
-\subsection*{Manuscript ID SMM-22-0007: Cross Validation Approaches for
-Penalized Cox Regression}
+\subsubsection*{Response to comments by Reviewer \#1}
 
-The authors wish to thank the reviewers for their careful reviews and thoughtful comments.
-
-\subsection*{Response to Reviewer 1's Comments}
-
-The authors compare four different cross-validation options for the important problem of penalized Cox regression, including two new proposed methods.
-The methodology is nicely described, and the illustration in Figure 1 should prove helpful to readers new to this area.
-The interpretations of the simulation results are helpful, and a good choice for the application dataset was made.
-The take-home message of the usefulness of the linear predictor approach was clearly made based on the work performed.
-However, there are a number of important questions and comments to mention here that should be addressed:
+Thank you very much for your careful reading of the paper and your insightful comments, particularly regarding issues relating to time-varying covariates.
+You raise a number of good points and we have modified the manuscript considerably based upon your feedback.
+In our opinion, these revisions have very much improved upon the original submission.
 
@@
 \re Thank you for this comment.
-We provided a few examples of penalty functions, such as the elastic net penalty, smoothly clipped absolute deviations (SCAD) penalty, and minimax concave penalty (MCP) on page 4.
-We also included a paragraph for more general application in the Discussion section.
+In the revision (page 4), we provide a few examples of penalty functions, such as the elastic net penalty, smoothly clipped absolute deviations (SCAD) penalty, and minimax concave penalty (MCP).
+We also mention this in the Discussion section.
 
 \item Though you aim to make various considerations in your simulation study, the impact of this work may very well be limited given consideration of only exponential survival distribution and independent censoring.
-    \begin{enumerate}
+Even common choices for a more complex distributional form, such as Weibull and Gompertz, might be considered too limiting.
+See, for example, Harden and Kropko (2019).
+
+\textbf{Response (distributional form):} You raise a good point -- we did not expect the results to change much depending on the distributional form, but this is something that deserves to be investigated.
+In the revised supplementary material, we include additional simulation results where the baseline hazard is generated from Weibull and Gompertz distributions.
+The overall conclusions remain fairly similar, however.
+This is mentioned in the revised manuscript in the text surrounding Table 1.
+
+\textbf{Response (independent censoring):} With regard to independent censoring, however, we feel that this would muddy the waters regarding whether a cross-validation method is working well or not.
+If the assumptions of the Cox model are not met, the partial likelihood is incorrect, and all models are wrong, which model should you choose?
+For example, suppose model A predicts well, but the coefficient estimates are biased, whereas model B estimates the coefficients more accurately but is worse at predicting the outcome.
+Is a cross-validation that favors model B better or worse than one that favors model A?
+These questions are interesting, but they have no clear answer and don't offer much help in terms of objectively determining which CV approach is best.
+See also point \ref{limitations} below.
+
 \item Extensions into time-varying covariates, and, preferably, joint modeling, would be useful as well.
@@
 
-        \re Thank you for making this comment.
-This work aims to make comparisons among cross-validation methods, which are more commonly used in the model selection context.
-In the Discussion section, we included a paragraph to state that the methods in this manuscript can be applied to any model selection process that involves the cox partial likelihood.
-While cross-validation and model selection play a key role for fitting penalized cox regression models, the process of fitting time-varying covariates and joint modeling with longitudinal covariates do not always involve model selection, therefore this extension is out of the main focus of this work.
+\re This is an excellent point -- additional issues may certainly arise when applying cross-validation to time-varying data.
+Our experience with high-dimensional Cox regression has tended to involve features that were only measured once due to the cost of these experiments, but as this cost goes down, that may certainly change.
+We feel that addressing these issues is beyond the scope of the current manuscript, but is definitely something worth discussing.
+The final paragraph of our discussion now points out these issues and discusses that future work is necessary to investigate the application of the proposed approach to these settings.
+\pb{ADD THIS PARAGRAPH}
 
-        \item Regarding distributional form, the exponential form here is quite limiting, and even common choices for a more complex distributional form, such as Weibull and Gompertz, might be considered too limiting.
-See, for example, Harden and Kropko (2019).
-
-        \re Thank you for this comment.
-We included additional simulation results in the supplementary material where the baseline hazard is generated from Weibull and Gompertz distribution.
-The overall conclusions do not change for both of these more complex distributional form.
-
-    \end{enumerate}
-
-\item Speaking of your Discussion section, there is no mention of any limitations of the current work.
+\item \label{limitations} Speaking of your Discussion section, there is no mention of any limitations of the current work.
 Aside from the joint modeling comment above, you should also talk about violations of linearity, and of non-proportional hazards.
@@
 
-\subsection*{Response to Reviewer 2's Comments}
-In this manuscript, the authors proposed two new cross-validation methods for Cox proportional hazards regression.
+\newpage
+
+\subsubsection*{Response to comments by Reviewer \#2}
 

### 2022-10-27 main.tex
Edits to responses

@@
 
-In this paper, we focus upon the LASSO penalty $P_{\lambda}(\bbeta) = \lambda \sum_{j} |\beta_{j}|$, although the methods that we analyze not only can be used for penalized Cox regression with any penalty, such as elastic net \citep{Zou2005}, smoothly clipped absolute deviations (SCAD) penalty \citep{FanLi2001}, and minimax concave penalty (MCP) \citep{Zhang2010}, but are also applicable to model selection in unpenalized Cox regression.
+In this paper, we focus upon the LASSO penalty $P_{\lambda}(\bbeta) = \lambda \sum_{j} |\beta_{j}|$ as it is most common, although the need to select regularization parameters arises in all penalized regression models, such as elastic net \citep{Zou2005}, smoothly clipped absolute deviations (SCAD) penalty \citep{FanLi2001}, and minimax concave penalty (MCP) \citep{Zhang2010}.
+Indeed, cross-validation can be used for model selection in Cox regression even without the presence of a penalty, so the relevance of this question is wide-ranging.
 LASSO-penalized Cox regression is particularly useful for high dimensional data where the number of covariates $p > n$, such as predicting overall survival in cancer patients based on genome-wide expression measurements.

### 2022-10-27 sub/smimr/1/scrub-url.R

@@ -1,4 +1,4 @@
-bib <- readLines('../../articles.bib')
+bib <- readLines('../../../articles.bib')
 bib <- bib[!grepl('eprint', bib)]
 url <- grepl("url\\s*=", bib)
 rproj <- grepl("R-project\\.org", bib)

### 2023-01-04 sub/smimr/response.tex
PB done editing responses

@@
 Is a cross-validation that favors model B better or worse than one that favors model A?
-These questions are interesting, but they have no clear answer and don't offer much help in terms of objectively determining which CV approach is best.
+These questions are potentially interesting, but they have no clear answer and don't offer much help in terms of objectively determining which CV approach is best.
 See also point \ref{limitations} below.
@@
 
-\re Thank you for the comments.
-The violation of linearity is usually handled via selecting or fitting a non-linear functional form for the covariates.
-The approaches that are discussed here are built upon Cox partial likelihood, which do not deal with non-proportional hazards and rare events, therefore those topics are out of the scope of this manuscript.
-We included sentences in the Discussion section to address those points.
+  \re Our point of view in this manuscript is that you have decided to fit a Cox proportional hazard model, but aren't sure what the best way to carry out cross-validation is.
+Certainly, if the Cox model is wrong in the first place, then it may be best to take a completely different approach to modeling, and this manuscript is not likely to be relevant.
+We don't really see the point of discussing issues like non-proportional hazards and non-linear effects, as they have already been covered in hundreds, if not thousands, of other works.
 
@@
 
-\re Thank you for the suggestion. We added the sentence to the first paragraph of Section 3.1.
+  \re Thank you for the suggestion; we have added the sentence as you suggest.
 
@@
 
-\re Thank you for the suggestion. We added $t_0$ index to $y$ notation.
+  \re Thank you for the suggestion; we have modified the notation according to your recommendation.
 
@@
 
-\re Thank you for this comment.
-We provided additional simulation results in the supplement with more choices of $t_0$ with Exponential, Weibull and Gompertz distributions.
-The overall conclusions do not change with various $t_0$.
+  \re We have now carried out additional simulation results in the supplement with more choices of $t_0$; the overall conclusions do not change in any meaningful way at different $t_0$.
 
@@
 
-\re Thank you for catching this. We clarified in the writing that this censoring rate only applies to settings in Table 1.
+\re Thank you for catching this. We have now clarified that this censoring rate only applies to Table 1.
 
@@
 
-\re \bd{To do: address this in Discussion}
+  \re We have added a comment in the Introduction clarifying our usage: throughout, we use the term ``conservative'' to mean a tendency to select a model with few predictors, or even the null model.'' We agree that in general one could mean different things by ``conservative'', although we felt that it was getting off topic to discuss this issue at length in the manuscript.
 
@@
 
-\re Thank you for the suggeston. We included the precision lasso method in the discussion section.
+  \re Thank you for the suggestion; we now mention the precision lasso method in the discussion section.
 
@@
 
-\re \bd{maybe the easiest way is to use more alpha values in this example?}
+  \re \pb{Need Biyue to edit Table 3 before I can reply here.}
 
 \item There are several typos sprinkled throughout the paper, so a careful proofreading is required.
-Some examples include: (i) Section 3.1, second paragraph, where the phrase “the value of $\beta$ used the generate the data” requires editing.
-(ii) Table 1, Scenario 4, would seem to be for
-strong signal, not weak. (iii) p.22, “more clear” should be changed to “clearer”.
+Some examples include: (i) Section 3.1, second paragraph, where the phrase ``the value of $\beta$ used the generate the data'' requires editing.
+(ii) Table 1, Scenario 4, would seem to be for strong signal, not weak.
+(iii) p.22, ``more clear'' should be changed to ``clearer''.
 
-\re Thank you for catching those. We made edits in the manuscript.
+  \re Thank you very much for catching these typos and pointing them out; these are all corrected in the revised manuscript.
 
@@
 
-\re \bd{maybe one advantage is that the deviance residuals can be applied to a more wide range of survival models other than the regression approaches.
-say, for a random forest model with survival outcome, they can't really use the linear predictor approach right?}
+  \re We completely agree that cross-validated deviance residuals are bad and should not be used for this purpose, period.
+We originally felt that it was worth pointing out the shortcomings of this approach, although we can see how it leads to confusion.
+Before throwing it out of the paper, however, we carried out another literature review and found that cross-validated linear predictors \emph{have} been proposed and \emph{are} used for some Cox model selection, particularly in the regression tree and random forest literature.
+
+  In the revised manuscript, we still describe cross-validated deviance residuals and point out their flaws as a metric for model selection, but no longer refer to it as a ``proposed method''.
+We really do feel that pointing out the flaws of certain approaches has just as much merit as pointing out the benefits of others, and believe that the paper is stronger with the comparison to deviance residuals included.
 
@@
 Specifically, data characteristics such as sample size and censoring rate are not consistent across simulation scenarios considered for each of different evaluation metrics in Section 3.
+I suggest presenting all the simulation scenarios varying different data characteristics at the beginning of Section 3 and discuss the comparative performance of the methods using each of model evaluation metrics in the following subsections.
 
-    \begin{enumerate}
-    \item I suggest presenting all the simulation scenarios varying different data characteristics at the beginning of Section 3 and discuss the comparative performance of the methods using each of model evaluation metrics in the following subsections.
-Addressing this comment will require additional computations for Sections 3.3-3.5
+  \re A considerable amount of time and effort went into identifying simulation scenarios that reveal interesting results about the methods being considered.
+Doing something like keeping the sample size the same while raising the number of predictors from 50 to 10,000 would only result in meaningless tables where all the methods fail.
 
-    \re \bd{I think there are definitely values to present a different variety of simulation scenarios. but maybe we should re-structure the writing a bit more}
+  In theory, perhaps, one simple set of simulation parameters will work to illustrate every single point.
+In reality, this is rarely true and settings must be adjusted to the situation at hand to most clearly demonstrate what is happening.
 
@@
 
-    \re We explored the impact of censoring in Section 3.1, varying from 40$\%$ and to 80$\%$.
-The censoring proportion has the greatest impact on the stability for the basic method, but does not impact the overall conclusions for the other methods.
+  \re The impact of censoring is explored in Section 3.1, where we vary the censoring rate from 40$\%$ and to 80$\%$.
+The main impact of censoring proportion with respect to the CV approaches considered here is that basic CV suffers greatly from stability issues when the censoring rate is high.
+This makes it difficult to have a high rate of censoring in general if we wish to compare with basic CV.
 
 \item Table 2 - Why only present the number of true positives here?
-I suggest presenting other operating characteristics such as TPR, NPR, PPV, etc...
-Then the results might help better justify the findings from the application.
+I suggest presenting other operating characteristics such as TPR, NPR, PPV, etc \ldots Then the results might help better justify the findings from the application.
 For example, depending on the results, referring to the simulation results could strengthen the argument given at the end of the second paragraph in Discussion (which is quite speculative in my opinion).
 
-    \re \bd{one argument would be we are evaluting the predictive performance. This comment can be addressed jointly with Reviewer 1's comment 8}
+  \re We are not proposing any new models.
+The model is exactly the same for all the entries in Table 2: lasso-penalized Cox regression.
+None of these methods (V\&VH, Basic, LinearPred, and DevResid) is going to be objectively superior in terms of variable selection accuracy (they all have identical AUROC, for example).
+They're using the same model to select variables, just stopping at different points along the path.
+Table 2 illustrates that by stopping early, the V\&VH method misses out on a number of true signals.
 
-\end{enumerate}
+  Obviously, it also selects fewer false positives as well.
+However, since there is no objectively correct answer to the balance between false positives and false negatives that a model should include, there is no way to decide upon these methods by comparing TPR/NPR/PPV/etc.
+There are, however, objective measures of prediction accuracy, which demonstrate that V\&VH stops too early -- this is point of Section 3.1.
+
+  To summarize, the logic of our presentation is: V\&VH stops too early (Section 3.1). What are the consequences of this? It misses true positives (Section 3.2).
 
@@
 
-\re The cross-validated linear predictor approach is implemented in the \texttt{ncvreg} package.
-The basic approach and the Verweij and Van Houwelingen's approach are available with the \texttt{glmnet} package.
-\bd{are we submitting the R codes as supplementary material?}
+\re Yes; the proposed cross-validated linear predictor approach is implemented in the \textbf{ncvreg} package.
+The basic approach and the Verweij and Van Houwelingen's approach are implemented in the \textbf{glmnet} package.
+The code to reproduce our simulations and analyses are available as a GitHub repository.
 
-\item Many places require major/minor clarification and corrections:
-\bd{the following points are all edited in the manuscript}
+\item Many places require major/minor clarification and corrections.
+
+  \re Thank you very much for catching and pointing out all of these issues! All of them have been fixed in the revised version of the manuscript.
+
 \begin{itemize}[align = left]
@@
 
-\re Thank you for the suggestion. We included definition for $\bbeta$ in this paragraph when first introducing the notation.
+%\re Thank you for the suggestion. We included definition for $\bbeta$ in this paragraph when first introducing the notation.
 
@@
 
-\re Thank you for the suggestion. We updated the manuscript by using bold betters for vectors.
+%\re Thank you for the suggestion. We updated the manuscript by using bold betters for vectors.
 
@@
 
-\re Thank you for this careful observation.
+%\re Thank you for this careful observation.
 In the figure, we used "Folder 1" for the simplicity of drawing.
@@
 
-\re Thank you for the suggestion. We replaced $k$ with $\tau$ in that formula
+%\re Thank you for the suggestion. We replaced $k$ with $\tau$ in that formula
 
@@
 
-\re No additional steps are needed to pool the cross-validated linear predictors over all K folds.
+%\re No additional steps are needed to pool the cross-validated linear predictors over all K folds.
 Please refer to the linear predictor diagram in Figure 1.
@@
 
-\re The numbers in the parentheses are standard deviation across all simulated samples, which is a metric for simulation errors.
+%\re The numbers in the parentheses are standard deviation across all simulated samples, which is a metric for simulation errors.
 
@@
 
-\re Thank you for the suggestion. We updated the caption to include defition for LOO.
+%\re Thank you for the suggestion. We updated the caption to include defition for LOO.
 
@@
 
-\re Thank you for catching this. We corrected this in the caption.
+%\re Thank you for catching this. We corrected this in the caption.
 
@@
 
-\re Thank you for the suggestion. We edited the caption of Figure 4 to clarify the legends.
+%\re Thank you for the suggestion. We edited the caption of Figure 4 to clarify the legends.
 
@@
 
-\re The column "Selected" represents the number of genes that are selected by the linear predictor cross-validation approach; each row represents the numbder of genes that fall into that function category.
+%\re The column "Selected" represents the number of genes that are selected by the linear predictor cross-validation approach; each row represents the numbder of genes that fall into that function category.
 $p$ is the p-value.
@@
 
-\subsection*{References}
+%% \subsection*{References}
 
-Harden JJ, Kropko J (2019). Simulating duration data for the Cox model, Political Science Research and Methods, 7(4): 921-928. https://doi.org/10.1017/psrm.2018.19
+%% Harden JJ, Kropko J (2019). Simulating duration data for the Cox model, Political Science Research and Methods, 7(4): 921-928. https://doi.org/10.1017/psrm.2018.19
 
-Wang H, Lengerich BJ, Aragam B, Xing EP (2019).
+%% Wang H, Lengerich BJ, Aragam B, Xing EP (2019).
 Precision Lasso: accounting for correlations and linear dependencies in high-dimensional genomic data.

### 2023-01-04 main.tex
PB done editing responses

@@
 We varied the sample size, dimension, censoring percentage, and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation.
-Accuracy was assesed via several criteria.
-The specifics of the different conditions considered in each simulation experiment are described below.
+The specifics of the simulation experiments are described below.
 
 Estimation accuracy was measured using the mean squared error of the coefficients ($\MSE$).
-We define $\hat{\bbeta}(\lambda)$ as the vector of coefficients estimated by a Cox model with LASSO penalty at $\lambda$, and define $\bbeta^*$ as the value of $\bbeta$ that is used the generate the data.
+We define $\hat{\bbeta}(\lambda)$ as the vector of coefficients estimated by a Cox model with LASSO penalty at $\lambda$, and define $\bbeta^*$ as the true value of $\bbeta$ (the value used to generate the data).
 Mean squared error is defined to be

### 2023-01-05 sub/smimr/response.tex
Revisions to final paragraph of discussion

@@
 The final paragraph of our discussion now points out these issues and discusses that future work is necessary to investigate the application of the proposed approach to these settings.
-\pb{ADD THIS PARAGRAPH}
 
@@
 \subsubsection*{Response to comments by Reviewer \#2}
+
+Thank you very much for your careful reading of the paper and your insightful comments, particularly regarding the cross-validated deviance residuals.
+This was the weakest aspect of the original submission, and by putting thought and work into addressing the issues you raised, we feel that we have significantly improved the manuscript.
 

### 2023-01-05 main.tex
p to p-value
Revisions to final paragraph of discussion

@@
 \toprule
-ID & Name & Selected & Enrichment & p \\
+ID & Name & Selected & Enrichment & $p$-value \\
 \midrule
@@
 
-Although we focused upon the penalized Cox regression with a LASSO penalty in this work, the methods that we analyzed are applicable in any context that involves conducting model selection for survival models that are built upon the Cox partial likelihood, such fitting penalized Cox regression with other penalty functions (elastic net \citep{Zou2005}, SCAD\citep{FanLi2001}, MCP\citep{Zhang2010}, precision lasso\citep{Wang2018}) and penalty structures (group selection, bi-level selection), selecting functional forms of covariates in unpenalized Cox regression, model selection for unpenalized Cox regression that involve time-varying covariates and longitudinal covariates.
-Since all the cross-validation approaches that are discussed in this manuscript depend on the Cox partial likelihood, they are subject to limitations such as non-proportional hazards and rare events that are tied to the Cox proportional hazard model.
+Although we focused upon the penalized Cox regression with a LASSO penalty in this work, the methods that we analyzed are applicable in any context that involves conducting model selection for survival models that are built upon the Cox partial likelihood, such fitting penalized Cox regression with other penalty functions (elastic net \citep{Zou2005}, SCAD \citep{FanLi2001}, MCP \citep{Zhang2010}, precision lasso \citep{Wang2018}), other penalty structures (group selection, bi-level selection), and selecting functional forms of covariates in unpenalized Cox regression.
+We note, however, that the focus of this article is restricted to fixed covariates.
+How best to carry out cross-validation for model selection in the presence of time-varying covariates, or in joint modeling of time-to-event and longitudinal outcomes, remains an open question worthy of additional future study.
 

### 2023-08-11 main.tex
update table 3 and upload new R code

@@
 \caption{\label{tb:shedden} The C Index and Brier Score of all elastic net models selected by cross-validation methods }
-\begin{tabular}{lccccc}
+\begin{tabular}{lcccccccc}
 \hline
-  & \multicolumn{5}{c}{Cross-validated C Index} \\ 
-  & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-6}
-V \& VH      & 0.639 & 0.637 & 0.638 & 0.637 & 0.637\\
-Basic       & 0.647 & 0.639 & 0.638 & 0.638 & 0.639 \\
-Linear Pred    & \textbf{0.650} & 0.647 & 0.638 & 0.638 & 0.639  \\
-Deviance Resid & 0.645 & 0.637 & 0.638 & 0.637 & 0.637 \\ 
+  & \multicolumn{8}{c}{Cross-validated C Index} \\ 
+  & $\alpha = 0.05$ & $\alpha = 0.1$ & $\alpha = 0.2$ & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-9}
+V \& VH     & 0.645 &0.645 &0.643 &0.630 &0.626 &0.617 &0.617 &0.617\\
+Basic       &0.652 &0.649 &0.650 &0.650 &0.649 &0.650 &0.651 &0.651 \\
+Linear Pred    &0.649 &0.648 &0.650 &0.650 &0.649 &0.648 &0.617 &0.617  \\
+Deviance Resid &0.648 &0.645 &0.647 &0.646 &0.626 &0.617 &0.617 &0.617 \\ 
 \addlinespace
-  & \multicolumn{5}{c}{Brier Score} \\ 
-  & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-6}
-V \& VH  & 0.243 & 0.244 & 0.244 & 0.245 & 0.245 \\
-Basic  & 0.233 & 0.243 & 0.244 & 0.244 & 0.244 \\
-Linear Pred & \textbf{0.229} & 0.232 & 0.244 & 0.244 & 0.244 \\
-Deviance Resid & 0.235 & 0.244 & 0.244 & 0.245 & 0.245 \\ 
+  & \multicolumn{8}{c}{Brier Score} \\ 
+  & $\alpha = 0.05$ & $\alpha = 0.1$ & $\alpha = 0.2$ & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-9}
+V \& VH &0.232 &0.231 &0.232 &0.243 &0.245 &0.249 &0.249 &0.249 \\
+Basic  &0.221 &0.226 &0.226 &0.225 &0.224 &0.221 &0.221 &0.221 \\
+Linear Pred & 0.220 &0.228 &0.226 &0.225 &0.224 &0.224 &0.249 &0.249 \\
+Deviance Resid &0.229 &0.231 &0.229 &0.228 &0.245 &0.249 &0.249 &0.249 \\ 
 \hline

### 2023-08-11 resubmit_code/shedden_elasticnet_git.R

@@ -1,0 +1,321 @@
+#------------------------
+#
+# Self contained R code
+# to analyze the Shedden Lung Cancer data set
+#
+#------------------------
+
+#------------------------
+# load/define functions that will be used in the analysis
+library(survival)
+library(ncvreg)
+library(glmnet)
+
+
+# Brier score for selected lambda
+
+brierSurv_lambda <- function(data,time,fit,lambda) {
+  predicted <- apply(data$x,1, 
+                     function(x)
+                     {predict(fit, x,type = "survival",
+                              lambda = lambda)(time)})
+  is_obs_after <- data$y[,1] > time
+  mean((is_obs_after - predicted)^2)
+}
+
+#------------------------
+# load data
+
+# Shedden Lung Cancer Data
+data <- readRDS(url("https://s3.amazonaws.com/pbreheny-data-sets/Shedden2008.rds"))
+X <- data$X
+S <- data$S
+Z <- data$Z
+
+# censor S by 72 month
+plot(survfit(S ~ 1), xlab = "month", ylab = "Survival Probability")
+abline(h = 0.5, col = "grey", lty = 2)
+
+# clean data
+ncolX <- ncol(X)
+n <- nrow(X)
+Z <- as.data.frame(Z)
+
+data_low <- data.frame(S,Z)
+data_low$Age <- as.numeric(data_low$Age)
+data_low$Sex[data_low$Sex == "Female"] <- 0
+data_low$Sex[data_low$Sex == "Male"] <- 1
+data_low$Sex <- as.numeric(data_low$Sex)
+data_low$AdjChemo <- as.numeric(data_low$AdjChemo == "Yes")
+
+x <- as.matrix(data.frame(data_low[,c(2,3,5)],X))
+x <- ncvreg::std(x)
+data <- list(y=S,x=x)
+
+# prespecify fold id
+set.seed(123456)
+nfold <- 10
+nevent <- sum(data$y[,2])
+sde <- sqrt(.Machine$double.eps)
+foldid <- numeric(0)
+foldid[data$y[,2] == 1] <-  ceiling(sample(1:nevent)/(nevent + sde) * nfold)
+foldid[data$y[,2] == 0] <-  ceiling(sample(1:(n-nevent))/(n - nevent + sde) * nfold)
+
+
+Alpha <- c(1,0.9,0.7,0.5,0.3,0.2,0.1,0.05)
+A <- length(Alpha)
+
[…]

### 2023-08-11 resubmit_code/table_1_high_dim_gompertz.R

@@ -1,0 +1,299 @@
+#------------------------------
+# Self-contained R code
+# to reproduce part of the simulations
+# in Table 1 of the Cross Validation 
+# Manuscript:
+# Explored 2 Scenarios here: 
+# high dim low signal
+# high dim high signal
+# Gompertz Distribution
+#------------------------------
+
+id <- as.numeric(commandArgs(TRUE)[1])
+library(survival)
+library(ncvreg)
+library(eha)
+
+#-------------------------------------------------------------------------
+# functions
+
+generatebeta <- function(p,nbeta,c = 1){
+  pos <- rep(c, ceiling(nbeta/2))
+  neg <- rep(-c, floor(nbeta/2))
+  beta <- c(pos,neg,rep(0,p-nbeta))
+}
+
+genSurvData <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,shape = 1, rate = 1 # 70% means 70% of the data are censored; 30% are observed
+){
+  p <- length(beta)
+  ## Generate column vector of coefficients
+  beta <- matrix(beta, ncol = 1)
+  x <- matrix(rnorm(n*p, sd = sd),n,p)
+  #standardize x first
+  x <- ncvreg::std(x)
+  #hazard <- h*exp(x %*% beta)
+  ## Calculate survival times
+  #time <- rexp(n, hazard)
+  
+  # use gompertz distribution
+  hazard_rate <- shape*exp(x %*% beta)
+  time <- eha::rgompertz(n, shape = hazard_rate, rate = rate, param = "rate")
+  
+  ## observations are censored ("event = rbinom(n, 1, 0.30)")
+  y <- Surv(time = time, event = rbinom(n, 1, 1-cpct))
+  return(list(x=x,y=y,beta = beta))
+}
+
+
+# Brier score for selected lambda
+brierSurv_lambda <- function(data,time,fit,lambda) {
+  predicted <- apply(data$x,1, 
+                     function(x)
+                     {predict(fit, x,type = "survival",
+                              lambda = lambda)(time)})
+  is_obs_after <- data$y[,1] > time
+  mean((is_obs_after - predicted)^2)
+}
+
+# Kullback-Leibler for selected lambda
[…]

### 2023-08-11 resubmit_code/table_1_high_dim_various_t.R

@@ -1,0 +1,292 @@
+#------------------------------
+# Self-contained R code
+# to reproduce part of the simulations
+# in Table 1 of the Cross Validation 
+# Manuscript:
+# Explored 2 Scenarios here: 
+# high dim low signal
+# high dim high signal
+#------------------------------
+
+id <- as.numeric(commandArgs(TRUE)[1])
+library(survival)
+library(ncvreg)
+
+#-------------------------------------------------------------------------
+# functions
+
+generatebeta <- function(p,nbeta,c = 1){
+  pos <- rep(c, ceiling(nbeta/2))
+  neg <- rep(-c, floor(nbeta/2))
+  beta <- c(pos,neg,rep(0,p-nbeta))
+}
+
+genSurvData <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,h = 1 # 70% means 70% of the data are censored; 30% are observed
+){
+  p <- length(beta)
+  ## Generate column vector of coefficients
+  beta <- matrix(beta, ncol = 1)
+  x <- matrix(rnorm(n*p, sd = sd),n,p)
+  #standardize x first
+  x <- ncvreg::std(x)
+  hazard <- h*exp(x %*% beta)
+  ## Calculate survival times
+  time <- rexp(n, hazard)
+  ## observations are censored ("event = rbinom(n, 1, 0.30)")
+  y <- Surv(time = time, event = rbinom(n, 1, 1-cpct))
+  return(list(x=x,y=y,beta = beta))
+}
+
+
+# Brier score for selected lambda
+brierSurv_lambda <- function(data,time,fit,lambda) {
+  predicted <- apply(data$x,1, 
+                     function(x)
+                     {predict(fit, x,type = "survival",
+                              lambda = lambda)(time)})
+  is_obs_after <- data$y[,1] > time
+  mean((is_obs_after - predicted)^2)
+}
+
+# Kullback-Leibler for selected lambda
+KL_lambda <- function(data,time,fit,lambda) {
+  predicted <- apply(data$x,1, 
+                     function(x)
+                     {predict(fit, x,type = "survival",
[…]

### 2023-08-11 resubmit_code/table_1_high_dim_weibull.R

@@ -1,0 +1,297 @@
+#------------------------------
+# Self-contained R code
+# to reproduce part of the simulations
+# in Table 1 of the Cross Validation 
+# Manuscript:
+# Explored 2 Scenarios here: 
+# high dim low signal
+# high dim high signal
+#------------------------------
+
+id <- as.numeric(commandArgs(TRUE)[1])
+library(survival)
+library(ncvreg)
+
+#-------------------------------------------------------------------------
+# functions
+
+generatebeta <- function(p,nbeta,c = 1){
+  pos <- rep(c, ceiling(nbeta/2))
+  neg <- rep(-c, floor(nbeta/2))
+  beta <- c(pos,neg,rep(0,p-nbeta))
+}
+
+genSurvData <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,scale = 1, shape = 1 # 70% means 70% of the data are censored; 30% are observed
+){
+  p <- length(beta)
+  ## Generate column vector of coefficients
+  beta <- matrix(beta, ncol = 1)
+  x <- matrix(rnorm(n*p, sd = sd),n,p)
+  #standardize x first
+  x <- ncvreg::std(x)
+  #hazard <- h*exp(x %*% beta)
+  ## Calculate survival times
+  #time <- rexp(n, hazard)
+  
+  # use weibull regression
+  hazard_rate <- (1/scale)*exp(x %*% beta/shape)
+  time <- rweibull(n, shape = shape, scale = 1/hazard_rate)
+  
+  ## observations are censored ("event = rbinom(n, 1, 0.30)")
+  y <- Surv(time = time, event = rbinom(n, 1, 1-cpct))
+  return(list(x=x,y=y,beta = beta))
+}
+
+
+# Brier score for selected lambda
+brierSurv_lambda <- function(data,time,fit,lambda) {
+  predicted <- apply(data$x,1, 
+                     function(x)
+                     {predict(fit, x,type = "survival",
+                              lambda = lambda)(time)})
+  is_obs_after <- data$y[,1] > time
+  mean((is_obs_after - predicted)^2)
+}
+
+# Kullback-Leibler for selected lambda
+KL_lambda <- function(data,time,fit,lambda) {
[…]

### 2023-08-14 sub/smimr/response.tex
Editing text surrounding Table 3 and response

@@
-\documentclass[12]{article}
+\documentclass{article}
 \usepackage[utf8]{inputenc}
@@
 Certainly, if the Cox model is wrong in the first place, then it may be best to take a completely different approach to modeling, and this manuscript is not likely to be relevant.
-We don't really see the point of discussing issues like non-proportional hazards and non-linear effects, as they have already been covered in hundreds, if not thousands, of other works.
+We don't really see the need to discuss issues like non-proportional hazards and non-linear effects, as they have already been covered in hundreds, if not thousands, of other manuscripts and textbooks.
 
@@
 
-  \re \pb{Need Biyue to edit Table 3 before I can reply here.}
+  \re This is certainly a fair point.
+In the revised manuscript, we include $\alpha=0.1$ and $\alpha=0.2$ in the comparison as well.
+In our experience (for all kinds of high-dimensional regression, not just survival models), there is rarely an obvious choice for $\alpha$: many choices of $\alpha$ perform similarly.
+For the case study presented in Section 4, it seems clear that lower values of $\alpha$ consistently outperform values near 1, but honestly, you could make a case for any value of $\alpha$ below 0.5.
+We focus on $\alpha=0.3$ for the sake of illustration, but the patterns (in terms of how the CV methods compare to each other) are similar for other choices of $\alpha$.
+In addition to modifying Table 3 itself, we have modified the text surrounding it to point out these issues, rather than the overly simplistic remark in the original submission that ``$\alpha=0.3$ is best.''
 
@@
 
-  \re Thank you very much for catching and pointing out all of these issues! All of them have been fixed in the revised version of the manuscript.
+  \re Thank you very much for catching and pointing out all of these issues!
+We truly appreciate your attention to detail in reading our work.
+All of them have been fixed in the revised version of the manuscript.
 

### 2023-08-14 main.tex
Editing text surrounding Table 3 and response

--- 2023-01-05 main.tex
@@
 In addition to the regularization parameter $\lam$, the elastic net model has an extra tuning parameter $\alpha$, where $\alpha=1$ corresponds to lasso and $\alpha=0$ corresponds to ridge regression.
-For $\alpha \in \{0.3, 0.5, 0.7, 0.9, 1\}$, we selected a regularization parameter $\lam$ that has the lowest cross-validation error for each of the four cross-validation methods described in Section~\ref{Sec:methods}.
-Although each method can be used to select a pair of $\alpha$ and $\lam$ values, the methods are on different scales and cannot be directly compared with each other.
-For this reason, Table~\ref{tb:shedden} reports the cross-validated C-Index and Brier Scores for each method's choice of $\lam$ at each value of $\alpha$.
-As shown in the table, the elastic net model with $\lam$ and $\alpha=0.3$ selected by the cross-validated linear predictor approach has the best performance in terms of C Index and Brier Score.
+For $\alpha \in \{0.1, 0.2, 0.3, 0.5, 0.7, 0.9, 1\}$, we selected the regularization parameter $\lam$ that has the lowest cross-validation error for each of the four cross-validation methods described in Section~\ref{Sec:methods}.
+As shown in Table~\ref{tb:shedden}, which reports the cross-validated C-Index and Brier Scores for each method, models with smaller values of $\alpha$ tend to perform better than models with $\alpha \approx 1$ (higher C-Index, lower Brier scores).
+As is typical in high-dimensional regression, there is no obvious best choice for $\alpha$, but for the sake of illustration, we focus on $\alpha=0.3$ in the remainder of this section, as it seems to consistently outperform $\alpha=1$ and there is little evidence of further gain by dropping $\alpha$ below 0.3.
 
@@
 \caption{\label{tb:shedden} The C Index and Brier Score of all elastic net models selected by cross-validation methods }
-\begin{tabular}{lccccc}
+\begin{tabular}{lccccccc}
 \hline
-  & \multicolumn{5}{c}{Cross-validated C Index} \\ 
-  & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-6}
-V \& VH      & 0.639 & 0.637 & 0.638 & 0.637 & 0.637\\
-Basic       & 0.647 & 0.639 & 0.638 & 0.638 & 0.639 \\
-Linear Pred    & \textbf{0.650} & 0.647 & 0.638 & 0.638 & 0.639  \\
-Deviance Resid & 0.645 & 0.637 & 0.638 & 0.637 & 0.637 \\ 
+  & \multicolumn{7}{c}{Cross-validated C Index} \\ 
+  & $\alpha = 0.1$ & $\alpha = 0.2$ & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-8}
+V \& VH        & 0.645 &0.643 &0.630 &0.626 &0.617 &0.617 &0.617 \\
+Basic          & 0.649 &0.650 &0.650 &0.649 &0.650 &0.651 &0.651 \\
+Linear Pred    & 0.648 &0.650 &0.650 &0.649 &0.648 &0.617 &0.617 \\
+Deviance Resid & 0.645 &0.647 &0.646 &0.626 &0.617 &0.617 &0.617 \\ 
 \addlinespace
-  & \multicolumn{5}{c}{Brier Score} \\ 
-  & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-6}
-V \& VH  & 0.243 & 0.244 & 0.244 & 0.245 & 0.245 \\
-Basic  & 0.233 & 0.243 & 0.244 & 0.244 & 0.244 \\
-Linear Pred & \textbf{0.229} & 0.232 & 0.244 & 0.244 & 0.244 \\
-Deviance Resid & 0.235 & 0.244 & 0.244 & 0.245 & 0.245 \\ 
+  & \multicolumn{7}{c}{Brier Score} \\ 
+  & $\alpha = 0.1$ & $\alpha = 0.2$ & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-8}
+V \& VH        &0.231 &0.232 &0.243 &0.245 &0.249 &0.249 &0.249 \\
+Basic          &0.226 &0.226 &0.225 &0.224 &0.221 &0.221 &0.221 \\
+Linear Pred    &0.228 &0.226 &0.225 &0.224 &0.224 &0.249 &0.249 \\
+Deviance Resid &0.231 &0.229 &0.228 &0.245 &0.249 &0.249 &0.249 \\ 
 \hline
@@
 
+%% \begin{table}[ht]
+%% \centering
+%% \caption{\label{tb:shedden} The C Index and Brier Score of all elastic net models selected by cross-validation methods }
+%% \begin{tabular}{lcccccccc}
+%% \hline
+%%   & \multicolumn{8}{c}{Cross-validated C Index} \\ 
+%%   & $\alpha = 0.05$ & $\alpha = 0.1$ & $\alpha = 0.2$ & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-9}
+%% V \& VH     & 0.645 &0.645 &0.643 &0.630 &0.626 &0.617 &0.617 &0.617\\
+%% Basic       &0.652 &0.649 &0.650 &0.650 &0.649 &0.650 &0.651 &0.651 \\
+%% Linear Pred    &0.649 &0.648 &0.650 &0.650 &0.649 &0.648 &0.617 &0.617  \\
+%% Deviance Resid &0.648 &0.645 &0.647 &0.646 &0.626 &0.617 &0.617 &0.617 \\ 
+%% \addlinespace
+%%   & \multicolumn{8}{c}{Brier Score} \\ 
+%%   & $\alpha = 0.05$ & $\alpha = 0.1$ & $\alpha = 0.2$ & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-9}
+%% V \& VH &0.232 &0.231 &0.232 &0.243 &0.245 &0.249 &0.249 &0.249 \\
+%% Basic  &0.221 &0.226 &0.226 &0.225 &0.224 &0.221 &0.221 &0.221 \\
+%% Linear Pred & 0.220 &0.228 &0.226 &0.225 &0.224 &0.224 &0.249 &0.249 \\
+%% Deviance Resid &0.229 &0.231 &0.229 &0.228 &0.245 &0.249 &0.249 &0.249 \\ 
+%% \hline
+%% \end{tabular}
+%% \end{table}
+
 For each fixed $\alpha$, we compared the cross-validation errors (CVE) from the four cross-validation methods along the solution path.
-At $\alpha = 0.3$, CVEs are rescaled and plotted for all four methods along the same grid of $\lambda$ values in Figure~\ref{fig:CVE}.
+At $\alpha = 0.3$, CVEs are normalized so that they can be compared on the same scale and plotted for all four methods along the same grid of $\lambda$ values in Figure~\ref{fig:CVE}.
 The selected $\lambda$ values, where a given CVE curve reaches its lowest point, are indicated with dotted lines.
@@
 
-Cross-validation is a crutial step in fitting penalized Cox regression models.
+Cross-validation is a crucial step in fitting penalized Cox regression models.
 In this paper, we propose two cross-validation methods for Cox Regression: cross-validated linear predictors and cross-validated deviance residuals.

### 2023-08-14 sub/smimr/response.tex
Done with reviewer response

@@
 \usepackage{amssymb,amsmath}
+\usepackage{hyperref}
+\hypersetup{colorlinks=true, citecolor=black, urlcolor=cyan}
 
@@
 The basic approach and the Verweij and Van Houwelingen's approach are implemented in the \textbf{glmnet} package.
-The code to reproduce our simulations and analyses are available as a GitHub repository.
+The code to reproduce our simulations and analyses are available at \url{https://github.com/biydai/Cross-Validation}; this repository is currently private, with the expectation of making it public once the paper is accepted.
 
@@
 All of them have been fixed in the revised version of the manuscript.
+Most of these points don't need any comment (we simply adopted your recommendation).
+The only exception is your comment ``Equation (6) and P8L10: I found the definition of $\hat{\eta}^{\text{cv}}$ confusing.
+Do we not need some step for pooling cross-validated linear predictors over all K folds?'' To which our reply is that no additional steps are needed; the linear predictors are simply concatenated in order to obtain the vector of cross-validated linear predictors, as shown in Figure 1.
+Each fold produces $n/K$ entries; at the end of all $K$ folds, we have all $n$ elements of the cross-validated linear predictor.
+We have made this more explicit in the text after equation (8).
 
-\begin{itemize}[align = left]
-\item[-]  P3L33: define beta
-
-%\re Thank you for the suggestion. We included definition for $\bbeta$ in this paragraph when first introducing the notation.
-
-\item[-] Distinguish between scalars and vectors by using bold letters for vectors
-
-%\re Thank you for the suggestion. We updated the manuscript by using bold betters for vectors.
-
-\item[-] Figure 1: Since “Folder 1” is highlighted for each panel, I think the figure corresponds to the specific case with k=1. If so, k should be replaced by 1 in expressions.
-
-%\re Thank you for this careful observation.
-In the figure, we used "Folder 1" for the simplicity of drawing.
-However, replacing $k$ with $1$ in the expressions in the figure may cause confusion to the readers.
-Therefore, we added description in the caption of Figure 1 to clarify that this illustration represents $k = 1$.
-
-\item[-]  P8L5: I would not use “k” in the expression because k is used for k-th “fold” in this paper.
-
-%\re Thank you for the suggestion. We replaced $k$ with $\tau$ in that formula
-
-\item[-]  Equation (6) and P8L10: I found the definition of $\hat{\eta}^{\text{cv}}$ confusing. Do we not need some step for pooling cross-validated linear predictors over all K folds?
-
-%\re No additional steps are needed to pool the cross-validated linear predictors over all K folds.
-Please refer to the linear predictor diagram in Figure 1.
-Suppose each fold has $n_k$ number of observations and the complete data have $n = \sum_{k=1}^{K}  n_k$ observations.
-The cross-validated linear predictor of the $k$th folder $\hat{\boldeta}^{\text{cv}}_k$ is a vector of length $n_k$ and the complete set of cross-validated linear predictors $\hat{\boldeta}^{\text{cv}}$ has a length of $n$.
-We also updated the vector $\boldeta$ by using bold.
-
-\item[-]  Table 1: what are the numbers in parentheses?
-
-%\re The numbers in the parentheses are standard deviation across all simulated samples, which is a metric for simulation errors.
-
-\item[-] Figure 4 caption: LOO is not defined
-
-%\re Thank you for the suggestion. We updated the caption to include defition for LOO.
-
-\item[-]  Figure 4 caption: (blue line) should be (green line)?
-
-%\re Thank you for catching this. We corrected this in the caption.
-
-\item[-]  Figure 4 legend: The “Method” can mean two different things in this paper – either (four different CV methods) or (10 fold CV, LOOCV).
-Therefore, the legend is very confusing because “Basic” here actually means basic CV with 10 folds.
-Although the text in the paper clarifies this, tables and figures need to be understandable by themselves without reference to the text.
-
-%\re Thank you for the suggestion. We edited the caption of Figure 4 to clarify the legends.
-
-\item[-]  Table 4, what are “Selected” and “P” here?
-
-%\re The column "Selected" represents the number of genes that are selected by the linear predictor cross-validation approach; each row represents the numbder of genes that fall into that function category.
-$p$ is the p-value.
-\bd{is the p-value adjusted?
-update $p$ with "p-value"?}
-
-\end{itemize}
 \end{enumerate}
-
-%% \subsection*{References}
-
-%% Harden JJ, Kropko J (2019). Simulating duration data for the Cox model, Political Science Research and Methods, 7(4): 921-928. https://doi.org/10.1017/psrm.2018.19
-
-%% Wang H, Lengerich BJ, Aragam B, Xing EP (2019).
-Precision Lasso: accounting for correlations and linear dependencies in high-dimensional genomic data.
-Bioinformatics, 35(7), 1181-1187.
-https://doi.org/10.1093/bioinformatics/bty750
-
 \end{document}

### 2023-08-14 main.tex
Done with reviewer response

@@
 \begin{equation*}
-  h_{i}(t) = h_{0}(t) \exp( X_{i}\Tr\bbeta),
+  h_{i}(t) = h_{0}(t) \exp( \x_i\Tr\bbeta),
 \end{equation*} 
-where $\bbeta$ is the coefficient vector for the model and $h_{0}$ is the baseline hazard and $\exp(X_i\Tr\bbeta)$ is the risk for a subject with covariates $X_i$ relative to that baseline.
+where $\bbeta$ is the coefficient vector for the model and $h_{0}$ is the baseline hazard and $\exp(\x_i\Tr\bbeta)$ is the risk for a subject with covariates $\x_i$ relative to that baseline.
 The estimation of the coefficients is obtained by maximizing a partial likelihood.
@@
   \label{eq:cox-pl-subj}
-  L_{i}(\bbeta) = \left \{\frac{\exp ( X_{i}\Tr\bbeta)}{\sum_{ k \in R(t_{i})}\exp ( X_{k}\Tr \bbeta)}\right \}^{\delta_{i}},
+  L_{i}(\bbeta) = \left \{\frac{\exp ( \x_i\Tr\bbeta)}{\sum_{ k \in R(t_{i})}\exp (\x_k \Tr \bbeta)}\right \}^{\delta_{i}},
 \end{equation}
@@
   \label{eq:res-m}
-  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(X_i\Tr\bbeta), \\
+  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(\x_i\Tr\bbeta), \\
   \label{eq:res-d}
@@
   \label{eq:cv-lp}
-  \hat{\eta}^{\text{cv}}_{i} = X_{i}\Tr\hat{\bbeta}^{-k}
+  \hat{\eta}^{\text{cv}}_{i} = \x_i\Tr\hat{\bbeta}^{-k}
 \end{equation} 
@@
 \begin{equation*}
-  L(\hat{\boldeta}^{\text{cv}}) = \prod_{i=1}^{n} \left\{ \frac{\exp (\hat{\eta}^{\text{cv}}_{i})}{\sum_{\tau \in R(t_{i})}\exp (\hat{\eta}^{\text{cv}}_\tau)} \right\} ^{\delta_i}.
+  L(\hat{\boldeta}^{\text{cv}}) = \prod_{i=1}^{n} \left\{ \frac{\exp (\hat{\eta}^{\text{cv}}_{i})}{\sum_{r \in R(t_{i})}\exp (\hat{\eta}^{\text{cv}}_r)} \right\} ^{\delta_i}.
 \end{equation*}
@@
 For these simulation studies, the covariate matrix $X$, consisting of $n$ observations and $p$ covariates, was generated from independent $\text{Normal}(0, 1)$ samples.
-Given $X_i$ and a coefficient vector $\bbeta$, survival times were generated from an exponential distribution with rate parameter $\exp(X_i \Tr \bbeta)$.
+Given $\x_i$ and a coefficient vector $\bbeta$, survival times were generated from an exponential distribution with rate parameter $\exp(\x_i \Tr \bbeta)$.
 Censoring status indicators were drawn independently from the binomial distribution.
@@
 \begin{equation}
-\text{Brier}(y(t_0), \hat{S}(t_0|\lambda,x)) = (y(t_0) - \hat{S}(t_0|\lambda,x))^2.
+\text{Brier}(y(t_0), \hat{S}(t_0|\lambda, \x)) = \{ y(t_0) - \hat{S}(t_0|\lambda, \x) \}^2.
 \end{equation}
@@
 With the same notation defined in the previous paragraph, the Kullback-Leibler score is defined to be
-\begin{equation}
-	\text{KL}(y(t_0), \hat{S}(t_0|\lambda,x)) = -\left\{ y(t_0)\log(\hat{S}(t_0|\lambda,x) + (1 - y(t_0))\log(1 - \hat{S}(t_0|\lambda,x)) \right\}.
-\end{equation}
+$$ \text{KL}(y(t_0), \hat{S}(t_0|\lambda, \x)) = -\left\{ y(t_0)\log(\hat{S}(t_0|\lambda, \x) + (1 - y(t_0))\log(1 - \hat{S}(t_0|\lambda, \x)) \right\}. $$
 As for the Brier score, KL scores were computed for all CV approaches at the median survival time.
@@
 
-\caption{\label{Tab:sim} Comparing the predictive accuracy of four cross-validation methods under scenarios with varying dimension and signal strength}
+\caption{\label{Tab:sim} Predictive accuracy of four cross-validation methods under scenarios with varying dimension and signal strength (standard deviations given in parentheses).}
 \centering
@@
 %		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
-%		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
+%		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = \x_i\hat{\beta}^{-k}$\;
 %	} 
@@
 %		Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
-%		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
+%		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = \x_i\hat{\beta}^{-k}$\;
 %	} 

### 2023-08-15 main.tex
Resubmitted to SMMR

@@
 The log partial likelihood is denoted by $\ell$, with $\ell^k$ and $\ell^{-k}$ defined similarly.
-Finally, let $\hat{\bbeta}^{-k}$ denote the penalized estimates that are obtained by using $L^{-k}(\bbeta)$ in \eqref{eq:obj}.
+Finally, let $\bbh^{-k}$ denote the penalized estimates that are obtained by using $L^{-k}(\bbeta)$ in \eqref{eq:obj}.
 Figure~\ref{Fig:methods} diagrams the relationships between the four cross-validation methods we will describe in this section.
 
-%%%%%%%%%%%%%%%%%%%%
 \subsection{Cross validated partial likelihood} 
 \label{Sec:cox-cv-existing}
-%%%%%%%%%%%%%%%%%%%%
 
-A direct approach to measure a model's predictive accuracy in $D_k$ is to evaluate the log partial likelihood at $\hat{\bbeta}^{-k}$ using the data in $D_k$.
+A direct approach to measure a model's predictive accuracy in $D_k$ is to evaluate the log partial likelihood at $\bbh^{-k}$ using the data in $D_k$.
 Cross-validation error (CVE) is then measured by the total deviance across all $K$ folds, which (up to a constant) is given by:
+
 \begin{equation}
   \label{eq:basic}
-  \CVE = -2 \sum_{k=1}^{K} \ell^{k}(\hat{\bbeta}^{-k});
+  \CVE = -2 \sum_{k=1}^{K} \ell^{k}(\bbh^{-k});
 \end{equation}
@@
   \label{eq:VVH}
-  \CVE = -2\sum_{k = 1}^K \left\{ \ell(\hat{\bbeta}^{- k})  - \ell^{-k}(\hat{\bbeta}^{- k}) \right\}. 
+  \CVE = -2\sum_{k = 1}^K \left\{ \ell(\bbh^{- k})  - \ell^{-k}(\bbh^{- k}) \right\}. 
 \end{equation}
-Here, $\ell(\hat{\bbeta}^{-k})$ is the log partial likelihood evaluated at $\hat{\bbeta}^{-k}$ using the entire data set $D$ and $\ell^{-k}(\hat{\bbeta}^{-k})$ is the log partial likelihood evaluated at $\hat{\bbeta}^{-k}$ on
+Here, $\ell(\bbh^{-k})$ is the log partial likelihood evaluated at $\bbh^{-k}$ using the entire data set $D$ and $\ell^{-k}(\bbh^{-k})$ is the log partial likelihood evaluated at $\bbh^{-k}$ on
 the training set $T_k$.
@@
 We refer to this cross-validation method as the Verweij and Van Houwelingen (V\&VH) approach.
-For many models, such as logistic regression or linear regression, the quantities $\ell(\hat{\bbeta}^{- k})  - \ell^{-k}(\hat{\bbeta}^{- k})$ and  $\ell^{k}(\hat{\bbeta}^{-k})$ are equivalent to each other -- in other words, the basic and V\&VH approaches agree.
+For many models, such as logistic regression or linear regression, the quantities $\ell(\bbh^{- k})  - \ell^{-k}(\bbh^{- k})$ and  $\ell^{k}(\bbh^{-k})$ are equivalent to each other -- in other words, the basic and V\&VH approaches agree.
 However, Verweij and Van Houwelingen's approach is more stable for Cox regression as there is always a large number of observations in the risk sets its calculations are based on.
@@
 
-In order to obtain a cross-validated version of these residuals, we would replace $\bbeta$ in the above equations with $\hat{\bbeta}^{-k}$.
+In order to obtain a cross-validated version of these residuals, we would replace $\bbeta$ in the above equations with $\bbh^{-k}$.
 Typically the Nelson-Aalen estimator is used to estimate $H_0$, although other approaches are also possible, such as the Breslow and Kalbfleisch-Prentice estimators \citep{Breslow1972, Kalbfleisch2011}.
@@
 We refer to this method as the \emph{cross-validated linear predictors} approach.
-To be more specific, we would first obtain $\hat{\bbeta}^{-k}$ from training set $T_{k}$; then the cross-validated linear predictors can be calculated for each observation $i$ from the test set $D_k$:
+To be more specific, we would first obtain $\bbh^{-k}$ from training set $T_{k}$; then the cross-validated linear predictors can be calculated for each observation $i$ from the test set $D_k$:
 \begin{equation}
   \label{eq:cv-lp}
-  \hat{\eta}^{\text{cv}}_{i} = \x_i\Tr\hat{\bbeta}^{-k}
+  \hat{\eta}^{\text{cv}}_{i} = \x_i\Tr\bbh^{-k}
 \end{equation} 
@@
 Estimation accuracy was measured using the mean squared error of the coefficients ($\MSE$).
-We define $\hat{\bbeta}(\lambda)$ as the vector of coefficients estimated by a Cox model with LASSO penalty at $\lambda$, and define $\bbeta^*$ as the true value of $\bbeta$ (the value used to generate the data).
+We define $\bbh(\lambda)$ as the vector of coefficients estimated by a Cox model with LASSO penalty at $\lambda$, and define $\bbeta^*$ as the true value of $\bbeta$ (the value used to generate the data).
 Mean squared error is defined to be
 \begin{equation}
-\MSE = E\| \hat{\bbeta}(\lambda) - \bbeta^* \| ^2.
+\MSE = E\| \bbh(\lambda) - \bbeta^* \| ^2.
 \end{equation}
@@
 For each generated data set, an oracle model was fit using Cox regression including only the true features (i.e., those with non-zero coefficients).
-Letting $\hat{\bbeta}^{\text{oracle}}$ denote the resulting estimates, the relative MSE, on the log scale, is given by
+Letting $\bbh^{\text{oracle}}$ denote the resulting estimates, the relative MSE, on the log scale, is given by
 \begin{equation}
-\log\left\{ \frac{\MSE(\hat{\bbeta}(\lambda^{\text{cv}}))}{\MSE(\hat{\bbeta}^{\text{oracle}})} \right\},
+\log\left\{ \frac{\MSE(\bbh(\lambda^{\text{cv}}))}{\MSE(\bbh^{\text{oracle}})} \right\},
 \end{equation}
 where $\lambda^{\text{cv}}$ denotes the value of $\lambda$ selected by cross-validation.
-MSE is estimated by averaging over $N$ replications respectively for $\hat{\bbeta}(\lambda^{\text{cv}})$ and $\hat{\bbeta}^{\text{oracle}}$, with the ratio taken afterwards.
+MSE is estimated by averaging over $N$ replications respectively for $\bbh(\lambda^{\text{cv}})$ and $\bbh^{\text{oracle}}$, with the ratio taken afterwards.
 The $\log$ is taken to correct for skewness.
@@
 We also measure the predictive accuracy via the mean squared prediction error, or Brier score \citep{VanHouwelingen2011}.
-Let $\hat{S}(t_0|\lambda,x)$ denote the predicted probability for an individual with covariates $x$ to survive beyond time $t_0$, and let $y(t_0) = \mathbb{1}\left\{ t > t_{0}\right\}$ indicate whether the individual actually survived beyond time $t_0$.
+Let $\hat{S}(t_0|\lambda, \x)$ denote the predicted probability for an individual with covariates $\x$ to survive beyond time $t_0$, and let $y(t_0) = \mathbb{1}\left\{ t > t_{0}\right\}$ indicate whether the individual actually survived beyond time $t_0$.
 Then the Brier score is defined to be
@@
 For each simulated data set, n = 120, p = 1000.
-Expected censoring percentage is 10$\%$.}
+Expected censoring percentage is 10\%.}
 \end{figure}

### 2023-08-21 manuscript_code/figure_2_auto.R
add past sim results as .RData to figures

@@ -7,8 +7,16 @@
 # in n = 120, p = 1000
 #------------------------------
 
-library(survival)
-library(ncvreg)
+
+rerun <- FALSE
+
+if (rerun == FALSE){
+  load("cor_auto.RData")
+}else{
+  
+  library(survival)
+  library(ncvreg)
+
 
 #-------------------------------------------------------------------------
 # functions
@@ -264,39 +272,15 @@ for(i in 1:N){
   }
 }
 
-#-------------------------------------------------------------------------
-# printing and saving results
-results <- list()
-for(i in 1:5){
-  results[[i]] <- rbind(
-    apply(gp[i,,],2,mean),
-    apply(ncv[i,,],2,mean),
-    apply(ug[i,,],2,mean),
-    apply(dvr[i,,],2,mean)
-  )
-  rownames(results[[i]]) <- c("VVH","LP","ST","DVR")
-  colnames(results[[i]]) <- rho_all
 }
-names(results) <- c("lambda",
-                    "MSE",
-                    "Brier",
-                    "KL",
-                    "CIndex")
-results$lambda
-results$MSE
-results$Brier
-results$KL
-results$CIndex
 
-save(gp,ncv,dvr,ug,oracle.mse,
-      file = "cor_auto.RData")
+
 
 #-------------------------------------------------------------------------
 # making Figure 2
 library(ggplot2)
 library(cowplot)
 
-load("A:\\cv_manuscript_code\\correlated\\cor_auto.RData")
 results <- list()
 for(i in 1:5){
   results[[i]] <- c(
@@ -321,7 +305,7 @@ mse <- log(c(apply(gp[2,,]/oracle.mse,2,mean),
 Brier <- (results$Brier)
 KL <- (results$KL)
 CIndex <- results$CIndex
-coef <- rho_all
+coef <- c(0,0.2,0.4,0.6,0.8)
 Method <- c(rep("V & VH",5),
             rep("Basic",5),rep("LinearPred",5),
             rep("DevResid",5))
@@ -369,14 +353,14 @@ CIndex <- ggplot(data = df, aes(x = coef,y = CIndex,group = Method))+
   theme(legend.position = "right")
 
 
-png(filename="figure_2_auto.png",
-   units="px",
-   width=1000*1.5,
-   height=350*1.5,
-   pointsize=12,
-   res = 120*1.3)
[…]

### 2023-08-21 manuscript_code/figure_2_ex.R

@@ -7,6 +7,13 @@
 # in n = 120, p = 1000
 #------------------------------
 
+rerun <- FALSE
+
+if (rerun == FALSE){
+  load("cor_exchange.RData")
+}else{
+
+
 library(survival)
 library(ncvreg)
 
@@ -262,40 +269,14 @@ for(i in 1:N){
     #print(c(i,j))
   }
 }
-
-#-------------------------------------------------------------------------
-# printing and saving results
-results <- list()
-for(i in 1:5){
-  results[[i]] <- rbind(
-    apply(gp[i,,],2,mean),
-    apply(ncv[i,,],2,mean),
-    apply(ug[i,,],2,mean),
-    apply(dvr[i,,],2,mean)
-  )
-  rownames(results[[i]]) <- c("VVH","LP","ST","DVR")
-  colnames(results[[i]]) <- rho_all
 }
-names(results) <- c("lambda",
-                    "MSE",
-                    "Brier",
-                    "KL",
-                    "CIndex")
-results$lambda
-results$MSE
-results$Brier
-results$KL
-results$CIndex
 
-save(gp,ncv,dvr,ug,oracle.mse,
-      file = "cor_exchange.RData")
 
 #-------------------------------------------------------------------------
 # making Figure 2
 library(ggplot2)
 library(cowplot)
 
-load("A:\\cv_manuscript_code\\correlated\\cor_exchange.RData")
 results <- list()
 for(i in 1:5){
   results[[i]] <- c(
@@ -320,7 +301,7 @@ mse <- log(c(apply(gp[2,,]/oracle.mse,2,mean),
 Brier <- (results$Brier)
 KL <- (results$KL)
 CIndex <- results$CIndex
-coef <- rho_all
+coef <- c(0,0.2,0.4,0.6,0.8)
 Method <- c(rep("V & VH",5),
             rep("Basic",5),rep("LinearPred",5),
             rep("DevResid",5))

### 2023-08-21 manuscript_code/figure_2_git.R

@@ -7,6 +7,12 @@
 # in n = 120, p = 1000
 #------------------------------
 
+rerun <- FALSE
+
+if (rerun == FALSE){
+  load("sim_midd.RData")
+}else{
+
 library(survival)
 library(ncvreg)
 
@@ -217,6 +223,9 @@ for(i in 1:N){
   }
 }
 
+}
+
+
 #-------------------------------------------------------------------------
 # making Figure 2
 library(ggplot2)
@@ -252,6 +261,9 @@ method <- c(rep("V & VH",6),
             rep("DevResid",6))
 df <- data.frame(lambda,mse,Brier,KL,CIndex,coef,method)
 df$method <- factor(df$method,levels = c("V & VH","Standard","LinearPred","DevResid"))
+
+
+
 
 
 lambda_p <- ggplot(data = df, aes(x = coef,y = lambda,group = method))+

### 2023-08-21 manuscript_code/figure_3_git.R

@@ -7,6 +7,16 @@
 # CV Method
 #------------------------------
 
+
+# increase number of folds (plot on the right)
+#----------------------------------------------------------------------------------------
+
+rerun <- FALSE
+
+if (rerun == FALSE){
+  load("scenario_3_1.RData")
+}else{
+
 # functions
 library(survival)
 library(ncvreg)
@@ -37,11 +47,6 @@ genSurvData <- function(n = 100, beta = c(1,-1),cpct = 0.70,sd = 1,h = 1 # 70% m
   y <- Surv(time = time, event = event)
   return(list(x=x,y=y,beta = beta))
 }
-
-
-
-# increase number of folds (plot on the right)
-#----------------------------------------------------------------------------------------
 
 n <- 120
 p <- 1000
@@ -163,41 +168,10 @@ for(i in 1:N){
     ncv[2,i,j] <- norm[ncv[1,i,j] == lambda]
     dvr[2,i,j] <- norm[dvr[1,i,j] == lambda]
   
-    print(c(i,j))
   }
 }
-
-results <- list()
-for(i in 1:2){
-  results[[i]] <- rbind(
-    apply(gp[i,,],2,mean),
-    apply(ncv[i,,],2,mean),
-    apply(ug[i,,],2,mean),
-    apply(dvr[i,,],2,mean)
-  )
-  rownames(results[[i]]) <- c("VVH","LP","ST","DVR")
-  colnames(results[[i]]) <- Nfold
 }
 
-for(i in 3:4){
-  results[[i]] <- rbind(
-    apply(gp[i-2,,],2,sd),
-    apply(ncv[i-2,,],2,sd),
-    apply(ug[i-2,,],2,sd),
-    apply(dvr[i-2,,],2,sd)
-  )
-  rownames(results[[i]]) <- c("VVH","LP","ST","DVR")
-  colnames(results[[i]]) <- Nfold
-}
-
-names(results) <- c("lambda",
-                    "MSE",
-                    "lambda_sd",
-                    "MSE_sd")
-results$lambda
-results$MSE
-results$lambda_sd
-results$MSE_sd
 
 # making plots
 
@@ -221,6 +195,175 @@ MSE_p
 
 # simulation on imbalance (plot on the left)
 #--------------------------------------------------------------------------------------------------------
+
+rerun <- FALSE
+
+if (rerun == FALSE){
+  load("scenario_3_2.RData")
+}else{
+
[…]

### 2023-08-21 manuscript_code/figure_4_git.R

@@ -6,6 +6,13 @@
 # in n = 120, p = 1000
 #------------------------------
 
+rerun <- FALSE
+
+if (rerun == FALSE){
+  load("sim_loo.RData")
+}else{
+
+
 library(survival)
 library(ncvreg)
 
@@ -223,10 +230,10 @@ for(i in 1:N){
     ncv.loo[2,i,j] <- norm[ncv.loo[1,i,j] == lambda]
     dvr.loo[2,i,j] <- norm[dvr.loo[1,i,j] == lambda]
     
-    print(c(i,j))
   }
 }
 
+}
 
 #-----------------------------------------------------------------------------------
 # load saved results and make plots

### 2023-08-21 manuscript_code/figure_5_git.R

@@ -3,8 +3,9 @@
 # to reproduce part of the simulations
 # in Figure 5 of the Cross Validation 
 # Manuscript:
-# Explored low dimension p =50
+# Explored low dimension p =50, p = 1000
 #------------------------------
+
 
 library(survival)
 library(ncvreg)
@@ -45,7 +46,16 @@ brierSurv_lambda <- function(data,time,fit,lambda) {
 }
 
 
-#-------------------------------------------------------------
+#------------------------------
+# Explored low dimension p =50
+#------------------------------
+
+rerun <- FALSE
+
+if (rerun == FALSE){
+  load("sim_CVAUC_low.RData")
+}else{
+
 N <- 200
 n <- 150
 p <- 50
@@ -120,6 +130,7 @@ for(i in 1:N){
     
   }
 }
+}
 
 # store results
 # lambda
@@ -153,7 +164,12 @@ df$Type <- c(rep("lambda",2400),rep("log(MSE Ratio)",2400),rep("Brier Score",240
 # Explored high dimension p =1000
 #------------------------------
 
-#-------------------------------------------------------------
+rerun <- FALSE
+
+if (rerun == FALSE){
+  load("sim_CVAUC_high.RData")
+}else{
+
 N <- 200
 n <- 120
 p <- 1000
@@ -228,7 +244,7 @@ for(i in 1:N){
     
   }
 }
-
+}
 
 # store results
 

### 2023-08-21 manuscript_code/table_1_git.R

@@ -65,6 +65,13 @@ AUC_lambda <- function(data,fit,lambda){
 
 #-------------------------------------------------------------------------
 # scenarios: 
+
+rerun <- FALSE
+
+if (rerun == FALSE){
+  load("sim_low.RData")
+}else{
+
 N <- 200
 p <- 50
 n <- 150
@@ -219,6 +226,7 @@ for(i in 1:N){
     
   }
 }
+}
 
 digit <- 3
 results_mean <- list()
@@ -286,6 +294,13 @@ cbind(
 
 #-------------------------------------------------------------------------
 # scenarios: 
+
+rerun <- FALSE
+
+if (rerun == FALSE){
+  load("sim_high.RData")
+}else{
+
 N <- 200
 p <- 10000
 n <- 400
@@ -442,7 +457,7 @@ for(i in 1:N){
     
   }
 }
-
+}
 
 digit <- 3
 results_mean <- list()

### 2023-08-21 sub/smimr/2/cover.tex

\documentclass{uiletter}
\usepackage{hyperref}
\hypersetup{colorlinks=true, citecolor=black, urlcolor=cyan}

\begin{document}

\StartLetter{}{Dear editors of {\em Statistical Methods in Medical Research},}

We are submitting a revised version of the manuscript ``Cross validation approaches for penalized Cox regression'' for your consideration to be published in \emph{Statistical Methods in Medical Research}.

This version has been revised and improved in several ways, most importantly by carrying out a more thorough literature search with respect to the deviance residual method. In our original submission, we pointed out that this was an approach one could try, but that it didn't work very well. This was a bit confusing, as it seemed like we were proposing an approach that didn't work very well (in addition to one that did). Since the original submission, we were able to find a number of other papers in which this approach has indeed been utilized, which makes our investigation of this method more clearly relevant.

In addition, we incorporated all the excellent feedback we received from the reviewers, and feel that we have addressed all their concerns in this revision.

We realize that we are quite late with this resubmission -- my coauthor Biyue moved to a new institution and due to a miscommunication, we each thought the other had resubmitted the paper. I communicated with Prof. Chris Metcalfe, who encouraged us to resubmit the paper despite this on August 12, 2023. Thank you very much for your consideration of this revised manuscript.

\EndLetter

\end{document}

### 2024-01-08 sub/smimr/response2/response.tex
New referee reports; addressing in-progress

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[a4paper,margin = 1in]{geometry}
\usepackage{enumitem}
\usepackage{color}
\usepackage{amssymb,amsmath}
\usepackage{hyperref}
\hypersetup{colorlinks=true, citecolor=black, urlcolor=cyan}

\newcommand{\re}{\textbf{Response: }}
\newcommand\pb[1]{{\color{red}#1}}
\newcommand\bd[1]{{\color{blue}#1}}
\newcommand{\bbeta}{\boldsymbol{\beta}}
\newcommand{\boldeta}{\boldsymbol\eta}

\begin{document}

\subsubsection*{Response to comments by Reviewer \#1}

Thank you very much for taking the time to carefully read our revised manuscript and provide additional feedback. \pb{more}

\begin{enumerate}[align = left]

\item \emph{On the point of dependent censoring, I agree that this may cause some issues regarding the evaluation of how well a cross-validation approach that depends on the partial likelihood is doing. But this does not mean you should avoid mentioning this as a shortcoming of the method in the Discussion.  One of the key points of a Discussion section is to convey to the reader guidelines on when your proposed method will work and when it may not. Figuring out a way to conduct cross-validation for penalized Cox regression under dependent censoring would be a great gap to point out for future work.  And dependent censoring happens far more than many papers take for granted. Please include some comments in the Discussion on why dependent censoring is an issue with this cross-validation approach and a gap that should be pursued in future research.}

  \re \pb{fixed}

\item \emph{On p. 12, just above Figure 1, I do not understand the sentence “We computed the C Index using the independent test set of 1000 individuals described above.”. I see where you mention the 1000 individuals for the independent test data for this simulation example, but this is an unusual scenario when n=120 (as in the first simulation, with results in Figure 1, or even sample sizes as large as 400 in the second simulation, with results in Table 1).  Please provide justification for this, as we would normally not be able to afford a hold-out sample so large (relative to the "observed" sample size) in typical analyses of data, and it would actually be a bad use of resources, as a larger percentage of the original data is needed to help more reliably develop the models and tune parameters. And this hold-out sample is also uncensored, which is also not realistic. Please justify all this. Feel free to mention external model validation as one setting where a larger holdout sample might be utilized, but this is really not portrayed to be the setting in your simulations.}

  \re \pb{explain this here; don't change paper}

\item \emph{Related to the simulation studies, please disregard casual mentioning of the first simulation, and so on, and clearly label the different simulations throughout Section 3 as Simulation 1, Simulation 2, and so on, with clear mentioning of this in both the text and the tables and figures in which the results of those simulations are displayed. This will make for much more straightforward reading/referencing.}

  \re \pb{in process of fixing}

\item \emph{Across all your simulation studies, please provide details so that others can reproduce your simulation results. If this will take up too much text, you can point them to a GitHub site where they can find sufficient details to allow for such reproducibility.}

  \re {mentioned GH repo; make sure public}

\item \emph{Still a few typos remain (e.g., “such fitting penalized Cox regression” on p. 25), so some additional proofreading is required, but it reads much more smoothly in the revision.  I would suggest finding consistency with how you display cross-validation (displayed both with and without a hyphen many times, or as CV).  And cross-validated as an adjective should always have a hyphen.  Another inconsistency is with LASSO vs. lasso, the latter used a few times in Section 4 and at least once in Section 5.}

  \re \pb{need to fix}

\end{enumerate}

\end{document}

### 2024-01-08 main.tex
New referee reports; addressing in-progress

@@
   \includegraphics[width=\textwidth]{figure_2_new.png}
-  \caption{\label{Fig:mse-brier-c} Simulation comparing 4 CV methods.
+  \caption{\label{Fig:mse-brier-c} Simulation comparing 4 CV methods as described in Section~\ref{Sec:accuracy}.
 The horizontal axis in all three plots is $s$, which represents the signal strength in the data.
@@
 
-For our first simulation study, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates.
+Figure~\ref{Fig:mse-brier-c} illustrates the results when the data contains $n = 120$ observations and $p = 1000$ covariates.
 The covariate vector $\bbeta$ is assumed to be sparse, with $\beta_j = 0$ for $j \in \{11, 12, \ldots, 1000\}$ and $\beta_j = s$ for $j \in \{1, 2, \ldots, 10 \}$, where $s$ is a prespecified constant that specifies the signal strength in the simulated data set.
@@
 $N = 200$ replications were used for each scenario.
-Figure~\ref{Fig:mse-brier-c} illustrates the results of this simulation study.
-For all four metrics, the overall pattern is the same (KL scores not shown in Figure~\ref{Fig:mse-brier-c} due to space constraints, but they were very similar to Brier scores).
+For all four metrics, the overall patterns in Figure~\ref{Fig:mse-brier-c} are the same (KL scores not shown in Figure~\ref{Fig:mse-brier-c} due to space constraints, but they were very similar to Brier scores).
 Although the four cross-validation methods performed relatively similarly, the basic and linear predictor approaches had the best performance across all metrics: lower MSE, better Brier scores, and higher C Index.
@@
 
-Although we focused upon the penalized Cox regression with a LASSO penalty in this work, the methods that we analyzed are applicable in any context that involves conducting model selection for survival models that are built upon the Cox partial likelihood, such fitting penalized Cox regression with other penalty functions (elastic net \citep{Zou2005}, SCAD \citep{FanLi2001}, MCP \citep{Zhang2010}, precision lasso \citep{Wang2018}), other penalty structures (group selection, bi-level selection), and selecting functional forms of covariates in unpenalized Cox regression.
+Although we focused upon the penalized Cox regression with a LASSO penalty in this work, the methods that we analyzed are applicable in any context that involves conducting model selection for survival models that are built upon the Cox partial likelihood, such as fitting penalized Cox regression with other penalty functions (elastic net \citep{Zou2005}, SCAD \citep{FanLi2001}, MCP \citep{Zhang2010}, precision lasso \citep{Wang2018}), other penalty structures (group selection, bi-level selection), and selecting functional forms of covariates in unpenalized Cox regression.
+
 We note, however, that the focus of this article is restricted to fixed covariates.
 How best to carry out cross-validation for model selection in the presence of time-varying covariates, or in joint modeling of time-to-event and longitudinal outcomes, remains an open question worthy of additional future study.
+In addition, we have assumed throughout that the Cox model and its assumptions regarding independent censoring are correct.
+If these assumptions are not met, then it is unclear which Cox model should be chosen (or indeed, whether Cox regression should be used at all).
 

### 2024-01-11 sub/smimr/response2/response.tex
Finished revisions based on referee feedback

@@
 \usepackage[utf8]{inputenc}
-\usepackage[a4paper,margin = 1in]{geometry}
+\usepackage[margin = 0.9in]{geometry}
 \usepackage{enumitem}
@@
 
-Thank you very much for taking the time to carefully read our revised manuscript and provide additional feedback. \pb{more}
+Thank you very much for taking the time to carefully read our revised manuscript and provide additional feedback.
 
@@
 
-  \re \pb{fixed}
+  \re The final paragraph of the discussion now comments on both time-varying covariates and dependent censoring.
 
@@
 
-  \re \pb{explain this here; don't change paper}
+  \re Our goal is to assess the true prediction accuracy of the model (using the entire population).
+For this, the larger the number we use, the better (lower Monte Carlo error).
+We could use 100,000 or 10,000,000 observations, but it doesn't make much difference -- 1,000 observations is adequate.
+What you are describing is the question of how an analyst might use the data to both fit a model and obtain an estimate of that model's predictive accuracy.
+This is important, but not the subject of our manuscript.
+Our paper is about \emph{model selection} (estimating the performance of different models in order to choose the best one), not \emph{model assessment} (having chosen a final model, estimating its prediction error).
 
@@
 
-  \re \pb{in process of fixing}
+  \re This is a good point; all legends now make references to the section of the manuscript in which the corresponding simulation is described.
 
@@
 
-  \re {mentioned GH repo; make sure public}
+  \re We absolutely agree! All code is available at the following url; we weren't sure where to put it in the manuscript because each journal handles this differently. \\
+  \url{https://github.com/biydai/Cross-Validation}.
 
@@
 
-  \re \pb{need to fix}
+  \re Thank you very much for bringing this to our attention, especially our inconsistencies with respect to hyphenating cross-validation and capitalizing LASSO.
+All of these issues have now been fixed.
 

### 2024-01-11 cross-validation.tex
Editing text surrounding Table 3 and response
Done with reviewer response
Finished revisions based on referee feedback

@@
 \usepackage{hyperref}
-\hypersetup{colorlinks=true, citecolor=black, urlcolor=cyan}
+\hypersetup{colorlinks=true, citecolor=black, linkcolor=., urlcolor=cyan}
 \usepackage{afterpage}
@@
 
+\newcommand{\x}{\mathbf{x}}
 \newcommand{\lam}{\lambda}
@@
 \newcommand{\by}{\boldsymbol{y}}
+\newcommand{\bbh}{\hat{\boldsymbol{\beta}}\phantom{}}
 \newcommand{\bbeta}{\boldsymbol{\beta}}
@@
 \newcommand{\btau}{\boldsymbol{\tau}}
-\newcommand{\bx}{\boldsymbol{x}}
 \newcommand{\beps}{\boldsymbol\epsilon}
@@
 
-\title{Cross Validation Approaches for Penalized Cox Regression}
-\author{Biyue Dai\\Department of Biostatistics\\University of Iowa
+\title{Cross-Validation Approaches for Penalized Cox Regression}
+\author{Biyue Dai \\ Division of Biostatistics \\ University of Minnesota
   \and
-  Patrick Breheny\\Department of Biostatistics\\University of Iowa}
+  Patrick Breheny \\ Department of Biostatistics \\ University of Iowa}
 \date{\today}
@@
 \section*{Abstract}
-Cross validation is the most common way of selecting tuning parameters in penalized regression, but its use in penalized Cox regression models has received relatively little attention in the literature.
-Due to its partial likelihood construction, carrying out cross validation for Cox models is not straightforward, and there are several potential approaches for implementation.
+Cross-validation is the most common way of selecting tuning parameters in penalized regression, but its use in penalized Cox regression models has received relatively little attention in the literature.
+Due to its partial likelihood construction, carrying out cross-validation for Cox models is not straightforward, and there are several potential approaches for implementation.
 Here, we propose a new approach based on cross-validating the linear predictors of the Cox model and compare it to approaches that have been proposed elsewhere.
 We show that the proposed approach offers an attractive balance of performance and numerical stability, and illustrate these advantages using simulated data as well as analyzing a high-dimensional study of gene expression and survival in lung cancer patients.
+
 \input{main}

### 2024-01-11 main.tex
Finished revisions based on referee feedback

@@
 
-In this paper, we propose an alternative way to carry out cross validation for Cox regression.
-Instead of applying cross-validation to the partial likelihood directly, we propose applying cross validation to the linear predictors of the regression model.
-Through simulation studies, we compare these proposed methods with existing cross-validation approaches for LASSO penalized Cox regression in both low- and high-dimensional settings.
+In this paper, we propose an alternative way to carry out cross-validation for Cox regression.
+Instead of applying cross-validation to the partial likelihood directly, we propose applying cross-validation to the linear predictors of the regression model.
+Through simulation studies, we compare these proposed methods with existing cross-validation approaches for lasso-penalized Cox regression in both low- and high-dimensional settings.
 We find that the linear predictor approach avoids the conservatism of the deviance residual and \citet{Verweij1993} approaches as well as the numerical instability of ordinary cross-validation.
@@
 
-In this paper, we focus upon the LASSO penalty $P_{\lambda}(\bbeta) = \lambda \sum_{j} |\beta_{j}|$ as it is most common, although the need to select regularization parameters arises in all penalized regression models, such as elastic net \citep{Zou2005}, smoothly clipped absolute deviations (SCAD) penalty \citep{FanLi2001}, and minimax concave penalty (MCP) \citep{Zhang2010}.
+In this paper, we focus upon the lasso penalty $P_{\lambda}(\bbeta) = \lambda \sum_{j} |\beta_{j}|$ as it is most common, although the need to select regularization parameters arises in all penalized regression models, such as elastic net \citep{Zou2005}, smoothly clipped absolute deviations (SCAD) penalty \citep{FanLi2001}, and minimax concave penalty (MCP) \citep{Zhang2010}.
 Indeed, cross-validation can be used for model selection in Cox regression even without the presence of a penalty, so the relevance of this question is wide-ranging.
-LASSO-penalized Cox regression is particularly useful for high dimensional data where the number of covariates $p > n$, such as predicting overall survival in cancer patients based on genome-wide expression measurements.
-LASSO estimates yield a sparse model where some coefficients are exactly zero.
+Lasso-penalized Cox regression is particularly useful for high dimensional data where the number of covariates $p > n$, such as predicting overall survival in cancer patients based on genome-wide expression measurements.
+Lasso estimates yield a sparse model where some coefficients are exactly zero.
 This sparsity pattern changes as we vary $\lam$: at large values of $\lambda$, most or all of the coefficients are 0, but as $\lam$ decreases, more covariates are selected.
-Selecting $\lambda$ is critical to LASSO estimation in the sense that the model's accuracy suffers if $\lam$ is either too large or too small.
+Selecting $\lambda$ is critical to lasso estimation in the sense that the model's accuracy suffers if $\lam$ is either too large or too small.
 
@@
 
-\subsection{Cross validated partial likelihood} 
+\subsection{Cross-validated partial likelihood} 
 \label{Sec:cox-cv-existing}
@@
   \includegraphics[height= 12cm ]{figure_1_new.png}
-  \caption{\label{Fig:methods} Illustration of $K$-fold cross validation methods for Cox regression when $k = 1$}
+  \caption{\label{Fig:methods} Illustration of $K$-fold cross-validation methods for Cox regression when $k = 1$}
 \end{figure}
@@
 %%%%%%%%%%%%%%%%%%%%
-\subsection{Cross validated deviance residuals}
+\subsection{Cross-validated deviance residuals}
 %%%%%%%%%%%%%%%%%%%%
 
-The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model, which prevents us from calculating prediction errors (residuals).
+The fundamental challenge of conducting cross-validation for Cox regression is that the baseline hazard is not estimated from the model, which prevents us from calculating prediction errors (residuals).
 However, if one takes additional steps to estimate the baseline hazard in addition to estimating the regression coefficients, then such residuals can be calculated.
@@
 
-The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross validated error, analogous to using the residual sum of squares in linear regression.
+The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross-validated error, analogous to using the residual sum of squares in linear regression.
 We refer to this method as the \emph{cross-validated deviance residuals} approach.
@@
 %%%%%%%%%%%%%%%%%%%%
-\subsection{Cross validated linear predictors}
+\subsection{Cross-validated linear predictors}
 \label{Sec:linear-predictor}
@@
 
-Simulation studies were conducted to compare how the methods described in Section~\ref{Sec:methods} behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for LASSO penalized regression.
+Simulation studies were conducted to compare how the methods described in Section~\ref{Sec:methods} behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for lasso penalized regression.
 For these simulation studies, the covariate matrix $X$, consisting of $n$ observations and $p$ covariates, was generated from independent $\text{Normal}(0, 1)$ samples.
@@
 Estimation accuracy was measured using the mean squared error of the coefficients ($\MSE$).
-We define $\bbh(\lambda)$ as the vector of coefficients estimated by a Cox model with LASSO penalty at $\lambda$, and define $\bbeta^*$ as the true value of $\bbeta$ (the value used to generate the data).
+We define $\bbh(\lambda)$ as the vector of coefficients estimated by a Cox model with lasso penalty at $\lambda$, and define $\bbeta^*$ as the true value of $\bbeta$ (the value used to generate the data).
 Mean squared error is defined to be
@@
 \end{equation}
-In our simulation study, for each data set, we generated another independent data set with 1000 uncensored observations.
+In our simulation study, for each data set, we generated another independent data set with 1000 uncensored observations to assess prediction accuracy.
 We computed Brier scores for all of the cross-validation methods for all individuals in this test set, setting $t_0$ to the median survival time.
@@
 A C Index of 0.5 means the prediction is no better than flipping a coin and C Index of 1 means perfect concordance.
-We computed the C Index using the independent test set of 1000 individuals described above.
+We computed the C Index using the same set of 1000 independent observations as the Brier score.
 
@@
 
-\caption{\label{Tab:sim} Predictive accuracy of four cross-validation methods under scenarios with varying dimension and signal strength (standard deviations given in parentheses).}
+\caption{\label{Tab:sim} Predictive accuracy of four cross-validation methods under scenarios with varying dimension and signal strength (standard deviations given in parentheses; Section~\ref{Sec:accuracy}.}
 \centering
@@
 \setlength{\tabcolsep}{3pt}
-\caption{\label{Tab:power} Number of non-zero features (true positives) selected by each cross-validation method, under scenarios with varying dimension and signal strength.
+\caption{\label{Tab:power} Number of non-zero features (true positives) selected by each cross-validation method, under scenarios with varying dimension and signal strength (Section~\ref{Sec:power}).
 For weak and strong signal strength, nonzero coefficients are set to 0.3 and 0.6 respectively.}
@@
   \includegraphics[width=\textwidth]{figure_3.png}
-  \caption{\label{Fig:stability} Simulations illustrating the stability issues with the basic approach.
+  \caption{\label{Fig:stability} Simulations illustrating the stability issues with the basic approach (Section~\ref{Sec:stability}).
 For the figure on the left, $n = 100$, fold assignments are unbalanced, and the vertical axis is the proportion of undefined partial likelihoods.
@@
 
-\subsection {Leave-One-Out Cross Validation}
+\subsection {Leave-One-Out Cross-Validation}
 \label{Sec:loocv}
 
-As discussed earlier, basic cross-validation cannot be used to carry out leave-one-out cross validation (LOOCV) for Cox regression models.
+As discussed earlier, basic cross-validation cannot be used to carry out leave-one-out cross-validation (LOOCV) for Cox regression models.
 The other three approaches (V\&VH, linear predictor, and deviance residual), however, are not subject to the stability issues described in Section~\ref{Sec:stability}, and can be used to implement LOOCV.
@@
   \includegraphics[width=\textwidth]{figure_4.png}
-  \caption{\label{Fig:loocv}Simulation comparing LOO (Leave-One-Out) and 10-fold CV across three approaches (V\&VH, linear predictor, and deviance residual).
+  \caption{\label{Fig:loocv} Simulation comparing LOO (Leave-One-Out) and 10-fold CV across three approaches (V\&VH, linear predictor, and deviance residual) as described in Section~\ref{Sec:loocv}.
 With $N = 200$ replications, the data sets generated have dimension $n = 120$ and $p = 1000$.
@@
 All of the cross-validation approaches described in Section~\ref{Sec:methods} use the Cox partial likelihood, in one way or another, as the criterion upon which model selection is based.
-In this section, we compare these likelihood-based approaches against an alternative approach of cross validation using concordance, also known as the cross-validated C Index.
+In this section, we compare these likelihood-based approaches against an alternative approach of cross-validation using concordance, also known as the cross-validated C Index.
 This is a widely used approach for conducting cross-validation with time-to-event outcomes, particularly for machine learning models \citep{Subramanian2011,Simon2011a}, and was recently added as an option in the \texttt{glmnet} package (\texttt{type.measure="C"}).
@@
   \includegraphics[height= 15cm ]{figure_auc.png}
-  \caption{\label{Fig:cvauc}Comparing Partial Likelihood and C Index both built from Cross-validated Linear Predictors.
+  \caption{\label{Fig:cvauc} Comparing Partial Likelihood and C Index both built from cross-validated Linear Predictors (Section~\ref{Sec:CIndex}).
 The horizontal axis represents the strength of signal $s$ in the data.
@@
 \section{Application to Real Data}
+\label{Sec:real}
 
@@
 \centering
-\caption{\label{tb:shedden} The C Index and Brier Score of all elastic net models selected by cross-validation methods }
+\caption{\label{tb:shedden} The C Index and Brier Score of all elastic net models selected by cross-validation methods (Section~\ref{Sec:real})}
 \begin{tabular}{lccccccc}
@@
   \includegraphics[height= 8cm ]{shedden.png}
-  \caption{\label{fig:CVE}Rescaled Cross Validated Errors at $\alpha = 0.3$.
+  \caption{\label{fig:CVE} Rescaled cross-validated errors at $\alpha = 0.3$ (Section~\ref{Sec:real}).
 The horizonal axis is $\log(\lambda)$.
@@
 \centering
-\caption{\label{Tab:go} Five most significantly enriched molecular functions among genes selected by the linear predictor method with the elastic net ($\alpha=0.3$).
+\caption{\label{Tab:go} Five most significantly enriched molecular functions among genes selected by the linear predictor method with the elastic net ($\alpha=0.3$) for the data described in Section~\ref{Sec:real}.
 The five terms have a false discovery rate of 3\% using the Benjamini-Hochberg method.
@@
 
-Although we focused upon the penalized Cox regression with a LASSO penalty in this work, the methods that we analyzed are applicable in any context that involves conducting model selection for survival models that are built upon the Cox partial likelihood, such as fitting penalized Cox regression with other penalty functions (elastic net \citep{Zou2005}, SCAD \citep{FanLi2001}, MCP \citep{Zhang2010}, precision lasso \citep{Wang2018}), other penalty structures (group selection, bi-level selection), and selecting functional forms of covariates in unpenalized Cox regression.
+Although we focused upon the penalized Cox regression with a lasso penalty in this work, the methods that we analyzed are applicable in any context that involves conducting model selection for survival models that are built upon the Cox partial likelihood, such as fitting penalized Cox regression with other penalty functions (elastic net \citep{Zou2005}, SCAD \citep{FanLi2001}, MCP \citep{Zhang2010}, precision lasso \citep{Wang2018}), other penalty structures (group selection, bi-level selection), and selecting functional forms of covariates in unpenalized Cox regression.
 

### 2024-01-11 sub/smimr/3/cover.tex
Formatting for SMIMR resubmission

--- 2023-08-21 sub/smimr/2/cover.tex
@@
 
-We are submitting a revised version of the manuscript ``Cross validation approaches for penalized Cox regression'' for your consideration to be published in \emph{Statistical Methods in Medical Research}.
+We are submitting another revision of the manuscript ``Cross-validation approaches for penalized Cox regression'' for your consideration to be published in \emph{Statistical Methods in Medical Research}.
 
-This version has been revised and improved in several ways, most importantly by carrying out a more thorough literature search with respect to the deviance residual method.
-In our original submission, we pointed out that this was an approach one could try, but that it didn't work very well.
-This was a bit confusing, as it seemed like we were proposing an approach that didn't work very well (in addition to one that did).
-Since the original submission, we were able to find a number of other papers in which this approach has indeed been utilized, which makes our investigation of this method more clearly relevant.
+The revisions here are relatively modest -- both reviewers were happy with the paper, although reviewer \#1 pointed out some places where clarifications that could be made and inconsistencies addressed.
+These have all been addressed now.
 
-In addition, we incorporated all the excellent feedback we received from the reviewers, and feel that we have addressed all their concerns in this revision.
-
-We realize that we are quite late with this resubmission -- my coauthor Biyue moved to a new institution and due to a miscommunication, we each thought the other had resubmitted the paper.
-I communicated with Prof.
-Chris Metcalfe, who encouraged us to resubmit the paper despite this on August 12, 2023.
 Thank you very much for your consideration of this revised manuscript.

### 2024-01-11 sub/smimr/2/scrub-url.R

@@ -1,6 +1,0 @@
-bib <- readLines('../../../articles.bib')
-bib <- bib[!grepl('eprint', bib)]
-url <- grepl("url\\s*=", bib)
-rproj <- grepl("R-project\\.org", bib)
-bib <- bib[!url | rproj]
-writeLines(bib, 'articles.bib')

### 2024-01-12 main.tex
update figure 1 and cv deviance residual notation

@@
 
-The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross-validated error, analogous to using the residual sum of squares in linear regression.
+The sum of squared cross-validated deviance residuals, $\sum_{i}{d}^{\text{cv}}_{i}^2$, are then used as the cross-validated error, analogous to using the residual sum of squares in linear regression.
 We refer to this method as the \emph{cross-validated deviance residuals} approach.

### 2024-01-12 supplementary.tex
Edits to abstract and intro
update figure 1 and cv deviance residual notation

@@
 \begin{align*}
-  m_{i}^{cv} &= \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}) \\
-  d_{i}^{cv} &= \text{sign}(m_{i}^{cv})\sqrt{-2(m_i^{cv} + \delta_{i}\log(\delta_{i} - m_i^{cv}))},
+  m_{i}^{\text{cv}} &= \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}) \\
+  d_{i}^{\text{cv}} &= \text{sign}(m_{i}^{\text{cv}})\sqrt{-2(m_i^{\text{cv}} + \delta_{i}\log(\delta_{i} - m_i^{\text{cv}}))},
 \end{align*}
 for $i$ in the test set $D_k$, with $\hat{\beta}^{-k}$ obtained from the training set $T_k$.
-The sum of squared cross-validated deviance residuals, $\sum_{i}(d_{i}^{cv})^2$, are then used as the cross validated error, analogously to using the residual sum of squares in linear regression.
+The sum of squared cross-validated deviance residuals, $\sum_{i}(d_{i}^{\text{cv}})^2$, are then used as the cross validated error, analogously to using the residual sum of squares in linear regression.
 In the manuscript, this method is referred to as the \emph{cross-validated deviance residuals} approach.
@@
   	\For{each obervation $i$ in the $k$th fold}{
-  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
-  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
+  	Compute the Cross-validated Martingale Residual: $m^{\text{cv}}_{i}  = \delta_{i} - \hat{H}_{0}(t_{i})\exp(x_i\Tr\hat\bbeta^{-k})$\;
+  	Compute the Deviance Residual: $ d^{\text{cv}}_{i} = \text{sgn}(m^{\text{cv}}_{i})\sqrt{-2(m^{\text{cv}}_{i} + \delta_{i}\log(\delta_{i} - m^{\text{cv}}_{i}))}$\;
   	}
  }
- Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
+ Compute the sum of squares: $\sum_{i}({d}^{\text{cv}}_{i})^2$
  }
@@
 
-\begin{table}[h]
+\begin{table}[ht]
 
@@
 
-\begin{table}[h]
+\begin{table}[ht]
 \setlength{\tabcolsep}{3pt}

### 2024-01-12 main.tex
fixing some minor errors introduced by previous commit

--- 2024-01-11 main.tex
@@
   \centering
-  \includegraphics[height= 12cm ]{figure_1_new.png}
+  \includegraphics[height= 12cm ]{figure_1_HD.png}
   \caption{\label{Fig:methods} Illustration of $K$-fold cross-validation methods for Cox regression when $k = 1$}
@@
 
-The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross-validated error, analogous to using the residual sum of squares in linear regression.
+The sum of squared cross-validated deviance residuals, $\sum_{i} (d^{\text{cv}}_{i})^2$, are then used as the cross-validated error, analogous to using the residual sum of squares in linear regression.
 We refer to this method as the \emph{cross-validated deviance residuals} approach.
